### 2013-02-13 article/main.tex
first fixing comments

%% bare_jrnl_compsoc.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE Computer
%% Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
% The Computer Society usually requires 10pt for submissions.
%
\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[12pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

%%extra packages ------ 
\usepackage{longtable}   
\usepackage{float} %allows the use of [H] to force positioning of figures
\usepackage{subfigure}
\usepackage{hyperref} %used in references now  --- it shoud be removed
\usepackage{amssymb}
\usepackage[neveradjust]{paralist} %compact lists

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  % \usepackage[nocompress]{cite}
\else
  % normal IEEE
  % \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
%\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
   \graphicspath{{figures}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
%\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
%   \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
   %\graphicspath{{./figures/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
   %\DeclareGraphicsExtensions{.eps}
%\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\else
%\usepackage[tight,footnotesize]{subfigure}
%\fi
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. Computer Society papers
% use a larger font and \sffamily font for their captions, hence the
% additional options needed under compsoc mode. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false]{caption}
%  \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false]{caption}
%  \usepackage[font=footnotesize]{subfig}
%\fi
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Software Development in Startup Companies: The Green-field Startup Model}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{FirstName~SecondName,~\IEEEmembership{Fellow,~IEEE,}
        BLA~BLA,~\IEEEmembership{Fellow,~IEEE,}
        and~BLA~BLA,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem Author info.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: see http://something/contact.html
\IEEEcompsocthanksitem Author info.}% <-this % stops a space
\thanks{}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
Software startups are newly created companies with no operating history and extremely fast in producing cutting-edge technologies. Despite their increasing importance in economy there are only few scientific studies attempting to address software engineering (SE) issues, especially for early-stage startups. This research aims to understand the software development strategies conducted by practitioners, in the period of time that goes from idea conception to the first open beta release. A \textit{grounded theory} approach guided the execution of a case study to explore the state-of-practice. As result, the Greenfield startup model (GSM)  reveals the urgent priority of startups of releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early collected user feedbacks. Nevertheless, the need of shortening time-to-market, by speeding-up the development through low-precision and product-centric engineering activities, is counterbalanced by the need of restructuring the product before setting off for further growth.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

% Note that keywords are not normally used for peer review papers.
\begin{keywords}
Software Development, Startups, Grounded Theory.
\end{keywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
% is not selected <OR> if conference mode is selected - because compsoc
% conference papers position the abstract like regular (non-compsoc)
% papers do!
\IEEEdisplaynotcompsoctitleabstractindextext
% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
% compsoc under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\section{Introduction}
\label{intro}
% Computer Society journal papers do something a tad strange with the very
% first section heading (almost always called "Introduction"). They place it
% ABOVE the main text! IEEEtran.cls currently does not do this for you.
% However, You can achieve this effect by making LaTeX jump through some
% hoops via something like:
%
%\ifCLASSOPTIONcompsoc
%  \noindent\raisebox{2\baselineskip}[0pt][0pt]%
%  {\parbox{\columnwidth}{\section{Introduction}\label{sec:introduction}%
%  \global\everypar=\everypar}}%
%  \vspace{-1\baselineskip}\vspace{-\parskip}\par
%\else
%  \section{Introduction}\label{sec:introduction}\par
%\fi
%
% Admittedly, this is a hack and may well be fragile, but seems to do the
% trick for me. Note the need to keep any \label that may be used right
% after \section in the above as the hack puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A}{n} impressive number of new software startups are launched worldwide every day as a result of an increase of new markets, accessible technologies, and venture capital \cite{8491286}.  With the term \textit{software startups} we refer to those temporary organizations focused on the creation of high-tech and innovative products, with little or no operating history, aiming to grow by  aggressively scaling their business in highly scalable markets.

New ventures such as \textit{Facebook}, \textit{Linkedin}, \textit{Spotify}, \textit{Pinterest}, \textit{Instagram}, \textit{Groupon} and \textit{Dropbox}, to name a few, are good examples of startups that evolved into successful businesses. But, despite many successful stories, the great majority of startups fail within two years from their creation, primarily due to self-destruction rather than competition \cite{Crowne2002}. Operating in a chaotic, rapidly evolving and uncertain domain, software startups face intense time-pressure from the market and are exposed to tough competition \cite{Maccormack2001, Eisenhardt1998}. In order to succeed in this environment, it is crucial to choose the right features to build and be able to quickly adapt the product to new requests constrained by very limited amount of resources \cite{Sutton2000}.

From a software engineering perspective development in startups is challenging since they develop software in a context where processes can hardly follow a prescriptive methodology \cite{Sutton2000, Coleman2005}. Despite startups share some characteristics with similar domains (e.g. small and web companies), the combination of different factors makes the specific development context quite unique \cite{Blank2005, Sutton2000}. Research is needed to support their engineering activities \cite{Coleman2005}, guiding practitioners in taking decisions and avoiding choices that could easily lead the whole business to failure \cite{Kajko-Mattsson2008}.
However, despite the impressive size of the startup ecosystem \cite{ISI:000243253000007}, the state-of-the-art presents a gap. Through a \textit{Systematic Mapping Study} (SMS) of the literature, conducted by Paternoster et al. \cite{SMS}, only few publications were found related to engineering activities in startups. Moreover the majority of these studies do not possess the relevance and rigour required to form a consistent body of knowledge in the state-of-the-art and in the state-of-practice.

This research aims to understand how software development strategies are engineered by practitioners in startup companies in terms of level of: structure, planning and control of software projects, in the period of time  that goes from idea conception to the first open beta release of the software product.

With a cross-sectional study we focused the research context in a limited time-frame to highlight the nature of uncertainty in the development activities and time pressure from the market, as discussed in \cite{Maccormack2001, Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and CTOs of startups, covering a wide spectrum of thematics and iteratively adjusting the direction of the research according to the emerging evidences.

By means of the Green-field startup model (GSM), we captured the underlying phenomenon of software development in early-stage startups. The framework is based on a systematic procedure  which is grounded into empirical concepts obtained by conduction of the case-study. GSM presents the most significant themes of the development strategies that characterise startups' context.

The most urgent priority of software development is to shorten time-to-market. The greater part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).The first release of the product includes only a limited set of well suitable functionalities.

Nevertheless, by speeding-up the development process, startups accumulate technical debt, ignoring processes, relying on informal communication and replacing documentation with low-precision artifacts. For startups it is essential to look ahead within short-term deadlines and determine the right product/market fit driven by user feedbacks, using flexible and reactive development approaches and adopting mitigation strategies to contain the severity of accumulated technical debt.

Practitioners can use the GSM model to apply presented strategies to speed-up the development by taking in consideration the amount of technical debt accumulated during the first early-stage period. In this regard, we identified several commonalities between the issues related to software development in startups and the research focused at studying technical debt \cite{Nugroho2011,Izurieta2012}. On the other hand, researchers can take advantage from the GSM model to understand how the technical debt influences the future growth of startup companies.

The remain of this paper is structured as follow:  we provide relevant information about startups in section \ref{backg}. Section \ref{resmet} introduces the research questions and discusses the design and execution of the research process. Results are discussed in section \ref{res:gsm} which presents the GSM model. Section \ref{res:val}  compares results of this research to literature and frameworks. Section \ref{valt} discusses the validity threats considered during this study. Finally, this paper concludes in section \ref{conc}.

% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)

%\hfill mds

%\hfill January 11, 2007
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\section{Background}
\label{backg}

Modern entrepreneurship, which was born more than thirty years ago \cite{storey1982entrepreneurship}, has been boosted by the advent of the consumer internet markets in the middle of the nineties and culminated with the notorious \textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126}. Several years later, with the massification of the internet and mobile devices, we are now assisting to an impressive proliferation of software ventures - metaphorically referred as the \textit{startup bubble}. The easy access to vast potential markets and the low cost of services distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}. Inspired by stories of overwhelming successes, a large number of software businesses are created every day. However, the great majority of these companies fail within two years from their creation \cite{Crowne2002}.

Looking at the number of new business incubators which appeared in the last three years one can evaluate the importance of startups \cite{Grimaldi2005}. The wave of disruption in new technologies has led companies to be more and more competitive, forcing themselves to radical organizational and innovational renewals,  which bring many companies to the attempt of behaving like startups \cite{Christensen1997}.

The implementation of methodologies to structure and control the development activities in startups is a major challenge for engineers \cite{Coleman2008}. Generally, the management of software development is achieved through the introduction of software processes, which defineswhat steps the development organizations should take at each stage of production and provide assistance in making estimates, developing plans and measuring quality \cite{Chroust1996}. In the last decades, several models have been introduced to control software development activities. However, their application in startup companies doesn't report significant benefits \cite{Coleman2008a, Coleman2008, Sutton2000}.

In this context, software engineering (SE) faces complex and multifaceted obstacles in understanding how to manage development processes. Sutton defines startups as creative and flexible in nature and reluctant to introduce process or bureaucratic measures which may result ineffective \cite{Sutton2000}. Also Bach refers to startups as \textit{``a bunch of energetic and committed people without defined development processes''} \cite{Bach1998}. In fact, startups have very limited resources and typically wish to use them to support product development instead of establishing processes \cite{Coleman2008, Heitlager2007}. Some attempts to taylor lightweight processes to startups, reported basic failure of their application : \textit{``Everyone is busy, and software engineering practices are often one of the first places developers cut corners''} \cite{Martin2007}. Rejecting the notion of repeatable and controlled processes, startups prominently take advantage of unpredictable, reactive and low-precision \cite{surviving-os-cockburn} engineering practices \cite{Sutton2000, Tanabian2005, Chorev2006, Kakati2003}.

Moreover, most startups develop packaged applications rather than software for a specific client \cite{genome2012}. Issues related to this domain are addressed in literature by the area known as \textit{market-driven software development} (sometimes called \textit{packaged software development} or \textit{COTS software development} \cite{Alves2006}). Among other results, researchers emphasize the importance of time-to-market as a key strategic objective \cite{dagMDR,dagMDR,sawyer99} for companies operating in this domain.  Other peculiar aspects which influence the software development are related to requirements, which are reported to be often ``invented by the software company'' \cite{512553}, ``rarely documented'' \cite{Karlsson02challengesin}, and can be validated only after the product is released to market \cite{dahl2003,Keil:1995}. Under these circumstances, failure of product launches are largely due to ``products not meeting customer needs''  \cite{Alves2006}.

Accordingly, product-oriented practices help startups having flexible teams, with workflows that leave them the ability to quickly change the direction according to the targeted market  \cite{Heitlager2007,Sutton2000}. In this regard, many startups focus on team productivity, asserting more control to the employees instead of providing them rigid guidelines \cite{Tanabian2005, Chorev2006, Kakati2003}.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Lack of research in startups}
Despite some studies tried to address the above mentioned issues, \cite{SMS} found only a few SE works in this specific area, as confirmed by other studies \cite{Coleman2008, Coleman2008a, Coleman2007, Sutton2000}. Moreover the studies, identified in our systematic review, appear to be highly fragmented and spread across different areas rather than constituting a consistent \textit{body of knowledge}.

A research published in 2000 by Sutton \cite{Sutton2000} confirmed a general lack of  studies in this area, claiming that \textit{``software startups represent a segment that has been mostly neglected in process studies''} and it has been further confirmed with the empirical studies of Coleman et al. \cite{Coleman2008,Coleman2008a,Coleman2007} eight years later.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Definition of startup}
The word \textit{startup} appeared in the SE literature for the first time in 1994 in an article written by Carmel \cite{Camel1994a} where he studied the \textit{time-to-completion in young package firm}. He noticed how these companies were particularly innovative and successful, advocating for the need of more research investigating software development practices so as to replicate success and try to transfer it to other technology sectors.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Evolution of the process}

First insights reveal how software startups are product-oriented in the first period of their development phase \cite{Heitlager2007}. Despite good achievements at the beginning, software development and organizational management increase in complexity \cite{1456074,Banker1998} causing  deterioration of performance over time. Briefly, the necessity of establishing initial repeatable and scalable processes cannot be postponed forever.

A study of Kajko-Mattsson  \cite{Kajko-Mattsson2008}, which investigated a Swedish software startup, reported a heavy lack of requirements gathering process, minimal project management, lack of control over the change requests, absence of documentation to track the status and progress of the process and defective releases. Accordingly, Ambler et al. report how two startups approaching to an upcoming IPO started to require processes to focus on scalable solutions, in view of the growing company's size in terms of users and employees \cite{Ambler2002}. In this regard, Crowne, in \cite{Crowne2002}, specifies different stages through which software startups evolve. Starting without any established workflows, startups grow over time, creating and stabilizing processes to eventually improving them only when sufficiently mature.

But when startups have no time for training and orienting activities, as discussed in \cite{Sutton2000}, their main focus remains on team capabilities instead of prescriptive processes hiring people who can \textit{``hit the ground running''} \cite{Yoffie1999}. Empowering the team and focus on methodological attributes of the processes oriented in prototyping, proof-of-concepts, mock-ups and demos, to test basic functionalities, have been the primary priority in startups as described in \cite{Camel1994a}. Only when they grow, formal methodologies arise, followed by a conduction of quality assurance and long-term planning processes \cite{Yoffie1999}.

As studied in \cite{Tingling2007}, the maturity of a company affects the extent to which processes are adopted. The author reports how introducing Extreme Programming (XP) principles \cite{Beck:2004:EPE:1076267} in the development process was challenging because of the need of trained team-members for fully implementing the methodology. In fact, \cite{Silva2005} was able to start with all the XP practices in place only after six months of coaching the team, trying to enhance maturity from day-one. Nevertheless, even then, customization of practices were inevitably implemented to adapt the processes to the undertaken startups' context \cite{Deias}.

As partially discussed above, contributions to flexibility and reactiveness of the development process has been conducted prominently by means of \textit{Lean} \cite{Gautam2008} and \textit{Agile} \cite{Abrahamsson2002} methodologies (also reported in \cite{Taipale2010, Kuvinka2011}), where the extreme uncertain conditions lead startups to learn fast from trials and errors with a strong customer relationship in order to avoid wasting time in building wrong functionalities and prevent rapidly exhaustion of resources \cite{Midler2008, Hilmola2003, Sutton2000}. Customer involvement in software development has also been discussed in  \cite{Yogendra2002} as important factor to encourage an early alignment of business concerns to the technology strategies, because both are salient considerations to be successful \cite{Tingling2007}.

When startups take a development approach that is mainly product-oriented rather than process-oriented, it is essential to have a flexible team, with a workflow that helps them quickly to change direction according to the target market \cite{Sutton2000}. In this regard, many startups have been focused on team productivity, providing more control to the employees instead of providing them rigid guidelines \cite{Tanabian2005, Chorev2006, Kakati2003}.

Then, when startups are mature enough to support software process improvement (SPI), the solutions considered according to the state-of-the-art are oriented towards light-weight processes such as a design of development process based on XP, proposed by Zettel in \cite{Zettel2001}. The process consists of a set of activities and artifacts (in addition to some important roles) defined in order to identify responsibilities and tools to utilize. But, despite the promising benefits reported by Zettel, we were not able to identify any future evaluation in real-world settings.

Another attempt of SPI in startups has been conducted by Deakins et al. introducing a \textit{Helical Model for managing e-commerce development environment} \cite{Deakins2005}. Also in this case, the author prescribed broad guidelines for a rapid high-quality development process, which underwent limited testing only in academic settings.

The first publication mentioning the problem of \textit{one-size-fits-all}, related to the SPI  representations for startups, is described in \cite{Fayad1997}. The author reveals the problem in actuating the same \textit{best-practices} criteria for established companies in 10-person software startups. Thoroughly remarked in \cite{Sutton2000}, Sutton states that problems of SPI in software startups arise because of: the dynamism of the development process, which precludes repeatability; organizational maturity, that cannot be maintained from startups in view of lack of corporate direction; severe lack of resources, both human and technological for process definition, implementation, management, training \ldots; in conclusion, the primary benefits of SPI do not address startups, which instead of promoting product quality, aim to minimize time-to-market.

Additionally, the role of SPI has always been neglected because seen as an obstacle to the development team's creativity and flexibility as described in \cite{Coleman2008a} and to the need of a quick delivering product process environment. In fact, product quality is left aside in favour of minimal and suitable functionalities to shorten the time-to-market. As reported in two studies of Mater and Mirel \cite{Mater2000, Mirel2000}, quality aspects, mostly taken in consideration in internet startups, are oriented to usability and scalability, even though the market and application type heavily influences the quality-demand \cite{Coleman2008, Kim2005}.

Finally, to maintain the development activities, oriented to limited but suitable functionality, many studies suggest to externalize the complexity of parts of the project to third party solutions by means of outsourcing activities, software reuse and open-source strategies \cite{Hanna2010, Jansen2008, Wall2001, Bean2005}.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Technical debt}
Notwithstanding, a new interesting area of the SE research, trying to tackle the problem of the \textit{technical debt}, seems to bring interesting implications in studying development in software startups. The metaphoric neologism was originally introduced by Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the attention of SE researchers\footnote{To attest the fact that \textit{technical debt} is gaining traction among researchers, we mention two important contributions which characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt} organized by the Software Engineering Institute  and ICSE.}. The concept of \textit{technical debt} is well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea is that developers sometimes accept compromises a system in one dimension (e.g., modularity) to meet an urgent demand in some other dimension (e.g., a deadline), and that such compromises incur a ``debt'' on which ``interest'' has to be paid and which the ``principal'' should be repaid at some point for the long-term health of the project''}.  The compromise between high-speed and high-quality engineering is faced daily by software startups, not only in terms of architecture design but in multifaceted aspects (weak project management, testing, process control, \ldots).

In conclusion, since \textit{``all decisions related to product development are trade-off situations''} \cite{Hilmola2003}, generally startups optimize workflows to the dynamic context they are involved into. In fact they typically adopt any development style that might work to support their first needs in what is called the \textit{``Just do it''} school of software startups \cite{Ries2011}. Additionally, as remarked by Coleman, \textit{``many managers just decide to apply what they know, as their experience tells them it is merely common sense''} \cite{Coleman2008}.

In this paper we make use of context-specific terminologies, which will be used to facilitate the readability and clarify any ambiguities throughout the whole paper in view of the weak state-of-the-art presented in \cite{SMS} :

\begin{compactitem}
\item Software development strategy: the overall approach adopted by the company to carry out the product development.
\item Engineering activities: the activities needed to bring a product from idea to market. Traditional engineering activities are, among others, requirement engineering, design, architecture, implementation, testing.
\item Engineering elements: any practice, tool and artifacts contributing and supporting the engineering activities.
\item Quality attributes: those overall factors that affect run-time behavior, system design, and user experience. They represent areas of concern that have the potential for applications to impact across layers and tiers. Some of these attributes are related to the overall system design, while others are specific to run time, design time, or user centric issues \cite{Microsoft2009}.
\item Growth: an increase in the company size respect to the initial conditions in terms of either employees or users/customers, and product complexity in terms of handling an increasing number of feature requests.
\item Software product: any software product and software service.
\end{compactitem}
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\section{Research methodology}
\label{resmet}

\begin{figure*}[!t]
\centering
\includegraphics[width=4in]{figures/completemethodology2}
\caption{Grounded theory study process overview}\label{fig:gt:completemethodology}
\end{figure*}

The aim of this research is to \textit{understand how software development strategies are engineered by practitioners in startup companies in terms of level of: structure, planning and control of software projects, in the period of time  that goes from idea conception to the first open beta release of the software product}.

The above mentioned goal is structured according to the GQM paradigm \cite{Basili1992}.

Initially the boundaries of the research domain were set by means of non-systematic \textit{literature survey}, which provided the initial keywords used further in the research process. Moreover it helped us in defining the research questions (RQs):

\begin{compactitem}

\item RQ-1 : What is the current state-of-practice related to software development strategy in early-stage startups?
\begin{compactitem}

\item RQ-1.1 : How do startups structure and execute the main engineering activities?
\item RQ-1.2 : How are product quality attributes considered by startups?
\end{compactitem}

\end{compactitem}

We investigated the software development approach undertaken by practitioners of startups.  Following a \textit{Grounded Theory} (GT) methodology \cite{Glaser1978}, we executed 13 semi-structured interviews integrated with follow-up questionnaires. From this, we elaborated and extracted the Green-field startup model (GSM) explaining the underlying phenomenon of software development in startups.

\textit{Grounded Theory} methodology is described by Robson as \textit{``the best approach to answer the question - What is going on here?''} and defined by its creators \cite{ColinRobson2009} as \textit{``a set of well-developed categories ( e.g. themes, concepts ) that are systematically interrelated through statements of relations to form a theoretical framework that explains some relevant social, psychological, educational, nursing or other phenomenon''} .

Despite GT has its roots in the social sciences, it has been increasingly used by \textit{Information System} researchers and has been almost ignored by the SE research \cite{Coleman2007}. In 1997 Bertelsen advocated for a need of more qualitative research in SE \cite{Bertelsen1997}, and in the last decade we assisted to an increasing number of publications that used a GT-like approach in SE \cite{Coleman2008a,Coleman2007, Sulayman2012}.

Following the \textit{grounded theory} principles we attempted to capture the most relevant aspects of software development from startup practitioners, letting a theory emerge from the interviews and adjusting the research hypotheses and questions as we proceeded through. During these interviews we collected data related to engineering activities undertaken by startups in the time frame between the idea conception and the first open beta release of the product.  Then, we proceeded to the analysis of the data, finding important relations among concepts with a formal approach.

As suggested by Coleman, in view of the different versions of \textit{grounded theory} which emerged in the last years, researchers should indicate which \textit{``implementation''} of the theory is being used \cite{Coleman2007}. In fact, after the methodology was initially introduced, its original authors  had some divergences regarding how the theory should be formed by analyzing the data (Strauss and Corbin on one side, Glaser on the other).  Glaser advocated for a more \textit{puristic} approach, where the theoretical categories should \textit{``naturally emerge from the data''}, whilst Strauss and Corbin formulated an updated version of the methodology, which leaves to the researchers an higher degree of freedom. The latter approach empowers researchers' \textit{``theoretical sensitivity''} \cite{Corbin1990}, and encourages them to outline the research problem beforehand. Since the knowledge obtained from parallel execution of the SMS \cite{SMS} and our direct experience with startups companies provided a good initial level of knowledge, in this study we use the Corbin and Strauss approach \cite{Strauss1998}.

In this research we addressed technical aspects related to software development in startups, exploring their operational dynamics. In view of the lack of agreement on an unique definition of the word \textit{startup}, we delimited our initial contextual boundaries to newly created innovative and single-product software companies, in the time-frame that goes from the idea conception to the release of the first product in highly scalable markets. Moreover, since most startups are web-oriented companies \cite{tc-webstartups-link, Allen2003}, in our case study, we inquired web companies with little or no operating history.

A complete overview of the overall case study methodology and execution is shown in Figure \ref{fig:gt:completemethodology}, which presents how we tailored the general GT methodology to our specific needs.

The initial design of the case study is supported by the results of the \textit{systematic mapping study} (SMS) \cite{SMS} which contributed to define the initial questions for both questionnaires and semistructured interviews. The process depicted in Figure \ref{fig:gt:completemethodology} is evolutionary and affects the design at each new iteration.  In \textit{data collection} the empirical results are stored in a single \textit{case study database} and subsequently processed in \textit{data analysis} to form the theoretical categories.  At each iteration the new emergent theory is updated following a formal procedure (\textit{Theory generation}), and after verifying that the \textit{theoretical saturation}\footnote{The point at which executing more interviews wouldn't bring any additional value for constructing the theory.} of categories has been achieved, we finally proceeded to \textit{theory validation} or performed another complete cycle.

The whole procedure has been executed simultaneously in pair on the same screen by the first two authors, handling conflicts by reviewing the rationale of decision with the third and fourth authors. When necessary we performed an in-depth review of the design of the research and data collected during the execution process.

The detail process is thoroughly described in the following subsections which are structured according to the five macro phases depicted in bold in Figure \ref{fig:gt:completemethodology}.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Design and execution}

By means of the SMS \cite{SMS} we refined our research questions. The research questions have been narrowed down to outline the domain of our study, even though still broad enough to allow emerging new theories and to adjust the scope of the research \cite{Corbin1990}.

The sampling strategy took place in two distinct phases. At first we executed an initial \textit{convenience sampling} \cite{Dawson2009}, which led to the identification of eight companies. Then we included five additional startups during the theory formation process (\textit{theoretical sampling}) iteratively improving the sample according to the emerging theory.  Additional information about the companies are reported in Table \ref{t_interviews-stats}.

\begin{table*}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Companies overview} 
\label{t_interviews-stats}
\centering
\begin{tabular}{|c||c||c||c||c||c|}

\hline
    ID  &  Company age & Founding & Initial & Current & First product  \\
          &   (months)   & team   & developers  & employees      &  building time    \\
                    &      & &   &  & (months)   \\
   \hline
C1    & 11    & 4     & 2     & 11    & 6   \\
C2    & 5     & 2     & 2     & 6     & 3      \\
C3    & 18    & 4     & 4     & 4     & 6      \\
C4    & 17    & 3     & 2     & 11    & 6     \\
C5    & 20    & 2     & 1     & 4     & 12   \\
C6    & 30    & 3     & 2     & 4     & 1      \\
C7    & 12    & 2     & 1     & 7     & 4      \\
C8    & 24    & 4     & 3     & 16    & 4    \\
C9    & 5     & 5     & 4     & 5     & 1     \\
C10   & 43    & 6     & 4     & 9     & 4      \\
C11   & 36    & 3     & 3     & 6     & 2     \\
C12   & 12    & 3     & 3     & 3     & 3      \\
C13   & 24    & 2     & 2     & 20    & 3     \\

\hline
\end{tabular}
\end{table*}

As shown in column \textit{Company age}, most of the companies were founded within the last three years with only one exception (C10, which age was 3 years and 6 months). As can be observed by looking at the column \textit{Founding team}, the companies have been founded by an average number of  3 members, which in majority are developers, as can be seen by looking at the column \textit{Initial developers}. Moreover, the number of \textit{Current employees} shows how, to different degrees, companies expanded the initial teams, being able to bring the product to the market in less than 6 months from the idea conception\footnote{C5 building time is extended to 1 year mainly because the product has been developed part-time by one person only.} (see \textit{First product building time}).
Summarizing the above identified features, the sample of the case study is mainly composed by recently founded companies, working with web applications in different nations and market sectors, with very small development teams. Moreover, the growing team size and and the publicly available data suggest an healthy status of the businesses.

We executed the case study online, supported by tools for video conferencing recording each complete session. Each interview has been executed only once per company with the interviewee's role of CEO or CTO. A step-by-step workflow have been followed, which allowed us to perform the interviews, collect additional material, prepare the customized follow-up questionnaire and iteratively adjust the \textit{interview package} artifacts\footnote{The \textit{interview package} and {follow-up questionnaire} has been released under MIT License \cite{MITLicense} and it is available for download on Github \cite{GitHubInterviewPackage}. We encourage anyone who has interests in pushing this work forward, to fork the repository and contribute to it. If used in a research context, please inform us in order to be able to track where and how the \textit{interview package} has been used.}.

The follow-up questionnaires were designed to capture additional data, gather missing information and confirm interview results with triangulation. Questionnaires have been tailored to each startup, partially taking advantage of the repertory grid principles \cite{Edwards2009}. To obtain better chances of quickly obtaining responses, questionnaires have been sent to respondents immediately after transcribing the interview results. Two weeks after the conclusion of the last interview, we closed our follow-up questionnaires and collected the data.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Data collection}
The approach undertaken to collect the data, known as \textit{data triangulation}, integrates multiple data sources converging on the same phenomenon.
After transcribing the interview, we extracted conceptual relations, and then we integrated them with the questionnaire's results. A well structured case-study database allowed us to easily retrieve and seek for information, assembling the evidence from the case study reports, as described also in \cite{Yin1994}. The database has been stored and constructed using the qualitative data analysis software package AtlasTI\footnote{Available online at \url{http://www.atlasti.com/} .} - one of the most suitable tools for GT \cite{Coleman2007}.

Entering the field, we overlapped interviews with questionnaire results to adjust the data collection and take advantage of emergent themes and reveal possible inconsistencies. The data was analyzed simultaneously and with flexibility in mind in such a way that adjustments were made according to the emerging findings.
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Data analysis}
Before starting the analysis, a \textit{data ordering} procedure was necessary since interviews were spread across a multitude of topics. Therefore transcripts have been structured in thematic areas accordingly to different \textit{topic cards} used during the interviews. We proceeded horizontally between same thematic areas of transcripts to be able to identify a better number of similar concepts, rather than going through an entire transcript at one time.

Once the data were ordered, we executed the process of \textit{coding} interviews, following the steps listed below:

\begin{compactitem}

\item Labels were assigned to raw data, and a first low-level conceptualization was carried out using both \textit{in-vivo} and \textit{open coding} \cite{ColinRobson2009}.
\item Concepts were grouped together into theoretical categories and subcategories. By means of \textit{axial coding} we first described the different relations between subcategories, and then relations between subcategories and categories.
\item Categories were refined several times in the attempt to create different level of abstractions and adjusting concepts, aided by a simple knowledge management tool.
\item Consistency among categories were validated by exploring and analyzing links  among subcategories by means of the \textit{selective coding}.
\item The \textit{core category} - the one with the greatest explanatory power -  was identified by analyzing the causal relations between high-level categories.
\end{compactitem}

During data extraction we used the technique of \textit{in-vivo coding} since \textit{``direct quoting from the transcript give more expressive power to the data''} \cite{ColinRobson2009} combined with the more descriptive procedure of \textit{open coding}. Following the example of other grounded theories, developed in related areas such as Information Systems \cite{Orlikowski1993} and Software Process Improvement \cite{Coleman2006},  we performed the high-level conceptualization during creation of categories, in the process of refining axial and selective coding.

As we were iterating through the interviews, we analyzed new data accordingly, by updating \textit{codes} and \textit{categories} on necessity, and taking notes in the form of  memos to adjust the new \textit{emerging theory}. To improve the speed of the coding process, we analyzed all the transcripts one part at-the-time: transcripts followed the semi-structured interview format, and this approach enhanced the chances of finding similar codes across different transcripts, allowing us to quickly iterate and update categories.

Given the large number of codes, categories, properties, propositions and related questions that evolved from the analytical process \cite{Corbin1990}, an important activity that helped the coding process was \textit{memoing}. It constituted an important component involved in the formulation and revision of theory during the research process. For this purpose we made use of three different types of memos: code memos, theoretical memos and operational memos. The first type of memo is related to the conceptual labeling of open codes, whilst the second type concerns axial and selective coding, and thus focusing on paradigm features. Finally, operational memos contain directions relating to the evolving \textit{emerging theory}.

After the coding process the first representation of the \textit{experience map} identified in GT is constructed by means of a \textit{theoretical framework}. The theoretical framework is presented in the form of a network of categories and subcategories identified during the coding process. Precisely, categories and subcategories are linked together according to cause-effect relation \cite{Corbin1990}. During the formation of the theoretical framework the researchers operated by means of a  bottom-up approach. From empirical data and coding process, the framework was developed into two different levels: a detailed level representing the network of subcategories (identified mainly by \textit{axial coding} process), and high-level representing the network of main categories (identified mainly by \textit{selective coding} process).

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Theory generation}

As mentioned above, emergent theories have been tested integrating the sample with additional companies selected following the principle of \textit{theoretical sampling} \cite{Yin1994}.

The process of \textit{theory generation} took place at each iteration together with interview execution, where we systematically analyzed the \textit{theoretical framework} by means of the \textit{paradigm model} introduced by Corbin \cite{Corbin1990}.

The paradigm model is composed by:

\begin{compactitem}
\item \textit{Causal conditions}, the events which lead to the occurrence of the \textit{phenomenon}, that is our core category.
\item \textit{Context}, set of conditions in which the phenomenon can be extrapolated;
\item \textit{Intervening conditions}, the broader set of condition that the phenomenon can be generalized.
\item \textit{Action/interaction strategies}, the actions and responses that occur as the result of the phenomenon.
\item \textit{Consequences}, specification of the outcomes, both intended and unintended of the actions and interaction strategies.
\end{compactitem}

Within the limits of the critical bounding assumptions, the role of the generated theory is to explain, predict and understand the underling phenomenon.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%

\section{Results: Green-field startup model}
\label{res:gsm}

By analyzing the interview transcripts with the GT process we extracted the information necessary to form the theoretical model. The initial coding process (open and in-vivo) provided 630 unique codes, grounded to 1295 citations in the transcripts, divided according to the thematic areas.

The thematic areas were divided according to the development phases designed for the interview process. The complete list of  raw codes is presented in Appendix and constitutes the empirical foundation of the theoretical categories, which has been used to form the theoretical framework (Green-field startup model).

In this section we present the GSM model to capture the underlying phenomenon of software development in early-stage startups. The GSM model is based on a systematic procedure which is grounded into empirical concepts obtained by conduction of the case-study. Starting from 630 raw \textit{codes} extrapolated from the interview transcripts, the framework is formed by 128 sub-categories, clustered in 35 \textit{groups}, and finally in 7 categories at the highest level of abstraction.

By the means of a conceptual framework we want to provide explanations of the development strategies and engineering activities undertaken by startups. The knowledge, which we attempted to pack into a conceptual model, can be used by researchers and practitioners to improve their understanding of the phenomenon.
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Framework overview}
\label{res:gsm:frmov}
Main concepts representing underlying phenomena have been grouped together to form high-level categories. Figure \ref{fig:gsm} shows the network of causal relationships (represented by arrows) between categories (represented by blocks). Each node is linked to subsequent nodes which are called successors, and preceding nodes which are called predecessors. The relations between them are denoted by the direction of the interconnecting arrows. The network is read from left to right.

\begin{figure*}[!t]
\centering
\includegraphics[width=4in]{figures/high-level}
\caption{Green-field startup model}\label{fig:gsm}
\end{figure*}

In the explanation of the GSM model we emphasize labels of  theoretical categories using the \textit{italic} font style. For the main categories shown in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e. CATx) to facilitate the reading of this section.
The network is centered around the core category, \textit{speed-up development}, which is the most interconnected node in the framework reflecting the fact that \textit{``it is the one (category) with the greatest explanatory power''}  \cite{Strauss1998}. The meaning of the labels on the three arrows that reach the core category (\textit{efficiency}, \textit{effectiveness} and \textit{performance}) is discussed in detail in section \ref{valt}.

A contextual condition which characterizes, to some extent, each and every startup, is the \textit{severe lack of resources}. In fact, the capabilities of an early-stage startups' to support its development activities are constrained by an extremely limited access to human, time and intellectual resources, eventually leading to a general absence of structures and processes.

The \textit{severe lack of resources} forces the company to focus on implementing an essential set of functionalities and it is one of the main reasons why the \textit{product quality has low priority} respect to other more urgent needs\footnote{There are some exceptions where the quality aspects actually matter and such cases will be discussed later in \ref{valt}.}. At the same time, to be able to deal with such constraints, startups must depend on a small group of capable and motivated individuals.

As it has been unanimously expressed by respondents, from a software perspective, the most urgent priority is to \textit{speed-up the development} as much as possible by adopting an extremely flexible and effective \textit{evolutionary approach}. On the other hand the efficiency of teamwork is facilitated by the low attention initially given to architectural aspects related to product quality. This allows startups to have a faulty but functioning product, which can quickly be introduced to the market beginning with the in-house implementation of the prototype from day-one.

The initial employees are the ingredients which enable high levels of performance in software development. To support a fast-paced production environment, engineers are required to be highly committed, co-located, multi-role, and self-organized. In other words \textit{team is the catalyst of development}.
With an essential and flexible workflow, which relies on tacit knowledge instead of formal documentation, startups are able to achieve very short time-to-market. However, each line of code, written without following structures and processes, contributes to grow the \textit{accumulated technical debt}, which is further increased by having almost non-existing specifications, a minimal project management  and a severe lack of automated tests.

Despite the fact that the consequences of such debt are not clearly perceived in the initial stages (where finding the product/market fit as quickly as possible is the most important priority), startups, which survive to subsequent phases, will likely increment their user-base size, the product size, and the number of developers. This will require the company to eventually pay the \textit{accumulated technical debt}, confronting the fact that an \textit{initial growth hinders productivity}.

In the subsequent subsections we are going to explain in detail all the categories presented in Figure \ref{fig:gsm}, to conclude in subsection \ref{res:gsm:th} with the final theory. During the following explanations we are going to use identifiers of the companies presented in Table \ref{t_interviews-stats} (i.e. Cx) to highlight statements recorded from the interviewees.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Severe lack of resources}
\label{res:gsm:cat7}
The concept of \textit{severe lack of resources} characterizes the uncertainty of the development strategies in startups and it is composed by three subcategories: \textit{time-shortage}, \textit{limited human resources} and \textit{limited access to expertise}.

Since startups want to bring the product to market as quickly as possible, the resource they are the most deprived of is time. Startups operate under a constant time pressure, mainly generated by external sources (\textit{investor(s) pressure}, \textit{business pressure}) and sometimes internal necessities such as \textit{internal deadlines} and \textit{demo presentations at events}. In this regard, C3 commented: \textit{``Investors wanted to see product features, engineers wanted to make them better. Finally the time-to-market was considered more important and the team interests was somehow sacrificed.''}

In addition, to compensate the \textit{limited human resources},  practitioners need to empower \textit{multi-role and full stack engineers}, as confirmed by C1: \textit{``Everyone was involved in any tasks, from mobile to web development, organizing themselves in choosing the part to implement''}.

The extent to which startups have access to specialized knowledge - both internal and external to the company - is extremely reduced when compared to traditional established software companies. Therefore, in order to partially mitigate the \textit{limited access to expertise}, startups rely on the external aid of mentors or advisors. Under these strict limitations, most of the decisions related to software development are fundamentally trade-off situations (confirming Himola's results \cite{Hilmola2003}).
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Team is the catalyst of development}
\label{res:gsm:cat4}
Among the different aspects fostering the speed of the development process, the startups' main focus is on the characteristics of the initial team.

In startups \textit{developers have big responsibilities}. In fact, the \textit{limited human resources}, discussed in CAT7, causes the team-members to be active in every aspect of the development process, from the definition of functionalities to the final deployment.

Early engineers in the founding team of a startups are typically \textit{multi-role and full-stack engineers}: they are full-stack for being able to tackle different problems at different level of the technology stack (\textit{generalists developers instead of specialists}) and multi-role since usually \textit{engineers are responsible for marketing/sales/development}. As remarked by  C11: \textit{``Instead of hiring gurus in one technology, startups should hire young developers, generalists, that know how to quickly learn new technologies, and quickly move among a huge variety of tasks.''}.

Moreover, having a \textit{very small and co-located development team} enables members to operate with high coordination, control and communication, which heavily rely on \textit{tacit knowledge},  replacing most of the documentation with informal discussions. Practitioners reported that keeping the development team small helps startups in being fast and flexible, as remarked by C8: \textit{``If you have more than 10 people, it is absolutely impossible to be fast''}. Then, also \textit{previous working experience} and \textit{knowing each other before starting the company} enforce the efficiency of activities by \textit{limiting th need of formalities between team-members}.

In every software company, \textit{skilled developers are essential for high speed} in development. Especially in startups, the ``hacking culture'' and a tendency to the ``just-do-it'' approach allow the team to quickly move from the formulation of a feature idea to its implementation. In this regard, C1 comments: \textit{``we had a hacker culture/environment, people hacking stuff without formally analyzing it, but breaking it down and finding a way around. Ticket on trello and go.''}

A \textit{limited access to expertise} forces the team to rely mainly on their personal abilities, even though asking opinions to mentors is reported to be a viable practice to maintain feasible objectives. Furthermore \textit{teams work under constant pressure}, \textit{``up to 16 hours per day''}, mainly constrained by a tight \textit{time shortage}.

Finally, startups present founders-centric structures, and especially in the early-stage, it is clear that \textit{CTO/CEO background has high-impact} on the company's development approach. For instance, in case of an academic background, the CTO encourages the introduction of some architectural design before the development phase. Moreover, even though the CTO/CEO starts guiding initially the development process, most of the decision are taken consensually by all members of the team. Then, the CTO/CEO decides only in particular situations where some conflicts occur.
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Evolutionary approach}
\label{res:gsm:cat2}
Among different approaches to development, startups prefer to build an initial prototype and iteratively refine it over time, similarly to the concept of \textit{``evolutionary prototyping''} \ref{EvProt}.  The reason of using this approach is that startups want to validate the product against the market as soon as possible, finding the proper product/market fit. Indeed, they can focus on developing only parts of the system they want to validate instead of working on developing a whole new system. Then, as the prototype is released, users detect opportunities for new functionalities and improvements, providing their feedback to developers.

Since \textit{flexibility and reactiveness are the main priorities}, the most suitable class of software development approaches are clearly highly evolutionary in nature. In fact, as \textit{uncertain conditions make long-term planning not viable}, startups cannot base their work on assumptions without rapidly validating them, releasing the product to market. The uncertainty is first of all in the team composition. In fact, since the teams are typically very small and the project knowledge is not documented, even a little change in their composition (e.g. a developer falls in ill) can have a very high impact on the overall product development. Furthermore, startups operate in a continuous evolving environment in terms of competitors and targeted market sectors. Then, to obtain a competitive advantage in the market, startups typically make use of cutting-edge solutions, characterised by an evolution that cannot be foreseen in the long run. However, a special role in daily decisions is covered by user feedback and requests, which represent the main drivers to define the product features to provide in the short-term.

To obtain fast user responses and quickly \textit{validate the product}, startups \textit{build a functioning prototype and iterate on it} over time. In fact, paraphrasing C4, \textit{``[\ldots] you should start with something that is really rough and then polish it, fix it and iterate. We were under constant pressure. The aim was to understand as soon as possible the product market/fit iterating quickly, adjusting the product and releasing fast.''}

The company focus on building a small set of functionalities to include in the first version, and \textit{progressively rolls-out to larger number of people} with very \textit{small iterations} (confirmed again by C4: \textit{``we deploy from 5 to 20 times a day''}).

The scope of this evolutionary approach is to avoid wasting time in ``over-engineering the system'' and building complex functionalities that have not been tested on real users. By releasing a very small number of good-enough functionalities (see CAT3) the startup can verify the suitability of the features and understand how to adjust the direction of product development towards actual users' needs. The first version of the product is typically a prototype which contains the basic rough functionalities developed with the least possible effort that can validate critical features which can undermine the startup's survival in the short term (recalling  Ries' definition of \textit{Minimum Viable Product} (MVP) \cite{Ries2011}).

Thanks to a \textit{direct contact and observation of users}, \textit{automated feedback collection} and \textit{analysis of product metrics}. The product metrics mainly refer to the UX of the product to find any kind of frictions that might reduce the satisfaction level of the final users. Startups attempt to \textit{find what is valuable for customers} with activities in agreement with Blank's \textit{Customer validation} stage \cite{Blank2005} .

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Product quality has low priority}
\label{res:gsm:cat3}
The main interests of software startups, related to the product, are mainly concentrated on building a  \textit{limited number of suitable functionalities} rather than fulfilling restrictive non-functional requirements. This strategy allows them to quickly release simple products with less need of preliminary architectural studies.

Exploring quality aspects considered during the development process, the only important concerns are expressed in favour of the user experience\footnote{According to the ISO 9241-210 (Ergonomics of human-system interaction), UX is defined as \textit{``a person's perceptions and responses that result from the use or anticipated use of a product, system or service''.} Even though this definition leaves so much to interpretation, we can refer to it as presented in the glossary.}, in terms of \textit{ease of use, attractiveness of the UI} and most of all, \textit{smooth user-flow without interruptions}. The fact that \textit{UX is the only important quality} is remarked by C11: \textit{``When a user needs to think too much on what action should be done next, he will just close the application without returning''} and confirmed by C3: \textit{``If the product works, but it is not usable, it doesn't work''.}

The extent to which quality aspects are taken into account might heavily depend upon the market sector and the type of application. Nevertheless, realizing high level of UX  is often the most important attribute to consider for customer discovery of  evolutionary approaches in view of the \textit{limited human resources} and \textit{time shortage}, presented in CAT7. C4 confirms: \textit{``None of the quality aspects matter that much as the development speed does.''}

To achieve a good level of UX while dealing with lack of human resources and time shortages, startups can analyze similar products of bigger companies which can afford more rigorous usability studies. Then, the users' feedback and product metrics begin to have a central role in determining the level of UX achieved. Product metrics are typically web-based \textit{statistical hypothesis testing}, such as A/B testing.

Other than UX, some other factors can influence the quality concerns of development:

\begin{compactitem}

\item The \textit{efficiency emerges after using the product}, letting engineers avoid wasting time in excessive improvements of not-validated functionalities. In fact, the level of efficiency can be optimized after attesting the effectiveness of the minimal set of functionalities, obtained according to CAT2 and to the concept of MVP.
\item The \textit{product should be reasonably ready-to-scale} to be able to accommodate a potential growth of the user-base\footnote{Note that startups tackle fast growing markets which are particularly subject to sudden user growths}. Thanks to modern cloud services, startups can \textit{externalize complexity to third party solutions} achieving a good level of scalability with a reasonable effort.
\item Realizing high reliability is not an urgent priority since \textit{users are fault-tolerant in innovative beta products}. In these cases, users have typically a positive attitude towards the product, even though it presents unreliable behaviours. In this regard, the main focus of beta testing is reducing frictions between the product and the users, often incorporating usability testing. In fact, the beta release is typically the first time that the software is available outside of the organization that developed it. \footnote{Detailed discussion of the impact of innovative products on the user satisfaction is presented in Appendix according to the Kano model.}.
\end{compactitem}

Finally, startups implement a \textit{limited number of suitable functionalities}, which in view of \textit{limited human resources} and \textit{time shortage}, represent a viable strategy to focus on shortening time-to-market.%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Speed-up development}
\label{res:gsm:cat1}
As remarked above in the model, \textit{speed-up development} represents the core category of the GSM model. Firmly grounded as the primary objective of startups, it shows the most important characteristic of developing software in the early stage.

As mentioned before, in order to \textit{speed-up development}, startups adopt evolutionary approaches supported by a solid team focusing on implementing a minimal set of suitable functionalities. Startups \textit{keep simple and informal workflows} to be flexible and reactive, adapting to the fast changing environment. While a rigorously defined workflow generally consists of an established sequence of well-defined steps to follow, startups, by contrast, adopt informal workflows. This choice is mainly dictated by contextual factors characterized by the general conditions of unpredictability and uncertainty. The adoption of informal workflows is facilitated by the fact that teams are typically self-organized and \textit{developers have big responsibilities} .

Additionally possible planning activities are refrained by the aim to shorten time-to-market, as reported by C8: \textit{``Speed was the essence so we didn't plan out too many details''}. To deal with such unpredictability, startups prefer to take decisions as fast as possible, mainly by means of informal and frequent verbal discussions.

Despite Agile principles appear to be suitable to embrace changes, especially in the early-stage, development practices are often perceived as time-wasters and ignored to accommodate the need of releasing the product to market as quickly as possible. In this regard, to maintain the simplest and most informal approach, startups aim to \textit{find the product/market fit quickly} and implement a \textit{limited number of suitable functionalities}. This approach is considered possible also in view of a lack of systematic quality assurance activities, since only the user experience is considered as important and other quality aspects, such as efficiency, can be postponed after the first release.

Another strategy, which supports startups in quickly delivering products, is the \textit{externalization of complexity on third party solutions} by: making use of third party components (COTS) and open source solutions (both for product components either development tools and libraries); taking advantage of  external services; and deploying the product on external infrastructures, for the sake of presenting a \textit{product reasonably ready to scale} for a possible future growth.

Even though \textit{the use of well-integrated and simple tools} allows startups to automate many activities and reduce their completion time, drawbacks of such approaches are the increased concerns of interoperability issues. A category of tools is represented by advanced version control systems, which are not only used to manage the code-base, but for many other purposes such as: assigning tasks; trace responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.

Startups can further improve development speed by making \textit{use of standards and known technologies} which are widely recognized, well tested, and supported by strong communities. Moreover, the use of widely adopted standards and frameworks reduces the need of a formal architectural design since most of implementation solutions are well documented and ready-to-use. This is confirmed by C1: \textit{``as long as you use Ruby standards with the Rails framework, the language is clean itself and doesn't need much documentation''.}

Other important factors that positively impact the speed of development are the team's desire to: \textit{create disruptive technologies}; to \textit{demonstrate personal abilities}; and to \textit{have the product used in the market}. As reported by practitioners, these factors are essential to enhance the morale of developers and therefore to achieve higher team performances. By contrast, especially in the typical sprint-based environments of Agile, when a developer is not \textit{able to meet deadlines} the morale goes down, hindering the development speed.

Finally the constant pressure under which the company regularly operates, leads the team to often \textit{work overtime to meet deadlines}. But as reported by practitioners, such way of working can be an effective strategy only in the short term since can lead to poorly maintainable code and developers burnouts in the long run.
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Accumulated technical debt}
\label{res:gsm:cat5}
Startups achieve a very high development speed by radically ignoring aspects related to documentation, structures and processes. However, although these aspects are not considered important in the very first stages, they will become increasingly more important for the productivity in the long-term, as we illustrate in CAT6.

Instead of traditional requirement engineering activities, startups make use of \textit{informal specification of functionalities} by means of ticked-based tools to manage low-precision lists of features to implement, written in form of self-explanatory user stories \ref{AgilePlan}. Practitioners intensively use physical tools such as post-it notes and whiteboards, which help in making functionalities visible and prioritizing stories based on personal experiences. C4 comments \textit{``[\ldots] it is the only way. Too many people make the mistake of sitting down and write big specs and then they build it for four months, realizing the product is not valuable only at the end.''}

Since startups are risky businesses by nature, even less attention is given to the traditional phase of \textit{analysis}, which is replaced by a \textit{rough and quick feasibility study}   \textit{``only sometimes and however informally''}, as stated by C5.  First of all it is generally \textit{hard to analyze risks with cutting-edge technologies}. To find out the feasibility of such cutting-edge projects, startups attempt a first implementation with rough and informal specifications, assuming that project's complexity will remain limited to a small number of functionalities, as discussed in CAT3. Additionally by \textit{keeping the product as simple as possible} and learning from competitors' solutions and mistakes, practitioners can use their \textit{past experiences in similar contexts to help assessing feasibility} of the project. Finally, to avoid restrictions on the flexibility of the team, potential limiting decisions are taken only when strictly necessary and anyhow as late as possible.

Another important factor which contributes to \textit{accumulate the technical debt} is the general \textit{lack of architectural design}, substituted by \textit{high-level mockups and low-precision diagrams}, \textit{using modularization and frameworks from day 1} and \textit{describing critical interactions with third-party components only}. In particular, the use of well-known standards, frameworks and conventions limits the need of formal UML\ref{UML} diagrams and documentation, and provide a minimum level of maintainability costs. Additionally, since quality aspects are not a main concern (see \textit{limited number of functionalities} and {efficiency emerges after using the product} in CAT3) having a well-structured architecture remains a secondary priority, and design is conducted only when strictly necessary.

A similar attitude towards verification and validation brings startups to a severe \textit{lack of automated testing}, which is mainly replaced by manual smoke tests executed by: trying the product internally; seeking for critical faults; and letting early adopters report non-critical bugs.  Paraphrasing C3, \textit{``Trying the product internally allows us to get rid of 50\% of bugs of important functionalities. Meanwhile users report bugs of secondary functionalities, eventually allowing us to mitigate the lack of testing. Indeed, staying one week in production enables us to identify 90\% of bugs''.} The lack of complete automated tests is partially also motivated by the fact that \textit{users are fault-tolerant in innovative beta products} and  by the limited number of functionalities, which allow the team to easily control critical bugs. However, in certain cases where components of the system might cause loss of data or severe damages to the product or users, engineers realize a reasonable level of automatic testing. In such cases, aided by modern automatic tools, they can quickly assess the status of the system integration as they add new functionalities to the product.

Furthermore, a rigid project management is perceived as \textit{``waste of time''} which hinders the development speed since the \textit{uncertainty makes formal scheduling pointless} (C9  reports that \textit{``initial chaos helps to develop faster''}). Startups' \textit{minimal project management}  is supported by keeping: \textit{internal milestones short and informal}, \textit{low-precision task assignment mechanism}  and a very low attention for project metrics (paraphrasing C13, \textit{``the only track of progress was made by looking at closed tickets''}). In this context \textit{only a final release milestone is viable}, which helps practitioners to remain focused on short term goals and put new features in production. To further \textit{speed-up development}, startups \textit{simplify issue management by integrating it with feature management}. Then, the \textit{limited number of functionalities} and the {use of standard/known technologies} with a \textit{simple workflow} are the main reasons for not establishing a heavy project management process since the project and technologies can be autonomously managed by the development team (with characteristics of: co-location, self-organization and very small size) .

Finally, one of the categories, which most contributes in growing the \textit{accumulated technical debt}, is the substantial  use of informal and verbal communication channel on daily basis. The high co-location and the fast paced development approach increase the volume of \textit{tacit knowledge} and severe lack of any kind of documentation.  C4 observes in this regard that:  \textit{``[\ldots] the issue of having documentation and diagrams out of the source code is that you need to update them every time you change something. There is no time for it. Instead, there is a huge pay off in having a code that is understandable itself.''}
This approach is supported by the fact that startups make use of  \textit{simple and informal workflow}, \textit{standard/known technologies}, \textit{very small and co-located development team} with \textit{limited need of formalities}.
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Initial growth hinders performance}
\label{res:gsm:cat6}
The lack of attention given in the first phases to engineering activities allows startups to ship code extremely quickly (see CAT5). However, if successful, the initial product becomes more and more complex overtime, the number of user increases and the company size starts to grow. Under these circumstances the need of controlling the initial chaos forces the development team to return the \textit{accumulated technical debt} instead of focusing on new users' requests. Hence, \textit{initial growth hinders performance} in terms of number of new user stories implemented in a certain amount of time, which brings new suitable functionalities to the users.

When the number of users increases, customers become more quality demanding and also scalability issues start to arise.

Usually, \textit{company and user size grow} when new business events occur, such as: a \textit{new round of funding}, a possible \textit{acquisition}, the release of a \textit{competing product on the market}, or when the project is \textit{open for the first public release}. As a consequence, the \textit{increasing number of users} creates growing number of requests and expectations. Therefore, whereas the project lacks of minimal processes, suddenly, \textit{the current team is not able to manage increased complexity} of new functionalities to implement and maintain the codebase.

Subsequently, practitioners start considering the need of project management activities, also in view of \textit{hiring new staff members}, as discussed by C13: \textit{``(project management) is strictly necessary if you radically change the team or when team grows.  The informal communication and lack of documentation slow down the process afterwards''}.  Project management becomes further important when the \textit{focus shifts to business concerns}. Part of the effort, which was initially almost entirely dedicated to product development, is required to move to business activities. Moreover the availability of project information becomes an important issue because of the accumulated \textit{tacit knowledge}, which hinders the ability of new hires to quickly start the development of project tasks.

When the company faces growth, the partial and informal engineering activities that have been conducted during the first phases of \textit{rush} development (refer to CAT5 subcategories \textit{minimal project management}, \textit{informal specification of functionalities}, \textit{rough and quick feasibility study}, \textit{lack of automated testing}, \textit{tacit knowledge}), force the company to \textit{pay off the accumulated technical debt}. Under these circumstances, startups need to be able to cope with the renewed needs and expectations of both the company (internal necessities) and customers (product).

Another factor that slows down performance is caused by the fact that \textit{portion of codes needs to be rewritten} and \textit{substantial refactoring of the codebase} is required by the increasing product's demand .

Practitioners realize that some decisions taken (or not taken) during the \textit{rough and quick feasibility study} before starting the implementation, have led to negative consequences on the long term performance and maintainability of the product. Additionally, the initial decisions of the product design might not be able to satisfy the increased demands of the product's users and developers (\textit{lack of architectural design}). The combination of these factors leads to the need of \textit{re-engineer the product}.

By re-engineering the systems, startups aim to \textit{increase the scalability of the product/infrastructure} and starting to \textit{standardize the codebase with well-known frameworks}.

C7 reports that: \textit{``To mitigate this (lack of frameworks) I had to make a schema for other developers when we hired them. We had to do a big refactoring of the codebase, moving it from custom php to Django, normalizing the model and making it stick with the business strategy. I had the code in different php server communicating via JSON, some engineering horror. Now  that we are fixing it, it's really painful. We had to trash some code. However I don't regret that I didn't make this choice sooner, it was the only way''} .

The \textit{fear of changing a product, which is working}, arises when product complexity increases. In fact, the changes to the codebase, to support bug fixing, become highly interrelated with other functionalities and difficult to manage because the product is poorly engineered (see CAT5). Therefore, the fear arises from changing a validated product that might bring changes to the user responses.

The increasing number of feature requests causes the \textit{growing necessity of having a release plan}, rationalizing the rollout of new features.

Therefore, startups begin to \textit{partially replace informal communication with traceable systems} and \textit{introduce basic metrics for measuring project and team progress} in order to establish an initial structured workflow. In conclusion, because the increasing number of users causes issues in scalability and reliability, the \textit{need of more structured workflow} becomes of major relevance. C11 states: \textit{``Yet, it is still better to have a reasonable drop-down in performance when team grows than lose time in the beginning''}.
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Theory}
\label{res:gsm:th}
In order to explain and understand the development strategies in early-stage software startups we construct the theory generated and supported by the GSM model presented above. As discussed in section \ref{valt}, we applied the \textit{paradigm model} to form the final theory, introduced by Corbin \cite{Corbin1990}, here presented:\\

Focusing on limited number of suitable functionalities, and adopting partial and rapid evolutionary development approaches, early-stage software startups operate at high development speed, aided by skilled and highly co-located developers. Through these development strategies, early-stage software startups aim to find early product/market fit within extremely uncertain conditions and severe lack of resources. Nevertheless, by speeding-up the development process, they accumulate technical debt, causing an initial and temporary drop-down in performance before setting off for further growth. \\

The theory above is composed by different elements reflecting the \textit{paradigm model}. Here we structure the list of elements presented in the theory:

\begin{compactitem}

\item \textit{Causal conditions} are represented by three main conceptual categories: \textit{product quality has low priority}, \textit{evolutionary approach} and \textit{ team is the catalyst of development}.
\item \textit{Phenomenon} is represented by the core category \textit{speed-up development}.
\item \textit{Context} is limited to web early-stage software startups operating in conditions of severe lack of resources aiming to early find product/market fit.
\item \textit{Intervening conditions} are  summarized by the extremely uncertain development environment.
\item \textit{Action and interaction strategies} are represented by the accumulation of technical debt (see CAT5).
\item \textit{Consequences} lead to a temporary performance drop-down (see CAT6).
\end{compactitem}
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\section{Theory implications}
\label{sect:theory:impl}

In this subsection we present a list of the most relevant implications which emerge from the behaviour of early-stage startups, formally expressed in the model above, answering to RQ-2 (\textit{What is the current state-of-practice related to software development strategy in early-stage startups?}) through its sub-questions. Although the startups we inquired in the case study were spread across different nations and market sectors, we have found several patterns that emerged very clearly:

\begin{compactitem}

\item The most urgent priority of software development is to shorten time-to-market. When developing a new product, which has not been attempted before, it is crucial for the company to release the product soon and observe the users reactions. The company's chances to survive, are strictly related to the ability to find the right product/market fit as quickly as possible. Developing software without releasing it for a long period of time, coming up with very complex solutions often is not a viable option. Business demands, investors interests, market pressures and the general uncertain conditions force startups to quickly iterate on the product by exploring and experimenting new functionalities progressively rolled-out to early-adopters.

\item Startups do not apply any standard development methodology. When focused on building the first version of the product, startups do not observe any specific and standard development methodologies or processes\footnote{Is worth noticing however that when asked the first question \textit{``Did you use any specific development methodology?''} 10 startups out of 13 referred to some kind of customized light version of SCRUM or XP (see Subsection \ref{sect:res:gt:meth-compare} for more details).}.  From an accurate and in-depth investigation into day-by-day activities, it emerged that the number of  the observed development \textit{practices} is extremely small and never systematic (only little \textit{pair programming} sessions and rarely \textit{test driven development}). Despite the wide number of publications, which discuss the adoption of lightweight methodologies, we have found that especially in this first stage not only methodologies but even individual practices are basically neglected.

\item The closest development approach undertaken by early-stage startups is the \textit{Lean startup methodology}.  A highly evolutionary development approach, centered around the quick production of a functioning prototype and guided by customer feedbacks is a central aspect of the \textit{Lean startup methodology} pioneered by Ries \cite{Ries2011}. However, startups do not formally follow the cycle \textit{build-measure-learn} as it is proposed by the methodology, where they should set-up an experiment to \textit{test the riskiest element of the business idea}, nor make explicit use of the good practices suggested in the book. For example, when it comes to user feedbacks, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}. Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its subcategories.

\item The greatest part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots). Especially when bringing the first product to the market, the attitude of startups towards traditional engineering activities is strongly characterized by a general feeling of unnecessity. Every document, artifact or process, which does not produce a short-term tangible benefit to the software production is perceived as a waste of time. The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (Subsection \ref{sect:theory:debt}), which is partially mitigated by different elements supporting the development. Answering RQ-2.1, project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.  Thus, by radically ignoring activities related to documentation, structures and processes, startups achieve a very high initial development speed in the first stage, but on the other hand they accumulate a technical debt that needs to be at least partially fulfilled when the company and the product grow.

\item The first release of the product includes only a limited set of well suitable functionalities focused on user experience. The beta version is usually focused on a minimal and essential set of functionalities providing a reasonable user experience (UX). As confirmed by practitioners, the user must be able to accomplish a task on the product without major interruptions in the flow, which would cause frustration before being able to capture his attention. To realize UX in a short time frame, developers focus on a limited number of suitable functionalities, especially in the first version(s) of the product. Other product qualities are regarded as less important\footnote{Note that from a business perspective, one might argue that established companies might take advantage of the new idea and develop a better version of the product internally. Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}. Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process (see Appendix \ref{a:related:grey} for more details) while focusing effort on acquiring larger shares of a ``mass market'' sector.}: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important, even more than time-to-market \cite{Offutt2002}.}. Furthermore, thanks to cloud computing and more in general high availability of third party components, aspects such as performance, security and scalability can be achieved with a reasonable effort. Thus, given the general lack of quality concerns\footnote{The extent to which quality aspects are considered by the company, is strongly influenced by the market sector and the application type, discussed as confounding factors. However our case study has been carried out in startups working on innovative web application with very few quality concerns.}, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.

\item Engineering activities are supported by low-precision artifacts and collaborative tools integrated with the development environment. Startups take advantage of the simplicity of cutting-edge tools, which require a minimum learning and maintenance effort and provide a large payoff.  Tools are used for a vast number of tasks: ticked-based boards to manage user stories; systems to collect user feedback; tools for tracing product metrics; and scripts for managing deployments. The most prominent families of tools, which support software development in startups are modern version control systems. They are used by practitioners not only to manage the code base and the repositories, but for many other purposes such as: task assignment; tracing responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.  On the other hand, an excessive usage of  \textit{out-of-the-code} tools can negatively affect the development speed, as significantly reported by C11: \textit{``We realized we were losing too much time moving tickets around boards than actually working on stories. You need to find the balance between the time spent with tools and the amount of actual coding''}. This risk is often mitigated with an intense usage of offline manual tools: especially post-it notes and whiteboards are reported to be particularly effective to foster internal communication within the team.

\item The founding team is the most important determinant of speed. More than techniques and methodologies, the features of the team are the essential factors to achieve high-development speed. Discussed in the theoretical category the \textit{team is the catalyst of development}, this result has been confirmed in multiple occasions both in the interviews and in the questionnaires. What is truly important for the startup, among other factors, is the extremely high-colocation which enables continuous informal and colloquial coordination and discussions. The initial (small) development team usually spends most of the hours of the week working closely to implement the product, often extending the conventional office hours.  The large amount of communication, spontaneously held in informal places, has been proven to be an extremely effective element of several successful software projects \cite{Highsmith2000}. Hence, the vastly \textit{informal} development environment of startups acts as a catalyst of speed. Other important elements, which contribute in achieving such an initial high-speed, are: minimal set of functionalities with low concerns on product quality; evolutionary prototyping approach; smart use of advanced tools and frameworks; low precision artifacts; and the close colocation and capability of the team.

\item The initial lack of structures and processes negatively affects the performance when the company starts to grow. One of the primary objectives of startups is to expand their business into fast growing markets. In fact, if successful, the company will face a growth in terms of number of customers, employees and product functionalities. Then, the increased complexity arises the necessity of controlling the initially chaotic software development environment. Under delicate circumstances, where an increased market demand meets a poorly engineered product and process,  the development team need to start returning the accumulated technical debt. As reported by practitioners, introducing more structures, documentation and standard procedures, cause an initial drop-down in performance which severity depends on the speed of growth, the amount of technical debt and the capacity of the team to absorb it.

\item Startups bring the first product to market in a very short-time and practitioners are satisfied with the software development strategy. The companies inquired in the case study were typically able to release the first beta of the product in a short time frame. Although practitioners acknowledge the initial negative consequences of the technical debt on subsequent performances, they confirm that it is better spending some effort in later stage to recover the debt than losing time in over-engineering the production initially with the risk of never achieving any significant growth. 
\end{compactitem}

All of the above observations are extracted from the analysis of the model, which is systematically derived from empirical data of the case study. Indeed, with slightly different levels of adherence, the implication we presented are reflected in the behaviour of the great majority of companies in the sample. The results of this analysis, which has been focused on the software development strategy from the idea conception to the first release, indicate that startups are far from adopting standard or ad-hoc development methodologies, especially in this phase. The typical tendency is to focus on the team capability to implement and quickly iterate on a prototype, which is released very fast. Thus, in a context where even for the most \textit{lightweight} agile methodologies is hard to penetrate, we believe that it is premature to discuss issues related to SPI. Researchers should rather focus on the trade-off between development speed and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which appears to be the most important determinant for the future of software development in the company.

%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
\section{Discussion}
\label{res:val}

In this section we discuss the validity of the implications of this study, obtained by means of cross-methodological observations:

\begin{compactitem}
\item We perform a comparison between our framework and similar  models and frameworks.

\item We validate the correctness of the GT procedure by attesting that high-level relations between categories, evaluating if they are compliant with the empirical data we collected.

\item Finally we demonstrate how engineering elements independently reported in the follow-up questionnaire by practitioners, can be mapped to theoretical categories in the model.
\end{compactitem}

Additional validation procedures are presented in \cite{SMS}\footnote{The author performs a validation of the theoretical framework by triangulating the results of different methodologies with the existing literature. He highlights the areas which have been neglected by existing studies, providing possible directions for future studies. Finally he identifies and discusses confounding factors which could interfere with the theoretical model.}.
%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
\subsection{Comparison with other frameworks}
\label{sect:theory:validation:others}

To validate the generalization of the theory, in this section we describe conceptualizations defined from the framework that are in support of a previous model developed by Coleman in \cite{Coleman2007,Coleman2008a, Coleman2008}. We refer to Coleman's framework since it is the only author which has conducted a similar research in startups' context, even though with a different focus. The author investigated factors in software development that hinder initiatives of software process improvement (SPI) in much more mature companies.

In this subsection we illustrate the most important aspects considered by Coleman, examining similar and contrast factors in face of our \textit{theoretical framework}. We will consider only those categories strictly related to the core category, which relations, as described in GT methodology, generate complete explanatory power of the generated theory.

Coleman's framework aims to highlight how managers consider two distinct kind of processes: \textit{essentials} and \textit{non-essentials}. The \textit{essential} processes are the most closely linked to the product development such as requirements gathering, design and testing. The \textit{non-essential} processes are those that might be omitted such as planning, estimating and staging meetings. In particular, he discusses how practices are routinely removed: \textit{``With most methodologies and approaches, very few stick to the letter of them and they are always adapted, so we adapted ours to the way we wanted it to work for us, for our own size and scale''.}

Our theoretical framework explores the same challenges in software startups, where the act of tailoring processes leads developers to adopt only minimal practices, which are most suitable for the startups context. Furthermore, CTOs and CEOs' background has a great impact on the speed of the development process as described in Subsection \ref{sect:theory:team}. This is in accordance to Coleman's framework, which presents the background of founders and development managers as main factors that affect the management style of the development process.

Differently from Coleman's, our theoretical framework doesn't present how \textit{market requirements} can affect the conduction of processes. In this regard, Coleman describes how the more the definition of requirements are predictable the more well-defined workflows are established. This particular aspect is considered as a confounding factor\footnote{The  identification of confounding depends on the context in which the study is conducted and on the background knowledge of the researchers \cite{Pearl2011}.}  in our study since we were not able to ground this concept in the data.

But what characterizes Coleman's network is the \textit{cost of process} (core category) and all the factors that in management contributed to the lack of software process improvements (SPI).

The \textit{cost of process} represents the  lack of formal and prescriptive workflows in development, mainly conducted by means of verbal \textit{communication} limiting heavy \textit{documentation} and \textit{bureaucracy}.

The author reports practitioners' perception that \textit{documentation} alone would not have ensured a complete shared understanding of project requirements. Moreover, defined processes are perceived by managers as additional items with a negative impact on the \textit{Creativity} and \textit{Flexibility} of the development team. This is in accordance with the generated theory (see Subsection \ref{sect:theory:theory}), which bases the reasons of adopting evolutionary and low-precision engineering elements to the flexibility and reactivity attributes of the development process.

At this regard we can paraphrase a comment recorded by Coleman during an interview: \textit{``When we set up we had more supervisory and managerial roles in that group than we have now and we had to scale that back which has made things a lot more flexible. I do think you have to be nimble, quick and capable being responsive in our position. That works well and I don't want to lose it.''} \cite{Coleman2007}.

Verbal \textit{communication}, lack of heavy \textit{documentation} and \textit{bureaucracy} can easily be associated to the \textit{accumulated technical debt} category (see CAT5). According to Coleman's framework, they describes how startups experience the lack of main engineering activities and documentation. Additionally, the \textit{speed-up development} category expresses the same concept of \textit{flexibility} and \textit{process erosion} since they have direct relation to the subcategory of \textit{keep a simple and informal workflow}, as discussed in CAT1.

As also reported in our framework, the definition of a \textit{``minimum process''} is not a matter of poor knowledge and training, but rather it is the necessity of
 operating with solutions that let the company move faster. \textit{``One-size-fits-all''} solutions have always found difficult to penetrate small software organizations \cite{Staples2007}. In fact, also when startups began establishing any SPI process they experienced \textit{process erosions}, which resolved to a barely sufficient workflow to satisfy the organizational business needs.

Then, software startups privilege the use of agile principles in support of \textit{creativity} and \textit{flexibility} instead of SPI methods, whereby processes need to be predictable and repeatable. Nimble and ad-hoc solutions prevent the use of heavy bureaucracy and fast communication strategies, even though the accumulated tacit knowledge is hard to manage and to transfer to new hires as also discussed by the category \textit{CAT6}.

Also in \cite{Coleman2008}, the author describes how the \textit{``management approach''} is oriented towards \textit{``embrace and empower''} solutions in contrast with \textit{``command and control''} style, where there is an evidence of trust in development staff to carry out tasks with less direct supervision, greater delegation of responsibilities and a more generally consensual environment. Nevertheless, software development manager and founders have still impact on management style and indirectly on software development process. In the case of early-stage startups, founders are mainly software development managers as CEOs/CTOs and  technical practitioners at the same time. As Coleman identified the influence of the founders' and managers' background over the software development process, also GSM identifies how the  CEOs and CTOs background shapes the high-level strategies adopted in developing the initial product. Notwithstanding, team members remain self-organized, able to intervene in all the aspects of the development process without any direct supervision, as discussed in Subsection \ref{sect:theory:team}.

In conclusion, by the comparison of the theoretical framework with Coleman's framework we have found that despite Coleman's study contributes to define which factors hinder the introduction of SPI, his results are coherent with the theory generated within our research. In this regard, a wide variability of processes is a key factor for startups. In contrast with some SPI processes, such as \textit{Six Sigma} which objective is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, startups seem to follow the \textit{Fischer's Fundamental Theorem of natural selection} \cite{Fisher}. In fact, moving towards flexible and variable processes (easy to adapt to the uncertain conditions), increase odds of ``natural selection'' only when some ``restrictive conditions'' are met. In other words, even though ignoring SPI practices is advisable, not following SPI does not predict major adaptiveness of startups development to the market. But, to imply a major evolutionary product/market fit, certain conditions explained in GSM need to be considered.

Also Baskerville in \cite{Internet} refers to SPI as a development approach typically effective only in large-scale, long-term development efforts with stable and disciplined processes. By applying a grounded theory analysis in 10 companies he found that major causal factors that influence development are:

\begin{compactenum}

\item A desperate rush-to-market.
\item A new and unique software market environment.
\item A lack of experience developing software under the conditions this environment imposed.

\end{compactenum}

As discussed in CAT7, 1) can be mapped to \textit{time shortage} and 3) to \textit{limited access to expertise}, whilst  2) can refer to {uncertain conditions make long-term planning not viable} explained in CAT2. Even though with different research focus and context of study - understand how and why Internet-speed software development (i.e. an agile method oriented using daily builds aimed at developing a product with high speed) differs from traditional software development - Baskerville revealed similar causal factors.

Baskerville argues that the dawn of the Internet era has intensified software development problems by emphasizing shorter cycle times, stating to:

\begin{compactitem}
\item Speed-up development by releasing more often the software and ``implanting'' customers in the development environment (similar to the \textit{find the product/market fit} concept presented in CAT2). 
\item Make heavy use of simple tools and existing components (similar to the concepts expressed in CAT1)
\item Invest time in facilitating development of scalable systems by the use of simple but stable architectural solutions (see subcategory \textit{using modularization and frameworks from day 1} discussed in CAT5). 
\item Tailor the development process daily according to the intense demands for speed. In this regard, he argues that the trend is to skip phases or tasks that might impede the ability to deliver software quickly even though producing lower quality software.

\end{compactitem}

With wider research focus, another study, developed in 1986 by Brooks \cite{BrooksJr1987}, discussed what challenges are involved in constructing software products. In his study the author  divides difficulties in development into essence (inherent to the nature of the software), and accidents (difficulties attending software production but which are not inherent). In other words essence concerns the hard part of building a software through activities such as specification, design, testing. Accidents refer to the labor of representing the software or testing its representation.

The author claims that the major effort applied by engineers in the last decades was dedicated towards accidents problems, trying to exploit new strategies to enhance software performance, reliability and simplicity of development, such as the introduction of high-level languages for programming.  Despite the great achievements in improving development performance, the \textit{essence} property of the software remained unaltered.

Essence difficulties are inherent to: complexity, conformity, changeability and invisibility. Complexity  mainly refers to interaction of software modules and elements between each others in some non-linear fashion, which increases accordingly to the project size. Conformity refers to the duty of software to adapt to human institutions and systems. Changeability  refers to the constant pressure of software products to accommodate culture, market, laws and hardware transitions. Invisibility  refers to the difficulty of representing software in space.

Here we present the basic attacks (or mitigation strategies) presented in 1986 by Brooks on the conceptual essence ascertained within the GSM model description:

\begin{compactitem}

\item Buy versus build-  the most radical possible solution for constructing software is not to construct it at all, taking advantage of what others have already implemented. It is the main strategy, which enables startups to externalize complexity to third party solutions explained within CAT1,  \textit{speed-up development}.
\item Requirements refinement and rapid prototyping- avoid deciding precisely what to build but rather iteratively extract and refine the product requirements with customers and users. It represents the evolutionary development approach to maintain a fast release-cycles during development (CAT2).
\item Incremental development- starting from simple solutions allows organizations to early prototype and control complexity overtime. It represents the main purpose of focusing on \textit{limited number of suitable functionalities} examined in \textit{product quality has low priority}  category (CAT3).
\item Great team- people are the center of a software project and it is profoundly important to empower and liberate their creative mind. It is the main objective of empowering developers' capabilities described in \textit{team is the catalyst of development}(CAT4).
\end{compactitem}

Brooks' forecast about software development strategies is revealed to be extremely accurate, according to the state-of-practice in modern startups.
%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
\subsection{High-level relations validity}
\label{sect:theory:validation:stats}

In this subsection we validate the most important causal relations presented in our model by attesting their correctness  with a quantification of empirical trends observed in the data. In fact we identified some trends and recurrent patterns which we tried to synthesize in this section by thoroughly analyzing the complete set of data obtained from the case study and the GSM model. We illustrate our observations aided by simple statistical methods, which provides to researchers an additional tool to validate our model supported by further companies' data.

As previously discussed when presenting the GSM model, the most essential feature that startups aim to achieve is high speed.

The \textit{core category} (\textit{speed-up development}) is influenced by: the degree of quality concerns (contributes to the \textit{efficiency}); the undertaken development approach (provides \textit{effectiveness}); and finally the intrinsic characteristics of the team (guarantee \textit{performance}).

Before starting the evaluation of the impact of the three categories on \textit{speed-up development}, we grounded in the interview transcripts the fact that those relations were characterized in terms of performances\footnote{Time-to-market was the main concern, as confirmed in \cite{SMS}.}:

\begin{compactitem}

\item The team enhanced the overall \textit{performance} of the software development. In fact, recalling that startups deal with severe lack of resources, the team members are the company's main assets. They develop software with no need of formalities and structures, left to the team's own capabilities to efficiently develop the product. Moreover, the effectiveness is granted by continuous advices given by mentors and by a management style oriented towards self-organization, full-stack, and multi-role settings\footnote{Note that the importance of a skilled team for development performance is expressed in many studies such as \cite{surviving-os-cockburn, Highsmith2000}.} (described in CAT4 and confirmed, among others, by Sutton in \cite{Sutton2000}).

\item The evolutionary approach enhanced the \textit{effectiveness} in implementing the right functionalities. In fact, this approach improves the company's capabilities to adjust the trajectory of product development (in agreement to the concept of \textit{pivoting} discussed in \cite{Ries2011}).
\item Since product quality has low priority, the development was focused only on implementing a limited number of suitable functionalities, enhancing its \textit{efficiency}. In fact, an essential prototype (MVP) does not require to comply with heavy quality constraints, enabling the team to quickly implement functionalities, ready to be validated by final users.
\end{compactitem}

An optimal combination of these three categories enables startups to ship code extremely quickly. However, the price that startups have to pay for achieving such high speed is related to the extent of the \textit{accumulated technical debt}, which will finally contribute to hinder performance in later stages. This is reflected by the right-hand side of the high-level framework, here extracted in Figure \ref{fig:an:comp:high-level2}.

Accordingly, by analysing the data of the 13 companies collected with the empirical study, we observed two distinct trends which further empower the above illustrated relations:

\begin{compactenum}
\item \textit{The more} a capable team\footnote{The team capability is evaluated according to the characteristics presented in CAT4.} with few quality constraints uses an evolutionary approach, \textit{the faster} the company releases the product.
\item \textit{The faster} the company releases the product, \textit{the larger} the degree of \textit{accumulated technical debt}\footnote{The debt here is considered regardless of context-specific debt mitigation factors and tactics.}.
\end{compactenum}

To assess these hypotheses we defined three measures reflecting the theoretical model:

\begin{compactitem}

\item \textit{Execution speed}: a metric that represents the development speed of the startup during different phases of the first release, computed by means of a weighted average speed for each phase. It was obtained by analysing interview transcripts looking at subcategories of \textit{Speed-up development}.

\item \textit{Technical debt}: a metric that represents the extent to which processes are controlled, structured, planned and documented by means of engineering artifacts and practices. It has been computed by means of a weighted average of the debt accumulated in each development phase, observing subcategories of  \textit{accumulated technical debt} with consequences on the startups growth.

\item \textit{Potential capability}: a metric that represents the degree to which each company reflected the capability of reaction and flexibility to the dynamic environment during the development process, given by the three categories that (theoretically) mostly affect \textit{speed-up development}.
\end{compactitem}

We quantified these measures for each company involved in our empirical study by defining score metrics based on a set of rubrics and evaluating startups accordingly, through an analysis of interview transcripts, codes and questionnaire results. The complete statistical procedure and provided  rubrics  to quantify the measures are illustrated and discussed in Appendix.

The validation has been performed on the most critical  causal relationships of the model. According to the paradigm model those are the causal condition which are linked to the core-category. There are four high-level relationships  which have not been validated using numerical methods. However their correctness has been verified with a cross-analysis of the extant literature and by attesting the causal relations among codes and categories, through axial and selective coding processes.

%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
\subsection{Engineering elements and categories}
\label{sect:theory:validation:elements}

Finally, to provide a concrete example of how the model can be used, we show that the engineering elements (reported in Appendix) mentioned by respondents during the interviews can be mapped into the theoretical categories of the framework.

According to practitioners, the most highly rated categories, which contribute to \textit{speed-up development}, are activities and practices incorporated under the \textit{find the product/market fit quickly} category of the GSM model, followed by the \textit{use of well-integrated and simple tools}, \textit{skilled developers} and \textit{multi-role and full-stack engineers}. It is important to note, among others, how elements and categories related to the team capabilities are regarded as very important by practitioners. This confirms the primary role of the team (CAT4) in contributing to the primary objective of startups (i.e. shortening time-to-market).

With this last step we further evaluated the framework's explanatory power by incorporating low-level concepts extracted from the questionnaire responses into abstract categories of the model.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\section{Validity threats}
\label{valt}
The first threat related to external validity might be caused by a possible wrong selection of subjects interviewed for the study. This threat profoundly affects GT, since it is a qualitative research method, which makes the use of semi-structured interviews, in fact, centres the research on respondent's opinions. In order to mitigate this threat we selected interviewees which covered positions of CTOs' and CEOs'. Their perspectives on what it is taking place within the organization, were the only meaningful data taken in consideration in the study.  In many cases there was no supporting evidence to verify the opinion expressed. Questionnaire dedicated to ``data triangulation'' was not of primary support because of the little time practitioners were able to grant to it. However, researchers have to accept the veracity of what respondents reported during the interviews \cite{Coleman2008a}. Notwithstanding the issues surrounding semi-structured interviews are vital. Within our study, although the reality might be different from what has been described, the reported data are CTOs' and CEOs' perception, which is used to base their decisions.

Finally the comparison with similar frameworks also helped in establishing the domain to which the study's findings can be generalized. In particular the previous framework developed by Coleman in \cite{Coleman2008} has allowed researchers to find similarities, differences and broader reasoning related to factors that hinder mature processes to be established in startups. Moreover a comparison has been conducted also with the notorious Brook's framework describing the \textit{basic attacks} to the challenges of developing software systems and other extant literature retrieved by \cite{SMS}\footnote{Note that the generalizability of results is applied only to those startups that operate in software-intensive activities.}. In conclusion, regarding the validation of the subcategories, a continuous re-evaluation of the interview transcripts has been performed.

To enhance the internal validity, we created a three-dimensional research framework to perform a triangulation validity procedure. By means of grounded theory approach formed by a \textit{systematic mapping study}, \textit{interviews} and  \textit{follow-up questionnaires} we searched for convergence among different sources of information to confirm our theories.

By improving and validating the theory definition, we conducted a final comparison of the emergent theory with: extant literature, previously developed models and evaluation of empirical data. With the final model validation we highlighted and examined similarities, contrasts and explanations as discussed in \cite{Eisenhardt2007}.  In this regard Eisenhardt stated: ``Tying the emergent theory to existing literature enhances the internal validity, generalizability, and theoretical level of the theory building from case study research [\ldots] because the findings often rest on a very limited number of cases.''

Little ambiguities about direction of causal influences have been reduced by introducing arrows to indicate the direction of the causal relation in the GSM model, both in the high level and low level frameworks. Furthermore, by means of comparison between the GSM model and extant studies in the broader fields, we revealed that the theory resembles in wider startup development context. Nevertheless, we discovered important confounding factors which might mine the conclusion and generalizability of the study related to innovation, creativity and market requirements.

The first important threat to the construct validity of the study is a possible inadequate description of constructs. In order to dismiss this risk, the entire study constructs have been adapted to the terminologies utilized by practitioners and defined at an adequate level for each theoretical conceptualization\footnote{For instance defining \textit{Time shortage}, the researchers explained it in terms of \textit{Investor pressure, CEO/business pressure, Demo presentations at events and internal final deadline} used by most of the interviewees during the study.}. Moreover, during the coding of interview transcripts the researchers adopted wide and explanatory descriptive labels for theoretical categories, to capture the underlying phenomenon without losing relevant details.

The second important threat is caused by the fact that interviewees might already be aware of the possible emergent theories analyzed by researchers. To reduce the risk that some hypothesis could have been involved in the design of the research, we let mainly participants drive the direction of the case study.

Another threat associated with the construct validity is the evaluation apprehension, which deals with people's rejection attitude towards the assessment of their behaviour. To overcome this risk we didn't attained to first answers from practitioners, but rather went towards details of engineering elements and activities. Moreover, we developed a rigorous data collection protocol to create case study databases and employ integration of multiple data collection methods. Reporting bias was mitigated by packaging all the needed material for conducting new researches in other contexts in order to criticize our findings (interview package with instructions has been made available online\footnote{Available at \url{https://github.com/adv0r/BTH-Interview-Package} .}. Moreover, two academic supervisors with expertise in the area have conducted a peer-review analysis of framework's constructs.

To control distortion during analysis we made extensive use of memos. To demonstrate that we were aware of personal biases, memos provided us another important function in controlling the quality of data analysis. Through the use of memos and comparative analysis we were able to check if data fitted into the emerging theory and also countering subjectivity that ultimately enhanced the likelihood of producing accurate research findings.

Grounded theory has been already applied by other researches in similar contexts to attest relationships among conceptualizations of an examined phenomenon (see \cite{Coleman2007, Basri, Coleman2008a}). Those relationships between sampling and analysis should be verified in such a way that emerging findings remain consistent as further data are collected. To attest this, we  explored and tested:

\begin{compactitem}
\item Each category and the strength of relations between them.
\item Hypotheses, derived from and related to the emergent theory.
\item Deviant cases to ensure robustness and general applicability.
\end{compactitem}

In particular we were prepared to modify generated categories so that the new data could be adapted into the emerging theory according to the concepts of: \textit{theoretical sampling} and \textit{theoretical saturation}.

According to the \textit{theoretical sampling} concept, we adjusted our research design and the emergent theory until only marginal results were generated. Moreover to enhance reliability of the outcome conceptualizations and relations, the process of coding interviews has been systematically conducted following a detailed process. Then, we asked to other practitioners a confirmation to further validate the generated framework and enhance the reliability of the relations among the identified categories.

An important issue is related to the fact that the limited number of interviews might not represent the complete scenarios in our context of study. Nevertheless, this issue has been partially mitigated as result of the theoretical saturation concept. This experience is attested by Martin's statement: \textit{``By the time three or four sets of data have been analysed, the majority of useful concepts will have been discovered.''} \cite{giardinoEtAl}. To obtain major certainty of the obtained results, what we applied was the final validation.

Despite we couldn't apply any power analysis procedures, in grounded theory the sample size depended on when data saturation occurred. Ramer in \cite{Ramer}, comparing quantitative to qualitative researches, states: \textit{``reaching data saturation, which involves obtaining data until no new information emerges, is critical for obtaining applicability in qualitative research''}. To avoid interrupting the research prematurely, after attesting that no more relevant information could be gathered from executing additional interviews, we iterated grounded theory study one more time, verifying that the explanatory power of the core category was fulfilled.

The  approach of generating theory allowed us to check emerging categories and their properties by gathering new evidence within an iterative and evolutionary approach, in view of the wide context of the research area that was under investigation.
%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
\subsection{Future work}
\label{futwork}

The GSM model enables practitioners in startup companies to speed-up development empowering the role of the team, focusing on minimal functionalities and adopting evolutionary development approaches. Researchers can use the GSM model to reproduce the study with other software startups, and focus on mitigation strategies to minimize the impact of the accumulated technical debt over the need of shortening time-to-market. Moreover, the  mutual collaborative dialog between the research in the development strategies in software startups and technical debt can benefits in understanding how both influence the growth of a company after the release of the first product to the market.

%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
%\subsection{Subsection Heading Here}
%Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

%\subsubsection{Subsubsection Heading Here}
%Subsubsection text here.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%

% Note that IEEE CS typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.
% However, the Computer Society has been known to put floats at the bottom.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE CS papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% However, Computer Society journals sometimes do use bottom floats - bear
% this in mind when choosing appropriate optional arguments for the
% figure/table environments.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Conclusion}
\label{conc}
Startups are able to produce cutting-edge software products with a wide impact on the market, significantly contributing to the global economy. Software development, especially in the early-stage, is at the core of the company's daily activities. Despite their severely high failure-rate, we have found that the quick proliferation of startups is not supported by an adequate and scientific body of knowledge. To be able to intervene on the software development strategy with scientific and engineering approaches, the first step is to understand startups' behavior. Hence, in this research work, we provided an initial explanation of the underlying phenomenon by means of a grounded theory case study, focusing on early engineering activities, from the idea conception to the first open beta release of the product.

Through an intense exploratory research, conducted with a systematic approach, we produced a theoretical model grounded into the hindsight knowledge collected among startup practitioners with the aim of explaining how development strategies are engineered by practitioners. The explanatory capability and correctness of the model have been validated by means of systematic comparisons with the state-of-the-art results and empirical data. The systematic mapping of the literature \cite{SMS}, with 37 studies extracted from an initial sample of 943 items, revealed a multi-faceted state-of-the-art inadequate to support software development activities in startup companies. On the other hand, the case study conducted in 13 early-stage startups, provided a broad set of empirical evidences obtained by combining different research methodologies in a grounded theory approach.

The overall results of our research confirm that startups possess unique characteristics of uncertainty, lack of resources and time-pressure. These factors influence the software development to an extent that transform every decision related to the development strategies into a difficult trade-off for the company. Moreover, although startups share different characteristics with other similar SE domains (e.g. market-driven development, small companies and web engineering), the unique combination of coexisting factors poses a whole new set of challenges which need to be addressed by primary studies. Especially when bringing the first product to market, startups' most urgent priority is releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early feedback and collected metrics. In this stage, startups often discard any formal project management, documentation, analysis, planning, testing and other traditional process activities. In fact, practitioners take advantage of an evolutionary prototyping approach, using well-integrated tools and externalizing complexity to third party solutions.

However, the initial gain obtained in terms of flexibility and speed is counterbalanced by the need of restructuring the product and controlling the engineering activities when the company starts to grow. In fact, if successful, the startup will face a growth in terms of number of customers, employees and product functionalities that leads to the necessity to control the initially chaotic software development environment. In this context the most significant challenge for early-stage startups is finding a sweet spot between being fast enough to enter the market early while controlling the amount of accumulated technical debt.  In fact, as confirmed by analyzing other framework in SE, the use of prescriptive and rigid structures over the development activities would hinder performance since startups operate across the border between chaotic and complex domain. Therefore, from a software development perspective, working with low-precision engineering activities represents the only viable strategy that startups can follow in the early stage of their lifecycle.

In this research we discussed a number of novel challenges for both practitioners and researchers, while presenting a first set of concepts, terms and activities which set the software engineering scene for the rapidly increasing startup phenomenon. By making a comparison with Berry's definition of SE \cite{Berry1992}, we would like to see the rise of a new discipline - \textit{startup engineering} - which can be defined as  \textit{the use of scientific, engineering, managerial and systematic approaches with the aim of successfully developing software systems in startup companies}.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% use section* for acknowledgement
\ifCLASSOPTIONcompsoc
  % The Computer Society usually uses the plural form
  \section*{Acknowledgments}
\else
  % regular IEEE prefers the singular form
  \section*{Acknowledgment}
\fi

The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
\bibliography{tse}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
%\begin{thebibliography}{1}

%\bibitem{IEEEhowto:kopka}
%This is an example of a book reference
%H. Kopka and P.W. Daly, \emph{A Guide to {\LaTeX}}, third ed. Harlow, U.K.: Addison-Wesley, 1999.

%This is an example of a Transactions article reference
%D.S. Coming and O.G. Staadt, "Velocity-Aligned Discrete Oriented Polytopes for Dynamic Collision Detection," IEEE Trans. Visualization and Computer Graphics, vol. 14,  no. 1,  pp. 1-12,  Jan/Feb  2008, doi:10.1109/TVCG.2007.70405.

%This is an example of a article from a conference proceeding
%H. Goto, Y. Hasegawa, and M. Tanaka, "Efficient Scheduling Focusing on the Duality of MPL Representation," Proc. IEEE Symp. Computational Intelligence in Scheduling (SCIS '07), pp. 57-64, Apr. 2007, doi:10.1109/SCIS.2007.367670.

%This is an example of a PrePrint reference
%J.M.P. Martinez, R.B. Llavori, M.J.A. Cabo, and T.B. Pedersen, "Integrating Data Warehouses with Web Data: A Survey," IEEE Trans. Knowledge and Data Eng., preprint, 21 Dec. 2007, doi:10.1109/TKDE.2007.190746.

%Again, see the IEEEtrans_HOWTO.pdf for several more bibliographical examples. Also, more style examples
%can be seen at http://www.computer.org/author/style/transref.htm
%\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{pictures/cg}}]{Michael Shell}

%\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2013-02-13 article/cleanGdoc.sh

@@ -1,0 +1,13 @@
+#!/bin/bash
+#! echo "Cleaning google doc shit!"
+#! cp cleanGdoc.jar preamble/cleanGdoc.jar
+#! cp cleanGdoc.jar chapters/cleanGdoc.jar
+#! cp cleanGdoc.jar appendix/cleanGdoc.jar
+
+#!echo "*** Cleaning preamble/ ..."
+#! cd preamble/
+java -jar cleanGdoc.jar
+#! cd ..
+
+echo "Done."
+

### 2013-02-13 article/main.tex
adjusted the complete methodology figure

@@
 In \textit{data collection} the empirical results are stored in a single \textit{case study database} and subsequently processed in \textit{data analysis} to form the theoretical categories.
-At each iteration the new emergent theory is updated following a formal procedure (\textit{Theory generation}), and after verifying that the \textit{theoretical saturation}\footnote{The point at which executing more interviews wouldn't bring any additional value for constructing the theory.} of categories has been achieved, we finally proceeded to \textit{theory validation} or performed another complete cycle.
+At each iteration the new emergent theory is updated following a formal procedure (\textit{Theory generation}), and after verifying that the \textit{theoretical saturation}\footnote{The point at which executing more interviews wouldn't bring any additional value for constructing the theory.} of categories has been achieved, we finally proceeded \textit{theory validation} or performed another complete cycle.
 
@@
 \label{res:gsm:cat2}
-Among different approaches to development, startups prefer to build an initial prototype and iteratively refine it over time, similarly to the concept of \textit{``evolutionary prototyping''} \ref{EvProt}.
+Among different approaches to development, startups prefer to build an initial prototype and iteratively refine it over time, similarly to the concept of \textit{``evolutionary prototyping''} \cite{EvProt}.
 The reason of using this approach is that startups want to validate the product against the market as soon as possible, finding the proper product/market fit.
@@
 
-Instead of traditional requirement engineering activities, startups make use of \textit{informal specification of functionalities} by means of ticked-based tools to manage low-precision lists of features to implement, written in form of self-explanatory user stories \ref{AgilePlan}.
+Instead of traditional requirement engineering activities, startups make use of \textit{informal specification of functionalities} by means of ticked-based tools to manage low-precision lists of features to implement, written in form of self-explanatory user stories \cite{AgilePlan}.
 Practitioners intensively use physical tools such as post-it notes and whiteboards, which help in making functionalities visible and prioritizing stories based on personal experiences.
@@
 Another important factor which contributes to \textit{accumulate the technical debt} is the general \textit{lack of architectural design}, substituted by \textit{high-level mockups and low-precision diagrams}, \textit{using modularization and frameworks from day 1} and \textit{describing critical interactions with third-party components only}.
-In particular, the use of well-known standards, frameworks and conventions limits the need of formal UML\ref{UML} diagrams and documentation, and provide a minimum level of maintainability costs.
+In particular, the use of well-known standards, frameworks and conventions limits the need of formal UML \cite{UML} diagrams and documentation, and provide a minimum level of maintainability costs.
 Additionally, since quality aspects are not a main concern (see \textit{limited number of functionalities} and {efficiency emerges after using the product} in CAT3) having a well-structured architecture remains a secondary priority, and design is conducted only when strictly necessary.
@@
 
-In this subsection we present a list of the most relevant implications which emerge from the behaviour of early-stage startups, formally expressed in the model above, answering to RQ-2 (\textit{What is the current state-of-practice related to software development strategy in early-stage startups?}) through its sub-questions.
+In this subsection we present a list of the most relevant implications which emerge from the behaviour of early-stage startups, formally expressed in the GSM model, answering to our research question (\textit{What is the current state-of-practice related to software development strategy in early-stage startups?}) through its sub-questions.
 Although the startups we inquired in the case study were spread across different nations and market sectors, we have found several patterns that emerged very clearly:
@@
 \item Startups do not apply any standard development methodology.
-When focused on building the first version of the product, startups do not observe any specific and standard development methodologies or processes\footnote{Is worth noticing however that when asked the first question \textit{``Did you use any specific development methodology?''} 10 startups out of 13 referred to some kind of customized light version of SCRUM or XP (see Subsection \ref{sect:res:gt:meth-compare} for more details).}.
-From an accurate and in-depth investigation into day-by-day activities, it emerged that the number of  the observed development \textit{practices} is extremely small and never systematic (only little \textit{pair programming} sessions and rarely \textit{test driven development}).
+When focused on building the first version of the product, startups do not observe any specific and standard development methodologies or processes.
+From an accurate and in-depth investigation into daily activities, it emerged that the number of  the observed development \textit{practices} is extremely small and never systematic (only little \textit{pair programming} sessions and rarely \textit{test driven development}).
 Despite the wide number of publications, which discuss the adoption of lightweight methodologies, we have found that especially in this first stage not only methodologies but even individual practices are basically neglected.
@@
 Every document, artifact or process, which does not produce a short-term tangible benefit to the software production is perceived as a waste of time.
-The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (Subsection \ref{sect:theory:debt}), which is partially mitigated by different elements supporting the development.
-Answering RQ-2.1, project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
+The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (CAT6), which is partially mitigated by different elements supporting the development.
+Project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
 Thus, by radically ignoring activities related to documentation, structures and processes, startups achieve a very high initial development speed in the first stage, but on the other hand they accumulate a technical debt that needs to be at least partially fulfilled when the company and the product grow.
@@
 Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}.
-Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process (see Appendix \ref{a:related:grey} for more details) while focusing effort on acquiring larger shares of a ``mass market'' sector.}: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important, even more than time-to-market \cite{Offutt2002}.}.
+Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process while focusing effort on acquiring larger shares of a ``mass market'' sector.}: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important, even more than time-to-market \cite{Offutt2002}.}.
 Furthermore, thanks to cloud computing and more in general high availability of third party components, aspects such as performance, security and scalability can be achieved with a reasonable effort.
-Thus, given the general lack of quality concerns\footnote{The extent to which quality aspects are considered by the company, is strongly influenced by the market sector and the application type, discussed as confounding factors.
-However our case study has been carried out in startups working on innovative web application with very few quality concerns.}, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
+Thus, given the general lack of quality concerns, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
 
@@
 
-Additional validation procedures are presented in \cite{SMS}\footnote{The author performs a validation of the theoretical framework by triangulating the results of different methodologies with the existing literature.
-He highlights the areas which have been neglected by existing studies, providing possible directions for future studies.
-Finally he identifies and discusses confounding factors which could interfere with the theoretical model.}.
 %--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
@@
 Our theoretical framework explores the same challenges in software startups, where the act of tailoring processes leads developers to adopt only minimal practices, which are most suitable for the startups context.
-Furthermore, CTOs and CEOs' background has a great impact on the speed of the development process as described in Subsection \ref{sect:theory:team}.
+Furthermore, CTOs and CEOs' background has a great impact on the speed of the development process as described in CAT4.
 This is in accordance to Coleman's framework, which presents the background of founders and development managers as main factors that affect the management style of the development process.
@@
 Moreover, defined processes are perceived by managers as additional items with a negative impact on the \textit{Creativity} and \textit{Flexibility} of the development team.
-This is in accordance with the generated theory (see Subsection \ref{sect:theory:theory}), which bases the reasons of adopting evolutionary and low-precision engineering elements to the flexibility and reactivity attributes of the development process.
+This is in accordance with the generated theory, which bases the reasons of adopting evolutionary and low-precision engineering elements to the flexibility and reactivity attributes of the development process.
 
@@
 As Coleman identified the influence of the founders' and managers' background over the software development process, also GSM identifies how the  CEOs and CTOs background shapes the high-level strategies adopted in developing the initial product.
-Notwithstanding, team members remain self-organized, able to intervene in all the aspects of the development process without any direct supervision, as discussed in Subsection \ref{sect:theory:team}.
+Notwithstanding, team members remain self-organized, able to intervene in all the aspects of the development process without any direct supervision, as discussed in CAT4.
 
@@
 However, the price that startups have to pay for achieving such high speed is related to the extent of the \textit{accumulated technical debt}, which will finally contribute to hinder performance in later stages.
-This is reflected by the right-hand side of the high-level framework, here extracted in Figure \ref{fig:an:comp:high-level2}.
 

### 2013-02-14 article/main.tex
added validation complete

@@
 \author{FirstName~SecondName,~\IEEEmembership{Fellow,~IEEE,}
-        BLA~BLA,~\IEEEmembership{Fellow,~IEEE,}
-        and~BLA~BLA,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
+        ..~..,~\IEEEmembership{Fellow,~IEEE,}
+        and~...~....,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
 \IEEEcompsocitemizethanks{\IEEEcompsocthanksitem Author info.\protect\\
@@
 A \textit{grounded theory} approach guided the execution of a case study to explore the state-of-practice.
-As result, the Greenfield startup model (GSM)  reveals the urgent priority of startups of releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early collected user feedbacks.
+As result, the Greenfield startup model (GSM)  reveals the urgent priority of startups of releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early collected user feedback.
 Nevertheless, the need of shortening time-to-market, by speeding-up the development through low-precision and product-centric engineering activities, is counterbalanced by the need of restructuring the product before setting off for further growth.
@@
 Nevertheless, by speeding-up the development process, startups accumulate technical debt, ignoring processes, relying on informal communication and replacing documentation with low-precision artifacts.
-For startups it is essential to look ahead within short-term deadlines and determine the right product/market fit driven by user feedbacks, using flexible and reactive development approaches and adopting mitigation strategies to contain the severity of accumulated technical debt.
+For startups it is essential to look ahead within short-term deadlines and determine the right product/market fit driven by user feedback, using flexible and reactive development approaches and adopting mitigation strategies to contain the severity of accumulated technical debt.
 
@@
 Section \ref{resmet} introduces the research questions and discusses the design and execution of the research process.
-Results are discussed in section \ref{res:gsm} which presents the GSM model.
+Results are discussed in section \ref{res:gsm}, which presents the GSM model.
 Section \ref{res:val}  compares results of this research to literature and frameworks.
@@
 
-Looking at the number of new business incubators which appeared in the last three years one can evaluate the importance of startups \cite{Grimaldi2005}.
+Looking at the number of new business incubators which appeared in the last decade one can evaluate the importance of startups \cite{Grimaldi2005}.
 The wave of disruption in new technologies has led companies to be more and more competitive, forcing themselves to radical organizational and innovational renewals,  which bring many companies to the attempt of behaving like startups \cite{Christensen1997}.
@@
 The implementation of methodologies to structure and control the development activities in startups is a major challenge for engineers \cite{Coleman2008}.
-Generally, the management of software development is achieved through the introduction of software processes, which defineswhat steps the development organizations should take at each stage of production and provide assistance in making estimates, developing plans and measuring quality \cite{Chroust1996}.
+Generally, the management of software development is achieved through the introduction of software processes, which define what steps the development organizations should take at each stage of production and provide assistance in making estimates, developing plans and measuring quality \cite{Chroust1996}.
 In the last decades, several models have been introduced to control software development activities.
@@
 In fact, startups have very limited resources and typically wish to use them to support product development instead of establishing processes \cite{Coleman2008, Heitlager2007}.
-Some attempts to taylor lightweight processes to startups, reported basic failure of their application : \textit{``Everyone is busy, and software engineering practices are often one of the first places developers cut corners''} \cite{Martin2007}.
+Some attempts to taylor lightweight processes to startups reported basic failure of their application: \textit{``Everyone is busy, and software engineering practices are often one of the first places developers cut corners''} \cite{Martin2007}.
 Rejecting the notion of repeatable and controlled processes, startups prominently take advantage of unpredictable, reactive and low-precision \cite{surviving-os-cockburn} engineering practices \cite{Sutton2000, Tanabian2005, Chorev2006, Kakati2003}.
@@
 Issues related to this domain are addressed in literature by the area known as \textit{market-driven software development} (sometimes called \textit{packaged software development} or \textit{COTS software development} \cite{Alves2006}).
-Among other results, researchers emphasize the importance of time-to-market as a key strategic objective \cite{dagMDR,dagMDR,sawyer99} for companies operating in this domain.
+Among other results, researchers emphasize the importance of time-to-market as a key strategic objective \cite{dagMDR,sawyer99} for companies operating in this domain.
 Other peculiar aspects which influence the software development are related to requirements, which are reported to be often ``invented by the software company'' \cite{512553}, ``rarely documented'' \cite{Karlsson02challengesin}, and can be validated only after the product is released to market \cite{dahl2003,Keil:1995}.
@@
 %--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
-\subsection{Lack of research in startups}
-Despite some studies tried to address the above mentioned issues, \cite{SMS} found only a few SE works in this specific area, as confirmed by other studies \cite{Coleman2008, Coleman2008a, Coleman2007, Sutton2000}.
-Moreover the studies, identified in our systematic review, appear to be highly fragmented and spread across different areas rather than constituting a consistent \textit{body of knowledge}.
+\subsection{General lack of research in startups}
+Despite some studies tried to address the above mentioned issues, a SMS \cite{SMS} revealed only a few SE works in this specific area, as confirmed by other studies \cite{Coleman2008, Coleman2008a, Coleman2007, Sutton2000}.
+Moreover, the studies, identified in \cite{SMS}, appear to be highly fragmented and spread across different areas rather than constituting a consistent \textit{body of knowledge}.
 
@@
 %--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
-\subsection{Definition of startup}
+\subsection{Software development in startups}
+
 The word \textit{startup} appeared in the SE literature for the first time in 1994 in an article written by Carmel \cite{Camel1994a} where he studied the \textit{time-to-completion in young package firm}.
 He noticed how these companies were particularly innovative and successful, advocating for the need of more research investigating software development practices so as to replicate success and try to transfer it to other technology sectors.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
-\subsection{Evolution of the process}
 
@@
 
-Then, when startups are mature enough to support software process improvement (SPI), the solutions considered according to the state-of-the-art are oriented towards light-weight processes such as a design of development process based on XP, proposed by Zettel in \cite{Zettel2001}.
+%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+\subsection{Software process improvement in startups}
+
+When startups are mature enough to support software process improvement (SPI), the solutions considered according to the state-of-the-art are oriented towards light-weight processes such as a design of development process based on XP, proposed by Zettel in \cite{Zettel2001}.
 The process consists of a set of activities and artifacts (in addition to some important roles) defined in order to identify responsibilities and tools to utilize.
@@
 The process depicted in Figure \ref{fig:gt:completemethodology} is evolutionary and affects the design at each new iteration.
-In \textit{data collection} the empirical results are stored in a single \textit{case study database} and subsequently processed in \textit{data analysis} to form the theoretical categories.
-At each iteration the new emergent theory is updated following a formal procedure (\textit{Theory generation}), and after verifying that the \textit{theoretical saturation}\footnote{The point at which executing more interviews wouldn't bring any additional value for constructing the theory.} of categories has been achieved, we finally proceeded \textit{theory validation} or performed another complete cycle.
+In \textit{data collection} the empirical results are integrated and stored in a single \textit{case study database} and subsequently processed in \textit{data analysis} to form the theoretical categories.
+At each iteration the new emergent theory is updated following a formal procedure (\textit{Theory generation}), and after verifying that the \textit{theoretical saturation}\footnote{The point at which executing more interviews wouldn't bring any additional value for constructing the theory.} of categories has been achieved, we finally proceeded to the \textit{theory validation} or performed another complete cycle.
 
@@
 \item The closest development approach undertaken by early-stage startups is the \textit{Lean startup methodology}.
-A highly evolutionary development approach, centered around the quick production of a functioning prototype and guided by customer feedbacks is a central aspect of the \textit{Lean startup methodology} pioneered by Ries \cite{Ries2011}.
+A highly evolutionary development approach, centered around the quick production of a functioning prototype and guided by customer feedback is a central aspect of the \textit{Lean startup methodology} pioneered by Ries \cite{Ries2011}.
 However, startups do not formally follow the cycle \textit{build-measure-learn} as it is proposed by the methodology, where they should set-up an experiment to \textit{test the riskiest element of the business idea}, nor make explicit use of the good practices suggested in the book.
-For example, when it comes to user feedbacks, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}.
+For example, when it comes to user feedback, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}.
 Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its subcategories.
@@
 
-\item We validate the correctness of the GT procedure by attesting that high-level relations between categories, evaluating if they are compliant with the empirical data we collected.
-
-\item Finally we demonstrate how engineering elements independently reported in the follow-up questionnaire by practitioners, can be mapped to theoretical categories in the model.
+\item We validate the correctness of the GT procedure by attesting the relations with the GSM model's core category, evaluating if they are compliant with the empirical data we collected.
+
 \end{compactitem}
@@
 \item Great team- people are the center of a software project and it is profoundly important to empower and liberate their creative mind.
-It is the main objective of empowering developers' capabilities described in \textit{team is the catalyst of development}(CAT4).
+It is the main objective of empowering developers' capabilities described in \textit{team is the catalyst of development} (CAT4).
 \end{compactitem}
@@
 Brooks' forecast about software development strategies is revealed to be extremely accurate, according to the state-of-practice in modern startups.
+%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+\subsection{Theoretical categories and existing literature}
+\label{sect:theory:validation:sms}
+
+In this section we report the evaluation of our theory in contrast with main relevant contribution of studies retrieved within \cite{SMS}, discussing the categories of the GSM model.
+
+The comparison has been executed by identifying issues related to the theoretical categories in the retrieved relevant studies.
+We illustrate how the selected studies contributed to our framework validation as following: for each retrieved article we present the main results, mapping them to categories in accordance to the theoretical framework.
+Table \ref{tab:an:literature_comp} shows the main contributions of the 37 identified studies, where an 'X' is placed in correspondence to the theoretical category they considered.
+The table has been sorted by computing how many categories of our theoretical frameworks have been considered by the study (column \textit{Count}).
+Therefore in the highest part of the table there are the studies which somehow discuss the higher number of concepts identified in the early-stage startups framework.
+
+\begin{table*}[!t]
+\renewcommand{\arraystretch}{1.3}
+\caption{SMS \cite{SMS} and GSM} 
+\label{tab:an:literature_comp}
+\centering
+\begin{tabular}{|l||c||c||c||c||c||c||c||c||c|}
+
+    \hline
+    Author (year) & CAT1  & CAT2  & CAT3  & CAT4  & CAT5  & CAT6  & CAT7  & Count & Ref. \\
+\hline
+Sutton (2000) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Sutton2000} \\
+Kajko (2008) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Kajko-Mattsson2008} \\
+Crowne (2002) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Crowne2002} \\
+Coleman (2008) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Coleman2008a} \\
+Coleman (2008) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Coleman2008} \\
+Coleman (2007) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Coleman2007} \\
+Camel (1994) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Camel1994a} \\
+Yoffie (1999) & X     & X     & X     & X     & X     & X     &       & 6     & \cite{Yoffie1999} \\
+Zettel (2001) & X     & X     & X     & X     &       &       & X     & 5     & \cite{Zettel2001} \\
+Jansen (2008) & X     &       & X     &       & X     & X     & X     & 5     & \cite{Jansen2008} \\
+Heitlager (2007) &       &       & X     & X     & X     & X     & X     & 5     & \cite{Heitlager2007} \\
+Deias  (2002) & X     & X     &       & X     & X     & X     &       & 5     & \cite{Deias} \\
+Ambler (2002) & X     & X     &       &       & X     & X     & X     & 5     & \cite{Ambler2002} \\
+Wood (2005) & X     &       & X     & X     &       &       & X     & 4     & \cite{Wood2005} \\
+Tingling (2007) &       & X     &       & X     & X     & X     &       & 4     & \cite{Tingling2007} \\
+Taipale (2010) & X     & X     &       & X     & X     &       &       & 4     & \cite{Taipale2010} \\
+Silva (2005) & X     & X     &       & X     & X     &       &       & 4     & \cite{Silva2005} \\
+Mirel (2000) & X     &       & X     & X     & X     &       &       & 4     & \cite{Mirel2000} \\
+Midler (2008) & X     & X     &       & X     &       &       & X     & 4     & \cite{Midler2008} \\
+Tanabian (2005) & X     &       &       & X     &       &       & X     & 3     & \cite{Tanabian2005} \\
+Stanfill (2007) & X     &       &       &       & X     &       & X     & 3     & \cite{Stanfill2007} \\
+Mater (2000) & X     & X     &       &       &       & X     &       & 3     & \cite{Mater2000} \\
+Kuvinka (2011) & X     & X     &       & X     &       &       &       & 3     & \cite{Kuvinka2011} \\
+Deakins (2005) & X     & X     &       & X     &       &       &       & 3     & \cite{Deakins2005} \\
+Yogendra (2002) & X     & X     &       &       &       &       &       & 2     & \cite{Yogendra2002} \\
+Wall (2001) & X     &       &       &       &       &       & X     & 2     & \cite{Wall2001} \\
+Su-Chuang (2007) & X     &       &       & X     &       &       &       & 2     & \cite{Su-Chan2007} \\
+Steenhuis (2008) &       &       &       & X     &       &       & X     & 2     & \cite{Steenhuis2008} \\
+Sau-ling Lai (2010) &       & X     &       &       &       &       & X     & 2     & \cite{Lai2010} \\
+Kakati (2003) &       & X     &       & X     &       &       &       & 2     & \cite{Kakati2003} \\
+Himola (2003) & X     &       &       &       &       &       & X     & 2     & \cite{Hilmola2003} \\
+H\"{a}sel (2010) & X     &       &       & X     &       &       &       & 2     & \cite{Hasel2010} \\
+Hanna (2010) & X     &       &       &       &       &       & X     & 2     & \cite{Hanna2010} \\
+Bean (2005) & X     &       &       &       & X     &       &       & 2     & \cite{Bean2005} \\
+Kim (2005) &       & X     &       &       &       &       &       & 1     & \cite{Kim2005} \\
+Fayad (1997) &       &       &       & X     &       &       &       & 1     & \cite{Fayad1997} \\
+Chorev (2006) &       &       &       & X     &       &       &       & 1     & \cite{Chorev2006} \\
+
+\hline
+\end{tabular}
+\end{table*}
+
+According to our framework, only 7 studies, out of 37 (18.91\%), discuss aspects related to the overall software development process in early stage startups.
+The remaining part of the studies is only partially covering the phenomenon described by the framework.
+Looking at the results, we were able to extrapolate information which confirmed the GSM categories.
+
+We can see how the majority of the retrieved studies (29) mention issues related to \textit{speed-up development} which confirms, in fact, the importance of the \textit{core} category.
+On the other hand we can observe that only a limited number of studies mentions results related to CAT3, CAT5 and CAT6.
+This suggests the direction of possible future primary studies towards: the technical debt and its future consequences on performance; aspects related to functionality and quality of the first version of the product.
+
+These results confirm the relevance of development teams as widely discussed in the SE literature, here mapping 26 studies to CAT4 (\textit{team catalyst for development}).
+The importance of people has been widely discussed in other studies in SE advocating for the need of  \textit{empowering people}, which is the element with the greatest impact on software development.
+Furthermore, being humans \textit{non-linear variables}, they are unlikely to follow repeatable prescriptive methodologies.
+As stated by Highsmith, \textit{``A drawback of each and every methodology is to expect people to behave consistently over time, when is clearly not like that.''} \cite{Highsmith2000}.
+For more detailed discussion about the importance of people the reader is referred to some prominent authors: Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999}, Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph \cite{Adolph2011} and Cockburn \cite{people-first-order}.
+Especially in early-stage startups, where the whole company overlaps with the development team, factors related to people are without any doubt the most important ones (confirmed with the empirical data of  our case study.
+
+In addition to the theoretical framework proposed by Coleman previously discussed, in the next part of this subsection we present similarities and contrasts with other relevant studies selected with the SMS.
+
+Starting from the concept of \textit{severe lack of resources} many authors describe startups facing with inescapable constraints defined by engineering/business concerns and constant time-pressure \cite{Camel1994a, Cugola98softwareprocesses:}.
+Because of their young and inexperienced working conditions, startups present limited resources in terms of management and strategic alliances \cite{Sutton2000}.
+
+Crowne in \cite{Crowne2002} identifies four main stages of a startup's lifecycle: startup, stabilization, growth and maturity.
+Inexperienced practitioners operating in a chaotic environment characterizes the first phase, where there is no strategic plan for developing the product.
+But, moving towards the \textit{stabilization} phase, the product starts to be unreliable and requirements unmanageable.
+Consequently, when startups move to the \textit{growth} phase, they initiate structuring and controlling processes, which are incrementally integrated in the development environment.
+The \textit{maturity} represents the moment when the company is ready to introduce process improvement initiatives.
+This study profoundly corresponds to different conceptualization of the theoretical framework.
+Starting from the minimal and low-precision engineering elements, startups start to grow the accumulated technical debt which requires fast integration of more structured and standard processes overtime.
+Issues related to low-precision engineering practices during companies' growth are discussed in an experience paper by Ambler \cite{Ambler2002} where he reports the case studies of two growing internet startups approaching an IPO.
+Despite the two companies were facing business and team growth, attitudes against the introduction of processes to structure and model the system were evident.
+The need to re-architect and redevelop the codebase lead the researchers to suggest a tailored version of the RUP development framework.
+Nevertheless, even with the aid of CASE tools\footnote{CASE tools are a class of software that automates many of the activities involved in various engineering activities.
+See \url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, startups failed in practicing engineering activities in view of eccesive lack of flexibility.
+In fact, developers fast moved to simpler tools such as the whiteboard, where they were able to provide input and insights, question assumptions, share past experiences, and even talk about potential implementation strategies.
+In general the development remained highly iterative, where project teams started to have a little modeling and testing only when necessary.
+They didn't take the serial approach of completely developing and accepting the requirements model, analysis model and so on.
+This let them react swiftly and adapt rapidly according to changes into their highly competitive markets.
+The researchers report \textit{``modeling was streamlined''}, that is modeling what was needed and go straightforward to coding, avoiding analysis paralysis.
+
+The use of whiteboards in working environment is in support of the \textit{use of well-integrated and simple tools}.
+In both the startups studied by Ambler, documentation was minimal, finding the ``sweet spot'' to have just barely enough tools and artifacts to meet their needs and no more.
+Generally, the use of \textit{keep a simple and informal workflow} endorses the use of low-precision engineering practices which lead to speed-up development initially, but with negative effects on the accumulated technical debt of a minimal project management and consequently on the final company growth.
+
+Kajko-Mattsson et al.
+investigated a Swedish software startup company involved in mobile applications \cite{Kajko-Mattsson2008}, reporting the following issues: a lack of requirements gathering process; release cycles and their length were not defined; the releases and their scope were not planned; lack of control over the change requests; lack of process control in terms of  absence of any documentation to track the status and progress of the process; defective releases in terms of lack of testing; poor communication since the communication inside the company was informal and not documented at all.
+In the same study they designed a process improvement model, based on three main phases: \textit{Evaluate}, \textit{Plan}  and \textit{Change}.
+
+During the evaluation phase Kajko-Mattsson et al.
+tried to establish the development workflow status within the company, and the authors document a status of chaos where any initiatives of quantitative methodology assessment would have been a waste of time.
+Therefore, they applied a qualitative methodology approach to define a model of the release management process.
+In fact, they moved towards the plan phase where the researchers decided to develop a model to manage releases, introducing a process composed by: release scope preparation, release planning, release development, system testing, acceptance testing and release deployment.
+Despite the effort spent from the researcher to establish a development workflow, practices were not fully conducted even though some marginal benefits have been reported: planned major release from four to six months; planned minor releases from two to four weeks; encompassing urgent corrective releases from one to five days.
+Regarding the previously mentioned problems, only partial improvements have been achieved since the major obstacle was to motivate employees to change their habits.
+With perfect hindsight the researchers claim that understanding and adapting solutions by solving one problem at a time would have increased benefits.
+Then, despite the reported results, start establishing simple and informal workflow aided the organization to move from the stage of knowing nothing to knowing at least something about their development process.
+
+The integration of minimal project management during the development activities can foster the control over development productivity and control the \textit{accumulated technical debt}, but still with particular attention to the development workflow status within the company, designing specific and tailored solutions.
+
+In fact, prescribed development models such as SPI in startups are basically ignored by startups, as discussed in \cite{Zettel2001}.
+To overcome the need of structure and control to accelerate the development process, in his study, Zettel developed a \textit{``LIghtweight Process for E-business software development''} (LIPE).
+This lightweight development method integrates XP approach with ideas from the areas of software measurement for project control and process improvement as a compromise between ad-hoc and more rigorous approach.
+LIPE, as argued from the author, provides scalability of the development process with a certain degree of flexibility, omitting parts of prescribed practices.
+Nevertheless, we couldn't retrieve any empirical evaluation of the study.
+
+In \cite{Deakins2005}, the author designed a new spiral methodology evolved during fieldwork with a developer team that had limited resources managing innovative product in volatile e-commerce environments.
+He proposes fast evolutionary approach experimentation with flexible processes obtaining customer's feedback as soon as possible for continuous product improvement.
+Even in this case the model is compliant with the nature of a random, ad-hoc and visionary product development.
+In fact, evolutionary development approaches are typically more suitable  for web applications, as confirmed in \cite{Deshpande2001}.
+Then the practices are oriented towards rapid and iterative design with complete freedom to the developer team, enabled to operate anywhere within the organization, according also to the category \textit{team is the catalyst of development}.
+
+Carmel reports in \cite{Camel1994a} the need of flexible and rapid development solutions to shorten time-to-market, but even more important, the need of high-skilled team developers, which heavily impact on \textit{speeding-up development}.
+In fact, the author suggests that entrepreneurs need to look for a well formed, skilled core development team and not just a set of product ideas and features.
+Moreover the team must be empowered with full-stack and self-organization settings with the flexibility of minimal bureaucracy.
+A studied company stated: \textit{``we didn't need weeks or months of detailed modeling and documentation , but rather modeling the architecture a little and then either exploring strategies by providing users with code or simply starting work on the actual software itself''}.
+
+Eventually to focus on simple solutions, two other studies \cite{Wall2001, Bean2005} sugges to \textit{``reuse the wheel, instead of reinventing it''}, taking advantage of open source software whenever it is possible, building systems quickly and effectively as a result.
+The author conclude stating that even though the principles in leading-edge development processes are the same principles of yesteryear, the way in which the fundamentals are applied has changed from prescriptive to agile solutions with less concern to quality aspects, also confirmed in \cite{Mater2000}.
+
+Thus, by summarizing the comparison, we have found that most of the contributions of the studies are compliant with the behavioural model of the phenomenon we provided.
+
+%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+\subsection{Confounding factors from the literature}
+\label{sect:an:comp:cat-vs-literature:conf}
+
+The purpose of this subsection is to identify which confounding factors need to be taken into consideration when evaluating the theoretical framework.
+The confounding factors are those variables which and have not been considered in the model and might interfere with the theoretical framework positively or negatively \cite{ColinRobson2009}.
+Even though there are no systematic procedure to identify confounding factors, we report those that have been explicitly identified by the literature review, that are related to: \textit{creativity, innovation, market type and application type}.
+
+A prominent contribution investigating creativity and innovation in startups, is presented in \cite{Heitlager2007}.
+The study reports how development approaches oriented to the product in the early-stage life-cycle, called \textit{fluid} (see Figure \ref{fig:an:comp:cof:inn}), have main impact to innovative solutions initially.
+After a while the need of structuring and controlling the growing company size (along with \textit{transition} and \textit{specific} phases) arises the need of project management activities and software processes for long-term scalability issues.
+
+Based on the model presented in \cite{Heitlager2007}, it identifies the lifecycle stages of a startup (\textit{fluid}, \textit{transition} and \textit{specific}), corresponding for each stage a development approach (product-oriented and process-oriented) the level of innovation achieved.
+The study reports how product-oriented development, in contrast with forced and mature process imposition, gives degree of freedom to the development team, enhancing \textit{creativity} of developers and augmenting the \textit{innovation} capability of the company in the early-stage.
+
+The empirical model we presented has been obtained from the transcripts of interviews, where the theme of innovation only seldom emerged.
+However, if innovative and creative factors were considered in the model, the integration of results would have been straightforward, since the evolution of this model is totally compliant with our assumption of CAT5 and CAT6.
+From product-oriented development in the \textit{fluid} phase to the process-oriented development in the \textit{specific} phase, the \textit{innovation} confounding factor seems to behave initially the same as the \textit{speed-up development}.
+Nevertheless, we were unable to ground these evidences in the GT data even though they might present strict correlation with the choice of an early product-development orientation with low-precision engineering elements.
+
+Other two important factors, that are not grounded in our empirical data, are related to: \textit{a.} the requirements expected by a specific market sector (also called \textit{market requirements} in \cite{Coleman2007}); and  \textit{b.} the application type as reported in \cite{Sutton2000, Coleman2008, Coleman2008a}.
+The main impact of those confounding factors is related to the adoption of flexible and reactive solutions for the development process.
+In particular they refer to the necessity of fulfillment of quality concerns that goes beyond scalability and UX defined by the framework.
+Especially when requirements are rigidly imposed or the application domain is well-known, providing low-quality products to final users might determine the failure of a startup.
+
+For example security in critical applications such as safety-critical systems present quality concerns, which must be granted.
+But also minor risky application domains such as e-commerce systems have high-impact factors for obtaining the first customer reference, as discussed in \cite{Deakins2005,Kim2005,Silva2005}.
+This characterizes those applications, whose part of functionalities are already successfully implemented in other software systems or whose failure might report serious financing or social damages \cite{Bass2003}.Yet, limiting our context to early-stage software startups, which typically develop first cutting-edge product in beta versions, partially mitigates the high-quality demands of users.
+
+To understand how to obtain processes in accordance to a set of customer needs, correlating the proper levels of customers' satisfaction, we introduce the Kano model, described in \cite{Ullah2011} and extensively studied in product development.
+As shown in the Kano model, users preferences can be classified into four categories: \textit{basic}, \textit{excitement}, \textit{performance} and \textit{indifference} areas.
+
+The \textit{basic} area represents the space where customers are dissatisfied if those \textit{must-be} functionalities are not met.
+This is the case of similar functionalities already implemented in the current market with well-known solutions.
+The \textit{performance} area represents the space where customers' satisfaction is proportional to how many of functionalities are implemented.
+In the \textit{indifference} area the customer does not really care about the functionalities: the satisfaction remains invariant.
+This represents the case where startups have not (yet) achieved the right product/market fit.
+Finally, the \textit{excitement} area represents the space where customers will be more satisfied if functionalities are implemented, but not disappointed if they are not implemented.
+This is the case of extremely innovative product development.
+
+Well-known domains such as e-commerce applications or other \textit{``quality critical products''} belong to the \textit{basic} area, where if a quality concern such as security in bank transfer services is missing satisfaction of customers will be profoundly affected.
+But as discussed, in early-stage startups that typically develop first beta products, developers usually deal with cutting-edge technologies.
+Those innovative solutions, if in the right market, are likely to belong to  the \textit{excitement} area, where satisfaction can hardly be negatively affected by faulty products.
+In fact, in the theoretical framework (see Section \ref{sect:theory:detailed}), minimal functionalities and less concern about quality aspects have low-impact on customers' satisfaction, supporting startup's ability to mainly focus on \textit{speed-up development}.
+Moreover operating in the \textit{excitement area} is fully compliant with an evolutionary approach and minimal functionalities since from the first product released to the market, it is unlikely to lose users because of high-quality demand.
+
+Nevertheless, this aspect profoundly undermines the generalizability of the theoretical model according to the theory, in particular \textit{CAT3}.
+Therefore, future studies within wider startups' context are advisable to extrapolate insights from those situations where quality concerns cannot be postponed or externalized.
+
+Finally, also the experience of team-members is considered as an important confounding factor.
+Since only little evidence has been gathered during the interviews, ``lack of experience'' was not considered as a sub-category in \textit{severe lack of resources}.
+Notwithstanding, according to \cite{Crowne2002}, startups at the beginning rely on clever, but inexperienced developers.
+In fact, having team-members with deep experience is a ``double-edge sword'', as discussed in \cite{Yoffie1999}.
+From one hand experience might quickly provide structure and maturity to the development process, from the other it might cause huge challenges in managing human resources in view of the fact that self-confident overachievers will almost inevitably clash.
+Consequently the team management requires higher control and coordination activities that inevitably hinder the flexibility of the development environment at the beginning, that is essential in early-stage startups.
+Accordingly, also Coleman in \cite{Coleman2007,Coleman2008a,Coleman2008} states that the operating strategies towards \textit{``minimum processes''} in startups is not a matter of poor knowledge and training, but rather it is the necessity of operating with solutions that let the company move faster.
+
+Nevertheless, since the amount of the results did not allow ``the lack of experience'' to be in the theoretical framework (in view of the fact that not enough data were identified \footnote{A category is formed if the codes are representative of more than the 70\% of the analyzed transcripts.}), this confounding factor requires further investigation.
+
 %--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
@@
 
-Before starting the evaluation of the impact of the three categories on \textit{speed-up development}, we grounded in the interview transcripts the fact that those relations were characterized in terms of performances\footnote{Time-to-market was the main concern, as confirmed in \cite{SMS}.}:
+Before starting the evaluation of the impact of the three categories on \textit{speed-up development}, we grounded in the interview transcripts the fact that those relations were characterized in terms of performances since time-to-market was the main concern (as confirmed in \cite{SMS}):
 
@@
 
-\item \textit{Potential capability}: a metric that represents the degree to which each company reflected the capability of reaction and flexibility to the dynamic environment during the development process, given by the three categories that (theoretically) mostly affect \textit{speed-up development}.
+\item \textit{Potential capability}: a metric that represents the degree to which each company reflected the capability of reaction and flexibility to the dynamic environment during the development process, given by the three categories that mostly affect \textit{speed-up development}.
 \end{compactitem}
@@
 However their correctness has been verified with a cross-analysis of the extant literature and by attesting the causal relations among codes and categories, through axial and selective coding processes.
-
-%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
-\subsection{Engineering elements and categories}
-\label{sect:theory:validation:elements}
-
-Finally, to provide a concrete example of how the model can be used, we show that the engineering elements (reported in Appendix) mentioned by respondents during the interviews can be mapped into the theoretical categories of the framework.
-
-According to practitioners, the most highly rated categories, which contribute to \textit{speed-up development}, are activities and practices incorporated under the \textit{find the product/market fit quickly} category of the GSM model, followed by the \textit{use of well-integrated and simple tools}, \textit{skilled developers} and \textit{multi-role and full-stack engineers}.
-It is important to note, among others, how elements and categories related to the team capabilities are regarded as very important by practitioners.
-This confirms the primary role of the team (CAT4) in contributing to the primary objective of startups (i.e.
-shortening time-to-market).
-
-With this last step we further evaluated the framework's explanatory power by incorporating low-level concepts extracted from the questionnaire responses into abstract categories of the model.
 
@@
 
-Little ambiguities about direction of causal influences have been reduced by introducing arrows to indicate the direction of the causal relation in the GSM model, both in the high level and low level frameworks.
+Little ambiguities about direction of causal influences have been reduced by introducing arrows to indicate the direction of the causal relation in the GSM model.
 Furthermore, by means of comparison between the GSM model and extant studies in the broader fields, we revealed that the theory resembles in wider startup development context.

### 2013-02-19 article/main.tex
version 0.2 sent to michael

@@
 %\boldmath
-Software startups are newly created companies with no operating history and extremely fast in producing cutting-edge technologies.
+Software startups are newly created companies with no operating history and oriented in producing cutting-edge technologies.
+These companies develop software under highly uncertain conditions, tackling fast growing markets with severe lack of resources.
 Despite their increasing importance in economy there are only few scientific studies attempting to address software engineering (SE) issues, especially for early-stage startups.
@@
 
-With a cross-sectional study we focused the research context in a limited time-frame to highlight the nature of uncertainty in the development activities and time pressure from the market, as discussed in \cite{Maccormack2001, Eisenhardt1998}.
+With a cross-sectional study we focused the research context in a limited time-frame to emphasize the nature of uncertainty in the development activities and time pressure from the market, as discussed in \cite{Maccormack2001, Eisenhardt1998}.
 We performed semi-structured in-depth interviews with CEOs and CTOs of startups, covering a wide spectrum of thematics and iteratively adjusting the direction of the research according to the emerging evidences.
@@
 The most urgent priority of software development is to shorten time-to-market.
-The greater part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).The first release of the product includes only a limited set of well suitable functionalities.
+The greater part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).
+The first release of the product includes only a limited set of well suitable functionalities.
 
@@
 Section \ref{valt} discusses the validity threats considered during this study.
+Section \ref{sect:theory:impl} discusses the most relevant implications of the GSM model.
 Finally, this paper concludes in section \ref{conc}.
@@
 
-Modern entrepreneurship, which was born more than thirty years ago \cite{storey1982entrepreneurship}, has been boosted by the advent of the consumer internet markets in the middle of the nineties and culminated with the notorious \textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126}.
+Modern entrepreneurship has been boosted by the advent of the consumer internet markets in the middle of the nineties and culminated with the notorious \textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126, storey1982entrepreneurship}.
 Several years later, with the massification of the internet and mobile devices, we are now assisting to an impressive proliferation of software ventures - metaphorically referred as the \textit{startup bubble}.
@@
 
-Looking at the number of new business incubators which appeared in the last decade one can evaluate the importance of startups \cite{Grimaldi2005}.
+Looking at the number of new business incubators which appeared in the last decade one can estimate the importance of startups \cite{Grimaldi2005}.
 The wave of disruption in new technologies has led companies to be more and more competitive, forcing themselves to radical organizational and innovational renewals,  which bring many companies to the attempt of behaving like startups \cite{Christensen1997}.
@@
 \subsection{Technical debt}
-Notwithstanding, a new interesting area of the SE research, trying to tackle the problem of the \textit{technical debt}, seems to bring interesting implications in studying development in software startups.
+Notwithstanding, a new area of the SE research, trying to tackle the problem of the \textit{technical debt},  brings and encompasses different implications in studying development in software startups \cite{Tom2013}.
 The metaphoric neologism was originally introduced by Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the attention of SE researchers\footnote{To attest the fact that \textit{technical debt} is gaining traction among researchers, we mention two important contributions which characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt} organized by the Software Engineering Institute  and ICSE.}.
@@
 
-The aim of this research is to \textit{understand how software development strategies are engineered by practitioners in startup companies in terms of level of: structure, planning and control of software projects, in the period of time  that goes from idea conception to the first open beta release of the software product}.
-
-The above mentioned goal is structured according to the GQM paradigm \cite{Basili1992}.
+The aim of this research is to \textit{understand how software development strategies are engineered by practitioners in startup companies in terms of level of: structure, planning and control of software projects, in the period of time  that goes from idea conception to the first open beta release of the software product} \footnote{The above mentioned goal is structured according to the GQM paradigm \cite{Basili1992}.}.
 
@@
 
-By means of the SMS \cite{SMS} we refined our research questions.
+This case study started based on a Systematic Mapping Study (SMS) of the literature regarding software development in startups \cite{SMS}.
+By means of this SMS we refined our research questions.
 The research questions have been narrowed down to outline the domain of our study, even though still broad enough to allow emerging new theories and to adjust the scope of the research \cite{Corbin1990}.
@@
 By analyzing the interview transcripts with the GT process we extracted the information necessary to form the theoretical model.
-The initial coding process (open and in-vivo) provided 630 unique codes, grounded to 1295 citations in the transcripts, divided according to the thematic areas.
+The initial coding process (open and in-vivo) provided 630 unique codes \footnote{Available at \url{https://github.com/adv0r/BTH-Interview-Package}.}, grounded to 1295 citations in the transcripts, divided according to the thematic areas.
 
@@
 \item \textit{Intervening conditions} are  summarized by the extremely uncertain development environment.
-\item \textit{Action and interaction strategies} are represented by the accumulation of technical debt (see CAT5).
-\item \textit{Consequences} lead to a temporary performance drop-down (see CAT6).
+\item \textit{Action and interaction strategies} are represented by the accumulation of technical debt.
+\item \textit{Consequences} lead to a temporary performance drop-down.
 \end{compactitem}
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
-\section{Theory implications}
-\label{sect:theory:impl}
-
-In this subsection we present a list of the most relevant implications which emerge from the behaviour of early-stage startups, formally expressed in the GSM model, answering to our research question (\textit{What is the current state-of-practice related to software development strategy in early-stage startups?}) through its sub-questions.
-Although the startups we inquired in the case study were spread across different nations and market sectors, we have found several patterns that emerged very clearly:
+
+%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+\section{Validation}
+\label{res:val}
+
+In this section we discuss the validity of the GSM model by means of cross-methodological observations:
 
 \begin{compactitem}
-
-\item The most urgent priority of software development is to shorten time-to-market.
-When developing a new product, which has not been attempted before, it is crucial for the company to release the product soon and observe the users reactions.
-The company's chances to survive, are strictly related to the ability to find the right product/market fit as quickly as possible.
-Developing software without releasing it for a long period of time, coming up with very complex solutions often is not a viable option.
-Business demands, investors interests, market pressures and the general uncertain conditions force startups to quickly iterate on the product by exploring and experimenting new functionalities progressively rolled-out to early-adopters.
-
-\item Startups do not apply any standard development methodology.
-When focused on building the first version of the product, startups do not observe any specific and standard development methodologies or processes.
-From an accurate and in-depth investigation into daily activities, it emerged that the number of  the observed development \textit{practices} is extremely small and never systematic (only little \textit{pair programming} sessions and rarely \textit{test driven development}).
-Despite the wide number of publications, which discuss the adoption of lightweight methodologies, we have found that especially in this first stage not only methodologies but even individual practices are basically neglected.
-
-\item The closest development approach undertaken by early-stage startups is the \textit{Lean startup methodology}.
-A highly evolutionary development approach, centered around the quick production of a functioning prototype and guided by customer feedback is a central aspect of the \textit{Lean startup methodology} pioneered by Ries \cite{Ries2011}.
-However, startups do not formally follow the cycle \textit{build-measure-learn} as it is proposed by the methodology, where they should set-up an experiment to \textit{test the riskiest element of the business idea}, nor make explicit use of the good practices suggested in the book.
-For example, when it comes to user feedback, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}.
-Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its subcategories.
-
-\item The greatest part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).
-Especially when bringing the first product to the market, the attitude of startups towards traditional engineering activities is strongly characterized by a general feeling of unnecessity.
-Every document, artifact or process, which does not produce a short-term tangible benefit to the software production is perceived as a waste of time.
-The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (CAT6), which is partially mitigated by different elements supporting the development.
-Project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
-Thus, by radically ignoring activities related to documentation, structures and processes, startups achieve a very high initial development speed in the first stage, but on the other hand they accumulate a technical debt that needs to be at least partially fulfilled when the company and the product grow.
-
-\item The first release of the product includes only a limited set of well suitable functionalities focused on user experience.
-The beta version is usually focused on a minimal and essential set of functionalities providing a reasonable user experience (UX).
-As confirmed by practitioners, the user must be able to accomplish a task on the product without major interruptions in the flow, which would cause frustration before being able to capture his attention.
-To realize UX in a short time frame, developers focus on a limited number of suitable functionalities, especially in the first version(s) of the product.
-Other product qualities are regarded as less important\footnote{Note that from a business perspective, one might argue that established companies might take advantage of the new idea and develop a better version of the product internally.
-Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}.
-Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process while focusing effort on acquiring larger shares of a ``mass market'' sector.}: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important, even more than time-to-market \cite{Offutt2002}.}.
-Furthermore, thanks to cloud computing and more in general high availability of third party components, aspects such as performance, security and scalability can be achieved with a reasonable effort.
-Thus, given the general lack of quality concerns, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
-
-\item Engineering activities are supported by low-precision artifacts and collaborative tools integrated with the development environment.
-Startups take advantage of the simplicity of cutting-edge tools, which require a minimum learning and maintenance effort and provide a large payoff.
-Tools are used for a vast number of tasks: ticked-based boards to manage user stories; systems to collect user feedback; tools for tracing product metrics; and scripts for managing deployments.
-The most prominent families of tools, which support software development in startups are modern version control systems.
-They are used by practitioners not only to manage the code base and the repositories, but for many other purposes such as: task assignment; tracing responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.
-On the other hand, an excessive usage of  \textit{out-of-the-code} tools can negatively affect the development speed, as significantly reported by C11: \textit{``We realized we were losing too much time moving tickets around boards than actually working on stories.
-You need to find the balance between the time spent with tools and the amount of actual coding''}.
-This risk is often mitigated with an intense usage of offline manual tools: especially post-it notes and whiteboards are reported to be particularly effective to foster internal communication within the team.
-
-\item The founding team is the most important determinant of speed.
-More than techniques and methodologies, the features of the team are the essential factors to achieve high-development speed.
-Discussed in the theoretical category the \textit{team is the catalyst of development}, this result has been confirmed in multiple occasions both in the interviews and in the questionnaires.
-What is truly important for the startup, among other factors, is the extremely high-colocation which enables continuous informal and colloquial coordination and discussions.
-The initial (small) development team usually spends most of the hours of the week working closely to implement the product, often extending the conventional office hours.
-The large amount of communication, spontaneously held in informal places, has been proven to be an extremely effective element of several successful software projects \cite{Highsmith2000}.
-Hence, the vastly \textit{informal} development environment of startups acts as a catalyst of speed.
-Other important elements, which contribute in achieving such an initial high-speed, are: minimal set of functionalities with low concerns on product quality; evolutionary prototyping approach; smart use of advanced tools and frameworks; low precision artifacts; and the close colocation and capability of the team.
-
-\item The initial lack of structures and processes negatively affects the performance when the company starts to grow.
-One of the primary objectives of startups is to expand their business into fast growing markets.
-In fact, if successful, the company will face a growth in terms of number of customers, employees and product functionalities.
-Then, the increased complexity arises the necessity of controlling the initially chaotic software development environment.
-Under delicate circumstances, where an increased market demand meets a poorly engineered product and process,  the development team need to start returning the accumulated technical debt.
-As reported by practitioners, introducing more structures, documentation and standard procedures, cause an initial drop-down in performance which severity depends on the speed of growth, the amount of technical debt and the capacity of the team to absorb it.
-
-\item Startups bring the first product to market in a very short-time and practitioners are satisfied with the software development strategy.
-The companies inquired in the case study were typically able to release the first beta of the product in a short time frame.
-Although practitioners acknowledge the initial negative consequences of the technical debt on subsequent performances, they confirm that it is better spending some effort in later stage to recover the debt than losing time in over-engineering the production initially with the risk of never achieving any significant growth.
-\end{compactitem}
-
-All of the above observations are extracted from the analysis of the model, which is systematically derived from empirical data of the case study.
-Indeed, with slightly different levels of adherence, the implication we presented are reflected in the behaviour of the great majority of companies in the sample.
-The results of this analysis, which has been focused on the software development strategy from the idea conception to the first release, indicate that startups are far from adopting standard or ad-hoc development methodologies, especially in this phase.
-The typical tendency is to focus on the team capability to implement and quickly iterate on a prototype, which is released very fast.
-Thus, in a context where even for the most \textit{lightweight} agile methodologies is hard to penetrate, we believe that it is premature to discuss issues related to SPI.
-Researchers should rather focus on the trade-off between development speed and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which appears to be the most important determinant for the future of software development in the company.
-
-%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
-\section{Discussion}
-\label{res:val}
-
-In this section we discuss the validity of the implications of this study, obtained by means of cross-methodological observations:
-
-\begin{compactitem}
-\item We perform a comparison between our framework and similar  models and frameworks.
-
-\item We validate the correctness of the GT procedure by attesting the relations with the GSM model's core category, evaluating if they are compliant with the empirical data we collected.
+\item Comparison between the GSM model and similar  models and frameworks.
+
+\item Comparison between the GSM model and the extant literature. We highlight the areas which have been neglected by existing studies, providing possible directions for future studies.
+\item Discussion of confounding factors which could interfere with the GSM model.
+\item Validation of the correctness of the GT procedure by attesting relations between the GSM categories, evaluating if they are compliant with the empirical data we collected.
 
@@
 
-In this subsection we illustrate the most important aspects considered by Coleman, examining similar and contrast factors in face of our \textit{theoretical framework}.
+In this subsection we illustrate the most important aspects considered by Coleman, examining similar and contrast factors in face of the GSM model.
 We will consider only those categories strictly related to the core category, which relations, as described in GT methodology, generate complete explanatory power of the generated theory.
@@
 
-Our theoretical framework explores the same challenges in software startups, where the act of tailoring processes leads developers to adopt only minimal practices, which are most suitable for the startups context.
+The GSM model explores the same challenges in software startups, where the act of tailoring processes leads developers to adopt only minimal practices, which are most suitable for the startups context.
 Furthermore, CTOs and CEOs' background has a great impact on the speed of the development process as described in CAT4.
@@
 
-Differently from Coleman's, our theoretical framework doesn't present how \textit{market requirements} can affect the conduction of processes.
+Differently from Coleman's, the GSM model doesn't present how \textit{market requirements} can affect the conduction of processes.
 In this regard, Coleman describes how the more the definition of requirements are predictable the more well-defined workflows are established.
@@
 
-As also reported in our framework, the definition of a \textit{``minimum process''} is not a matter of poor knowledge and training, but rather it is the necessity of
+As also reported in the GSM model, the definition of a \textit{``minimum process''} is not a matter of poor knowledge and training, but rather it is the necessity of
  operating with solutions that let the company move faster.
@@
 
-In conclusion, by the comparison of the theoretical framework with Coleman's framework we have found that despite Coleman's study contributes to define which factors hinder the introduction of SPI, his results are coherent with the theory generated within our research.
+In conclusion, by the comparison of the GSM model with Coleman's framework we have found that despite Coleman's study contributes to define which factors hinder the introduction of SPI, his results are coherent with the theory generated within our research.
 In this regard, a wide variability of processes is a key factor for startups.
@@
 \item Speed-up development by releasing more often the software and ``implanting'' customers in the development environment (similar to the \textit{find the product/market fit} concept presented in CAT2).
-\item Make heavy use of simple tools and existing components (similar to the concepts expressed in CAT1)
+\item Make heavy use of simple tools and existing components (similar to the concepts expressed in CAT1).
 \item Invest time in facilitating development of scalable systems by the use of simple but stable architectural solutions (see subcategory \textit{using modularization and frameworks from day 1} discussed in CAT5).
@@
 
-In this section we report the evaluation of our theory in contrast with main relevant contribution of studies retrieved within \cite{SMS}, discussing the categories of the GSM model.
+In this section we report the evaluation of our theory in contrast with main relevant contribution of studies retrieved and discussed within the SMS in \cite{SMS}, discussing the categories of the GSM model.
 
 The comparison has been executed by identifying issues related to the theoretical categories in the retrieved relevant studies.
-We illustrate how the selected studies contributed to our framework validation as following: for each retrieved article we present the main results, mapping them to categories in accordance to the theoretical framework.
+We illustrate how the selected studies contributed to the GSM model validation as following: for each retrieved article we present the main results, mapping them to categories in accordance to the GSM model.
 Table \ref{tab:an:literature_comp} shows the main contributions of the 37 identified studies, where an 'X' is placed in correspondence to the theoretical category they considered.
-The table has been sorted by computing how many categories of our theoretical frameworks have been considered by the study (column \textit{Count}).
+The table has been sorted by computing how many categories of the GSM model have been considered by the study (column \textit{Count}).
 Therefore in the highest part of the table there are the studies which somehow discuss the higher number of concepts identified in the early-stage startups framework.
@@
 
-According to our framework, only 7 studies, out of 37 (18.91\%), discuss aspects related to the overall software development process in early stage startups.
+According to the GSM model, only 7 studies, out of 37 (18.91\%), discuss aspects related to the overall software development process in early stage startups.
 The remaining part of the studies is only partially covering the phenomenon described by the framework.
@@
 The \textit{maturity} represents the moment when the company is ready to introduce process improvement initiatives.
-This study profoundly corresponds to different conceptualization of the theoretical framework.
+This study profoundly corresponds to different conceptualization of the GSM model.
 Starting from the minimal and low-precision engineering elements, startups start to grow the accumulated technical debt which requires fast integration of more structured and standard processes overtime.
@@
 %--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
-\subsection{Confounding factors from the literature}
+\subsection{Confounding factors}
 \label{sect:an:comp:cat-vs-literature:conf}
 
-The purpose of this subsection is to identify which confounding factors need to be taken into consideration when evaluating the theoretical framework.
+The purpose of this subsection is to identify which confounding factors need to be taken into consideration when evaluating the GSM model.
 The confounding factors are those variables which and have not been considered in the model and might interfere with the theoretical framework positively or negatively \cite{ColinRobson2009}.
@@
 A prominent contribution investigating creativity and innovation in startups, is presented in \cite{Heitlager2007}.
-The study reports how development approaches oriented to the product in the early-stage life-cycle, called \textit{fluid} (see Figure \ref{fig:an:comp:cof:inn}), have main impact to innovative solutions initially.
+The study reports how development approaches oriented to the product in the early-stage life-cycle, called \textit{fluid} have main impact to innovative solutions initially.
 After a while the need of structuring and controlling the growing company size (along with \textit{transition} and \textit{specific} phases) arises the need of project management activities and software processes for long-term scalability issues.
@@
 Those innovative solutions, if in the right market, are likely to belong to  the \textit{excitement} area, where satisfaction can hardly be negatively affected by faulty products.
-In fact, in the theoretical framework (see Section \ref{sect:theory:detailed}), minimal functionalities and less concern about quality aspects have low-impact on customers' satisfaction, supporting startup's ability to mainly focus on \textit{speed-up development}.
+In fact, in the GSM model, minimal functionalities and less concern about quality aspects have low-impact on customers' satisfaction, supporting startup's ability to mainly focus on \textit{speed-up development}.
 Moreover operating in the \textit{excitement area} is fully compliant with an evolutionary approach and minimal functionalities since from the first product released to the market, it is unlikely to lose users because of high-quality demand.
@@
 
-Nevertheless, since the amount of the results did not allow ``the lack of experience'' to be in the theoretical framework (in view of the fact that not enough data were identified \footnote{A category is formed if the codes are representative of more than the 70\% of the analyzed transcripts.}), this confounding factor requires further investigation.
+Nevertheless, since the amount of the results did not allow ``the lack of experience'' to be in the GSM model (in view of the fact that not enough data were identified \footnote{A category is formed if the codes are representative of more than the 70\% of the analyzed transcripts.}), this confounding factor requires further investigation.
 
 %--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
-\subsection{High-level relations validity}
+\subsection{Static assessment of the GSM}
 \label{sect:theory:validation:stats}
@@
 However their correctness has been verified with a cross-analysis of the extant literature and by attesting the causal relations among codes and categories, through axial and selective coding processes.
+
+%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+\section{Discussion}
+\label{sect:theory:impl}
+
+In this section we present a list of the most relevant implications which emerge from the behaviour of early-stage startups, formally expressed in the GSM model, answering to our research question (\textit{What is the current state-of-practice related to software development strategy in early-stage startups?}) through its sub-questions.
+Although the startups we inquired in the case study were spread across different nations and market sectors, we have found several patterns that emerged very clearly:
+
+\begin{compactitem}
+
+\item The most urgent priority of software development is to shorten time-to-market.
+When developing a new product, which has not been attempted before, it is crucial for the company to release the product soon and observe the users reactions.
+The company's chances to survive, are strictly related to the ability to find the right product/market fit as quickly as possible.
+Developing software without releasing it for a long period of time, coming up with very complex solutions often is not a viable option.
+Business demands, investors interests, market pressures and the general uncertain conditions force startups to quickly iterate on the product by exploring and experimenting new functionalities progressively rolled-out to early-adopters.
+
+\item Startups do not apply any standard development methodology.
+When focused on building the first version of the product, startups do not observe any specific and standard development methodologies or processes.
+From an accurate and in-depth investigation into daily activities, it emerged that the number of  the observed development \textit{practices} is extremely small and never systematic (only little \textit{pair programming} sessions and rarely \textit{test driven development}).
+Despite the wide number of publications, which discuss the adoption of lightweight methodologies, we have found that especially in this first stage not only methodologies but even individual practices are basically neglected.
+
+\item The closest development approach undertaken by early-stage startups is the \textit{Lean startup methodology}.
+A highly evolutionary development approach, centered around the quick production of a functioning prototype and guided by customer feedback is a central aspect of the \textit{Lean startup methodology} pioneered by Ries \cite{Ries2011}.
+However, startups do not formally follow the cycle \textit{build-measure-learn} as it is proposed by the methodology, where they should set-up an experiment to \textit{test the riskiest element of the business idea}, nor make explicit use of the good practices suggested in the book.
+For example, when it comes to user feedback, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}.
+Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its subcategories.
+
+\item The greatest part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).
+Especially when bringing the first product to the market, the attitude of startups towards traditional engineering activities is strongly characterized by a general feeling of unnecessity.
+Every document, artifact or process, which does not produce a short-term tangible benefit to the software production is perceived as a waste of time.
+The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (CAT6), which is partially mitigated by different elements supporting the development.
+Project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
+Thus, by radically ignoring activities related to documentation, structures and processes, startups achieve a very high initial development speed in the first stage, but on the other hand they accumulate a technical debt that needs to be at least partially fulfilled when the company and the product grow.
+
+\item The first release of the product includes only a limited set of well suitable functionalities focused on user experience.
+The beta version is usually focused on a minimal and essential set of functionalities providing a reasonable user experience (UX).
+As confirmed by practitioners, the user must be able to accomplish a task on the product without major interruptions in the flow, which would cause frustration before being able to capture his attention.
+To realize UX in a short time frame, developers focus on a limited number of suitable functionalities, especially in the first version(s) of the product.
+Other product qualities are regarded as less important: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important, even more than time-to-market \cite{Offutt2002}.}.
+From a business perspective, one might argue that established companies might take advantage of the new idea and develop a better version of the product internally.
+Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}.
+Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process while focusing effort on acquiring larger shares of a ``mass market'' sector.
+Furthermore, thanks to cloud computing and more in general high availability of third party components, aspects such as performance, security and scalability can be achieved with a reasonable effort.
+Thus, given the general lack of quality concerns, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
+
+\item Engineering activities are supported by low-precision artifacts and collaborative tools integrated with the development environment.
+Startups take advantage of the simplicity of cutting-edge tools, which require a minimum learning and maintenance effort and provide a large payoff.
+Tools are used for a vast number of tasks: ticked-based boards to manage user stories; systems to collect user feedback; tools for tracing product metrics; and scripts for managing deployments.
+The most prominent families of tools, which support software development in startups are modern version control systems.
+They are used by practitioners not only to manage the code base and the repositories, but for many other purposes such as: task assignment; tracing responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.
+On the other hand, an excessive usage of  \textit{out-of-the-code} tools can negatively affect the development speed, as significantly reported by C11: \textit{``We realized we were losing too much time moving tickets around boards than actually working on stories.
+You need to find the balance between the time spent with tools and the amount of actual coding''}.
+This risk is often mitigated with an intense usage of offline manual tools: especially post-it notes and whiteboards are reported to be particularly effective to foster internal communication within the team.
+
+\item The founding team is the most important determinant of speed.
+More than techniques and methodologies, the features of the team are the essential factors to achieve high-development speed.
+Discussed in the theoretical category the \textit{team is the catalyst of development}, this result has been confirmed in multiple occasions both in the interviews and in the questionnaires.
+What is truly important for the startup, among other factors, is the extremely high-colocation which enables continuous informal and colloquial coordination and discussions.
+The initial (small) development team usually spends most of the hours of the week working closely to implement the product, often extending the conventional office hours.
+The large amount of communication, spontaneously held in informal places, has been proven to be an extremely effective element of several successful software projects \cite{Highsmith2000}.
+Hence, the vastly \textit{informal} development environment of startups acts as a catalyst of speed.
+Other important elements, which contribute in achieving such an initial high-speed, are: minimal set of functionalities with low concerns on product quality; evolutionary prototyping approach; smart use of advanced tools and frameworks; low precision artifacts; and the close colocation and capability of the team.
+
+\item The initial lack of structures and processes negatively affects the performance when the company starts to grow.
+One of the primary objectives of startups is to expand their business into fast growing markets.
+In fact, if successful, the company will face a growth in terms of number of customers, employees and product functionalities.
+Then, the increased complexity arises the necessity of controlling the initially chaotic software development environment.
+Under delicate circumstances, where an increased market demand meets a poorly engineered product and process,  the development team need to start returning the accumulated technical debt.
+As reported by practitioners, introducing more structures, documentation and standard procedures, cause an initial drop-down in performance which severity depends on the speed of growth, the amount of technical debt and the capacity of the team to absorb it.
+
+\item Startups bring the first product to market in a very short-time and practitioners are satisfied with the software development strategy.
+The companies inquired in the case study were typically able to release the first beta of the product in a short time frame.
+Although practitioners acknowledge the initial negative consequences of the technical debt on subsequent performances, they confirm that it is better spending some effort in later stage to recover the debt than losing time in over-engineering the production initially with the risk of never achieving any significant growth.
+\end{compactitem}
+
+All of the above observations are extracted from the analysis of the model, which is systematically derived from empirical data of the case study.
+Indeed, with slightly different levels of adherence, the implication we presented are reflected in the behaviour of the great majority of companies in the sample.
+The results of this analysis, which has been focused on the software development strategy from the idea conception to the first release, indicate that startups are far from adopting standard or ad-hoc development methodologies, especially in this phase.
+The typical tendency is to focus on the team capability to implement and quickly iterate on a prototype, which is released very fast.
+Thus, in a context where even for the most \textit{lightweight} agile methodologies is hard to penetrate, we believe that it is premature to discuss issues related to SPI.
+Researchers should rather focus on the trade-off between development speed and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which appears to be the most important determinant for the future of software development in the company.
 
@@
 Moreover, we developed a rigorous data collection protocol to create case study databases and employ integration of multiple data collection methods.
-Reporting bias was mitigated by packaging all the needed material for conducting new researches in other contexts in order to criticize our findings (interview package with instructions has been made available online\footnote{Available at \url{https://github.com/adv0r/BTH-Interview-Package} .}.
+Reporting bias was mitigated by packaging all the needed material for conducting new researches in other contexts in order to criticize our findings (interview package with instructions has been made available online\footnote{Available at \url{https://github.com/adv0r/BTH-Interview-Package}.}.
 Moreover, two academic supervisors with expertise in the area have conducted a peer-review analysis of framework's constructs.

### 2013-02-19 article/main.tex
Changes before Carmines' last push

--- 2013-02-14 article/main.tex
@@
 As result, the Greenfield startup model (GSM)  reveals the urgent priority of startups of releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early collected user feedback.
-Nevertheless, the need of shortening time-to-market, by speeding-up the development through low-precision and product-centric engineering activities, is counterbalanced by the need of restructuring the product before setting off for further growth.
+Nevertheless, the need of shortening time-to-market, by speeding-up the development through low-precision and product-centric engineering activities, is
+counterbalanced by the need of restructuring the product before setting off for further growth.
 \end{abstract}
@@
 small and web companies), the combination of different factors makes the specific development context quite unique \cite{Blank2005, Sutton2000}.
-Research is needed to support their engineering activities \cite{Coleman2005}, guiding practitioners in taking decisions and avoiding choices that could easily lead the whole business to failure \cite{Kajko-Mattsson2008}.
+Therefore, research is needed to support their engineering activities \cite{Coleman2005}, guiding practitioners in taking decisions and avoiding choices that could easily lead the whole business to failure \cite{Kajko-Mattsson2008}.
 However, despite the impressive size of the startup ecosystem \cite{ISI:000243253000007}, the state-of-the-art presents a gap.
@@
 By means of the Green-field startup model (GSM), we captured the underlying phenomenon of software development in early-stage startups.
-The framework is based on a systematic procedure  which is grounded into empirical concepts obtained by conduction of the case-study.
+The framework is based on a systematic procedure  which is grounded into empirical concepts obtained by a case-study.
 GSM presents the most significant themes of the development strategies that characterise startups' context.
 
-The most urgent priority of software development is to shorten time-to-market.
-The greater part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).The first release of the product includes only a limited set of well suitable functionalities.
+The results of our study suggest that the most urgent priority of software development in startups is to shorten time-to-market.
+The greater part of engineering activities are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing).The first release of the product includes only a limited set of well suitable functionalities.
 
@@
 
-The remain of this paper is structured as follow:  we provide relevant information about startups in section \ref{backg}.
+The remainder of this paper is structured as follows:  we provide relevant information about startups in section \ref{backg}.
 Section \ref{resmet} introduces the research questions and discusses the design and execution of the research process.
 Results are discussed in section \ref{res:gsm}, which presents the GSM model.
-Section \ref{res:val}  compares results of this research to literature and frameworks.
+Section \ref{res:val}  compares results of this research to literature and frameworks presented in this field.
 Section \ref{valt} discusses the validity threats considered during this study.
@@
 
+%MUN: This is maybe only a minor concern, but it may be relevant for certain reviewers: The references you use
+% in the introduction are not very recent. Look at the list of papers [1-15], most of them are published before 2005.
+% I understand that you use some references to strengthen your general arguments (which can refer to older publications),
+% but your introduction should also show the relevance of your research right now. This is best shown by referring to
+% recent (max 5 years old) literature. Think about if you can add more recent (2010-now), well published, papers in the
+% introduction.
+
 % You must have at least 2 lines in the paragraph with the drop letter
@@
 
-Modern entrepreneurship, which was born more than thirty years ago \cite{storey1982entrepreneurship}, has been boosted by the advent of the consumer internet markets in the middle of the nineties and culminated with the notorious \textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126}.
+Modern entrepreneurship, which was born more than thirty years ago \cite{storey1982entrepreneurship}, has been boosted by the advent of the consumer internet markets in the middle of the 1990s and culminated with the notorious \textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126}.
 Several years later, with the massification of the internet and mobile devices, we are now assisting to an impressive proliferation of software ventures - metaphorically referred as the \textit{startup bubble}.
 The easy access to vast potential markets and the low cost of services distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}.
-Inspired by stories of overwhelming successes, a large number of software businesses are created every day.
+Inspired by success stories, a large number of software businesses are created every day.
 However, the great majority of these companies fail within two years from their creation \cite{Crowne2002}.
@@
 Some attempts to taylor lightweight processes to startups reported basic failure of their application: \textit{``Everyone is busy, and software engineering practices are often one of the first places developers cut corners''} \cite{Martin2007}.
-Rejecting the notion of repeatable and controlled processes, startups prominently take advantage of unpredictable, reactive and low-precision \cite{surviving-os-cockburn} engineering practices \cite{Sutton2000, Tanabian2005, Chorev2006, Kakati2003}.
+Rejecting the notion of repeatable and controlled processes, startups prominently take advantage of unpredictable, reactive and low-precision \cite{
+surviving-os-cockburn} engineering practices \cite{Sutton2000, Tanabian2005, Chorev2006, Kakati2003}.
 
@@
 
-The word \textit{startup} appeared in the SE literature for the first time in 1994 in an article written by Carmel \cite{Camel1994a} where he studied the \textit{time-to-completion in young package firm}.
+The word \textit{startup} appeared in the SE literature for the first time in 1994 in an article written by Carmel \cite{Camel1994a} where he studied the \textit{time-to-completion in a young package firm}.
+%MUN: If this is a quote from
+%Carmel, put it in quotation marks, like you did in the previous section with Suttons quote. Generally, check that you
+%are consistent how you use quotations throughout the paper.
 He noticed how these companies were particularly innovative and successful, advocating for the need of more research investigating software development practices so as to replicate success and try to transfer it to other technology sectors.
@@
 
-A study of Kajko-Mattsson  \cite{Kajko-Mattsson2008}, which investigated a Swedish software startup, reported a heavy lack of requirements gathering process, minimal project management, lack of control over the change requests, absence of documentation to track the status and progress of the process and defective releases.
+A study of Kajko-Mattsson \cite{Kajko-Mattsson2008}, which investigated a Swedish software startup, reported a heavy lack of requirements gathering processes, minimal project management, lack of control over the change requests, absence of documentation to track the status and progress of the process, and defective releases.
 Accordingly, Ambler et al.
 report how two startups approaching to an upcoming IPO started to require processes to focus on scalable solutions, in view of the growing company's size in terms of users and employees \cite{Ambler2002}.
-In this regard, Crowne, in \cite{Crowne2002}, specifies different stages through which software startups evolve.
+In this regard, Crowne \cite{Crowne2002} specifies different stages through which software startups evolve.
 Starting without any established workflows, startups grow over time, creating and stabilizing processes to eventually improving them only when sufficiently mature.
 
-But when startups have no time for training and orienting activities, as discussed in \cite{Sutton2000}, their main focus remains on team capabilities instead of prescriptive processes hiring people who can \textit{``hit the ground running''} \cite{Yoffie1999}.
-Empowering the team and focus on methodological attributes of the processes oriented in prototyping, proof-of-concepts, mock-ups and demos, to test basic functionalities, have been the primary priority in startups as described in \cite{Camel1994a}.
-Only when they grow, formal methodologies arise, followed by a conduction of quality assurance and long-term planning processes \cite{Yoffie1999}.
-
-As studied in \cite{Tingling2007}, the maturity of a company affects the extent to which processes are adopted.
-The author reports how introducing Extreme Programming (XP) principles \cite{Beck:2004:EPE:1076267} in the development process was challenging because of the need of trained team-members for fully implementing the methodology.
-In fact, \cite{Silva2005} was able to start with all the XP practices in place only after six months of coaching the team, trying to enhance maturity from day-one.
+But when startups have no time for training and orienting activities, as discussed in \cite{Sutton2000}, their main focus remains on team capabilities instead of prescriptive processes, hiring people who can \textit{``hit the ground running''} \cite{Yoffie1999}.
+Empowering the team and focus on methodological attributes of the processes oriented in prototyping, proof-of-concepts, mock-ups and demos, to test basic functionalities, have been the primary priority in startups \cite{Camel1994a}.
+Only when they grow, formal methodologies arise, followed by performing quality assurance and long-term planning processes \cite{Yoffie1999}.
+
+The maturity of a company affects the extent to which processes are adopted.
+Tingling \cite{Tingling2007} reports how introducing Extreme Programming (XP) principles \cite{Beck:2004:EPE:1076267} in the development process was challenging because of the need of trained team-members for fully implementing the methodology.
+Da Silva and Kon \cite{Silva2005} were able to start with all the XP practices in place only after six months of coaching the team, trying to enhance maturity from day-one.
 Nevertheless, even then, customization of practices were inevitably implemented to adapt the processes to the undertaken startups' context \cite{Deias}.
@@
 As partially discussed above, contributions to flexibility and reactiveness of the development process has been conducted prominently by means of \textit{Lean} \cite{Gautam2008} and \textit{Agile} \cite{Abrahamsson2002} methodologies (also reported in \cite{Taipale2010, Kuvinka2011}), where the extreme uncertain conditions lead startups to learn fast from trials and errors with a strong customer relationship in order to avoid wasting time in building wrong functionalities and prevent rapidly exhaustion of resources \cite{Midler2008, Hilmola2003, Sutton2000}.
-Customer involvement in software development has also been discussed in  \cite{Yogendra2002} as important factor to encourage an early alignment of business concerns to the technology strategies, because both are salient considerations to be successful \cite{Tingling2007}.
+Customer involvement in software development has also been discussed by Yogendra \cite{Yogendra2002} as an important factor to encourage an early alignment of business concerns to the technology strategies, because both are salient considerations to be successful \cite{Tingling2007}.
 
@@
 
-When startups are mature enough to support software process improvement (SPI), the solutions considered according to the state-of-the-art are oriented towards light-weight processes such as a design of development process based on XP, proposed by Zettel in \cite{Zettel2001}.
+When startups are mature enough to support software process improvement (SPI), the solutions considered according to the state-of-the-art are oriented towards light-weight processes such as a design of development process based on XP, proposed by Zettel \cite{Zettel2001}.
 The process consists of a set of activities and artifacts (in addition to some important roles) defined in order to identify responsibilities and tools to utilize.
@@
 
-The first publication mentioning the problem of \textit{one-size-fits-all}, related to the SPI  representations for startups, is described in \cite{Fayad1997}.
-The author reveals the problem in actuating the same \textit{best-practices} criteria for established companies in 10-person software startups.
-Thoroughly remarked in \cite{Sutton2000}, Sutton states that problems of SPI in software startups arise because of: the dynamism of the development process, which precludes repeatability; organizational maturity, that cannot be maintained from startups in view of lack of corporate direction; severe lack of resources, both human and technological for process definition, implementation, management, training \ldots; in conclusion, the primary benefits of SPI do not address startups, which instead of promoting product quality, aim to minimize time-to-market.
-
-Additionally, the role of SPI has always been neglected because seen as an obstacle to the development team's creativity and flexibility as described in \cite{Coleman2008a} and to the need of a quick delivering product process environment.
-In fact, product quality is left aside in favour of minimal and suitable functionalities to shorten the time-to-market.
+The problem of \textit{one-size-fits-all}, related to the SPI  representations for startups, is described by Fayad \cite{Fayad1997}.
+The author discusses the problem in actuating the same \textit{best-practices} criteria for established companies in 10-person software startups.
+Thoroughly remarked in \cite{Sutton2000}, Sutton states that problems of SPI in software startups arise because of: the dynamism of the development process, which precludes repeatability; organizational maturity, that cannot be maintained from startups in view of lack of corporate direction; severe lack of resources, both human and technological for process definition, implementation, management, and training.
+In conclusion, the primary benefits of SPI do not address startups, which instead of promoting product quality, aim to minimize time-to-market.
+
+Additionally, the role of SPI has been neglected because seen as an obstacle to the development team's creativity and flexibility \cite{Coleman2008a}, and to the need of a quick delivering product process environment.
+%MUN: the second
+%part of the sentence is just a statement without reference. If it refers to a statement by Coleman, put his reference
+%at the end OR find the correct reference OR delete the statement.
+Product quality is left aside in favor of minimal and suitable functionalities to shorten the time-to-market.
 As reported in two studies of Mater and Mirel \cite{Mater2000, Mirel2000}, quality aspects, mostly taken in consideration in internet startups, are oriented to usability and scalability, even though the market and application type heavily influences the quality-demand \cite{Coleman2008, Kim2005}.
 
-Finally, to maintain the development activities, oriented to limited but suitable functionality, many studies suggest to externalize the complexity of parts of the project to third party solutions by means of outsourcing activities, software reuse and open-source strategies \cite{Hanna2010, Jansen2008, Wall2001, Bean2005}.
+To maintain the development activities, oriented to limited but suitable functionality, studies suggest to externalize the complexity of parts of the project to third party solutions by means of outsourcing activities, software reuse and open-source strategies \cite{Hanna2010, Jansen2008, Wall2001, Bean2005}.
+%MUN: Can you split up the references?
+I.e.
+write:
+%"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source strategies \cite{Bla}.
 
@@
 The metaphoric neologism was originally introduced by Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the attention of SE researchers\footnote{To attest the fact that \textit{technical debt} is gaining traction among researchers, we mention two important contributions which characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt} organized by the Software Engineering Institute  and ICSE.}.
-The concept of \textit{technical debt} is well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea is that developers sometimes accept compromises a system in one dimension (e.g., modularity) to meet an urgent demand in some other dimension (e.g., a deadline), and that such compromises incur a ``debt'' on which ``interest'' has to be paid and which the ``principal'' should be repaid at some point for the long-term health of the project''}.
-The compromise between high-speed and high-quality engineering is faced daily by software startups, not only in terms of architecture design but in multifaceted aspects (weak project management, testing, process control, \ldots).
-
-In conclusion, since \textit{``all decisions related to product development are trade-off situations''} \cite{Hilmola2003}, generally startups optimize workflows to the dynamic context they are involved into.
-In fact they typically adopt any development style that might work to support their first needs in what is called the \textit{``Just do it''} school of software startups \cite{Ries2011}.
+The concept of \textit{technical debt} is well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea is that developers sometimes accept compromises in a system in one dimension (e.g., modularity) to meet an urgent demand in some other dimension (e.g., a deadline), and that such compromises incur a ``debt'' on
+which ``interest'' has to be paid and which the ``principal'' should be repaid at some point for the long-term health of the project''}.
+The compromise between high-speed and high-quality engineering is faced daily by software startups, not only in terms of architecture design but in multifaceted aspects (weak project management, testing, process control).
+
+Since \textit{``all decisions related to product development are trade-off situations''} \cite{Hilmola2003}, generally startups optimize workflows to the dynamic context they are involved into.
+They typically adopt any development style that might work to support their first needs in what is called the \textit{``Just do it''} school of software startups \cite{Ries2011}.
 Additionally, as remarked by Coleman, \textit{``many managers just decide to apply what they know, as their experience tells them it is merely common sense''} \cite{Coleman2008}.
 
-In this paper we make use of context-specific terminologies, which will be used to facilitate the readability and clarify any ambiguities throughout the whole paper in view of the weak state-of-the-art presented in \cite{SMS} :
+\subsection{Terminology}
+In order to set a common ground and to prevent ambiguity, we use the following terminology throughout the paper:
 
@@
 \end{figure*}
+%MUN: I think you can increase the font size in the figure a bit. Try thereby to keep the 
+%height of the figure, expanding however the width (use horizontal space since it is a two 
+%column figure).
+%Write "Data analysis" horizontally
 
@@
 
-Initially the boundaries of the research domain were set by means of non-systematic \textit{literature survey}, which provided the initial keywords used further in the research process.
+Initially the boundaries of the research domain were set by means of non-systematic literature survey, which provided the initial keywords used further in the research process.
 Moreover it helped us in defining the research questions (RQs):
@@
 
-Despite GT has its roots in the social sciences, it has been increasingly used by \textit{Information System} researchers and has been almost ignored by the SE research \cite{Coleman2007}.
+Despite GT has its roots in the social sciences, it has been increasingly used by Information System researchers and has been almost ignored by the SE research \cite{Coleman2007}.
 In 1997 Bertelsen advocated for a need of more qualitative research in SE \cite{Bertelsen1997}, and in the last decade we assisted to an increasing number of publications that used a GT-like approach in SE \cite{Coleman2008a,Coleman2007, Sulayman2012}.
@@
 The latter approach empowers researchers' \textit{``theoretical sensitivity''} \cite{Corbin1990}, and encourages them to outline the research problem beforehand.
-Since the knowledge obtained from parallel execution of the SMS \cite{SMS} and our direct experience with startups companies provided a good initial level of knowledge, in this study we use the Corbin and Strauss approach \cite{Strauss1998}.
+Since the knowledge obtained from parallel execution of the SMS \cite{SMS} and our direct experience with startups companies provided
+a good initial level of knowledge, in this study we use the Corbin and Strauss approach \cite{Strauss1998}.
 
@@
 Then, to obtain a competitive advantage in the market, startups typically make use of cutting-edge solutions, characterised by an evolution that cannot be foreseen in the long run.
-However, a special role in daily decisions is covered by user feedback and requests, which represent the main drivers to define the product features to provide in the short-term.
+However, a special role in daily decisions is covered by user feedback
+and requests, which represent the main drivers to define the product features to provide in the short-term.
 
@@
 Indeed, staying one week in production enables us to identify 90\% of bugs''.} The lack of complete automated tests is partially also motivated by the fact that \textit{users are fault-tolerant in innovative beta products} and  by the limited number of functionalities, which allow the team to easily control critical bugs.
-However, in certain cases where components of the system might cause loss of data or severe damages to the product or users, engineers realize a reasonable level of automatic testing.
-In such cases, aided by modern automatic tools, they can quickly assess the status of the system integration as they add new functionalities to the product.
+However, in certain cases where components of the system might cause loss of data or severe damages to the product or users, engineers realize a reasonable level of
+automatic testing. In such cases, aided by modern automatic tools, they can quickly assess the status of the system integration as they add new functionalities to the product.
 
@@
 To further \textit{speed-up development}, startups \textit{simplify issue management by integrating it with feature management}.
-Then, the \textit{limited number of functionalities} and the {use of standard/known technologies} with a \textit{simple workflow} are the main reasons for not establishing a heavy project management process since the project and technologies can be autonomously managed by the development team (with characteristics of: co-location, self-organization and very small size) .
+Then, the \textit{limited number of functionalities} and the {use of standard/known technologies} with a \textit{simple workflow} are the main reasons for
+not establishing a heavy project management process since the project and technologies can be autonomously managed by the development team (with characteristics of: co-location, self-organization and very small size) .
 
@@
 For example, when it comes to user feedback, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}.
-Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its subcategories.
+Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its
+subcategories.
 
@@
 The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (CAT6), which is partially mitigated by different elements supporting the development.
-Project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
+Project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-
+precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary
+bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
 Thus, by radically ignoring activities related to documentation, structures and processes, startups achieve a very high initial development speed in the first stage, but on the other hand they accumulate a technical debt that needs to be at least partially fulfilled when the company and the product grow.
@@
 Other product qualities are regarded as less important\footnote{Note that from a business perspective, one might argue that established companies might take advantage of the new idea and develop a better version of the product internally.
-Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}.
+Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of
+users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}.
 Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process while focusing effort on acquiring larger shares of a ``mass market'' sector.}: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important, even more than time-to-market \cite{Offutt2002}.}.
 Furthermore, thanks to cloud computing and more in general high availability of third party components, aspects such as performance, security and scalability can be achieved with a reasonable effort.
-Thus, given the general lack of quality concerns, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
+Thus,
+given the general lack of quality concerns, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
 
@@
 They are used by practitioners not only to manage the code base and the repositories, but for many other purposes such as: task assignment; tracing responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.
-On the other hand, an excessive usage of  \textit{out-of-the-code} tools can negatively affect the development speed, as significantly reported by C11: \textit{``We realized we were losing too much time moving tickets around boards than actually working on stories.
+On the other hand, an excessive usage of  \textit{out-of-the-code} tools can negatively affect the development speed, as significantly reported by C11: \
+textit{``We realized we were losing too much time moving tickets around boards than actually working on stories.
 You need to find the balance between the time spent with tools and the amount of actual coding''}.
@@
 Thus, in a context where even for the most \textit{lightweight} agile methodologies is hard to penetrate, we believe that it is premature to discuss issues related to SPI.
-Researchers should rather focus on the trade-off between development speed and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which appears to be the most important determinant for the future of software development in the company.
+Researchers should rather focus on the trade-off between development speed and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which
+appears to be the most important determinant for the future of software development in the company.
 
@@
 As Coleman identified the influence of the founders' and managers' background over the software development process, also GSM identifies how the  CEOs and CTOs background shapes the high-level strategies adopted in developing the initial product.
-Notwithstanding, team members remain self-organized, able to intervene in all the aspects of the development process without any direct supervision, as discussed in CAT4.
+Notwithstanding, team members remain self-organized, able to intervene in all the
+aspects of the development process without any direct supervision, as discussed in CAT4.
 
@@
 In other words, even though ignoring SPI practices is advisable, not following SPI does not predict major adaptiveness of startups development to the market.
-But, to imply a major evolutionary product/market fit, certain conditions explained in GSM need to be considered.
+But, to imply a major evolutionary product/market fit, certain
+conditions explained in GSM need to be considered.
 
@@
 For more detailed discussion about the importance of people the reader is referred to some prominent authors: Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999}, Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph \cite{Adolph2011} and Cockburn \cite{people-first-order}.
-Especially in early-stage startups, where the whole company overlaps with the development team, factors related to people are without any doubt the most important ones (confirmed with the empirical data of  our case study.
+Especially in early-stage startups, where
+the whole company overlaps with the development team, factors related to people are without any doubt the most important ones (confirmed with the empirical data of  our case study.
 
@@
 See \url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, startups failed in practicing engineering activities in view of eccesive lack of flexibility.
-In fact, developers fast moved to simpler tools such as the whiteboard, where they were able to provide input and insights, question assumptions, share past experiences, and even talk about potential implementation strategies.
+In fact, developers fast moved to simpler tools such as the whiteboard, where they were able to provide input and insights, question assumptions, share
+past experiences, and even talk about potential implementation strategies.
 In general the development remained highly iterative, where project teams started to have a little modeling and testing only when necessary.
@@
 Despite the effort spent from the researcher to establish a development workflow, practices were not fully conducted even though some marginal benefits have been reported: planned major release from four to six months; planned minor releases from two to four weeks; encompassing urgent corrective releases from one to five days.
-Regarding the previously mentioned problems, only partial improvements have been achieved since the major obstacle was to motivate employees to change their habits.
+Regarding the previously mentioned problems, only
+partial improvements have been achieved since the major obstacle was to motivate employees to change their habits.
 With perfect hindsight the researchers claim that understanding and adapting solutions by solving one problem at a time would have increased benefits.
@@
 In fact, in the theoretical framework (see Section \ref{sect:theory:detailed}), minimal functionalities and less concern about quality aspects have low-impact on customers' satisfaction, supporting startup's ability to mainly focus on \textit{speed-up development}.
-Moreover operating in the \textit{excitement area} is fully compliant with an evolutionary approach and minimal functionalities since from the first product released to the market, it is unlikely to lose users because of high-quality demand.
+Moreover operating in the \textit{excitement area} is fully compliant with an evolutionary approach and minimal functionalities since from the first product released
+to the market, it is unlikely to lose users because of high-quality demand.
 
@@
 Consequently the team management requires higher control and coordination activities that inevitably hinder the flexibility of the development environment at the beginning, that is essential in early-stage startups.
-Accordingly, also Coleman in \cite{Coleman2007,Coleman2008a,Coleman2008} states that the operating strategies towards \textit{``minimum processes''} in startups is not a matter of poor knowledge and training, but rather it is the necessity of operating with solutions that let the company move faster.
+Accordingly, also Coleman in \cite{Coleman2007,
+Coleman2008a,Coleman2008} states that the operating strategies towards \textit{``minimum processes''} in startups is not a matter of poor knowledge and training, but rather it is the necessity of operating with solutions that let the company move faster.
 
@@
 Notwithstanding the issues surrounding semi-structured interviews are vital.
-Within our study, although the reality might be different from what has been described, the reported data are CTOs' and CEOs' perception, which is used to base their decisions.
+Within our study, although the reality might
+be different from what has been described, the reported data are CTOs' and CEOs' perception, which is used to base their decisions.
 
@@
 Especially when bringing the first product to market, startups' most urgent priority is releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early feedback and collected metrics.
-In this stage, startups often discard any formal project management, documentation, analysis, planning, testing and other traditional process activities.
-In fact, practitioners take advantage of an evolutionary prototyping approach, using well-integrated tools and externalizing complexity to third party solutions.
+In this stage, startups often discard any formal project management, documentation, analysis, planning, testing and other traditional process
+activities. In fact, practitioners take advantage of an evolutionary prototyping approach, using well-integrated tools and externalizing complexity to third party solutions.
 
@@
 In fact, as confirmed by analyzing other framework in SE, the use of prescriptive and rigid structures over the development activities would hinder performance since startups operate across the border between chaotic and complex domain.
-Therefore, from a software development perspective, working with low-precision engineering activities represents the only viable strategy that startups can follow in the early stage of their lifecycle.
+Therefore, from a software development perspective, working with low-precision engineering activities represents the only viable strategy that startups
+can follow in the early stage of their lifecycle.
 

### 2013-02-19 article/main.tex

@@
 As result, the Greenfield startup model (GSM)  reveals the urgent priority of startups of releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early collected user feedback.
-Nevertheless, the need of shortening time-to-market, by speeding-up the development through low-precision and product-centric engineering activities, is
-counterbalanced by the need of restructuring the product before setting off for further growth.
+Nevertheless, the need of shortening
+time-to-market, by speeding-up the development through low-precision and product-centric engineering activities, is counterbalanced by the need of restructuring the product before setting off for further growth.
 \end{abstract}
@@
 The results of our study suggest that the most urgent priority of software development in startups is to shorten time-to-market.
-The greater part of engineering activities are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing).The first release of the product includes only a limited set of well suitable functionalities.
+The greater part of engineering activities are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing).
+The first release of the product includes only a limited set of well suitable functionalities.
 
@@
 
-Modern entrepreneurship, which was born more than thirty years ago \cite{storey1982entrepreneurship}, has been boosted by the advent of the consumer internet markets in the middle of the 1990s and culminated with the notorious \textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126}.
+Modern entrepreneurship has been boosted by the advent of the consumer internet markets in the middle of the 1990s and culminated with the notorious \textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126, storey1982entrepreneurship}.
 Several years later, with the massification of the internet and mobile devices, we are now assisting to an impressive proliferation of software ventures - metaphorically referred as the \textit{startup bubble}.
@@
 \subsection{Technical debt}
-Notwithstanding, a new interesting area of the SE research, trying to tackle the problem of the \textit{technical debt}, seems to bring interesting implications in studying development in software startups.
+Notwithstanding, a new area of the SE research, trying to tackle the problem of the \textit{technical debt},  brings and encompasses different implications in studying development in software startups \cite{Tom2013}.
 The metaphoric neologism was originally introduced by Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the attention of SE researchers\footnote{To attest the fact that \textit{technical debt} is gaining traction among researchers, we mention two important contributions which characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt} organized by the Software Engineering Institute  and ICSE.}.
-The concept of \textit{technical debt} is well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea is that developers sometimes accept compromises in a system in one dimension (e.g., modularity) to meet an urgent demand in some other dimension (e.g., a deadline), and that such compromises incur a ``debt'' on
-which ``interest'' has to be paid and which the ``principal'' should be repaid at some point for the long-term health of the project''}.
+The concept of \textit{technical debt} is well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea is that developers sometimes accept compromises in a system in one dimension (e.g., modularity) to meet an urgent demand in some other dimension (e.g., a deadline), and that such compromises incur a ``debt'
+' on which ``interest'' has to be paid and which the ``principal'' should be repaid at some point for the long-term health of the project''}.
 The compromise between high-speed and high-quality engineering is faced daily by software startups, not only in terms of architecture design but in multifaceted aspects (weak project management, testing, process control).
@@
 \item \textit{Intervening conditions} are  summarized by the extremely uncertain development environment.
-\item \textit{Action and interaction strategies} are represented by the accumulation of technical debt (see CAT5).
-\item \textit{Consequences} lead to a temporary performance drop-down (see CAT6).
+\item \textit{Action and interaction strategies} are represented by the accumulation of technical debt.
+\item \textit{Consequences} lead to a temporary performance drop-down.
 \end{compactitem}
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
-\section{Theory implications}
-\label{sect:theory:impl}
-
-In this subsection we present a list of the most relevant implications which emerge from the behaviour of early-stage startups, formally expressed in the GSM model, answering to our research question (\textit{What is the current state-of-practice related to software development strategy in early-stage startups?}) through its sub-questions.
-Although the startups we inquired in the case study were spread across different nations and market sectors, we have found several patterns that emerged very clearly:
+
+%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+\section{Validation}
+\label{res:val}
+
+In this section we discuss the validity of the GSM model by means of cross-methodological observations:
 
 \begin{compactitem}
-
-\item The most urgent priority of software development is to shorten time-to-market.
-When developing a new product, which has not been attempted before, it is crucial for the company to release the product soon and observe the users reactions.
-The company's chances to survive, are strictly related to the ability to find the right product/market fit as quickly as possible.
-Developing software without releasing it for a long period of time, coming up with very complex solutions often is not a viable option.
-Business demands, investors interests, market pressures and the general uncertain conditions force startups to quickly iterate on the product by exploring and experimenting new functionalities progressively rolled-out to early-adopters.
-
-\item Startups do not apply any standard development methodology.
-When focused on building the first version of the product, startups do not observe any specific and standard development methodologies or processes.
-From an accurate and in-depth investigation into daily activities, it emerged that the number of  the observed development \textit{practices} is extremely small and never systematic (only little \textit{pair programming} sessions and rarely \textit{test driven development}).
-Despite the wide number of publications, which discuss the adoption of lightweight methodologies, we have found that especially in this first stage not only methodologies but even individual practices are basically neglected.
-
-\item The closest development approach undertaken by early-stage startups is the \textit{Lean startup methodology}.
-A highly evolutionary development approach, centered around the quick production of a functioning prototype and guided by customer feedback is a central aspect of the \textit{Lean startup methodology} pioneered by Ries \cite{Ries2011}.
-However, startups do not formally follow the cycle \textit{build-measure-learn} as it is proposed by the methodology, where they should set-up an experiment to \textit{test the riskiest element of the business idea}, nor make explicit use of the good practices suggested in the book.
-For example, when it comes to user feedback, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}.
-Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its
-subcategories.
-
-\item The greatest part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).
-Especially when bringing the first product to the market, the attitude of startups towards traditional engineering activities is strongly characterized by a general feeling of unnecessity.
-Every document, artifact or process, which does not produce a short-term tangible benefit to the software production is perceived as a waste of time.
-The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (CAT6), which is partially mitigated by different elements supporting the development.
-Project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-
-precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary
-bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
-Thus, by radically ignoring activities related to documentation, structures and processes, startups achieve a very high initial development speed in the first stage, but on the other hand they accumulate a technical debt that needs to be at least partially fulfilled when the company and the product grow.
-
-\item The first release of the product includes only a limited set of well suitable functionalities focused on user experience.
-The beta version is usually focused on a minimal and essential set of functionalities providing a reasonable user experience (UX).
-As confirmed by practitioners, the user must be able to accomplish a task on the product without major interruptions in the flow, which would cause frustration before being able to capture his attention.
-To realize UX in a short time frame, developers focus on a limited number of suitable functionalities, especially in the first version(s) of the product.
-Other product qualities are regarded as less important\footnote{Note that from a business perspective, one might argue that established companies might take advantage of the new idea and develop a better version of the product internally.
-Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of
-users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}.
-Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process while focusing effort on acquiring larger shares of a ``mass market'' sector.}: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important, even more than time-to-market \cite{Offutt2002}.}.
-Furthermore, thanks to cloud computing and more in general high availability of third party components, aspects such as performance, security and scalability can be achieved with a reasonable effort.
-Thus,
-given the general lack of quality concerns, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
-
-\item Engineering activities are supported by low-precision artifacts and collaborative tools integrated with the development environment.
-Startups take advantage of the simplicity of cutting-edge tools, which require a minimum learning and maintenance effort and provide a large payoff.
-Tools are used for a vast number of tasks: ticked-based boards to manage user stories; systems to collect user feedback; tools for tracing product metrics; and scripts for managing deployments.
-The most prominent families of tools, which support software development in startups are modern version control systems.
-They are used by practitioners not only to manage the code base and the repositories, but for many other purposes such as: task assignment; tracing responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.
-On the other hand, an excessive usage of  \textit{out-of-the-code} tools can negatively affect the development speed, as significantly reported by C11: \
-textit{``We realized we were losing too much time moving tickets around boards than actually working on stories.
-You need to find the balance between the time spent with tools and the amount of actual coding''}.
-This risk is often mitigated with an intense usage of offline manual tools: especially post-it notes and whiteboards are reported to be particularly effective to foster internal communication within the team.
-
-\item The founding team is the most important determinant of speed.
-More than techniques and methodologies, the features of the team are the essential factors to achieve high-development speed.
-Discussed in the theoretical category the \textit{team is the catalyst of development}, this result has been confirmed in multiple occasions both in the interviews and in the questionnaires.
-What is truly important for the startup, among other factors, is the extremely high-colocation which enables continuous informal and colloquial coordination and discussions.
-The initial (small) development team usually spends most of the hours of the week working closely to implement the product, often extending the conventional office hours.
-The large amount of communication, spontaneously held in informal places, has been proven to be an extremely effective element of several successful software projects \cite{Highsmith2000}.
-Hence, the vastly \textit{informal} development environment of startups acts as a catalyst of speed.
-Other important elements, which contribute in achieving such an initial high-speed, are: minimal set of functionalities with low concerns on product quality; evolutionary prototyping approach; smart use of advanced tools and frameworks; low precision artifacts; and the close colocation and capability of the team.
-
-\item The initial lack of structures and processes negatively affects the performance when the company starts to grow.
-One of the primary objectives of startups is to expand their business into fast growing markets.
-In fact, if successful, the company will face a growth in terms of number of customers, employees and product functionalities.
-Then, the increased complexity arises the necessity of controlling the initially chaotic software development environment.
-Under delicate circumstances, where an increased market demand meets a poorly engineered product and process,  the development team need to start returning the accumulated technical debt.
-As reported by practitioners, introducing more structures, documentation and standard procedures, cause an initial drop-down in performance which severity depends on the speed of growth, the amount of technical debt and the capacity of the team to absorb it.
-
-\item Startups bring the first product to market in a very short-time and practitioners are satisfied with the software development strategy.
-The companies inquired in the case study were typically able to release the first beta of the product in a short time frame.
-Although practitioners acknowledge the initial negative consequences of the technical debt on subsequent performances, they confirm that it is better spending some effort in later stage to recover the debt than losing time in over-engineering the production initially with the risk of never achieving any significant growth.
-\end{compactitem}
-
-All of the above observations are extracted from the analysis of the model, which is systematically derived from empirical data of the case study.
-Indeed, with slightly different levels of adherence, the implication we presented are reflected in the behaviour of the great majority of companies in the sample.
-The results of this analysis, which has been focused on the software development strategy from the idea conception to the first release, indicate that startups are far from adopting standard or ad-hoc development methodologies, especially in this phase.
-The typical tendency is to focus on the team capability to implement and quickly iterate on a prototype, which is released very fast.
-Thus, in a context where even for the most \textit{lightweight} agile methodologies is hard to penetrate, we believe that it is premature to discuss issues related to SPI.
-Researchers should rather focus on the trade-off between development speed and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which
-appears to be the most important determinant for the future of software development in the company.
-
-%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
-\section{Discussion}
-\label{res:val}
-
-In this section we discuss the validity of the implications of this study, obtained by means of cross-methodological observations:
-
-\begin{compactitem}
-\item We perform a comparison between our framework and similar  models and frameworks.
-
-\item We validate the correctness of the GT procedure by attesting the relations with the GSM model's core category, evaluating if they are compliant with the empirical data we collected.
+\item Comparison between the GSM model and similar  models and frameworks.
+
+\item Comparison between the GSM model and the extant literature. We highlight the areas which have been neglected by existing studies, providing possible directions for future studies.
+\item Discussion of confounding factors which could interfere with the GSM model.
+\item Validation of the correctness of the GT procedure by attesting relations between the GSM categories, evaluating if they are compliant with the empirical data we collected.
 
@@
 In other words, even though ignoring SPI practices is advisable, not following SPI does not predict major adaptiveness of startups development to the market.
-But, to imply a major evolutionary product/market fit, certain
-conditions explained in GSM need to be considered.
+But, to imply a major evolutionary product/market fit, certain conditions
+explained in GSM need to be considered.
 
@@
 Those innovative solutions, if in the right market, are likely to belong to  the \textit{excitement} area, where satisfaction can hardly be negatively affected by faulty products.
-In fact, in the theoretical framework (see Section \ref{sect:theory:detailed}), minimal functionalities and less concern about quality aspects have low-impact on customers' satisfaction, supporting startup's ability to mainly focus on \textit{speed-up development}.
-Moreover operating in the \textit{excitement area} is fully compliant with an evolutionary approach and minimal functionalities since from the first product released
-to the market, it is unlikely to lose users because of high-quality demand.
+In fact, in the GSM model, minimal functionalities and less concern about quality aspects have low-impact on customers' satisfaction, supporting startup's ability to mainly focus on \textit{speed-up development}.
+Moreover operating in the \textit{excitement area} is fully compliant with an evolutionary approach and minimal functionalities since from the first product released to the market, it is unlikely to lose users because
+of high-quality demand.
 
@@
 However their correctness has been verified with a cross-analysis of the extant literature and by attesting the causal relations among codes and categories, through axial and selective coding processes.
+
+%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+\section{Discussion}
+\label{sect:theory:impl}
+
+In this section we present a list of the most relevant implications which emerge from the behaviour of early-stage startups, formally expressed in the GSM model, answering to our research question (\textit{What is the current state-of-practice related to software development strategy in early-stage startups?}) through its sub-questions.
+Although the startups we inquired in the case study were spread across different nations and market sectors, we have found several patterns that emerged very clearly:
+
+\begin{compactitem}
+
+\item The most urgent priority of software development is to shorten time-to-market.
+When developing a new product, which has not been attempted before, it is crucial for the company to release the product soon and observe the users reactions.
+The company's chances to survive, are strictly related to the ability to find the right product/market fit as quickly as possible.
+Developing software without releasing it for a long period of time, coming up with very complex solutions often is not a viable option.
+Business demands, investors interests, market pressures and the general uncertain conditions force startups to quickly iterate on the product by exploring and experimenting new functionalities progressively rolled-out to early-adopters.
+
+\item Startups do not apply any standard development methodology.
+When focused on building the first version of the product, startups do not observe any specific and standard development methodologies or processes.
+From an accurate and in-depth investigation into daily activities, it emerged that the number of  the observed development \textit{practices} is extremely small and never systematic (only little \textit{pair programming} sessions and rarely \textit{test driven development}).
+Despite the wide number of publications, which discuss the adoption of lightweight methodologies, we have found that especially in this first stage not only methodologies but even individual practices are basically neglected.
+
+\item The closest development approach undertaken by early-stage startups is the \textit{Lean startup methodology}.
+A highly evolutionary development approach, centered around the quick production of a functioning prototype and guided by customer feedback is a central aspect of the \textit{Lean startup methodology} pioneered by Ries \cite{Ries2011}.
+However, startups do not formally follow the cycle \textit{build-measure-learn} as it is proposed by the methodology, where they should set-up an experiment to \textit{test the riskiest element of the business idea}, nor make explicit use of the good practices suggested in the book.
+For example, when it comes to user feedback, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}.
+Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its
+subcategories.
+
+\item The greatest part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).
+Especially when bringing the first product to the market, the attitude of startups towards traditional engineering activities is strongly characterized by a general feeling of unnecessity.
+Every document, artifact or process, which does not produce a short-term tangible benefit to the software production is perceived as a waste of time.
+The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (CAT6), which is partially mitigated by different elements supporting the development.
+Project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-
+precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary
+bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
+Thus, by radically ignoring activities related to documentation, structures and processes, startups achieve a very high initial development speed in the first stage, but on the other hand they accumulate a technical debt that needs to be at least partially fulfilled when the company and the product grow.
+
+\item The first release of the product includes only a limited set of well suitable functionalities focused on user experience.
+The beta version is usually focused on a minimal and essential set of functionalities providing a reasonable user experience (UX).
+As confirmed by practitioners, the user must be able to accomplish a task on the product without major interruptions in the flow, which would cause frustration before being able to capture his attention.
+To realize UX in a short time frame, developers focus on a limited number of suitable functionalities, especially in the first version(s) of the product.
+Other product qualities are regarded as less important: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important,
+even more than time-to-market \cite{Offutt2002}.}.
+From a business perspective, one might argue that established companies might take advantage of the new idea and develop a better version of the product internally.
+Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}.
+Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process while focusing effort on acquiring larger shares of a ``mass market'' sector.
+Furthermore, thanks to cloud computing and more in general high availability of third party components, aspects such as performance, security and scalability can be achieved with a reasonable effort.
+Thus, given the general lack
+of quality concerns, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
+
+\item Engineering activities are supported by low-precision artifacts and collaborative tools integrated with the development environment.
+Startups take advantage of the simplicity of cutting-edge tools, which require a minimum learning and maintenance effort and provide a large payoff.
+Tools are used for a vast number of tasks: ticked-based boards to manage user stories; systems to collect user feedback; tools for tracing product metrics; and scripts for managing deployments.
+The most prominent families of tools, which support software development in startups are modern version control systems.
+They are used by practitioners not only to manage the code base and the repositories, but for many other purposes such as: task assignment; tracing responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.
+On the other hand, an excessive usage of  \textit{out-of-the-code} tools can negatively affect the development speed, as significantly reported by C11: \
+textit{``We realized we were losing too much time moving tickets around boards than actually working on stories.
+You need to find the balance between the time spent with tools and the amount of actual coding''}.
+This risk is often mitigated with an intense usage of offline manual tools: especially post-it notes and whiteboards are reported to be particularly effective to foster internal communication within the team.
+
+\item The founding team is the most important determinant of speed.
+More than techniques and methodologies, the features of the team are the essential factors to achieve high-development speed.
+Discussed in the theoretical category the \textit{team is the catalyst of development}, this result has been confirmed in multiple occasions both in the interviews and in the questionnaires.
+What is truly important for the startup, among other factors, is the extremely high-colocation which enables continuous informal and colloquial coordination and discussions.
+The initial (small) development team usually spends most of the hours of the week working closely to implement the product, often extending the conventional office hours.
+The large amount of communication, spontaneously held in informal places, has been proven to be an extremely effective element of several successful software projects \cite{Highsmith2000}.
+Hence, the vastly \textit{informal} development environment of startups acts as a catalyst of speed.
+Other important elements, which contribute in achieving such an initial high-speed, are: minimal set of functionalities with low concerns on product quality; evolutionary prototyping approach; smart use of advanced tools and frameworks; low precision artifacts; and the close colocation and capability of the team.
+
+\item The initial lack of structures and processes negatively affects the performance when the company starts to grow.
+One of the primary objectives of startups is to expand their business into fast growing markets.
+In fact, if successful, the company will face a growth in terms of number of customers, employees and product functionalities.
+Then, the increased complexity arises the necessity of controlling the initially chaotic software development environment.
+Under delicate circumstances, where an increased market demand meets a poorly engineered product and process,  the development team need to start returning the accumulated technical debt.
+As reported by practitioners, introducing more structures, documentation and standard procedures, cause an initial drop-down in performance which severity depends on the speed of growth, the amount of technical debt and the capacity of the team to absorb it.
+
+\item Startups bring the first product to market in a very short-time and practitioners are satisfied with the software development strategy.
+The companies inquired in the case study were typically able to release the first beta of the product in a short time frame.
+Although practitioners acknowledge the initial negative consequences of the technical debt on subsequent performances, they confirm that it is better spending some effort in later stage to recover the debt than losing time in over-engineering the production initially with the risk of never achieving any significant growth.
+\end{compactitem}
+
+All of the above observations are extracted from the analysis of the model, which is systematically derived from empirical data of the case study.
+Indeed, with slightly different levels of adherence, the implication we presented are reflected in the behaviour of the great majority of companies in the sample.
+The results of this analysis, which has been focused on the software development strategy from the idea conception to the first release, indicate that startups are far from adopting standard or ad-hoc development methodologies, especially in this phase.
+The typical tendency is to focus on the team capability to implement and quickly iterate on a prototype, which is released very fast.
+Thus, in a context where even for the most \textit{lightweight} agile methodologies is hard to penetrate, we believe that it is premature to discuss issues related to SPI.
+Researchers should rather focus on the trade-off between development speed and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which
+appears to be the most important determinant for the future of software development in the company.
 

### 2013-02-21 article/main.tex
MUN review and comments

@@
 \usepackage[neveradjust]{paralist} %compact lists
+\usepackage{amsthm}
+
+\newtheorem*{theory}{Theory}
 
@@
 The metaphoric neologism was originally introduced by Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the attention of SE researchers\footnote{To attest the fact that \textit{technical debt} is gaining traction among researchers, we mention two important contributions which characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt} organized by the Software Engineering Institute  and ICSE.}.
-The concept of \textit{technical debt} is well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea is that developers sometimes accept compromises in a system in one dimension (e.g., modularity) to meet an urgent demand in some other dimension (e.g., a deadline), and that such compromises incur a ``debt'
+The concept of \textit{technical debt} is well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea is that developers sometimes accept compromises in a system in one dimension (e.g., modularity) to meet an urgent demand in some other dimension (e.g., a deadline), and that such compromises incur a
+``debt'
 ' on which ``interest'' has to be paid and which the ``principal'' should be repaid at some point for the long-term health of the project''}.
@@
 %height of the figure, expanding however the width (use horizontal space since it is a two 
-%column figure).
-%Write "Data analysis" horizontally
+%column figure). Change also in figure:
+% - Write "Data Analysis" horizontally
+% - Change names of macro-phases to upper-case, i.e. Design and Execution, Data Collection, Data Analysis, etc.
+% - Add section number to macro-phase name, i.e. 3.1 Design and Execution, 3.2 Data Collection
+% - Change "Discussion and Validation" into "Theory Validation" as you name it in the text.
+% Also, add a section (3.5) that describes shortly what you did in "Theory Validation"
 
@@
 
-Following the \textit{grounded theory} principles we attempted to capture the most relevant aspects of software development from startup practitioners, letting a theory emerge from the interviews and adjusting the research hypotheses and questions as we proceeded through.
+Following the GT principles we attempted to capture the most relevant aspects of software development from startup practitioners, letting a theory emerge from the interviews and adjusting the research hypotheses and questions as we proceeded through.
 During these interviews we collected data related to engineering activities undertaken by startups in the time frame between the idea conception and the first open beta release of the product.
@@
 
-As suggested by Coleman, in view of the different versions of \textit{grounded theory} which emerged in the last years, researchers should indicate which \textit{``implementation''} of the theory is being used \cite{Coleman2007}.
-In fact, after the methodology was initially introduced, its original authors  had some divergences regarding how the theory should be formed by analyzing the data (Strauss and Corbin on one side, Glaser on the other).
+As suggested by Coleman, in view of the different versions of GT which emerged in the last years, researchers should indicate which \textit{``implementation''} of the theory is being used \cite{Coleman2007}.
+After the methodology was initially introduced, its original authors  had divergences regarding how the theory should be formed by analyzing the data (Strauss and Corbin on one side, Glaser on the other).
 Glaser advocated for a more \textit{puristic} approach, where the theoretical categories should \textit{``naturally emerge from the data''}, whilst Strauss and Corbin formulated an updated version of the methodology, which leaves to the researchers an higher degree of freedom.
@@
 
-In this research we addressed technical aspects related to software development in startups, exploring their operational dynamics.
+In this research we address technical aspects related to software development in startups, exploring their operational dynamics.
 In view of the lack of agreement on an unique definition of the word \textit{startup}, we delimited our initial contextual boundaries to newly created innovative and single-product software companies, in the time-frame that goes from the idea conception to the release of the first product in highly scalable markets.
@@
 
-The initial design of the case study is supported by the results of the \textit{systematic mapping study} (SMS) \cite{SMS} which contributed to define the initial questions for both questionnaires and semistructured interviews.
+The initial design of the case study is supported by the results of the \textit{Systematic Mapping Study} (SMS) \cite{SMS} which contributed to define the initial questions for both questionnaires and semi-structured interviews.
 The process depicted in Figure \ref{fig:gt:completemethodology} is evolutionary and affects the design at each new iteration.
-In \textit{data collection} the empirical results are integrated and stored in a single \textit{case study database} and subsequently processed in \textit{data analysis} to form the theoretical categories.
-At each iteration the new emergent theory is updated following a formal procedure (\textit{Theory generation}), and after verifying that the \textit{theoretical saturation}\footnote{The point at which executing more interviews wouldn't bring any additional value for constructing the theory.} of categories has been achieved, we finally proceeded to the \textit{theory validation} or performed another complete cycle.
+In \textit{Data Collection} the empirical results are integrated and stored in a single case study database and subsequently processed in \textit{Data Analysis} to form the theoretical categories.
+At each iteration the new emergent theory is updated following a formal procedure (\textit{Theory Generation}), and after verifying that \textit{Theoretical Saturation}\footnote{The point at which executing more interviews wouldn't bring any additional value for constructing the theory.} of categories has been achieved, we finally proceeded to \textit{Theory Validation}.
 
@@
 
-The detail process is thoroughly described in the following subsections which are structured according to the five macro phases depicted in bold in Figure \ref{fig:gt:completemethodology}.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
-\subsection{Design and execution}
-
-This case study started based on a Systematic Mapping Study (SMS) of the literature regarding software development in startups \cite{SMS}.
+The process details are described in the following subsections, structured according to the five macro phases depicted in bold in Figure \ref{fig:gt:completemethodology}.
+
+%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+\subsection{Design and Execution}
+\label{desex}
+
+This case study started based on a SMS of the literature regarding software development in startups \cite{SMS}.
 By means of this SMS we refined our research questions.
@@
 
-The sampling strategy took place in two distinct phases.
-At first we executed an initial \textit{convenience sampling} \cite{Dawson2009}, which led to the identification of eight companies.
-Then we included five additional startups during the theory formation process (\textit{theoretical sampling}) iteratively improving the sample according to the emerging theory.
-Additional information about the companies are reported in Table \ref{t_interviews-stats}.
+The sampling of candidate companies took place in two distinct phases.
+At first we executed an initial convenience sampling \cite{Dawson2009}, which led to the identification of eight companies.
+Then we included five additional startups during the theory formation process (theoretical sampling) iteratively improving the sample according to the emerging theory.
+The characteristics of the sampled companies are reported in Table \ref{t_interviews-stats}.
 
@@
 
-As shown in column \textit{Company age}, most of the companies were founded within the last three years with only one exception (C10, which age was 3 years and 6 months).
-As can be observed by looking at the column \textit{Founding team}, the companies have been founded by an average number of  3 members, which in majority are developers, as can be seen by looking at the column \textit{Initial developers}.
-Moreover, the number of \textit{Current employees} shows how, to different degrees, companies expanded the initial teams, being able to bring the product to the market in less than 6 months from the idea conception\footnote{C5 building time is extended to 1 year mainly because the product has been developed part-time by one person only.} (see \textit{First product building time}).
-Summarizing the above identified features, the sample of the case study is mainly composed by recently founded companies, working with web applications in different nations and market sectors, with very small development teams.
-Moreover, the growing team size and and the publicly available data suggest an healthy status of the businesses.
-
-We executed the case study online, supported by tools for video conferencing recording each complete session.
-Each interview has been executed only once per company with the interviewee's role of CEO or CTO.
-A step-by-step workflow have been followed, which allowed us to perform the interviews, collect additional material, prepare the customized follow-up questionnaire and iteratively adjust the \textit{interview package} artifacts\footnote{The \textit{interview package} and {follow-up questionnaire} has been released under MIT License \cite{MITLicense} and it is available for download on Github \cite{GitHubInterviewPackage}.
-We encourage anyone who has interests in pushing this work forward, to fork the repository and contribute to it.
-If used in a research context, please inform us in order to be able to track where and how the \textit{interview package} has been used.}.
-
-The follow-up questionnaires were designed to capture additional data, gather missing information and confirm interview results with triangulation.
-Questionnaires have been tailored to each startup, partially taking advantage of the repertory grid principles \cite{Edwards2009}.
+All companies, except C10, were founded within the last three years (2009-2012), %MUN: I've shortened the text and
+%added absolute numbers. Is 2009-2012 correct?
+by an average of 3 starting members, which were in majority developers.
+Moreover, the number of current employees shows how, to different degrees, companies expanded the initial teams.
+All companies, except C5, brought their first product to market in less than 6 months from the idea conception.
+The products consist of web-applications, launched in different nations and market sectors.
+The growing team size and the publicly available data suggest an healthy status of the businesses.
+
+We executed the case study online, supported by tools for video conferencing, recording each complete session.
+The interview subjects were the companies' CEOs or CTOs.
+We followed a step-by-step workflow, consisting of the actual interview, collection of additional material, preparation of the customized follow-up questionnaire and the iterative adjustment of the interview package artifacts\footnote{The interview package and follow-up questionnaire has been released under MIT License \cite{MITLicense} and it is available for download on Github \cite{GitHubInterviewPackage}.
+We encourage anyone interested in pushing this work forward, to fork the repository and contribute to it.
+If used in a research context, please inform us in order to be able to track where and how the interview package has been used.}.
+
+The follow-up questionnaires were designed to capture additional data, gather missing information and confirm interview results by triangulation.
+The questionnaires have been tailored to each startup, partially taking advantage of the repertory grid principles \cite{Edwards2009}.
 To obtain better chances of quickly obtaining responses, questionnaires have been sent to respondents immediately after transcribing the interview results.
 Two weeks after the conclusion of the last interview, we closed our follow-up questionnaires and collected the data.
+%MUN: I don't understand this: you
+% collected the data from questionnaires after ALL interviews have been finished, but on the other hand you 
+% said the overall process was iterative (data analysis and triangulation let themes emerge which in turn were used
+% in subsequent interviews).
 
@@
 \subsection{Data collection}
-The approach undertaken to collect the data, known as \textit{data triangulation}, integrates multiple data sources converging on the same phenomenon.
-After transcribing the interview, we extracted conceptual relations, and then we integrated them with the questionnaire's results.
+The data collection was designed to allow triangulation, which integrates multiple data sources converging on the same phenomenon.
+After transcribing the interview, we extracted conceptual relations, integrating them with the questionnaire's results.
 A well structured case-study database allowed us to easily retrieve and seek for information, assembling the evidence from the case study reports, as described also in \cite{Yin1994}.
 The database has been stored and constructed using the qualitative data analysis software package AtlasTI\footnote{Available online at \url{http://www.atlasti.com/} .} - one of the most suitable tools for GT \cite{Coleman2007}.
-
-Entering the field, we overlapped interviews with questionnaire results to adjust the data collection and take advantage of emergent themes and reveal possible inconsistencies.
+We overlapped interviews with questionnaire results to adjust the data collection and take advantage of emergent themes and reveal possible inconsistencies.
 The data was analyzed simultaneously and with flexibility in mind in such a way that adjustments were made according to the emerging findings.
@@
 \subsection{Data analysis}
-Before starting the analysis, a \textit{data ordering} procedure was necessary since interviews were spread across a multitude of topics.
-Therefore transcripts have been structured in thematic areas accordingly to different \textit{topic cards} used during the interviews.
+Before starting the analysis, a data ordering procedure was necessary since interviews were spread across a multitude of topics.
+Therefore transcripts have been structured in thematic areas accordingly to different topic cards used during the interviews.
 We proceeded horizontally between same thematic areas of transcripts to be able to identify a better number of similar concepts, rather than going through an entire transcript at one time.
 
-Once the data were ordered, we executed the process of \textit{coding} interviews, following the steps listed below:
+Once the data were ordered, we executed the process of coding interviews, following the steps listed below:
 
@@
 
-\item Labels were assigned to raw data, and a first low-level conceptualization was carried out using both \textit{in-vivo} and \textit{open coding} \cite{ColinRobson2009}.
+\item Labels were assigned to raw data, and a first low-level conceptualization was carried out using both in-vivo and open coding \cite{ColinRobson2009}.
 \item Concepts were grouped together into theoretical categories and subcategories.
-By means of \textit{axial coding} we first described the different relations between subcategories, and then relations between subcategories and categories.
-\item Categories were refined several times in the attempt to create different level of abstractions and adjusting concepts, aided by a simple knowledge management tool.
-\item Consistency among categories were validated by exploring and analyzing links  among subcategories by means of the \textit{selective coding}.
-\item The \textit{core category} - the one with the greatest explanatory power -  was identified by analyzing the causal relations between high-level categories.
+By means of axial coding we first described the different relations between subcategories, and then relations between subcategories and categories.
+\item Categories were refined several times in the attempt to create different levels of abstraction and adjusting concepts, aided by a simple knowledge management tool.
+\item Consistency among categories was validated by exploring and analyzing links among subcategories by means of selective coding.
+\item The core category - the one with the greatest explanatory power -  was identified by analyzing the causal relations between high-level categories.
 \end{compactitem}
 
-During data extraction we used the technique of \textit{in-vivo coding} since \textit{``direct quoting from the transcript give more expressive power to the data''} \cite{ColinRobson2009} combined with the more descriptive procedure of \textit{open coding}.
+During data extraction we used the technique of in-vivo coding since \textit{``direct quoting from the transcript gives more expressive power to the data''} \cite{ColinRobson2009} combined with the more descriptive procedure of open coding.
 Following the example of other grounded theories, developed in related areas such as Information Systems \cite{Orlikowski1993} and Software Process Improvement \cite{Coleman2006}, we performed the high-level conceptualization during creation of categories, in the process of refining axial and selective coding.
 
-As we were iterating through the interviews, we analyzed new data accordingly, by updating \textit{codes} and \textit{categories} on necessity, and taking notes in the form of  memos to adjust the new \textit{emerging theory}.
+As we were iterating through the interviews, we analyzed new data accordingly, by updating codes and categories on necessity, and taking notes in the form of memos to adjust the new emerging theory.
 To improve the speed of the coding process, we analyzed all the transcripts one part at-the-time: transcripts followed the semi-structured interview format, and this approach enhanced the chances of finding similar codes across different transcripts, allowing us to quickly iterate and update categories.
 
-Given the large number of codes, categories, properties, propositions and related questions that evolved from the analytical process \cite{Corbin1990}, an important activity that helped the coding process was \textit{memoing}.
+Given the large number of codes, categories, properties, propositions and related questions that evolved from the analytical process \cite{Corbin1990}, an important activity that helped the coding process was memoing.
 It constituted an important component involved in the formulation and revision of theory during the research process.
@@
 The first type of memo is related to the conceptual labeling of open codes, whilst the second type concerns axial and selective coding, and thus focusing on paradigm features.
-Finally, operational memos contain directions relating to the evolving \textit{emerging theory}.
-
-After the coding process the first representation of the \textit{experience map} identified in GT is constructed by means of a \textit{theoretical framework}.
-The theoretical framework is presented in the form of a network of categories and subcategories identified during the coding process.
-Precisely, categories and subcategories are linked together according to cause-effect relation \cite{Corbin1990}.
+Finally, operational memos contain directions relating to the evolving emerging theory.
+
+After the coding process the first representation of the experience map identified in GT is constructed by means of a theoretical framework.
+The theoretical framework is presented in the form of a network of categories and subcategories that are linked together according to cause-effect relationships \cite{Corbin1990}.
 During the formation of the theoretical framework the researchers operated by means of a  bottom-up approach.
-From empirical data and coding process, the framework was developed into two different levels: a detailed level representing the network of subcategories (identified mainly by \textit{axial coding} process), and high-level representing the network of main categories (identified mainly by \textit{selective coding} process).
+From the empirical data and coding process, the framework developed into two different levels: a detailed level representing the network of subcategories (identified mainly by the axial coding process), and a high-level representation of the main categories network (identified mainly by the selective coding process).
 
@@
 
-As mentioned above, emergent theories have been tested integrating the sample with additional companies selected following the principle of \textit{theoretical sampling} \cite{Yin1994}.
-
-The process of \textit{theory generation} took place at each iteration together with interview execution, where we systematically analyzed the \textit{theoretical framework} by means of the \textit{paradigm model} introduced by Corbin \cite{Corbin1990}.
-
+As mentioned in Section \ref{desex}, emergent theories have been tested by integrating additional companies into the sample, selected following the principle of theoretical sampling \cite{Yin1994}.
+
+The process of theory generation took place at each iteration together with interview execution, where we systematically analyzed the theoretical framework by means of the paradigm model introduced by Corbin \cite{Corbin1990}.
 The paradigm model is composed by:
@@
 \begin{compactitem}
-\item \textit{Causal conditions}, the events which lead to the occurrence of the \textit{phenomenon}, that is our core category.
-\item \textit{Context}, set of conditions in which the phenomenon can be extrapolated;
-\item \textit{Intervening conditions}, the broader set of condition that the phenomenon can be generalized.
-\item \textit{Action/interaction strategies}, the actions and responses that occur as the result of the phenomenon.
-\item \textit{Consequences}, specification of the outcomes, both intended and unintended of the actions and interaction strategies.
+\item \textit{Causal conditions}: the events which lead to the occurrence of the phenomenon, that is our core category
+\item \textit{Context}: set of conditions in which the phenomenon can be extrapolated
+\item \textit{Intervening conditions}: the broader set of conditions from that the phenomenon can be generalized
+\item \textit{Action/interaction strategies}: the actions and responses that occur as the result of the phenomenon
+\item \textit{Consequences}: specification of the outcomes, both intended and unintended of the actions and interaction strategies
 \end{compactitem}
 
-Within the limits of the critical bounding assumptions, the role of the generated theory is to explain, predict and understand the underling phenomenon.
+Within the limits of the critical bounding assumptions, the role of the generated theory is to explain, predict and understand the underlying phenomenon.
+
+\subsection{Theory Validation}
+%MUN: put here what I wrote in the comment in Section "Theory validation" (i.e. the section following the GSM)
 
@@
 The initial coding process (open and in-vivo) provided 630 unique codes \footnote{Available at \url{https://github.com/adv0r/BTH-Interview-Package}.}, grounded to 1295 citations in the transcripts, divided according to the thematic areas.
-
 The thematic areas were divided according to the development phases designed for the interview process.
-The complete list of  raw codes is presented in Appendix and constitutes the empirical foundation of the theoretical categories, which has been used to form the theoretical framework (Green-field startup model).
-
-In this section we present the GSM model to capture the underlying phenomenon of software development in early-stage startups.
-The GSM model is based on a systematic procedure which is grounded into empirical concepts obtained by conduction of the case-study.
-Starting from 630 raw \textit{codes} extrapolated from the interview transcripts, the framework is formed by 128 sub-categories, clustered in 35 \textit{groups}, and finally in 7 categories at the highest level of abstraction.
-
-By the means of a conceptual framework we want to provide explanations of the development strategies and engineering activities undertaken by startups.
-The knowledge, which we attempted to pack into a conceptual model, can be used by researchers and practitioners to improve their understanding of the phenomenon.
+The raw codes constitute the empirical foundation of the theoretical categories, which has been used to form the Green-field startup model (GSM).
+
+The GSM captures the underlying phenomenon of software development in early-stage startups.
+The framework is formed by 128 sub-categories, clustered in 35 groups, and finally in 7 categories at the highest level of abstraction.
+By the means of the GSM we provide explanations of the development strategies and engineering activities undertaken by startups.
 %--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
@@
 \label{res:gsm:frmov}
-Main concepts representing underlying phenomena have been grouped together to form high-level categories.
+The main concepts representing the underlying phenomenon have been grouped together to form high-level categories.
 Figure \ref{fig:gsm} shows the network of causal relationships (represented by arrows) between categories (represented by blocks).
-Each node is linked to subsequent nodes which are called successors, and preceding nodes which are called predecessors.
+Each node is linked to subsequent nodes which are called successors, and preceding nodes, called predecessors.
 The relations between them are denoted by the direction of the interconnecting arrows.
@@
 
-In the explanation of the GSM model we emphasize labels of  theoretical categories using the \textit{italic} font style.
+In forthcoming the explanation of the GSM we emphasize labels of theoretical categories using the \textit{italic} font style.
 For the main categories shown in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e.
@@
 The network is centered around the core category, \textit{speed-up development}, which is the most interconnected node in the framework reflecting the fact that \textit{``it is the one (category) with the greatest explanatory power''}  \cite{Strauss1998}.
-The meaning of the labels on the three arrows that reach the core category (\textit{efficiency}, \textit{effectiveness} and \textit{performance}) is discussed in detail in section \ref{valt}.
+The meaning of the labels on the three arrows that reach the core category (efficiency, effectiveness and performance) is discussed in detail in section \ref{valt}.
+%MUN: this refers to the validity threats section.
+Do you mean rather the "Theory Validation" section?
 
 A contextual condition which characterizes, to some extent, each and every startup, is the \textit{severe lack of resources}.
-In fact, the capabilities of an early-stage startups' to support its development activities are constrained by an extremely limited access to human, time and intellectual resources, eventually leading to a general absence of structures and processes.
-
-The \textit{severe lack of resources} forces the company to focus on implementing an essential set of functionalities and it is one of the main reasons why the \textit{product quality has low priority} respect to other more urgent needs\footnote{There are some exceptions where the quality aspects actually matter and such cases will be discussed later in \ref{valt}.}.
-At the same time, to be able to deal with such constraints, startups must depend on a small group of capable and motivated individuals.
+In fact, the capabilities of an early-stage startup to support its development activities are constrained by an extremely limited access to human, time and intellectual resources, eventually leading to a general absence of structures and processes.
+
+The \textit{severe lack of resources} forces the company to focus on implementing an essential set of functionalities and it is one of the main reasons why the \textit{product quality has low priority} respect to other more urgent needs\footnote{There are some exceptions where the quality aspects actually matter and such cases will be discussed in section \ref{valt} %MUN: this refers to the validity threats section.
+Do you mean rather the "Theory Validation" section?
+.}. At the same time, to be able to deal with such constraints, startups must depend on a small group of capable and motivated individuals.
 
@@
 In other words \textit{team is the catalyst of development}.
-With an essential and flexible workflow, which relies on tacit knowledge instead of formal documentation, startups are able to achieve very short time-to-market.
-However, each line of code, written without following structures and processes, contributes to grow the \textit{accumulated technical debt}, which is further increased by having almost non-existing specifications, a minimal project management  and a severe lack of automated tests.
+With an essential and flexible workflow, which relies on tacit knowledge instead of formal documentation, startups are able to achieve very short time-to-market cycles.
+However, each line of code, written without following structures and processes, contributes to grow the \textit{accumulated technical debt}, which is further increased by having almost non-existing specifications, a minimal project management and a lack of automated tests.
 
@@
 
-In the subsequent subsections we are going to explain in detail all the categories presented in Figure \ref{fig:gsm}, to conclude in subsection \ref{res:gsm:th} with the final theory.
-During the following explanations we are going to use identifiers of the companies presented in Table \ref{t_interviews-stats} (i.e.
+In the following subsections we are going to explain in detail all the categories presented in Figure \ref{fig:gsm}, and conclude in subsection \ref{res:gsm:th} with the final theory.
+In the explanations we use identifiers of the companies presented in Table \ref{t_interviews-stats} (i.e.
 Cx) to highlight statements recorded from the interviewees.
@@
 In startups \textit{developers have big responsibilities}.
-In fact, the \textit{limited human resources}, discussed in CAT7, causes the team-members to be active in every aspect of the development process, from the definition of functionalities to the final deployment.
-
-Early engineers in the founding team of a startups are typically \textit{multi-role and full-stack engineers}: they are full-stack for being able to tackle different problems at different level of the technology stack (\textit{generalists developers instead of specialists}) and multi-role since usually \textit{engineers are responsible for marketing/sales/development}.
-As remarked by  C11: \textit{``Instead of hiring gurus in one technology, startups should hire young developers, generalists, that know how to quickly learn new technologies, and quickly move among a huge variety of tasks.''}.
+In fact, the \textit{limited human resources}, discussed in CAT7, cause the team-members to be active in every aspect of the development process, from the definition of functionalities to the final deployment.
+
+Early engineers in the founding team of a startups are typically \textit{multi-role and full-stack engineers}: they are full-stack for being able to tackle different problems at different levels of the technology stack (\textit{generalists developers instead of specialists}) and multi-role since usually \textit{engineers are responsible for marketing/sales/development}.
+As remarked by  C11: \textit{``Instead of hiring gurus in one technology, startups should hire young developers, generalists, that know how to quickly learn new technologies, and quickly move among a huge variety of tasks.''}
 
@@
 Practitioners reported that keeping the development team small helps startups in being fast and flexible, as remarked by C8: \textit{``If you have more than 10 people, it is absolutely impossible to be fast''}.
-Then, also \textit{previous working experience} and \textit{knowing each other before starting the company} enforce the efficiency of activities by \textit{limiting th need of formalities between team-members}.
+Then, also \textit{previous working experience} and \textit{knowing each other before starting the company} enforce the efficiency of activities by \textit{limiting the need of formalities between team-members}.
 
@@
 Especially in startups, the ``hacking culture'' and a tendency to the ``just-do-it'' approach allow the team to quickly move from the formulation of a feature idea to its implementation.
-In this regard, C1 comments: \textit{``we had a hacker culture/environment, people hacking stuff without formally analyzing it, but breaking it down and finding a way around.
+In this regard, C1 comments: \textit{``We had a hacker culture/environment, people hacking stuff without formally analyzing it, but breaking it down and finding a way around.
 Ticket on trello and go.''}
@@
 A \textit{limited access to expertise} forces the team to rely mainly on their personal abilities, even though asking opinions to mentors is reported to be a viable practice to maintain feasible objectives.
-Furthermore \textit{teams work under constant pressure}, \textit{``up to 16 hours per day''}, mainly constrained by a tight \textit{time shortage}.
+Furthermore \textit{teams work under constant pressure}, \textit{``up to 16 hours per day''}, %MUN: add Cx to this statement since it is a citation?
+mainly constrained by a tight \textit{time shortage}.
 
@@
 The uncertainty is first of all in the team composition.
-In fact, since the teams are typically very small and the project knowledge is not documented, even a little change in their composition (e.g.
-a developer falls in ill) can have a very high impact on the overall product development.
+Since the teams are typically very small and the project knowledge is not documented, even a little change in their composition (e.g.
+a developer falls ill) can have a very high impact on the overall product development.
 Furthermore, startups operate in a continuous evolving environment in terms of competitors and targeted market sectors.
@@
 However, a special role in daily decisions is covered by user feedback
-and requests, which represent the main drivers to define the product features to provide in the short-term.
+and requests, which represent the main drivers for defining the product features in the short-term.
 
 To obtain fast user responses and quickly \textit{validate the product}, startups \textit{build a functioning prototype and iterate on it} over time.
-In fact, paraphrasing C4, \textit{``[\ldots] you should start with something that is really rough and then polish it, fix it and iterate.
+Paraphrasing C4, \textit{``[\ldots] you should start with something that is really rough and then polish it, fix it and iterate.
 We were under constant pressure.
@@
 
-The company focus on building a small set of functionalities to include in the first version, and \textit{progressively rolls-out to larger number of people} with very \textit{small iterations} (confirmed again by C4: \textit{``we deploy from 5 to 20 times a day''}).
-
-The scope of this evolutionary approach is to avoid wasting time in ``over-engineering the system'' and building complex functionalities that have not been tested on real users.
+The companies focus on building a small set of functionalities to include in the first version, and \textit{progressively roll-out to a larger number of people} with very \textit{small iterations} (confirmed again by C4: \textit{``we deploy from 5 to 20 times a day''}).
+
+The objective of this evolutionary approach is to avoid wasting time in ``over-engineering the system'' and building complex functionalities that have not been tested on real users.
 By releasing a very small number of good-enough functionalities (see CAT3) the startup can verify the suitability of the features and understand how to adjust the direction of product development towards actual users' needs.
-The first version of the product is typically a prototype which contains the basic rough functionalities developed with the least possible effort that can validate critical features which can undermine the startup's survival in the short term (recalling  Ries' definition of \textit{Minimum Viable Product} (MVP) \cite{Ries2011}).
-
-Thanks to a \textit{direct contact and observation of users}, \textit{automated feedback collection} and \textit{analysis of product metrics}.
-The product metrics mainly refer to the UX of the product to find any kind of frictions that might reduce the satisfaction level of the final users.
-Startups attempt to \textit{find what is valuable for customers} with activities in agreement with Blank's \textit{Customer validation} stage \cite{Blank2005} .
+The first version of the product is typically a prototype which contains the basic functionalities developed with the least possible effort that can validate critical features, enabling the startup's survival in the short term (recalling  Ries' definition of \textit{Minimum Viable Product} (MVP) \cite{Ries2011}).
+
+Supported by \textit{direct contact and observation of users}, \textit{automated feedback collection} and \textit{analysis of product metrics}, startups attempt to \textit{find what is valuable for customers} with activities in agreement with Blank's \textit{Customer validation} stage \cite{Blank2005}.
 
@@
 \label{res:gsm:cat3}
-The main interests of software startups, related to the product, are mainly concentrated on building a  \textit{limited number of suitable functionalities} rather than fulfilling restrictive non-functional requirements.
+The interests of software startups, related to the product, are mainly concentrated on building a  \textit{limited number of suitable functionalities} rather than fulfilling restrictive non-functional requirements.
 This strategy allows them to quickly release simple products with less need of preliminary architectural studies.
 
-Exploring quality aspects considered during the development process, the only important concerns are expressed in favour of the user experience\footnote{According to the ISO 9241-210 (Ergonomics of human-system interaction), UX is defined as \textit{``a person's perceptions and responses that result from the use or anticipated use of a product, system or service''.} Even though this definition leaves so much to interpretation, we can refer to it as presented in the glossary.}, in terms of \textit{ease of use, attractiveness of the UI} and most of all, \textit{smooth user-flow without interruptions}.
+Exploring quality aspects considered during the development process, the only important concerns are expressed towards the user experience\footnote{According to the ISO 9241-210 (Ergonomics of human-system interaction), UX is defined as \textit{``a person's perceptions and responses that result from the use or anticipated use of a product, system or service''.} Even though this definition leaves so much to interpretation, we can refer to it as presented in the glossary.}, in terms of \textit{ease of use, attractiveness of the UI} and most of all, \textit{smooth user-flow without interruptions}.
 The fact that \textit{UX is the only important quality} is remarked by C11: \textit{``When a user needs to think too much on what action should be done next, he will just close the application without returning''} and confirmed by C3: \textit{``If the product works, but it is not usable, it doesn't work''.}
@@
 Product metrics are typically web-based \textit{statistical hypothesis testing}, such as A/B testing.
+%MUN: add reference or explain shortly what A/B testing is
 
@@
 \item The \textit{efficiency emerges after using the product}, letting engineers avoid wasting time in excessive improvements of not-validated functionalities.
-In fact, the level of efficiency can be optimized after attesting the effectiveness of the minimal set of functionalities, obtained according to CAT2 and to the concept of MVP.
+The level of efficiency can be optimized after attesting the effectiveness of the minimal set of functionalities, obtained according to CAT2 and to the concept of MVP.
 \item The \textit{product should be reasonably ready-to-scale} to be able to accommodate a potential growth of the user-base\footnote{Note that startups tackle fast growing markets which are particularly subject to sudden user growths}.
@@
 \footnote{Detailed discussion of the impact of innovative products on the user satisfaction is presented in Appendix according to the Kano model.}.
+%MUN: do you use the Kano model in Theory Validation? Change text to refer to the correct section.
 \end{compactitem}
@@
 \label{res:gsm:cat1}
-As remarked above in the model, \textit{speed-up development} represents the core category of the GSM model.
+As remarked initially, \textit{speed-up development} represents the core category of the GSM.
 Firmly grounded as the primary objective of startups, it shows the most important characteristic of developing software in the early stage.
@@
 
-Additionally possible planning activities are refrained by the aim to shorten time-to-market, as reported by C8: \textit{``Speed was the essence so we didn't plan out too many details''}.
+Additionally, possible planning activities are refrained by the aim to shorten time-to-market, as reported by C8: \textit{``Speed was the essence so we didn't plan out too many details''}.
 To deal with such unpredictability, startups prefer to take decisions as fast as possible, mainly by means of informal and frequent verbal discussions.
@@
 
-Another strategy, which supports startups in quickly delivering products, is the \textit{externalization of complexity on third party solutions} by: making use of third party components (COTS) and open source solutions (both for product components either development tools and libraries); taking advantage of  external services; and deploying the product on external infrastructures, for the sake of presenting a \textit{product reasonably ready to scale} for a possible future growth.
+Another strategy, which supports startups in quickly delivering products, is the \textit{externalization of complexity on third party solutions} by: making use of third party components (COTS) and open source solutions (both for product components, and development tools and libraries); taking advantage of  external services; and deploying the product on external infrastructures, for the sake of delivering a \textit{product reasonably ready to scale} for a possible future growth.
 
@@
 Finally the constant pressure under which the company regularly operates, leads the team to often \textit{work overtime to meet deadlines}.
-But as reported by practitioners, such way of working can be an effective strategy only in the short term since can lead to poorly maintainable code and developers burnouts in the long run.
+But as reported by practitioners, such way of working can be an effective strategy only in the short term since it can lead to poorly maintainable code and developer burnouts in the long run.
 %--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
@@
 In particular, the use of well-known standards, frameworks and conventions limits the need of formal UML \cite{UML} diagrams and documentation, and provide a minimum level of maintainability costs.
-Additionally, since quality aspects are not a main concern (see \textit{limited number of functionalities} and {efficiency emerges after using the product} in CAT3) having a well-structured architecture remains a secondary priority, and design is conducted only when strictly necessary.
+Additionally, since quality aspects are not a main concern (see \textit{limited number of functionalities} and \textit{efficiency emerges after using the product} in CAT3) having a well-structured architecture remains a secondary priority, and design is conducted only when strictly necessary.
 
@@
 Then, the \textit{limited number of functionalities} and the {use of standard/known technologies} with a \textit{simple workflow} are the main reasons for
-not establishing a heavy project management process since the project and technologies can be autonomously managed by the development team (with characteristics of: co-location, self-organization and very small size) .
-
-Finally, one of the categories, which most contributes in growing the \textit{accumulated technical debt}, is the substantial  use of informal and verbal communication channel on daily basis.
+not establishing a heavy project management process since the project and technologies can be autonomously managed by the development team (with characteristics of: co-location, self-organization and very small size).
+%MUN: do you mean small team size?
+
+Finally, one of the categories, which most contributes in growing the \textit{accumulated technical debt}, is the substantial  use of informal and verbal communication channels on a daily basis.
 The high co-location and the fast paced development approach increase the volume of \textit{tacit knowledge} and severe lack of any kind of documentation.
@@
 The lack of attention given in the first phases to engineering activities allows startups to ship code extremely quickly (see CAT5).
-However, if successful, the initial product becomes more and more complex overtime, the number of user increases and the company size starts to grow.
+However, if successful, the initial product becomes more and more complex over time, the number of users increases and the company size starts to grow.
 Under these circumstances the need of controlling the initial chaos forces the development team to return the \textit{accumulated technical debt} instead of focusing on new users' requests.
@@
 When the number of users increases, customers become more quality demanding and also scalability issues start to arise.
-
 Usually, \textit{company and user size grow} when new business events occur, such as: a \textit{new round of funding}, a possible \textit{acquisition}, the release of a \textit{competing product on the market}, or when the project is \textit{open for the first public release}.
-As a consequence, the \textit{increasing number of users} creates growing number of requests and expectations.
+As a consequence, the \textit{increasing number of users} creates a growing number of requests and expectations.
 Therefore, whereas the project lacks of minimal processes, suddenly, \textit{the current team is not able to manage increased complexity} of new functionalities to implement and maintain the codebase.
 
-Subsequently, practitioners start considering the need of project management activities, also in view of \textit{hiring new staff members}, as discussed by C13: \textit{``(project management) is strictly necessary if you radically change the team or when team grows.
+Subsequently, practitioners start considering the need of project management activities, also in view of \textit{hiring new staff members}, as discussed by C13: \textit{``(project management) is strictly necessary if you radically change the team or when the team grows.
 The informal communication and lack of documentation slow down the process afterwards''}.
@@
 
-Another factor that slows down performance is caused by the fact that \textit{portion of codes needs to be rewritten} and \textit{substantial refactoring of the codebase} is required by the increasing product's demand .
+Another factor that slows down performance is caused by the fact that \textit{portion of codes needs to be rewritten} and \textit{substantial refactoring of the codebase} is required by increasing product demands.
 
@@
 We had to do a big refactoring of the codebase, moving it from custom php to Django, normalizing the model and making it stick with the business strategy.
-I had the code in different php server communicating via JSON, some engineering horror.
+I had the code in different php servers communicating via JSON, some engineering horror.
 Now  that we are fixing it, it's really painful.
 We had to trash some code.
-However I don't regret that I didn't make this choice sooner, it was the only way''} .
+However I don't regret that I didn't make this choice sooner, it was the only way''}.
 
 The \textit{fear of changing a product, which is working}, arises when product complexity increases.
-In fact, the changes to the codebase, to support bug fixing, become highly interrelated with other functionalities and difficult to manage because the product is poorly engineered (see CAT5).
+The changes to the codebase, to support bug fixing, become highly interrelated with other functionalities and difficult to manage because the product is poorly engineered (see CAT5).
 Therefore, the fear arises from changing a validated product that might bring changes to the user responses.
@@
 \label{res:gsm:th}
-In order to explain and understand the development strategies in early-stage software startups we construct the theory generated and supported by the GSM model presented above.
-As discussed in section \ref{valt}, we applied the \textit{paradigm model} to form the final theory, introduced by Corbin \cite{Corbin1990}, here presented:\\
-
+In order to explain and understand the development strategies in early-stage software startups we construct the theory generated and supported by the above presented GSM:
+
+\begin{theory}
 Focusing on limited number of suitable functionalities, and adopting partial and rapid evolutionary development approaches, early-stage software startups operate at high development speed, aided by skilled and highly co-located developers.
@@
 Nevertheless, by speeding-up the development process, they accumulate technical debt, causing an initial and temporary drop-down in performance before setting off for further growth.
-\\
-
-The theory above is composed by different elements reflecting the \textit{paradigm model}. Here we structure the list of elements presented in the theory:
+\end{theory}
+
+As discussed in section \ref{valt}%MUN: this refers to the validity threats section. Do you mean rather the "Theory Validation" section?
+, we formed the theory by applying the paradigm model, introduced by Corbin \cite{Corbin1990}. The theory is composed by different elements reflecting the paradigm model:
 
@@
 
-\item \textit{Causal conditions} are represented by three main conceptual categories: \textit{product quality has low priority}, \textit{evolutionary approach} and \textit{ team is the catalyst of development}.
+\item \textit{Causal conditions} are represented by three main conceptual categories: \textit{product quality has low priority}, \textit{evolutionary approach} and \textit{team is the catalyst of development}.
 \item \textit{Phenomenon} is represented by the core category \textit{speed-up development}.
@@
 %--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
-\section{Validation}
+\section{Theory Validation}
 \label{res:val}
 
-In this section we discuss the validity of the GSM model by means of cross-methodological observations:
+In this section we discuss the validity of the GSM by means of cross-methodological observations:
+%MUN: expand this discussion a bit (motivation why you do this like you do it) and move to Section 3.5.
+Then write here just that you perform theory validation as discussed in Section 3.5 (and remove the bullet list below).
 
@@
 
-To validate the generalization of the theory, in this section we describe conceptualizations defined from the framework that are in support of a previous model developed by Coleman in \cite{Coleman2007,Coleman2008a, Coleman2008}.
-We refer to Coleman's framework since it is the only author which has conducted a similar research in startups' context, even though with a different focus.
-The author investigated factors in software development that hinder initiatives of software process improvement (SPI) in much more mature companies.
-
-In this subsection we illustrate the most important aspects considered by Coleman, examining similar and contrast factors in face of the GSM model.
-We will consider only those categories strictly related to the core category, which relations, as described in GT methodology, generate complete explanatory power of the generated theory.
-
-Coleman's framework aims to highlight how managers consider two distinct kind of processes: \textit{essentials} and \textit{non-essentials}.
-The \textit{essential} processes are the most closely linked to the product development such as requirements gathering, design and testing.
-The \textit{non-essential} processes are those that might be omitted such as planning, estimating and staging meetings.
-In particular, he discusses how practices are routinely removed: \textit{``With most methodologies and approaches, very few stick to the letter of them and they are always adapted, so we adapted ours to the way we wanted it to work for us, for our own size and scale''.}
-
-The GSM model explores the same challenges in software startups, where the act of tailoring processes leads developers to adopt only minimal practices, which are most suitable for the startups context.
-Furthermore, CTOs and CEOs' background has a great impact on the speed of the development process as described in CAT4.
+To validate the generalization of the theory, we describe conceptualizations defined from the GSM that are in support of a previous model developed by Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}.
+We refer to Coleman's work since only he has, to the best of our knowledge, conducted a similar research in startups' context, even though with a different focus.
+Coleman investigated factors in software development that hinder initiatives of software process improvement (SPI) in much more mature companies.
+%MUN: so he did not do research in a startup context?
+You need to clarify what you mean by "more mature companies"...
+I assume you still mean startups, but startups at a later stage, i.e.
+after first product release.
+
+In this subsection we illustrate the most important aspects considered by Coleman, examining similar and contrasting factors with respect to the GSM.
+We consider only those categories strictly related to the core category, which relations, as described in GT methodology, generate complete explanatory power of the generated theory.
+
+Coleman's framework aims to highlight how managers consider two distinct kinds of processes: \textit{essentials} and \textit{non-essentials}.
+The \textit{essential} processes are the most closely linked to product development, such as requirements gathering, design and testing.
+The \textit{non-essential} processes are those that might be omitted, such as planning, estimating and staging meetings.
+In particular, he discusses how practices are routinely removed: \textit{``With most methodologies and approaches, very few stick to the letter of them and they are always adapted, so we adapted ours to the way we wanted it to work for us, for our own size and scale''.}%MUN: add ref to Coleman's paper where he states this
+
+The GSM explores the same challenges in software startups, where the act of tailoring processes leads developers to adopt only minimal practices, which are most suitable for the startups context.
+Furthermore, CTOs' and CEOs' background has a great impact on the speed of the development process as described in CAT4.
 This is in accordance to Coleman's framework, which presents the background of founders and development managers as main factors that affect the management style of the development process.
 
-Differently from Coleman's, the GSM model doesn't present how \textit{market requirements} can affect the conduction of processes.
-In this regard, Coleman describes how the more the definition of requirements are predictable the more well-defined workflows are established.
+Differently from Coleman's framework, the GSM doesn't present how \textit{market requirements} can affect the conduction of processes.
+In this regard, Coleman describes how the more requirements definitions are predictable the more well-defined workflows are established.
 This particular aspect is considered as a confounding factor\footnote{The identification of confounding depends on the context in which the study is conducted and on the background knowledge of the researchers \cite{Pearl2011}.}  in our study since we were not able to ground this concept in the data.
@@
 But what characterizes Coleman's network is the \textit{cost of process} (core category) and all the factors that in management contributed to the lack of software process improvements (SPI).
-
 The \textit{cost of process} represents the  lack of formal and prescriptive workflows in development, mainly conducted by means of verbal \textit{communication} limiting heavy \textit{documentation} and \textit{bureaucracy}.
 
-The author reports practitioners' perception that \textit{documentation} alone would not have ensured a complete shared understanding of project requirements.
+Coleman reports the practitioners' perception that \textit{documentation} alone would not have ensured a complete shared understanding of project requirements.
 Moreover, defined processes are perceived by managers as additional items with a negative impact on the \textit{Creativity} and \textit{Flexibility} of the development team.
-This is in accordance with the generated theory, which bases the reasons of adopting evolutionary and low-precision engineering elements to the flexibility and reactivity attributes of the development process.
-
-At this regard we can paraphrase a comment recorded by Coleman during an interview: \textit{``When we set up we had more supervisory and managerial roles in that group than we have now and we had to scale that back which has made things a lot more flexible.
+This is in accordance with our generated theory, which bases the reasons of adopting evolutionary and low-precision engineering elements in the flexibility and reactivity attributes of the development process in startups.
+
+In this regard we can paraphrase a comment recorded by Coleman during an interview: \textit{``When we set up we had more supervisory and managerial roles in that group than we have now and we had to scale that back which has made things a lot more flexible.
 I do think you have to be nimble, quick and capable being responsive in our position.
@@
 Verbal \textit{communication}, lack of heavy \textit{documentation} and \textit{bureaucracy} can easily be associated to the \textit{accumulated technical debt} category (see CAT5).
-According to Coleman's framework, they describes how startups experience the lack of main engineering activities and documentation.
+According to Coleman's framework, they describe how startups experience the lack of main engineering activities and documentation.
 Additionally, the \textit{speed-up development} category expresses the same concept of \textit{flexibility} and \textit{process erosion} since they have direct relation to the subcategory of \textit{keep a simple and informal workflow}, as discussed in CAT1.
 
-As also reported in the GSM model, the definition of a \textit{``minimum process''} is not a matter of poor knowledge and training, but rather it is the necessity of
- operating with solutions that let the company move faster.
-\textit{``One-size-fits-all''} solutions have always found difficult to penetrate small software organizations \cite{Staples2007}.
-In fact, also when startups began establishing any SPI process they experienced \textit{process erosions}, which resolved to a barely sufficient workflow to satisfy the organizational business needs.
-
-Then, software startups privilege the use of agile principles in support of \textit{creativity} and \textit{flexibility} instead of SPI methods, whereby processes need to be predictable and repeatable.
-Nimble and ad-hoc solutions prevent the use of heavy bureaucracy and fast communication strategies, even though the accumulated tacit knowledge is hard to manage and to transfer to new hires as also discussed by the category \textit{CAT6}.
-
-Also in \cite{Coleman2008}, the author describes how the \textit{``management approach''} is oriented towards \textit{``embrace and empower''} solutions in contrast with \textit{``command and control''} style, where there is an evidence of trust in development staff to carry out tasks with less direct supervision, greater delegation of responsibilities and a more generally consensual environment.
-Nevertheless, software development manager and founders have still impact on management style and indirectly on software development process.
+As also reported in the GSM, the definition of a \textit{``minimum process''} is not a matter of poor knowledge and training, but rather it is the necessity of operating with solutions that let the company move faster.
+\textit{``One-size-fits-all''} solutions have always found difficulty in penetrating small software organizations \cite{Staples2007}.
+When startups began establishing any SPI process, they experienced \textit{process erosions}%MUN: add ref (Coleman?).
+You should add, for each category you use from Coleman, add the reference to the paper.
+This makes it clear that the category comes from him (and is not one of our categories)
+, which resolved to a barely sufficient workflow to satisfy the organizational business needs.
+
+Then, software startups favor the use of agile principles in support of \textit{creativity} and \textit{flexibility} instead of SPI methods, whereby processes need to be predictable and repeatable.
+Nimble and ad-hoc solutions prevent the use of heavy bureaucracy and fast communication strategies%MUN: seems to contradict each other: "prevent the use of heavy bureaucracy AND fast communication strategies
+, even though the accumulated tacit knowledge is hard to manage and to transfer to new hires as also discussed in category \textit{CAT6}.
+
+Coleman describes how the \textit{``management approach''} is oriented towards \textit{``embrace and empower''} solutions in contrast with \textit{``command and control''} style, where there is an evidence for trust in the  development staff to carry out tasks with less direct supervision, greater delegation of responsibilities and a more generally consensual environment \cite{Coleman2008}.
+Nevertheless, software development managers and founders have still impact on management style and indirectly on the software development process.
 In the case of early-stage startups, founders are mainly software development managers as CEOs/CTOs and  technical practitioners at the same time.
-As Coleman identified the influence of the founders' and managers' background over the software development process, also GSM identifies how the  CEOs and CTOs background shapes the high-level strategies adopted in developing the initial product.
+As Coleman identified the influence of the founders' and managers' background on the software development process, the GSM similarly identifies that the CEOs and CTOs background shapes the high-level strategies adopted in developing the initial product.
 Notwithstanding, team members remain self-organized, able to intervene in all the
@@
 
-In conclusion, by the comparison of the GSM model with Coleman's framework we have found that despite Coleman's study contributes to define which factors hinder the introduction of SPI, his results are coherent with the theory generated within our research.
+In conclusion, by the comparison of the GSM with Coleman's framework, we have found that despite Coleman's study contributes to define which factors hinder the introduction of SPI%MUN: add: "of SPI in star I ditups"?
+, his results are coherent with the theory generated within our research%MUN: add "our research, focused at explaining software development in general in startups." ?
+.
 In this regard, a wide variability of processes is a key factor for startups.
 In contrast with some SPI processes, such as \textit{Six Sigma} which objective is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, startups seem to follow the \textit{Fischer's Fundamental Theorem of natural selection} \cite{Fisher}.
-In fact, moving towards flexible and variable processes (easy to adapt to the uncertain conditions), increase odds of ``natural selection'' only when some ``restrictive conditions'' are met.
-In other words, even though ignoring SPI practices is advisable, not following SPI does not predict major adaptiveness of startups development to the market.
-But, to imply a major evolutionary product/market fit, certain conditions
-explained in GSM need to be considered.
-
-Also Baskerville in \cite{Internet} refers to SPI as a development approach typically effective only in large-scale, long-term development efforts with stable and disciplined processes.
+Moving towards flexible and variable processes (easy to adapt to the uncertain conditions), increase the odds of ``natural selection'' only when some ``restrictive conditions'' are met.
+In other words, even though ignoring SPI practices is advisable%MUN: under which circumstances?
+And the following part,I did not fully understand.
+Try to clarify.
+ not following SPI does not predict major adaptiveness of startups development to the market.
+But, to imply a major evolutionary product/market fit, certain conditions explained in GSM need to be considered.
+
+%MUN: As I wrote already in the previous draft, you need to clarify at the beginning of this subsection which frameworks you are going to compare GSM to.
+Now, you say that Coleman is the only author who did research, comparable to yours, into startups.
+Now there comes Baskerville and Brooks.
+Write at the beginning of this section, shortly, which frameworks/authors you compare GSM to and why, then create subsections for each.
+Also Baskerville \cite{Internet} refers to SPI as a development approach typically effective only in large-scale, long-term development efforts with stable and disciplined processes.
 By applying a grounded theory analysis in 10 companies he found that major causal factors that influence development are:
@@
 
-As discussed in CAT7, 1) can be mapped to \textit{time shortage} and 3) to \textit{limited access to expertise}, whilst  2) can refer to {uncertain conditions make long-term planning not viable} explained in CAT2.
-Even though with different research focus and context of study - understand how and why Internet-speed software development (i.e.
-an agile method oriented using daily builds aimed at developing a product with high speed) differs from traditional software development - Baskerville revealed similar causal factors.
+As discussed in CAT7, 1) can be mapped to \textit{time shortage} and 3) to \textit{limited access to expertise}, whilst  2) can refer to \textit{uncertain conditions make long-term planning not viable}, explained in CAT2.
+Even though with different research focus and context of study % - understand how and why Internet-speed software development (i.e.
+an agile method oriented using daily builds aimed at developing a product with high speed) differs from traditional software development - MUN: put this half-sentence where you introduce Baskerville's work
+ Baskerville revealed similar causal factors as the GSM.
 
@@
 
-With wider research focus, another study, developed in 1986 by Brooks \cite{BrooksJr1987}, discussed what challenges are involved in constructing software products.
-In his study the author  divides difficulties in development into essence (inherent to the nature of the software), and accidents (difficulties attending software production but which are not inherent).
-In other words essence concerns the hard part of building a software through activities such as specification, design, testing.
+With a wider research focus, Brooks \cite{BrooksJr1987} discussed what challenges are involved in constructing software products.
+Brooks divides difficulties in development into essence (inherent to the nature of the software), and accidents (difficulties attending software production but which are not inherent).
+In other words essence, concerns the hard part of building a software through activities such as specification, design, testing.
 Accidents refer to the labor of representing the software or testing its representation.
 
-The author claims that the major effort applied by engineers in the last decades was dedicated towards accidents problems, trying to exploit new strategies to enhance software performance, reliability and simplicity of development, such as the introduction of high-level languages for programming.
+Brooks claims that the major effort applied by engineers in the last decades was dedicated towards accidents problems, trying to exploit new strategies to enhance software performance, reliability and simplicity of development, such as the introduction of high-level languages for programming.
 Despite the great achievements in improving development performance, the \textit{essence} property of the software remained unaltered.
@@
 
-In this section we report the evaluation of our theory in contrast with main relevant contribution of studies retrieved and discussed within the SMS in \cite{SMS}, discussing the categories of the GSM model.
-
-The comparison has been executed by identifying issues related to the theoretical categories in the retrieved relevant studies.
-We illustrate how the selected studies contributed to the GSM model validation as following: for each retrieved article we present the main results, mapping them to categories in accordance to the GSM model.
+In this subsection we report the evaluation of our theory in contrast with the main contributions of the studies retrieved and discussed within the SMS \cite{SMS}.
+For each retrieved article we present the main results, mapping them to categories in accordance to the GSM model.
 Table \ref{tab:an:literature_comp} shows the main contributions of the 37 identified studies, where an 'X' is placed in correspondence to the theoretical category they considered.
-The table has been sorted by computing how many categories of the GSM model have been considered by the study (column \textit{Count}).
-Therefore in the highest part of the table there are the studies which somehow discuss the higher number of concepts identified in the early-stage startups framework.
+The table is sorted by the number of GSM categories covered by the studies.
 
@@
 Sutton (2000) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Sutton2000} \\
-Kajko (2008) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Kajko-Mattsson2008} \\
+Kajko-Mattson (2008) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Kajko-Mattsson2008} \\
 Crowne (2002) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Crowne2002} \\
@@
 Taipale (2010) & X     & X     &       & X     & X     &       &       & 4     & \cite{Taipale2010} \\
-Silva (2005) & X     & X     &       & X     & X     &       &       & 4     & \cite{Silva2005} \\
+da Silva (2005) & X     & X     &       & X     & X     &       &       & 4     & \cite{Silva2005} \\
 Mirel (2000) & X     &       & X     & X     & X     &       &       & 4     & \cite{Mirel2000} \\
@@
 Chorev (2006) &       &       &       & X     &       &       &       & 1     & \cite{Chorev2006} \\
-
+%MUN: add a row which shows the sum of each column. This shows how much categories are covered overall
 \hline
@@
 
-According to the GSM model, only 7 studies, out of 37 (18.91\%), discuss aspects related to the overall software development process in early stage startups.
-The remaining part of the studies is only partially covering the phenomenon described by the framework.
-Looking at the results, we were able to extrapolate information which confirmed the GSM categories.
-
-We can see how the majority of the retrieved studies (29) mention issues related to \textit{speed-up development} which confirms, in fact, the importance of the \textit{core} category.
-On the other hand we can observe that only a limited number of studies mentions results related to CAT3, CAT5 and CAT6.
-This suggests the direction of possible future primary studies towards: the technical debt and its future consequences on performance; aspects related to functionality and quality of the first version of the product.
-
-These results confirm the relevance of development teams as widely discussed in the SE literature, here mapping 26 studies to CAT4 (\textit{team catalyst for development}).
-The importance of people has been widely discussed in other studies in SE advocating for the need of  \textit{empowering people}, which is the element with the greatest impact on software development.
-Furthermore, being humans \textit{non-linear variables}, they are unlikely to follow repeatable prescriptive methodologies.
-As stated by Highsmith, \textit{``A drawback of each and every methodology is to expect people to behave consistently over time, when is clearly not like that.''} \cite{Highsmith2000}.
-For more detailed discussion about the importance of people the reader is referred to some prominent authors: Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999}, Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph \cite{Adolph2011} and Cockburn \cite{people-first-order}.
-Especially in early-stage startups, where
-the whole company overlaps with the development team, factors related to people are without any doubt the most important ones (confirmed with the empirical data of  our case study.
-
-In addition to the theoretical framework proposed by Coleman previously discussed, in the next part of this subsection we present similarities and contrasts with other relevant studies selected with the SMS.
-
-Starting from the concept of \textit{severe lack of resources} many authors describe startups facing with inescapable constraints defined by engineering/business concerns and constant time-pressure \cite{Camel1994a, Cugola98softwareprocesses:}.
+Only 7 studies out of 37, discuss aspects related to the overall software development process in early stage startups.
+The remaining part of the studies is only partially covering the phenomenon described by the GSM.
+Looking at the results%MUN: which results?
+The papers' main contribution or...?
+, we were able to extrapolate information which confirmed the GSM categories.
+
+We can see how the majority of the retrieved studies (29) mention issues related to \textit{speed-up development} (CAT1) which confirms the importance of the core category.
+On the other hand, we can observe that only a limited number of studies mentions results related to \textit{product quality has low priority} (CAT3), \textit{accumulated technical debt} (CAT5), and \textit{initial grow hinders performance} (CAT6), suggesting directions of possible future primary studies.
+
+These results confirm the relevance of development teams as widely discussed in the SE literature, here mapping 26 studies to \textit{team is the catalyst of development} (CAT4).
+The importance of people has been widely discussed in other studies in SE (Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999}, Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph \cite{Adolph2011} and Cockburn \cite{people-first-order}), advocating for the need of empowering people, which is the element with the greatest impact on software development.
+Humans, being non-linear variables%MUN: add ref
+, are unlikely to follow repeatable prescriptive methodologies.
+As stated by Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and every methodology is to expect people to behave consistently over time, when is clearly not like that''}.
+Especially in early-stage startups, where the whole company overlaps with the development team, the GSM confirms the importance of factors related to people.
+
+%MUN: Think of structuring this long subsection. Have a subsubsection for each CAT, i.e. 5.2.1 Severe lack of resources (CAT7)
+%I've added to each category you mention the corresponding CATx since that identifier you use in table 2. I did not check if you discuss each and every category. Do you? If yes, ok. If not, why not?
+Starting from the category \textit{severe lack of resources} (CAT7), many authors describe startups facing with inescapable constraints defined by engineering/business concerns and constant time-pressure \cite{Camel1994a, Cugola98softwareprocesses:}.
 Because of their young and inexperienced working conditions, startups present limited resources in terms of management and strategic alliances \cite{Sutton2000}.
 
-Crowne in \cite{Crowne2002} identifies four main stages of a startup's lifecycle: startup, stabilization, growth and maturity.
+Crowne \cite{Crowne2002} identifies four main stages of a startup's lifecycle: startup, stabilization, growth and maturity.
 Inexperienced practitioners operating in a chaotic environment characterizes the first phase, where there is no strategic plan for developing the product.
-But, moving towards the \textit{stabilization} phase, the product starts to be unreliable and requirements unmanageable.
-Consequently, when startups move to the \textit{growth} phase, they initiate structuring and controlling processes, which are incrementally integrated in the development environment.
-The \textit{maturity} represents the moment when the company is ready to introduce process improvement initiatives.
-This study profoundly corresponds to different conceptualization of the GSM model.
+But, moving towards the stabilization phase, the product starts to be unreliable and requirements unmanageable.
+Consequently, when startups move to the growth phase, they initiate structuring and controlling processes, which are incrementally integrated in the development environment.
+The maturity phase represents the moment when the company is ready to introduce process improvement initiatives.
+Crowne's study profoundly corresponds to different conceptualizations in the GSM.
 Starting from the minimal and low-precision engineering elements, startups start to grow the accumulated technical debt which requires fast integration of more structured and standard processes overtime.
-Issues related to low-precision engineering practices during companies' growth are discussed in an experience paper by Ambler \cite{Ambler2002} where he reports the case studies of two growing internet startups approaching an IPO.
+Issues related to low-precision engineering practices during companies' growth are discussed in an experience paper by Ambler \cite{Ambler2002}.
+He reports on two case studies of growing internet startups approaching an IPO.%MUN:explain acronym
 Despite the two companies were facing business and team growth, attitudes against the introduction of processes to structure and model the system were evident.
-The need to re-architect and redevelop the codebase lead the researchers to suggest a tailored version of the RUP development framework.
+The need to re-engineer the codebase lead the researchers to suggest a tailored version of the RUP%MUN: add reference
+development framework.
 Nevertheless, even with the aid of CASE tools\footnote{CASE tools are a class of software that automates many of the activities involved in various engineering activities.
 See \url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, startups failed in practicing engineering activities in view of eccesive lack of flexibility.
-In fact, developers fast moved to simpler tools such as the whiteboard, where they were able to provide input and insights, question assumptions, share
+Developers fast moved to simpler tools such as using a whiteboard, where they were able to provide input and insights, question assumptions, share
 past experiences, and even talk about potential implementation strategies.
-In general the development remained highly iterative, where project teams started to have a little modeling and testing only when necessary.
+%MUN: Candidate for removal since details can be found in Ambler anyway: "In general the development remained highly iterative, where project teams started to have a little modeling and testing only when necessary.
 They didn't take the serial approach of completely developing and accepting the requirements model, analysis model and so on.
 This let them react swiftly and adapt rapidly according to changes into their highly competitive markets.
-The researchers report \textit{``modeling was streamlined''}, that is modeling what was needed and go straightforward to coding, avoiding analysis paralysis.
-
-The use of whiteboards in working environment is in support of the \textit{use of well-integrated and simple tools}.
-In both the startups studied by Ambler, documentation was minimal, finding the ``sweet spot'' to have just barely enough tools and artifacts to meet their needs and no more.
+The researchers report \textit{``modeling was streamlined''}, that is modeling what was needed and go straightforward to coding, avoiding analysis paralysis."
+
+The use of whiteboards in working environments supports \textit{use of well-integrated and simple tools}.
+In both the startups studied by Ambler, documentation was minimal, finding the ``sweet spot'' to have just barely enough tools and artifacts to meet their needs.
 Generally, the use of \textit{keep a simple and informal workflow} endorses the use of low-precision engineering practices which lead to speed-up development initially, but with negative effects on the accumulated technical debt of a minimal project management and consequently on the final company growth.
@@
 Kajko-Mattsson et al.
-investigated a Swedish software startup company involved in mobile applications \cite{Kajko-Mattsson2008}, reporting the following issues: a lack of requirements gathering process; release cycles and their length were not defined; the releases and their scope were not planned; lack of control over the change requests; lack of process control in terms of  absence of any documentation to track the status and progress of the process; defective releases in terms of lack of testing; poor communication since the communication inside the company was informal and not documented at all.
-In the same study they designed a process improvement model, based on three main phases: \textit{Evaluate}, \textit{Plan}  and \textit{Change}.
-
-During the evaluation phase Kajko-Mattsson et al.
-tried to establish the development workflow status within the company, and the authors document a status of chaos where any initiatives of quantitative methodology assessment would have been a waste of time.
-Therefore, they applied a qualitative methodology approach to define a model of the release management process.
-In fact, they moved towards the plan phase where the researchers decided to develop a model to manage releases, introducing a process composed by: release scope preparation, release planning, release development, system testing, acceptance testing and release deployment.
-Despite the effort spent from the researcher to establish a development workflow, practices were not fully conducted even though some marginal benefits have been reported: planned major release from four to six months; planned minor releases from two to four weeks; encompassing urgent corrective releases from one to five days.
+\cite{Kajko-Mattsson2008} investigated a Swedish software startup company involved in mobile applications, reporting the following issues: a lack of requirements gathering process; release cycles and their length were not defined; the releases and their scope were not planned; lack of control over the change requests; lack of process control in terms of  absence of any documentation to track the status and progress of the process; defective releases in terms of lack of testing; poor communication since the communication inside the company was informal and not documented at all.
+In the same study they designed a process improvement model, based on three main phases: Evaluate, Plan and Change.
+
+%MUN: the whole paragraph needs to be summarized to its essence. Details can be read in the referred paper
+In the evaluation phase, the status of the development workflow within the company  is analyzed.
+The researchers documented a state of chaos where any initiatives of quantitative methodology assessment would have been a waste of time.
+Therefore, they applied a qualitative methodology to define a model of the release management process.
+They moved towards the plan phase where the researchers decided to develop a model to manage releases, introducing a process composed by: release scope preparation, release planning, release development, system testing, acceptance testing and release deployment.
+Despite the effort spent by the researchers to establish a development workflow, practices were not fully conducted even though some marginal benefits have been reported: planned major release from four to six months; planned minor releases from two to four weeks; encompassing urgent corrective releases from one to five days.
 Regarding the previously mentioned problems, only
 partial improvements have been achieved since the major obstacle was to motivate employees to change their habits.
-With perfect hindsight the researchers claim that understanding and adapting solutions by solving one problem at a time would have increased benefits.
-Then, despite the reported results, start establishing simple and informal workflow aided the organization to move from the stage of knowing nothing to knowing at least something about their development process.
-
-The integration of minimal project management during the development activities can foster the control over development productivity and control the \textit{accumulated technical debt}, but still with particular attention to the development workflow status within the company, designing specific and tailored solutions.
-
-In fact, prescribed development models such as SPI in startups are basically ignored by startups, as discussed in \cite{Zettel2001}.
-To overcome the need of structure and control to accelerate the development process, in his study, Zettel developed a \textit{``LIghtweight Process for E-business software development''} (LIPE).
-This lightweight development method integrates XP approach with ideas from the areas of software measurement for project control and process improvement as a compromise between ad-hoc and more rigorous approach.
-LIPE, as argued from the author, provides scalability of the development process with a certain degree of flexibility, omitting parts of prescribed practices.
-Nevertheless, we couldn't retrieve any empirical evaluation of the study.
-
-In \cite{Deakins2005}, the author designed a new spiral methodology evolved during fieldwork with a developer team that had limited resources managing innovative product in volatile e-commerce environments.
-He proposes fast evolutionary approach experimentation with flexible processes obtaining customer's feedback as soon as possible for continuous product improvement.
-Even in this case the model is compliant with the nature of a random, ad-hoc and visionary product development.
+With perfect hindsight the researchers claim that understanding and adapting solutions by solving one problem at a time would have increased the benefits.
+Then, despite the reported results, starting the establishment of a simple and informal workflow aided the organization to move from the stage of knowing nothing to knowing at least something about their development process.
+
+The integration of minimal project management during the development activities can foster the control over development productivity and control the \textit{accumulated technical debt} (CAT5), but still with particular attention to the development workflow status within the company, designing specific and tailored solutions.
+
+Prescribed development models such as SPI in startups are basically ignored by startups, as discussed in \cite{Zettel2001}.%MUN: <-- rephrase, not clear to me
+To overcome the need of structure and to accelerate the development process, Zettel developed a \textit{``LIghtweight Process for E-business software development''} (LIPE).
+This lightweight development method integrates the XP approach with ideas from areas of software measurement for project control and process improvement as a compromise between an ad-hoc and a more rigorous approach.
+Zettel argues that LIPE provides scalability of the development process with a certain degree of flexibility, omitting parts of prescribed practices.
+Nevertheless, we couldn't identify any empirical evaluation of LIPE that supports its suitability in a startup context.
+
+Deakins \cite{Deakins2005} designed a new spiral methodology, evolved during fieldwork with a developer team that had limited resources managing an innovative product in volatile e-commerce environments.
+He proposes fast evolutionary approach experimentation with flexible processes obtaining customer's feedback as soon as possible for continuous product improvement.%MUN: <-- rephrase, sentence not parsable
+Even in this case the model%MUN: which model?
+is compliant with the nature of a random, ad-hoc and visionary product development.
 In fact, evolutionary development approaches are typically more suitable  for web applications, as confirmed in \cite{Deshpande2001}.
-Then the practices are oriented towards rapid and iterative design with complete freedom to the developer team, enabled to operate anywhere within the organization, according also to the category \textit{team is the catalyst of development}.
-
-Carmel reports in \cite{Camel1994a} the need of flexible and rapid development solutions to shorten time-to-market, but even more important, the need of high-skilled team developers, which heavily impact on \textit{speeding-up development}.
-In fact, the author suggests that entrepreneurs need to look for a well formed, skilled core development team and not just a set of product ideas and features.
-Moreover the team must be empowered with full-stack and self-organization settings with the flexibility of minimal bureaucracy.
-A studied company stated: \textit{``we didn't need weeks or months of detailed modeling and documentation , but rather modeling the architecture a little and then either exploring strategies by providing users with code or simply starting work on the actual software itself''}.
-
-Eventually to focus on simple solutions, two other studies \cite{Wall2001, Bean2005} sugges to \textit{``reuse the wheel, instead of reinventing it''}, taking advantage of open source software whenever it is possible, building systems quickly and effectively as a result.
-The author conclude stating that even though the principles in leading-edge development processes are the same principles of yesteryear, the way in which the fundamentals are applied has changed from prescriptive to agile solutions with less concern to quality aspects, also confirmed in \cite{Mater2000}.
-
+Then, the practices are oriented towards rapid and iterative design with complete freedom to the developer team, enabled to operate anywhere within the organization, according also to the category \textit{team is the catalyst of development}.%MUN: <-- the whole paragraph is not that clear to me.
+Try to start with the category (team is catalyst), then write about Deakins spiral model
+
+Carmel \cite{Camel1994a} reports on the need of flexible and rapid development solutions to shorten time-to-market, but even more important, the need of high-skilled team developers, which has an impact on \textit{speed-up development} (CAT1).
+He suggests that entrepreneurs need to look for a well formed, skilled core development team and not just a set of product ideas and features.
+Moreover the team must be empowered with full-stack and self-organization settings, paired with the flexibility of minimal bureaucracy.
+A studied company stated: \textit{``we didn't need weeks or months of detailed modeling and documentation, but rather modeling the architecture a little and then either exploring strategies by providing users with code or simply starting work on the actual software itself''}.
+
+%MUN: To which category does the following paragraph belong?
+Eventually to focus on simple solutions%MUN: solutions on what?
+, two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the wheel, instead of reinventing it''}, taking advantage of open source software whenever it is possible, building systems quickly and effectively as a result.
+Even though the principles in leading-edge development processes are the same principles of yesteryear, the way in which the fundamentals are applied has changed from prescriptive to agile solutions, with less concern on quality aspects, also confirmed by Mater \cite{Mater2000}.
+
+%MUN: When you subdivide this sections, as proposed above, finish with a  subsubsection "Summary", summarizing the supported you have found in literature for your theoretical categories.
+Nothing extensive, 3-4 sentences to conclude this part of the validation, similar as below but a bit more explanatory (what does "compliant with the behavioural model of the phenomenon" mean?)
 Thus, by summarizing the comparison, we have found that most of the contributions of the studies are compliant with the behavioural model of the phenomenon we provided.
@@
 
-The purpose of this subsection is to identify which confounding factors need to be taken into consideration when evaluating the GSM model.
-The confounding factors are those variables which and have not been considered in the model and might interfere with the theoretical framework positively or negatively \cite{ColinRobson2009}.
-Even though there are no systematic procedure to identify confounding factors, we report those that have been explicitly identified by the literature review, that are related to: \textit{creativity, innovation, market type and application type}.
-
-A prominent contribution investigating creativity and innovation in startups, is presented in \cite{Heitlager2007}.
-The study reports how development approaches oriented to the product in the early-stage life-cycle, called \textit{fluid} have main impact to innovative solutions initially.
-After a while the need of structuring and controlling the growing company size (along with \textit{transition} and \textit{specific} phases) arises the need of project management activities and software processes for long-term scalability issues.
-
+The purpose of this subsection is to identify which confounding factors need to be taken into consideration when evaluating the GSM.
+Confounding factors are those variables which and have not been considered in the model and might interfere with the theoretical framework positively or negatively \cite{ColinRobson2009}.
+We report those factors that have been explicitly identified by the literature review%MUN: do you mean the SMS?
+Then refer to it.
+I remember you mentioned CF in the results (GSM), maybe also refer to the respective sections
+, related to: creativity and innovation, market requirements and application type, and developer experience.
+
+%MUN: Again, structure this into subsubsections according to the above factors. I did it, but check if you agree...
+\subsubsection{Creativity and innovation} 
+Heitlager et al.
+\cite{Heitlager2007} contributed to the investigation of creativity and innovation in startups.
+%MUN: the rest of this paragraph is too detailed.
+I'd remove this "The study reports how development approaches oriented to the product in the early-stage life-cycle, called \textit{fluid} have main impact to innovative solutions initially.
+After a while the need of structuring and controlling the growing company size (along with \textit{transition} and \textit{specific} phases) arises the need of project management activities and software processes for long-term scalability issues."
+
+%MUN: Please reformulate the first sentence, I cannot parse it
 Based on the model presented in \cite{Heitlager2007}, it identifies the lifecycle stages of a startup (\textit{fluid}, \textit{transition} and \textit{specific}), corresponding for each stage a development approach (product-oriented and process-oriented) the level of innovation achieved.
-The study reports how product-oriented development, in contrast with forced and mature process imposition, gives degree of freedom to the development team, enhancing \textit{creativity} of developers and augmenting the \textit{innovation} capability of the company in the early-stage.
-
-The empirical model we presented has been obtained from the transcripts of interviews, where the theme of innovation only seldom emerged.
+The study reports how product-oriented development, in contrast with forced and mature process imposition, provides a degree of freedom to the development team, that enhances the creativity of developers and augments the innovation capability of the company in the early-stage.
+
+In the interviews we conducted, and on which the GSM is grounded, the theme of innovation only seldom emerged.
 However, if innovative and creative factors were considered in the model, the integration of results would have been straightforward, since the evolution of this model is totally compliant with our assumption of CAT5 and CAT6.
-From product-oriented development in the \textit{fluid} phase to the process-oriented development in the \textit{specific} phase, the \textit{innovation} confounding factor seems to behave initially the same as the \textit{speed-up development}.
-Nevertheless, we were unable to ground these evidences in the GT data even though they might present strict correlation with the choice of an early product-development orientation with low-precision engineering elements.
-
-Other two important factors, that are not grounded in our empirical data, are related to: \textit{a.} the requirements expected by a specific market sector (also called \textit{market requirements} in \cite{Coleman2007}); and  \textit{b.} the application type as reported in \cite{Sutton2000, Coleman2008, Coleman2008a}.
-The main impact of those confounding factors is related to the adoption of flexible and reactive solutions for the development process.
-In particular they refer to the necessity of fulfillment of quality concerns that goes beyond scalability and UX defined by the framework.
+%MUN: not clear to me what you mean by "evolution of this model" and what "compliant with our assumption" means.
+Also, don't use only CATx in this context, but cite the category name AND CATx
+From product-oriented development in the \textit{fluid} phase to the process-oriented development in the \textit{specific} phase, the innovation confounding factor seems to behave initially the same as the \textit{speed-up development}.%MUN: <-- please provide some motivation...
+not obvious
+ Nevertheless, we were unable to ground these evidences in the GT data even though they might present strict correlation%MUN: "present strict correlation" What does that mean, I don't understand
+ with the choice of an early product-development orientation with low-precision engineering elements.
+
+\subsubsection{Market requirements and application type} 
+Market requirements \cite{Coleman2007} and application type \cite{Sutton2000, Coleman2008, Coleman2008a} are two further potential confounding factors to the GSM.
+Their main impact is related to the adoption of flexible and reactive solutions for the development process.
+In particular, they refer to the necessity of fulfillment of quality concerns that goes beyond scalability and UX defined by the GSM.
 Especially when requirements are rigidly imposed or the application domain is well-known, providing low-quality products to final users might determine the failure of a startup.
 
-For example security in critical applications such as safety-critical systems present quality concerns, which must be granted.
-But also minor risky application domains such as e-commerce systems have high-impact factors for obtaining the first customer reference, as discussed in \cite{Deakins2005,Kim2005,Silva2005}.
-This characterizes those applications, whose part of functionalities are already successfully implemented in other software systems or whose failure might report serious financing or social damages \cite{Bass2003}.Yet, limiting our context to early-stage software startups, which typically develop first cutting-edge product in beta versions, partially mitigates the high-quality demands of users.
-
-To understand how to obtain processes in accordance to a set of customer needs, correlating the proper levels of customers' satisfaction, we introduce the Kano model, described in \cite{Ullah2011} and extensively studied in product development.
+For example, safety-critical systems possess quality concerns that must be considered.
+But also minor risky application domains such as e-commerce systems have high-impact factors for obtaining the first customer reference,%MUN: <-- cannot parse it.
+Remove the first part and write: "Even e-commerce systems...".
+Then rephrase the second part of the sentence to make it clearer.
+ as discussed in \cite{Deakins2005,Kim2005,Silva2005}. This%MUN: what is "this"?
+ characterizes those applications, whose part of functionalities are already successfully implemented in other software systems or whose failure might inflict serious financial or social damages \cite{Bass2003}.
+Yet, limiting our context to early-stage software startups, which typically develop first cutting-edge products in beta versions, partially mitigates the high quality demands of users.
+
+%MUN: The following 4 paragraphs explain how you use the Kano model to position startups w.r.t.
+customer expectations and quality concerns.
+It's ok, but it adds little value to the validation of GSM.
+I'd remove it.
+% To understand how to obtain processes in accordance to a set of customer needs, correlating the proper levels of customers' satisfaction, we introduce the Kano model, described in \cite{Ullah2011} and extensively studied in product development.
 As shown in the Kano model, users preferences can be classified into four categories: \textit{basic}, \textit{excitement}, \textit{performance} and \textit{indifference} areas.
-
-The \textit{basic} area represents the space where customers are dissatisfied if those \textit{must-be} functionalities are not met.
+% 
+% The \textit{basic} area represents the space where customers are dissatisfied if those \textit{must-be} functionalities are not met.
 This is the case of similar functionalities already implemented in the current market with well-known solutions.
@@
 This is the case of extremely innovative product development.
-
-Well-known domains such as e-commerce applications or other \textit{``quality critical products''} belong to the \textit{basic} area, where if a quality concern such as security in bank transfer services is missing satisfaction of customers will be profoundly affected.
+% 
+% Well-known domains such as e-commerce applications or other \textit{``quality critical products''} belong to the \textit{basic} area, where if a quality concern such as security in bank transfer services is missing satisfaction of customers will be profoundly affected.
 But as discussed, in early-stage startups that typically develop first beta products, developers usually deal with cutting-edge technologies.
@@
 Moreover operating in the \textit{excitement area} is fully compliant with an evolutionary approach and minimal functionalities since from the first product released to the market, it is unlikely to lose users because
-of high-quality demand.
-
-Nevertheless, this aspect profoundly undermines the generalizability of the theoretical model according to the theory, in particular \textit{CAT3}.
+% of high-quality demand.
+% 
+% Nevertheless, this aspect profoundly undermines the generalizability of the theoretical model according to the theory, in particular \textit{CAT3}.
 Therefore, future studies within wider startups' context are advisable to extrapolate insights from those situations where quality concerns cannot be postponed or externalized.
 
+\subsubsection{Developer experience} 
+%MUN: Ups, a confounding factor you did not mention initially. Add it...
 Finally, also the experience of team-members is considered as an important confounding factor.
 Since only little evidence has been gathered during the interviews, ``lack of experience'' was not considered as a sub-category in \textit{severe lack of resources}.
-Notwithstanding, according to \cite{Crowne2002}, startups at the beginning rely on clever, but inexperienced developers.
-In fact, having team-members with deep experience is a ``double-edge sword'', as discussed in \cite{Yoffie1999}.
-From one hand experience might quickly provide structure and maturity to the development process, from the other it might cause huge challenges in managing human resources in view of the fact that self-confident overachievers will almost inevitably clash.
-Consequently the team management requires higher control and coordination activities that inevitably hinder the flexibility of the development environment at the beginning, that is essential in early-stage startups.
-Accordingly, also Coleman in \cite{Coleman2007,
+Nevertheless, startups rely at the beginning on clever, but inexperienced developers \cite{Crowne2002}.
+Having team-members with deep experience is a ``double-edge sword'' \cite{Yoffie1999}.
+On one hand, experience might quickly provide structure and maturity to the development process, on the other it might cause challenges in managing self-confident overachievers that almost inevitably clash.
+Consequently, team management requires higher control and coordination activities that inevitably hinder the flexibility of the development environment, that is essential in early-stage startups.
+Accordingly, also Coleman \cite{Coleman2007,
 Coleman2008a,Coleman2008} states that the operating strategies towards \textit{``minimum processes''} in startups is not a matter of poor knowledge and training, but rather it is the necessity of operating with solutions that let the company move faster.
 
-Nevertheless, since the amount of the results did not allow ``the lack of experience'' to be in the GSM model (in view of the fact that not enough data were identified \footnote{A category is formed if the codes are representative of more than the 70\% of the analyzed transcripts.}), this confounding factor requires further investigation.
+%MUN: removed below paragraph since it does not add much
+% Nevertheless, since the amount of the results did not allow ``the lack of experience'' to be in the GSM model (in view of the fact that not enough data were identified \footnote{A category is formed if the codes are representative of more than the 70\% of the analyzed transcripts.}), this confounding factor requires further investigation.
 
@@
 
-In this subsection we validate the most important causal relations presented in our model by attesting their correctness  with a quantification of empirical trends observed in the data.
-In fact we identified some trends and recurrent patterns which we tried to synthesize in this section by thoroughly analyzing the complete set of data obtained from the case study and the GSM model.
-We illustrate our observations aided by simple statistical methods, which provides to researchers an additional tool to validate our model supported by further companies' data.
-
-As previously discussed when presenting the GSM model, the most essential feature that startups aim to achieve is high speed.
-
-The \textit{core category} (\textit{speed-up development}) is influenced by: the degree of quality concerns (contributes to the \textit{efficiency}); the undertaken development approach (provides \textit{effectiveness}); and finally the intrinsic characteristics of the team (guarantee \textit{performance}).
-
+%MUN: explain in the research method section (new section on how you performed validation), exactly how you did static validation and what statistical methods you used on what data, then refer to that section from here (no need to repeat again here).
+In this subsection we validate the most important causal relations presented in our model by attesting their correctness  with a quantification of empirical trends observed in the interview data.
+We identify trends and recurrent patterns by thoroughly analyzing and synthesizing the complete set of data obtained from the case study on which the GSM is built upon.
+We illustrate our observations aided by simple statistical methods, providing researchers an additional tool, supported by further companies' data, to validate our model.
+
+We argued in section \ref{res:gsm:frmov}, that the most essential feature startups aim to achieve is high speed.
+As illustrated in Figure \ref{fig:gsm}, the core category (\textit{speed-up development}) is influenced by: the degree of quality concerns (contributes to the \textit{efficiency}); the undertaken development approach (provides \textit{effectiveness}); and finally the intrinsic characteristics of the team (guarantee \textit{performance}).
+
+%MUN: for the paragraphs below, you need to explain what and why you do this.
+What do you mean by "we grounded in the interview transcripts the fact that those relations were characterized in terms of performances since time-to-market was the main concern." Not clear to me.
 Before starting the evaluation of the impact of the three categories on \textit{speed-up development}, we grounded in the interview transcripts the fact that those relations were characterized in terms of performances since time-to-market was the main concern (as confirmed in \cite{SMS}):
@@
 
-\item The team enhanced the overall \textit{performance} of the software development.
-In fact, recalling that startups deal with severe lack of resources, the team members are the company's main assets.
+\item The team enhances the overall \textit{performance} of software development.
+Recalling that startups deal with severe lack of resources, the team members are the company's main assets.
 They develop software with no need of formalities and structures, left to the team's own capabilities to efficiently develop the product.
-Moreover, the effectiveness is granted by continuous advices given by mentors and by a management style oriented towards self-organization, full-stack, and multi-role settings\footnote{Note that the importance of a skilled team for development performance is expressed in many studies such as \cite{surviving-os-cockburn, Highsmith2000}.} (described in CAT4 and confirmed, among others, by Sutton in \cite{Sutton2000}).
-
-\item The evolutionary approach enhanced the \textit{effectiveness} in implementing the right functionalities.
-In fact, this approach improves the company's capabilities to adjust the trajectory of product development (in agreement to the concept of \textit{pivoting} discussed in \cite{Ries2011}).
-\item Since product quality has low priority, the development was focused only on implementing a limited number of suitable functionalities, enhancing its \textit{efficiency}.
-In fact, an essential prototype (MVP) does not require to comply with heavy quality constraints, enabling the team to quickly implement functionalities, ready to be validated by final users.
+Moreover, the effectiveness is granted by continuous advice given by mentors and by a management style oriented towards self-organization, full-stack, and multi-role settings.
+The importance of a skilled team for development performance is expressed in many studies such as \cite{surviving-os-cockburn, Highsmith2000}, described in CAT4 and confirmed by Sutton \cite{Sutton2000}.
+\item The evolutionary approach enhances the \textit{effectiveness} in implementing the correct functionalities.
+This approach improves the company's capabilities to adjust the trajectory of product development (in agreement to the concept of pivoting discussed by Ries \cite{Ries2011}).
+\item Since product quality has low priority, the development is focused only on implementing a limited number of suitable functionalities, enhancing its \textit{efficiency}.
+An essential prototype (MVP) does not require to comply with heavy quality constraints, enabling the team to quickly implement functionalities, ready to be validated by final users.
 \end{compactitem}
@@
 
-Accordingly, by analysing the data of the 13 companies collected with the empirical study, we observed two distinct trends which further empower the above illustrated relations:
+Accordingly, by analysing the data of the 13 companies, we observed two distinct trends which further empower the above illustrated relations:
+%MUN: You need to be more precise with terminology.
+Above you state that the two items below are "trends which further empower relations", but then, further below, you write "to assess these hypotheses".
+Observed trends != Hypotheses.
+This is also why you need to be extremely explicit and precise when explaining what the a) aim, b) method, and c) outcome/result of this section is.
+I'm aware that it is difficult to boil down what you wrote in the thesis to this limited space, but I think it is doable (and relevant for supporting validity of GSM)
 
 \begin{compactenum}
-\item \textit{The more} a capable team\footnote{The team capability is evaluated according to the characteristics presented in CAT4.} with few quality constraints uses an evolutionary approach, \textit{the faster} the company releases the product.
-\item \textit{The faster} the company releases the product, \textit{the larger} the degree of \textit{accumulated technical debt}\footnote{The debt here is considered regardless of context-specific debt mitigation factors and tactics.}.
+\item The more a capable \footnote{The team capability is evaluated according to the characteristics presented in CAT4.} team, with few quality constraints, uses an evolutionary approach, the faster the company releases the product.
+\item The faster the company releases the product, the larger the degree of \textit{accumulated technical debt}\footnote{The debt here is considered regardless of context-specific debt mitigation factors and tactics.}.
 \end{compactenum}
@@
 The complete statistical procedure and provided rubrics to quantify the measures are illustrated and discussed in Appendix.
-
+%MUN: make sure it is clear that the PROCEDURE material is available in the ONLINE appendix and refer to it. Still, you need to report the actual RESULTS of you statistical analysis here... 
 The validation has been performed on the most critical  causal relationships of the model.
@@
 %--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
-\section{Discussion}
+\section{Discussion} %MUN:Implications of the GSM
 \label{sect:theory:impl}
+
+%MUN: There are two issues with this section: it is too long and it basically reiterates (in a concise manner) what you have said in the preceding 12 pages.
+The real value of this section (suggested by my proposed name above), would be if you can state in 0.5 (MAX 1 page), what are the implications of your results on THEORY (or state-of-art, research) and PRACTICE (state-of-practice, people who want to embark a startup).
+I'd propose you have a short intro here, then have two subsections: State-of-art, State-of-practice where you state the implications of GSM (or use a better subdivision if you come up with one, or use none if you don't exceed 0.5 pages).
+These implications need to logically follow from what you have said before (and the data you collected in the interviews and, when necessary, by reference to the SMS).
+Don't have a long bullet list, but structure your arguments in a logical flow.
+This is the section in which you have to show, based on your previous rigorous and well done analysis, the essence / gained knowledge, indicating what are the possible next steps for research in this area.
+You can be creative/visionary here as long as your arguments follow from the previous analysis.
 
@@
 %--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
-\section{Validity threats}
+\section{Threats to Validity}
 \label{valt}
+%MUN: Also here, introduce some structure: first, explain shortly what types of validity threats exist for the research methodology you applied.
+Then, discuss each type individually.
+Think about which threat type is the most prominent one for your research methodology and start with that one.
+I did not review this section in detail now since it will change considerably after your restructuring.
+
+%External
 The first threat related to external validity might be caused by a possible wrong selection of subjects interviewed for the study.
@@
 
+%Internal
 To enhance the internal validity, we created a three-dimensional research framework to perform a triangulation validity procedure.
@@
 
+%Construct
 The first important threat to the construct validity of the study is a possible inadequate description of constructs.
@@
 \label{futwork}
+%MUN: Don't have future work as a subsection in threats to validity. Either merge it in "Implications of the GSM" OR into "Conclusions".
 
@@
 Software development, especially in the early-stage, is at the core of the company's daily activities.
-Despite their severely high failure-rate, we have found that the quick proliferation of startups is not supported by an adequate and scientific body of knowledge.
+Despite their severely high failure-rate, we have found that the quick proliferation of startups is not supported by an adequate and scientific body of knowledge.%MUN: <-- last sentence: isn't that a conclusion from the SMS?
  To be able to intervene on the software development strategy with scientific and engineering approaches, the first step is to understand startups' behavior.
@@
 
-Through an intense exploratory research, conducted with a systematic approach, we produced a theoretical model grounded into the hindsight knowledge collected among startup practitioners with the aim of explaining how development strategies are engineered by practitioners.
+Through an exploratory research, conducted by a systematic approach, we produced a theoretical model grounded into the hindsight knowledge collected among practitioners with the aim of explaining how development strategies are engineered startups.
 The explanatory capability and correctness of the model have been validated by means of systematic comparisons with the state-of-the-art results and empirical data.
 The systematic mapping of the literature \cite{SMS}, with 37 studies extracted from an initial sample of 943 items, revealed a multi-faceted state-of-the-art inadequate to support software development activities in startup companies.
+%MUN: <-- you mix again SMS and GT here...
+fix it
 On the other hand, the case study conducted in 13 early-stage startups, provided a broad set of empirical evidences obtained by combining different research methodologies in a grounded theory approach.
 
-The overall results of our research confirm that startups possess unique characteristics of uncertainty, lack of resources and time-pressure.
-These factors influence the software development to an extent that transform every decision related to the development strategies into a difficult trade-off for the company.
+The overall results of our research confirm that startups possess unique %MUN: why unique? 
+characteristics of uncertainty, lack of resources and time-pressure.
+These factors influence the software development to an extent that transforms every decision related to the development strategies into a difficult trade-off for the company.
 Moreover, although startups share different characteristics with other similar SE domains (e.g.
@@
 In this stage, startups often discard any formal project management, documentation, analysis, planning, testing and other traditional process
-activities. In fact, practitioners take advantage of an evolutionary prototyping approach, using well-integrated tools and externalizing complexity to third party solutions.
+activities. Practitioners take advantage of an evolutionary prototyping approach, using well-integrated tools and externalizing complexity to third party solutions.
 
 However, the initial gain obtained in terms of flexibility and speed is counterbalanced by the need of restructuring the product and controlling the engineering activities when the company starts to grow.
-In fact, if successful, the startup will face a growth in terms of number of customers, employees and product functionalities that leads to the necessity to control the initially chaotic software development environment.
+If successful, the startup will face a growth in terms of number of customers, employees and product functionalities that leads to the necessity to control the initially chaotic software development environment.
 In this context the most significant challenge for early-stage startups is finding a sweet spot between being fast enough to enter the market early while controlling the amount of accumulated technical debt.
-In fact, as confirmed by analyzing other framework in SE, the use of prescriptive and rigid structures over the development activities would hinder performance since startups operate across the border between chaotic and complex domain.
-Therefore, from a software development perspective, working with low-precision engineering activities represents the only viable strategy that startups
-can follow in the early stage of their lifecycle.
+As confirmed by analyzing other frameworks in SE, the use of prescriptive and rigid structures over the development activities would hinder performance since startups operate across the border between the chaotic and complex domain %MUN: Not sure whether you have discussed/mentioned the meaning of chaotic/complex domain in the paper, it think not.
+Hence you need to rephrase that (but you can not introduce new concepts here in the conclusion).
+Therefore, from a software development perspective, working with low-precision engineering activities represents the only viable strategy that startups can follow in the early stage of their lifecycle.
 
@@
 By making a comparison with Berry's definition of SE \cite{Berry1992}, we would like to see the rise of a new discipline - \textit{startup engineering} - which can be defined as  \textit{the use of scientific, engineering, managerial and systematic approaches with the aim of successfully developing software systems in startup companies}.
+%MUN: Nice ;-)
 

### 2013-02-22 article/main.tex
Formatted to 80 chars line-lenght

@@
 \markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
-{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
+{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer 
+Society Journals}
 % The only time the second header will appear is for the odd numbered pages
@@
 % or like this to get the Computer Society new two part style.
-%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007 IEEE}%
-%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
+%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007  
+% IEEE}%
+%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer  
+% Society\hfill}}
 % Remember, if you use this you must call \IEEEpubidadjcol in the second
@@
 %\boldmath
-Software startups are newly created companies with no operating history and oriented in producing cutting-edge technologies.
-These companies develop software under highly uncertain conditions, tackling fast growing markets with severe lack of resources.
-Despite their increasing importance in economy there are only few scientific studies attempting to address software engineering (SE) issues, especially for early-stage startups.
-This research aims to understand the software development strategies conducted by practitioners, in the period of time that goes from idea conception to the first open beta release.
-A \textit{grounded theory} approach guided the execution of a case study to explore the state-of-practice.
-As result, the Greenfield startup model (GSM)  reveals the urgent priority of startups of releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early collected user feedback.
-Nevertheless, the need of shortening
-time-to-market, by speeding-up the development through low-precision and product-centric engineering activities, is counterbalanced by the need of restructuring the product before setting off for further growth.
+Software startups are newly created companies with no operating history and 
+oriented in producing cutting-edge technologies. These companies develop 
+software under highly uncertain conditions, tackling fast growing markets with 
+severe lack of resources. Despite their increasing importance in economy there 
+are only few scientific studies attempting to address software engineering (SE) 
+issues, especially for early-stage startups. This research aims to understand 
+the software development strategies conducted by practitioners, in the period of 
+time that goes from idea conception to the first open beta release. A 
+\textit{grounded theory} approach guided the execution of a case study to 
+explore the state-of-practice. As result, the Greenfield startup model (GSM)  
+reveals the urgent priority of startups of releasing the product as quickly as 
+possible to verify the product/market fit and to adjust the business and product 
+trajectory according to early collected user feedback. Nevertheless, the need of 
+shortening time-to-market, by speeding-up the development through low-precision 
+and product-centric engineering activities, is counterbalanced by the need of 
+restructuring the product before setting off for further growth.
 \end{abstract}
@@
 
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+%------------------------------------------------------------------------------%
 \section{Introduction}
@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{A}{n} impressive number of new software startups are launched worldwide every day as a result of an increase of new markets, accessible technologies, and venture capital \cite{8491286}.
-With the term \textit{software startups} we refer to those temporary organizations focused on the creation of high-tech and innovative products, with little or no operating history, aiming to grow by  aggressively scaling their business in highly scalable markets.
-
-New ventures such as \textit{Facebook}, \textit{Linkedin}, \textit{Spotify}, \textit{Pinterest}, \textit{Instagram}, \textit{Groupon} and \textit{Dropbox}, to name a few, are good examples of startups that evolved into successful businesses.
-But, despite many successful stories, the great majority of startups fail within two years from their creation, primarily due to self-destruction rather than competition \cite{Crowne2002}.
-Operating in a chaotic, rapidly evolving and uncertain domain, software startups face intense time-pressure from the market and are exposed to tough competition \cite{Maccormack2001, Eisenhardt1998}.
-In order to succeed in this environment, it is crucial to choose the right features to build and be able to quickly adapt the product to new requests constrained by very limited amount of resources \cite{Sutton2000}.
-
-From a software engineering perspective development in startups is challenging since they develop software in a context where processes can hardly follow a prescriptive methodology \cite{Sutton2000, Coleman2005}.
-Despite startups share some characteristics with similar domains (e.g.
-small and web companies), the combination of different factors makes the specific development context quite unique \cite{Blank2005, Sutton2000}.
-Therefore, research is needed to support their engineering activities \cite{Coleman2005}, guiding practitioners in taking decisions and avoiding choices that could easily lead the whole business to failure \cite{Kajko-Mattsson2008}.
-However, despite the impressive size of the startup ecosystem \cite{ISI:000243253000007}, the state-of-the-art presents a gap.
-Through a \textit{Systematic Mapping Study} (SMS) of the literature, conducted by Paternoster et al.
-\cite{SMS}, only few publications were found related to engineering activities in startups.
-Moreover the majority of these studies do not possess the relevance and rigour required to form a consistent body of knowledge in the state-of-the-art and in the state-of-practice.
-
-This research aims to understand how software development strategies are engineered by practitioners in startup companies in terms of level of: structure, planning and control of software projects, in the period of time  that goes from idea conception to the first open beta release of the software product.
-
-With a cross-sectional study we focused the research context in a limited time-frame to emphasize the nature of uncertainty in the development activities and time pressure from the market, as discussed in \cite{Maccormack2001, Eisenhardt1998}.
-We performed semi-structured in-depth interviews with CEOs and CTOs of startups, covering a wide spectrum of thematics and iteratively adjusting the direction of the research according to the emerging evidences.
-
-By means of the Green-field startup model (GSM), we captured the underlying phenomenon of software development in early-stage startups.
-The framework is based on a systematic procedure  which is grounded into empirical concepts obtained by a case-study.
-GSM presents the most significant themes of the development strategies that characterise startups' context.
-
-The results of our study suggest that the most urgent priority of software development in startups is to shorten time-to-market.
-The greater part of engineering activities are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing).
-The first release of the product includes only a limited set of well suitable functionalities.
-
-Nevertheless, by speeding-up the development process, startups accumulate technical debt, ignoring processes, relying on informal communication and replacing documentation with low-precision artifacts.
-For startups it is essential to look ahead within short-term deadlines and determine the right product/market fit driven by user feedback, using flexible and reactive development approaches and adopting mitigation strategies to contain the severity of accumulated technical debt.
-
-Practitioners can use the GSM model to apply presented strategies to speed-up the development by taking in consideration the amount of technical debt accumulated during the first early-stage period.
-In this regard, we identified several commonalities between the issues related to software development in startups and the research focused at studying technical debt \cite{Nugroho2011,Izurieta2012}.
-On the other hand, researchers can take advantage from the GSM model to understand how the technical debt influences the future growth of startup companies.
-
-The remainder of this paper is structured as follows:  we provide relevant information about startups in section \ref{backg}.
-Section \ref{resmet} introduces the research questions and discusses the design and execution of the research process.
-Results are discussed in section \ref{res:gsm}, which presents the GSM model.
-Section \ref{res:val}  compares results of this research to literature and frameworks presented in this field.
-Section \ref{valt} discusses the validity threats considered during this study.
-Section \ref{sect:theory:impl} discusses the most relevant implications of the GSM model.
+\IEEEPARstart{A}{n} impressive number of new software startups are launched 
+worldwide every day as a result of an increase of new markets, accessible 
+technologies, and venture capital \cite{8491286}.  With the term 
+\textit{software startups} we refer to those temporary organizations focused on 
+the creation of high-tech and innovative products, with little or no operating 
+history, aiming to grow by  aggressively scaling their business in highly 
+scalable markets.
+
+New ventures such as \textit{Facebook}, \textit{Linkedin}, \textit{Spotify}, 
+\textit{Pinterest}, \textit{Instagram}, \textit{Groupon} and \textit{Dropbox}, 
+to name a few, are good examples of startups that evolved into successful 
+businesses. But, despite many successful stories, the great majority of startups 
+fail within two years from their creation, primarily due to self-destruction 
+rather than competition \cite{Crowne2002}. Operating in a chaotic, rapidly 
+evolving and uncertain domain, software startups face intense time-pressure from 
+the market and are exposed to tough competition \cite{Maccormack2001, 
+Eisenhardt1998}. In order to succeed in this environment, it is crucial to 
+choose the right features to build and be able to quickly adapt the product to 
+new requests constrained by very limited amount of resources \cite{Sutton2000}.
+
+From a software engineering perspective development in startups is challenging 
+since they develop software in a context where processes can hardly follow a 
+prescriptive methodology \cite{Sutton2000, Coleman2005}. Despite startups share 
+some characteristics with similar domains (e.g. small and web companies), the 
+combination of different factors makes the specific development context quite 
+unique \cite{Blank2005, Sutton2000}. Therefore, research is needed to support 
+their engineering activities \cite{Coleman2005}, guiding practitioners in taking 
+decisions and avoiding choices that could easily lead the whole business to 
+failure \cite{Kajko-Mattsson2008}.
+However, despite the impressive size of the startup ecosystem 
+\cite{ISI:000243253000007}, the state-of-the-art presents a gap. Through a 
+\textit{Systematic Mapping Study} (SMS) of the literature, conducted by 
+Paternoster et al. \cite{SMS}, only few publications were found related to 
+engineering activities in startups. Moreover the majority of these studies do 
+not possess the relevance and rigour required to form a consistent body of 
+knowledge in the state-of-the-art and in the state-of-practice.
+
+This research aims to understand how software development strategies are 
+engineered by practitioners in startup companies in terms of level of: 
+structure, planning and control of software projects, in the period of time  
+that goes from idea conception to the first open beta release of the software 
+product.
+
+With a cross-sectional study we focused the research context in a limited 
+time-frame to emphasize the nature of uncertainty in the development activities 
+and time pressure from the market, as discussed in \cite{Maccormack2001, 
+Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and 
+CTOs of startups, covering a wide spectrum of thematics and iteratively 
+adjusting the direction of the research according to the emerging evidences.
+
+By means of the Green-field startup model (GSM), we captured the underlying 
+phenomenon of software development in early-stage startups. The framework is 
+based on a systematic procedure  which is grounded into empirical concepts 
+obtained by a case-study. GSM presents the most significant themes of the 
+development strategies that characterise startups' context.
+
+The results of our study suggest that the most urgent priority of software 
+development in startups is to shorten time-to-market. The greater part of 
+engineering activities are focused on the implementation while only little 
+attention is given to more conventional activities (project management, 
+requirement specifications, analysis, architecture design, automatic testing). 
+The first release of the product includes only a limited set of well suitable 
+functionalities.
+
+Nevertheless, by speeding-up the development process, startups accumulate 
+technical debt, ignoring processes, relying on informal communication and 
+replacing documentation with low-precision artifacts. For startups it is 
+essential to look ahead within short-term deadlines and determine the right 
+product/market fit driven by user feedback, using flexible and reactive 
+development approaches and adopting mitigation strategies to contain the 
+severity of accumulated technical debt.
+
+Practitioners can use the GSM to apply presented strategies to speed-up the 
+development by taking in consideration the amount of technical debt accumulated 
+during the first early-stage period. In this regard, we identified several 
+commonalities between the issues related to software development in startups and 
+the research focused at studying technical debt \cite{Nugroho2011,Izurieta2012}. 
+On the other hand, researchers can take advantage from the GSM to 
+understand how the technical debt influences the future growth of startup 
+companies.
+
+The remainder of this paper is structured as follows:  we provide relevant 
+information about startups in section \ref{backg}. Section \ref{resmet} 
+introduces the research questions and discusses the design and execution of the 
+research process. Results are discussed in section \ref{res:gsm}, which presents 
+the GSM. Section \ref{res:val}  compares results of this research to 
+literature and frameworks presented in this field. Section \ref{valt} discusses 
+the validity threats considered during this study. Section 
+\ref{sect:theory:impl} discusses the most relevant implications of the GSM. 
 Finally, this paper concludes in section \ref{conc}.
 
-%MUN: This is maybe only a minor concern, but it may be relevant for certain reviewers: The references you use
-% in the introduction are not very recent. Look at the list of papers [1-15], most of them are published before 2005.
-% I understand that you use some references to strengthen your general arguments (which can refer to older publications),
-% but your introduction should also show the relevance of your research right now. This is best shown by referring to
-% recent (max 5 years old) literature. Think about if you can add more recent (2010-now), well published, papers in the
-% introduction.
+%MUN: This is maybe only a minor concern, but it may be relevant for certain 
+%reviewers: The references you use in the introduction are not very recent. 
+%Look at the list of papers [1-15], most of them are published before 2005.
+%I understand that you use some references to strengthen your general arguments 
+%(which can refer to older publications), but your introduction should also 
+%show the relevance of your research right now. This is best shown by 
+%referring to recent (max 5 years old) literature. Think about if you can add 
+%more recent (2010-now), well published, papers in the introduction.
 
@@
 %\hfill January 11, 2007
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+%------------------------------------------------------------------------------%
 \section{Background}
@@
 
-Modern entrepreneurship has been boosted by the advent of the consumer internet markets in the middle of the 1990s and culminated with the notorious \textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126, storey1982entrepreneurship}.
-Several years later, with the massification of the internet and mobile devices, we are now assisting to an impressive proliferation of software ventures - metaphorically referred as the \textit{startup bubble}.
-The easy access to vast potential markets and the low cost of services distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}.
-Inspired by success stories, a large number of software businesses are created every day.
-However, the great majority of these companies fail within two years from their creation \cite{Crowne2002}.
-
-Looking at the number of new business incubators which appeared in the last decade one can estimate the importance of startups \cite{Grimaldi2005}.
-The wave of disruption in new technologies has led companies to be more and more competitive, forcing themselves to radical organizational and innovational renewals,  which bring many companies to the attempt of behaving like startups \cite{Christensen1997}.
-
-The implementation of methodologies to structure and control the development activities in startups is a major challenge for engineers \cite{Coleman2008}.
-Generally, the management of software development is achieved through the introduction of software processes, which define what steps the development organizations should take at each stage of production and provide assistance in making estimates, developing plans and measuring quality \cite{Chroust1996}.
-In the last decades, several models have been introduced to control software development activities.
-However, their application in startup companies doesn't report significant benefits \cite{Coleman2008a, Coleman2008, Sutton2000}.
-
-In this context, software engineering (SE) faces complex and multifaceted obstacles in understanding how to manage development processes.
-Sutton defines startups as creative and flexible in nature and reluctant to introduce process or bureaucratic measures which may result ineffective \cite{Sutton2000}.
-Also Bach refers to startups as \textit{``a bunch of energetic and committed people without defined development processes''} \cite{Bach1998}.
-In fact, startups have very limited resources and typically wish to use them to support product development instead of establishing processes \cite{Coleman2008, Heitlager2007}.
-Some attempts to taylor lightweight processes to startups reported basic failure of their application: \textit{``Everyone is busy, and software engineering practices are often one of the first places developers cut corners''} \cite{Martin2007}.
-Rejecting the notion of repeatable and controlled processes, startups prominently take advantage of unpredictable, reactive and low-precision \cite{
-surviving-os-cockburn} engineering practices \cite{Sutton2000, Tanabian2005, Chorev2006, Kakati2003}.
-
-Moreover, most startups develop packaged applications rather than software for a specific client \cite{genome2012}.
-Issues related to this domain are addressed in literature by the area known as \textit{market-driven software development} (sometimes called \textit{packaged software development} or \textit{COTS software development} \cite{Alves2006}).
-Among other results, researchers emphasize the importance of time-to-market as a key strategic objective \cite{dagMDR,sawyer99} for companies operating in this domain.
-Other peculiar aspects which influence the software development are related to requirements, which are reported to be often ``invented by the software company'' \cite{512553}, ``rarely documented'' \cite{Karlsson02challengesin}, and can be validated only after the product is released to market \cite{dahl2003,Keil:1995}.
-Under these circumstances, failure of product launches are largely due to ``products not meeting customer needs''  \cite{Alves2006}.
-
-Accordingly, product-oriented practices help startups having flexible teams, with workflows that leave them the ability to quickly change the direction according to the targeted market  \cite{Heitlager2007,Sutton2000}.
-In this regard, many startups focus on team productivity, asserting more control to the employees instead of providing them rigid guidelines \cite{Tanabian2005, Chorev2006, Kakati2003}.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+Modern entrepreneurship has been boosted by the advent of the consumer internet 
+markets in the middle of the 1990s and culminated with the notorious 
+\textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126, 
+storey1982entrepreneurship}. Several years later, with the massification of the 
+internet and mobile devices, we are now assisting to an impressive proliferation 
+of software ventures - metaphorically referred as the \textit{startup bubble}. 
+The easy access to vast potential markets and the low cost of services 
+distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}. 
+Inspired by success stories, a large number of software businesses are created 
+every day. However, the great majority of these companies fail within two years 
+from their creation \cite{Crowne2002}.
+
+Looking at the number of new business incubators which appeared in the last 
+decade one can estimate the importance of startups \cite{Grimaldi2005}. The wave 
+of disruption in new technologies has led companies to be more and more 
+competitive, forcing themselves to radical organizational and innovational 
+renewals,  which bring many companies to the attempt of behaving like startups 
+\cite{Christensen1997}.
+
+The implementation of methodologies to structure and control the development 
+activities in startups is a major challenge for engineers \cite{Coleman2008}. 
+Generally, the management of software development is achieved through the 
+introduction of software processes, which define what steps the development 
+organizations should take at each stage of production and provide assistance in 
+making estimates, developing plans and measuring quality \cite{Chroust1996}. In 
+the last decades, several models have been introduced to control software 
+development activities. However, their application in startup companies doesn't 
+report significant benefits \cite{Coleman2008a, Coleman2008, Sutton2000}.
+
+In this context, software engineering (SE) faces complex and multifaceted 
+obstacles in understanding how to manage development processes. Sutton defines 
+startups as creative and flexible in nature and reluctant to introduce process 
+or bureaucratic measures which may result ineffective \cite{Sutton2000}. Also 
+Bach refers to startups as \textit{``a bunch of energetic and committed people 
+without defined development processes''} \cite{Bach1998}. In fact, startups have 
+very limited resources and typically wish to use them to support product 
+development instead of establishing processes \cite{Coleman2008, Heitlager2007}. 
+Some attempts to taylor lightweight processes to startups reported basic failure 
+of their application: \textit{``Everyone is busy, and software engineering 
+practices are often one of the first places developers cut corners''} 
+\cite{Martin2007}. Rejecting the notion of repeatable and controlled processes, 
+startups prominently take advantage of unpredictable, reactive and low-precision 
+\cite{surviving-os-cockburn} engineering practices \cite{Sutton2000, 
+Tanabian2005, Chorev2006, Kakati2003}.
+
+Moreover, most startups develop packaged applications rather than software for 
+a specific client \cite{genome2012}. Issues related to this domain are addressed 
+in literature by the area known as \textit{market-driven software development} 
+(sometimes called \textit{packaged software development} or \textit{COTS 
+software development} \cite{Alves2006}). Among other results, researchers 
+emphasize the importance of time-to-market as a key strategic objective 
+\cite{dagMDR,sawyer99} for companies operating in this domain.  Other peculiar 
+aspects which influence the software development are related to requirements, 
+which are reported to be often ``invented by the software company'' 
+\cite{512553}, ``rarely documented'' \cite{Karlsson02challengesin}, and can be 
+validated only after the product is released to market 
+\cite{dahl2003,Keil:1995}. Under these circumstances, failure of product 
+launches are largely due to ``products not meeting customer needs''  
+\cite{Alves2006}.
+
+Accordingly, product-oriented practices help startups having flexible teams, 
+with workflows that leave them the ability to quickly change the direction 
+according to the targeted market  \cite{Heitlager2007,Sutton2000}. In this 
+regard, many startups focus on team productivity, asserting more control to the 
+employees instead of providing them rigid guidelines \cite{Tanabian2005, 
+Chorev2006, Kakati2003}.
+
+%------------------------------------------------------------------------------%
 \subsection{General lack of research in startups}
-Despite some studies tried to address the above mentioned issues, a SMS \cite{SMS} revealed only a few SE works in this specific area, as confirmed by other studies \cite{Coleman2008, Coleman2008a, Coleman2007, Sutton2000}.
-Moreover, the studies, identified in \cite{SMS}, appear to be highly fragmented and spread across different areas rather than constituting a consistent \textit{body of knowledge}.
-
-A research published in 2000 by Sutton \cite{Sutton2000} confirmed a general lack of  studies in this area, claiming that \textit{``software startups represent a segment that has been mostly neglected in process studies''} and it has been further confirmed with the empirical studies of Coleman et al.
+Despite some studies tried to address the above mentioned issues, a SMS 
+\cite{SMS} revealed only a few SE works in this specific area, as confirmed by 
+other studies \cite{Coleman2008, Coleman2008a, Coleman2007, Sutton2000}. 
+Moreover, the studies, identified in \cite{SMS}, appear to be highly fragmented 
+and spread across different areas rather than constituting a consistent 
+\textit{body of knowledge}.
+
+A research published in 2000 by Sutton \cite{Sutton2000} confirmed a general 
+lack of  studies in this area, claiming that \textit{``software startups 
+represent a segment that has been mostly neglected in process studies''} and it 
+has been further confirmed with the empirical studies of Coleman et al. 
 \cite{Coleman2008,Coleman2008a,Coleman2007} eight years later.
 
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+%------------------------------------------------------------------------------%
 \subsection{Software development in startups}
 
-The word \textit{startup} appeared in the SE literature for the first time in 1994 in an article written by Carmel \cite{Camel1994a} where he studied the \textit{time-to-completion in a young package firm}.
-%MUN: If this is a quote from
-%Carmel, put it in quotation marks, like you did in the previous section with Suttons quote. Generally, check that you
-%are consistent how you use quotations throughout the paper.
-He noticed how these companies were particularly innovative and successful, advocating for the need of more research investigating software development practices so as to replicate success and try to transfer it to other technology sectors.
-
-First insights reveal how software startups are product-oriented in the first period of their development phase \cite{Heitlager2007}.
-Despite good achievements at the beginning, software development and organizational management increase in complexity \cite{1456074,Banker1998} causing  deterioration of performance over time.
-Briefly, the necessity of establishing initial repeatable and scalable processes cannot be postponed forever.
-
-A study of Kajko-Mattsson \cite{Kajko-Mattsson2008}, which investigated a Swedish software startup, reported a heavy lack of requirements gathering processes, minimal project management, lack of control over the change requests, absence of documentation to track the status and progress of the process, and defective releases.
-Accordingly, Ambler et al.
-report how two startups approaching to an upcoming IPO started to require processes to focus on scalable solutions, in view of the growing company's size in terms of users and employees \cite{Ambler2002}.
-In this regard, Crowne \cite{Crowne2002} specifies different stages through which software startups evolve.
-Starting without any established workflows, startups grow over time, creating and stabilizing processes to eventually improving them only when sufficiently mature.
-
-But when startups have no time for training and orienting activities, as discussed in \cite{Sutton2000}, their main focus remains on team capabilities instead of prescriptive processes, hiring people who can \textit{``hit the ground running''} \cite{Yoffie1999}.
-Empowering the team and focus on methodological attributes of the processes oriented in prototyping, proof-of-concepts, mock-ups and demos, to test basic functionalities, have been the primary priority in startups \cite{Camel1994a}.
-Only when they grow, formal methodologies arise, followed by performing quality assurance and long-term planning processes \cite{Yoffie1999}.
+The word \textit{startup} appeared in the SE literature for the first time in 
+1994 in an article written by Carmel \cite{Camel1994a} where he studied the 
+\textit{time-to-completion in a young package firm}. %MUN: If this is a quote 
+%from Carmel, put it in quotation marks, like you did in the previous section 
+%with Suttons quote. Generally, check that you %are consistent how you use 
+%quotations throughout the paper.
+He noticed how these companies were particularly innovative and successful, 
+advocating for the need of more research investigating software development 
+practices so as to replicate success and try to transfer it to other technology 
+sectors.
+
+First insights reveal how software startups are product-oriented in the first 
+period of their development phase \cite{Heitlager2007}. Despite good 
+achievements at the beginning, software development and organizational 
+management increase in complexity \cite{1456074,Banker1998} causing  
+deterioration of performance over time. Briefly, the necessity of establishing 
+initial repeatable and scalable processes cannot be postponed forever.
+
+A study of Kajko-Mattsson \cite{Kajko-Mattsson2008}, which investigated a 
+Swedish software startup, reported a heavy lack of requirements gathering 
+processes, minimal project management, lack of control over the change requests, 
+absence of documentation to track the status and progress of the process, and 
+defective releases. Accordingly, Ambler et al. report how two startups 
+approaching to an upcoming IPO started to require processes to focus on scalable 
+solutions, in view of the growing company's size in terms of users and employees 
+\cite{Ambler2002}. In this regard, Crowne \cite{Crowne2002} specifies different 
+stages through which software startups evolve. Starting without any established 
+workflows, startups grow over time, creating and stabilizing processes to 
+eventually improving them only when sufficiently mature.
+
+But when startups have no time for training and orienting activities, as 
+discussed in \cite{Sutton2000}, their main focus remains on team capabilities 
+instead of prescriptive processes, hiring people who can \textit{``hit the 
+ground running''} \cite{Yoffie1999}. Empowering the team and focus on 
+methodological attributes of the processes oriented in prototyping, 
+proof-of-concepts, mock-ups and demos, to test basic functionalities, have been 
+the primary priority in startups \cite{Camel1994a}. Only when they grow, formal 
+methodologies arise, followed by performing quality assurance and long-term 
+planning processes \cite{Yoffie1999}.
 
 The maturity of a company affects the extent to which processes are adopted. 
-Tingling \cite{Tingling2007} reports how introducing Extreme Programming (XP) principles \cite{Beck:2004:EPE:1076267} in the development process was challenging because of the need of trained team-members for fully implementing the methodology.
-Da Silva and Kon \cite{Silva2005} were able to start with all the XP practices in place only after six months of coaching the team, trying to enhance maturity from day-one.
-Nevertheless, even then, customization of practices were inevitably implemented to adapt the processes to the undertaken startups' context \cite{Deias}.
-
-As partially discussed above, contributions to flexibility and reactiveness of the development process has been conducted prominently by means of \textit{Lean} \cite{Gautam2008} and \textit{Agile} \cite{Abrahamsson2002} methodologies (also reported in \cite{Taipale2010, Kuvinka2011}), where the extreme uncertain conditions lead startups to learn fast from trials and errors with a strong customer relationship in order to avoid wasting time in building wrong functionalities and prevent rapidly exhaustion of resources \cite{Midler2008, Hilmola2003, Sutton2000}.
-Customer involvement in software development has also been discussed by Yogendra \cite{Yogendra2002} as an important factor to encourage an early alignment of business concerns to the technology strategies, because both are salient considerations to be successful \cite{Tingling2007}.
-
-When startups take a development approach that is mainly product-oriented rather than process-oriented, it is essential to have a flexible team, with a workflow that helps them quickly to change direction according to the target market \cite{Sutton2000}.
-In this regard, many startups have been focused on team productivity, providing more control to the employees instead of providing them rigid guidelines \cite{Tanabian2005, Chorev2006, Kakati2003}.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+Tingling \cite{Tingling2007} reports how introducing Extreme Programming (XP) 
+principles \cite{Beck:2004:EPE:1076267} in the development process was 
+challenging because of the need of trained team-members for fully implementing 
+the methodology. Da Silva and Kon \cite{Silva2005} were able to start with all 
+the XP practices in place only after six months of coaching the team, trying to 
+enhance maturity from day-one. Nevertheless, even then, customization of 
+practices were inevitably implemented to adapt the processes to the undertaken 
+startups' context \cite{Deias}.
+
+As partially discussed above, contributions to flexibility and reactiveness of 
+the development process has been conducted prominently by means of \textit{Lean} 
+\cite{Gautam2008} and \textit{Agile} \cite{Abrahamsson2002} methodologies (also 
+reported in \cite{Taipale2010, Kuvinka2011}), where the extreme uncertain 
+conditions lead startups to learn fast from trials and errors with a strong 
+customer relationship in order to avoid wasting time in building wrong 
+functionalities and prevent rapidly exhaustion of resources \cite{Midler2008, 
+Hilmola2003, Sutton2000}. Customer involvement in software development has also 
+been discussed by Yogendra \cite{Yogendra2002} as an important factor to 
+encourage an early alignment of business concerns to the technology strategies, 
+because both are salient considerations to be successful \cite{Tingling2007}.
+
+When startups take a development approach that is mainly product-oriented 
+rather than process-oriented, it is essential to have a flexible team, with a 
+workflow that helps them quickly to change direction according to the target 
+market \cite{Sutton2000}. In this regard, many startups have been focused on 
+team productivity, providing more control to the employees instead of providing 
+them rigid guidelines \cite{Tanabian2005, Chorev2006, Kakati2003}.
+
+%------------------------------------------------------------------------------%
 \subsection{Software process improvement in startups}
 
-When startups are mature enough to support software process improvement (SPI), the solutions considered according to the state-of-the-art are oriented towards light-weight processes such as a design of development process based on XP, proposed by Zettel \cite{Zettel2001}.
-The process consists of a set of activities and artifacts (in addition to some important roles) defined in order to identify responsibilities and tools to utilize.
-But, despite the promising benefits reported by Zettel, we were not able to identify any future evaluation in real-world settings.
+When startups are mature enough to support software process improvement (SPI), 
+the solutions considered according to the state-of-the-art are oriented towards 
+light-weight processes such as a design of development process based on XP, 
+proposed by Zettel \cite{Zettel2001}. The process consists of a set of 
+activities and artifacts (in addition to some important roles) defined in order 
+to identify responsibilities and tools to utilize. But, despite the promising 
+benefits reported by Zettel, we were not able to identify any future evaluation 
+in real-world settings.
 
 Another attempt of SPI in startups has been conducted by Deakins et al. 
-introducing a \textit{Helical Model for managing e-commerce development environment} \cite{Deakins2005}.
-Also in this case, the author prescribed broad guidelines for a rapid high-quality development process, which underwent limited testing only in academic settings.
-
-The problem of \textit{one-size-fits-all}, related to the SPI  representations for startups, is described by Fayad \cite{Fayad1997}.
-The author discusses the problem in actuating the same \textit{best-practices} criteria for established companies in 10-person software startups.
-Thoroughly remarked in \cite{Sutton2000}, Sutton states that problems of SPI in software startups arise because of: the dynamism of the development process, which precludes repeatability; organizational maturity, that cannot be maintained from startups in view of lack of corporate direction; severe lack of resources, both human and technological for process definition, implementation, management, and training.
-In conclusion, the primary benefits of SPI do not address startups, which instead of promoting product quality, aim to minimize time-to-market.
-
-Additionally, the role of SPI has been neglected because seen as an obstacle to the development team's creativity and flexibility \cite{Coleman2008a}, and to the need of a quick delivering product process environment.
-%MUN: the second
-%part of the sentence is just a statement without reference. If it refers to a statement by Coleman, put his reference
-%at the end OR find the correct reference OR delete the statement.
-Product quality is left aside in favor of minimal and suitable functionalities to shorten the time-to-market.
-As reported in two studies of Mater and Mirel \cite{Mater2000, Mirel2000}, quality aspects, mostly taken in consideration in internet startups, are oriented to usability and scalability, even though the market and application type heavily influences the quality-demand \cite{Coleman2008, Kim2005}.
-
-To maintain the development activities, oriented to limited but suitable functionality, studies suggest to externalize the complexity of parts of the project to third party solutions by means of outsourcing activities, software reuse and open-source strategies \cite{Hanna2010, Jansen2008, Wall2001, Bean2005}.
-%MUN: Can you split up the references?
-I.e.
-write:
-%"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source strategies \cite{Bla}.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+introducing a \textit{Helical Model for managing e-commerce development 
+environment} \cite{Deakins2005}. Also in this case, the author prescribed broad 
+guidelines for a rapid high-quality development process, which underwent limited 
+testing only in academic settings.
+
+The problem of \textit{one-size-fits-all}, related to the SPI  representations 
+for startups, is described by Fayad \cite{Fayad1997}. The author discusses the 
+problem in actuating the same \textit{best-practices} criteria for established 
+companies in 10-person software startups. Thoroughly remarked in 
+\cite{Sutton2000}, Sutton states that problems of SPI in software startups arise 
+because of: the dynamism of the development process, which precludes 
+repeatability; organizational maturity, that cannot be maintained from startups 
+in view of lack of corporate direction; severe lack of resources, both human and 
+technological for process definition, implementation, management, and training. 
+In conclusion, the primary benefits of SPI do not address startups, which 
+instead of promoting product quality, aim to minimize time-to-market.
+
+Additionally, the role of SPI has been neglected because seen as an obstacle to 
+the development team's creativity and flexibility \cite{Coleman2008a}, and to 
+the need of a quick delivering product process environment. %MUN: the second
+%part of the sentence is just a statement without reference. If it refers to a 
+%statement by Coleman, put his reference at the end OR find the correct 
+%reference OR delete the statement.
+Product quality is left aside in favor of minimal and suitable functionalities 
+to shorten the time-to-market. As reported in two studies of Mater and Mirel 
+\cite{Mater2000, Mirel2000}, quality aspects, mostly taken in consideration in 
+internet startups, are oriented to usability and scalability, even though the 
+market and application type heavily influences the quality-demand 
+\cite{Coleman2008, Kim2005}.
+
+To maintain the development activities, oriented to limited but suitable 
+functionality, studies suggest to externalize the complexity of parts of the 
+project to third party solutions by means of outsourcing activities, software 
+reuse and open-source strategies \cite{Hanna2010, Jansen2008, Wall2001, 
+Bean2005}. %MUN: Can you split up the references? I.e. write:
+%"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source 
+%strategies \cite{Bla}.
+
+%------------------------------------------------------------------------------%
 \subsection{Technical debt}
-Notwithstanding, a new area of the SE research, trying to tackle the problem of the \textit{technical debt},  brings and encompasses different implications in studying development in software startups \cite{Tom2013}.
-The metaphoric neologism was originally introduced by Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the attention of SE researchers\footnote{To attest the fact that \textit{technical debt} is gaining traction among researchers, we mention two important contributions which characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt} organized by the Software Engineering Institute  and ICSE.}.
-The concept of \textit{technical debt} is well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea is that developers sometimes accept compromises in a system in one dimension (e.g., modularity) to meet an urgent demand in some other dimension (e.g., a deadline), and that such compromises incur a
-``debt'
-' on which ``interest'' has to be paid and which the ``principal'' should be repaid at some point for the long-term health of the project''}.
-The compromise between high-speed and high-quality engineering is faced daily by software startups, not only in terms of architecture design but in multifaceted aspects (weak project management, testing, process control).
-
-Since \textit{``all decisions related to product development are trade-off situations''} \cite{Hilmola2003}, generally startups optimize workflows to the dynamic context they are involved into.
-They typically adopt any development style that might work to support their first needs in what is called the \textit{``Just do it''} school of software startups \cite{Ries2011}.
-Additionally, as remarked by Coleman, \textit{``many managers just decide to apply what they know, as their experience tells them it is merely common sense''} \cite{Coleman2008}.
+Notwithstanding, a new area of the SE research, trying to tackle the problem of 
+the \textit{technical debt},  brings and encompasses different implications in 
+studying development in software startups \cite{Tom2013}. The metaphoric 
+neologism was originally introduced by Cunningham in 1992 
+\cite{TechnicalDebtCunn} and has recently attracted the attention of SE 
+researchers\footnote{To attest the fact that \textit{technical debt} is gaining 
+traction among researchers, we mention two important contributions which 
+characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} 
+and a dedicated workshop \cite{workshopDebt} organized by the Software 
+Engineering Institute  and ICSE.}. The concept of \textit{technical debt} is 
+well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea 
+is that developers sometimes accept compromises in a system in one dimension 
+(e.g., modularity) to meet an urgent demand in some other dimension (e.g., a 
+deadline), and that such compromises incur a ``debt'' on which ``interest'' has 
+to be paid and which the ``principal'' should be repaid at some point for the 
+long-term health of the project''}.  The compromise between high-speed and 
+high-quality engineering is faced daily by software startups, not only in terms 
+of architecture design but in multifaceted aspects (weak project management, 
+testing, process control).
+
+Since \textit{``all decisions related to product development are trade-off 
+situations''} \cite{Hilmola2003}, generally startups optimize workflows to the 
+dynamic context they are involved into. They typically adopt any development 
+style that might work to support their first needs in what is called the 
+\textit{``Just do it''} school of software startups \cite{Ries2011}. 
+Additionally, as remarked by Coleman, \textit{``many managers just decide to 
+apply what they know, as their experience tells them it is merely common 
+sense''} \cite{Coleman2008}.
 
 \subsection{Terminology}
-In order to set a common ground and to prevent ambiguity, we use the following terminology throughout the paper:
+In order to set a common ground and to prevent ambiguity, we use the following 
+terminology throughout the paper:
 
 \begin{compactitem}
-\item Software development strategy: the overall approach adopted by the company to carry out the product development.
-\item Engineering activities: the activities needed to bring a product from idea to market.
-Traditional engineering activities are, among others, requirement engineering, design, architecture, implementation, testing.
-\item Engineering elements: any practice, tool and artifacts contributing and supporting the engineering activities.
-\item Quality attributes: those overall factors that affect run-time behavior, system design, and user experience.
-They represent areas of concern that have the potential for applications to impact across layers and tiers.
-Some of these attributes are related to the overall system design, while others are specific to run time, design time, or user centric issues \cite{Microsoft2009}.
-\item Growth: an increase in the company size respect to the initial conditions in terms of either employees or users/customers, and product complexity in terms of handling an increasing number of feature requests.
+\item Software development strategy: the overall approach adopted by the 
+company to carry out the product development.
+\item Engineering activities: the activities needed to bring a product from 
+idea to market. Traditional engineering activities are, among others, 
+requirement engineering, design, architecture, implementation, testing.
+\item Engineering elements: any practice, tool and artifacts contributing and 
+supporting the engineering activities.
+\item Quality attributes: those overall factors that affect run-time behavior, 
+system design, and user experience. They represent areas of concern that have 
+the potential for applications to impact across layers and tiers. Some of these 
+attributes are related to the overall system design, while others are specific 
+to run time, design time, or user centric issues \cite{Microsoft2009}.
+\item Growth: an increase in the company size respect to the initial conditions 
+in terms of either employees or users/customers, and product complexity in terms 
+of handling an increasing number of feature requests.
 \item Software product: any software product and software service.
 \end{compactitem}
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+%------------------------------------------------------------------------------%
 \section{Research methodology}
@@
 \includegraphics[width=4in]{figures/completemethodology2}
-\caption{Grounded theory study process overview}\label{fig:gt:completemethodology}
+\caption{Grounded theory study process overview}
+\label{fig:gt:completemethodology}
 \end{figure*}
-%MUN: I think you can increase the font size in the figure a bit. Try thereby to keep the 
-%height of the figure, expanding however the width (use horizontal space since it is a two 
-%column figure). Change also in figure:
+%MUN: I think you can increase the font size in the figure a bit. Try thereby 
+%to keep the height of the figure, expanding however the width (use horizontal 
+%space since it is a two column figure). Change also in figure:
 % - Write "Data Analysis" horizontally
-% - Change names of macro-phases to upper-case, i.e. Design and Execution, Data Collection, Data Analysis, etc.
-% - Add section number to macro-phase name, i.e. 3.1 Design and Execution, 3.2 Data Collection
-% - Change "Discussion and Validation" into "Theory Validation" as you name it in the text.
-% Also, add a section (3.5) that describes shortly what you did in "Theory Validation"
-
-The aim of this research is to \textit{understand how software development strategies are engineered by practitioners in startup companies in terms of level of: structure, planning and control of software projects, in the period of time  that goes from idea conception to the first open beta release of the software product} \footnote{The above mentioned goal is structured according to the GQM paradigm \cite{Basili1992}.}.
-
-Initially the boundaries of the research domain were set by means of non-systematic literature survey, which provided the initial keywords used further in the research process.
-Moreover it helped us in defining the research questions (RQs):
+% - Change names of macro-phases to upper-case, i.e. Design and Execution, 
+%   Data Collection, Data Analysis, etc.
+% - Add section number to macro-phase name, i.e. 3.1 Design and Execution, 3.2 
+%   Data Collection
+% - Change "Discussion and Validation" into "Theory Validation" as you name it 
+%   in the text.
+%Also, add a section (3.5) that describes shortly what you did in "Theory 
+%Validation"
+
+The aim of this research is to \textit{understand how software development 
+strategies are engineered by practitioners in startup companies in terms of 
+level of: structure, planning and control of software projects, in the period of 
+time  that goes from idea conception to the first open beta release of the 
+software product} \footnote{The above mentioned goal is structured according to 
+the GQM paradigm \cite{Basili1992}.}.
+
+Initially the boundaries of the research domain were set by means of 
+non-systematic literature survey, which provided the initial keywords used 
+further in the research process. Moreover it helped us in defining the research 
+questions (RQs):
 
@@
 
-\item RQ-1 : What is the current state-of-practice related to software development strategy in early-stage startups?
+\item RQ-1 : What is the current state-of-practice related to software 
+development strategy in early-stage startups?
 \begin{compactitem}
 
-\item RQ-1.1 : How do startups structure and execute the main engineering activities?
+\item RQ-1.1 : How do startups structure and execute the main engineering 
+activities?
 \item RQ-1.2 : How are product quality attributes considered by startups?
@@
 
-We investigated the software development approach undertaken by practitioners of startups.
-Following a \textit{Grounded Theory} (GT) methodology \cite{Glaser1978}, we executed 13 semi-structured interviews integrated with follow-up questionnaires.
-From this, we elaborated and extracted the Green-field startup model (GSM) explaining the underlying phenomenon of software development in startups.
-
-\textit{Grounded Theory} methodology is described by Robson as \textit{``the best approach to answer the question - What is going on here?''} and defined by its creators \cite{ColinRobson2009} as \textit{``a set of well-developed categories ( e.g.
-themes, concepts ) that are systematically interrelated through statements of relations to form a theoretical framework that explains some relevant social, psychological, educational, nursing or other phenomenon''} .
-
-Despite GT has its roots in the social sciences, it has been increasingly used by Information System researchers and has been almost ignored by the SE research \cite{Coleman2007}.
-In 1997 Bertelsen advocated for a need of more qualitative research in SE \cite{Bertelsen1997}, and in the last decade we assisted to an increasing number of publications that used a GT-like approach in SE \cite{Coleman2008a,Coleman2007, Sulayman2012}.
-
-Following the GT principles we attempted to capture the most relevant aspects of software development from startup practitioners, letting a theory emerge from the interviews and adjusting the research hypotheses and questions as we proceeded through.
-During these interviews we collected data related to engineering activities undertaken by startups in the time frame between the idea conception and the first open beta release of the product.
-Then, we proceeded to the analysis of the data, finding important relations among concepts with a formal approach.
-
-As suggested by Coleman, in view of the different versions of GT which emerged in the last years, researchers should indicate which \textit{``implementation''} of the theory is being used \cite{Coleman2007}.
-After the methodology was initially introduced, its original authors  had divergences regarding how the theory should be formed by analyzing the data (Strauss and Corbin on one side, Glaser on the other).
-Glaser advocated for a more \textit{puristic} approach, where the theoretical categories should \textit{``naturally emerge from the data''}, whilst Strauss and Corbin formulated an updated version of the methodology, which leaves to the researchers an higher degree of freedom.
-The latter approach empowers researchers' \textit{``theoretical sensitivity''} \cite{Corbin1990}, and encourages them to outline the research problem beforehand.
-Since the knowledge obtained from parallel execution of the SMS \cite{SMS} and our direct experience with startups companies provided
-a good initial level of knowledge, in this study we use the Corbin and Strauss approach \cite{Strauss1998}.
-
-In this research we address technical aspects related to software development in startups, exploring their operational dynamics.
-In view of the lack of agreement on an unique definition of the word \textit{startup}, we delimited our initial contextual boundaries to newly created innovative and single-product software companies, in the time-frame that goes from the idea conception to the release of the first product in highly scalable markets.
-Moreover, since most startups are web-oriented companies \cite{tc-webstartups-link, Allen2003}, in our case study, we inquired web companies with little or no operating history.
-
-A complete overview of the overall case study methodology and execution is shown in Figure \ref{fig:gt:completemethodology}, which presents how we tailored the general GT methodology to our specific needs.
-
-The initial design of the case study is supported by the results of the \textit{Systematic Mapping Study} (SMS) \cite{SMS} which contributed to define the initial questions for both questionnaires and semi-structured interviews.
-The process depicted in Figure \ref{fig:gt:completemethodology} is evolutionary and affects the design at each new iteration.
-In \textit{Data Collection} the empirical results are integrated and stored in a single case study database and subsequently processed in \textit{Data Analysis} to form the theoretical categories.
-At each iteration the new emergent theory is updated following a formal procedure (\textit{Theory Generation}), and after verifying that \textit{Theoretical Saturation}\footnote{The point at which executing more interviews wouldn't bring any additional value for constructing the theory.} of categories has been achieved, we finally proceeded to \textit{Theory Validation}.
-
-The whole procedure has been executed simultaneously in pair on the same screen by the first two authors, handling conflicts by reviewing the rationale of decision with the third and fourth authors.
-When necessary we performed an in-depth review of the design of the research and data collected during the execution process.
-
-The process details are described in the following subsections, structured according to the five macro phases depicted in bold in Figure \ref{fig:gt:completemethodology}.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+We investigated the software development approach undertaken by practitioners 
+of startups. Following a \textit{Grounded Theory} (GT) methodology 
+\cite{Glaser1978}, we executed 13 semi-structured interviews integrated with 
+follow-up questionnaires. From this, we elaborated and extracted the Green-field 
+startup model (GSM) explaining the underlying phenomenon of software development 
+in startups.
+
+\textit{Grounded Theory} methodology is described by Robson as \textit{``the 
+best approach to answer the question - What is going on here?''} and defined by 
+its creators \cite{ColinRobson2009} as \textit{``a set of well-developed 
+categories ( e.g. themes, concepts ) that are systematically interrelated 
+through statements of relations to form a theoretical framework that explains 
+some relevant social, psychological, educational, nursing or other 
+phenomenon''}.
+
+Despite GT has its roots in the social sciences, it has been increasingly used 
+by Information System researchers and has been almost ignored by the SE research 
+\cite{Coleman2007}. In 1997 Bertelsen advocated for a need of more qualitative 
+research in SE \cite{Bertelsen1997}, and in the last decade we assisted to an 
+increasing number of publications that used a GT-like approach in SE 
+\cite{Coleman2008a,Coleman2007, Sulayman2012}.
+
+Following the GT principles we attempted to capture the most relevant aspects 
+of software development from startup practitioners, letting a theory emerge 
+from the interviews and adjusting the research hypotheses and questions as we 
+proceeded through. During these interviews we collected data related to 
+engineering activities undertaken by startups in the time frame between the idea 
+conception and the first open beta release of the product.  Then, we proceeded 
+to the analysis of the data, finding important relations among concepts with a 
+formal approach.
+
+As suggested by Coleman, in view of the different versions of GT which emerged 
+in the last years, researchers should indicate which \textit{``implementation''} 
+of the theory is being used \cite{Coleman2007}. After the methodology was 
+initially introduced, its original authors  had divergences regarding how the 
+theory should be formed by analyzing the data (Strauss and Corbin on one side, 
+Glaser on the other).  Glaser advocated for a more \textit{puristic} approach, 
+where the theoretical categories should \textit{``naturally emerge from the 
+data''}, whilst Strauss and Corbin formulated an updated version of the 
+methodology, which leaves to the researchers an higher degree of freedom. The 
+latter approach empowers researchers' \textit{``theoretical sensitivity''} 
+\cite{Corbin1990}, and encourages them to outline the research problem 
+beforehand. Since the knowledge obtained from parallel execution of the SMS 
+\cite{SMS} and our direct experience with startups companies provided a good 
+initial level of knowledge, in this study we use the Corbin and Strauss 
+approach \cite{Strauss1998}.
+
+In this research we address technical aspects related to software development 
+in startups, exploring their operational dynamics. In view of the lack of 
+agreement on an unique definition of the word \textit{startup}, we delimited our 
+initial contextual boundaries to newly created innovative and single-product 
+software companies, in the time-frame that goes from the idea conception to the 
+release of the first product in highly scalable markets. Moreover, since most 
+startups are web-oriented companies \cite{tc-webstartups-link, Allen2003}, in 
+our case study, we inquired web companies with little or no operating history.
+
+A complete overview of the overall case study methodology and execution is 
+shown in Figure \ref{fig:gt:completemethodology}, which presents how we tailored 
+the general GT methodology to our specific needs.
+
+The initial design of the case study is supported by the results of the 
+\textit{Systematic Mapping Study} (SMS) \cite{SMS} which contributed to define 
+the initial questions for both questionnaires and semi-structured interviews. 
+The process depicted in Figure \ref{fig:gt:completemethodology} is evolutionary 
+and affects the design at each new iteration. In \textit{Data Collection} the 
+empirical results are integrated and stored in a single case study database and 
+subsequently processed in \textit{Data Analysis} to form the theoretical 
+categories.  At each iteration the new emergent theory is updated following a 
+formal procedure (\textit{Theory Generation}), and after verifying that 
+\textit{Theoretical Saturation}\footnote{The point at which executing more 
+interviews wouldn't bring any additional value for constructing the theory.} of 
+categories has been achieved, we finally proceeded to \textit{Theory 
+Validation}.
+
+The whole procedure has been executed simultaneously in pair on the same screen 
+by the first two authors, handling conflicts by reviewing the rationale of 
+decision with the third and fourth authors. When necessary we performed an 
+in-depth review of the design of the research and data collected during the 
+execution process.
+
+The process details are described in the following subsections, structured 
+according to the five macro phases depicted in bold in Figure 
+\ref{fig:gt:completemethodology}.
+
+%------------------------------------------------------------------------------%
 \subsection{Design and Execution}
@@
 
-This case study started based on a SMS of the literature regarding software development in startups \cite{SMS}.
-By means of this SMS we refined our research questions.
-The research questions have been narrowed down to outline the domain of our study, even though still broad enough to allow emerging new theories and to adjust the scope of the research \cite{Corbin1990}.
-
-The sampling of candidate companies took place in two distinct phases.
-At first we executed an initial convenience sampling \cite{Dawson2009}, which led to the identification of eight companies.
-Then we included five additional startups during the theory formation process (theoretical sampling) iteratively improving the sample according to the emerging theory.
-The characteristics of the sampled companies are reported in Table \ref{t_interviews-stats}.
+This case study started based on a SMS of the literature regarding software 
+development in startups \cite{SMS}. By means of this SMS we refined our research 
+questions. The research questions have been narrowed down to outline the domain 
+of our study, even though still broad enough to allow emerging new theories and 
+to adjust the scope of the research \cite{Corbin1990}.
+
+The sampling of candidate companies took place in two distinct phases. At first 
+we executed an initial convenience sampling \cite{Dawson2009}, which led to the 
+identification of eight companies. Then we included five additional startups 
+during the theory formation process (theoretical sampling) iteratively improving 
+the sample according to the emerging theory. The characteristics of the sampled 
+companies are reported in Table \ref{t_interviews-stats}.
 
@@
 
-All companies, except C10, were founded within the last three years (2009-2012), %MUN: I've shortened the text and
-%added absolute numbers. Is 2009-2012 correct?
+All companies, except C10, were founded within the last three years (2009-2012), 
+%MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct?
 by an average of 3 starting members, which were in majority developers. 
-Moreover, the number of current employees shows how, to different degrees, companies expanded the initial teams.
-All companies, except C5, brought their first product to market in less than 6 months from the idea conception.
-The products consist of web-applications, launched in different nations and market sectors.
-The growing team size and the publicly available data suggest an healthy status of the businesses.
-
-We executed the case study online, supported by tools for video conferencing, recording each complete session.
-The interview subjects were the companies' CEOs or CTOs.
-We followed a step-by-step workflow, consisting of the actual interview, collection of additional material, preparation of the customized follow-up questionnaire and the iterative adjustment of the interview package artifacts\footnote{The interview package and follow-up questionnaire has been released under MIT License \cite{MITLicense} and it is available for download on Github \cite{GitHubInterviewPackage}.
-We encourage anyone interested in pushing this work forward, to fork the repository and contribute to it.
-If used in a research context, please inform us in order to be able to track where and how the interview package has been used.}.
-
-The follow-up questionnaires were designed to capture additional data, gather missing information and confirm interview results by triangulation.
-The questionnaires have been tailored to each startup, partially taking advantage of the repertory grid principles \cite{Edwards2009}.
-To obtain better chances of quickly obtaining responses, questionnaires have been sent to respondents immediately after transcribing the interview results.
-Two weeks after the conclusion of the last interview, we closed our follow-up questionnaires and collected the data.
-%MUN: I don't understand this: you
-% collected the data from questionnaires after ALL interviews have been finished, but on the other hand you 
-% said the overall process was iterative (data analysis and triangulation let themes emerge which in turn were used
-% in subsequent interviews).
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+Moreover, the number of current employees shows how, to different degrees, 
+companies expanded the initial teams. All companies, except C5, brought their 
+first product to market in less than 6 months from the idea conception. The 
+products consist of web-applications, launched in different nations and market 
+sectors. The growing team size and the publicly available data suggest an 
+healthy status of the businesses.
+
+We executed the case study online, supported by tools for video conferencing, 
+recording each complete session. The interview subjects were the companies' CEOs 
+or CTOs. We followed a step-by-step workflow, consisting of the actual 
+interview, collection of additional material, preparation of the customized 
+follow-up questionnaire and the iterative adjustment of the interview package 
+artifacts\footnote{The interview package and follow-up questionnaire has been 
+released under MIT License \cite{MITLicense} and it is available for download on 
+Github \cite{GitHubInterviewPackage}. We encourage anyone interested in pushing 
+this work forward, to fork the repository and contribute to it. If used in a 
+research context, please inform us in order to be able to track where and how 
+the interview package has been used.}.
+
+The follow-up questionnaires were designed to capture additional data, gather 
+missing information and confirm interview results by triangulation. The 
+questionnaires have been tailored to each startup, partially taking advantage of 
+the repertory grid principles \cite{Edwards2009}. To obtain better chances of 
+quickly obtaining responses, questionnaires have been sent to respondents 
+immediately after transcribing the interview results. Two weeks after the 
+conclusion of the last interview, we closed our follow-up questionnaires and 
+collected the data. %MUN: I don't understand this: you
+%collected the data from questionnaires after ALL interviews have been 
+%finished, but on the other hand you said the overall process was iterative 
+%(data analysis and triangulation let themes emerge which in turn were used
+%in subsequent interviews).
+
+%------------------------------------------------------------------------------%
 \subsection{Data collection}
-The data collection was designed to allow triangulation, which integrates multiple data sources converging on the same phenomenon.
-After transcribing the interview, we extracted conceptual relations, integrating them with the questionnaire's results.
-A well structured case-study database allowed us to easily retrieve and seek for information, assembling the evidence from the case study reports, as described also in \cite{Yin1994}.
-The database has been stored and constructed using the qualitative data analysis software package AtlasTI\footnote{Available online at \url{http://www.atlasti.com/} .} - one of the most suitable tools for GT \cite{Coleman2007}.
-We overlapped interviews with questionnaire results to adjust the data collection and take advantage of emergent themes and reveal possible inconsistencies.
-The data was analyzed simultaneously and with flexibility in mind in such a way that adjustments were made according to the emerging findings.
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+The data collection was designed to allow triangulation, which integrates 
+multiple data sources converging on the same phenomenon. After transcribing the 
+interview, we extracted conceptual relations, integrating them with the 
+questionnaire's results. A well structured case-study database allowed us to 
+easily retrieve and seek for information, assembling the evidence from the case 
+study reports, as described also in \cite{Yin1994}. The database has been stored 
+and constructed using the qualitative data analysis software package 
+AtlasTI\footnote{Available online at \url{http://www.atlasti.com/}.} - one of 
+the most suitable tools for GT \cite{Coleman2007}. 
+We overlapped interviews with questionnaire results to adjust the data 
+collection and take advantage of emergent themes and reveal possible 
+inconsistencies. The data was analyzed simultaneously and with flexibility in 
+mind in such a way that adjustments were made according to the emerging 
+findings.
+%------------------------------------------------------------------------------%
 \subsection{Data analysis}
-Before starting the analysis, a data ordering procedure was necessary since interviews were spread across a multitude of topics.
-Therefore transcripts have been structured in thematic areas accordingly to different topic cards used during the interviews.
-We proceeded horizontally between same thematic areas of transcripts to be able to identify a better number of similar concepts, rather than going through an entire transcript at one time.
-
-Once the data were ordered, we executed the process of coding interviews, following the steps listed below:
+Before starting the analysis, a data ordering procedure was necessary since 
+interviews were spread across a multitude of topics. Therefore transcripts have 
+been structured in thematic areas accordingly to different topic cards used 
+during the interviews. We proceeded horizontally between same thematic areas of 
+transcripts to be able to identify a better number of similar concepts, rather 
+than going through an entire transcript at one time.
+
+Once the data were ordered, we executed the process of coding interviews, 
+following the steps listed below:
 
@@
 
-\item Labels were assigned to raw data, and a first low-level conceptualization was carried out using both in-vivo and open coding \cite{ColinRobson2009}.
-\item Concepts were grouped together into theoretical categories and subcategories.
-By means of axial coding we first described the different relations between subcategories, and then relations between subcategories and categories.
-\item Categories were refined several times in the attempt to create different levels of abstraction and adjusting concepts, aided by a simple knowledge management tool.
-\item Consistency among categories was validated by exploring and analyzing links among subcategories by means of selective coding.
-\item The core category - the one with the greatest explanatory power -  was identified by analyzing the causal relations between high-level categories.
+\item Labels were assigned to raw data, and a first low-level conceptualization 
+was carried out using both in-vivo and open coding \cite{ColinRobson2009}.
+\item Concepts were grouped together into theoretical categories and 
+subcategories. By means of axial coding we first described the different 
+relations between subcategories, and then relations between subcategories and 
+categories.
+\item Categories were refined several times in the attempt to create different 
+levels of abstraction and adjusting concepts, aided by a simple knowledge 
+management tool.
+\item Consistency among categories was validated by exploring and analyzing 
+links among subcategories by means of selective coding.
+\item The core category - the one with the greatest explanatory power - was 
+identified by analyzing the causal relations between high-level categories.
 \end{compactitem}
 
-During data extraction we used the technique of in-vivo coding since \textit{``direct quoting from the transcript gives more expressive power to the data''} \cite{ColinRobson2009} combined with the more descriptive procedure of open coding.
-Following the example of other grounded theories, developed in related areas such as Information Systems \cite{Orlikowski1993} and Software Process Improvement \cite{Coleman2006}, we performed the high-level conceptualization during creation of categories, in the process of refining axial and selective coding.
-
-As we were iterating through the interviews, we analyzed new data accordingly, by updating codes and categories on necessity, and taking notes in the form of memos to adjust the new emerging theory.
-To improve the speed of the coding process, we analyzed all the transcripts one part at-the-time: transcripts followed the semi-structured interview format, and this approach enhanced the chances of finding similar codes across different transcripts, allowing us to quickly iterate and update categories.
-
-Given the large number of codes, categories, properties, propositions and related questions that evolved from the analytical process \cite{Corbin1990}, an important activity that helped the coding process was memoing.
-It constituted an important component involved in the formulation and revision of theory during the research process.
-For this purpose we made use of three different types of memos: code memos, theoretical memos and operational memos.
-The first type of memo is related to the conceptual labeling of open codes, whilst the second type concerns axial and selective coding, and thus focusing on paradigm features.
-Finally, operational memos contain directions relating to the evolving emerging theory.
-
-After the coding process the first representation of the experience map identified in GT is constructed by means of a theoretical framework.
-The theoretical framework is presented in the form of a network of categories and subcategories that are linked together according to cause-effect relationships \cite{Corbin1990}.
-During the formation of the theoretical framework the researchers operated by means of a  bottom-up approach.
-From the empirical data and coding process, the framework developed into two different levels: a detailed level representing the network of subcategories (identified mainly by the axial coding process), and a high-level representation of the main categories network (identified mainly by the selective coding process).
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+During data extraction we used the technique of in-vivo coding since 
+\textit{``direct quoting from the transcript gives more expressive power to the 
+data''} \cite{ColinRobson2009} combined with the more descriptive procedure of 
+open coding. Following the example of other grounded theories, developed in 
+related areas such as Information Systems \cite{Orlikowski1993} and Software 
+Process Improvement \cite{Coleman2006}, we performed the high-level 
+conceptualization during creation of categories, in the process of refining 
+axial and selective coding.
+
+As we were iterating through the interviews, we analyzed new data accordingly, 
+by updating codes and categories on necessity, and taking notes in the form of 
+memos to adjust the new emerging theory. To improve the speed of the coding 
+process, we analyzed all the transcripts one part at-the-time: transcripts 
+followed the semi-structured interview format, and this approach enhanced the 
+chances of finding similar codes across different transcripts, allowing us to 
+quickly iterate and update categories.
+
+Given the large number of codes, categories, properties, propositions and 
+related questions that evolved from the analytical process \cite{Corbin1990}, 
+an important activity that helped the coding process was memoing. It constituted 
+an important component involved in the formulation and revision of theory 
+during the research process. For this purpose we made use of three different 
+types of memos: code memos, theoretical memos and operational memos. The first 
+type of memo is related to the conceptual labeling of open codes, whilst the 
+second type concerns axial and selective coding, and thus focusing on paradigm 
+features. Finally, operational memos contain directions relating to the evolving 
+emerging theory.
+
+After the coding process the first representation of the experience map 
+identified in GT is constructed by means of a theoretical framework. The 
+theoretical framework is presented in the form of a network of categories and 
+subcategories that are linked together according to cause-effect relationships 
+\cite{Corbin1990}. During the formation of the theoretical framework the 
+researchers operated by means of a  bottom-up approach. From the empirical data 
+and coding process, the framework developed into two different levels: a 
+detailed level representing the network of subcategories (identified mainly by 
+the axial coding process), and a high-level representation of the main 
+categories network (identified mainly by the selective coding process).
+
+%------------------------------------------------------------------------------%
 \subsection{Theory generation}
 
-As mentioned in Section \ref{desex}, emergent theories have been tested by integrating additional companies into the sample, selected following the principle of theoretical sampling \cite{Yin1994}.
-
-The process of theory generation took place at each iteration together with interview execution, where we systematically analyzed the theoretical framework by means of the paradigm model introduced by Corbin \cite{Corbin1990}.
-The paradigm model is composed by:
+As mentioned in Section \ref{desex}, emergent theories have been tested by 
+integrating additional companies into the sample, selected following the 
+principle of theoretical sampling \cite{Yin1994}.
+
+The process of theory generation took place at each iteration together with 
+interview execution, where we systematically analyzed the theoretical framework 
+by means of the paradigm model introduced by Corbin \cite{Corbin1990}. The 
+paradigm model is composed by:
 
 \begin{compactitem}
-\item \textit{Causal conditions}: the events which lead to the occurrence of the phenomenon, that is our core category
-\item \textit{Context}: set of conditions in which the phenomenon can be extrapolated
-\item \textit{Intervening conditions}: the broader set of conditions from that the phenomenon can be generalized
-\item \textit{Action/interaction strategies}: the actions and responses that occur as the result of the phenomenon
-\item \textit{Consequences}: specification of the outcomes, both intended and unintended of the actions and interaction strategies
+\item \textit{Causal conditions}: the events which lead to the occurrence of 
+the phenomenon, that is our core category
+\item \textit{Context}: set of conditions in which the phenomenon can be 
+extrapolated
+\item \textit{Intervening conditions}: the broader set of conditions from that 
+the phenomenon can be generalized
+\item \textit{Action/interaction strategies}: the actions and responses that 
+occur as the result of the phenomenon
+\item \textit{Consequences}: specification of the outcomes, both intended and 
+unintended of the actions and interaction strategies
 \end{compactitem}
 
-Within the limits of the critical bounding assumptions, the role of the generated theory is to explain, predict and understand the underlying phenomenon.
+Within the limits of the critical bounding assumptions, the role of the 
+generated theory is to explain, predict and understand the underlying 
+phenomenon.
 
 \subsection{Theory Validation}
-%MUN: put here what I wrote in the comment in Section "Theory validation" (i.e. the section following the GSM)
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+%MUN: put here what I wrote in the comment in Section "Theory validation" (i.e. 
+%the section following the GSM)
+
+%------------------------------------------------------------------------------%
 
@@
 
-By analyzing the interview transcripts with the GT process we extracted the information necessary to form the theoretical model.
-The initial coding process (open and in-vivo) provided 630 unique codes \footnote{Available at \url{https://github.com/adv0r/BTH-Interview-Package}.}, grounded to 1295 citations in the transcripts, divided according to the thematic areas.
-The thematic areas were divided according to the development phases designed for the interview process.
-The raw codes constitute the empirical foundation of the theoretical categories, which has been used to form the Green-field startup model (GSM).
-
-The GSM captures the underlying phenomenon of software development in early-stage startups.
-The framework is formed by 128 sub-categories, clustered in 35 groups, and finally in 7 categories at the highest level of abstraction.
-By the means of the GSM we provide explanations of the development strategies and engineering activities undertaken by startups.
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+By analyzing the interview transcripts with the GT process we extracted the 
+information necessary to form the theoretical model. The initial coding process 
+(open and in-vivo) provided 630 unique codes\footnote{Available at 
+\url{https://github.com/adv0r/BTH-Interview-Package}.}, grounded to 1295 
+citations in the transcripts, divided according to the thematic areas. The 
+thematic areas were divided according to the development phases designed for the 
+interview process. The raw codes constitute the empirical foundation of the 
+theoretical categories, which has been used to form the Green-field startup 
+model (GSM).
+
+The GSM captures the underlying phenomenon of software development in 
+early-stage startups. The framework is formed by 128 sub-categories, clustered 
+in 35 groups, and finally in 7 categories at the highest level of abstraction. 
+By the means of the GSM we provide explanations of the development strategies 
+and engineering activities undertaken by startups.
+%------------------------------------------------------------------------------%
 \subsection{Framework overview}
 \label{res:gsm:frmov}
-The main concepts representing the underlying phenomenon have been grouped together to form high-level categories.
-Figure \ref{fig:gsm} shows the network of causal relationships (represented by arrows) between categories (represented by blocks).
-Each node is linked to subsequent nodes which are called successors, and preceding nodes, called predecessors.
-The relations between them are denoted by the direction of the interconnecting arrows.
-The network is read from left to right.
+The main concepts representing the underlying phenomenon have been grouped 
+together to form high-level categories. Figure \ref{fig:gsm} shows the network 
+of causal relationships (represented by arrows) between categories (represented 
+by blocks). Each node is linked to subsequent nodes which are called successors, 
+and preceding nodes, called predecessors. The relations between them are denoted 
+by the direction of the interconnecting arrows. The network is read from left to 
+right.
 
@@
 
-In forthcoming the explanation of the GSM we emphasize labels of theoretical categories using the \textit{italic} font style.
-For the main categories shown in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e.
+In forthcoming the explanation of the GSM we emphasize labels of theoretical 
+categories using the \textit{italic} font style. For the main categories shown 
+in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e. 
 CATx) to facilitate the reading of this section.
-The network is centered around the core category, \textit{speed-up development}, which is the most interconnected node in the framework reflecting the fact that \textit{``it is the one (category) with the greatest explanatory power''}  \cite{Strauss1998}.
-The meaning of the labels on the three arrows that reach the core category (efficiency, effectiveness and performance) is discussed in detail in section \ref{valt}.
-%MUN: this refers to the validity threats section.
-Do you mean rather the "Theory Validation" section?
-
-A contextual condition which characterizes, to some extent, each and every startup, is the \textit{severe lack of resources}.
-In fact, the capabilities of an early-stage startup to support its development activities are constrained by an extremely limited access to human, time and intellectual resources, eventually leading to a general absence of structures and processes.
-
-The \textit{severe lack of resources} forces the company to focus on implementing an essential set of functionalities and it is one of the main reasons why the \textit{product quality has low priority} respect to other more urgent needs\footnote{There are some exceptions where the quality aspects actually matter and such cases will be discussed in section \ref{valt} %MUN: this refers to the validity threats section.
-Do you mean rather the "Theory Validation" section?
-.}. At the same time, to be able to deal with such constraints, startups must depend on a small group of capable and motivated individuals.
-
-As it has been unanimously expressed by respondents, from a software perspective, the most urgent priority is to \textit{speed-up the development} as much as possible by adopting an extremely flexible and effective \textit{evolutionary approach}.
-On the other hand the efficiency of teamwork is facilitated by the low attention initially given to architectural aspects related to product quality.
-This allows startups to have a faulty but functioning product, which can quickly be introduced to the market beginning with the in-house implementation of the prototype from day-one.
-
-The initial employees are the ingredients which enable high levels of performance in software development.
-To support a fast-paced production environment, engineers are required to be highly committed, co-located, multi-role, and self-organized.
-In other words \textit{team is the catalyst of development}.
-With an essential and flexible workflow, which relies on tacit knowledge instead of formal documentation, startups are able to achieve very short time-to-market cycles.
-However, each line of code, written without following structures and processes, contributes to grow the \textit{accumulated technical debt}, which is further increased by having almost non-existing specifications, a minimal project management and a lack of automated tests.
-
-Despite the fact that the consequences of such debt are not clearly perceived in the initial stages (where finding the product/market fit as quickly as possible is the most important priority), startups, which survive to subsequent phases, will likely increment their user-base size, the product size, and the number of developers.
-This will require the company to eventually pay the \textit{accumulated technical debt}, confronting the fact that an \textit{initial growth hinders productivity}.
-
-In the following subsections we are going to explain in detail all the categories presented in Figure \ref{fig:gsm}, and conclude in subsection \ref{res:gsm:th} with the final theory.
-In the explanations we use identifiers of the companies presented in Table \ref{t_interviews-stats} (i.e.
-Cx) to highlight statements recorded from the interviewees.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+The network is centered around the core category, \textit{speed-up 
+development}, which is the most interconnected node in the framework reflecting 
+the fact that \textit{``it is the one (category) with the greatest explanatory 
+power''}  \cite{Strauss1998}. The meaning of the labels on the three arrows that 
+reach the core category (efficiency, effectiveness and performance) is discussed 
+in detail in section \ref{valt}. %MUN: this refers to the validity threats 
+%section. Do you mean rather the "Theory Validation" section?
+
+A contextual condition which characterizes, to some extent, each and every 
+startup, is the \textit{severe lack of resources}. In fact, the capabilities of 
+an early-stage startup to support its development activities are constrained by 
+an extremely limited access to human, time and intellectual resources, 
+eventually leading to a general absence of structures and processes.
+
+The \textit{severe lack of resources} forces the company to focus on 
+implementing an essential set of functionalities and it is one of the main 
+reasons why the \textit{product quality has low priority} respect to other more 
+urgent needs\footnote{There are some exceptions where the quality aspects 
+actually matter and such cases will be discussed in section \ref{valt}%MUN: 
+%this refers to the validity threats section. Do you mean rather the "Theory 
+%Validation" section?
+.}. At the same time, to be able to deal with such constraints, startups must 
+depend on a small group of capable and motivated individuals.
+
+As it has been unanimously expressed by respondents, from a software 
+perspective, the most urgent priority is to \textit{speed-up the development} as 
+much as possible by adopting an extremely flexible and effective 
+\textit{evolutionary approach}. On the other hand the efficiency of teamwork is 
+facilitated by the low attention initially given to architectural aspects 
+related to product quality. This allows startups to have a faulty but 
+functioning product, which can quickly be introduced to the market beginning 
+with the in-house implementation of the prototype from day-one.
+
+The initial employees are the ingredients which enable high levels of 
+performance in software development. To support a fast-paced production 
+environment, engineers are required to be highly committed, co-located, 
+multi-role, and self-organized. In other words \textit{team is the catalyst of 
+development}.
+With an essential and flexible workflow, which relies on tacit knowledge 
+instead of formal documentation, startups are able to achieve very short 
+time-to-market cycles. However, each line of code, written without following 
+structures and processes, contributes to grow the \textit{accumulated technical 
+debt}, which is further increased by having almost non-existing specifications, 
+a minimal project management and a lack of automated tests.
+
+Despite the fact that the consequences of such debt are not clearly perceived 
+in the initial stages (where finding the product/market fit as quickly as 
+possible is the most important priority), startups, which survive to subsequent 
+phases, will likely increment their user-base size, the product size, and the 
+number of developers. This will require the company to eventually pay the 
+\textit{accumulated technical debt}, confronting the fact that an 
+\textit{initial growth hinders productivity}.
+
+In the following subsections we explain in detail all the categories presented 
+in Figure \ref{fig:gsm}, and conclude in subsection \ref{res:gsm:th} with the 
+final theory. In the explanations we use identifiers of the companies presented 
+in Table \ref{t_interviews-stats} (i.e. Cx) to highlight statements recorded 
+from the interviewees.
+
+%------------------------------------------------------------------------------%
 \subsection{Severe lack of resources}
 \label{res:gsm:cat7}
-The concept of \textit{severe lack of resources} characterizes the uncertainty of the development strategies in startups and it is composed by three subcategories: \textit{time-shortage}, \textit{limited human resources} and \textit{limited access to expertise}.
-
-Since startups want to bring the product to market as quickly as possible, the resource they are the most deprived of is time.
-Startups operate under a constant time pressure, mainly generated by external sources (\textit{investor(s) pressure}, \textit{business pressure}) and sometimes internal necessities such as \textit{internal deadlines} and \textit{demo presentations at events}.
-In this regard, C3 commented: \textit{``Investors wanted to see product features, engineers wanted to make them better.
-Finally the time-to-market was considered more important and the team interests was somehow sacrificed.''}
-
-In addition, to compensate the \textit{limited human resources},  practitioners need to empower \textit{multi-role and full stack engineers}, as confirmed by C1: \textit{``Everyone was involved in any tasks, from mobile to web development, organizing themselves in choosing the part to implement''}.
-
-The extent to which startups have access to specialized knowledge - both internal and external to the company - is extremely reduced when compared to traditional established software companies.
-Therefore, in order to partially mitigate the \textit{limited access to expertise}, startups rely on the external aid of mentors or advisors.
-Under these strict limitations, most of the decisions related to software development are fundamentally trade-off situations (confirming Himola's results \cite{Hilmola2003}).
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+The concept of \textit{severe lack of resources} characterizes the uncertainty 
+of the development strategies in startups and it is composed by three 
+subcategories: \textit{time-shortage}, \textit{limited human resources} and 
+\textit{limited access to expertise}.
+
+Since startups want to bring the product to market as quickly as possible, the 
+resource they are the most deprived of is time. Startups operate under a 
+constant time pressure, mainly generated by external sources 
+(\textit{investor(s) pressure}, \textit{business pressure}) and sometimes 
+internal necessities such as \textit{internal deadlines} and \textit{demo 
+presentations at events}. In this regard, C3 commented: \textit{``Investors 
+wanted to see product features, engineers wanted to make them better. Finally 
+the time-to-market was considered more important and the team interests was 
+somehow sacrificed.''}
+
+In addition, to compensate the \textit{limited human resources},  practitioners 
+need to empower \textit{multi-role and full stack engineers}, as confirmed by 
+C1: \textit{``Everyone was involved in any tasks, from mobile to web 
+development, organizing themselves in choosing the part to implement''}.
+
+The extent to which startups have access to specialized knowledge - both 
+internal and external to the company - is extremely reduced when compared to 
+traditional established software companies. Therefore, in order to partially 
+mitigate the \textit{limited access to expertise}, startups rely on the external 
+aid of mentors or advisors. Under these strict limitations, most of the 
+decisions related to software development are fundamentally trade-off 
+situations (confirming Himola's results \cite{Hilmola2003}).
+%------------------------------------------------------------------------------%
 \subsection{Team is the catalyst of development}
 \label{res:gsm:cat4}
-Among the different aspects fostering the speed of the development process, the startups' main focus is on the characteristics of the initial team.
-
-In startups \textit{developers have big responsibilities}.
-In fact, the \textit{limited human resources}, discussed in CAT7, cause the team-members to be active in every aspect of the development process, from the definition of functionalities to the final deployment.
-
-Early engineers in the founding team of a startups are typically \textit{multi-role and full-stack engineers}: they are full-stack for being able to tackle different problems at different levels of the technology stack (\textit{generalists developers instead of specialists}) and multi-role since usually \textit{engineers are responsible for marketing/sales/development}.
-As remarked by  C11: \textit{``Instead of hiring gurus in one technology, startups should hire young developers, generalists, that know how to quickly learn new technologies, and quickly move among a huge variety of tasks.''}
-
-Moreover, having a \textit{very small and co-located development team} enables members to operate with high coordination, control and communication, which heavily rely on \textit{tacit knowledge},  replacing most of the documentation with informal discussions.
-Practitioners reported that keeping the development team small helps startups in being fast and flexible, as remarked by C8: \textit{``If you have more than 10 people, it is absolutely impossible to be fast''}.
-Then, also \textit{previous working experience} and \textit{knowing each other before starting the company} enforce the efficiency of activities by \textit{limiting the need of formalities between team-members}.
-
-In every software company, \textit{skilled developers are essential for high speed} in development.
-Especially in startups, the ``hacking culture'' and a tendency to the ``just-do-it'' approach allow the team to quickly move from the formulation of a feature idea to its implementation.
-In this regard, C1 comments: \textit{``We had a hacker culture/environment, people hacking stuff without formally analyzing it, but breaking it down and finding a way around.
+Among the different aspects fostering the speed of the development process, the 
+startups' main focus is on the characteristics of the initial team.
+
+In startups \textit{developers have big responsibilities}. In fact, the 
+\textit{limited human resources}, discussed in CAT7, cause the team-members to 
+be active in every aspect of the development process, from the definition of 
+functionalities to the final deployment.
+
+Early engineers in the founding team of a startups are typically 
+\textit{multi-role and full-stack engineers}: they are full-stack for being able 
+to tackle different problems at different levels of the technology stack 
+(\textit{generalists developers instead of specialists}) and multi-role since 
+usually \textit{engineers are responsible for marketing/sales/development}. As 
+remarked by  C11: \textit{``Instead of hiring gurus in one technology, startups 
+should hire young developers, generalists, that know how to quickly learn new 
+technologies, and quickly move among a huge variety of tasks.''}
+
+Moreover, having a \textit{very small and co-located development team} enables 
+members to operate with high coordination, control and communication, which 
+heavily rely on \textit{tacit knowledge},  replacing most of the documentation 
+with informal discussions. Practitioners reported that keeping the development 
+team small helps startups in being fast and flexible, as remarked by C8: 
+\textit{``If you have more than 10 people, it is absolutely impossible to be 
+fast''}. Then, also \textit{previous working experience} and \textit{knowing 
+each other before starting the company} enforce the efficiency of activities by 
+\textit{limiting the need of formalities between team-members}.
+
+In every software company, \textit{skilled developers are essential for high 
+speed} in development. Especially in startups, the ``hacking culture'' and a 
+tendency to the ``just-do-it'' approach allow the team to quickly move from the 
+formulation of a feature idea to its implementation. In this regard, C1 
+comments: \textit{``We had a hacker culture/environment, people hacking stuff 
+without formally analyzing it, but breaking it down and finding a way around. 
 Ticket on trello and go.''}
 
-A \textit{limited access to expertise} forces the team to rely mainly on their personal abilities, even though asking opinions to mentors is reported to be a viable practice to maintain feasible objectives.
-Furthermore \textit{teams work under constant pressure}, \textit{``up to 16 hours per day''}, %MUN: add Cx to this statement since it is a citation?
+A \textit{limited access to expertise} forces the team to rely mainly on their 
+personal abilities, even though asking opinions to mentors is reported to be a 
+viable practice to maintain feasible objectives. Furthermore \textit{teams work 
+under constant pressure}, \textit{``up to 16 hours per day''}, %MUN: add Cx to 
+%this statement since it is a citation? 
 mainly constrained by a tight \textit{time shortage}.
 
-Finally, startups present founders-centric structures, and especially in the early-stage, it is clear that \textit{CTO/CEO background has high-impact} on the company's development approach.
-For instance, in case of an academic background, the CTO encourages the introduction of some architectural design before the development phase.
-Moreover, even though the CTO/CEO starts guiding initially the development process, most of the decision are taken consensually by all members of the team.
-Then, the CTO/CEO decides only in particular situations where some conflicts occur.
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+Finally, startups present founders-centric structures, and especially in the 
+early-stage, it is clear that \textit{CTO/CEO background has high-impact} on the 
+company's development approach. For instance, in case of an academic background, 
+the CTO encourages the introduction of some architectural design before the 
+development phase. Moreover, even though the CTO/CEO starts guiding initially 
+the development process, most of the decision are taken consensually by all 
+members of the team. Then, the CTO/CEO decides only in particular situations 
+where some conflicts occur.
+%------------------------------------------------------------------------------%
 \subsection{Evolutionary approach}
 \label{res:gsm:cat2}
-Among different approaches to development, startups prefer to build an initial prototype and iteratively refine it over time, similarly to the concept of \textit{``evolutionary prototyping''} \cite{EvProt}.
-The reason of using this approach is that startups want to validate the product against the market as soon as possible, finding the proper product/market fit.
-Indeed, they can focus on developing only parts of the system they want to validate instead of working on developing a whole new system.
-Then, as the prototype is released, users detect opportunities for new functionalities and improvements, providing their feedback to developers.
-
-Since \textit{flexibility and reactiveness are the main priorities}, the most suitable class of software development approaches are clearly highly evolutionary in nature.
-In fact, as \textit{uncertain conditions make long-term planning not viable}, startups cannot base their work on assumptions without rapidly validating them, releasing the product to market.
-The uncertainty is first of all in the team composition.
-Since the teams are typically very small and the project knowledge is not documented, even a little change in their composition (e.g.
-a developer falls ill) can have a very high impact on the overall product development.
-Furthermore, startups operate in a continuous evolving environment in terms of competitors and targeted market sectors.
-Then, to obtain a competitive advantage in the market, startups typically make use of cutting-edge solutions, characterised by an evolution that cannot be foreseen in the long run.
-However, a special role in daily decisions is covered by user feedback
-and requests, which represent the main drivers for defining the product features in the short-term.
-
-To obtain fast user responses and quickly \textit{validate the product}, startups \textit{build a functioning prototype and iterate on it} over time.
-Paraphrasing C4, \textit{``[\ldots] you should start with something that is really rough and then polish it, fix it and iterate.
-We were under constant pressure.
-The aim was to understand as soon as possible the product market/fit iterating quickly, adjusting the product and releasing fast.''}
-
-The companies focus on building a small set of functionalities to include in the first version, and \textit{progressively roll-out to a larger number of people} with very \textit{small iterations} (confirmed again by C4: \textit{``we deploy from 5 to 20 times a day''}).
-
-The objective of this evolutionary approach is to avoid wasting time in ``over-engineering the system'' and building complex functionalities that have not been tested on real users.
-By releasing a very small number of good-enough functionalities (see CAT3) the startup can verify the suitability of the features and understand how to adjust the direction of product development towards actual users' needs.
-The first version of the product is typically a prototype which contains the basic functionalities developed with the least possible effort that can validate critical features, enabling the startup's survival in the short term (recalling  Ries' definition of \textit{Minimum Viable Product} (MVP) \cite{Ries2011}).
-
-Supported by \textit{direct contact and observation of users}, \textit{automated feedback collection} and \textit{analysis of product metrics}, startups attempt to \textit{find what is valuable for customers} with activities in agreement with Blank's \textit{Customer validation} stage \cite{Blank2005}.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+Among different approaches to development, startups prefer to build an initial 
+prototype and iteratively refine it over time, similarly to the concept of 
+\textit{``evolutionary prototyping''} \cite{EvProt}. The reason of using this 
+approach is that startups want to validate the product against the market as 
+soon as possible, finding the proper product/market fit. Indeed, they can focus 
+on developing only parts of the system they want to validate instead of working 
+on developing a whole new system. Then, as the prototype is released, users 
+detect opportunities for new functionalities and improvements, providing their 
+feedback to developers.
+
+Since \textit{flexibility and reactiveness are the main priorities}, the most 
+suitable class of software development approaches are clearly highly 
+evolutionary in nature. In fact, as \textit{uncertain conditions make long-term 
+planning not viable}, startups cannot base their work on assumptions without 
+rapidly validating them, releasing the product to market. The uncertainty is 
+first of all in the team composition. Since the teams are typically very small 
+and the project knowledge is not documented, even a little change in their 
+composition (e.g. a developer falls ill) can have a very high impact on the 
+overall product development. Furthermore, startups operate in a continuous 
+evolving environment in terms of competitors and targeted market sectors. Then, 
+to obtain a competitive advantage in the market, startups typically make use of 
+cutting-edge solutions, characterised by an evolution that cannot be foreseen in 
+the long run. However, a special role in daily decisions is covered by user 
+feedback and requests, which represent the main drivers for defining the product 
+features in the short-term.
+
+To obtain fast user responses and quickly \textit{validate the product}, 
+startups \textit{build a functioning prototype and iterate on it} over time. 
+Paraphrasing C4, \textit{``[\ldots] you should start with something that is 
+really rough and then polish it, fix it and iterate. We were under constant 
+pressure. The aim was to understand as soon as possible the product market/fit 
+iterating quickly, adjusting the product and releasing fast.''}
+
+The companies focus on building a small set of functionalities to include in 
+the first version, and \textit{progressively roll-out to a larger number of 
+people} with very \textit{small iterations} (confirmed again by C4: \textit{``we 
+deploy from 5 to 20 times a day''}).
+
+The objective of this evolutionary approach is to avoid wasting time in 
+``over-engineering the system'' and building complex functionalities that have 
+not been tested on real users. By releasing a very small number of good-enough 
+functionalities (see CAT3) the startup can verify the suitability of the 
+features and understand how to adjust the direction of product development 
+towards actual users' needs. The first version of the product is typically a 
+prototype which contains the basic functionalities developed with the least 
+possible effort that can validate critical features, enabling the startup's 
+survival in the short term (recalling  Ries' definition of \textit{Minimum 
+Viable Product} (MVP) \cite{Ries2011}).
+
+Supported by \textit{direct contact and observation of users}, 
+\textit{automated feedback collection} and \textit{analysis of product metrics}, 
+startups attempt to \textit{find what is valuable for customers} with activities 
+in agreement with Blank's \textit{Customer validation} stage \cite{Blank2005}.
+
+%------------------------------------------------------------------------------%
 \subsection{Product quality has low priority}
 \label{res:gsm:cat3}
-The interests of software startups, related to the product, are mainly concentrated on building a  \textit{limited number of suitable functionalities} rather than fulfilling restrictive non-functional requirements.
-This strategy allows them to quickly release simple products with less need of preliminary architectural studies.
-
-Exploring quality aspects considered during the development process, the only important concerns are expressed towards the user experience\footnote{According to the ISO 9241-210 (Ergonomics of human-system interaction), UX is defined as \textit{``a person's perceptions and responses that result from the use or anticipated use of a product, system or service''.} Even though this definition leaves so much to interpretation, we can refer to it as presented in the glossary.}, in terms of \textit{ease of use, attractiveness of the UI} and most of all, \textit{smooth user-flow without interruptions}.
-The fact that \textit{UX is the only important quality} is remarked by C11: \textit{``When a user needs to think too much on what action should be done next, he will just close the application without returning''} and confirmed by C3: \textit{``If the product works, but it is not usable, it doesn't work''.}
-
-The extent to which quality aspects are taken into account might heavily depend upon the market sector and the type of application.
-Nevertheless, realizing high level of UX  is often the most important attribute to consider for customer discovery of  evolutionary approaches in view of the \textit{limited human resources} and \textit{time shortage}, presented in CAT7.
-C4 confirms: \textit{``None of the quality aspects matter that much as the development speed does.''}
-
-To achieve a good level of UX while dealing with lack of human resources and time shortages, startups can analyze similar products of bigger companies which can afford more rigorous usability studies.
-Then, the users' feedback and product metrics begin to have a central role in determining the level of UX achieved.
-Product metrics are typically web-based \textit{statistical hypothesis testing}, such as A/B testing.
-%MUN: add reference or explain shortly what A/B testing is
-
-Other than UX, some other factors can influence the quality concerns of development:
+The interests of software startups, related to the product, are mainly 
+concentrated on building a  \textit{limited number of suitable functionalities} 
+rather than fulfilling restrictive non-functional requirements. This strategy 
+allows them to quickly release simple products with less need of preliminary 
+architectural studies.
+
+Exploring quality aspects considered during the development process, the only 
+important concerns are expressed towards the user experience\footnote{According 
+to the ISO 9241-210 (Ergonomics of human-system interaction), UX is defined as 
+\textit{``a person's perceptions and responses that result from the use or 
+anticipated use of a product, system or service''.} Even though this definition 
+leaves so much to interpretation, we can refer to it as presented in the 
+glossary.}, in terms of \textit{ease of use, attractiveness of the UI} and most 
+of all, \textit{smooth user-flow without interruptions}. The fact that 
+\textit{UX is the only important quality} is remarked by C11: \textit{``When a 
+user needs to think too much on what action should be done next, he will just 
+close the application without returning''} and confirmed by C3: \textit{``If the 
+product works, but it is not usable, it doesn't work''.}
+
+The extent to which quality aspects are taken into account might heavily depend 
+upon the market sector and the type of application. Nevertheless, realizing high 
+level of UX  is often the most important attribute to consider for customer 
+discovery of  evolutionary approaches in view of the \textit{limited human 
+resources} and \textit{time shortage}, presented in CAT7. C4 confirms: 
+\textit{``None of the quality aspects matter that much as the development speed 
+does.''}
+
+To achieve a good level of UX while dealing with lack of human resources and 
+time shortages, startups can analyze similar products of bigger companies which 
+can afford more rigorous usability studies. Then, the users' feedback and 
+product metrics begin to have a central role in determining the level of UX 
+achieved. Product metrics are typically web-based \textit{statistical hypothesis 
+testing}, such as A/B testing. %MUN: add reference or explain shortly what A/B 
+%testing is
+
+Other than UX, some other factors can influence the quality concerns of 
+development:
 
@@
 
-\item The \textit{efficiency emerges after using the product}, letting engineers avoid wasting time in excessive improvements of not-validated functionalities.
-The level of efficiency can be optimized after attesting the effectiveness of the minimal set of functionalities, obtained according to CAT2 and to the concept of MVP.
-\item The \textit{product should be reasonably ready-to-scale} to be able to accommodate a potential growth of the user-base\footnote{Note that startups tackle fast growing markets which are particularly subject to sudden user growths}.
-Thanks to modern cloud services, startups can \textit{externalize complexity to third party solutions} achieving a good level of scalability with a reasonable effort.
-\item Realizing high reliability is not an urgent priority since \textit{users are fault-tolerant in innovative beta products}.
-In these cases, users have typically a positive attitude towards the product, even though it presents unreliable behaviours.
-In this regard, the main focus of beta testing is reducing frictions between the product and the users, often incorporating usability testing.
-In fact, the beta release is typically the first time that the software is available outside of the organization that developed it.
-\footnote{Detailed discussion of the impact of innovative products on the user satisfaction is presented in Appendix according to the Kano model.}.
-%MUN: do you use the Kano model in Theory Validation? Change text to refer to the correct section.
+\item The \textit{efficiency emerges after using the product}, letting 
+engineers avoid wasting time in excessive improvements of not-validated 
+functionalities. The level of efficiency can be optimized after attesting the 
+effectiveness of the minimal set of functionalities, obtained according to CAT2 
+and to the concept of MVP.
+\item The \textit{product should be reasonably ready-to-scale} to be able to 
+accommodate a potential growth of the user-base\footnote{Note that startups 
+tackle fast growing markets which are particularly subject to sudden user 
+growths}. Thanks to modern cloud services, startups can \textit{externalize 
+complexity to third party solutions} achieving a good level of scalability with 
+a reasonable effort.
+\item Realizing high reliability is not an urgent priority since \textit{users 
+are fault-tolerant in innovative beta products}. In these cases, users have 
+typically a positive attitude towards the product, even though it presents 
+unreliable behaviours. In this regard, the main focus of beta testing is 
+reducing frictions between the product and the users, often incorporating 
+usability testing. In fact, the beta release is typically the first time that 
+the software is available outside of the organization that developed it. 
+\footnote{Detailed discussion of the impact of innovative products on the user 
+satisfaction is presented in Appendix according to the Kano model.}.
+%MUN: do you use the Kano model in Theory Validation? Change text to refer to 
+%the correct section. [After I've read the respective section, I'd suggest to 
+%remove the Kano discussion altogether to shorten the paper. Keep the footnote, 
+%reference the correct section, but remove "according to the Kano model]
 \end{compactitem}
 
-Finally, startups implement a \textit{limited number of suitable functionalities}, which in view of \textit{limited human resources} and \textit{time shortage}, represent a viable strategy to focus on shortening time-to-market.%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+Finally, startups implement a \textit{limited number of suitable 
+functionalities}, which in view of \textit{limited human resources} and 
+\textit{time shortage}, represent a viable strategy to focus on shortening 
+time-to-market.
+%------------------------------------------------------------------------------%
 \subsection{Speed-up development}
 \label{res:gsm:cat1}
-As remarked initially, \textit{speed-up development} represents the core category of the GSM.
-Firmly grounded as the primary objective of startups, it shows the most important characteristic of developing software in the early stage.
-
-As mentioned before, in order to \textit{speed-up development}, startups adopt evolutionary approaches supported by a solid team focusing on implementing a minimal set of suitable functionalities.
-Startups \textit{keep simple and informal workflows} to be flexible and reactive, adapting to the fast changing environment.
-While a rigorously defined workflow generally consists of an established sequence of well-defined steps to follow, startups, by contrast, adopt informal workflows.
-This choice is mainly dictated by contextual factors characterized by the general conditions of unpredictability and uncertainty.
-The adoption of informal workflows is facilitated by the fact that teams are typically self-organized and \textit{developers have big responsibilities} .
-
-Additionally, possible planning activities are refrained by the aim to shorten time-to-market, as reported by C8: \textit{``Speed was the essence so we didn't plan out too many details''}.
-To deal with such unpredictability, startups prefer to take decisions as fast as possible, mainly by means of informal and frequent verbal discussions.
-
-Despite Agile principles appear to be suitable to embrace changes, especially in the early-stage, development practices are often perceived as time-wasters and ignored to accommodate the need of releasing the product to market as quickly as possible.
-In this regard, to maintain the simplest and most informal approach, startups aim to \textit{find the product/market fit quickly} and implement a \textit{limited number of suitable functionalities}.
-This approach is considered possible also in view of a lack of systematic quality assurance activities, since only the user experience is considered as important and other quality aspects, such as efficiency, can be postponed after the first release.
-
-Another strategy, which supports startups in quickly delivering products, is the \textit{externalization of complexity on third party solutions} by: making use of third party components (COTS) and open source solutions (both for product components, and development tools and libraries); taking advantage of  external services; and deploying the product on external infrastructures, for the sake of delivering a \textit{product reasonably ready to scale} for a possible future growth.
-
-Even though \textit{the use of well-integrated and simple tools} allows startups to automate many activities and reduce their completion time, drawbacks of such approaches are the increased concerns of interoperability issues.
-A category of tools is represented by advanced version control systems, which are not only used to manage the code-base, but for many other purposes such as: assigning tasks; trace responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.
-
-Startups can further improve development speed by making \textit{use of standards and known technologies} which are widely recognized, well tested, and supported by strong communities.
-Moreover, the use of widely adopted standards and frameworks reduces the need of a formal architectural design since most of implementation solutions are well documented and ready-to-use.
-This is confirmed by C1: \textit{``as long as you use Ruby standards with the Rails framework, the language is clean itself and doesn't need much documentation''.}
-
-Other important factors that positively impact the speed of development are the team's desire to: \textit{create disruptive technologies}; to \textit{demonstrate personal abilities}; and to \textit{have the product used in the market}.
-As reported by practitioners, these factors are essential to enhance the morale of developers and therefore to achieve higher team performances.
-By contrast, especially in the typical sprint-based environments of Agile, when a developer is not \textit{able to meet deadlines} the morale goes down, hindering the development speed.
-
-Finally the constant pressure under which the company regularly operates, leads the team to often \textit{work overtime to meet deadlines}.
-But as reported by practitioners, such way of working can be an effective strategy only in the short term since it can lead to poorly maintainable code and developer burnouts in the long run.
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+As remarked initially, \textit{speed-up development} represents the core 
+category of the GSM. Firmly grounded as the primary objective of startups, it 
+shows the most important characteristic of developing software in the early 
+stage.
+
+As mentioned before, in order to \textit{speed-up development}, startups adopt 
+evolutionary approaches supported by a solid team focusing on implementing a 
+minimal set of suitable functionalities. Startups \textit{keep simple and 
+informal workflows} to be flexible and reactive, adapting to the fast changing 
+environment. While a rigorously defined workflow generally consists of an 
+established sequence of well-defined steps to follow, startups, by contrast, 
+adopt informal workflows. This choice is mainly dictated by contextual factors 
+characterized by the general conditions of unpredictability and uncertainty. The 
+adoption of informal workflows is facilitated by the fact that teams are 
+typically self-organized and \textit{developers have big responsibilities} .
+
+Additionally, possible planning activities are refrained by the aim to shorten 
+time-to-market, as reported by C8: \textit{``Speed was the essence so we didn't 
+plan out too many details''}. To deal with such unpredictability, startups 
+prefer to take decisions as fast as possible, mainly by means of informal and 
+frequent verbal discussions.
+
+Despite Agile principles appear to be suitable to embrace changes, especially 
+in the early-stage, development practices are often perceived as time-wasters 
+and ignored to accommodate the need of releasing the product to market as 
+quickly as possible. In this regard, to maintain the simplest and most informal 
+approach, startups aim to \textit{find the product/market fit quickly} and 
+implement a \textit{limited number of suitable functionalities}. This approach 
+is considered possible also in view of a lack of systematic quality assurance 
+activities, since only the user experience is considered as important and other 
+quality aspects, such as efficiency, can be postponed after the first release.
+
+Another strategy, which supports startups in quickly delivering products, is 
+the \textit{externalization of complexity on third party solutions} by: making 
+use of third party components (COTS) and open source solutions (both for 
+product components, and development tools and libraries); taking advantage of  
+external services; and deploying the product on external infrastructures, for 
+the sake of delivering a \textit{product reasonably ready to scale} for a 
+possible future growth.
+
+Even though \textit{the use of well-integrated and simple tools} allows 
+startups to automate many activities and reduce their completion time, drawbacks 
+of such approaches are the increased concerns of interoperability issues. A 
+category of tools is represented by advanced version control systems, which are 
+not only used to manage the code-base, but for many other purposes such as: 
+assigning tasks; trace responsibilities; configuration management; issue 
+management; automatic deployment; discussions and code reviews.
+
+Startups can further improve development speed by making \textit{use of 
+standards and known technologies} which are widely recognized, well tested, and 
+supported by strong communities. Moreover, the use of widely adopted standards 
+and frameworks reduces the need of a formal architectural design since most of 
+implementation solutions are well documented and ready-to-use. This is confirmed 
+by C1: \textit{``as long as you use Ruby standards with the Rails framework, the 
+language is clean itself and doesn't need much documentation''.}
+
+Other important factors that positively impact the speed of development are the 
+team's desire to: \textit{create disruptive technologies}; to 
+\textit{demonstrate personal abilities}; and to \textit{have the product used in 
+the market}. As reported by practitioners, these factors are essential to 
+enhance the morale of developers and therefore to achieve higher team 
+performances. By contrast, especially in the typical sprint-based environments 
+of Agile, when a developer is not \textit{able to meet deadlines} the morale 
+goes down, hindering the development speed.
+
+Finally the constant pressure under which the company regularly operates, leads 
+the team to often \textit{work overtime to meet deadlines}. But as reported by 
+practitioners, such way of working can be an effective strategy only in the 
+short term since it can lead to poorly maintainable code and developer burnouts 
+in the long run.
+%------------------------------------------------------------------------------%
 \subsection{Accumulated technical debt}
 \label{res:gsm:cat5}
-Startups achieve a very high development speed by radically ignoring aspects related to documentation, structures and processes.
-However, although these aspects are not considered important in the very first stages, they will become increasingly more important for the productivity in the long-term, as we illustrate in CAT6.
-
-Instead of traditional requirement engineering activities, startups make use of \textit{informal specification of functionalities} by means of ticked-based tools to manage low-precision lists of features to implement, written in form of self-explanatory user stories \cite{AgilePlan}.
-Practitioners intensively use physical tools such as post-it notes and whiteboards, which help in making functionalities visible and prioritizing stories based on personal experiences.
-C4 comments \textit{``[\ldots] it is the only way.
-Too many people make the mistake of sitting down and write big specs and then they build it for four months, realizing the product is not valuable only at the end.''}
-
-Since startups are risky businesses by nature, even less attention is given to the traditional phase of \textit{analysis}, which is replaced by a \textit{rough and quick feasibility study}   \textit{``only sometimes and however informally''}, as stated by C5.
-First of all it is generally \textit{hard to analyze risks with cutting-edge technologies}.
-To find out the feasibility of such cutting-edge projects, startups attempt a first implementation with rough and informal specifications, assuming that project's complexity will remain limited to a small number of functionalities, as discussed in CAT3.
-Additionally by \textit{keeping the product as simple as possible} and learning from competitors' solutions and mistakes, practitioners can use their \textit{past experiences in similar contexts to help assessing feasibility} of the project.
-Finally, to avoid restrictions on the flexibility of the team, potential limiting decisions are taken only when strictly necessary and anyhow as late as possible.
-
-Another important factor which contributes to \textit{accumulate the technical debt} is the general \textit{lack of architectural design}, substituted by \textit{high-level mockups and low-precision diagrams}, \textit{using modularization and frameworks from day 1} and \textit{describing critical interactions with third-party components only}.
-In particular, the use of well-known standards, frameworks and conventions limits the need of formal UML \cite{UML} diagrams and documentation, and provide a minimum level of maintainability costs.
-Additionally, since quality aspects are not a main concern (see \textit{limited number of functionalities} and \textit{efficiency emerges after using the product} in CAT3) having a well-structured architecture remains a secondary priority, and design is conducted only when strictly necessary.
-
-A similar attitude towards verification and validation brings startups to a severe \textit{lack of automated testing}, which is mainly replaced by manual smoke tests executed by: trying the product internally; seeking for critical faults; and letting early adopters report non-critical bugs.
-Paraphrasing C3, \textit{``Trying the product internally allows us to get rid of 50\% of bugs of important functionalities.
-Meanwhile users report bugs of secondary functionalities, eventually allowing us to mitigate the lack of testing.
-Indeed, staying one week in production enables us to identify 90\% of bugs''.} The lack of complete automated tests is partially also motivated by the fact that \textit{users are fault-tolerant in innovative beta products} and  by the limited number of functionalities, which allow the team to easily control critical bugs.
-However, in certain cases where components of the system might cause loss of data or severe damages to the product or users, engineers realize a reasonable level of
-automatic testing. In such cases, aided by modern automatic tools, they can quickly assess the status of the system integration as they add new functionalities to the product.
-
-Furthermore, a rigid project management is perceived as \textit{``waste of time''} which hinders the development speed since the \textit{uncertainty makes formal scheduling pointless} (C9  reports that \textit{``initial chaos helps to develop faster''}).
-Startups' \textit{minimal project management}  is supported by keeping: \textit{internal milestones short and informal}, \textit{low-precision task assignment mechanism}  and a very low attention for project metrics (paraphrasing C13, \textit{``the only track of progress was made by looking at closed tickets''}).
-In this context \textit{only a final release milestone is viable}, which helps practitioners to remain focused on short term goals and put new features in production.
-To further \textit{speed-up development}, startups \textit{simplify issue management by integrating it with feature management}.
-Then, the \textit{limited number of functionalities} and the {use of standard/known technologies} with a \textit{simple workflow} are the main reasons for
-not establishing a heavy project management process since the project and technologies can be autonomously managed by the development team (with characteristics of: co-location, self-organization and very small size).
-%MUN: do you mean small team size?
-
-Finally, one of the categories, which most contributes in growing the \textit{accumulated technical debt}, is the substantial  use of informal and verbal communication channels on a daily basis.
-The high co-location and the fast paced development approach increase the volume of \textit{tacit knowledge} and severe lack of any kind of documentation.
-C4 observes in this regard that:  \textit{``[\ldots] the issue of having documentation and diagrams out of the source code is that you need to update them every time you change something.
-There is no time for it.
-Instead, there is a huge pay off in having a code that is understandable itself.''}
-This approach is supported by the fact that startups make use of  \textit{simple and informal workflow}, \textit{standard/known technologies}, \textit{very small and co-located development team} with \textit{limited need of formalities}.
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+Startups achieve a very high development speed by radically ignoring aspects 
+related to documentation, structures and processes. However, although these 
+aspects are not considered important in the very first stages, they will become 
+increasingly more important for the productivity in the long-term, as we 
+illustrate in CAT6.
+
+Instead of traditional requirement engineering activities, startups make use of 
+\textit{informal specification of functionalities} by means of ticked-based 
+tools to manage low-precision lists of features to implement, written in form of 
+self-explanatory user stories \cite{AgilePlan}. Practitioners intensively use 
+physical tools such as post-it notes and whiteboards, which help in making 
+functionalities visible and prioritizing stories based on personal experiences. 
+C4 comments \textit{``[\ldots] it is the only way. Too many people make the 
+mistake of sitting down and write big specs and then they build it for four 
+months, realizing the product is not valuable only at the end.''}
+
+Since startups are risky businesses by nature, even less attention is given to 
+the traditional phase of \textit{analysis}, which is replaced by a \textit{rough 
+and quick feasibility study}   \textit{``only sometimes and however 
+informally''}, as stated by C5.  First of all it is generally \textit{hard to 
+analyze risks with cutting-edge technologies}. To find out the feasibility of 
+such cutting-edge projects, startups attempt a first implementation with rough 
+and informal specifications, assuming that project's complexity will remain 
+limited to a small number of functionalities, as discussed in CAT3. Additionally 
+by \textit{keeping the product as simple as possible} and learning from 
+competitors' solutions and mistakes, practitioners can use their \textit{past 
+experiences in similar contexts to help assessing feasibility} of the project. 
+Finally, to avoid restrictions on the flexibility of the team, potential 
+limiting decisions are taken only when strictly necessary and anyhow as late as 
+possible.
+
+Another important factor which contributes to \textit{accumulate the technical 
+debt} is the general \textit{lack of architectural design}, substituted by 
+\textit{high-level mockups and low-precision diagrams}, \textit{using 
+modularization and frameworks from day 1} and \textit{describing critical 
+interactions with third-party components only}. In particular, the use of 
+well-known standards, frameworks and conventions limits the need of formal UML 
+\cite{UML} diagrams and documentation, and provide a minimum level of 
+maintainability costs. Additionally, since quality aspects are not a main 
+concern (see \textit{limited number of functionalities} and \textit{efficiency 
+emerges after using the product} in CAT3) having a well-structured architecture 
+remains a secondary priority, and design is conducted only when strictly 
+necessary.
+
+A similar attitude towards verification and validation brings startups to a 
+severe \textit{lack of automated testing}, which is mainly replaced by manual 
+smoke tests executed by: trying the product internally; seeking for critical 
+faults; and letting early adopters report non-critical bugs.  Paraphrasing C3, 
+\textit{``Trying the product internally allows us to get rid of 50\% of bugs of 
+important functionalities. Meanwhile users report bugs of secondary 
+functionalities, eventually allowing us to mitigate the lack of testing. Indeed, 
+staying one week in production enables us to identify 90\% of bugs''.} The lack 
+of complete automated tests is partially also motivated by the fact that 
+\textit{users are fault-tolerant in innovative beta products} and  by the 
+limited number of functionalities, which allow the team to easily control 
+critical bugs. However, in certain cases where components of the system might 
+cause loss of data or severe damages to the product or users, engineers realize 
+a reasonable level of automatic testing. In such cases, aided by modern 
+automatic tools, they can quickly assess the status of the system integration as 
+they add new functionalities to the product.
+
+Furthermore, a rigid project management is perceived as \textit{``waste of 
+time''} which hinders the development speed since the \textit{uncertainty makes 
+formal scheduling pointless} (C9  reports that \textit{``initial chaos helps to 
+develop faster''}). Startups' \textit{minimal project management}  is supported 
+by keeping: \textit{internal milestones short and informal}, 
+\textit{low-precision task assignment mechanism}  and a very low attention for 
+project metrics (paraphrasing C13, \textit{``the only track of progress was made 
+by looking at closed tickets''}). In this context \textit{only a final release 
+milestone is viable}, which helps practitioners to remain focused on short term 
+goals and put new features in production. To further \textit{speed-up 
+development}, startups \textit{simplify issue management by integrating it with 
+feature management}. Then, the \textit{limited number of functionalities} and 
+the {use of standard/known technologies} with a \textit{simple workflow} are the 
+main reasons for 
+not establishing a heavy project management process since the project and 
+technologies can be autonomously managed by the development team (with 
+characteristics of: co-location, self-organization and very small size). %MUN: 
+%do you mean small team size?
+
+Finally, one of the categories, which most contributes in growing the 
+\textit{accumulated technical debt}, is the substantial  use of informal and 
+verbal communication channels on a daily basis. The high co-location and the 
+fast paced development approach increase the volume of \textit{tacit knowledge} 
+and severe lack of any kind of documentation.  C4 observes in this regard that:  
+\textit{``[\ldots] the issue of having documentation and diagrams out of the 
+source code is that you need to update them every time you change something. 
+There is no time for it. Instead, there is a huge pay off in having a code that 
+is understandable itself.''}
+This approach is supported by the fact that startups make use of  
+\textit{simple and informal workflow}, \textit{standard/known technologies}, 
+\textit{very small and co-located development team} with \textit{limited need of 
+formalities}.
+%------------------------------------------------------------------------------%
 \subsection{Initial growth hinders performance}
 \label{res:gsm:cat6}
-The lack of attention given in the first phases to engineering activities allows startups to ship code extremely quickly (see CAT5).
-However, if successful, the initial product becomes more and more complex over time, the number of users increases and the company size starts to grow.
-Under these circumstances the need of controlling the initial chaos forces the development team to return the \textit{accumulated technical debt} instead of focusing on new users' requests.
-Hence, \textit{initial growth hinders performance} in terms of number of new user stories implemented in a certain amount of time, which brings new suitable functionalities to the users.
-
-When the number of users increases, customers become more quality demanding and also scalability issues start to arise.
-Usually, \textit{company and user size grow} when new business events occur, such as: a \textit{new round of funding}, a possible \textit{acquisition}, the release of a \textit{competing product on the market}, or when the project is \textit{open for the first public release}.
-As a consequence, the \textit{increasing number of users} creates a growing number of requests and expectations.
-Therefore, whereas the project lacks of minimal processes, suddenly, \textit{the current team is not able to manage increased complexity} of new functionalities to implement and maintain the codebase.
-
-Subsequently, practitioners start considering the need of project management activities, also in view of \textit{hiring new staff members}, as discussed by C13: \textit{``(project management) is strictly necessary if you radically change the team or when the team grows.
-The informal communication and lack of documentation slow down the process afterwards''}.
-Project management becomes further important when the \textit{focus shifts to business concerns}.
-Part of the effort, which was initially almost entirely dedicated to product development, is required to move to business activities.
-Moreover the availability of project information becomes an important issue because of the accumulated \textit{tacit knowledge}, which hinders the ability of new hires to quickly start the development of project tasks.
-
-When the company faces growth, the partial and informal engineering activities that have been conducted during the first phases of \textit{rush} development (refer to CAT5 subcategories \textit{minimal project management}, \textit{informal specification of functionalities}, \textit{rough and quick feasibility study}, \textit{lack of automated testing}, \textit{tacit knowledge}), force the company to \textit{pay off the accumulated technical debt}.
-Under these circumstances, startups need to be able to cope with the renewed needs and expectations of both the company (internal necessities) and customers (product).
-
-Another factor that slows down performance is caused by the fact that \textit{portion of codes needs to be rewritten} and \textit{substantial refactoring of the codebase} is required by increasing product demands.
-
-Practitioners realize that some decisions taken (or not taken) during the \textit{rough and quick feasibility study} before starting the implementation, have led to negative consequences on the long term performance and maintainability of the product.
-Additionally, the initial decisions of the product design might not be able to satisfy the increased demands of the product's users and developers (\textit{lack of architectural design}).
-The combination of these factors leads to the need of \textit{re-engineer the product}.
-
-By re-engineering the systems, startups aim to \textit{increase the scalability of the product/infrastructure} and starting to \textit{standardize the codebase with well-known frameworks}.
-
-C7 reports that: \textit{``To mitigate this (lack of frameworks) I had to make a schema for other developers when we hired them.
-We had to do a big refactoring of the codebase, moving it from custom php to Django, normalizing the model and making it stick with the business strategy.
-I had the code in different php servers communicating via JSON, some engineering horror.
-Now  that we are fixing it, it's really painful.
-We had to trash some code.
-However I don't regret that I didn't make this choice sooner, it was the only way''}.
-
-The \textit{fear of changing a product, which is working}, arises when product complexity increases.
-The changes to the codebase, to support bug fixing, become highly interrelated with other functionalities and difficult to manage because the product is poorly engineered (see CAT5).
-Therefore, the fear arises from changing a validated product that might bring changes to the user responses.
-
-The increasing number of feature requests causes the \textit{growing necessity of having a release plan}, rationalizing the rollout of new features.
-
-Therefore, startups begin to \textit{partially replace informal communication with traceable systems} and \textit{introduce basic metrics for measuring project and team progress} in order to establish an initial structured workflow.
-In conclusion, because the increasing number of users causes issues in scalability and reliability, the \textit{need of more structured workflow} becomes of major relevance.
-C11 states: \textit{``Yet, it is still better to have a reasonable drop-down in performance when team grows than lose time in the beginning''}.
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+The lack of attention given in the first phases to engineering activities 
+allows startups to ship code extremely quickly (see CAT5). However, if 
+successful, the initial product becomes more and more complex over time, the 
+number of users increases and the company size starts to grow. Under these 
+circumstances the need of controlling the initial chaos forces the development 
+team to return the \textit{accumulated technical debt} instead of focusing on 
+new users' requests. Hence, \textit{initial growth hinders performance} in terms 
+of number of new user stories implemented in a certain amount of time, which 
+brings new suitable functionalities to the users.
+
+When the number of users increases, customers become more quality demanding and 
+also scalability issues start to arise. Usually, \textit{company and user size 
+grow} when new business events occur, such as: a \textit{new round of funding}, 
+a possible \textit{acquisition}, the release of a \textit{competing product on 
+the market}, or when the project is \textit{open for the first public release}. 
+As a consequence, the \textit{increasing number of users} creates a growing 
+number of requests and expectations. Therefore, whereas the project lacks of 
+minimal processes, suddenly, \textit{the current team is not able to manage 
+increased complexity} of new functionalities to implement and maintain the 
+codebase.
+
+Subsequently, practitioners start considering the need of project management 
+activities, also in view of \textit{hiring new staff members}, as discussed by 
+C13: \textit{``(project management) is strictly necessary if you radically 
+change the team or when the team grows.  The informal communication and lack of 
+documentation slow down the process afterwards''}.  Project management becomes 
+further important when the \textit{focus shifts to business concerns}. Part of 
+the effort, which was initially almost entirely dedicated to product 
+development, is required to move to business activities. Moreover the 
+availability of project information becomes an important issue because of the 
+accumulated \textit{tacit knowledge}, which hinders the ability of new hires to 
+quickly start the development of project tasks.
+
+When the company faces growth, the partial and informal engineering activities 
+that have been conducted during the first phases of \textit{rush} development 
+(refer to CAT5 subcategories \textit{minimal project management}, 
+\textit{informal specification of functionalities}, \textit{rough and quick 
+feasibility study}, \textit{lack of automated testing}, \textit{tacit 
+knowledge}), force the company to \textit{pay off the accumulated technical 
+debt}. Under these circumstances, startups need to be able to cope with the 
+renewed needs and expectations of both the company (internal necessities) and 
+customers (product).
+
+Another factor that slows down performance is caused by the fact that 
+\textit{portion of codes needs to be rewritten} and \textit{substantial 
+refactoring of the codebase} is required by increasing product demands.
+
+Practitioners realize that some decisions taken (or not taken) during the 
+\textit{rough and quick feasibility study} before starting the implementation, 
+have led to negative consequences on the long term performance and 
+maintainability of the product. Additionally, the initial decisions of the 
+product design might not be able to satisfy the increased demands of the 
+product's users and developers (\textit{lack of architectural design}). The 
+combination of these factors leads to the need of \textit{re-engineer the 
+product}.
+
+By re-engineering the systems, startups aim to \textit{increase the scalability 
+of the product/infrastructure} and starting to \textit{standardize the codebase 
+with well-known frameworks}.
+
+C7 reports that: \textit{``To mitigate this (lack of frameworks) I had to make 
+a schema for other developers when we hired them. We had to do a big 
+refactoring of the codebase, moving it from custom php to Django, normalizing 
+the model and making it stick with the business strategy. I had the code in 
+different php servers communicating via JSON, some engineering horror. Now  that 
+we are fixing it, it's really painful. We had to trash some code. However I 
+don't regret that I didn't make this choice sooner, it was the only way''}.
+
+The \textit{fear of changing a product, which is working}, arises when product 
+complexity increases. The changes to the codebase, to support bug fixing, become 
+highly interrelated with other functionalities and difficult to manage because 
+the product is poorly engineered (see CAT5). Therefore, the fear arises from 
+changing a validated product that might bring changes to the user responses.
+
+The increasing number of feature requests causes the \textit{growing necessity 
+of having a release plan}, rationalizing the rollout of new features.
+
+Therefore, startups begin to \textit{partially replace informal communication 
+with traceable systems} and \textit{introduce basic metrics for measuring 
+project and team progress} in order to establish an initial structured workflow. 
+In conclusion, because the increasing number of users causes issues in 
+scalability and reliability, the \textit{need of more structured workflow} 
+becomes of major relevance. C11 states: \textit{``Yet, it is still better to 
+have a reasonable drop-down in performance when team grows than lose time in the 
+beginning''}.
+%------------------------------------------------------------------------------%
 \subsection{Theory}
 \label{res:gsm:th}
-In order to explain and understand the development strategies in early-stage software startups we construct the theory generated and supported by the above presented GSM:
+In order to explain and understand the development strategies in early-stage 
+software startups we construct the theory generated and supported by the above 
+presented GSM:
 
 \begin{theory}
-Focusing on limited number of suitable functionalities, and adopting partial and rapid evolutionary development approaches, early-stage software startups operate at high development speed, aided by skilled and highly co-located developers.
-Through these development strategies, early-stage software startups aim to find early product/market fit within extremely uncertain conditions and severe lack of resources.
-Nevertheless, by speeding-up the development process, they accumulate technical debt, causing an initial and temporary drop-down in performance before setting off for further growth.
+Focusing on limited number of suitable functionalities, and adopting partial 
+and rapid evolutionary development approaches, early-stage software startups 
+operate at high development speed, aided by skilled and highly co-located 
+developers. Through these development strategies, early-stage software startups 
+aim to find early product/market fit within extremely uncertain conditions and 
+severe lack of resources. Nevertheless, by speeding-up the development process, 
+they accumulate technical debt, causing an initial and temporary drop-down in 
+performance before setting off for further growth.
 \end{theory}
 
-As discussed in section \ref{valt}%MUN: this refers to the validity threats section. Do you mean rather the "Theory Validation" section?
-, we formed the theory by applying the paradigm model, introduced by Corbin \cite{Corbin1990}. The theory is composed by different elements reflecting the paradigm model:
+As discussed in section \ref{valt} %MUN: this refers to the validity threats 
+%section. Do you mean rather the "Theory Validation" section?
+, we formed the theory by applying the paradigm model, introduced by Corbin 
+\cite{Corbin1990}. The theory is composed by different elements reflecting the 
+paradigm model:
 
@@
 
-\item \textit{Causal conditions} are represented by three main conceptual categories: \textit{product quality has low priority}, \textit{evolutionary approach} and \textit{team is the catalyst of development}.
-\item \textit{Phenomenon} is represented by the core category \textit{speed-up development}.
-\item \textit{Context} is limited to web early-stage software startups operating in conditions of severe lack of resources aiming to early find product/market fit.
-\item \textit{Intervening conditions} are  summarized by the extremely uncertain development environment.
-\item \textit{Action and interaction strategies} are represented by the accumulation of technical debt.
+\item \textit{Causal conditions} are represented by three main conceptual 
+categories: \textit{product quality has low priority}, \textit{evolutionary 
+approach} and \textit{team is the catalyst of development}.
+\item \textit{Phenomenon} is represented by the core category \textit{speed-up 
+development}.
+\item \textit{Context} is limited to web early-stage software startups 
+operating in conditions of severe lack of resources aiming to early find 
+product/market fit.
+\item \textit{Intervening conditions} are  summarized by the extremely 
+uncertain development environment.
+\item \textit{Action and interaction strategies} are represented by the 
+accumulation of technical debt.
 \item \textit{Consequences} lead to a temporary performance drop-down.
@@
 
-%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+%------------------------------------------------------------------------------%
 \section{Theory Validation}
@@
 
-In this section we discuss the validity of the GSM by means of cross-methodological observations:
-%MUN: expand this discussion a bit (motivation why you do this like you do it) and move to Section 3.5.
-Then write here just that you perform theory validation as discussed in Section 3.5 (and remove the bullet list below).
+In this section we discuss the validity of the GSM by means of 
+cross-methodological observations:
+%MUN: expand this discussion a bit (motivation why you do this like you do it) 
+%and move to Section 3.5. Then write here just that you perform theory 
+%validation as discussed in Section 3.5 (and remove the bullet list below).
 
 \begin{compactitem}
-\item Comparison between the GSM model and similar  models and frameworks.
-
-\item Comparison between the GSM model and the extant literature. We highlight the areas which have been neglected by existing studies, providing possible directions for future studies.
-\item Discussion of confounding factors which could interfere with the GSM model.
-\item Validation of the correctness of the GT procedure by attesting relations between the GSM categories, evaluating if they are compliant with the empirical data we collected.
+\item Comparison between the GSM and similar  models and frameworks.
+\item Comparison between the GSM and literature. We highlight 
+the areas which have been neglected by existing studies, providing possible 
+directions for future studies.
+\item Discussion of confounding factors which could interfere with the GSM.
+\item Validation of the correctness of the GT procedure by attesting relations 
+between the GSM categories, evaluating if they are compliant with the empirical 
+data we collected.
 
@@
 
-%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+%------------------------------------------------------------------------------%
 \subsection{Comparison with other frameworks}
@@
 
-To validate the generalization of the theory, we describe conceptualizations defined from the GSM that are in support of a previous model developed by Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}.
-We refer to Coleman's work since only he has, to the best of our knowledge, conducted a similar research in startups' context, even though with a different focus.
-Coleman investigated factors in software development that hinder initiatives of software process improvement (SPI) in much more mature companies.
-%MUN: so he did not do research in a startup context?
-You need to clarify what you mean by "more mature companies"...
-I assume you still mean startups, but startups at a later stage, i.e.
-after first product release.
-
-In this subsection we illustrate the most important aspects considered by Coleman, examining similar and contrasting factors with respect to the GSM.
-We consider only those categories strictly related to the core category, which relations, as described in GT methodology, generate complete explanatory power of the generated theory.
-
-Coleman's framework aims to highlight how managers consider two distinct kinds of processes: \textit{essentials} and \textit{non-essentials}.
-The \textit{essential} processes are the most closely linked to product development, such as requirements gathering, design and testing.
-The \textit{non-essential} processes are those that might be omitted, such as planning, estimating and staging meetings.
-In particular, he discusses how practices are routinely removed: \textit{``With most methodologies and approaches, very few stick to the letter of them and they are always adapted, so we adapted ours to the way we wanted it to work for us, for our own size and scale''.}%MUN: add ref to Coleman's paper where he states this
-
-The GSM explores the same challenges in software startups, where the act of tailoring processes leads developers to adopt only minimal practices, which are most suitable for the startups context.
-Furthermore, CTOs' and CEOs' background has a great impact on the speed of the development process as described in CAT4.
-This is in accordance to Coleman's framework, which presents the background of founders and development managers as main factors that affect the management style of the development process.
-
-Differently from Coleman's framework, the GSM doesn't present how \textit{market requirements} can affect the conduction of processes.
-In this regard, Coleman describes how the more requirements definitions are predictable the more well-defined workflows are established.
-This particular aspect is considered as a confounding factor\footnote{The identification of confounding depends on the context in which the study is conducted and on the background knowledge of the researchers \cite{Pearl2011}.}  in our study since we were not able to ground this concept in the data.
-
-But what characterizes Coleman's network is the \textit{cost of process} (core category) and all the factors that in management contributed to the lack of software process improvements (SPI).
-The \textit{cost of process} represents the  lack of formal and prescriptive workflows in development, mainly conducted by means of verbal \textit{communication} limiting heavy \textit{documentation} and \textit{bureaucracy}.
-
-Coleman reports the practitioners' perception that \textit{documentation} alone would not have ensured a complete shared understanding of project requirements.
-Moreover, defined processes are perceived by managers as additional items with a negative impact on the \textit{Creativity} and \textit{Flexibility} of the development team.
-This is in accordance with our generated theory, which bases the reasons of adopting evolutionary and low-precision engineering elements in the flexibility and reactivity attributes of the development process in startups.
-
-In this regard we can paraphrase a comment recorded by Coleman during an interview: \textit{``When we set up we had more supervisory and managerial roles in that group than we have now and we had to scale that back which has made things a lot more flexible.
-I do think you have to be nimble, quick and capable being responsive in our position.
-That works well and I don't want to lose it.''} \cite{Coleman2007}.
-
-Verbal \textit{communication}, lack of heavy \textit{documentation} and \textit{bureaucracy} can easily be associated to the \textit{accumulated technical debt} category (see CAT5).
-According to Coleman's framework, they describe how startups experience the lack of main engineering activities and documentation.
-Additionally, the \textit{speed-up development} category expresses the same concept of \textit{flexibility} and \textit{process erosion} since they have direct relation to the subcategory of \textit{keep a simple and informal workflow}, as discussed in CAT1.
-
-As also reported in the GSM, the definition of a \textit{``minimum process''} is not a matter of poor knowledge and training, but rather it is the necessity of operating with solutions that let the company move faster.
-\textit{``One-size-fits-all''} solutions have always found difficulty in penetrating small software organizations \cite{Staples2007}.
-When startups began establishing any SPI process, they experienced \textit{process erosions}%MUN: add ref (Coleman?).
-You should add, for each category you use from Coleman, add the reference to the paper.
-This makes it clear that the category comes from him (and is not one of our categories)
-, which resolved to a barely sufficient workflow to satisfy the organizational business needs.
-
-Then, software startups favor the use of agile principles in support of \textit{creativity} and \textit{flexibility} instead of SPI methods, whereby processes need to be predictable and repeatable.
-Nimble and ad-hoc solutions prevent the use of heavy bureaucracy and fast communication strategies%MUN: seems to contradict each other: "prevent the use of heavy bureaucracy AND fast communication strategies
-, even though the accumulated tacit knowledge is hard to manage and to transfer to new hires as also discussed in category \textit{CAT6}.
-
-Coleman describes how the \textit{``management approach''} is oriented towards \textit{``embrace and empower''} solutions in contrast with \textit{``command and control''} style, where there is an evidence for trust in the  development staff to carry out tasks with less direct supervision, greater delegation of responsibilities and a more generally consensual environment \cite{Coleman2008}.
-Nevertheless, software development managers and founders have still impact on management style and indirectly on the software development process.
-In the case of early-stage startups, founders are mainly software development managers as CEOs/CTOs and  technical practitioners at the same time.
-As Coleman identified the influence of the founders' and managers' background on the software development process, the GSM similarly identifies that the CEOs and CTOs background shapes the high-level strategies adopted in developing the initial product.
-Notwithstanding, team members remain self-organized, able to intervene in all the
-aspects of the development process without any direct supervision, as discussed in CAT4.
-
-In conclusion, by the comparison of the GSM with Coleman's framework, we have found that despite Coleman's study contributes to define which factors hinder the introduction of SPI%MUN: add: "of SPI in star I ditups"?
-, his results are coherent with the theory generated within our research%MUN: add "our research, focused at explaining software development in general in startups." ?
-.
-In this regard, a wide variability of processes is a key factor for startups.
-In contrast with some SPI processes, such as \textit{Six Sigma} which objective is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, startups seem to follow the \textit{Fischer's Fundamental Theorem of natural selection} \cite{Fisher}.
-Moving towards flexible and variable processes (easy to adapt to the uncertain conditions), increase the odds of ``natural selection'' only when some ``restrictive conditions'' are met.
-In other words, even though ignoring SPI practices is advisable%MUN: under which circumstances?
-And the following part,I did not fully understand.
-Try to clarify.
- not following SPI does not predict major adaptiveness of startups development to the market.
-But, to imply a major evolutionary product/market fit, certain conditions explained in GSM need to be considered.
-
-%MUN: As I wrote already in the previous draft, you need to clarify at the beginning of this subsection which frameworks you are going to compare GSM to.
-Now, you say that Coleman is the only author who did research, comparable to yours, into startups.
-Now there comes Baskerville and Brooks.
-Write at the beginning of this section, shortly, which frameworks/authors you compare GSM to and why, then create subsections for each.
-Also Baskerville \cite{Internet} refers to SPI as a development approach typically effective only in large-scale, long-term development efforts with stable and disciplined processes.
-By applying a grounded theory analysis in 10 companies he found that major causal factors that influence development are:
+To validate the generalization of the theory, we describe conceptualizations 
+defined from the GSM that are in support of a previous model developed by 
+Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}. We refer to Coleman's work 
+since only he has, to the best of our knowledge, conducted a similar research in 
+startups' context, even though with a different focus. Coleman investigated 
+factors in software development that hinder initiatives of software process 
+improvement (SPI) in much more mature companies. %MUN: so he did not do research 
+%in a startup context? You need to clarify what you mean by "more mature 
+%companies"... I assume you still mean startups, but startups at a later stage, 
+%i.e. after first product release.
+
+In this subsection we illustrate the most important aspects considered by 
+Coleman, examining similar and contrasting factors with respect to the GSM. We 
+consider only those categories strictly related to the core category, which 
+relations, as described in GT methodology, generate complete explanatory power 
+of the generated theory.
+
+Coleman's framework aims to highlight how managers consider two distinct kinds 
+of processes: \textit{essentials} and \textit{non-essentials}. The 
+\textit{essential} processes are the most closely linked to product development, 
+such as requirements gathering, design and testing. The \textit{non-essential} 
+processes are those that might be omitted, such as planning, estimating and 
+staging meetings. In particular, he discusses how practices are routinely 
+removed: \textit{``With most methodologies and approaches, very few stick to the 
+letter of them and they are always adapted, so we adapted ours to the way we 
+wanted it to work for us, for our own size and scale''.} %MUN: add ref to 
+%Coleman's paper where he states this
+
+The GSM explores the same challenges in software startups, where the act of 
+tailoring processes leads developers to adopt only minimal practices, which are 
+most suitable for the startups context. Furthermore, CTOs' and CEOs' background 
+has a great impact on the speed of the development process as described in CAT4. 
+This is in accordance to Coleman's framework, which presents the background of 
+founders and development managers as main factors that affect the management 
+style of the development process.
+
+Differently from Coleman's framework, the GSM doesn't present how 
+\textit{market requirements} can affect the conduction of processes. In this 
+regard, Coleman describes how the more requirements definitions are predictable 
+the more well-defined workflows are established. This particular aspect is 
+considered as a confounding factor\footnote{The identification of confounding 
+depends on the context in which the study is conducted and on the background 
+knowledge of the researchers \cite{Pearl2011}.}  in our study since we were not 
+able to ground this concept in the data.
+
+But what characterizes Coleman's network is the \textit{cost of process} (core 
+category) and all the factors that in management contributed to the lack of 
+software process improvements (SPI). The \textit{cost of process} represents the 
+ lack of formal and prescriptive workflows in development, mainly conducted by 
+means of verbal \textit{communication} limiting heavy \textit{documentation} and 
+\textit{bureaucracy}.
+
+Coleman reports the practitioners' perception that \textit{documentation} alone 
+would not have ensured a complete shared understanding of project requirements. 
+Moreover, defined processes are perceived by managers as additional items with a 
+negative impact on the \textit{Creativity} and \textit{Flexibility} of the 
+development team. This is in accordance with our generated theory, which bases 
+the reasons of adopting evolutionary and low-precision engineering elements in 
+the flexibility and reactivity attributes of the development process in 
+startups.
+
+In this regard we can paraphrase a comment recorded by Coleman during an 
+interview: \textit{``When we set up we had more supervisory and managerial roles 
+in that group than we have now and we had to scale that back which has made 
+things a lot more flexible. I do think you have to be nimble, quick and capable 
+being responsive in our position. That works well and I don't want to lose 
+it.''} \cite{Coleman2007}.
+
+Verbal \textit{communication}, lack of heavy \textit{documentation} and 
+\textit{bureaucracy} can easily be associated to the \textit{accumulated 
+technical debt} category (see CAT5). According to Coleman's framework, they 
+describe how startups experience the lack of main engineering activities and 
+documentation. Additionally, the \textit{speed-up development} category 
+expresses the same concept of \textit{flexibility} and \textit{process erosion} 
+since they have direct relation to the subcategory of \textit{keep a simple and 
+informal workflow}, as discussed in CAT1.
+
+As also reported in the GSM, the definition of a \textit{``minimum process''} 
+is not a matter of poor knowledge and training, but rather it is the necessity 
+of operating with solutions that let the company move faster. 
+\textit{``One-size-fits-all''} solutions have always found difficulty in 
+penetrating small software organizations \cite{Staples2007}. When startups began 
+establishing any SPI process, they experienced \textit{process erosions} %MUN: 
+%add ref (Coleman?). You should add, for each category you use from Coleman, 
+%add the reference to the paper. This makes it clear that the category comes 
+%from him (and is not one of our categories)
+, which resolved to a barely sufficient workflow to satisfy the organizational 
+business needs.
+
+Then, software startups favor the use of agile principles in support of 
+\textit{creativity} and \textit{flexibility} instead of SPI methods, whereby 
+processes need to be predictable and repeatable. Nimble and ad-hoc solutions 
+prevent the use of heavy bureaucracy and fast communication strategies %MUN: 
+%seems to contradict each other: "ad-hoc solutions prevent the use of heavy 
+%bureaucracy AND fast communication strategies" I agree that ad-hoc solutions 
+%do not produce heavy bureaucracy, but they allow fast communication 
+%strategies... I may be wrong, but then you need to explain it better
+, even though the accumulated tacit knowledge is hard to manage and to transfer 
+to new hires as also discussed in category \textit{CAT6}.
+
+Coleman describes how the \textit{``management approach''} is oriented towards 
+\textit{``embrace and empower''} solutions in contrast with \textit{``command 
+and control''} style, where there is an evidence for trust in the  development 
+staff to carry out tasks with less direct supervision, greater delegation of 
+responsibilities and a more generally consensual environment \cite{Coleman2008}. 
+Nevertheless, software development managers and founders have still impact on 
+management style and indirectly on the software development process. In the case 
+of early-stage startups, founders are mainly software development managers as 
+CEOs/CTOs and  technical practitioners at the same time. As Coleman identified 
+the influence of the founders' and managers' background on the software 
+development process, the GSM similarly identifies that the CEOs and CTOs 
+background shapes the high-level strategies adopted in developing the initial 
+product. Notwithstanding, team members remain self-organized, able to intervene 
+in all the aspects of the development process without any direct supervision, as 
+discussed in CAT4.
+
+In conclusion, by the comparison of the GSM with Coleman's framework, we have 
+found that despite Coleman's study contributes to define which factors hinder 
+the introduction of SPI %MUN: add: "of SPI in startups"?
+, his results are coherent with the theory generated within our research %MUN: 
+%add "our research, focused at explaining software development in general in 
+%startups." ?
+. In this regard, a wide variability of processes is a key factor for startups. 
+In contrast with some SPI processes, such as \textit{Six Sigma} which objective 
+is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, 
+startups seem to follow the \textit{Fischer's Fundamental Theorem of natural 
+selection} \cite{Fisher}. Moving towards flexible and variable processes (easy 
+to adapt to the uncertain conditions), increase the odds of ``natural 
+selection'' only when some ``restrictive conditions'' are met. In other words, 
+even though ignoring SPI practices is advisable %MUN: under which 
+%circumstances? This statement sounds too general, i.e. don't do SPI ever...
+%And the following part,I did not fully understand. Try to clarify.
+not following SPI does not predict major adaptiveness of startups development 
+to the market. But, to imply a major evolutionary product/market fit, certain 
+conditions explained in GSM need to be considered.
+
+%MUN: As I wrote already in the previous draft, you need to clarify at the 
+%beginning of this subsection which frameworks you are going to compare GSM to. 
+%Now, you say that Coleman is the only author who did research, comparable to 
+%yours, into startups. Now there comes Baskerville and Brooks. Write at the 
+%beginning of this section, shortly, which frameworks/authors you compare GSM 
+%to and why, then create subsections for each. 
+Also Baskerville \cite{Internet} refers to SPI as a development approach 
+typically effective only in large-scale, long-term development efforts with 
+stable and disciplined processes. By applying a grounded theory analysis in 10 
+companies he found that major causal factors that influence development are:
 
@@
 \item A new and unique software market environment.
-\item A lack of experience developing software under the conditions this environment imposed.
+\item A lack of experience developing software under the conditions this 
+environment imposed.
 
@@
 
-As discussed in CAT7, 1) can be mapped to \textit{time shortage} and 3) to \textit{limited access to expertise}, whilst  2) can refer to \textit{uncertain conditions make long-term planning not viable}, explained in CAT2.
-Even though with different research focus and context of study % - understand how and why Internet-speed software development (i.e.
-an agile method oriented using daily builds aimed at developing a product with high speed) differs from traditional software development - MUN: put this half-sentence where you introduce Baskerville's work
- Baskerville revealed similar causal factors as the GSM.
-
-Baskerville argues that the dawn of the Internet era has intensified software development problems by emphasizing shorter cycle times, stating to:
+As discussed in CAT7, 1) can be mapped to \textit{time shortage} and 3) to 
+\textit{limited access to expertise}, whilst  2) can refer to \textit{uncertain 
+conditions make long-term planning not viable}, explained in CAT2. Even though 
+with different research focus and context of study % - understand how and why 
+%Internet-speed software development (i.e. an agile method oriented using daily 
+%builds aimed at developing a product with high speed) differs from traditional 
+%software development - MUN: put this sentence where you introduce 
+%Baskerville's work
+, Baskerville revealed similar causal factors as the GSM.
+
+Baskerville argues that the dawn of the Internet era has intensified software 
+development problems by emphasizing shorter cycle times, stating to:
 
 \begin{compactitem}
-\item Speed-up development by releasing more often the software and ``implanting'' customers in the development environment (similar to the \textit{find the product/market fit} concept presented in CAT2).
-\item Make heavy use of simple tools and existing components (similar to the concepts expressed in CAT1).
-\item Invest time in facilitating development of scalable systems by the use of simple but stable architectural solutions (see subcategory \textit{using modularization and frameworks from day 1} discussed in CAT5).
-\item Tailor the development process daily according to the intense demands for speed.
-In this regard, he argues that the trend is to skip phases or tasks that might impede the ability to deliver software quickly even though producing lower quality software.
+\item Speed-up development by releasing more often the software and 
+``implanting'' customers in the development environment (similar to the 
+\textit{find the product/market fit} concept presented in CAT2). 
+\item Make heavy use of simple tools and existing components (similar to the 
+concepts expressed in CAT1).
+\item Invest time in facilitating development of scalable systems by the use of 
+simple but stable architectural solutions (see subcategory \textit{using 
+modularization and frameworks from day 1} discussed in CAT5). 
+\item Tailor the development process daily according to the intense demands for 
+speed. In this regard, he argues that the trend is to skip phases or tasks that 
+might impede the ability to deliver software quickly even though producing lower 
+quality software.
 
@@
 
-With a wider research focus, Brooks \cite{BrooksJr1987} discussed what challenges are involved in constructing software products.
-Brooks divides difficulties in development into essence (inherent to the nature of the software), and accidents (difficulties attending software production but which are not inherent).
-In other words essence, concerns the hard part of building a software through activities such as specification, design, testing.
-Accidents refer to the labor of representing the software or testing its representation.
-
-Brooks claims that the major effort applied by engineers in the last decades was dedicated towards accidents problems, trying to exploit new strategies to enhance software performance, reliability and simplicity of development, such as the introduction of high-level languages for programming.
-Despite the great achievements in improving development performance, the \textit{essence} property of the software remained unaltered.
-
-Essence difficulties are inherent to: complexity, conformity, changeability and invisibility.
-Complexity  mainly refers to interaction of software modules and elements between each others in some non-linear fashion, which increases accordingly to the project size.
-Conformity refers to the duty of software to adapt to human institutions and systems.
-Changeability  refers to the constant pressure of software products to accommodate culture, market, laws and hardware transitions.
-Invisibility  refers to the difficulty of representing software in space.
-
-Here we present the basic attacks (or mitigation strategies) presented in 1986 by Brooks on the conceptual essence ascertained within the GSM model description:
+With a wider research focus, Brooks \cite{BrooksJr1987} discussed what 
+challenges are involved in constructing software products. Brooks divides 
+difficulties in development into essence (inherent to the nature of the 
+software), and accidents (difficulties attending software production but which 
+are not inherent). In other words essence, concerns the hard part of building a 
+software through activities such as specification, design, testing. Accidents 
+refer to the labor of representing the software or testing its representation.
+
+Brooks claims that the major effort applied by engineers in the last decades 
+was dedicated towards accidents problems, trying to exploit new strategies to 
+enhance software performance, reliability and simplicity of development, such as 
+the introduction of high-level languages for programming.  Despite the great 
+achievements in improving development performance, the \textit{essence} property 
+of the software remained unaltered.
+
+Essence difficulties are inherent to: complexity, conformity, changeability and 
+invisibility. Complexity  mainly refers to interaction of software modules and 
+elements between each others in some non-linear fashion, which increases 
+accordingly to the project size. Conformity refers to the duty of software to 
+adapt to human institutions and systems. Changeability  refers to the constant 
+pressure of software products to accommodate culture, market, laws and hardware 
+transitions. Invisibility  refers to the difficulty of representing software in 
+space.
+
+Here we present the basic attacks (or mitigation strategies) presented in 1986 
+by Brooks on the conceptual essence ascertained within the GSM model 
+description:
 
@@
 
-\item Buy versus build-  the most radical possible solution for constructing software is not to construct it at all, taking advantage of what others have already implemented.
-It is the main strategy, which enables startups to externalize complexity to third party solutions explained within CAT1,  \textit{speed-up development}.
-\item Requirements refinement and rapid prototyping- avoid deciding precisely what to build but rather iteratively extract and refine the product requirements with customers and users.
-It represents the evolutionary development approach to maintain a fast release-cycles during development (CAT2).
-\item Incremental development- starting from simple solutions allows organizations to early prototype and control complexity overtime.
-It represents the main purpose of focusing on \textit{limited number of suitable functionalities} examined in \textit{product quality has low priority}  category (CAT3).
-\item Great team- people are the center of a software project and it is profoundly important to empower and liberate their creative mind.
-It is the main objective of empowering developers' capabilities described in \textit{team is the catalyst of development} (CAT4).
+\item Buy versus build-  the most radical possible solution for constructing 
+software is not to construct it at all, taking advantage of what others have 
+already implemented. It is the main strategy, which enables startups to 
+externalize complexity to third party solutions explained within CAT1,  
+\textit{speed-up development}.
+\item Requirements refinement and rapid prototyping- avoid deciding precisely 
+what to build but rather iteratively extract and refine the product requirements 
+with customers and users. It represents the evolutionary development approach to 
+maintain a fast release-cycles during development (CAT2).
+\item Incremental development- starting from simple solutions allows 
+organizations to early prototype and control complexity overtime. It represents 
+the main purpose of focusing on \textit{limited number of suitable 
+functionalities} examined in \textit{product quality has low priority}  category 
+(CAT3).
+\item Great team- people are the center of a software project and it is 
+profoundly important to empower and liberate their creative mind. It is the main 
+objective of empowering developers' capabilities described in \textit{team is 
+the catalyst of development} (CAT4).
 \end{compactitem}
 
-Brooks' forecast about software development strategies is revealed to be extremely accurate, according to the state-of-practice in modern startups.
-%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+Brooks' forecast about software development strategies is revealed to be 
+extremely accurate, according to the state-of-practice in modern startups.
+%------------------------------------------------------------------------------%
 \subsection{Theoretical categories and existing literature}
@@
 
-In this subsection we report the evaluation of our theory in contrast with the main contributions of the studies retrieved and discussed within the SMS \cite{SMS}.
-For each retrieved article we present the main results, mapping them to categories in accordance to the GSM model.
-Table \ref{tab:an:literature_comp} shows the main contributions of the 37 identified studies, where an 'X' is placed in correspondence to the theoretical category they considered.
-The table is sorted by the number of GSM categories covered by the studies.
+In this subsection we report the evaluation of our theory in contrast with the 
+main contributions of the studies retrieved and discussed within the SMS 
+\cite{SMS}. For each retrieved article we present the main results, mapping them 
+to categories in accordance to the GSM model. Table \ref{tab:an:literature_comp} 
+shows the main contributions of the 37 identified studies, where an 'X' is 
+placed in correspondence to the theoretical category they considered. The table 
+is sorted by the number of GSM categories covered by the studies.
 
@@
 Sutton (2000) & X & X & X & X & X & X & X & 7 & \cite{Sutton2000} \\
-Kajko-Mattson (2008) & X     & X     & X     & X     & X     & X     & X     & 7     & \cite{Kajko-Mattsson2008} \\
+Kajko-Mattson (2008) & X & X & X & X & X & X & X & 7 & 
+\cite{Kajko-Mattsson2008} \\
 Crowne (2002) & X & X & X & X & X & X & X & 7 & \cite{Crowne2002} \\
@@
 Chorev (2006) &  &  &  & X &  &  &  & 1 & \cite{Chorev2006} \\
-%MUN: add a row which shows the sum of each column. This shows how much categories are covered overall
+%MUN: add a row which shows the sum of each column. This shows how much 
+%categories are covered overall
 \hline
@@
 
-Only 7 studies out of 37, discuss aspects related to the overall software development process in early stage startups.
-The remaining part of the studies is only partially covering the phenomenon described by the GSM.
-Looking at the results%MUN: which results?
-The papers' main contribution or...?
+Only 7 studies out of 37, discuss aspects related to the overall software 
+development process in early stage startups. The remaining part of the studies 
+is only partially covering the phenomenon described by the GSM. Looking at the 
+results %MUN: which results? The papers' main contribution or...?
 , we were able to extrapolate information which confirmed the GSM categories.
 
-We can see how the majority of the retrieved studies (29) mention issues related to \textit{speed-up development} (CAT1) which confirms the importance of the core category.
-On the other hand, we can observe that only a limited number of studies mentions results related to \textit{product quality has low priority} (CAT3), \textit{accumulated technical debt} (CAT5), and \textit{initial grow hinders performance} (CAT6), suggesting directions of possible future primary studies.
-
-These results confirm the relevance of development teams as widely discussed in the SE literature, here mapping 26 studies to \textit{team is the catalyst of development} (CAT4).
-The importance of people has been widely discussed in other studies in SE (Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999}, Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph \cite{Adolph2011} and Cockburn \cite{people-first-order}), advocating for the need of empowering people, which is the element with the greatest impact on software development.
-Humans, being non-linear variables%MUN: add ref
-, are unlikely to follow repeatable prescriptive methodologies.
-As stated by Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and every methodology is to expect people to behave consistently over time, when is clearly not like that''}.
-Especially in early-stage startups, where the whole company overlaps with the development team, the GSM confirms the importance of factors related to people.
-
-%MUN: Think of structuring this long subsection. Have a subsubsection for each CAT, i.e. 5.2.1 Severe lack of resources (CAT7)
-%I've added to each category you mention the corresponding CATx since that identifier you use in table 2. I did not check if you discuss each and every category. Do you? If yes, ok. If not, why not?
-Starting from the category \textit{severe lack of resources} (CAT7), many authors describe startups facing with inescapable constraints defined by engineering/business concerns and constant time-pressure \cite{Camel1994a, Cugola98softwareprocesses:}.
-Because of their young and inexperienced working conditions, startups present limited resources in terms of management and strategic alliances \cite{Sutton2000}.
-
-Crowne \cite{Crowne2002} identifies four main stages of a startup's lifecycle: startup, stabilization, growth and maturity.
-Inexperienced practitioners operating in a chaotic environment characterizes the first phase, where there is no strategic plan for developing the product.
-But, moving towards the stabilization phase, the product starts to be unreliable and requirements unmanageable.
-Consequently, when startups move to the growth phase, they initiate structuring and controlling processes, which are incrementally integrated in the development environment.
-The maturity phase represents the moment when the company is ready to introduce process improvement initiatives.
-Crowne's study profoundly corresponds to different conceptualizations in the GSM.
-Starting from the minimal and low-precision engineering elements, startups start to grow the accumulated technical debt which requires fast integration of more structured and standard processes overtime.
-Issues related to low-precision engineering practices during companies' growth are discussed in an experience paper by Ambler \cite{Ambler2002}.
-He reports on two case studies of growing internet startups approaching an IPO.%MUN:explain acronym
-Despite the two companies were facing business and team growth, attitudes against the introduction of processes to structure and model the system were evident.
-The need to re-engineer the codebase lead the researchers to suggest a tailored version of the RUP%MUN: add reference
-development framework.
-Nevertheless, even with the aid of CASE tools\footnote{CASE tools are a class of software that automates many of the activities involved in various engineering activities.
-See \url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, startups failed in practicing engineering activities in view of eccesive lack of flexibility.
-Developers fast moved to simpler tools such as using a whiteboard, where they were able to provide input and insights, question assumptions, share
+We can see how the majority of the retrieved studies (29) mention issues 
+related to \textit{speed-up development} (CAT1) which confirms the importance of 
+the core category. On the other hand, we can observe that only a limited number 
+of studies mentions results related to \textit{product quality has low priority} 
+(CAT3), \textit{accumulated technical debt} (CAT5), and \textit{initial grow 
+hinders performance} (CAT6), suggesting directions of possible future primary 
+studies.
+
+These results confirm the relevance of development teams as widely discussed in 
+the SE literature, here mapping 26 studies to \textit{team is the catalyst of 
+development} (CAT4). The importance of people has been widely discussed in other 
+studies in SE (Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999}, 
+Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph 
+\cite{Adolph2011} and Cockburn \cite{people-first-order}), advocating for the 
+need of empowering people, which is the element with the greatest impact on 
+software development. Humans, being non-linear variables %MUN: add ref
+, are unlikely to follow repeatable prescriptive methodologies. As stated by 
+Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and every 
+methodology is to expect people to behave consistently over time, when is 
+clearly not like that''}. Especially in early-stage startups, where the whole 
+company overlaps with the development team, the GSM confirms the importance of 
+factors related to people.
+
+%MUN: Think of structuring this long subsection. Have a subsubsection for each 
+%CAT, i.e. 5.2.1 Severe lack of resources (CAT7)
+%I've added to each category you mention the corresponding CATx since that  
+%identifier you use in table 2. I did not check if you discuss each and every 
+%category. Do you? If yes, ok. If not, why not?
+Starting from the category \textit{severe lack of resources} (CAT7), many 
+authors describe startups facing with inescapable constraints defined by 
+engineering/business concerns and constant time-pressure \cite{Camel1994a, 
+Cugola98softwareprocesses:}. Because of their young and inexperienced working 
+conditions, startups present limited resources in terms of management and 
+strategic alliances \cite{Sutton2000}.
+
+Crowne \cite{Crowne2002} identifies four main stages of a startup's lifecycle: 
+startup, stabilization, growth and maturity. Inexperienced practitioners 
+operating in a chaotic environment characterizes the first phase, where there is 
+no strategic plan for developing the product. But, moving towards the 
+stabilization phase, the product starts to be unreliable and requirements 
+unmanageable. Consequently, when startups move to the growth phase, they 
+initiate structuring and controlling processes, which are incrementally 
+integrated in the development environment. The maturity phase represents the 
+moment when the company is ready to introduce process improvement initiatives. 
+Crowne's study profoundly corresponds to different conceptualizations in the 
+GSM. Starting from the minimal and low-precision engineering elements, startups 
+start to grow the accumulated technical debt which requires fast integration of 
+more structured and standard processes overtime.
+Issues related to low-precision engineering practices during companies' growth 
+are discussed in an experience paper by Ambler \cite{Ambler2002}. He reports on 
+two case studies of growing internet startups approaching an IPO. %MUN:explain 
+%acronym
+Despite the two companies were facing business and team growth, attitudes 
+against the introduction of processes to structure and model the system were 
+evident. The need to re-engineer the codebase lead the researchers to suggest a 
+tailored version of the RUP %MUN: add reference
+development framework. Nevertheless, even with the aid of CASE 
+tools\footnote{CASE tools are a class of software that automates many of the 
+activities involved in various engineering activities. See 
+\url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, 
+startups failed in practicing engineering activities in view of eccesive lack of 
+flexibility. Developers fast moved to simpler tools such as using a whiteboard, 
+where they were able to provide input and insights, question assumptions, share 
 past experiences, and even talk about potential implementation strategies. 
-%MUN: Candidate for removal since details can be found in Ambler anyway: "In general the development remained highly iterative, where project teams started to have a little modeling and testing only when necessary.
-They didn't take the serial approach of completely developing and accepting the requirements model, analysis model and so on.
-This let them react swiftly and adapt rapidly according to changes into their highly competitive markets.
-The researchers report \textit{``modeling was streamlined''}, that is modeling what was needed and go straightforward to coding, avoiding analysis paralysis."
-
-The use of whiteboards in working environments supports \textit{use of well-integrated and simple tools}.
-In both the startups studied by Ambler, documentation was minimal, finding the ``sweet spot'' to have just barely enough tools and artifacts to meet their needs.
-Generally, the use of \textit{keep a simple and informal workflow} endorses the use of low-precision engineering practices which lead to speed-up development initially, but with negative effects on the accumulated technical debt of a minimal project management and consequently on the final company growth.
-
-Kajko-Mattsson et al.
-\cite{Kajko-Mattsson2008} investigated a Swedish software startup company involved in mobile applications, reporting the following issues: a lack of requirements gathering process; release cycles and their length were not defined; the releases and their scope were not planned; lack of control over the change requests; lack of process control in terms of  absence of any documentation to track the status and progress of the process; defective releases in terms of lack of testing; poor communication since the communication inside the company was informal and not documented at all.
-In the same study they designed a process improvement model, based on three main phases: Evaluate, Plan and Change.
-
-%MUN: the whole paragraph needs to be summarized to its essence. Details can be read in the referred paper
-In the evaluation phase, the status of the development workflow within the company  is analyzed.
-The researchers documented a state of chaos where any initiatives of quantitative methodology assessment would have been a waste of time.
-Therefore, they applied a qualitative methodology to define a model of the release management process.
-They moved towards the plan phase where the researchers decided to develop a model to manage releases, introducing a process composed by: release scope preparation, release planning, release development, system testing, acceptance testing and release deployment.
-Despite the effort spent by the researchers to establish a development workflow, practices were not fully conducted even though some marginal benefits have been reported: planned major release from four to six months; planned minor releases from two to four weeks; encompassing urgent corrective releases from one to five days.
-Regarding the previously mentioned problems, only
-partial improvements have been achieved since the major obstacle was to motivate employees to change their habits.
-With perfect hindsight the researchers claim that understanding and adapting solutions by solving one problem at a time would have increased the benefits.
-Then, despite the reported results, starting the establishment of a simple and informal workflow aided the organization to move from the stage of knowing nothing to knowing at least something about their development process.
-
-The integration of minimal project management during the development activities can foster the control over development productivity and control the \textit{accumulated technical debt} (CAT5), but still with particular attention to the development workflow status within the company, designing specific and tailored solutions.
-
-Prescribed development models such as SPI in startups are basically ignored by startups, as discussed in \cite{Zettel2001}.%MUN: <-- rephrase, not clear to me
-To overcome the need of structure and to accelerate the development process, Zettel developed a \textit{``LIghtweight Process for E-business software development''} (LIPE).
-This lightweight development method integrates the XP approach with ideas from areas of software measurement for project control and process improvement as a compromise between an ad-hoc and a more rigorous approach.
-Zettel argues that LIPE provides scalability of the development process with a certain degree of flexibility, omitting parts of prescribed practices.
-Nevertheless, we couldn't identify any empirical evaluation of LIPE that supports its suitability in a startup context.
-
-Deakins \cite{Deakins2005} designed a new spiral methodology, evolved during fieldwork with a developer team that had limited resources managing an innovative product in volatile e-commerce environments.
-He proposes fast evolutionary approach experimentation with flexible processes obtaining customer's feedback as soon as possible for continuous product improvement.%MUN: <-- rephrase, sentence not parsable
+%MUN: Candidate for removal since details can be found in Ambler anyway: "In 
+%general the development remained highly iterative, where project teams started 
+%to have a little modeling and testing only when necessary. They didn't take 
+%the serial approach of completely developing and accepting the requirements 
+%model, analysis model and so on. This let them react swiftly and adapt 
+%rapidly according to changes into their highly competitive markets. The 
+%researchers report \textit{``modeling was streamlined''}, that is modeling 
+%what was needed and go straightforward to coding, avoiding analysis 
+%paralysis."
+
+The use of whiteboards in working environments supports \textit{use of 
+well-integrated and simple tools}. In both the startups studied by Ambler, 
+documentation was minimal, finding the ``sweet spot'' to have just barely enough 
+tools and artifacts to meet their needs. Generally, the use of \textit{keep a 
+simple and informal workflow} endorses the use of low-precision engineering 
+practices which lead to speed-up development initially, but with negative 
+effects on the accumulated technical debt of a minimal project management and 
+consequently on the final company growth.
+
+Kajko-Mattsson et al. \cite{Kajko-Mattsson2008} investigated a Swedish software 
+startup company involved in mobile applications, reporting the following issues: 
+a lack of requirements gathering process; release cycles and their length were 
+not defined; the releases and their scope were not planned; lack of control over 
+the change requests; lack of process control in terms of  absence of any 
+documentation to track the status and progress of the process; defective 
+releases in terms of lack of testing; poor communication since the communication 
+inside the company was informal and not documented at all. In the same study 
+they designed a process improvement model, based on three main phases: Evaluate, 
+Plan and Change.
+
+%MUN: the whole paragraph needs to be summarized to its essence. Details can be 
+%read in the referred paper
+In the evaluation phase, the status of the development workflow within the 
+company  is analyzed. The researchers documented a state of chaos where any 
+initiatives of quantitative methodology assessment would have been a waste of 
+time. Therefore, they applied a qualitative methodology to define a model of the 
+release management process. They moved towards the plan phase where the 
+researchers decided to develop a model to manage releases, introducing a process 
+composed by: release scope preparation, release planning, release development, 
+system testing, acceptance testing and release deployment. Despite the effort 
+spent by the researchers to establish a development workflow, practices were not 
+fully conducted even though some marginal benefits have been reported: planned 
+major release from four to six months; planned minor releases from two to four 
+weeks; encompassing urgent corrective releases from one to five days. Regarding 
+the previously mentioned problems, only 
+partial improvements have been achieved since the major obstacle was to 
+motivate employees to change their habits. With perfect hindsight the 
+researchers claim that understanding and adapting solutions by solving one 
+problem at a time would have increased the benefits. Then, despite the reported 
+results, starting the establishment of a simple and informal workflow aided the 
+organization to move from the stage of knowing nothing to knowing at least 
+something about their development process.
+
+The integration of minimal project management during the development activities 
+can foster the control over development productivity and control the 
+\textit{accumulated technical debt} (CAT5), but still with particular attention 
+to the development workflow status within the company, designing specific and 
+tailored solutions.
+
+Prescribed development models such as SPI in startups are basically ignored by 
+startups, as discussed in \cite{Zettel2001}. %MUN: <-- rephrase, not clear to me
+To overcome the need of structure and to accelerate the development process, 
+Zettel developed a \textit{``LIghtweight Process for E-business software 
+development''} (LIPE). This lightweight development method integrates the XP 
+approach with ideas from areas of software measurement for project control and 
+process improvement as a compromise between an ad-hoc and a more rigorous 
+approach. Zettel argues that LIPE provides scalability of the development 
+process with a certain degree of flexibility, omitting parts of prescribed 
+practices. Nevertheless, we couldn't identify any empirical evaluation of LIPE 
+that supports its suitability in a startup context.
+
+Deakins \cite{Deakins2005} designed a new spiral methodology, evolved during 
+fieldwork with a developer team that had limited resources managing an 
+innovative product in volatile e-commerce environments. He proposes fast 
+evolutionary approach experimentation with flexible processes obtaining 
+customer's feedback as soon as possible for continuous product improvement.%MUN: 
+%<-- rephrase, sentence not parsable
 Even in this case the model%MUN: which model?
-is compliant with the nature of a random, ad-hoc and visionary product development.
-In fact, evolutionary development approaches are typically more suitable  for web applications, as confirmed in \cite{Deshpande2001}.
-Then, the practices are oriented towards rapid and iterative design with complete freedom to the developer team, enabled to operate anywhere within the organization, according also to the category \textit{team is the catalyst of development}.%MUN: <-- the whole paragraph is not that clear to me.
-Try to start with the category (team is catalyst), then write about Deakins spiral model
-
-Carmel \cite{Camel1994a} reports on the need of flexible and rapid development solutions to shorten time-to-market, but even more important, the need of high-skilled team developers, which has an impact on \textit{speed-up development} (CAT1).
-He suggests that entrepreneurs need to look for a well formed, skilled core development team and not just a set of product ideas and features.
-Moreover the team must be empowered with full-stack and self-organization settings, paired with the flexibility of minimal bureaucracy.
-A studied company stated: \textit{``we didn't need weeks or months of detailed modeling and documentation, but rather modeling the architecture a little and then either exploring strategies by providing users with code or simply starting work on the actual software itself''}.
+is compliant with the nature of a random, ad-hoc and visionary product 
+development. In fact, evolutionary development approaches are typically more 
+suitable  for web applications, as confirmed in \cite{Deshpande2001}. Then, the 
+practices are oriented towards rapid and iterative design with complete freedom 
+to the developer team, enabled to operate anywhere within the organization, 
+according also to the category \textit{team is the catalyst of 
+development}.%MUN: <-- the whole paragraph is not that clear to me. Try to start 
+%with the category (team is catalyst), then write about Deakins spiral model
+
+Carmel \cite{Camel1994a} reports on the need of flexible and rapid development 
+solutions to shorten time-to-market, but even more important, the need of 
+high-skilled team developers, which has an impact on \textit{speed-up 
+development} (CAT1). He suggests that entrepreneurs need to look for a well 
+formed, skilled core development team and not just a set of product ideas and 
+features. Moreover the team must be empowered with full-stack and 
+self-organization settings, paired with the flexibility of minimal bureaucracy. 
+A studied company stated: \textit{``we didn't need weeks or months of detailed 
+modeling and documentation, but rather modeling the architecture a little and 
+then either exploring strategies by providing users with code or simply starting 
+work on the actual software itself''}.
 
 %MUN: To which category does the following paragraph belong?
-Eventually to focus on simple solutions%MUN: solutions on what?
-, two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the wheel, instead of reinventing it''}, taking advantage of open source software whenever it is possible, building systems quickly and effectively as a result.
-Even though the principles in leading-edge development processes are the same principles of yesteryear, the way in which the fundamentals are applied has changed from prescriptive to agile solutions, with less concern on quality aspects, also confirmed by Mater \cite{Mater2000}.
-
-%MUN: When you subdivide this sections, as proposed above, finish with a  subsubsection "Summary", summarizing the supported you have found in literature for your theoretical categories.
-Nothing extensive, 3-4 sentences to conclude this part of the validation, similar as below but a bit more explanatory (what does "compliant with the behavioural model of the phenomenon" mean?)
-Thus, by summarizing the comparison, we have found that most of the contributions of the studies are compliant with the behavioural model of the phenomenon we provided.
-
-%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+Eventually to focus on simple solutions %MUN: solutions on what?
+, two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the 
+wheel, instead of reinventing it''}, taking advantage of open source software 
+whenever it is possible, building systems quickly and effectively as a result. 
+Even though the principles in leading-edge development processes are the same 
+principles of yesteryear, the way in which the fundamentals are applied has 
+changed from prescriptive to agile solutions, with less concern on quality 
+aspects, also confirmed by Mater \cite{Mater2000}.
+
+%MUN: When you subdivide this sections, as proposed above, finish with a  
+%subsubsection "Summary", summarizing the supported you have found in 
+%literature for your theoretical categories. Nothing extensive, 3-4 sentences 
+%to conclude this part of the validation, similar as below but a bit more 
+%explanatory (what does "compliant with the behavioural model of the 
+%phenomenon" mean?)
+
+Thus, by summarizing the comparison, we have found that most of the 
+contributions of the studies are compliant with the behavioural model of the 
+phenomenon we provided.
+
+%------------------------------------------------------------------------------%
 \subsection{Confounding factors}
@@
 
-The purpose of this subsection is to identify which confounding factors need to be taken into consideration when evaluating the GSM.
-Confounding factors are those variables which and have not been considered in the model and might interfere with the theoretical framework positively or negatively \cite{ColinRobson2009}.
-We report those factors that have been explicitly identified by the literature review%MUN: do you mean the SMS?
-Then refer to it.
-I remember you mentioned CF in the results (GSM), maybe also refer to the respective sections
-, related to: creativity and innovation, market requirements and application type, and developer experience.
-
-%MUN: Again, structure this into subsubsections according to the above factors. I did it, but check if you agree...
+The purpose of this subsection is to identify which confounding factors need to 
+be taken into consideration when evaluating the GSM. Confounding factors are 
+those variables which and have not been considered in the model and might 
+interfere with the theoretical framework positively or negatively 
+\cite{ColinRobson2009}. We report those factors that have been explicitly 
+identified by the literature review %MUN: do you mean the SMS? Then refer to it. 
+%I remember you mentioned CF in the results (GSM), maybe also refer to the 
+%respective sections
+, related to: creativity and innovation, market requirements and application 
+type, and developer experience.
+
+%MUN: Again, structure this into subsubsections according to the above factors. 
+%I did it, but check if you agree...
 \subsubsection{Creativity and innovation} 
-Heitlager et al.
-\cite{Heitlager2007} contributed to the investigation of creativity and innovation in startups.
-%MUN: the rest of this paragraph is too detailed.
-I'd remove this "The study reports how development approaches oriented to the product in the early-stage life-cycle, called \textit{fluid} have main impact to innovative solutions initially.
-After a while the need of structuring and controlling the growing company size (along with \textit{transition} and \textit{specific} phases) arises the need of project management activities and software processes for long-term scalability issues."
+Heitlager et al. \cite{Heitlager2007} contributed to the investigation of 
+creativity and innovation in startups. 
+%MUN: the rest of this paragraph is too detailed. I'd remove this "The study 
+%reports how development approaches oriented to the product in the early-stage 
+%life-cycle, called \textit{fluid} have main impact to innovative solutions 
+%initially. After a while the need of structuring and controlling the growing 
+%company size (along with \textit{transition} and \textit{specific} phases) 
+%arises the need of project management activities and software processes for 
+%long-term scalability issues."
 
 %MUN: Please reformulate the first sentence, I cannot parse it
-Based on the model presented in \cite{Heitlager2007}, it identifies the lifecycle stages of a startup (\textit{fluid}, \textit{transition} and \textit{specific}), corresponding for each stage a development approach (product-oriented and process-oriented) the level of innovation achieved.
-The study reports how product-oriented development, in contrast with forced and mature process imposition, provides a degree of freedom to the development team, that enhances the creativity of developers and augments the innovation capability of the company in the early-stage.
-
-In the interviews we conducted, and on which the GSM is grounded, the theme of innovation only seldom emerged.
-However, if innovative and creative factors were considered in the model, the integration of results would have been straightforward, since the evolution of this model is totally compliant with our assumption of CAT5 and CAT6.
-%MUN: not clear to me what you mean by "evolution of this model" and what "compliant with our assumption" means.
-Also, don't use only CATx in this context, but cite the category name AND CATx
-From product-oriented development in the \textit{fluid} phase to the process-oriented development in the \textit{specific} phase, the innovation confounding factor seems to behave initially the same as the \textit{speed-up development}.%MUN: <-- please provide some motivation...
-not obvious
- Nevertheless, we were unable to ground these evidences in the GT data even though they might present strict correlation%MUN: "present strict correlation" What does that mean, I don't understand
- with the choice of an early product-development orientation with low-precision engineering elements.
+Based on the model presented in \cite{Heitlager2007}, it identifies the 
+lifecycle stages of a startup (\textit{fluid}, \textit{transition} and 
+\textit{specific}), corresponding for each stage a development approach 
+(product-oriented and process-oriented) the level of innovation achieved. The 
+study reports how product-oriented development, in contrast with forced and 
+mature process imposition, provides a degree of freedom to the development team, 
+that enhances the creativity of developers and augments the innovation 
+capability of the company in the early-stage.
+
+In the interviews we conducted, and on which the GSM is grounded, the theme of 
+innovation only seldom emerged. However, if innovative and creative factors were 
+considered in the model, the integration of results would have been 
+straightforward, since the evolution of this model is totally compliant with our 
+assumption of CAT5 and CAT6. %MUN: not clear to me what you mean by "evolution 
+%of this model" and what "compliant with our assumption" means. Also, don't use 
+%only CATx in this context, but cite the category name AND CATx
+From product-oriented development in the \textit{fluid} phase to the 
+process-oriented development in the \textit{specific} phase, the innovation 
+confounding factor seems to behave initially the same as the \textit{speed-up 
+development}.%MUN: <-- please provide some motivation... not obvious
+Nevertheless, we were unable to ground these evidences in the GT data even 
+though they might present strict correlation %MUN: "present strict correlation" 
+%What does that mean, I don't understand
+with the choice of an early product-development orientation with low-precision 
+engineering elements.
 
 \subsubsection{Market requirements and application type} 
-Market requirements \cite{Coleman2007} and application type \cite{Sutton2000, Coleman2008, Coleman2008a} are two further potential confounding factors to the GSM.
-Their main impact is related to the adoption of flexible and reactive solutions for the development process.
-In particular, they refer to the necessity of fulfillment of quality concerns that goes beyond scalability and UX defined by the GSM.
-Especially when requirements are rigidly imposed or the application domain is well-known, providing low-quality products to final users might determine the failure of a startup.
-
-For example, safety-critical systems possess quality concerns that must be considered.
-But also minor risky application domains such as e-commerce systems have high-impact factors for obtaining the first customer reference,%MUN: <-- cannot parse it.
-Remove the first part and write: "Even e-commerce systems...".
-Then rephrase the second part of the sentence to make it clearer.
- as discussed in \cite{Deakins2005,Kim2005,Silva2005}. This%MUN: what is "this"?
- characterizes those applications, whose part of functionalities are already successfully implemented in other software systems or whose failure might inflict serious financial or social damages \cite{Bass2003}.
-Yet, limiting our context to early-stage software startups, which typically develop first cutting-edge products in beta versions, partially mitigates the high quality demands of users.
-
-%MUN: The following 4 paragraphs explain how you use the Kano model to position startups w.r.t.
-customer expectations and quality concerns.
-It's ok, but it adds little value to the validation of GSM.
-I'd remove it.
-% To understand how to obtain processes in accordance to a set of customer needs, correlating the proper levels of customers' satisfaction, we introduce the Kano model, described in \cite{Ullah2011} and extensively studied in product development.
-As shown in the Kano model, users preferences can be classified into four categories: \textit{basic}, \textit{excitement}, \textit{performance} and \textit{indifference} areas.
-% 
-% The \textit{basic} area represents the space where customers are dissatisfied if those \textit{must-be} functionalities are not met.
-This is the case of similar functionalities already implemented in the current market with well-known solutions.
-The \textit{performance} area represents the space where customers' satisfaction is proportional to how many of functionalities are implemented.
-In the \textit{indifference} area the customer does not really care about the functionalities: the satisfaction remains invariant.
-This represents the case where startups have not (yet) achieved the right product/market fit.
-Finally, the \textit{excitement} area represents the space where customers will be more satisfied if functionalities are implemented, but not disappointed if they are not implemented.
-This is the case of extremely innovative product development.
-% 
-% Well-known domains such as e-commerce applications or other \textit{``quality critical products''} belong to the \textit{basic} area, where if a quality concern such as security in bank transfer services is missing satisfaction of customers will be profoundly affected.
-But as discussed, in early-stage startups that typically develop first beta products, developers usually deal with cutting-edge technologies.
-Those innovative solutions, if in the right market, are likely to belong to  the \textit{excitement} area, where satisfaction can hardly be negatively affected by faulty products.
-In fact, in the GSM model, minimal functionalities and less concern about quality aspects have low-impact on customers' satisfaction, supporting startup's ability to mainly focus on \textit{speed-up development}.
-Moreover operating in the \textit{excitement area} is fully compliant with an evolutionary approach and minimal functionalities since from the first product released to the market, it is unlikely to lose users because
+Market requirements \cite{Coleman2007} and application type \cite{Sutton2000, 
+Coleman2008, Coleman2008a} are two further potential confounding factors to the 
+GSM. Their main impact is related to the adoption of flexible and reactive 
+solutions for the development process. In particular, they refer to the 
+necessity of fulfillment of quality concerns that goes beyond scalability and UX 
+defined by the GSM. Especially when requirements are rigidly imposed or the 
+application domain is well-known, providing low-quality products to final users 
+might determine the failure of a startup.
+
+For example, safety-critical systems possess quality concerns that must be 
+considered. But also minor risky application domains such as e-commerce systems 
+have high-impact factors for obtaining the first customer reference, %MUN: <-- 
+%cannot parse it. Remove the first part and write: "Even e-commerce 
+%systems...". Then rephrase the second part of the sentence to make it clearer.
+as discussed in \cite{Deakins2005,Kim2005,Silva2005}. This %MUN: what is "this"?
+characterizes those applications, whose part of functionalities are already 
+successfully implemented in other software systems or whose failure might 
+inflict serious financial or social damages \cite{Bass2003}. Yet, limiting our 
+context to early-stage software startups, which typically develop first 
+cutting-edge products in beta versions, partially mitigates the high quality 
+demands of users.
+
+%MUN: The following 4 paragraphs explain how you use the Kano model to position 
+%startups w.r.t. customer expectations and quality concerns. It's ok, but it 
+%adds little value to the validation of GSM. I'd remove it.
+%To understand how to obtain processes in accordance to a set of customer 
+%needs, correlating the proper levels of customers' satisfaction, we 
+%introduce the Kano model, described in \cite{Ullah2011} and extensively 
+%studied in product development.  As shown in the Kano model, users 
+%preferences can be classified into four categories: \textit{basic}, 
+%\textit{excitement}, \textit{performance} % and \textit{indifference} areas.
+% 
+%The \textit{basic} area represents the space where customers are dissatisfied 
+%if those \textit{must-be} functionalities are not met. This is the case of 
+%similar functionalities already implemented in the current market with 
+%well-known solutions. The \textit{performance} area represents the space where 
+%customers' satisfaction is proportional to how many of functionalities are 
+%implemented. In the \textit{indifference} area the customer does not really 
+%care about the functionalities: the satisfaction remains invariant. This 
+%represents the case where startups have not (yet) achieved the right 
+%product/market fit. Finally, the \textit{excitement} area represents the space 
+%where customers will be more satisfied if functionalities are implemented, but 
+%not disappointed if they are not implemented. This is the case of extremely 
+%innovative product development.
+% 
+%Well-known domains such as e-commerce applications or other \textit{``quality 
+%critical products''} belong to the \textit{basic} area, where if a quality 
+%concern such as security in bank transfer services is missing satisfaction of 
+%customers will be profoundly affected. But as discussed, in early-stage 
+%startups that typically develop first beta products, developers usually deal 
+%with cutting-edge technologies. Those innovative solutions, if in the right 
+%market, are likely to belong to  the \textit{excitement} area, where 
+%satisfaction can hardly be negatively affected by faulty products. In fact, in 
+%the GSM model, minimal functionalities and less concern about quality aspects 
+%have low-impact on customers' satisfaction, supporting startup's ability to 
+%mainly focus on \textit{speed-up development}. Moreover operating in the 
+%\textit{excitement area} is fully compliant with an evolutionary approach and 
+%minimal functionalities since from the first product released to the market, 
+%it is unlikely to lose users because 
 % of high-quality demand.
 % 
-% Nevertheless, this aspect profoundly undermines the generalizability of the theoretical model according to the theory, in particular \textit{CAT3}.
-Therefore, future studies within wider startups' context are advisable to extrapolate insights from those situations where quality concerns cannot be postponed or externalized.
+%Nevertheless, this aspect profoundly undermines the generalizability of the 
+%theoretical model according to the theory, in particular \textit{CAT3}. 
+%therefore, future studies within wider startups' context are advisable to 
+%extrapolate insights from those situations where quality concerns cannot be 
+%postponed or externalized.
 
@@
 %MUN: Ups, a confounding factor you did not mention initially. Add it...
-Finally, also the experience of team-members is considered as an important confounding factor.
-Since only little evidence has been gathered during the interviews, ``lack of experience'' was not considered as a sub-category in \textit{severe lack of resources}.
-Nevertheless, startups rely at the beginning on clever, but inexperienced developers \cite{Crowne2002}.
-Having team-members with deep experience is a ``double-edge sword'' \cite{Yoffie1999}.
-On one hand, experience might quickly provide structure and maturity to the development process, on the other it might cause challenges in managing self-confident overachievers that almost inevitably clash.
-Consequently, team management requires higher control and coordination activities that inevitably hinder the flexibility of the development environment, that is essential in early-stage startups.
-Accordingly, also Coleman \cite{Coleman2007,
-Coleman2008a,Coleman2008} states that the operating strategies towards \textit{``minimum processes''} in startups is not a matter of poor knowledge and training, but rather it is the necessity of operating with solutions that let the company move faster.
+Finally, also the experience of team-members is considered as an important 
+confounding factor. Since only little evidence has been gathered during the 
+interviews, ``lack of experience'' was not considered as a sub-category in 
+\textit{severe lack of resources}. Nevertheless, startups rely at the beginning 
+on clever, but inexperienced developers \cite{Crowne2002}. Having team-members 
+with deep experience is a ``double-edge sword'' \cite{Yoffie1999}. On one hand, 
+experience might quickly provide structure and maturity to the development 
+process, on the other it might cause challenges in managing self-confident 
+overachievers that almost inevitably clash. Consequently, team management 
+requires higher control and coordination activities that inevitably hinder the 
+flexibility of the development environment, that is essential in early-stage 
+startups. Accordingly, also Coleman \cite{Coleman2007,
+Coleman2008a,Coleman2008} states that the operating strategies towards 
+\textit{``minimum processes''} in startups is not a matter of poor knowledge and 
+training, but rather it is the necessity of operating with solutions that let 
+the company move faster.
 
 %MUN: removed below paragraph since it does not add much
-% Nevertheless, since the amount of the results did not allow ``the lack of experience'' to be in the GSM model (in view of the fact that not enough data were identified \footnote{A category is formed if the codes are representative of more than the 70\% of the analyzed transcripts.}), this confounding factor requires further investigation.
-
-%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%%--------------------------%
+%Nevertheless, since the amount of the results did not allow ``the lack of 
+%experience'' to be in the GSM model (in view of the fact that not enough data 
+%were identified \footnote{A category is formed if the codes are representative 
+%of more than the 70\% of the analyzed transcripts.}), this confounding factor 
+%requires further investigation.
+
+%------------------------------------------------------------------------------%
 \subsection{Static assessment of the GSM}
@@
 
-%MUN: explain in the research method section (new section on how you performed validation), exactly how you did static validation and what statistical methods you used on what data, then refer to that section from here (no need to repeat again here).
-In this subsection we validate the most important causal relations presented in our model by attesting their correctness  with a quantification of empirical trends observed in the interview data.
-We identify trends and recurrent patterns by thoroughly analyzing and synthesizing the complete set of data obtained from the case study on which the GSM is built upon.
-We illustrate our observations aided by simple statistical methods, providing researchers an additional tool, supported by further companies' data, to validate our model.
-
-We argued in section \ref{res:gsm:frmov}, that the most essential feature startups aim to achieve is high speed.
-As illustrated in Figure \ref{fig:gsm}, the core category (\textit{speed-up development}) is influenced by: the degree of quality concerns (contributes to the \textit{efficiency}); the undertaken development approach (provides \textit{effectiveness}); and finally the intrinsic characteristics of the team (guarantee \textit{performance}).
+%MUN: explain in the research method section (new section on how you performed 
+%validation), exactly how you did static validation and what statistical 
+%methods you used on what data, then refer to that section from here (no need 
+%to repeat again here).
+In this subsection we validate the most important causal relations presented in 
+our model by attesting their correctness  with a quantification of empirical 
+trends observed in the interview data. We identify trends and recurrent patterns 
+by thoroughly analyzing and synthesizing the complete set of data obtained from 
+the case study on which the GSM is built upon. We illustrate our observations 
+aided by simple statistical methods, providing researchers an additional tool, 
+supported by further companies' data, to validate our model.
+
+We argued in section \ref{res:gsm:frmov}, that the most essential feature 
+startups aim to achieve is high speed. As illustrated in Figure \ref{fig:gsm}, 
+the core category (\textit{speed-up development}) is influenced by: the degree 
+of quality concerns (contributes to the \textit{efficiency}); the undertaken 
+development approach (provides \textit{effectiveness}); and finally the 
+intrinsic characteristics of the team (guarantee \textit{performance}).
 
 %MUN: for the paragraphs below, you need to explain what and why you do this. 
-What do you mean by "we grounded in the interview transcripts the fact that those relations were characterized in terms of performances since time-to-market was the main concern." Not clear to me.
-Before starting the evaluation of the impact of the three categories on \textit{speed-up development}, we grounded in the interview transcripts the fact that those relations were characterized in terms of performances since time-to-market was the main concern (as confirmed in \cite{SMS}):
+%What do you mean by "we grounded in the interview transcripts the fact that 
+%those relations were characterized in terms of performances since 
+%time-to-market was the main concern." Not clear to me.
+Before starting the evaluation of the impact of the three categories on 
+\textit{speed-up development}, we grounded in the interview transcripts the fact 
+that those relations were characterized in terms of performances since 
+time-to-market was the main concern (as confirmed in \cite{SMS}):
 
@@
 
-\item The team enhances the overall \textit{performance} of software development.
-Recalling that startups deal with severe lack of resources, the team members are the company's main assets.
-They develop software with no need of formalities and structures, left to the team's own capabilities to efficiently develop the product.
-Moreover, the effectiveness is granted by continuous advice given by mentors and by a management style oriented towards self-organization, full-stack, and multi-role settings.
-The importance of a skilled team for development performance is expressed in many studies such as \cite{surviving-os-cockburn, Highsmith2000}, described in CAT4 and confirmed by Sutton \cite{Sutton2000}.
-\item The evolutionary approach enhances the \textit{effectiveness} in implementing the correct functionalities.
-This approach improves the company's capabilities to adjust the trajectory of product development (in agreement to the concept of pivoting discussed by Ries \cite{Ries2011}).
-\item Since product quality has low priority, the development is focused only on implementing a limited number of suitable functionalities, enhancing its \textit{efficiency}.
-An essential prototype (MVP) does not require to comply with heavy quality constraints, enabling the team to quickly implement functionalities, ready to be validated by final users.
+\item The team enhances the overall \textit{performance} of software 
+development. Recalling that startups deal with severe lack of resources, the 
+team members are the company's main assets. They develop software with no need 
+of formalities and structures, left to the team's own capabilities to 
+efficiently develop the product. Moreover, the effectiveness is granted by 
+continuous advice given by mentors and by a management style oriented towards 
+self-organization, full-stack, and multi-role settings. The importance of a 
+skilled team for development performance is expressed in many studies such as 
+\cite{surviving-os-cockburn, Highsmith2000}, described in CAT4 and confirmed by 
+Sutton \cite{Sutton2000}.
+\item The evolutionary approach enhances the \textit{effectiveness} in 
+implementing the correct functionalities. This approach improves the company's 
+capabilities to adjust the trajectory of product development (in agreement to 
+the concept of pivoting discussed by Ries \cite{Ries2011}).
+\item Since product quality has low priority, the development is focused only on 
+implementing a limited number of suitable functionalities, enhancing its 
+\textit{efficiency}. An essential prototype (MVP) does not require to comply 
+with heavy quality constraints, enabling the team to quickly implement 
+functionalities, ready to be validated by final users.
 \end{compactitem}
 
-An optimal combination of these three categories enables startups to ship code extremely quickly.
-However, the price that startups have to pay for achieving such high speed is related to the extent of the \textit{accumulated technical debt}, which will finally contribute to hinder performance in later stages.
-
-Accordingly, by analysing the data of the 13 companies, we observed two distinct trends which further empower the above illustrated relations:
-%MUN: You need to be more precise with terminology.
-Above you state that the two items below are "trends which further empower relations", but then, further below, you write "to assess these hypotheses".
-Observed trends != Hypotheses.
-This is also why you need to be extremely explicit and precise when explaining what the a) aim, b) method, and c) outcome/result of this section is.
-I'm aware that it is difficult to boil down what you wrote in the thesis to this limited space, but I think it is doable (and relevant for supporting validity of GSM)
+An optimal combination of these three categories enables startups to ship code 
+extremely quickly. However, the price that startups have to pay for achieving 
+such high speed is related to the extent of the \textit{accumulated technical 
+debt}, which will finally contribute to hinder performance in later stages.
+
+Accordingly, by analysing the data of the 13 companies, we observed two 
+distinct trends which further empower the above illustrated relations:
+%MUN: You need to be more precise with terminology. Above you state that the 
+%two items below are "trends which further empower relations", but then, 
+%further below, you write "to assess these hypotheses". Observed trends != 
+%Hypotheses. This is also why you need to be extremely explicit and precise 
+%when explaining what the a) aim, b) method, and c) outcome/result of this 
+%section is. I'm aware that it is difficult to boil down what you wrote in the 
+%thesis to this limited space, but I think it is doable (and relevant for 
+%supporting validity of GSM)
 
 \begin{compactenum}
-\item The more a capable \footnote{The team capability is evaluated according to the characteristics presented in CAT4.} team, with few quality constraints, uses an evolutionary approach, the faster the company releases the product.
-\item The faster the company releases the product, the larger the degree of \textit{accumulated technical debt}\footnote{The debt here is considered regardless of context-specific debt mitigation factors and tactics.}.
+\item The more a capable \footnote{The team capability is evaluated according 
+to the characteristics presented in CAT4.} team, with few quality constraints, 
+uses an evolutionary approach, the faster the company releases the product.
+\item The faster the company releases the product, the larger the degree of 
+\textit{accumulated technical debt}\footnote{The debt here is considered 
+regardless of context-specific debt mitigation factors and tactics.}.
 \end{compactenum}
 
-To assess these hypotheses we defined three measures reflecting the theoretical model:
+To assess these hypotheses we defined three measures reflecting the theoretical 
+model:
 
@@
 
-\item \textit{Execution speed}: a metric that represents the development speed of the startup during different phases of the first release, computed by means of a weighted average speed for each phase.
-It was obtained by analysing interview transcripts looking at subcategories of \textit{Speed-up development}.
-
-\item \textit{Technical debt}: a metric that represents the extent to which processes are controlled, structured, planned and documented by means of engineering artifacts and practices.
-It has been computed by means of a weighted average of the debt accumulated in each development phase, observing subcategories of  \textit{accumulated technical debt} with consequences on the startups growth.
-
-\item \textit{Potential capability}: a metric that represents the degree to which each company reflected the capability of reaction and flexibility to the dynamic environment during the development process, given by the three categories that mostly affect \textit{speed-up development}.
+\item \textit{Execution speed}: a metric that represents the development speed 
+of the startup during different phases of the first release, computed by means 
+of a weighted average speed for each phase. It was obtained by analysing 
+interview transcripts looking at subcategories of \textit{Speed-up development}.
+
+\item \textit{Technical debt}: a metric that represents the extent to which 
+processes are controlled, structured, planned and documented by means of 
+engineering artifacts and practices. It has been computed by means of a weighted 
+average of the debt accumulated in each development phase, observing 
+subcategories of  \textit{accumulated technical debt} with consequences on the 
+startups growth.
+
+\item \textit{Potential capability}: a metric that represents the degree to 
+which each company reflected the capability of reaction and flexibility to the 
+dynamic environment during the development process, given by the three 
+categories that mostly affect \textit{speed-up development}.
 \end{compactitem}
 
-We quantified these measures for each company involved in our empirical study by defining score metrics based on a set of rubrics and evaluating startups accordingly, through an analysis of interview transcripts, codes and questionnaire results.
-The complete statistical procedure and provided rubrics to quantify the measures are illustrated and discussed in Appendix.
-%MUN: make sure it is clear that the PROCEDURE material is available in the ONLINE appendix and refer to it. Still, you need to report the actual RESULTS of you statistical analysis here... 
-The validation has been performed on the most critical  causal relationships of the model.
-According to the paradigm model those are the causal condition which are linked to the core-category.
-There are four high-level relationships  which have not been validated using numerical methods.
-However their correctness has been verified with a cross-analysis of the extant literature and by attesting the causal relations among codes and categories, through axial and selective coding processes.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+We quantified these measures for each company involved in our empirical study 
+by defining score metrics based on a set of rubrics and evaluating startups 
+accordingly, through an analysis of interview transcripts, codes and 
+questionnaire results. The complete statistical procedure and provided rubrics 
+to quantify the measures are illustrated and discussed in Appendix.
+%MUN: make sure it is clear that the PROCEDURE material is available in the 
+%ONLINE appendix and refer to it. Still, you need to report the actual RESULTS 
+%of you statistical analysis here... 
+The validation has been performed on the most critical  causal relationships of 
+the model. According to the paradigm model those are the causal condition which 
+are linked to the core-category. There are four high-level relationships  which 
+have not been validated using numerical methods. However their correctness has 
+been verified with a cross-analysis of the extant literature and by attesting 
+the causal relations among codes and categories, through axial and selective 
+coding processes.
+
+%------------------------------------------------------------------------------%
 \section{Discussion} %MUN:Implications of the GSM
@@
 
-%MUN: There are two issues with this section: it is too long and it basically reiterates (in a concise manner) what you have said in the preceding 12 pages.
-The real value of this section (suggested by my proposed name above), would be if you can state in 0.5 (MAX 1 page), what are the implications of your results on THEORY (or state-of-art, research) and PRACTICE (state-of-practice, people who want to embark a startup).
-I'd propose you have a short intro here, then have two subsections: State-of-art, State-of-practice where you state the implications of GSM (or use a better subdivision if you come up with one, or use none if you don't exceed 0.5 pages).
-These implications need to logically follow from what you have said before (and the data you collected in the interviews and, when necessary, by reference to the SMS).
-Don't have a long bullet list, but structure your arguments in a logical flow.
-This is the section in which you have to show, based on your previous rigorous and well done analysis, the essence / gained knowledge, indicating what are the possible next steps for research in this area.
-You can be creative/visionary here as long as your arguments follow from the previous analysis.
-
-In this section we present a list of the most relevant implications which emerge from the behaviour of early-stage startups, formally expressed in the GSM model, answering to our research question (\textit{What is the current state-of-practice related to software development strategy in early-stage startups?}) through its sub-questions.
-Although the startups we inquired in the case study were spread across different nations and market sectors, we have found several patterns that emerged very clearly:
+%MUN: There are two issues with this section: it is too long and it basically 
+%reiterates (in a concise manner) what you have said in the preceding 12 pages. 
+%The real value of this section (suggested by my proposed name above), would be 
+%if you can state in 0.5 (MAX 1 page), what are the implications of your 
+%results on THEORY (or state-of-art, research) and PRACTICE (state-of-practice, 
+%people who want to embark a startup). I'd propose you have a short intro here, 
+%then have two subsections: State-of-art, State-of-practice where you state the 
+%implications of GSM (or use a better subdivision if you come up with one, or 
+%use none if you don't exceed 0.5 pages). These implications need to logically 
+%follow from what you have said before (and the data you collected in the 
+%interviews and, when necessary, by reference to the SMS). Don't have a long 
+%bullet list, but structure your arguments in a logical flow. This is the 
+%section in which you have to show, based on your previous rigorous and well 
+%done analysis, the essence / gained knowledge, indicating what are the possible 
+%next steps for research in this area. You can be creative/visionary here as 
+%long as your arguments follow from the previous analysis.
+
+In this section we present a list of the most relevant implications which 
+emerge from the behaviour of early-stage startups, formally expressed in the GSM 
+model, answering to our research question (\textit{What is the current 
+state-of-practice related to software development strategy in early-stage 
+startups?}) through its sub-questions. Although the startups we inquired in the 
+case study were spread across different nations and market sectors, we have 
+found several patterns that emerged very clearly:
 
@@
 
-\item The most urgent priority of software development is to shorten time-to-market.
-When developing a new product, which has not been attempted before, it is crucial for the company to release the product soon and observe the users reactions.
-The company's chances to survive, are strictly related to the ability to find the right product/market fit as quickly as possible.
-Developing software without releasing it for a long period of time, coming up with very complex solutions often is not a viable option.
-Business demands, investors interests, market pressures and the general uncertain conditions force startups to quickly iterate on the product by exploring and experimenting new functionalities progressively rolled-out to early-adopters.
-
-\item Startups do not apply any standard development methodology.
-When focused on building the first version of the product, startups do not observe any specific and standard development methodologies or processes.
-From an accurate and in-depth investigation into daily activities, it emerged that the number of  the observed development \textit{practices} is extremely small and never systematic (only little \textit{pair programming} sessions and rarely \textit{test driven development}).
-Despite the wide number of publications, which discuss the adoption of lightweight methodologies, we have found that especially in this first stage not only methodologies but even individual practices are basically neglected.
-
-\item The closest development approach undertaken by early-stage startups is the \textit{Lean startup methodology}.
-A highly evolutionary development approach, centered around the quick production of a functioning prototype and guided by customer feedback is a central aspect of the \textit{Lean startup methodology} pioneered by Ries \cite{Ries2011}.
-However, startups do not formally follow the cycle \textit{build-measure-learn} as it is proposed by the methodology, where they should set-up an experiment to \textit{test the riskiest element of the business idea}, nor make explicit use of the good practices suggested in the book.
-For example, when it comes to user feedback, startups do not explicitly follow the step-by-step process of \textit{customer development} formally defined by \cite{Blank2005}.
-Instead, startups absorb and implement the high-level principles that outlines the methodology, reflected in the model by the theoretical category \textit{find the product/market fit quickly} and its
-subcategories.
-
-\item The greatest part of engineering activities of startups are focused on the implementation while only little attention is given to more conventional activities (project management, requirement specifications, analysis, architecture design, automatic testing, \ldots).
-Especially when bringing the first product to the market, the attitude of startups towards traditional engineering activities is strongly characterized by a general feeling of unnecessity.
-Every document, artifact or process, which does not produce a short-term tangible benefit to the software production is perceived as a waste of time.
-The most visible downside of this kind of approach is a growing \textit{accumulated technical debt} (CAT6), which is partially mitigated by different elements supporting the development.
-Project management is substituted by the use of informal short-term deadlines supported by online collaborative tools with a kanban-board-like fashion; feature list are managed through ticket based systems supporting low-
-precision lists of requirement specifications in the form of user stories; after the essential set of functionality have been defined, making estimations and prioritization of features is not of much help; the traditional analysis process is, to a great extent, neglected (analysis is informally conducted for specific critical functionalities, usually by studying similar software solutions and rarely documenting decision-making processes); the design of the architecture is usually replaced by the use of well-known frameworks and naive diagrams on the whiteboard; advanced and modern version control systems work well for the configuration management; high-volumes of informal communication and the use of coding standards reduce the need of formal documentation; in the majority of startups writing automated test cases is perceived as a waste too (since the product reliability is not a primary concern,  engineers identify major bugs by testing the product internally and letting early adopters identify secondary
-bugs); finally, the infrastructural complexity of deployment is typically left to third party cloud applications.
-Thus, by radically ignoring activities related to documentation, structures and processes, startups achieve a very high initial development speed in the first stage, but on the other hand they accumulate a technical debt that needs to be at least partially fulfilled when the company and the product grow.
-
-\item The first release of the product includes only a limited set of well suitable functionalities focused on user experience.
-The beta version is usually focused on a minimal and essential set of functionalities providing a reasonable user experience (UX).
-As confirmed by practitioners, the user must be able to accomplish a task on the product without major interruptions in the flow, which would cause frustration before being able to capture his attention.
-To realize UX in a short time frame, developers focus on a limited number of suitable functionalities, especially in the first version(s) of the product.
-Other product qualities are regarded as less important: for instance, obtaining a high-reliability is not a priority since early adopters of innovative product are keen to accept minor bugs in exchange of new functionalities\footnote{These results further highlight differences between software development in startups and traditional web engineering, where quality aspects are rated as very important,
-even more than time-to-market \cite{Offutt2002}.}.
-From a business perspective, one might argue that established companies might take advantage of the new idea and develop a better version of the product internally.
-Despite this might represent a risk, established companies usually cannot afford to operate in a market of \textit{innovators} and \textit{early adopters} group of users - described by the Roger's bell curve in \cite{Rogers} - because of the little return of investment they would gain, as discussed by Christensen in \cite{Christensen2003}.
-Instead, established companies have more interests in acquiring a startup and all the knowledge it has gained during the \textit{customer development} process while focusing effort on acquiring larger shares of a ``mass market'' sector.
-Furthermore, thanks to cloud computing and more in general high availability of third party components, aspects such as performance, security and scalability can be achieved with a reasonable effort.
-Thus, given the general lack
-of quality concerns, and supported by the fact that the optimal strategy to improve the UX is by having the product used by real customers, startups can effectively focus on achieving short time-to-market.
-
-\item Engineering activities are supported by low-precision artifacts and collaborative tools integrated with the development environment.
-Startups take advantage of the simplicity of cutting-edge tools, which require a minimum learning and maintenance effort and provide a large payoff.
-Tools are used for a vast number of tasks: ticked-based boards to manage user stories; systems to collect user feedback; tools for tracing product metrics; and scripts for managing deployments.
-The most prominent families of tools, which support software development in startups are modern version control systems.
-They are used by practitioners not only to manage the code base and the repositories, but for many other purposes such as: task assignment; tracing responsibilities; configuration management; issue management; automatic deployment; discussions and code reviews.
-On the other hand, an excessive usage of  \textit{out-of-the-code} tools can negatively affect the development speed, as significantly reported by C11: \
-textit{``We realized we were losing too much time moving tickets around boards than actually working on stories.
-You need to find the balance between the time spent with tools and the amount of actual coding''}.
-This risk is often mitigated with an intense usage of offline manual tools: especially post-it notes and whiteboards are reported to be particularly effective to foster internal communication within the team.
-
-\item The founding team is the most important determinant of speed.
-More than techniques and methodologies, the features of the team are the essential factors to achieve high-development speed.
-Discussed in the theoretical category the \textit{team is the catalyst of development}, this result has been confirmed in multiple occasions both in the interviews and in the questionnaires.
-What is truly important for the startup, among other factors, is the extremely high-colocation which enables continuous informal and colloquial coordination and discussions.
-The initial (small) development team usually spends most of the hours of the week working closely to implement the product, often extending the conventional office hours.
-The large amount of communication, spontaneously held in informal places, has been proven to be an extremely effective element of several successful software projects \cite{Highsmith2000}.
-Hence, the vastly \textit{informal} development environment of startups acts as a catalyst of speed.
-Other important elements, which contribute in achieving such an initial high-speed, are: minimal set of functionalities with low concerns on product quality; evolutionary prototyping approach; smart use of advanced tools and frameworks; low precision artifacts; and the close colocation and capability of the team.
-
-\item The initial lack of structures and processes negatively affects the performance when the company starts to grow.
-One of the primary objectives of startups is to expand their business into fast growing markets.
-In fact, if successful, the company will face a growth in terms of number of customers, employees and product functionalities.
-Then, the increased complexity arises the necessity of controlling the initially chaotic software development environment.
-Under delicate circumstances, where an increased market demand meets a poorly engineered product and process,  the development team need to start returning the accumulated technical debt.
-As reported by practitioners, introducing more structures, documentation and standard procedures, cause an initial drop-down in performance which severity depends on the speed of growth, the amount of technical debt and the capacity of the team to absorb it.
-
-\item Startups bring the first product to market in a very short-time and practitioners are satisfied with the software development strategy.
-The companies inquired in the case study were typically able to release the first beta of the product in a short time frame.
-Although practitioners acknowledge the initial negative consequences of the technical debt on subsequent performances, they confirm that it is better spending some effort in later stage to recover the debt than losing time in over-engineering the production initially with the risk of never achieving any significant growth.
+\item The most urgent priority of software development is to shorten 
+time-to-market. When developing a new product, which has not been attempted 
+before, it is crucial for the company to release the product soon and observe 
+the users reactions. The company's chances to survive, are strictly related to 
+the ability to find the right product/market fit as quickly as possible. 
+Developing software without releasing it for a long period of time, coming up 
+with very complex solutions often is not a viable option. Business demands, 
+investors interests, market pressures and the general uncertain conditions force 
+startups to quickly iterate on the product by exploring and experimenting new 
+functionalities progressively rolled-out to early-adopters.
+
+\item Startups do not apply any standard development methodology. When focused 
+on building the first version of the product, startups do not observe any 
+specific and standard development methodologies or processes.  From an accurate 
+and in-depth investigation into daily activities, it emerged that the number of  
+the observed development \textit{practices} is extremely small and never 
+systematic (only little \textit{pair programming} sessions and rarely 
+\textit{test driven development}). Despite the wide number of publications, 
+which discuss the adoption of lightweight methodologies, we have found that 
+especially in this first stage not only methodologies but even individual 
+practices are basically neglected.
+
+\item The closest development approach undertaken by early-stage startups is 
+the \textit{Lean startup methodology}.  A highly evolutionary development 
+approach, centered around the quick production of a functioning prototype and 
+guided by customer feedback is a central aspect of the \textit{Lean startup 
+methodology} pioneered by Ries \cite{Ries2011}. However, startups do not 
+formally follow the cycle \textit{build-measure-learn} as it is proposed by the 
+methodology, where they should set-up an experiment to \textit{test the riskiest 
+element of the business idea}, nor make explicit use of the good practices 
+suggested in the book. For example, when it comes to user feedback, startups do 
+not explicitly follow the step-by-step process of \textit{customer development} 
+formally defined by \cite{Blank2005}. Instead, startups absorb and implement 
+the high-level principles that outlines the methodology, reflected in the model 
+by the theoretical category \textit{find the product/market fit quickly} and 
+its subcategories.
+
+\item The greatest part of engineering activities of startups are focused on 
+the implementation while only little attention is given to more conventional 
+activities (project management, requirement specifications, analysis, 
+architecture design, automatic testing, \ldots). Especially when bringing the 
+first product to the market, the attitude of startups towards traditional 
+engineering activities is strongly characterized by a general feeling of 
+unnecessity. Every document, artifact or process, which does not produce a 
+short-term tangible benefit to the software production is perceived as a waste 
+of time. The most visible downside of this kind of approach is a growing 
+\textit{accumulated technical debt} (CAT6), which is partially mitigated by 
+different elements supporting the development. Project management is substituted 
+by the use of informal short-term deadlines supported by online collaborative 
+tools with a kanban-board-like fashion; feature list are managed through ticket 
+based systems supporting low-precision lists of requirement specifications in 
+the form of user stories; after the essential set of functionality have been 
+defined, making estimations and prioritization of features is not of much help; 
+the traditional analysis process is, to a great extent, neglected (analysis is 
+informally conducted for specific critical functionalities, usually by studying 
+similar software solutions and rarely documenting decision-making processes); 
+the design of the architecture is usually replaced by the use of well-known 
+frameworks and naive diagrams on the whiteboard; advanced and modern version 
+control systems work well for the configuration management; high-volumes of 
+informal communication and the use of coding standards reduce the need of formal 
+documentation; in the majority of startups writing automated test cases is 
+perceived as a waste too (since the product reliability is not a primary 
+concern,  engineers identify major bugs by testing the product internally and 
+letting early adopters identify secondary bugs); finally, the infrastructural 
+complexity of deployment is typically left to third party cloud applications.  
+Thus, by radically ignoring activities related to documentation, structures and 
+processes, startups achieve a very high initial development speed in the first 
+stage, but on the other hand they accumulate a technical debt that needs to be 
+at least partially fulfilled when the company and the product grow.
+
+\item The first release of the product includes only a limited set of well 
+suitable functionalities focused on user experience. The beta version is usually 
+focused on a minimal and essential set of functionalities providing a reasonable 
+user experience (UX). As confirmed by practitioners, the user must be able to 
+accomplish a task on the product without major interruptions in the flow, which 
+would cause frustration before being able to capture his attention. To realize 
+UX in a short time frame, developers focus on a limited number of suitable 
+functionalities, especially in the first version(s) of the product. Other 
+product qualities are regarded as less important: for instance, obtaining a 
+high-reliability is not a priority since early adopters of innovative product 
+are keen to accept minor bugs in exchange of new functionalities\footnote{These 
+results further highlight differences between software development in startups 
+and traditional web engineering, where quality aspects are rated as very 
+important, even more than time-to-market \cite{Offutt2002}.}. From a business 
+perspective, one might argue that established companies might take advantage of 
+the new idea and develop a better version of the product internally. Despite 
+this might represent a risk, established companies usually cannot afford to 
+operate in a market of \textit{innovators} and \textit{early adopters} group of 
+users - described by the Roger's bell curve in \cite{Rogers} - because of the 
+little return of investment they would gain, as discussed by Christensen in 
+\cite{Christensen2003}. Instead, established companies have more interests in 
+acquiring a startup and all the knowledge it has gained during the 
+\textit{customer development} process while focusing effort on acquiring larger 
+shares of a ``mass market'' sector. Furthermore, thanks to cloud computing and 
+more in general high availability of third party components, aspects such as 
+performance, security and scalability can be achieved with a reasonable effort. 
+Thus, given the general lack of quality concerns, and supported by the fact that 
+the optimal strategy to improve the UX is by having the product used by real 
+customers, startups can effectively focus on achieving short time-to-market.
+
+\item Engineering activities are supported by low-precision artifacts and 
+collaborative tools integrated with the development environment. Startups take 
+advantage of the simplicity of cutting-edge tools, which require a minimum 
+learning and maintenance effort and provide a large payoff.  Tools are used for 
+a vast number of tasks: ticked-based boards to manage user stories; systems to 
+collect user feedback; tools for tracing product metrics; and scripts for 
+managing deployments. The most prominent families of tools, which support 
+software development in startups are modern version control systems. They are 
+used by practitioners not only to manage the code base and the repositories, but 
+for many other purposes such as: task assignment; tracing responsibilities; 
+configuration management; issue management; automatic deployment; discussions 
+and code reviews.  On the other hand, an excessive usage of  
+\textit{out-of-the-code} tools can negatively affect the development speed, as 
+significantly reported by C11: \textit{``We realized we were losing too much 
+time moving tickets around boards than actually working on stories. You need to 
+find the balance between the time spent with tools and the amount of actual 
+coding''}. This risk is often mitigated with an intense usage of offline manual 
+tools: especially post-it notes and whiteboards are reported to be particularly 
+effective to foster internal communication within the team.
+
+\item The founding team is the most important determinant of speed. More than 
+techniques and methodologies, the features of the team are the essential factors 
+to achieve high-development speed. Discussed in the theoretical category the 
+\textit{team is the catalyst of development}, this result has been confirmed in 
+multiple occasions both in the interviews and in the questionnaires. What is 
+truly important for the startup, among other factors, is the extremely 
+high-colocation which enables continuous informal and colloquial coordination 
+and discussions. The initial (small) development team usually spends most of the 
+hours of the week working closely to implement the product, often extending the 
+conventional office hours.  The large amount of communication, spontaneously 
+held in informal places, has been proven to be an extremely effective element of 
+several successful software projects \cite{Highsmith2000}. Hence, the vastly 
+\textit{informal} development environment of startups acts as a catalyst of 
+speed. Other important elements, which contribute in achieving such an initial 
+high-speed, are: minimal set of functionalities with low concerns on product 
+quality; evolutionary prototyping approach; smart use of advanced tools and 
+frameworks; low precision artifacts; and the close colocation and capability of 
+the team.
+
+\item The initial lack of structures and processes negatively affects the 
+performance when the company starts to grow. One of the primary objectives of 
+startups is to expand their business into fast growing markets. In fact, if 
+successful, the company will face a growth in terms of number of customers, 
+employees and product functionalities. Then, the increased complexity arises the 
+necessity of controlling the initially chaotic software development environment. 
+Under delicate circumstances, where an increased market demand meets a poorly 
+engineered product and process,  the development team need to start returning 
+the accumulated technical debt. As reported by practitioners, introducing more 
+structures, documentation and standard procedures, cause an initial drop-down in 
+performance which severity depends on the speed of growth, the amount of 
+technical debt and the capacity of the team to absorb it.
+
+\item Startups bring the first product to market in a very short-time and 
+practitioners are satisfied with the software development strategy. The 
+companies inquired in the case study were typically able to release the first 
+beta of the product in a short time frame. Although practitioners acknowledge 
+the initial negative consequences of the technical debt on subsequent 
+performances, they confirm that it is better spending some effort in later stage 
+to recover the debt than losing time in over-engineering the production 
+initially with the risk of never achieving any significant growth. 
 \end{compactitem}
 
-All of the above observations are extracted from the analysis of the model, which is systematically derived from empirical data of the case study.
-Indeed, with slightly different levels of adherence, the implication we presented are reflected in the behaviour of the great majority of companies in the sample.
-The results of this analysis, which has been focused on the software development strategy from the idea conception to the first release, indicate that startups are far from adopting standard or ad-hoc development methodologies, especially in this phase.
-The typical tendency is to focus on the team capability to implement and quickly iterate on a prototype, which is released very fast.
-Thus, in a context where even for the most \textit{lightweight} agile methodologies is hard to penetrate, we believe that it is premature to discuss issues related to SPI.
-Researchers should rather focus on the trade-off between development speed and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which
-appears to be the most important determinant for the future of software development in the company.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+All of the above observations are extracted from the analysis of the model, 
+which is systematically derived from empirical data of the case study. Indeed, 
+with slightly different levels of adherence, the implication we presented are 
+reflected in the behaviour of the great majority of companies in the sample. The 
+results of this analysis, which has been focused on the software development 
+strategy from the idea conception to the first release, indicate that startups 
+are far from adopting standard or ad-hoc development methodologies, especially 
+in this phase. The typical tendency is to focus on the team capability to 
+implement and quickly iterate on a prototype, which is released very fast. Thus, 
+in a context where even for the most \textit{lightweight} agile methodologies is 
+hard to penetrate, we believe that it is premature to discuss issues related to 
+SPI. Researchers should rather focus on the trade-off between development speed 
+and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which 
+appears to be the most important determinant for the future of 
+software development in the company.
+
+%------------------------------------------------------------------------------%
 \section{Threats to Validity}
 \label{valt}
-%MUN: Also here, introduce some structure: first, explain shortly what types of validity threats exist for the research methodology you applied.
-Then, discuss each type individually.
-Think about which threat type is the most prominent one for your research methodology and start with that one.
-I did not review this section in detail now since it will change considerably after your restructuring.
+%MUN: Also here, introduce some structure: first, explain shortly what types of 
+%validity threats exist for the research methodology you applied. Then, discuss 
+%each type individually. Think about which threat type is the most prominent 
+%one for your research methodology and start with that one. I did not review 
+%this section in detail now since it will change considerably after your 
+%restructuring.
 
 %External
-The first threat related to external validity might be caused by a possible wrong selection of subjects interviewed for the study.
-This threat profoundly affects GT, since it is a qualitative research method, which makes the use of semi-structured interviews, in fact, centres the research on respondent's opinions.
-In order to mitigate this threat we selected interviewees which covered positions of CTOs' and CEOs'.
-Their perspectives on what it is taking place within the organization, were the only meaningful data taken in consideration in the study.
-In many cases there was no supporting evidence to verify the opinion expressed.
-Questionnaire dedicated to ``data triangulation'' was not of primary support because of the little time practitioners were able to grant to it.
-However, researchers have to accept the veracity of what respondents reported during the interviews \cite{Coleman2008a}.
-Notwithstanding the issues surrounding semi-structured interviews are vital.
-Within our study, although the reality might
-be different from what has been described, the reported data are CTOs' and CEOs' perception, which is used to base their decisions.
-
-Finally the comparison with similar frameworks also helped in establishing the domain to which the study's findings can be generalized.
-In particular the previous framework developed by Coleman in \cite{Coleman2008} has allowed researchers to find similarities, differences and broader reasoning related to factors that hinder mature processes to be established in startups.
-Moreover a comparison has been conducted also with the notorious Brook's framework describing the \textit{basic attacks} to the challenges of developing software systems and other extant literature retrieved by \cite{SMS}\footnote{Note that the generalizability of results is applied only to those startups that operate in software-intensive activities.}.
-In conclusion, regarding the validation of the subcategories, a continuous re-evaluation of the interview transcripts has been performed.
+The first threat related to external validity might be caused by a possible 
+wrong selection of subjects interviewed for the study. This threat profoundly 
+affects GT, since it is a qualitative research method, which makes the use of 
+semi-structured interviews, in fact, centres the research on respondent's 
+opinions. In order to mitigate this threat we selected interviewees which 
+covered positions of CTOs' and CEOs'. Their perspectives on what it is taking 
+place within the organization, were the only meaningful data taken in 
+consideration in the study.  In many cases there was no supporting evidence to 
+verify the opinion expressed. Questionnaire dedicated to ``data triangulation'' 
+was not of primary support because of the little time practitioners were able to 
+grant to it. However, researchers have to accept the veracity of what 
+respondents reported during the interviews \cite{Coleman2008a}. Notwithstanding 
+the issues surrounding semi-structured interviews are vital. Within our study, 
+although the reality might be different from what has been described, the 
+reported data are CTOs' and CEOs' perception, which is used to base their 
+decisions.
+
+Finally the comparison with similar frameworks also helped in establishing the 
+domain to which the study's findings can be generalized. In particular the 
+previous framework developed by Coleman in \cite{Coleman2008} has allowed 
+researchers to find similarities, differences and broader reasoning related to 
+factors that hinder mature processes to be established in startups. Moreover a 
+comparison has been conducted also with the notorious Brook's framework 
+describing the \textit{basic attacks} to the challenges of developing software 
+systems and other extant literature retrieved by \cite{SMS}\footnote{Note that 
+the generalizability of results is applied only to those startups that operate 
+in software-intensive activities.}. In conclusion, regarding the validation of 
+the subcategories, a continuous re-evaluation of the interview transcripts has 
+been performed.
 
 %Internal
-To enhance the internal validity, we created a three-dimensional research framework to perform a triangulation validity procedure.
-By means of grounded theory approach formed by a \textit{systematic mapping study}, \textit{interviews} and  \textit{follow-up questionnaires} we searched for convergence among different sources of information to confirm our theories.
-
-By improving and validating the theory definition, we conducted a final comparison of the emergent theory with: extant literature, previously developed models and evaluation of empirical data.
-With the final model validation we highlighted and examined similarities, contrasts and explanations as discussed in \cite{Eisenhardt2007}.
-In this regard Eisenhardt stated: ``Tying the emergent theory to existing literature enhances the internal validity, generalizability, and theoretical level of the theory building from case study research [\ldots] because the findings often rest on a very limited number of cases.''
-
-Little ambiguities about direction of causal influences have been reduced by introducing arrows to indicate the direction of the causal relation in the GSM model.
-Furthermore, by means of comparison between the GSM model and extant studies in the broader fields, we revealed that the theory resembles in wider startup development context.
-Nevertheless, we discovered important confounding factors which might mine the conclusion and generalizability of the study related to innovation, creativity and market requirements.
+To enhance the internal validity, we created a three-dimensional research 
+framework to perform a triangulation validity procedure. By means of grounded 
+theory approach formed by a \textit{systematic mapping study}, 
+\textit{interviews} and  \textit{follow-up questionnaires} we searched for 
+convergence among different sources of information to confirm our theories.
+
+By improving and validating the theory definition, we conducted a final 
+comparison of the emergent theory with: extant literature, previously developed 
+models and evaluation of empirical data. With the final model validation we 
+highlighted and examined similarities, contrasts and explanations as discussed 
+in \cite{Eisenhardt2007}.  In this regard Eisenhardt stated: ``Tying the 
+emergent theory to existing literature enhances the internal validity, 
+generalizability, and theoretical level of the theory building from case study 
+research [\ldots] because the findings often rest on a very limited number of 
+cases.''
+
+Little ambiguities about direction of causal influences have been reduced by 
+introducing arrows to indicate the direction of the causal relation in the GSM 
+model. Furthermore, by means of comparison between the GSM model and extant 
+studies in the broader fields, we revealed that the theory resembles in wider 
+startup development context. Nevertheless, we discovered important confounding 
+factors which might mine the conclusion and generalizability of the study 
+related to innovation, creativity and market requirements.
 
 %Construct
-The first important threat to the construct validity of the study is a possible inadequate description of constructs.
-In order to dismiss this risk, the entire study constructs have been adapted to the terminologies utilized by practitioners and defined at an adequate level for each theoretical conceptualization\footnote{For instance defining \textit{Time shortage}, the researchers explained it in terms of \textit{Investor pressure, CEO/business pressure, Demo presentations at events and internal final deadline} used by most of the interviewees during the study.}.
-Moreover, during the coding of interview transcripts the researchers adopted wide and explanatory descriptive labels for theoretical categories, to capture the underlying phenomenon without losing relevant details.
-
-The second important threat is caused by the fact that interviewees might already be aware of the possible emergent theories analyzed by researchers.
-To reduce the risk that some hypothesis could have been involved in the design of the research, we let mainly participants drive the direction of the case study.
-
-Another threat associated with the construct validity is the evaluation apprehension, which deals with people's rejection attitude towards the assessment of their behaviour.
-To overcome this risk we didn't attained to first answers from practitioners, but rather went towards details of engineering elements and activities.
-Moreover, we developed a rigorous data collection protocol to create case study databases and employ integration of multiple data collection methods.
-Reporting bias was mitigated by packaging all the needed material for conducting new researches in other contexts in order to criticize our findings (interview package with instructions has been made available online\footnote{Available at \url{https://github.com/adv0r/BTH-Interview-Package}.}.
-Moreover, two academic supervisors with expertise in the area have conducted a peer-review analysis of framework's constructs.
-
-To control distortion during analysis we made extensive use of memos.
-To demonstrate that we were aware of personal biases, memos provided us another important function in controlling the quality of data analysis.
-Through the use of memos and comparative analysis we were able to check if data fitted into the emerging theory and also countering subjectivity that ultimately enhanced the likelihood of producing accurate research findings.
-
-Grounded theory has been already applied by other researches in similar contexts to attest relationships among conceptualizations of an examined phenomenon (see \cite{Coleman2007, Basri, Coleman2008a}).
-Those relationships between sampling and analysis should be verified in such a way that emerging findings remain consistent as further data are collected.
-To attest this, we  explored and tested:
+The first important threat to the construct validity of the study is a 
+possible inadequate description of constructs. In order to dismiss this risk, 
+the entire study constructs have been adapted to the terminologies utilized by 
+practitioners and defined at an adequate level for each theoretical 
+conceptualization\footnote{For instance defining \textit{Time shortage}, the 
+researchers explained it in terms of \textit{Investor pressure, CEO/business 
+pressure, Demo presentations at events and internal final deadline} used by most 
+of the interviewees during the study.}. Moreover, during the coding of interview 
+transcripts the researchers adopted wide and explanatory descriptive labels for 
+theoretical categories, to capture the underlying phenomenon without losing 
+relevant details.
+
+The second important threat is caused by the fact that interviewees might 
+already be aware of the possible emergent theories analyzed by researchers. To 
+reduce the risk that some hypothesis could have been involved in the design of 
+the research, we let mainly participants drive the direction of the case study.
+
+Another threat associated with the construct validity is the evaluation 
+apprehension, which deals with people's rejection attitude towards the 
+assessment of their behaviour. To overcome this risk we didn't attained to first 
+answers from practitioners, but rather went towards details of engineering 
+elements and activities. Moreover, we developed a rigorous data collection 
+protocol to create case study databases and employ integration of multiple data 
+collection methods. Reporting bias was mitigated by packaging all the needed 
+material for conducting new researches in other contexts in order to criticize 
+our findings (interview package with instructions has been made available 
+online\footnote{Available at 
+\url{https://github.com/adv0r/BTH-Interview-Package}.}. Moreover, two academic 
+supervisors with expertise in the area have conducted a peer-review analysis of 
+framework's constructs.
+
+To control distortion during analysis we made extensive use of memos. To 
+demonstrate that we were aware of personal biases, memos provided us another 
+important function in controlling the quality of data analysis. Through the use 
+of memos and comparative analysis we were able to check if data fitted into the 
+emerging theory and also countering subjectivity that ultimately enhanced the 
+likelihood of producing accurate research findings.
+
+Grounded theory has been already applied by other researches in similar 
+contexts to attest relationships among conceptualizations of an examined 
+phenomenon (see \cite{Coleman2007, Basri, Coleman2008a}). Those relationships 
+between sampling and analysis should be verified in such a way that emerging 
+findings remain consistent as further data are collected. To attest this, we  
+explored and tested:
 
@@
 
-In particular we were prepared to modify generated categories so that the new data could be adapted into the emerging theory according to the concepts of: \textit{theoretical sampling} and \textit{theoretical saturation}.
-
-According to the \textit{theoretical sampling} concept, we adjusted our research design and the emergent theory until only marginal results were generated.
-Moreover to enhance reliability of the outcome conceptualizations and relations, the process of coding interviews has been systematically conducted following a detailed process.
-Then, we asked to other practitioners a confirmation to further validate the generated framework and enhance the reliability of the relations among the identified categories.
-
-An important issue is related to the fact that the limited number of interviews might not represent the complete scenarios in our context of study.
-Nevertheless, this issue has been partially mitigated as result of the theoretical saturation concept.
-This experience is attested by Martin's statement: \textit{``By the time three or four sets of data have been analysed, the majority of useful concepts will have been discovered.''} \cite{giardinoEtAl}.
-To obtain major certainty of the obtained results, what we applied was the final validation.
-
-Despite we couldn't apply any power analysis procedures, in grounded theory the sample size depended on when data saturation occurred.
-Ramer in \cite{Ramer}, comparing quantitative to qualitative researches, states: \textit{``reaching data saturation, which involves obtaining data until no new information emerges, is critical for obtaining applicability in qualitative research''}.
-To avoid interrupting the research prematurely, after attesting that no more relevant information could be gathered from executing additional interviews, we iterated grounded theory study one more time, verifying that the explanatory power of the core category was fulfilled.
-
-The  approach of generating theory allowed us to check emerging categories and their properties by gathering new evidence within an iterative and evolutionary approach, in view of the wide context of the research area that was under investigation.
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+In particular we were prepared to modify generated categories so that the new 
+data could be adapted into the emerging theory according to the concepts of: 
+\textit{theoretical sampling} and \textit{theoretical saturation}.
+
+According to the \textit{theoretical sampling} concept, we adjusted our 
+research design and the emergent theory until only marginal results were 
+generated. Moreover to enhance reliability of the outcome conceptualizations and 
+relations, the process of coding interviews has been systematically conducted 
+following a detailed process. Then, we asked to other practitioners a 
+confirmation to further validate the generated framework and enhance the 
+reliability of the relations among the identified categories.
+
+An important issue is related to the fact that the limited number of interviews 
+might not represent the complete scenarios in our context of study. 
+Nevertheless, this issue has been partially mitigated as result of the 
+theoretical saturation concept. This experience is attested by Martin's 
+statement: \textit{``By the time three or four sets of data have been analysed, 
+the majority of useful concepts will have been discovered.''} 
+\cite{giardinoEtAl}. To obtain major certainty of the obtained results, what we 
+applied was the final validation.
+
+Despite we couldn't apply any power analysis procedures, in grounded theory the 
+sample size depended on when data saturation occurred. Ramer in \cite{Ramer}, 
+comparing quantitative to qualitative researches, states: \textit{``reaching 
+data saturation, which involves obtaining data until no new information emerges, 
+is critical for obtaining applicability in qualitative research''}. To avoid 
+interrupting the research prematurely, after attesting that no more relevant 
+information could be gathered from executing additional interviews, we iterated 
+grounded theory study one more time, verifying that the explanatory power of the 
+core category was fulfilled.
+
+The  approach of generating theory allowed us to check emerging categories 
+and their properties by gathering new evidence within an iterative and 
+evolutionary approach, in view of the wide context of the research area that was 
+under investigation.
+%------------------------------------------------------------------------------%
 \subsection{Future work}
 \label{futwork}
-%MUN: Don't have future work as a subsection in threats to validity. Either merge it in "Implications of the GSM" OR into "Conclusions".
-
-The GSM model enables practitioners in startup companies to speed-up development empowering the role of the team, focusing on minimal functionalities and adopting evolutionary development approaches.
-Researchers can use the GSM model to reproduce the study with other software startups, and focus on mitigation strategies to minimize the impact of the accumulated technical debt over the need of shortening time-to-market.
-Moreover, the  mutual collaborative dialog between the research in the development strategies in software startups and technical debt can benefits in understanding how both influence the growth of a company after the release of the first product to the market.
-
-%--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------%
+%MUN: Don't have future work as a subsection in threats to validity. Either 
+%merge it in "Implications of the GSM" OR into "Conclusions".
+
+The GSM model enables practitioners in startup companies to speed-up 
+development empowering the role of the team, focusing on minimal functionalities 
+and adopting evolutionary development approaches. Researchers can use the GSM 
+model to reproduce the study with other software startups, and focus on 
+mitigation strategies to minimize the impact of the accumulated technical debt 
+over the need of shortening time-to-market. Moreover, the  mutual collaborative 
+dialog between the research in the development strategies in software startups 
+and technical debt can benefits in understanding how both influence the growth 
+of a company after the release of the first product to the market.
+
+%------------------------------------------------------------------------------%
 %\subsection{Subsection Heading Here}
@@
 \label{conc}
-Startups are able to produce cutting-edge software products with a wide impact on the market, significantly contributing to the global economy.
-Software development, especially in the early-stage, is at the core of the company's daily activities.
-Despite their severely high failure-rate, we have found that the quick proliferation of startups is not supported by an adequate and scientific body of knowledge.%MUN: <-- last sentence: isn't that a conclusion from the SMS?
- To be able to intervene on the software development strategy with scientific and engineering approaches, the first step is to understand startups' behavior.
-Hence, in this research work, we provided an initial explanation of the underlying phenomenon by means of a grounded theory case study, focusing on early engineering activities, from the idea conception to the first open beta release of the product.
-
-Through an exploratory research, conducted by a systematic approach, we produced a theoretical model grounded into the hindsight knowledge collected among practitioners with the aim of explaining how development strategies are engineered startups.
-The explanatory capability and correctness of the model have been validated by means of systematic comparisons with the state-of-the-art results and empirical data.
-The systematic mapping of the literature \cite{SMS}, with 37 studies extracted from an initial sample of 943 items, revealed a multi-faceted state-of-the-art inadequate to support software development activities in startup companies.
-%MUN: <-- you mix again SMS and GT here...
-fix it
-On the other hand, the case study conducted in 13 early-stage startups, provided a broad set of empirical evidences obtained by combining different research methodologies in a grounded theory approach.
-
-The overall results of our research confirm that startups possess unique %MUN: why unique? 
-characteristics of uncertainty, lack of resources and time-pressure.
-These factors influence the software development to an extent that transforms every decision related to the development strategies into a difficult trade-off for the company.
-Moreover, although startups share different characteristics with other similar SE domains (e.g.
-market-driven development, small companies and web engineering), the unique combination of coexisting factors poses a whole new set of challenges which need to be addressed by primary studies.
-Especially when bringing the first product to market, startups' most urgent priority is releasing the product as quickly as possible to verify the product/market fit and to adjust the business and product trajectory according to early feedback and collected metrics.
-In this stage, startups often discard any formal project management, documentation, analysis, planning, testing and other traditional process
-activities. Practitioners take advantage of an evolutionary prototyping approach, using well-integrated tools and externalizing complexity to third party solutions.
-
-However, the initial gain obtained in terms of flexibility and speed is counterbalanced by the need of restructuring the product and controlling the engineering activities when the company starts to grow.
-If successful, the startup will face a growth in terms of number of customers, employees and product functionalities that leads to the necessity to control the initially chaotic software development environment.
-In this context the most significant challenge for early-stage startups is finding a sweet spot between being fast enough to enter the market early while controlling the amount of accumulated technical debt.
-As confirmed by analyzing other frameworks in SE, the use of prescriptive and rigid structures over the development activities would hinder performance since startups operate across the border between the chaotic and complex domain %MUN: Not sure whether you have discussed/mentioned the meaning of chaotic/complex domain in the paper, it think not.
-Hence you need to rephrase that (but you can not introduce new concepts here in the conclusion).
-Therefore, from a software development perspective, working with low-precision engineering activities represents the only viable strategy that startups can follow in the early stage of their lifecycle.
-
-In this research we discussed a number of novel challenges for both practitioners and researchers, while presenting a first set of concepts, terms and activities which set the software engineering scene for the rapidly increasing startup phenomenon.
-By making a comparison with Berry's definition of SE \cite{Berry1992}, we would like to see the rise of a new discipline - \textit{startup engineering} - which can be defined as  \textit{the use of scientific, engineering, managerial and systematic approaches with the aim of successfully developing software systems in startup companies}.
-%MUN: Nice ;-)
+Startups are able to produce cutting-edge software products with a wide impact 
+on the market, significantly contributing to the global economy. Software 
+development, especially in the early-stage, is at the core of the company's 
+daily activities. Despite their severely high failure-rate, we have found that 
+the quick proliferation of startups is not supported by an adequate and 
+scientific body of knowledge. %MUN: <-- last sentence: isn't that a conclusion 
+%from the SMS?
+To be able to intervene on the software development strategy with scientific and 
+engineering approaches, the first step is to understand startups' behavior. 
+Hence, in this research work, we provided an initial explanation of the 
+underlying phenomenon by means of a grounded theory case study, focusing on 
+early engineering activities, from the idea conception to the first open beta 
+release of the product.
+
+Through an exploratory research, conducted by a systematic approach, we 
+produced a theoretical model grounded into the hindsight knowledge collected 
+among practitioners with the aim of explaining how development strategies are 
+engineered startups. The explanatory capability and correctness of the model 
+have been validated by means of systematic comparisons with the state-of-the-art 
+results and empirical data. The systematic mapping of the literature \cite{SMS}, 
+with 37 studies extracted from an initial sample of 943 items, revealed a 
+multi-faceted state-of-the-art inadequate to support software development 
+activities in startup companies. %MUN: <-- you mix again SMS and GT here... fix 
+%it
+On the other hand, the case study conducted in 13 early-stage startups, 
+provided a broad set of empirical evidences obtained by combining different 
+research methodologies in a grounded theory approach.
+
+The overall results of our research confirm that startups possess unique %MUN: 
+%why unique? 
+characteristics of uncertainty, lack of resources and time-pressure. These 
+factors influence the software development to an extent that transforms every 
+decision related to the development strategies into a difficult trade-off for 
+the company. Moreover, although startups share different characteristics with 
+other similar SE domains (e.g. market-driven development, small companies and 
+web engineering), the unique combination of coexisting factors poses a whole new 
+set of challenges which need to be addressed by primary studies. Especially when 
+bringing the first product to market, startups' most urgent priority is 
+releasing the product as quickly as possible to verify the product/market fit 
+and to adjust the business and product trajectory according to early feedback 
+and collected metrics. In this stage, startups often discard any formal project 
+management, documentation, analysis, planning, testing and other traditional 
+process activities. Practitioners take advantage of an evolutionary prototyping 
+approach, using well-integrated tools and externalizing complexity to third 
+party solutions.
+
+However, the initial gain obtained in terms of flexibility and speed is 
+counterbalanced by the need of restructuring the product and controlling the 
+engineering activities when the company starts to grow. If successful, the 
+startup will face a growth in terms of number of customers, employees and 
+product functionalities that leads to the necessity to control the initially 
+chaotic software development environment. In this context the most significant 
+challenge for early-stage startups is finding a sweet spot between being fast 
+enough to enter the market early while controlling the amount of accumulated 
+technical debt.  As confirmed by analyzing other frameworks in SE, the use of 
+prescriptive and rigid structures over the development activities would hinder 
+performance since startups operate across the border between the chaotic and 
+complex domain. %MUN: Not sure whether you have discussed/mentioned the meaning 
+%of chaotic/complex domain in the paper, I think not. Hence you need to 
+%rephrase that (but you can not introduce new concepts here in the conclusion). 
+Therefore, from a software development perspective, working with low-precision 
+engineering activities represents the only viable strategy that startups can 
+follow in the early stage of their lifecycle.
+
+In this research we discussed a number of novel challenges for both 
+practitioners and researchers, while presenting a first set of concepts, terms 
+and activities which set the software engineering scene for the rapidly 
+increasing startup phenomenon. By making a comparison with Berry's definition of 
+SE \cite{Berry1992}, we would like to see the rise of a new discipline - 
+\textit{startup engineering} - which can be defined as  \textit{the use of 
+scientific, engineering, managerial and systematic approaches with the aim of 
+successfully developing software systems in startup companies}. %MUN: Nice ;-)
 
@@
 \bibliography{tse}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-%\begin{thebibliography}{1}
-
-%\bibitem{IEEEhowto:kopka}
-%This is an example of a book reference
-%H. Kopka and P.W. Daly, \emph{A Guide to {\LaTeX}}, third ed. Harlow, U.K.: Addison-Wesley, 1999.
-
-%This is an example of a Transactions article reference
-%D.S.
-Coming and O.G.
-Staadt, "Velocity-Aligned Discrete Oriented Polytopes for Dynamic Collision Detection," IEEE Trans.
-Visualization and Computer Graphics, vol. 14,  no. 1,  pp.
-1-12,  Jan/Feb  2008, doi:10.1109/TVCG.2007.70405.
-
-%This is an example of a article from a conference proceeding
-%H.
-Goto, Y.
-Hasegawa, and M.
-Tanaka, "Efficient Scheduling Focusing on the Duality of MPL Representation," Proc.
-IEEE Symp.
-Computational Intelligence in Scheduling (SCIS '07), pp.
-57-64, Apr.
-2007, doi:10.1109/SCIS.2007.367670.
-
-%This is an example of a PrePrint reference
-%J.M.P.
-Martinez, R.B.
-Llavori, M.J.A.
-Cabo, and T.B.
-Pedersen, "Integrating Data Warehouses with Web Data: A Survey," IEEE Trans.
-Knowledge and Data Eng., preprint, 21 Dec.
-2007, doi:10.1109/TKDE.2007.190746.
-
-%Again, see the IEEEtrans_HOWTO.pdf for several more bibliographical examples. Also, more style examples
-%can be seen at http://www.computer.org/author/style/transref.htm
-%\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-%\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{pictures/cg}}]{Michael Shell}
-
-%\end{IEEEbiography}
 

### 2013-02-24 article/main.tex
fixed comments version 0.3

@@
 
+%CG: Not fixed. There is a huge gap on the topic, the refs in the introduction 
+%are those retrieved by the SMS.
+
 %\hfill mds
@@
 1994 in an article written by Carmel \cite{Camel1994a} where he studied the 
-\textit{time-to-completion in a young package firm}. %MUN: If this is a quote 
+time-to-completion in a young package firm. %MUN: If this is a quote 
 %from Carmel, put it in quotation marks, like you did in the previous section 
@@
 %quotations throughout the paper. 
+%CG: fixed, btw  it was just to emphasize his the focus of his study. 
 He noticed how these companies were particularly innovative and successful, 
@@
 Additionally, the role of SPI has been neglected because seen as an obstacle to 
-the development team's creativity and flexibility \cite{Coleman2008a}, and to 
-the need of a quick delivering product process environment. %MUN: the second
-%part of the sentence is just a statement without reference. If it refers to a 
+the development team's creativity and flexibility, and to 
+the need of a quick delivering product process environment \cite{Coleman2008a}.%MUN: the second part of the sentence is just a statement without reference. If it refers to a 
 %statement by Coleman, put his reference at the end OR find the correct 
 %reference OR delete the statement.
+%CG fixed
 Product quality is left aside in favor of minimal and suitable functionalities 
@@
 functionality, studies suggest to externalize the complexity of parts of the 
-project to third party solutions by means of outsourcing activities, software 
-reuse and open-source strategies \cite{Hanna2010, Jansen2008, Wall2001, 
-Bean2005}. %MUN: Can you split up the references? I.e. write:
+project to third party solutions by means of outsourcing activities 
+\cite{Hanna2010}, software reuse \cite{Jansen2008} and open-source strategies , \cite{Wall2001, Bean2005}. %MUN: Can you split up the references? I.e. write:
 %"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source 
 %strategies \cite{Bla}.
-
+%CG: fixed
 %------------------------------------------------------------------------------%
@@
 \centering
-\includegraphics[width=4in]{figures/completemethodology2}
+\includegraphics[width=5in]{figures/completemethodology2}
 \caption{Grounded theory study process overview}
@@
 %to keep the height of the figure, expanding however the width (use horizontal
-%space since it is a two column figure). Change also in figure:
-% - Write "Data Analysis" horizontally
-% - Change names of macro-phases to upper-case, i.e. Design and Execution, 
-%   Data Collection, Data Analysis, etc.
-% - Add section number to macro-phase name, i.e. 3.1 Design and Execution, 3.2 
-%   Data Collection
-% - Change "Discussion and Validation" into "Theory Validation" as you name it 
-%   in the text.
-%Also, add a section (3.5) that describes shortly what you did in "Theory 
-%Validation"
+%space since it is a two column figure). Change also in figure:  - Write "Data
+%Analysis" horizontally  - Change names of macro-phases to upper-case, i.e.
+%Design and Execution,     Data Collection, Data Analysis, etc.  - Add section
+%number to macro-phase name, i.e. 3.1 Design and Execution, 3.2     Data
+%Collection  - Change "Discussion and Validation" into "Theory Validation" as
+%you name it     in the text. Also, add a section (3.5) that describes shortly
+%what you did in "Theory  Validation" CG: fixed... fontsize cannot be increased
+%uniformly among all the boxes, then I skipped that part...
 
@@
 All companies, except C10, were founded within the last three years (2009-2012), 
-%MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct?
+%MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct? CG: right
 by an average of 3 starting members, which were in majority developers. 
@@
 quickly obtaining responses, questionnaires have been sent to respondents 
-immediately after transcribing the interview results. Two weeks after the 
-conclusion of the last interview, we closed our follow-up questionnaires and 
-collected the data. %MUN: I don't understand this: you
+immediately after transcribing the interview results. Most of the follow-up questionnaires have been collected within one day after the 
+conclusion of the last interview.   %MUN: I don't understand this: you
 %collected the data from questionnaires after ALL interviews have been 
@@
 %(data analysis and triangulation let themes emerge which in turn were used
-%in subsequent interviews).
+%in subsequent interviews). CG: fixed...
 
@@
 
+%------------------------------------------------------------------------------%
 \subsection{Theory Validation}
+\label{rm:val}
 %MUN: put here what I wrote in the comment in Section "Theory validation" (i.e.
 %the section following the GSM)
+%CG: fixed
+
+Presenting GT is challenging for a researcher, who must pay attention to structure the level of details included, and to the way data is portrayed to display evidence of the emergent categories.
+In order to assess our research and determine if the GT is sufficiently grounded, we used a systematic technique to assess the theory, following evaluation lists and factors presented in  \cite{Strauss1998}.
+
+First we had to consider the following factors:
+
+\par\vspace {0.4cm}
+\begin{compactitem}
+\item \textit{Fit}, that describes the consistence of the theory with the real data.
+\item \textit{Understanding}, that describes the degree to which participants agree with the theory.
+\item \textit{Generality}, that describes the level of abstraction of the theory, and how sufficiently it can be adopted by practitioners without losing its relevance.
+\item \textit{Control}, that describes how well the theory enables practitioners to fully understand the described situation.
+\end{compactitem}
+\par\vspace {0.4cm}
+
+In order to achieve a high-level of fit between the theory and the raw empirical data, we started GT without any assumption on how the emergent patterns would influence the research direction.
+Ideas and perceptions evolved during the mapping study and interviews.
+In addition, constant comparative methods, overturning of some early categories as new data came to light, generation and testing of interim hypotheses, and constant re-evaluation of the interview transcripts, ensured that researchers' bias was minimized and theoretical fit maintained.
+
+To represent understandable emergent theories, concepts and categories were carefully developed using appropriate and self-explanatory codes.
+The developing theory was not immediately presented to the participants to prevent potential bias in the responses.
+But rather, the theoretical framework has been presented and described to startups practitioners not involved in the data collection, whose reaction was positively expressed.
+
+From generality perspective, we developed a theory from abstract categories, which allowed us to create a general guide to constantly changing situations.
+Nevertheless, we took attention on maintaining a reasonable level of abstraction, keeping focus on their sensitizing characteristics and providing low-level subgroup of categories.
+
+Finally, we provided a comprehensive set of categories with detailed interrelations to explain how the \textit{core-category} is achieved and the reasons of its consequences, in order to enable practitioners to independently understand and analyze situations, predict changes and their consequences, and to be capable of revising actions or the theory itself.
+Using both methodological tools and features provided by \textit{AtlasTI}, we explored and tested:
+
+\par\vspace {0.4cm}
+\begin{compactitem}
+\item Each category and the strength of relations between them.
+\item Hypotheses, derived from and related to the emergent theory.
+\item Deviant cases to ensure robustness and general applicability.
+\end{compactitem}
+\par\vspace {0.4cm}
+
+Additionally, Strauss and Corbin provided a list of questions to assist in determining how well the findings are grounded \cite{Strauss1998}. They questions are as follows:
+
+\par\vspace {0.4cm}
+\begin{compactenum}
+\item Are concepts generated, and are the concepts systematically related?
+\item Are there many conceptual linkages and are the categories well developed?
+\item Is variation\footnote{Variation refers to the variety of contexts which theory can be applied.} built into the theory and are the conditions under which variation can be found built into the study and explained?
+\item Are the conditions under which variation can be found built into the study and explained?
+\item Has the process been taken into account, and does the theory stand the test of time?
+\item Do the theoretical findings seem significant, and to what extent?
+\end{compactenum}
+\par\vspace {0.4cm}
+
+Answering the above questions, the concepts were generated according to the coding process.
+They were systematically related through the use of a network diagram (see figure \ref{fig:gt:completemethodology}), by which we established the linkages and relations between the concepts.
+At each iteration of the grounded theory process, we considered and examined a concept within different conditions and dimensions, trying to incorporate opinions of a broader as possible range of practitioners.
+
+All the linkages and categories were constructed by the use of Atlas.TI and compared according to the data analysis process.
+Variations of the generated theory were applied according to raw grounded data with the aid of codes and memos.
+Moreover during this study extensive explanation of the concepts were presented in-vivo statements as reported by practitioners.
+
+The process has been designed by the researches in different steps, explaining and making clear the purpose of each of them.
+Thus, the same processes enable other research to conduct the same study within similar contexts.
+Moreover \textit{comparison with the state-of-the-art} has been applied in order to validate the theory and enhance odds of its applicability within a wider time-frame.
+By this comparison we  highlight the areas which have been neglected by existing studies, providing possible directions for future studies (see subsection \ref{sect:theory:validation:others}).
+Moreover we study the confounding factors which could interfere with the application of the GSM model (see subsection \ref{sect:an:comp:cat-vs-literature:conf}).
+
+Then, we performed an \textit{evaluation with empirical data} (see subsection \ref{sect:theory:validation:stats}), quantifying scores for each company involved in our empirical study by defining metrics based on a set of rubrics and evaluating startups accordingly, through an analysis of interview transcripts and codes.
+The complete statistical procedure and provided  rubrics  to quantify the measures are illustrated in Appendix.
 
@@
 reach the core category (efficiency, effectiveness and performance) is discussed 
-in detail in section \ref{valt}. %MUN: this refers to the validity threats 
+in detail in section \ref{res:val}. %MUN: this refers to the validity threats 
 %section. Do you mean rather the "Theory Validation" section?
+%CG:fixed
 
@@
 urgent needs\footnote{There are some exceptions where the quality aspects 
-actually matter and such cases will be discussed in section \ref{valt}%MUN: 
+actually matter and such cases will be discussed in section \ref{res:val}%MUN: 
 %this refers to the validity threats section. Do you mean rather the "Theory 
 %Validation" section?
+%CG fixed
 .}. At the same time, to be able to deal with such constraints, startups must 
@@
 %this statement since it is a citation? 
+%CG: not fixed. It represents a subcategory, expressed by the 70% of the interviewed startups
 mainly constrained by a tight \textit{time shortage}.
@@
 achieved. Product metrics are typically web-based \textit{statistical hypothesis 
-testing}, such as A/B testing. %MUN: add reference or explain shortly what A/B 
+testing}, such as A/B testing \cite{AB}. %MUN: add reference or explain shortly what A/B 
 %testing is
+%CG: fixed
 
@@
 \footnote{Detailed discussion of the impact of innovative products on the user 
-satisfaction is presented in Appendix according to the Kano model.}.
+satisfaction is presented in subsection \ref{sect:an:comp:cat-vs-literature:conf}}.
 %MUN: do you use the Kano model in Theory Validation? Change text to refer to 
@@
 %reference the correct section, but remove "according to the Kano model]
+%CG: fixed
 \end{compactitem}
@@
 technologies can be autonomously managed by the development team (with 
-characteristics of: co-location, self-organization and very small size). %MUN: 
+characteristics of: co-location, self-organization and very small team size). %MUN: 
 %do you mean small team size?
+%CG: fixed
 
@@
 
-As discussed in section \ref{valt} %MUN: this refers to the validity threats 
+As discussed in section \ref{res:val} %MUN: this refers to the validity threats 
 %section. Do you mean rather the "Theory Validation" section?
+%CG:fixed
 , we formed the theory by applying the paradigm model, introduced by Corbin 
@@
 In this section we discuss the validity of the GSM by means of 
-cross-methodological observations:
+cross-methodological observations, as discussed in section \ref{rm:val}.
 %MUN: expand this discussion a bit (motivation why you do this like you do it) 
@@
 %validation as discussed in Section 3.5 (and remove the bullet list below).
-
-\begin{compactitem}
-\item Comparison between the GSM and similar  models and frameworks.
-\item Comparison between the GSM and literature. We highlight 
-the areas which have been neglected by existing studies, providing possible 
-directions for future studies.
-\item Discussion of confounding factors which could interfere with the GSM.
-\item Validation of the correctness of the GT procedure by attesting relations 
-between the GSM categories, evaluating if they are compliant with the empirical 
-data we collected.
-
-\end{compactitem}
+%CG:fixed
 
@@
 To validate the generalization of the theory, we describe conceptualizations 
-defined from the GSM that are in support of a previous model developed by 
-Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}. We refer to Coleman's work 
-since only he has, to the best of our knowledge, conducted a similar research in 
-startups' context, even though with a different focus. Coleman investigated 
+defined from the GSM that are in support of previous models developed by 
+Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}, Baskerville \cite{Internet} and finally Brooks \cite{BrooksJr1987}.
+
+Starting from the Coleman's work, we refer to him since only he has, to the best of our knowledge, 
+conducted a similar research in startups' context, even though with a 
+different focus. Coleman investigated 
 factors in software development that hinder initiatives of software process 
-improvement (SPI) in much more mature companies. %MUN: so he did not do research 
+improvement (SPI) in a later stage, representing also companies in expansion phase with more than 100 employees. %MUN: so he did not do research 
 %in a startup context? You need to clarify what you mean by "more mature 
@@
 %i.e. after first product release.
+%CG: fixed
 
@@
 letter of them and they are always adapted, so we adapted ours to the way we 
-wanted it to work for us, for our own size and scale''.} %MUN: add ref to 
+wanted it to work for us, for our own size and scale''} \cite{Coleman2008}. %MUN: add ref to 
 %Coleman's paper where he states this
+%CG fixed
 
@@
 penetrating small software organizations \cite{Staples2007}. When startups began 
-establishing any SPI process, they experienced \textit{process erosions} %MUN: 
+establishing any SPI process, they experienced \textit{process erosions} \cite{Coleman2008} %MUN: 
 %add ref (Coleman?). You should add, for each category you use from Coleman, 
@@
 %from him (and is not one of our categories)
+%CG: fixed
 , which resolved to a barely sufficient workflow to satisfy the organizational 
@@
 processes need to be predictable and repeatable. Nimble and ad-hoc solutions 
-prevent the use of heavy bureaucracy and fast communication strategies %MUN: 
+prevent the use of heavy bureaucracy and formal communication strategies %MUN: 
 %seems to contradict each other: "ad-hoc solutions prevent the use of heavy 
@@
 %strategies... I may be wrong, but then you need to explain it better
+%CG:fixed, you are right :) ...  
 , even though the accumulated tacit knowledge is hard to manage and to transfer 
@@
 found that despite Coleman's study contributes to define which factors hinder
-the introduction of SPI %MUN: add: "of SPI in startups"?
-, his results are coherent with the theory generated within our research %MUN: 
-%add "our research, focused at explaining software development in general in 
-%startups." ?
-. In this regard, a wide variability of processes is a key factor for startups. 
-In contrast with some SPI processes, such as \textit{Six Sigma} which objective 
-is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, 
-startups seem to follow the \textit{Fischer's Fundamental Theorem of natural 
-selection} \cite{Fisher}. Moving towards flexible and variable processes (easy 
-to adapt to the uncertain conditions), increase the odds of ``natural 
-selection'' only when some ``restrictive conditions'' are met. In other words, 
-even though ignoring SPI practices is advisable %MUN: under which 
-%circumstances? This statement sounds too general, i.e. don't do SPI ever...
-%And the following part,I did not fully understand. Try to clarify.
-not following SPI does not predict major adaptiveness of startups development 
-to the market. But, to imply a major evolutionary product/market fit, certain 
+the introduction of SPI in startups %MUN: add: "of SPI in startups"? CG: fixed ,
+his results are coherent with the theory generated within our research, focusing
+at explaining the overall strategies within software development in
+startups%MUN:  %add "our research, focused at explaining software development in
+general in  %startups." ? CG:fixed . In this regard, a wide variability of
+processes is a key factor for startups.  In contrast with some SPI processes,
+such as \textit{Six Sigma} which objective  is minimizing unpredictability in
+the definition of workflows \cite{Sixsigma},  startups seem to follow the
+\textit{Fischer's Fundamental Theorem of natural  selection} \cite{Fisher}.
+Moving towards flexible and variable processes (easy  to adapt to the uncertain
+conditions), increase the odds of ``natural  selection'' only when some
+``restrictive conditions'' are met. In other words,  even though ignoring SPI
+practices is advisable, under circumstances of high uncertainty and severe lack
+of resources, %MUN: under which  %circumstances? This statement sounds too
+general, i.e. don't do SPI ever... %CG: fixed %And the following part,I did not
+fully understand. Try to clarify. CG:fixed  not following SPI does not predict
+major adaptiveness of startups development  towards the customer's needs.
+Nevertheless, to imply a major evolutionary product/market fit, certain
 conditions explained in GSM need to be considered.
@@
 %to and why, then create subsections for each.
+
+%CG: partially fixed. Agreed on the beginning of the section, not agreed on the division in subsections. Let's discuss it by a skype call.
+
 Also Baskerville \cite{Internet} refers to SPI as a development approach
 typically effective only in large-scale, long-term development efforts with
-stable and disciplined processes. By applying a grounded theory analysis in 10 
-companies he found that major causal factors that influence development are:
+stable and disciplined processes.Internet-speed software development (i.e. an
+agile method oriented using daily builds aimed at developing a product with high
+speed) differs from traditional software development. Baskerville, by applying a
+grounded theory analysis in 10 companies, found that major causal factors that
+influence development are:
 
@@
 %software development - MUN: put this sentence where you introduce 
-%Baskerville's work
+%Baskerville's work. %CG: fixed
 , Baskerville revealed similar causal factors as the GSM.
@@
 Coleman (2007) & X & X & X & X & X & X & X & 7 & \cite{Coleman2007} \\
-Camel (1994) & X & X & X & X & X & X & X & 7 & \cite{Camel1994a} \\
+Carmel (1994) & X & X & X & X & X & X & X & 7 & \cite{Camel1994a} \\
 Yoffie (1999) & X & X & X & X & X & X & & 6 & \cite{Yoffie1999} \\
@@
 Chorev (2006) &  &  &  & X &  &  &  & 1 & \cite{Chorev2006} \\
+\hline
+Count & 29 & 22 & 13 & 26 & 18 & 14 & 20 & &\\
 %MUN: add a row which shows the sum of each column. This shows how much 
-%categories are covered overall
+%categories are covered overall %CG: fixed
 \hline
@@
 is only partially covering the phenomenon described by the GSM. Looking at the 
-results %MUN: which results? The papers' main contribution or...?
+papers' main contribution %MUN: which results? The papers' main contribution or...? %CG: fixed
 , we were able to extrapolate information which confirmed the GSM categories.
@@
 need of empowering people, which is the element with the greatest impact on 
-software development. Humans, being non-linear variables %MUN: add ref
+software development. Humans, being non-linear variables \cite{Human} %MUN: add ref %CG:fixed
 , are unlikely to follow repeatable prescriptive methodologies. As stated by 
@@
 %identifier you use in table 2. I did not check if you discuss each and every 
-%category. Do you? If yes, ok. If not, why not?
+%category. Do you? If yes, ok. If not, why not? CG: partially fixed, the categories CAT3 and CAT6 are not well covered as discussed above.
 Starting from the category \textit{severe lack of resources} (CAT7), many 
@@
 are discussed in an experience paper by Ambler \cite{Ambler2002}. He reports on 
-two case studies of growing internet startups approaching an IPO. %MUN:explain 
-%acronym
+two case studies of growing Internet startups approaching an initial public offering (IPO). %MUN:explain 
+%acronym %CG: fixed
 Despite the two companies were facing business and team growth, attitudes 
@@
 evident. The need to re-engineer the codebase lead the researchers to suggest a 
-tailored version of the RUP %MUN: add reference
+tailored version of the Rational Unified Process (RUP) %MUN: add reference
+ %CG: not fixed, not sure if needed since Ambler has the detailed 
+ %explaination of his process, similar to RUP... anyway a dedicated resource would be an IBM website
 development framework. Nevertheless, even with the aid of CASE 
@@
 \url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, 
-startups failed in practicing engineering activities in view of eccesive lack of 
-flexibility. Developers fast moved to simpler tools such as using a whiteboard, 
+startups failed in practicing engineering activities in view of excessive lack of 
+flexibility. Developers fast moved to simpler tools such as using a white-board, 
 where they were able to provide input and insights, question assumptions, share 
@@
 %what was needed and go straightforward to coding, avoiding analysis 
-%paralysis."
+%paralysis." %CG: ok
 
@@
 %MUN: the whole paragraph needs to be summarized to its essence. Details can be 
-%read in the referred paper
+%read in the referred paper. CG: partially fixed, removed the detailed process description
 In the evaluation phase, the status of the development workflow within the 
@@
 time. Therefore, they applied a qualitative methodology to define a model of the 
-release management process. They moved towards the plan phase where the 
-researchers decided to develop a model to manage releases, introducing a process 
-composed by: release scope preparation, release planning, release development, 
-system testing, acceptance testing and release deployment. Despite the effort 
+release management process. 
+%They moved towards the plan phase where the 
+%researchers decided to develop a model to manage releases, introducing a process 
+%composed by: release scope preparation, release planning, release development, 
+%system testing, acceptance testing and release deployment. 
+Despite the effort 
 spent by the researchers to establish a development workflow, practices were not 
@@
 
-Prescribed development models such as SPI in startups are basically ignored by 
-startups, as discussed in \cite{Zettel2001}. %MUN: <-- rephrase, not clear to me
+Prescribed development models, oriented to SPI, are basically ignored by 
+startups, as discussed in \cite{Zettel2001}. %MUN: <-- rephrase, not clear to me %CG:fixed
 To overcome the need of structure and to accelerate the development process, 
@@
 
-Deakins \cite{Deakins2005} designed a new spiral methodology, evolved during 
+Deakins \cite{Deakins2005} designed a new spiral methodology, constructed during a
 fieldwork with a developer team that had limited resources managing an 
-innovative product in volatile e-commerce environments. He proposes fast 
-evolutionary approach experimentation with flexible processes obtaining 
+innovative product in volatile e-commerce environments. This methodology represents a fast 
+evolutionary approach experimentation with flexible processes. It aims to obtain
 customer's feedback as soon as possible for continuous product improvement.%MUN: 
-%<-- rephrase, sentence not parsable
-Even in this case the model%MUN: which model?
-is compliant with the nature of a random, ad-hoc and visionary product 
-development. In fact, evolutionary development approaches are typically more 
-suitable  for web applications, as confirmed in \cite{Deshpande2001}. Then, the 
-practices are oriented towards rapid and iterative design with complete freedom 
-to the developer team, enabled to operate anywhere within the organization, 
-according also to the category \textit{team is the catalyst of 
-development}.%MUN: <-- the whole paragraph is not that clear to me. Try to start 
+%<-- rephrase, sentence not parsable CG:fixed
+Even in this case the spiral methodology%MUN: which model? CG:fixed
+is compliant with the nature of a random, ad-hoc and \textit{evolutionary development approach} (CAT 2). In fact, evolutionary development approaches are typically more 
+suitable  for web applications, as confirmed in \cite{Deshpande2001}.
+Moreover, they provide practices oriented towards rapid and iterative design with complete freedom 
+to the developer team, enabled to operate anywhere within the organization ( 
+compliant also to the category.%MUN: <-- the whole paragraph is not that clear to me. Try to start 
 %with the category (team is catalyst), then write about Deakins spiral model
+%CG: partially fixed, the focus here is on the evolutionary development approach
 
@@
 
-%MUN: To which category does the following paragraph belong?
-Eventually to focus on simple solutions %MUN: solutions on what?
+%MUN: To which category does the following paragraph belong? CG: they represent subcategories of speed-up development 
+Eventually to externalize complexity of the developing product %MUN: solutions on what? CG: fixed
 , two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the 
@@
 %explanatory (what does "compliant with the behavioural model of the 
-%phenomenon" mean?)
+%phenomenon" mean?).
+CG: not fixed, when I wrote it I tried to have an overview of the main contributions and make it fluently by papers description.
+Creating subsections would mean to divide papers from different focuses.
+This is a trade-off where we took a decision in the past.
+Take it in consideration, but we can discuss it during the next meeting.
 
 Thus, by summarizing the comparison, we have found that most of the 
-contributions of the studies are compliant with the behavioural model of the 
+contributions of the studies are compliant with the behavioral model of the 
 phenomenon we provided.
@@
 \cite{ColinRobson2009}. We report those factors that have been explicitly 
-identified by the literature review %MUN: do you mean the SMS? Then refer to it. 
+identified by the SMS \cite{SMS} %MUN: do you mean the SMS? Then refer to it. CG: fixed 
 %I remember you mentioned CF in the results (GSM), maybe also refer to the 
-%respective sections
+%respective sections CG: not parsed.
 , related to: creativity and innovation, market requirements and application 
@@
 %arises the need of project management activities and software processes for 
-%long-term scalability issues."
-
-%MUN: Please reformulate the first sentence, I cannot parse it
+%long-term scalability issues."  CG: ok.
+
+%MUN: Please reformulate the first sentence, I cannot parse it. CG: fixed
 Based on the model presented in \cite{Heitlager2007}, it identifies the 
 lifecycle stages of a startup (\textit{fluid}, \textit{transition} and 
-\textit{specific}), corresponding for each stage a development approach 
-(product-oriented and process-oriented) the level of innovation achieved. The 
+\textit{specific}). At each stage, the model presents different development approaches 
+(product-oriented and process-oriented), according to the level of innovation achieved. The 
 study reports how product-oriented development, in contrast with forced and 
@@
 innovation only seldom emerged. However, if innovative and creative factors were 
-considered in the model, the integration of results would have been 
-straightforward, since the evolution of this model is totally compliant with our 
-assumption of CAT5 and CAT6. %MUN: not clear to me what you mean by "evolution 
+considered in this model, the integration of results would have been 
+straightforward, since it is totally compliant with the description of how the \textit{accumulated technical debt} CAT5 causes the need of establishing a process in the long run (see how the \textit{initial growth hinders performance} CAT6).
+%MUN: not clear to me what you mean by "evolution
 %of this model" and what "compliant with our assumption" means. Also, don't use 
-%only CATx in this context, but cite the category name AND CATx
-From product-oriented development in the \textit{fluid} phase to the 
-process-oriented development in the \textit{specific} phase, the innovation 
+%only CATx in this context, but cite the category name AND CATx. CG: fixed
+From the \textit{fluid} phase to the \textit{specific} phase, the innovation 
 confounding factor seems to behave initially the same as the \textit{speed-up 
-development}.%MUN: <-- please provide some motivation... not obvious
+development}, both highly focused towards a product-oriented development. %MUN: <-- please provide some motivation... not obvious. CG: fixed
 Nevertheless, we were unable to ground these evidences in the GT data even 
-though they might present strict correlation %MUN: "present strict correlation" 
-%What does that mean, I don't understand
-with the choice of an early product-development orientation with low-precision 
+though they are two factors that might be positively correlated %MUN: "present strict correlation" 
+%What does that mean, I don't understand. CG: fixed.
+to the choice of an early product-development orientation and the need of low-precision 
 engineering elements.
@@
 For example, safety-critical systems possess quality concerns that must be 
-considered. But also minor risky application domains such as e-commerce systems 
-have high-impact factors for obtaining the first customer reference, %MUN: <-- 
+considered. Even in e-commerce systems 
+factors related to quality have high-impact effects on obtaining the first customer reference, %MUN: <-- 
 %cannot parse it. Remove the first part and write: "Even e-commerce 
-%systems...". Then rephrase the second part of the sentence to make it clearer.
-as discussed in \cite{Deakins2005,Kim2005,Silva2005}. This %MUN: what is "this"?
-characterizes those applications, whose part of functionalities are already 
+%systems...". Then rephrase the second part of the sentence to make it clearer. CG: fixed
+as discussed in \cite{Deakins2005,Kim2005,Silva2005}. Quality attributes especially %MUN: what is "this"? CG: fixed
+characterize those applications, whose part of functionalities are already 
 successfully implemented in other software systems or whose failure might 
@@
 %extrapolate insights from those situations where quality concerns cannot be 
-%postponed or externalized.
+%postponed or externalized. CG: ok
 
 \subsubsection{Developer experience} 
-%MUN: Ups, a confounding factor you did not mention initially. Add it...
+%MUN: Ups, a confounding factor you did not mention initially. Add it... CG: fixed
 Finally, also the experience of team-members is considered as an important 
@@
 %of more than the 70\% of the analyzed transcripts.}), this confounding factor 
-%requires further investigation.
+%requires further investigation. CG: ok
 
@@
 %methods you used on what data, then refer to that section from here (no need 
-%to repeat again here).
+%to repeat again here). CG: fixed
 In this subsection we validate the most important causal relations presented in 
@@
 
-%MUN: for the paragraphs below, you need to explain what and why you do this. 
+%MUN: for the paragraphs below, you need to explain what and why you do this. CG: fixed, explained in methodology 
 %What do you mean by "we grounded in the interview transcripts the fact that 
 %those relations were characterized in terms of performances since 
-%time-to-market was the main concern." Not clear to me.
+%time-to-market was the main concern." Not clear to me. CG: fixed, performance for us is efficiency AND effectiveness
 Before starting the evaluation of the impact of the three categories on 
 \textit{speed-up development}, we grounded in the interview transcripts the fact 
-that those relations were characterized in terms of performances since 
+that those relations were characterized in terms of "effectiveness" and "efficiency" since 
 time-to-market was the main concern (as confirmed in \cite{SMS}):
@@
 
-Accordingly, by analysing the data of the 13 companies, we observed two 
-distinct trends which further empower the above illustrated relations:
+Accordingly, by analyzing the data of the 13 companies, we formulated two 
+hypotheses which further describe the above illustrated relations:
 %MUN: You need to be more precise with terminology. Above you state that the 
@@
 %thesis to this limited space, but I think it is doable (and relevant for 
-%supporting validity of GSM)
+%supporting validity of GSM). CG: fixed
 
@@
 %ONLINE appendix and refer to it. Still, you need to report the actual RESULTS 
-%of you statistical analysis here... 
+%of you statistical analysis here... CG: partially fixed. we always referred to appendix without references until now.
+
+We conducted statistical tests using the analysis of variance to assess the existence of relations between the dimensions.
+First we defined two null hypotheses (H0): $H0_1$ = \textit{Startups do not release the product faster when a capable team adopt a more evolutionary approach AND with less quality constraints}; $H0_2$ = \textit{The execution speed does not increase the amount of accumulated technical debt}.
+Then we tested $H0_1$ and $H0_2$  with an one-tailed test using Pearson's product moment correlation coefficient\footnote{To perform the Pearson's correlation we verified two assumptions: a) data is normally distributed; b) data is on interval scale.}, with positive association analysis, fixing the level of confidence to 95\% which means we reject H0 in case the p-value is lower than 0.05.
+
+We conclude that, in our sample:
+\par\vspace {0.4cm}
+\begin{compactenum}
+\item Higher values of \textit{Execution speed} are strongly associated\footnote{Clear statistical significance, p-value: 0.002073.} with \textit{higher} values for  \textit{Technical debt}.
+
+\item Higher values of  \textit{Execution speed}  are strongly associated\footnote{Clear statistical significance, p-value: 0.004549.} with  \textit{higher values} for  \textit{Potential capability}.
+\end{compactenum}
+\par\vspace {0.4cm}
+
 The validation has been performed on the most critical  causal relationships of 
@@
 %long as your arguments follow from the previous analysis.
-
-In this section we present a list of the most relevant implications which 
-emerge from the behaviour of early-stage startups, formally expressed in the GSM 
+%CG: not fixed, I have shortened the content of one page.
+But still we cannot create the subsections you proposed since the implications are all derived from the model.
+Then, we can still shorten it if you want.
+Let's discuss it during the next skype call.
+
+In this section we present relevant implications which 
+emerge from the behavior of early-stage startups, formally expressed in the GSM 
 model, answering to our research question (\textit{What is the current 
@@
 
-\begin{compactitem}
-
-\item The most urgent priority of software development is to shorten 
+%------------------------------------------------------------------------------%
+\subsection{Shortening time-to-market}
+
+The most urgent priority of software development is to shorten 
 time-to-market. When developing a new product, which has not been attempted 
@@
 
-\item Startups do not apply any standard development methodology. When focused 
+%------------------------------------------------------------------------------%
+\subsection{Light-weight methodology}
+Startups do not apply any standard development methodology. When focused 
 on building the first version of the product, startups do not observe any 
@@
 
-\item The closest development approach undertaken by early-stage startups is 
+The closest development approach undertaken by early-stage startups is 
 the \textit{Lean startup methodology}.  A highly evolutionary development 
@@
 the high-level principles that outlines the methodology, reflected in the model 
-by the theoretical category \textit{find the product/market fit quickly} and 
-its subcategories.
-
-\item The greatest part of engineering activities of startups are focused on 
+by the theoretical category \textit{find the product/market fit quickly}.
+
+The greatest part of engineering activities of startups are focused on 
 the implementation while only little attention is given to more conventional 
@@
 engineering activities is strongly characterized by a general feeling of 
-unnecessity. Every document, artifact or process, which does not produce a 
+unnecessary. Every document, artifact or process, which does not produce a 
 short-term tangible benefit to the software production is perceived as a waste 
-of time. The most visible downside of this kind of approach is a growing 
-\textit{accumulated technical debt} (CAT6), which is partially mitigated by 
-different elements supporting the development. Project management is substituted 
-by the use of informal short-term deadlines supported by online collaborative 
-tools with a kanban-board-like fashion; feature list are managed through ticket 
-based systems supporting low-precision lists of requirement specifications in 
-the form of user stories; after the essential set of functionality have been 
-defined, making estimations and prioritization of features is not of much help; 
-the traditional analysis process is, to a great extent, neglected (analysis is 
-informally conducted for specific critical functionalities, usually by studying 
-similar software solutions and rarely documenting decision-making processes); 
-the design of the architecture is usually replaced by the use of well-known 
-frameworks and naive diagrams on the whiteboard; advanced and modern version 
-control systems work well for the configuration management; high-volumes of 
-informal communication and the use of coding standards reduce the need of formal 
-documentation; in the majority of startups writing automated test cases is 
-perceived as a waste too (since the product reliability is not a primary 
-concern,  engineers identify major bugs by testing the product internally and 
-letting early adopters identify secondary bugs); finally, the infrastructural 
-complexity of deployment is typically left to third party cloud applications.  
-Thus, by radically ignoring activities related to documentation, structures and 
-processes, startups achieve a very high initial development speed in the first 
-stage, but on the other hand they accumulate a technical debt that needs to be 
-at least partially fulfilled when the company and the product grow.
-
-\item The first release of the product includes only a limited set of well 
+of time.
+
+%------------------------------------------------------------------------------%
+\subsection{Focus on minimal set of functionalities}
+The first release of the product includes only a limited set of well 
 suitable functionalities focused on user experience. The beta version is usually 
@@
 
-\item Engineering activities are supported by low-precision artifacts and 
-collaborative tools integrated with the development environment. Startups take 
-advantage of the simplicity of cutting-edge tools, which require a minimum 
-learning and maintenance effort and provide a large payoff.  Tools are used for 
-a vast number of tasks: ticked-based boards to manage user stories; systems to 
-collect user feedback; tools for tracing product metrics; and scripts for 
-managing deployments. The most prominent families of tools, which support 
-software development in startups are modern version control systems. They are 
-used by practitioners not only to manage the code base and the repositories, but 
-for many other purposes such as: task assignment; tracing responsibilities; 
-configuration management; issue management; automatic deployment; discussions 
-and code reviews.  On the other hand, an excessive usage of  
-\textit{out-of-the-code} tools can negatively affect the development speed, as 
-significantly reported by C11: \textit{``We realized we were losing too much 
-time moving tickets around boards than actually working on stories. You need to 
-find the balance between the time spent with tools and the amount of actual 
-coding''}. This risk is often mitigated with an intense usage of offline manual 
-tools: especially post-it notes and whiteboards are reported to be particularly 
-effective to foster internal communication within the team.
-
-\item The founding team is the most important determinant of speed. More than 
+%------------------------------------------------------------------------------%
+\subsection{Empowering the team members}
+The founding team is the most important determinant of speed. More than 
 techniques and methodologies, the features of the team are the essential factors 
@@
 
-\item The initial lack of structures and processes negatively affects the 
+%------------------------------------------------------------------------------%
+\subsection{Pay back of the technical debt}
+The initial lack of structures and processes negatively affects the 
 performance when the company starts to grow. One of the primary objectives of 
@@
 
-\item Startups bring the first product to market in a very short-time and 
-practitioners are satisfied with the software development strategy. The 
-companies inquired in the case study were typically able to release the first 
-beta of the product in a short time frame. Although practitioners acknowledge 
-the initial negative consequences of the technical debt on subsequent 
-performances, they confirm that it is better spending some effort in later stage 
-to recover the debt than losing time in over-engineering the production 
-initially with the risk of never achieving any significant growth. 
-\end{compactitem}
-
 All of the above observations are extracted from the analysis of the model, 
@@
 with slightly different levels of adherence, the implication we presented are 
-reflected in the behaviour of the great majority of companies in the sample. The 
+reflected in the behavior of the great majority of companies in the sample. The 
 results of this analysis, which has been focused on the software development 
@@
 %------------------------------------------------------------------------------%
+\subsection{Future work}
+\label{futwork}
+%MUN: Don't have future work as a subsection in threats to validity. Either 
+%merge it in "Implications of the GSM" OR into "Conclusions". CG: fixed
+
+The GSM model enables practitioners in startup companies to speed-up 
+development empowering the role of the team, focusing on minimal functionalities 
+and adopting evolutionary development approaches. Researchers can use the GSM 
+model to reproduce the study with other software startups, and focus on 
+mitigation strategies to minimize the impact of the accumulated technical debt 
+over the need of shortening time-to-market. Moreover, the  mutual collaborative 
+dialog between the research in the development strategies in software startups 
+and technical debt can benefits in understanding how both influence the growth 
+of a company after the release of the first product to the market.
+
+%------------------------------------------------------------------------------%
 \section{Threats to Validity}
@@
 %restructuring.
-
+%CG: fixed
+
+%------------------------------------------------------------------------------%
+In this section we discuss the validity of the overall results.  We structure the validity in four parts, accordingly with Wholin taxonomy \cite{Wohlin2000}:
+
+\begin{compactitem}
+
+\item External validity - which requires establishing the domain to which the study's findings can be generalized (partially discussed above).
+
+\item Internal validity - which is enhanced by establishing causal relationships whereby we attest that certain conditions lead to other conditions, as distinguished from spurious relationships\footnote{A \textit{spurious} relationship is a relationship in which two events have no direct causal connection, but rather due to either a coincidence or the presence of confounding factors.}.
+\item Construct validity - which is enforced by clearly specifing operational procedures, and providing an adequate description of constructs\footnote{Constructs are complex ideas that we as humans form in order to summarize observations about things that we cannot see directly.
+For example a construct is \textit{``job satisfaction''}, which to be defined, requires a higher level of abstraction than a concrete object(e.g.
+a book).} that are used to generate the theory.
+\item Conclusion validity - which represent the degree to which conclusions are reasonable, according to the defined relationships in the data.
+It involves ensuring adequate sampling procedures, appropriate analytical methods and reliable measurement procedures.
+Reliability requires to demonstrate that the operations of a study - such as data collection procedures - can be repeated by other researchers, obtaining the same results.
+\end{compactitem}
+
+%------------------------------------------------------------------------------%
+\subsection{External validity}
 %External
@@
 
+%------------------------------------------------------------------------------%
+\subsection{Internal validity}
 %Internal
@@
 
+%------------------------------------------------------------------------------%
+\subsection{Construct validity}
 %Construct
@@
 
+%------------------------------------------------------------------------------%
+\subsection{Conclusion validity}
+
 Grounded theory has been already applied by other researches in similar 
@@
 under investigation.
-%------------------------------------------------------------------------------%
-\subsection{Future work}
-\label{futwork}
-%MUN: Don't have future work as a subsection in threats to validity. Either 
-%merge it in "Implications of the GSM" OR into "Conclusions".
-
-The GSM model enables practitioners in startup companies to speed-up 
-development empowering the role of the team, focusing on minimal functionalities 
-and adopting evolutionary development approaches. Researchers can use the GSM 
-model to reproduce the study with other software startups, and focus on 
-mitigation strategies to minimize the impact of the accumulated technical debt 
-over the need of shortening time-to-market. Moreover, the  mutual collaborative 
-dialog between the research in the development strategies in software startups 
-and technical debt can benefits in understanding how both influence the growth 
-of a company after the release of the first product to the market.
 
@@
 development, especially in the early-stage, is at the core of the company's 
-daily activities. Despite their severely high failure-rate, we have found that 
+daily activities. Despite their severely high failure-rate, SMS \cite{SMS} have found that 
 the quick proliferation of startups is not supported by an adequate and 
 scientific body of knowledge. %MUN: <-- last sentence: isn't that a conclusion 
-%from the SMS?
+%from the SMS? CG: fixed
 To be able to intervene on the software development strategy with scientific and 
@@
 have been validated by means of systematic comparisons with the state-of-the-art 
-results and empirical data. The systematic mapping of the literature \cite{SMS}, 
-with 37 studies extracted from an initial sample of 943 items, revealed a 
+results and empirical data. The systematic mapping study \cite{SMS}, 
+with 37 studies extracted, revealed a 
 multi-faceted state-of-the-art inadequate to support software development 
 activities in startup companies. %MUN: <-- you mix again SMS and GT here... fix 
-%it
+%it CG: not fixed, I would keep it since important part of the initial GT methodology
 On the other hand, the case study conducted in 13 early-stage startups, 
@@
 
-The overall results of our research confirm that startups possess unique %MUN: 
-%why unique? 
+The overall results of our research confirm that startups possess %MUN: 
+%why unique? CG: fixed.
 characteristics of uncertainty, lack of resources and time-pressure. These 
@@
 enough to enter the market early while controlling the amount of accumulated 
-technical debt.  As confirmed by analyzing other frameworks in SE, the use of 
-prescriptive and rigid structures over the development activities would hinder 
-performance since startups operate across the border between the chaotic and 
-complex domain. %MUN: Not sure whether you have discussed/mentioned the meaning 
+technical debt.  %MUN: Not sure whether you have discussed/mentioned the meaning 
 %of chaotic/complex domain in the paper, I think not. Hence you need to 
-%rephrase that (but you can not introduce new concepts here in the conclusion). 
+%rephrase that (but you can not introduce new concepts here in the conclusion). %CG: fixed. 
 Therefore, from a software development perspective, working with low-precision 

### 2013-02-24 article/main.tex
fixed comments version 0.2

@@
 Even in this case the spiral methodology%MUN: which model? CG:fixed
-is compliant with the nature of a random, ad-hoc and \textit{evolutionary development approach} (CAT 2). In fact, evolutionary development approaches are typically more 
+is compliant with the nature of a random, ad-hoc and \textit{evolutionary development approach} (CAT2). In fact, evolutionary development approaches are typically more 
 suitable  for web applications, as confirmed in \cite{Deshpande2001}.
 Moreover, they provide practices oriented towards rapid and iterative design with complete freedom 
-to the developer team, enabled to operate anywhere within the organization ( 
-compliant also to the category.%MUN: <-- the whole paragraph is not that clear to me. Try to start 
+to the developer team, enabled to operate anywhere within the organization (compliant also to the category.%MUN: <-- the whole paragraph is not that clear to me. Try to start 
 %with the category (team is catalyst), then write about Deakins spiral model
@@
 %------------------------------------------------------------------------------%
-\subsection{Pay back of the technical debt}
+\subsection{Paying back the technical debt}
 The initial lack of structures and processes negatively affects the 

### 2013-02-26 article/main.tex
Small fixes and comments marked as MUN1

@@
 %from Carmel, put it in quotation marks, like you did in the previous section 
-%with Suttons quote. Generally, check that you %are consistent how you use 
+%with Suttons quote. Generally, check that you are consistent how you use 
 %quotations throughout the paper. 
 %CG: fixed, btw  it was just to emphasize his the focus of his study. 
+%MUN1: Ok. As you may have noticed, I have removed a lot of cases were you use 
+%italics. If you emphasize *too* much, it it rather distracting than helpful 
+%for the reader. Furthermore, it conflicts with you convention you use to 
+%highlight high level categories of GSM.
 He noticed how these companies were particularly innovative and successful, 
@@
 the development team's creativity and flexibility, and to 
-the need of a quick delivering product process environment \cite{Coleman2008a}.%MUN: the second part of the sentence is just a statement without reference. If it refers to a 
-%statement by Coleman, put his reference at the end OR find the correct 
-%reference OR delete the statement.
+the need of a quick delivering product process environment \cite{Coleman2008a}.
+%MUN: the second part of the sentence is just a statement without reference. If 
+%it refers to a statement by Coleman, put his reference at the end OR find the 
+%correct reference OR delete the statement.
 %CG fixed
@@
 project to third party solutions by means of outsourcing activities 
-\cite{Hanna2010}, software reuse \cite{Jansen2008} and open-source strategies , \cite{Wall2001, Bean2005}. %MUN: Can you split up the references? I.e. write:
+\cite{Hanna2010}, software reuse \cite{Jansen2008} and open-source strategies , 
+\cite{Wall2001, Bean2005}. 
+%MUN: Can you split up the references? I.e. write:
 %"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source 
@@
 %you name it     in the text. Also, add a section (3.5) that describes shortly
-%what you did in "Theory  Validation" CG: fixed... fontsize cannot be increased
-%uniformly among all the boxes, then I skipped that part...
+%what you did in "Theory  Validation" 
+%CG: fixed... fontsize cannot be increased uniformly among all the boxes, then 
+%I skipped that part...
+%MUN1: OK, but in the figure, the correct section number for theory validation 
+%is 3.5, not 3.4
 
@@
 All companies, except C10, were founded within the last three years (2009-2012), 
-%MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct? CG: right
+%MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct? 
+%CG: right
 by an average of 3 starting members, which were in majority developers. 
@@
 quickly obtaining responses, questionnaires have been sent to respondents 
-immediately after transcribing the interview results. Most of the follow-up questionnaires have been collected within one day after the 
-conclusion of the last interview.   %MUN: I don't understand this: you
+immediately after transcribing the interview results. Most of the follow-up  
+questionnaires have been collected within one day after the conclusion of the 
+last interview.   
+%MUN: I don't understand this: you
 %collected the data from questionnaires after ALL interviews have been 
@@
 %(data analysis and triangulation let themes emerge which in turn were used
-%in subsequent interviews). CG: fixed...
+%in subsequent interviews). 
+%CG: fixed...
 
@@
 %CG: fixed
-
-Presenting GT is challenging for a researcher, who must pay attention to structure the level of details included, and to the way data is portrayed to display evidence of the emergent categories.
-In order to assess our research and determine if the GT is sufficiently grounded, we used a systematic technique to assess the theory, following evaluation lists and factors presented in  \cite{Strauss1998}.
-
-First we had to consider the following factors:
+%MUN1: This description is way too long and verbose. It should be MAX 1 column, 
+%and focus on what you present in Section 5. Theory Validation. Until the 
+%second-last paragraph you don't talk about that at all! Remove the "theory" on 
+%GT validation and what you did *during* the GT process. Keep what you need 
+%to motivate
+% - 5.1 Comparison of frameworks
+% - 5.2 Theoretical categories and existing literature
+% - 5.3 Confounding factors
+% - 5.4 Static assessment of the GSM
+%You need to explain why you did this and how it contributes to the validation 
+%of the theory.
+
+Presenting GT is challenging for a researcher, who must pay attention to  
+structure the included level of detail, and to the way data is portrayed to 
+display evidence of the emergent categories.  In order to assess our research 
+and determine if the GT is sufficiently grounded, we used a systematic technique 
+to assess the theory, following evaluation lists and factors presented in  
+\cite{Strauss1998}.
+
+We consider the following factors:
 
@@
 \begin{compactitem}
-\item \textit{Fit}, that describes the consistence of the theory with the real data.
-\item \textit{Understanding}, that describes the degree to which participants agree with the theory.
-\item \textit{Generality}, that describes the level of abstraction of the theory, and how sufficiently it can be adopted by practitioners without losing its relevance.
-\item \textit{Control}, that describes how well the theory enables practitioners to fully understand the described situation.
+\item \textit{Fit}, that describes the consistency of the theory with the real 
+data.
+\item \textit{Understanding}, that describes the degree to which participants 
+agree with the theory.
+\item \textit{Generality}, that describes the level of abstraction of the 
+theory, and how sufficiently it can be adopted by practitioners without losing 
+its relevance.
+\item \textit{Control}, that describes how well the theory enables 
+practitioners to fully understand the described situation.
 \end{compactitem}
@@
 
-In order to achieve a high-level of fit between the theory and the raw empirical data, we started GT without any assumption on how the emergent patterns would influence the research direction.
-Ideas and perceptions evolved during the mapping study and interviews.
-In addition, constant comparative methods, overturning of some early categories as new data came to light, generation and testing of interim hypotheses, and constant re-evaluation of the interview transcripts, ensured that researchers' bias was minimized and theoretical fit maintained.
-
-To represent understandable emergent theories, concepts and categories were carefully developed using appropriate and self-explanatory codes.
-The developing theory was not immediately presented to the participants to prevent potential bias in the responses.
-But rather, the theoretical framework has been presented and described to startups practitioners not involved in the data collection, whose reaction was positively expressed.
-
-From generality perspective, we developed a theory from abstract categories, which allowed us to create a general guide to constantly changing situations.
-Nevertheless, we took attention on maintaining a reasonable level of abstraction, keeping focus on their sensitizing characteristics and providing low-level subgroup of categories.
-
-Finally, we provided a comprehensive set of categories with detailed interrelations to explain how the \textit{core-category} is achieved and the reasons of its consequences, in order to enable practitioners to independently understand and analyze situations, predict changes and their consequences, and to be capable of revising actions or the theory itself.
-Using both methodological tools and features provided by \textit{AtlasTI}, we explored and tested:
+In order to achieve a high-level of fit between the theory and the raw 
+empirical data, we started GT without any assumption on how the emergent 
+patterns would influence the research direction. Ideas and perceptions evolved 
+during the mapping study and interviews. In addition, constant comparative 
+methods, overturning of some early categories as new data came to light, 
+generation and testing of interim hypotheses, and constant re-evaluation of the 
+interview transcripts, ensured that researchers' bias was minimized and 
+theoretical fit maintained.
+
+To represent understandable emergent theories, concepts and categories were 
+carefully developed using appropriate and self-explanatory codes. The developing 
+theory was not immediately presented to the participants to prevent potential 
+bias in the responses. But rather, the theoretical framework has been presented 
+and described to startups practitioners not involved in the data collection, 
+whose reaction was positively expressed.
+
+From generality perspective, we developed a theory from abstract categories, 
+which allowed us to create a general guide to constantly changing situations. 
+Nevertheless, we took attention on maintaining a reasonable level of 
+abstraction, keeping focus on their sensitizing characteristics and providing 
+low-level subgroup of categories.
+
+Finally, we provided a comprehensive set of categories with detailed 
+interrelations to explain how the \textit{core-category} is achieved and the 
+reasons of its consequences, in order to enable practitioners to independently 
+understand and analyze situations, predict changes and their consequences, and 
+to be capable of revising actions or the theory itself. Using both 
+methodological tools and features provided by \textit{AtlasTI}, we explored and 
+tested:
 
@@
 
-Additionally, Strauss and Corbin provided a list of questions to assist in determining how well the findings are grounded \cite{Strauss1998}. They questions are as follows:
+Additionally, Strauss and Corbin provided a list of questions to assist in 
+determining how well the findings are grounded \cite{Strauss1998}. They 
+questions are as follows:
 
@@
 \item Are there many conceptual linkages and are the categories well developed?
-\item Is variation\footnote{Variation refers to the variety of contexts which theory can be applied.} built into the theory and are the conditions under which variation can be found built into the study and explained?
-\item Are the conditions under which variation can be found built into the study and explained?
-\item Has the process been taken into account, and does the theory stand the test of time?
+\item Is variation\footnote{Variation refers to the variety of contexts to 
+which the theory can be applied.} built into the theory and are the conditions 
+under which variation can be found built into the study and explained?
+\item Are the conditions under which variation can be found built into the 
+study and explained?
+\item Has the process been taken into account, and does the theory stand the 
+test of time?
 \item Do the theoretical findings seem significant, and to what extent?
@@
 
-Answering the above questions, the concepts were generated according to the coding process.
-They were systematically related through the use of a network diagram (see figure \ref{fig:gt:completemethodology}), by which we established the linkages and relations between the concepts.
-At each iteration of the grounded theory process, we considered and examined a concept within different conditions and dimensions, trying to incorporate opinions of a broader as possible range of practitioners.
-
-All the linkages and categories were constructed by the use of Atlas.TI and compared according to the data analysis process.
-Variations of the generated theory were applied according to raw grounded data with the aid of codes and memos.
-Moreover during this study extensive explanation of the concepts were presented in-vivo statements as reported by practitioners.
-
-The process has been designed by the researches in different steps, explaining and making clear the purpose of each of them.
-Thus, the same processes enable other research to conduct the same study within similar contexts.
-Moreover \textit{comparison with the state-of-the-art} has been applied in order to validate the theory and enhance odds of its applicability within a wider time-frame.
-By this comparison we  highlight the areas which have been neglected by existing studies, providing possible directions for future studies (see subsection \ref{sect:theory:validation:others}).
-Moreover we study the confounding factors which could interfere with the application of the GSM model (see subsection \ref{sect:an:comp:cat-vs-literature:conf}).
-
-Then, we performed an \textit{evaluation with empirical data} (see subsection \ref{sect:theory:validation:stats}), quantifying scores for each company involved in our empirical study by defining metrics based on a set of rubrics and evaluating startups accordingly, through an analysis of interview transcripts and codes.
-The complete statistical procedure and provided  rubrics  to quantify the measures are illustrated in Appendix.
+Answering the above questions, the concepts were generated according to the 
+coding process. They were systematically related through the use of a network 
+diagram (see Figure \ref{fig:gt:completemethodology}), by which we established 
+the relations between the concepts. At each iteration of the 
+grounded theory process, we considered and examined a concept within different 
+conditions and dimensions, trying to incorporate data from a broader range of 
+practitioners.
+
+All the linkages and categories were constructed by the use of Atlas.TI and 
+compared according to the data analysis process. Variations of the generated 
+theory were applied according to raw grounded data with the aid of codes and 
+memos. Moreover during this study extensive explanation of the concepts were 
+presented in-vivo statements as reported by practitioners.
+
+The process has been designed by the researches in different steps, explaining 
+and making clear the purpose of each of them. Thus, the same processes enable 
+other research to conduct the same study within similar contexts. Moreover 
+\textit{comparison with the state-of-the-art} has been applied in order to 
+validate the theory and enhance odds of its applicability within a wider 
+time-frame. By this comparison we  highlight the areas which have been neglected 
+by existing studies, providing possible directions for future studies (see 
+subsection \ref{sect:theory:validation:others}). Moreover we study the 
+confounding factors which could interfere with the application of the GSM model 
+(see subsection \ref{sect:an:comp:cat-vs-literature:conf}).
+
+Then, we performed an \textit{evaluation with empirical data} (see subsection 
+\ref{sect:theory:validation:stats}), quantifying scores for each company 
+involved in our empirical study by defining metrics based on a set of rubrics 
+and evaluating startups accordingly, through an analysis of interview 
+transcripts and codes. The complete statistical procedure and provided  rubrics  
+to quantify the measures are illustrated in Appendix.
 
@@
 %CG:fixed
+%MUN1: this should be \ref{sect:theory:validation:stats}, right? Section 5 is 
+%quite long and you should point to the correct subsection
 
@@
 %CG fixed
+%MUN1: Same here, point to the correct subsection.
 .}. At the same time, to be able to deal with such constraints, startups must 
@@
 %this statement since it is a citation? 
-%CG: not fixed. It represents a subcategory, expressed by the 70% of the interviewed startups
+%CG: not fixed. It represents a subcategory, expressed by the 70% of the 
+%interviewed startups
+%MUN1: and how should the reader know that? See my problem with italics: you 
+%use italics for both interview statements and (sub-)categories. 
 mainly constrained by a tight \textit{time shortage}.
@@
 achieved. Product metrics are typically web-based \textit{statistical hypothesis 
-testing}, such as A/B testing \cite{AB}. %MUN: add reference or explain shortly what A/B 
-%testing is
+testing}, such as A/B testing \cite{AB}. 
+%MUN: add reference or explain shortly what A/B testing is
 %CG: fixed
@@
 technologies can be autonomously managed by the development team (with 
-characteristics of: co-location, self-organization and very small team size). %MUN: 
-%do you mean small team size?
+characteristics of: co-location, self-organization and very small team size). 
+%MUN: do you mean small team size?
 %CG: fixed
@@
 %CG:fixed
+%MUN1: Ok, but fix section 3.5... you see that you don't motivate 5.1-5.4 at 
+%all there and that the current content of 3.5 is not related to what you 
+%actually do here as validation?
 
@@
 defined from the GSM that are in support of previous models developed by 
-Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}, Baskerville \cite{Internet} and finally Brooks \cite{BrooksJr1987}.
-
-Starting from the Coleman's work, we refer to him since only he has, to the best of our knowledge, 
-conducted a similar research in startups' context, even though with a 
-different focus. Coleman investigated 
-factors in software development that hinder initiatives of software process 
-improvement (SPI) in a later stage, representing also companies in expansion phase with more than 100 employees. %MUN: so he did not do research 
+Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}, Baskerville 
+\cite{Internet} and finally Brooks \cite{BrooksJr1987}.
+
+Starting from the Coleman's work, we refer to him since only he has,  to the 
+best of our knowledge, conducted a similar research in startups' context, even 
+though with a different focus. Coleman investigated factors in software 
+development that hinder initiatives of software process improvement (SPI) in a 
+later stage, representing also companies in expansion phase with more than 100 
+employees. %MUN: so he did not do research 
 %in a startup context? You need to clarify what you mean by "more mature 
@@
 letter of them and they are always adapted, so we adapted ours to the way we 
-wanted it to work for us, for our own size and scale''} \cite{Coleman2008}. %MUN: add ref to 
-%Coleman's paper where he states this
+wanted it to work for us, for our own size and scale''} \cite{Coleman2008}. 
+%MUN: add ref to Coleman's paper where he states this
 %CG fixed
@@
 penetrating small software organizations \cite{Staples2007}. When startups began 
-establishing any SPI process, they experienced \textit{process erosions} \cite{Coleman2008} %MUN: 
+establishing any SPI process, they experienced \textit{process erosions}  
+\cite{Coleman2008} %MUN: 
 %add ref (Coleman?). You should add, for each category you use from Coleman, 
@@
 found that despite Coleman's study contributes to define which factors hinder
-the introduction of SPI in startups %MUN: add: "of SPI in startups"? CG: fixed ,
+the introduction of SPI in startups 
+%MUN: add: "of SPI in startups"? CG: fixed
 his results are coherent with the theory generated within our research, focusing
-at explaining the overall strategies within software development in
-startups%MUN:  %add "our research, focused at explaining software development in
-general in  %startups." ? CG:fixed . In this regard, a wide variability of
-processes is a key factor for startups.  In contrast with some SPI processes,
-such as \textit{Six Sigma} which objective  is minimizing unpredictability in
-the definition of workflows \cite{Sixsigma},  startups seem to follow the
-\textit{Fischer's Fundamental Theorem of natural  selection} \cite{Fisher}.
+at explaining the overall strategies within software development in startups.
+%MUN: add "our research, focused at explaining software development in
+%general in startups." ? CG:fixed
+In this regard, a wide variability of processes is a key factor for startups.  
+In contrast with some SPI processes, such as \textit{Six Sigma} which objective 
+is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, 
+startups seem to follow the \textit{Fischer's Fundamental Theorem of natural  
+selection} \cite{Fisher}.
 Moving towards flexible and variable processes (easy  to adapt to the uncertain
@@
 of resources, %MUN: under which  %circumstances? This statement sounds too
-general, i.e. don't do SPI ever... %CG: fixed %And the following part,I did not
-fully understand. Try to clarify. CG:fixed  not following SPI does not predict
+%general, i.e. don't do SPI ever... %CG: fixed 
+%And the following part,I did not fully understand. Try to clarify. CG:fixed  
+not following SPI does not predict
 major adaptiveness of startups development towards the customer's needs.
@@
 
-%CG: partially fixed. Agreed on the beginning of the section, not agreed on the division in subsections. Let's discuss it by a skype call.
+%CG: partially fixed. Agreed on the beginning of the section, not agreed on  the 
+%division in subsections. Let's discuss it by a skype call.
 
@@
 typically effective only in large-scale, long-term development efforts with
-stable and disciplined processes.Internet-speed software development (i.e. an
+stable and disciplined processes. Internet-speed software development (i.e. an
 agile method oriented using daily builds aimed at developing a product with high
@@
 is only partially covering the phenomenon described by the GSM. Looking at the 
-papers' main contribution %MUN: which results? The papers' main contribution or...? %CG: fixed
+papers' main contribution%MUN: which results? The papers' main contribution 
+%or...? %CG: fixed
 , we were able to extrapolate information which confirmed the GSM categories.
@@
 need of empowering people, which is the element with the greatest impact on 
-software development. Humans, being non-linear variables \cite{Human} %MUN: add ref %CG:fixed
+software development. Humans, being non-linear variables \cite{Human}
+%MUN: add ref %CG:fixed
 , are unlikely to follow repeatable prescriptive methodologies. As stated by 
@@
 %identifier you use in table 2. I did not check if you discuss each and every 
-%category. Do you? If yes, ok. If not, why not? CG: partially fixed, the categories CAT3 and CAT6 are not well covered as discussed above.
+%category. Do you? If yes, ok. If not, why not? 
+%CG: partially fixed, the categories CAT3 and CAT6 are not well covered as 
+%discussed above.
 Starting from the category \textit{severe lack of resources} (CAT7), many 
@@
 are discussed in an experience paper by Ambler \cite{Ambler2002}. He reports on 
-two case studies of growing Internet startups approaching an initial public offering (IPO). %MUN:explain 
-%acronym %CG: fixed
+two case studies of growing Internet startups approaching an initial public  
+offering (IPO). %MUN:explain %acronym %CG: fixed
 Despite the two companies were facing business and team growth, attitudes 
@@
 evident. The need to re-engineer the codebase lead the researchers to suggest a 
-tailored version of the Rational Unified Process (RUP) %MUN: add reference
+tailored version of the Rational Unified Process (RUP) 
+%MUN: add reference
 %CG: not fixed, not sure if needed since Ambler has the detailed 
- %explaination of his process, similar to RUP... anyway a dedicated resource would be an IBM website
+%explaination of his process, similar to RUP... anyway a dedicated resource 
+%would be an IBM website
 development framework. Nevertheless, even with the aid of CASE 
@@
 %MUN: the whole paragraph needs to be summarized to its essence. Details can be 
-%read in the referred paper. CG: partially fixed, removed the detailed process description
+%read in the referred paper. 
+%CG: partially fixed, removed the detailed process description
 In the evaluation phase, the status of the development workflow within the 
@@
 %system testing, acceptance testing and release deployment. 
-Despite the effort 
-spent by the researchers to establish a development workflow, practices were not 
-fully conducted even though some marginal benefits have been reported: planned 
-major release from four to six months; planned minor releases from two to four 
-weeks; encompassing urgent corrective releases from one to five days. Regarding 
-the previously mentioned problems, only 
+Despite the effort spent by the researchers to establish a development workflow, 
+practices were not fully conducted even though some marginal benefits have been 
+reported: planned major release from four to six months; planned minor releases 
+from two to four weeks; encompassing urgent corrective releases from one to five 
+days. 
+Regarding the previously mentioned problems, only 
 partial improvements have been achieved since the major obstacle was to 
@@
 
-Prescribed development models, oriented to SPI, are basically ignored by 
-startups, as discussed in \cite{Zettel2001}. %MUN: <-- rephrase, not clear to me %CG:fixed
+Prescribed development models, oriented towards SPI, are basically ignored by 
+startups, as discussed in \cite{Zettel2001}. 
+%MUN: <-- rephrase, not clear to me %CG:fixed
 To overcome the need of structure and to accelerate the development process, 
@@
 
-Deakins \cite{Deakins2005} designed a new spiral methodology, constructed during a
-fieldwork with a developer team that had limited resources managing an 
-innovative product in volatile e-commerce environments. This methodology represents a fast 
-evolutionary approach experimentation with flexible processes. It aims to obtain
-customer's feedback as soon as possible for continuous product improvement.%MUN: 
-%<-- rephrase, sentence not parsable CG:fixed
-Even in this case the spiral methodology%MUN: which model? CG:fixed
-is compliant with the nature of a random, ad-hoc and \textit{evolutionary development approach} (CAT2). In fact, evolutionary development approaches are typically more 
-suitable  for web applications, as confirmed in \cite{Deshpande2001}.
-Moreover, they provide practices oriented towards rapid and iterative design with complete freedom 
-to the developer team, enabled to operate anywhere within the organization (compliant also to the category.%MUN: <-- the whole paragraph is not that clear to me. Try to start 
+Deakins \cite{Deakins2005} designed a new spiral methodology, constructed  
+during a fieldwork with a developer team that had limited resources managing an 
+innovative product in volatile e-commerce environment. This methodology 
+represents a fast evolutionary approach with flexible processes. 
+It aims to obtain customer's feedback as soon as possible for continuous 
+product improvement.%MUN: %<-- rephrase, sentence not parsable CG:fixed
+Even in this case, the spiral methodology%MUN: which model? CG:fixed
+is compliant with the nature of a random, ad-hoc and  \textit{evolutionary 
+development approach} (CAT2). Evolutionary development approaches are 
+typically more suitable for web applications, as confirmed in 
+\cite{Deshpande2001}.
+Moreover, they provide practices oriented towards rapid and iterative design 
+with complete freedom to the developer team, enabled to operate anywhere within 
+the organization.
+%MUN: <-- the whole paragraph is not that clear to me. Try to start 
 %with the category (team is catalyst), then write about Deakins spiral model
@@
 
-%MUN: To which category does the following paragraph belong? CG: they represent subcategories of speed-up development 
-Eventually to externalize complexity of the developing product %MUN: solutions on what? CG: fixed
+%MUN: To which category does the following paragraph belong? 
+%CG: they represent subcategories of speed-up development 
+Eventually to externalize complexity of the developing product
+%MUN: solutions on what? CG: fixed
 , two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the 
@@
 %phenomenon" mean?). 
-CG: not fixed, when I wrote it I tried to have an overview of the main contributions and make it fluently by papers description.
-Creating subsections would mean to divide papers from different focuses.
-This is a trade-off where we took a decision in the past.
-Take it in consideration, but we can discuss it during the next meeting.
+%CG: not fixed, when I wrote it I tried to have an overview of the main 
+%contributions and make it fluently by papers description. Creating subsections 
+%would mean to divide papers from different focuses. This is a trade-off where 
+%we took a decision in the past. Take it in consideration, but we can discuss 
+%it during the next meeting.
+%MUN1: Why would creating subsubsections like 5.2.1 Severe lack of resources 
+%(CAT7) remove the flow? This subdivision would give the reader clues what the 
+%subsubsection is about, he can relate it to Table 2, and provides some 
+%navigational support. Now there are nearly 2p of flowing text with now clue in 
+%what order/priority the literature is covered. I think you don't even need to 
+%rewrite much, just add the subsubsections and a proper summary.
 
@@
 \cite{ColinRobson2009}. We report those factors that have been explicitly 
-identified by the SMS \cite{SMS} %MUN: do you mean the SMS? Then refer to it. CG: fixed 
+identified by the SMS \cite{SMS} %MUN: do you mean the SMS? Then refer to it.
+%CG: fixed 
 %I remember you mentioned CF in the results (GSM), maybe also refer to the 
 %respective sections CG: not parsed.
+%MUN1: me neither... forget what I wrote :-)
 , related to: creativity and innovation, market requirements and application 
@@
 %long-term scalability issues."  CG: ok.
-
 %MUN: Please reformulate the first sentence, I cannot parse it. CG: fixed
-Based on the model presented in \cite{Heitlager2007}, it identifies the 
-lifecycle stages of a startup (\textit{fluid}, \textit{transition} and 
-\textit{specific}). At each stage, the model presents different development approaches 
-(product-oriented and process-oriented), according to the level of innovation achieved. The 
-study reports how product-oriented development, in contrast with forced and 
-mature process imposition, provides a degree of freedom to the development team, 
-that enhances the creativity of developers and augments the innovation 
-capability of the company in the early-stage.
+The proposed model identifies the three lifecycle stages of a startup: 
+\textit{fluid}, \textit{transition} and \textit{specific}. At each stage, the 
+model presents different development approaches (product-oriented and 
+process-oriented), according to the level of innovation achieved. The study 
+reports how product-oriented development, in contrast with forced and mature 
+process imposition, provides a degree of freedom to the development team, that 
+enhances the creativity of developers and augments the innovation capability of 
+the company in the early-stage.
 
@@
 considered in this model, the integration of results would have been 
-straightforward, since it is totally compliant with the description of how the \textit{accumulated technical debt} CAT5 causes the need of establishing a process in the long run (see how the \textit{initial growth hinders performance} CAT6).
-%MUN: not clear to me what you mean by "evolution
+straightforward, since it is compliant with the description of how the 
+\textit{accumulated technical debt} (CAT5) causes the need of establishing a 
+process in the long run (see how the \textit{initial growth hinders performance} 
+(CAT6)). %MUN: not clear to me what you mean by "evolution 
 %of this model" and what "compliant with our assumption" means. Also, don't use 
@@
 confounding factor seems to behave initially the same as the \textit{speed-up 
-development}, both highly focused towards a product-oriented development. %MUN: <-- please provide some motivation... not obvious. CG: fixed
+development}, both highly focused towards a product-oriented development. 
+%MUN: <-- please provide some motivation... not obvious. CG: fixed
 Nevertheless, we were unable to ground these evidences in the GT data even 
-though they are two factors that might be positively correlated %MUN: "present strict correlation" 
-%What does that mean, I don't understand. CG: fixed.
-to the choice of an early product-development orientation and the need of low-precision 
-engineering elements.
+though those are factors that might be positively correlated 
+%MUN: "present strict correlation" %What does that mean, I don't understand. 
+%CG: fixed.
+to the choice of an early product-development orientation and the need of 
+low-precision engineering elements.
 
@@
 For example, safety-critical systems possess quality concerns that must be 
-considered. Even in e-commerce systems 
-factors related to quality have high-impact effects on obtaining the first customer reference, %MUN: <-- 
-%cannot parse it. Remove the first part and write: "Even e-commerce 
-%systems...". Then rephrase the second part of the sentence to make it clearer. CG: fixed
-as discussed in \cite{Deakins2005,Kim2005,Silva2005}. Quality attributes especially %MUN: what is "this"? CG: fixed
+considered. Even in e-commerce systems factors related to quality have 
+high-impact effects on obtaining the first customer reference, 
+%MUN: <-- cannot parse it. Remove the first part and write: "Even e-commerce 
+%systems...". Then rephrase the second part of the sentence to make it clearer. 
+%CG: fixed
+as discussed in \cite{Deakins2005,Kim2005,Silva2005}. Quality attributes  
+especially %MUN: what is "this"? CG: fixed
 characterize those applications, whose part of functionalities are already 
@@
 \subsubsection{Developer experience} 
-%MUN: Ups, a confounding factor you did not mention initially. Add it... CG: fixed
+%MUN: Ups, a confounding factor you did not mention initially. Add it... 
+%CG: fixed
 Finally, also the experience of team-members is considered as an important 
@@
 our model by attesting their correctness  with a quantification of empirical 
-trends observed in the interview data. We identify trends and recurrent patterns 
-by thoroughly analyzing and synthesizing the complete set of data obtained from 
-the case study on which the GSM is built upon. We illustrate our observations 
-aided by simple statistical methods, providing researchers an additional tool, 
-supported by further companies' data, to validate our model.
+trends observed in the interview data. 
+%MUN1: The following is part of the method and should be in 3.5: "We identify 
+%trends and recurrent patterns by thoroughly analyzing and synthesizing the 
+%complete set of data obtained from the case study on which the GSM is built 
+%upon. We illustrate our observations aided by simple statistical methods, 
+%providing researchers an additional tool, supported by further companies' 
+%data, to validate our model."
 
@@
 
-%MUN: for the paragraphs below, you need to explain what and why you do this. CG: fixed, explained in methodology 
+%MUN: for the paragraphs below, you need to explain what and why you do this. 
+%CG: fixed, explained in methodology 
 %What do you mean by "we grounded in the interview transcripts the fact that 
 %those relations were characterized in terms of performances since 
-%time-to-market was the main concern." Not clear to me. CG: fixed, performance for us is efficiency AND effectiveness
+%time-to-market was the main concern." Not clear to me. 
+%CG: fixed, performance for us is efficiency AND effectiveness
 Before starting the evaluation of the impact of the three categories on 
 \textit{speed-up development}, we grounded in the interview transcripts the fact 
-that those relations were characterized in terms of "effectiveness" and "efficiency" since 
-time-to-market was the main concern (as confirmed in \cite{SMS}):
+that those relations were characterized in terms of "effectiveness" and 
+"efficiency" since time-to-market was the main concern (as confirmed in 
+\cite{SMS}):
 
 \begin{compactitem}
-
 \item The team enhances the overall \textit{performance} of software 
@@
 %ONLINE appendix and refer to it. Still, you need to report the actual RESULTS 
-%of you statistical analysis here... CG: partially fixed. we always referred to appendix without references until now.
-
-We conducted statistical tests using the analysis of variance to assess the existence of relations between the dimensions.
-First we defined two null hypotheses (H0): $H0_1$ = \textit{Startups do not release the product faster when a capable team adopt a more evolutionary approach AND with less quality constraints}; $H0_2$ = \textit{The execution speed does not increase the amount of accumulated technical debt}.
-Then we tested $H0_1$ and $H0_2$  with an one-tailed test using Pearson's product moment correlation coefficient\footnote{To perform the Pearson's correlation we verified two assumptions: a) data is normally distributed; b) data is on interval scale.}, with positive association analysis, fixing the level of confidence to 95\% which means we reject H0 in case the p-value is lower than 0.05.
-
-We conclude that, in our sample:
-\par\vspace {0.4cm}
+%of you statistical analysis here... 
+%CG: partially fixed. we always referred to appendix without references until 
+%now.
+
+%MUN1: I don't really understand why you add the second clause in H01 
+%"AND with less quality constraints" --> just makes the hypothesis more 
+%difficult to understand. 
+We conducted statistical tests  using the analysis of variance to assess the 
+existence of relations between the dimensions. First we defined two null 
+hypotheses:
+
+\begin{itemize}
+\item[$H0_1$:] \textit{Startups do not release the product faster 
+when a capable team adopts a more evolutionary approach AND with less quality 
+constraints}
+\item[$H0_2$:] \textit{The execution speed does not increase the amount 
+of accumulated technical debt}. 
+\end{itemize}
+
+Then we tested $H0_1$ and $H0_2$  with an 
+one-tailed test using Pearson's product moment correlation 
+coefficient\footnote{To perform the Pearson's correlation we verified two 
+assumptions: a) data is normally distributed; b) data is on interval scale.}, 
+with positive association analysis, fixing the level of confidence to 95\%.
+We reject both $H0_1$ and $H0_2$, and conclude that in our sample:
 \begin{compactenum}
-\item Higher values of \textit{Execution speed} are strongly associated\footnote{Clear statistical significance, p-value: 0.002073.} with \textit{higher} values for  \textit{Technical debt}.
-
-\item Higher values of  \textit{Execution speed}  are strongly associated\footnote{Clear statistical significance, p-value: 0.004549.} with  \textit{higher values} for  \textit{Potential capability}.
+\item Higher values of \textit{Execution speed} are strongly associated 
+(p-value: 0.002) with higher values for \textit{Technical debt}.
+
+\item Higher values of \textit{Execution speed} are 
+strongly associated (p-value: 0.005) with higher values for  
+\textit{Potential capability}.
 \end{compactenum}
-\par\vspace {0.4cm}
 
@@
 the model. According to the paradigm model those are the causal condition which 
-are linked to the core-category. There are four high-level relationships  which 
-have not been validated using numerical methods. However their correctness has 
-been verified with a cross-analysis of the extant literature and by attesting 
-the causal relations among codes and categories, through axial and selective 
-coding processes.
+are linked to the core-category (see Figure \ref{fig:gsm}). There are four 
+high-level relationships  which have not been validated by quantitative 
+methods. However their correctness has been verified by a cross-analysis with 
+existing literature and by attesting the causal relations among codes and 
+categories, through axial and selective coding processes.
 
@@
 %long as your arguments follow from the previous analysis.
-%CG: not fixed, I have shortened the content of one page.
-But still we cannot create the subsections you proposed since the implications are all derived from the model.
-Then, we can still shorten it if you want.
-Let's discuss it during the next skype call.
+%CG: not fixed, I have shortened the content of one page. But still we cannot 
+%create the subsections you proposed since the implications are all derived 
+%from the model. Then, we can still shorten it if you want. Let's discuss it 
+%during the next skype call.
+
+%MUN1: Ok, the subdivision is better than the plain bullet list and I'm fine 
+%with the subsection names. Still, I would rename the section: "Discussion" --> 
+%"Implications of the GSM"
 
 In this section we present relevant implications which 
-emerge from the behavior of early-stage startups, formally expressed in the GSM 
-model, answering to our research question (\textit{What is the current 
+emerge from the behavior of early-stage startups, formally expressed in the GSM, 
+answering to our research question (\textit{What is the current 
 state-of-practice related to software development strategy in early-stage 
@@
 case study were spread across different nations and market sectors, we have 
-found several patterns that emerged very clearly:
+found several patterns that emerged very clearly.
 
@@
 
-The most urgent priority of software development is to shorten 
+The most urgent priority of software development in startups is to shorten 
 time-to-market. When developing a new product, which has not been attempted 
@@
 functionalities progressively rolled-out to early-adopters.
+%MUN1: Ok, I'm now "inventing" possible implications in the sections below... 
+%this is your job and you are much more qualified to do that than me :-)
+
+%MUN1:Do we (research) know enough about how to shorten time-to-market without 
+%accumulating technical debt? What are possible venues for research?
 
@@
 methodology, where they should set-up an experiment to \textit{test the riskiest 
-element of the business idea}, nor make explicit use of the good practices 
-suggested in the book. For example, when it comes to user feedback, startups do 
+element of the business idea}, nor make explicit use of the suggested good 
+practices.
+%MUN1: Why don't follow startups the build-measure-learn cycle proposed by 
+%Ries? Well, naively I'd say that since he published in 2011, his ideas do not 
+%have yet transferred to industry... Any other reason you can think of? Is it 
+%still too heavyweight and in conflict with the highest priority of shorting 
+%time-to-market? Is Lean still to Heavy for startups (in their early phases)?
+For example, when it comes to user feedback, startups do 
 not explicitly follow the step-by-step process of \textit{customer development}
-formally defined by \cite{Blank2005}. Instead, startups absorb and implement 
-the high-level principles that outlines the methodology, reflected in the model 
-by the theoretical category \textit{find the product/market fit quickly}.
-
-The greatest part of engineering activities of startups are focused on 
-the implementation while only little attention is given to more conventional 
-activities (project management, requirement specifications, analysis, 
-architecture design, automatic testing, \ldots). Especially when bringing the 
-first product to the market, the attitude of startups towards traditional 
-engineering activities is strongly characterized by a general feeling of 
-unnecessary. Every document, artifact or process, which does not produce a 
-short-term tangible benefit to the software production is perceived as a waste 
-of time.
+formally defined by Blank \cite{Blank2005}. Instead, startups absorb and 
+implement the high-level principles that outlines the methodology%MUN1:<-- 
+%which methodology?
+, reflected in the GSM by the theoretical category \textit{find the 
+product/market fit quickly}.
+
+%MUN1: paragraph below is candidate for removal
+%The greatest part of engineering activities of startups are focused on 
+%the implementation while only little attention is given to more conventional 
+%activities (project management, requirement specifications, analysis, 
+%architecture design, automatic testing, \ldots). Especially when bringing the 
+%first product to the market, the attitude of startups towards traditional 
+%engineering activities is strongly characterized by a general feeling of 
+%unnecessary. Every document, artifact or process, which does not produce a 
+%short-term tangible benefit to the software production is perceived as a waste 
+%of time.
 
@@
 focused on a minimal and essential set of functionalities providing a reasonable 
-user experience (UX). As confirmed by practitioners, the user must be able to 
+user experience (UX). 
+%MUN1: This rises the question, how do startups prioritize requirements. 
+%There's abundant amount of literature on how to select/prioritize reqs., but 
+%do they work in startups? Are they also too heavyweight or could they be 
+%adopted by startups in a modified form in order to increase the probabilities 
+%of success in the market?
+As confirmed by practitioners, the user must be able to 
 accomplish a task on the product without major interruptions in the flow, which 
@@
 perspective, one might argue that established companies might take advantage of 
-the new idea and develop a better version of the product internally. Despite 
-this might represent a risk, established companies usually cannot afford to 
-operate in a market of \textit{innovators} and \textit{early adopters} group of 
-users - described by the Roger's bell curve in \cite{Rogers} - because of the 
-little return of investment they would gain, as discussed by Christensen in 
-\cite{Christensen2003}. Instead, established companies have more interests in 
-acquiring a startup and all the knowledge it has gained during the 
+the new idea and develop a better version of the product internally.
+Despite this might represent a risk, established companies usually cannot 
+afford to operate in a market of \textit{innovators} and \textit{early adopters} 
+group of users - described by the Roger's bell curve in \cite{Rogers} - because 
+of the little return of investment they would gain, as discussed by Christensen 
+in \cite{Christensen2003}. Instead, established companies have more interests 
+in acquiring a startup and all the knowledge it has gained during the 
 \textit{customer development} process while focusing effort on acquiring larger 
-shares of a ``mass market'' sector. Furthermore, thanks to cloud computing and 
-more in general high availability of third party components, aspects such as 
-performance, security and scalability can be achieved with a reasonable effort. 
-Thus, given the general lack of quality concerns, and supported by the fact that 
-the optimal strategy to improve the UX is by having the product used by real 
-customers, startups can effectively focus on achieving short time-to-market.
+shares of a ``mass market'' sector. 
+%MUN: candidate for shorting
+%Furthermore, thanks to cloud computing and 
+%more in general high availability of third party components, aspects such as 
+%performance, security and scalability can be achieved with a reasonable 
+%effort. 
+%Thus, given the general lack of quality concerns, and supported by the fact 
+%that the optimal strategy to improve the UX is by having the product used by 
+%real customers, startups can effectively focus on achieving short 
+%time-to-market.
 
@@
 high-colocation which enables continuous informal and colloquial coordination 
-and discussions. The initial (small) development team usually spends most of the 
-hours of the week working closely to implement the product, often extending the 
-conventional office hours.  The large amount of communication, spontaneously 
-held in informal places, has been proven to be an extremely effective element of 
+and discussions. 
+%$MUN1: candidate for removal
+%The initial (small) development team usually spends most of the 
+%hours of the week working closely to implement the product, often extending 
+%the conventional office hours.
+The large amount of communication, spontaneously held in informal places, has 
+been proven to be an extremely effective element of 
 several successful software projects \cite{Highsmith2000}. Hence, the vastly 
 \textit{informal} development environment of startups acts as a catalyst of 
-speed. Other important elements, which contribute in achieving such an initial 
-high-speed, are: minimal set of functionalities with low concerns on product 
-quality; evolutionary prototyping approach; smart use of advanced tools and 
-frameworks; low precision artifacts; and the close colocation and capability of 
-the team.
+speed.
+%MUN1: candidate for removal
+%Other important elements, which contribute in achieving such an initial 
+%high-speed, are: minimal set of functionalities with low concerns on product 
+%quality; evolutionary prototyping approach; smart use of advanced tools and 
+%frameworks; low precision artifacts; and the close colocation and capability 
+%of the team.
 
@@
 performance when the company starts to grow. One of the primary objectives of 
-startups is to expand their business into fast growing markets. In fact, if 
+startups is to expand their business into fast growing markets. If 
 successful, the company will face a growth in terms of number of customers, 
-employees and product functionalities. Then, the increased complexity arises the 
+employees and product capabilities. Then, the increased complexity arises the 
 necessity of controlling the initially chaotic software development environment. 
@@
 technical debt and the capacity of the team to absorb it. 
-
+%MUN1: what does literature say about technical debt? Are there strategies that 
+%maybe applicable to startups and if not why? Is there more research needed? 
+%Obviously, shortening time-to-market and preventing technical debt are two 
+%conflicting goals, but is there a chance to find a sweet spot where 
+%development speed can be balanced to achieve market success and still reducing 
+%the performance bump that hampers growth?
+
+\subsection{Summary}
 All of the above observations are extracted from the analysis of the model, 
 which is systematically derived from empirical data of the case study. Indeed, 
-with slightly different levels of adherence, the implication we presented are 
+with slightly different levels of adherence, the implications we presented are 
 reflected in the behavior of the great majority of companies in the sample. The 
@@
 %------------------------------------------------------------------------------%
-\subsection{Future work}
-\label{futwork}
+%\subsection{Future work}
+%label{futwork}
 %MUN: Don't have future work as a subsection in threats to validity. Either 
@@
 
-The GSM model enables practitioners in startup companies to speed-up 
-development empowering the role of the team, focusing on minimal functionalities 
-and adopting evolutionary development approaches. Researchers can use the GSM 
-model to reproduce the study with other software startups, and focus on 
-mitigation strategies to minimize the impact of the accumulated technical debt 
-over the need of shortening time-to-market. Moreover, the  mutual collaborative 
-dialog between the research in the development strategies in software startups 
-and technical debt can benefits in understanding how both influence the growth 
-of a company after the release of the first product to the market.
+%MUN1: The paragraph below does say little about future work but rather 
+%presents the contributions of GSM. In "Future work" you should rather state 
+%what you consider needed to work on in relation with the research you have 
+%done. Granted, replicating your case study and validating/extending GSM is 
+%future work too, but it is not what *you* would do. Hence I'd suggest to 
+%remove this section. Check that you state the contributions of GSM in the 
+%conclusions.
+
+%The GSM model enables practitioners in startup companies to speed-up 
+%development empowering the role of the team, focusing on minimal 
+%functionalities and adopting evolutionary development approaches. Researchers 
+%can use the GSM model to reproduce the study with other software startups, and 
+%focus on mitigation strategies to minimize the impact of the accumulated 
+%technical debt over the need of shortening time-to-market. Moreover, the  
+%mutual collaborative dialog between the research in the development strategies 
+%in software startups and technical debt can benefits in understanding how both 
+%influence the growth of a company after the release of the first product to the 
+%market.
 
@@
 
-%------------------------------------------------------------------------------%
-In this section we discuss the validity of the overall results.  We structure the validity in four parts, accordingly with Wholin taxonomy \cite{Wohlin2000}:
-
-\begin{compactitem}
-
-\item External validity - which requires establishing the domain to which the study's findings can be generalized (partially discussed above).
-
-\item Internal validity - which is enhanced by establishing causal relationships whereby we attest that certain conditions lead to other conditions, as distinguished from spurious relationships\footnote{A \textit{spurious} relationship is a relationship in which two events have no direct causal connection, but rather due to either a coincidence or the presence of confounding factors.}.
-\item Construct validity - which is enforced by clearly specifing operational procedures, and providing an adequate description of constructs\footnote{Constructs are complex ideas that we as humans form in order to summarize observations about things that we cannot see directly.
-For example a construct is \textit{``job satisfaction''}, which to be defined, requires a higher level of abstraction than a concrete object(e.g.
-a book).} that are used to generate the theory.
-\item Conclusion validity - which represent the degree to which conclusions are reasonable, according to the defined relationships in the data.
-It involves ensuring adequate sampling procedures, appropriate analytical methods and reliable measurement procedures.
-Reliability requires to demonstrate that the operations of a study - such as data collection procedures - can be repeated by other researchers, obtaining the same results.
-\end{compactitem}
+%MUN1: After reading the discussion on validity threats, I think you have
+%chosen the wrong framework (Wohlin) to discuss threats. His framework is 
+%targeted at experiments and even though it has been applied in other contexts, 
+%other taxonomies are a better fit (see Yin and also Runeson). You tried to 
+%squeeze your threats and strategies into Wohlin's taxonomy which seems in some 
+%cases quite artificial and forced. I think it is better if you adopt Yin's 
+%taxonomy of validity threats. If you read Yin (p 33-38) you'll see that his 
+%taxonomy is a much better fit and that you can easier place and motivate your 
+%mitigation strategies.
+%------------------------------------------------------------------------------%
+In this section we discuss the validity of the overall results. %MUN1: threats 
+%are discussed w.r.t. the research methodology, not the results... 
+We structure the discussion according to Wohlin's taxonomy \cite{Wohlin2000}.
+
+%MUN1: Ok, what I meant was, not to explain the 4 categories in *general*, but 
+%to explain (shortly) their relevance w.r.t. GT. I guess my comment above was 
+%misleading...
+%\begin{compactitem}
+
+%\item External validity - which requires establishing the domain to which the 
+%study's findings can be generalized (partially discussed above).
+
+%\item Internal validity - which is enhanced by establishing causal 
+%relationships whereby we attest that certain conditions lead to other 
+%conditions, as distinguished from spurious relationships\footnote{A 
+%\textit{spurious} relationship is a relationship in which two events have no 
+%direct causal connection, but rather due to either a coincidence or the 
+%presence of confounding factors.}.
+%\item Construct validity - which is enforced by clearly specifing operational 
+%procedures, and providing an adequate description of 
+%constructs\footnote{Constructs are complex ideas that we as humans form in 
+%order to summarize observations about things that we cannot see directly. For 
+%example a construct is \textit{``job satisfaction''}, which to be defined, 
+%requires a higher level of abstraction than a concrete object(e.g. a book).} 
+%that are used to generate the theory.
+%\item Conclusion validity - which represent the degree to which conclusions 
+%are reasonable, according to the defined relationships in the data. It 
+%involves ensuring adequate sampling procedures, appropriate analytical methods 
+%and reliable measurement procedures. Reliability requires to demonstrate that 
+%the operations of a study - such as data collection procedures - can be 
+%repeated by other researchers, obtaining the same results.
+%\end{compactitem}
 
@@
 %External
-The first threat related to external validity might be caused by a possible 
-wrong selection of subjects interviewed for the study. This threat profoundly 
-affects GT, since it is a qualitative research method, which makes the use of 
-semi-structured interviews, in fact, centres the research on respondent's 
-opinions. In order to mitigate this threat we selected interviewees which 
-covered positions of CTOs' and CEOs'. Their perspectives on what it is taking 
+One threat to external validity is the selection of subjects interviewed 
+for the study. This threat profoundly affects GT, since it is a qualitative 
+research method, which makes the use of semi-structured interviews, and centers 
+the research on respondent's opinions. 
+In order to mitigate this threat we selected interviewees which 
+covered positions of CTOs' and CEOs'. Their perspectives on what is taking 
 place within the organization, were the only meaningful data taken in 
 consideration in the study.  In many cases there was no supporting evidence to 
-verify the opinion expressed. Questionnaire dedicated to ``data triangulation'' 
-was not of primary support because of the little time practitioners were able to 
-grant to it. However, researchers have to accept the veracity of what 
-respondents reported during the interviews \cite{Coleman2008a}. Notwithstanding 
-the issues surrounding semi-structured interviews are vital. Within our study, 
-although the reality might be different from what has been described, the 
-reported data are CTOs' and CEOs' perception, which is used to base their 
-decisions.
-
-Finally the comparison with similar frameworks also helped in establishing the 
+verify the opinion expressed. The questionnaire dedicated to ``data 
+triangulation'' was not of primary support because of the limited time 
+practitioners were able to grant to it. However, researchers have to accept the 
+veracity of what respondents reported during the interviews \cite{Coleman2008a}. 
+%MUN1: candidate for removal
+%Notwithstanding, the issues surrounding semi-structured interviews are vital. 
+%Within our study, although the reality might be different from what has been 
+%described, the reported data are CTOs' and CEOs' perception, which is used to 
+%base their decisions.
+
+The comparison of the GSM with similar frameworks helped in establishing the 
 domain to which the study's findings can be generalized. In particular the 
-previous framework developed by Coleman in \cite{Coleman2008} has allowed 
-researchers to find similarities, differences and broader reasoning related to 
-factors that hinder mature processes to be established in startups. Moreover a 
-comparison has been conducted also with the notorious Brook's framework 
-describing the \textit{basic attacks} to the challenges of developing software 
-systems and other extant literature retrieved by \cite{SMS}\footnote{Note that 
-the generalizability of results is applied only to those startups that operate 
-in software-intensive activities.}. In conclusion, regarding the validation of 
-the subcategories, a continuous re-evaluation of the interview transcripts has 
-been performed.
+previous framework developed by Coleman \cite{Coleman2008} has allowed 
+us to find similarities, differences and broader reasoning related to 
+factors that hinder maturing processes in startups. Moreover, we analyzed 
+Brook's framework describing the \textit{basic attacks} to the challenges of 
+developing software systems and other literature covered by the SMS on startups.
+\cite{SMS}\footnote{Note that the generalizability of results is applied only to 
+those startups that operate in software-intensive activities.}. 
+%MUN1: candidate for removal: detail on RM / seems out of place
+%In conclusion, regarding the validation of 
+%the subcategories, a continuous re-evaluation of the interview transcripts has 
+%been performed.
 
@@
 To enhance the internal validity, we created a three-dimensional research 
-framework to perform a triangulation validity procedure. By means of grounded 
-theory approach formed by a \textit{systematic mapping study}, 
-\textit{interviews} and  \textit{follow-up questionnaires} we searched for 
-convergence among different sources of information to confirm our theories.
-
-By improving and validating the theory definition, we conducted a final 
-comparison of the emergent theory with: extant literature, previously developed 
-models and evaluation of empirical data. With the final model validation we 
-highlighted and examined similarities, contrasts and explanations as discussed 
-in \cite{Eisenhardt2007}.  In this regard Eisenhardt stated: ``Tying the 
-emergent theory to existing literature enhances the internal validity, 
-generalizability, and theoretical level of the theory building from case study 
-research [\ldots] because the findings often rest on a very limited number of 
-cases.''
-
-Little ambiguities about direction of causal influences have been reduced by 
-introducing arrows to indicate the direction of the causal relation in the GSM 
-model. Furthermore, by means of comparison between the GSM model and extant 
-studies in the broader fields, we revealed that the theory resembles in wider 
-startup development context. Nevertheless, we discovered important confounding 
-factors which might mine the conclusion and generalizability of the study 
-related to innovation, creativity and market requirements.
+framework. By means of a grounded theory approach, formed by a 
+\textit{systematic mapping study}, \textit{interviews} and \textit{follow-up 
+questionnaires}, we searched for convergence among different sources of 
+information to confirm the generated theory.
+
+To validate the GSM we conducted a comparison of the emergent theory with 
+existing literature and previously developed models. Furthermore we 
+quantitatively evaluated three important relationships in the model. With the 
+theory validation we highlighted and examined similarities, contrasts and 
+explanations as discussed in \cite{Eisenhardt2007}. In this regard Eisenhardt 
+stated: ``Tying the emergent theory to existing literature enhances the internal 
+validity, generalizability, and theoretical level of the theory building from 
+a case study research [\ldots] because the findings often rest on a very limited 
+number of cases.''
+
+%MUN1: candidate for shortening
+%Little ambiguities about direction of causal influences have been reduced by 
+%introducing arrows to indicate the direction of the causal relation in the GSM 
+%model. Furthermore, by means of comparison between the GSM model and extant 
+%studies in the broader fields, we revealed that the theory resembles in wider 
+%startup development context. 
+Nevertheless, we identified important confounding factors, related to 
+innovation, market requirements and developer experience, which might 
+impact the internal validity of the GSM.
 
@@
 %Construct
-The first important threat to the construct validity of the study is a 
-possible inadequate description of constructs. In order to dismiss this risk, 
-the entire study constructs have been adapted to the terminologies utilized by 
-practitioners and defined at an adequate level for each theoretical 
-conceptualization\footnote{For instance defining \textit{Time shortage}, the 
-researchers explained it in terms of \textit{Investor pressure, CEO/business 
-pressure, Demo presentations at events and internal final deadline} used by most 
-of the interviewees during the study.}. Moreover, during the coding of interview 
-transcripts the researchers adopted wide and explanatory descriptive labels for 
-theoretical categories, to capture the underlying phenomenon without losing 
-relevant details.
-
+One threat to this study is a possible inadequate description of constructs. In 
+order to diminish this risk, the entire study constructs have been adapted to 
+the terminology utilized by practitioners and defined at an adequate level for 
+each theoretical conceptualization. For instance, we defined \textit{Time 
+shortage} in terms of \textit{Investor pressure, CEO/business 
+pressure, Demo presentations at events and internal final deadline} as used by 
+most of the interviewees in the study. 
+Moreover, during the coding of interview transcripts, we adopted explanatory 
+descriptive labels for theoretical categories, to capture the underlying 
+phenomenon without losing relevant details.
+
+%MUN1: I get the threat, but I don't understand its mitigation: let the 
+%interviewees drive the study?! Try to reformulate...
 The second important threat is caused by the fact that interviewees might 
@@
 
-Another threat associated with the construct validity is the evaluation 
-apprehension, which deals with people's rejection attitude towards the 
-assessment of their behaviour. To overcome this risk we didn't attained to first 
-answers from practitioners, but rather went towards details of engineering 
-elements and activities. Moreover, we developed a rigorous data collection 
-protocol to create case study databases and employ integration of multiple data 
-collection methods. Reporting bias was mitigated by packaging all the needed 
+Another threat associated with construct validity is evaluation 
+apprehension, which describes people's rejection attitude towards the 
+assessment of their behavior. 
+To address this risk we didn't attained to first answers from practitioners, but 
+rather went towards details of engineering elements and activities. %MUN1: <-- 
+%I don't understand this...
+Moreover, we developed a rigorous data collection protocol to create case study 
+databases and employ integration of multiple data collection methods.%MUN1: <-- 
+%Since you write "Moreover, ..." the reader gets the impression that this is a 
+%further mitigation strategy to evaluation apprehension. I don't see however 
+%how this mitigates this risk. Fix...
+Reporting bias was mitigated by packaging all the needed 
 material for conducting new researches in other contexts in order to criticize 
@@
 online\footnote{Available at 
-\url{https://github.com/adv0r/BTH-Interview-Package}.}. Moreover, two academic 
-supervisors with expertise in the area have conducted a peer-review analysis of 
-framework's constructs.
+\url{https://github.com/adv0r/BTH-Interview-Package}.}. %MUN1: --> What exactly 
+%is reporting bias? Let's assume that it means that you only report results, 
+%i.e. interviewee statements, that are favorable to your theory. How does 
+%publishing the auxiliary study material help to mitigate this risk? Publishing 
+%the raw data (what you not do) would only allow researchers to check the 
+%conclusions you made. The published auxiliary material helps to replicate the 
+%study, i.e. makes is possible to test your theory in a different context. 
+Moreover, two researchers not involved in the execution of the study conducted 
+a peer-review analysis of the model's constructs.
 
@@
 In particular we were prepared to modify generated categories so that the new 
-data could be adapted into the emerging theory according to the concepts of: 
-\textit{theoretical sampling} and \textit{theoretical saturation}.
-
-According to the \textit{theoretical sampling} concept, we adjusted our 
+data could be adapted into the emerging theory according to the concepts of 
+theoretical sampling and theoretical saturation.
+
+According to the theoretical sampling concept, we adjusted our 
 research design and the emergent theory until only marginal results were 
@@
 development, especially in the early-stage, is at the core of the company's 
-daily activities. Despite their severely high failure-rate, SMS \cite{SMS} have found that 
-the quick proliferation of startups is not supported by an adequate and 
-scientific body of knowledge. %MUN: <-- last sentence: isn't that a conclusion 
-%from the SMS? CG: fixed
+daily activities. Despite their severely high failure-rate, the SMS \cite{SMS} 
+found that the quick proliferation of startups is not supported by an adequate 
+and scientific body of knowledge. %MUN: <-- last sentence: isn't that a 
+%conclusion from the SMS? CG: fixed
 To be able to intervene on the software development strategy with scientific and 
@@
 activities in startup companies. %MUN: <-- you mix again SMS and GT here... fix 
-%it CG: not fixed, I would keep it since important part of the initial GT methodology
+%it CG: not fixed, I would keep it since important part of the initial GT 
+%methodology
+%MUN1: Ok, I understand that. But this paper is about the GT, ONLY the GT. The 
+%SMS is published elsewhere. Maybe it is ok how you use the SMS now... let's 
+%see what Tony says...
 On the other hand, the case study conducted in 13 early-stage startups, 

### 2013-03-03 article/main.tex
fixed comments version 0.4

@@
-%% bare_jrnl_compsoc.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE Computer
-%% Society journal paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
-
+%% bare_jrnl_compsoc.tex %% V1.3 %% 2007/01/11 %% by Michael Shell %% See: %%
+http://www.michaelshell.org/ %% for current contact information. %% %% This is a
+skeleton file demonstrating the use of IEEEtran.cls %% (requires IEEEtran.cls
+version 1.7 or later) with an IEEE Computer %% Society journal paper. %% %%
+Support sites: %% http://www.michaelshell.org/tex/ieeetran/ %%
+http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/ %% and %%
+http://www.ieee.org/
+
+%%************************************************************************* %%
+Legal Notice: %% This code is offered as-is without any warranty either
+expressed or %% implied; without even the implied warranty of MERCHANTABILITY or
+%% FITNESS FOR A PARTICULAR PURPOSE!  %% User assumes all risk. %% In no event
+shall IEEE or any contributor to this code be liable for %% any damages or
+losses, including, but not limited to, incidental, %% consequential, or any
+other damages, resulting from the use or misuse %% of any information contained
+here. %% %% All comments are the opinions of their respective authors and are
+not %% necessarily endorsed by the IEEE. %% %% This work is distributed under
+the LaTeX Project Public License (LPPL) %% ( http://www.latex-project.org/ )
+version 1.3, and may be freely used, %% distributed and modified. A copy of the
+LPPL, version 1.3, is included %% in the base LaTeX documentation of all
+distributions of LaTeX released %% 2003/12/01 or later. %% Retain all
+contribution notices and credits. %% ** Modified files should be clearly
+indicated as such, including  ** %% ** renaming them and changing author support
+contact information. ** %% %% File list of work: IEEEtran.cls,
+IEEEtran_HOWTO.pdf, bare_adv.tex, %%                    bare_conf.tex,
+bare_jrnl.tex, bare_jrnl_compsoc.tex
 %%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
-% The Computer Society usually requires 10pt for submissions.
-%
-\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran}
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[12pt,journal,compsoc]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-%%extra packages ------ 
-\usepackage{longtable}   
-\usepackage{float} %allows the use of [H] to force positioning of figures
-\usepackage{subfigure}
+
+% *** Authors should verify (and, if needed, correct) their LaTeX system  *** %
+*** with the testflow diagnostic prior to trusting their LaTeX platform *** %
+*** with production work. IEEE's font choices can trigger bugs that do  *** %
+*** not appear when using other class files.                            *** %
+The testflow support page is at: % http://www.michaelshell.org/tex/testflow/
+
+% Note that the a4paper option is mainly intended so that authors in % countries
+using A4 can easily print to A4 and see how their papers will % look in print -
+the typesetting of the document will not typically be % affected with changes in
+paper size (but the bottom and side margins will). % Use the testflow package
+mentioned above to verify correct handling of % both paper sizes by the user's
+LaTeX system. % % Also note that the "draftcls" or "draftclsnofoot", not
+"draft", option % should be used if it is desired that the figures are to be
+displayed in % draft mode. % % The Computer Society usually requires 10pt for
+submissions. % \documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran} % % If
+IEEEtran.cls has not been installed into the LaTeX system files, % manually
+specify the path to it like: %
+\documentclass[12pt,journal,compsoc]{../sty/IEEEtran}
+
+% Some very useful LaTeX packages include: (uncomment the ones you want to load)
+
+% *** MISC UTILITY PACKAGES *** % %\usepackage{ifpdf} % Heiko Oberdiek's
+ifpdf.sty is very useful if you need conditional % compilation based on whether
+the output is pdf or dvi. % usage: % \ifpdf %   % pdf code % \else %   % dvi
+code % \fi % The latest version of ifpdf.sty can be obtained from: %
+http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/ % Also, note that
+IEEEtran.cls V1.7 and later provides a builtin % \ifCLASSINFOpdf conditional
+that works the same way. % When switching from latex to pdflatex and vice-versa,
+the compiler may % have to be run twice to clear warning/error messages.
+
+%%extra packages ------  \usepackage{longtable}    \usepackage{float} %allows
+the use of [H] to force positioning of figures \usepackage{subfigure}
 \usepackage{hyperref} %used in references now  --- it shoud be removed
-\usepackage{amssymb}
-\usepackage[neveradjust]{paralist} %compact lists
+\usepackage{amssymb} \usepackage[neveradjust]{paralist} %compact lists
 \usepackage{amsthm}
@@
 
-% *** CITATION PACKAGES ***
-%
-\ifCLASSOPTIONcompsoc
-  % IEEE Computer Society needs nocompress option
-  % requires cite.sty v4.0 or later (November 2003)
-  % \usepackage[nocompress]{cite}
-\else
-  % normal IEEE
-  % \usepackage{cite}
-\fi
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-%
-% Note that some packages require special options to format as the Computer
-% Society requires. In particular, Computer Society  papers do not use
-% compressed citation ranges as is done in typical IEEE papers
-% (e.g., [1]-[4]). Instead, they list every citation separately in order
-% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
-% package with the nocompress option which is supported by cite.sty v4.0
-% and later. Note also the use of a CLASSOPTION conditional provided by
-% IEEEtran.cls V1.7 and later.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-%\ifCLASSINFOpdf
-   \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-   \graphicspath{{figures}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-%\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-%   \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-   %\graphicspath{{./figures/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-   %\DeclareGraphicsExtensions{.eps}
-%\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
+% *** CITATION PACKAGES *** % \ifCLASSOPTIONcompsoc   % IEEE Computer Society
+needs nocompress option   % requires cite.sty v4.0 or later (November 2003)   %
+\usepackage[nocompress]{cite} \else   % normal IEEE   % \usepackage{cite} \fi %
+cite.sty was written by Donald Arseneau % V1.6 and later of IEEEtran pre-defines
+the format of the cite.sty package % \cite{} output to follow that of IEEE.
+Loading the cite package will % result in citation numbers being automatically
+sorted and properly % "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6]
+without using % cite.sty will become [1], [2], [5]--[7], [9] using cite.sty.
+cite.sty's % \cite will automatically add leading space, if needed. Use
+cite.sty's % noadjust option (cite.sty V3.8 and later) if you want to turn this
+off. % cite.sty is already installed on most LaTeX systems. Be sure and use %
+version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does % not
+currently provide for hyperlinked citations. % The latest version can be
+obtained at: % http://www.ctan.org/tex-archive/macros/latex/contrib/cite/ % The
+documentation is contained in the cite.sty file itself. % % Note that some
+packages require special options to format as the Computer % Society requires.
+In particular, Computer Society  papers do not use % compressed citation ranges
+as is done in typical IEEE papers % (e.g., [1]-[4]). Instead, they list every
+citation separately in order % (e.g., [1], [2], [3], [4]). To get the latter we
+need to load the cite % package with the nocompress option which is supported by
+cite.sty v4.0 % and later. Note also the use of a CLASSOPTION conditional
+provided by % IEEEtran.cls V1.7 and later.
+
+% *** GRAPHICS RELATED PACKAGES *** % %\ifCLASSINFOpdf
+\usepackage[pdftex]{graphicx}   % declare the path(s) where your graphic files
+are    \graphicspath{{figures}}   % and their extensions so you won't have to
+specify these with   % every instance of \includegraphics   %
+\DeclareGraphicsExtensions{.pdf,.jpeg,.png} %\else   % or other class option
+(dvipsone, dvipdf, if not using dvips). graphicx   % will default to the driver
+specified in the system graphics.cfg if no   % driver is specified. %
+\usepackage[dvips]{graphicx}   % declare the path(s) where your graphic files
+are    %\graphicspath{{./figures/}}   % and their extensions so you won't have
+to specify these with   % every instance of \includegraphics
+%\DeclareGraphicsExtensions{.eps} %\fi % graphicx was written by David Carlisle
+and Sebastian Rahtz. It is % required if you want graphics, photos, etc.
+graphicx.sty is already % installed on most LaTeX systems. The latest version
+and documentation can % be obtained at:  % http://www.ctan.org/tex-
+archive/macros/latex/required/graphics/ % Another good source of documentation
+is "Using Imported Graphics in % LaTeX2e" by Keith Reckdahl which can be found
+as epslatex.ps or % epslatex.pdf at: http://www.ctan.org/tex-archive/info/ % %
+latex, and pdflatex in dvi mode, support graphics in encapsulated % postscript
+(.eps) format. pdflatex in pdf mode supports graphics % in .pdf, .jpeg, .png and
+.mps (metapost) formats. Users should ensure % that all non-photo figures use a
+vector format (.eps, .pdf, .mps) and % not a bitmapped formats (.jpeg, .png).
+IEEE frowns on bitmapped formats % which can result in "jaggedy"/blurry
+rendering of lines and letters as % well as large increases in file sizes. % %
+You can find documentation about the pdfTeX application at: %
+http://www.tug.org/applications/pdftex
+
+% *** MATH PACKAGES *** % %\usepackage[cmex10]{amsmath} % A popular package from
+the American Mathematical Society that provides % many useful and powerful
+commands for dealing with mathematics. If using % it, be sure to load this
+package with the cmex10 option to ensure that % only type 1 fonts will utilized
+at all point sizes. Without this option, % it is possible that some math
+symbols, particularly those within % footnotes, will be rendered in bitmap form
+which will result in a % document that can not be IEEE Xplore compliant! % %
+Also, note that the amsmath package sets \interdisplaylinepenalty to 10000 %
+thus preventing page breaks from occurring within multiline equations. Use:
+%\interdisplaylinepenalty=2500 % after loading amsmath to restore such page
+breaks as IEEEtran.cls normally % does. amsmath.sty is already installed on most
+LaTeX systems. The latest % version and documentation can be obtained at: %
+http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
+
+% *** SPECIALIZED LIST PACKAGES *** % %\usepackage{algorithmic} %
+algorithmic.sty was written by Peter Williams and Rogerio Brito. % This package
+provides an algorithmic environment fo describing algorithms. % You can use the
+algorithmic environment in-text or within a figure % environment to provide for
+a floating algorithm. Do NOT use the algorithm % floating environment provided
+by algorithm.sty (by the same authors) or % algorithm2e.sty (by Christophe
+Fiorio) as IEEE does not use dedicated % algorithm float types and packages that
+provide these will not provide % correct IEEE style captions. The latest version
+and documentation of % algorithmic.sty can be obtained at: % http://www.ctan.org
+/tex-archive/macros/latex/contrib/algorithms/ % There is also a support site at:
+% http://algorithms.berlios.de/index.html % Also of interest may be the
+(relatively newer and more customizable) % algorithmicx.sty package by Szasz
+Janos: % http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
+
+% *** ALIGNMENT PACKAGES *** % \usepackage{array} % Frank Mittelbach's and David
+Carlisle's array.sty patches and improves % the standard LaTeX2e array and
+tabular environments to provide better % appearance and additional user
+controls. As the default LaTeX2e table % generation code is lacking to the point
+of almost being broken with % respect to the quality of the end results, all
+users are strongly % advised to use an enhanced (at the very least that provided
+by array.sty) % set of table tools. array.sty is already installed on most
+systems. The % latest version and documentation can be obtained at: %
+http://www.ctan.org/tex-archive/macros/latex/required/tools/
+
+%\usepackage{mdwmath} %\usepackage{mdwtab} % Also highly recommended is Mark
+Wooding's extremely powerful MDW tools, % especially mdwmath.sty and mdwtab.sty
+which are used to format equations % and tables, respectively. The MDWtools set
+is already installed on most % LaTeX systems. The lastest version and
+documentation is available at: % http://www.ctan.org/tex-
+archive/macros/latex/contrib/mdwtools/
 
@@
 
-\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%\usepackage[tight,normalsize,sf,SF]{subfigure}
-%\else
-%\usepackage[tight,footnotesize]{subfigure}
-%\fi
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. Computer Society papers
-% use a larger font and \sffamily font for their captions, hence the
-% additional options needed under compsoc mode. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false]{caption}
-%  \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false]{caption}
-%  \usepackage[font=footnotesize]{subfig}
-%\fi
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
-% This package may be useful when used in conjunction with IEEEtran.cls'
-% captionsoff option. Some IEEE journals/societies require that submissions
-% have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.3.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% For subfigure.sty:
-% \let\MYorigsubfigure\subfigure
-% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat/subfig command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a 
-% page by themselves.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
-
-\begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Software Development in Startup Companies: The Green-field Startup Model}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-%
-%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
-% lists the Computer Society journals use for "first footnote" author
-% affiliations. Use \IEEEcompsocthanksitem which works much like \item
-% for each affiliation group. When not in compsoc mode,
-% \IEEEcompsocitemizethanks becomes like \thanks and
-% \IEEEcompsocthanksitem becomes a line break with idention. This
-% facilitates dual compilation, although admittedly the differences in the
-% desired content of \author between the different types of papers makes a
-% one-size-fits-all approach a daunting prospect. For instance, compsoc 
-% journal papers have the author affiliations above the "Manuscript
-% received ..."  text while in non-compsoc journals this is reversed. Sigh.
+\usepackage{eqparbox} % Also of notable interest is Scott Pakin's eqparbox
+package for creating % (automatically sized) equal width boxes - aka "natural
+width parboxes". % Available at: % http://www.ctan.org/tex-
+archive/macros/latex/contrib/eqparbox/
+
+% *** SUBFIGURE PACKAGES *** %\ifCLASSOPTIONcompsoc
+%\usepackage[tight,normalsize,sf,SF]{subfigure} %\else
+%\usepackage[tight,footnotesize]{subfigure} %\fi % subfigure.sty was written by
+Steven Douglas Cochran. This package makes it % easy to put subfigures in your
+figures. e.g., "Figure 1a and 1b". For IEEE % work, it is a good idea to load it
+with the tight package option to reduce % the amount of white space around the
+subfigures. Computer Society papers % use a larger font and \sffamily font for
+their captions, hence the % additional options needed under compsoc mode.
+subfigure.sty is already % installed on most LaTeX systems. The latest version
+and documentation can % be obtained at: % http://www.ctan.org/tex-
+archive/obsolete/macros/latex/contrib/subfigure/ % subfigure.sty has been
+superceeded by subfig.sty.
+
+%\ifCLASSOPTIONcompsoc %  \usepackage[caption=false]{caption} %
+\usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig} %\else %
+\usepackage[caption=false]{caption} %  \usepackage[font=footnotesize]{subfig}
+%\fi % subfig.sty, also written by Steven Douglas Cochran, is the modern %
+replacement for subfigure.sty. However, subfig.sty requires and % automatically
+loads Axel Sommerfeldt's caption.sty which will override % IEEEtran.cls handling
+of captions and this will result in nonIEEE style % figure/table captions. To
+prevent this problem, be sure and preload % caption.sty with its "caption=false"
+package option. This is will preserve % IEEEtran.cls handing of captions.
+Version 1.3 (2005/06/28) and later  % (recommended due to many improvements over
+1.2) of subfig.sty supports % the caption=false option directly:
+%\ifCLASSOPTIONcompsoc %
+\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+%\else %  \usepackage[caption=false,font=footnotesize]{subfig} %\fi % % The
+latest version and documentation can be obtained at: % http://www.ctan.org/tex-
+archive/macros/latex/contrib/subfig/ % The latest version and documentation of
+caption.sty can be obtained at: % http://www.ctan.org/tex-
+archive/macros/latex/contrib/caption/
+
+% *** FLOAT PACKAGES *** % %\usepackage{fixltx2e} % fixltx2e, the successor to
+the earlier fix2col.sty, was written by % Frank Mittelbach and David Carlisle.
+This package corrects a few problems % in the LaTeX2e kernel, the most notable
+of which is that in current % LaTeX2e releases, the ordering of single and
+double column floats is not % guaranteed to be preserved. Thus, an unpatched
+LaTeX2e can allow a % single column figure to be placed prior to an earlier
+double column % figure. The latest version and documentation can be found at: %
+http://www.ctan.org/tex-archive/macros/latex/base/
+
+%\usepackage{stfloats} % stfloats.sty was written by Sigitas Tolusis. This
+package gives LaTeX2e % the ability to do double column floats at the bottom of
+the page as well % as the top. (e.g., "\begin{figure*}[!b]" is not normally
+possible in % LaTeX2e). It also provides a command: %\fnbelowfloat % to enable
+the placement of footnotes below bottom floats (the standard % LaTeX2e kernel
+puts them above bottom floats). This is an invasive package % which rewrites
+many portions of the LaTeX2e float routines. It may not work % with other
+packages that modify the LaTeX2e float routines. The latest % version and
+documentation can be obtained at: % http://www.ctan.org/tex-
+archive/macros/latex/contrib/sttools/ % Documentation is contained in the
+stfloats.sty comments as well as in the % presfull.pdf file. Do not use the
+stfloats baselinefloat ability as IEEE % does not allow \baselineskip to
+stretch. Authors submitting work to the % IEEE should note that IEEE rarely uses
+double column equations and % that authors should try to avoid such use. Do not
+be tempted to use the % cuted.sty or midfloat.sty packages (also by Sigitas
+Tolusis) as IEEE does % not format its papers in such ways.
+
+%\ifCLASSOPTIONcaptionsoff %  \usepackage[nomarkers]{endfloat} %
+\let\MYoriglatexcaption\caption %
+\renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}} %\fi %
+endfloat.sty was written by James Darrell McCauley and Jeff Goldberg. % This
+package may be useful when used in conjunction with IEEEtran.cls' % captionsoff
+option. Some IEEE journals/societies require that submissions % have lists of
+figures/tables at the end of the paper and that % figures/tables without any
+captions are placed on a page by themselves at % the end of the document. If
+needed, the draftcls IEEEtran class option or % \CLASSINPUTbaselinestretch
+interface can be used to increase the line % spacing as well. Be sure and use
+the nomarkers option of endfloat to % prevent endfloat from "marking" where the
+figures would have been placed % in the text. The two hack lines of code above
+are a slight modification of % that suggested by in the endfloat docs (section
+8.3.1) to ensure that % the full captions always appear in the list of
+figures/tables - even if % the user used the short optional argument of
+\caption[]{}. % IEEE papers do not typically make use of \caption[]'s optional
+argument, % so this should not be an issue. A similar trick can be used to
+disable % captions of packages such as subfig.sty that lack options to turn off
+% the subcaptions: % For subfig.sty: % \let\MYorigsubfloat\subfloat %
+\renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}} % For subfigure.sty:
+% \let\MYorigsubfigure\subfigure %
+\renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}} % However, the
+above trick will not work if both optional arguments of % the \subfloat/subfig
+command are used. Furthermore, there needs to be a % description of each
+subfigure *somewhere* and endfloat does not add % subfigure captions to its list
+of figures. Thus, the best approach is to % avoid the use of subfigure captions
+(many IEEE journals avoid them anyway) % and instead reference/explain all the
+subfigures within the main caption. % The latest version of endfloat.sty and its
+documentation can obtained at: % http://www.ctan.org/tex-
+archive/macros/latex/contrib/endfloat/ % % The IEEEtran
+\ifCLASSOPTIONcaptionsoff conditional can also be used % later in the document,
+say, to conditionally put the References on a  % page by themselves.
+
+% *** PDF, URL AND HYPERLINK PACKAGES *** % \usepackage{url} % url.sty was
+written by Donald Arseneau. It provides better support for % handling and
+breaking URLs. url.sty is already installed on most LaTeX % systems. The latest
+version can be obtained at: % http://www.ctan.org/tex-
+archive/macros/latex/contrib/misc/ % Read the url.sty source comments for usage
+information. Basically, % \url{my_url_here}.
+
+% *** Do not adjust lengths that control margins, column widths, etc. *** % ***
+Do not use packages that alter fonts (such as pslatex).         *** % There
+should be no need to do such things with IEEEtran.cls V1.6 and later. % (Unless
+specifically asked to do so by the journal or conference you plan % to submit
+to, of course. )
+
+% correct bad hyphenation here \hyphenation{op-tical net-works semi-conduc-tor}
+
+\begin{document} % % paper title % can use linebreaks \\ within to get better
+formatting as desired \title{Software Development in Startup Companies: The
+Green-field Startup Model} % % % author names and IEEE memberships % note
+positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break % a
+structure at a ~ so this keeps an author's name from being broken across % two
+lines. % use \thanks{} to gain access to the first footnote area % a separate
+\thanks must be used for each paragraph as LaTeX2e's \thanks % was not built to
+handle multiple paragraphs % % %\IEEEcompsocitemizethanks is a special \thanks
+that produces the bulleted % lists the Computer Society journals use for "first
+footnote" author % affiliations. Use \IEEEcompsocthanksitem which works much
+like \item % for each affiliation group. When not in compsoc mode, %
+\IEEEcompsocitemizethanks becomes like \thanks and % \IEEEcompsocthanksitem
+becomes a line break with idention. This % facilitates dual compilation,
+although admittedly the differences in the % desired content of \author between
+the different types of papers makes a % one-size-fits-all approach a daunting
+prospect. For instance, compsoc  % journal papers have the author affiliations
+above the "Manuscript % received ..."  text while in non-compsoc journals this
+is reversed. Sigh.
 
@@
 and~...~....,~\IEEEmembership{Life~Fellow,~IEEE}% <-this % stops a space
-\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem Author info.\protect\\
-% note need leading \protect in front of \\ to get a newline within \thanks as
-% \\ is fragile and will error, could use \hfil\break instead.
-E-mail: see http://something/contact.html
-\IEEEcompsocthanksitem Author info.}% <-this % stops a space
-\thanks{}}
-
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
-
-% The paper headers
-\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
-{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer 
-Society Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-% 
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% The publisher's ID mark at the bottom of the page is less important with
-% Computer Society journal papers as those publications place the marks
-% outside of the main text columns and, therefore, unlike regular IEEE
-% journals, the available text space is not reduced by their presence.
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
-% or like this to get the Computer Society new two part style.
-%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007  
-% IEEE}%
-%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer  
-% Society\hfill}}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark (Computer Society jorunal
-% papers don't need this extra clearance.)
-
-% for Computer Society papers, we must declare the abstract and index terms
-% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
-% command as these need to go into the title area created by \maketitle.
-\IEEEcompsoctitleabstractindextext{%
-\begin{abstract}
-%\boldmath
-Software startups are newly created companies with no operating history and 
-oriented in producing cutting-edge technologies. These companies develop 
-software under highly uncertain conditions, tackling fast growing markets with 
-severe lack of resources. Despite their increasing importance in economy there 
-are only few scientific studies attempting to address software engineering (SE) 
-issues, especially for early-stage startups. This research aims to understand 
-the software development strategies conducted by practitioners, in the period of 
+\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem Author info.\protect\\ % note
+need leading \protect in front of \\ to get a newline within \thanks as % \\ is
+fragile and will error, could use \hfil\break instead. E-mail: see
+http://something/contact.html \IEEEcompsocthanksitem Author info.}% <-this %
+stops a space \thanks{}}
+
+% note the % following the last \IEEEmembership and also \thanks -  % these
+prevent an unwanted space from occurring between the last author name % and the
+end of the author line. i.e., if you had this: %  % \author{....lastname
+\thanks{...} \thanks{...} } %
+^------------^------------^----Do not want these spaces! % % a space would be
+appended to the last name and could cause every name on that % line to be
+shifted left slightly. This is one of those "LaTeX things". For % instance,
+"\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get % "AB" then you
+have to do: "\textbf{A}\textbf{B}" % \thanks is no different in this regard, so
+shield the last } of each \thanks % that ends a line with a % and do not let a
+space in before the next \thanks. % Spaces after \IEEEmembership other than the
+last one are OK (and needed) as % you are supposed to have spaces between the
+names. For what it is worth, % this is a minor point as most people would not
+even notice if the said evil % space somehow managed to creep in.
+
+% The paper headers \markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1,
+January~2007}% {Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls
+for Computer  Society Journals} % The only time the second header will appear is
+for the odd numbered pages % after the title page when using the twoside option.
+%  % *** Note that you probably will NOT want to include the author's *** % ***
+name in the headers of peer review papers.                   *** % You can use
+\ifCLASSOPTIONpeerreview for conditional compilation here if % you desire.
+
+% The publisher's ID mark at the bottom of the page is less important with %
+Computer Society journal papers as those publications place the marks % outside
+of the main text columns and, therefore, unlike regular IEEE % journals, the
+available text space is not reduced by their presence. % If you want to put a
+publisher's ID mark on the page you can do it like % this: %\IEEEpubid{0000--
+0000/00\$00.00~\copyright~2007 IEEE} % or like this to get the Computer Society
+new two part style. %\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--
+0000/00/\$00.00~\copyright~2007   % IEEE}%
+%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer   %
+Society\hfill}} % Remember, if you use this you must call \IEEEpubidadjcol in
+the second % column for its text to clear the IEEEpubid mark (Computer Society
+jorunal % papers don't need this extra clearance.)
+
+% for Computer Society papers, we must declare the abstract and index terms %
+PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran %
+command as these need to go into the title area created by \maketitle.
+\IEEEcompsoctitleabstractindextext{% \begin{abstract} %\boldmath Software
+startups are newly created companies with no operating history and  oriented in
+producing cutting-edge technologies. These companies develop  software under
+highly uncertain conditions, tackling fast growing markets with  severe lack of
+resources. Despite their increasing importance in economy there  are only few
+scientific studies attempting to address software engineering (SE)  issues,
+especially for early-stage startups. This research aims to understand  the
+software development strategies conducted by practitioners, in the period of
 time that goes from idea conception to the first open beta release. A
@@
 and product-centric engineering activities, is counterbalanced by the need of
-restructuring the product before setting off for further growth.
-\end{abstract}
-% IEEEtran.cls defaults to using nonbold math in the Abstract.
-% This preserves the distinction between vectors and scalars. However,
-% if the journal you are submitting to favors bold math in the abstract,
-% then you can use LaTeX's standard command \boldmath at the very start
-% of the abstract to achieve this. Many IEEE journals frown on math
-% in the abstract anyway. In particular, the Computer Society does
-% not want either math or citations to appear in the abstract.
+restructuring the product before setting off for further growth. \end{abstract}
+% IEEEtran.cls defaults to using nonbold math in the Abstract. % This preserves
+the distinction between vectors and scalars. However, % if the journal you are
+submitting to favors bold math in the abstract, % then you can use LaTeX's
+standard command \boldmath at the very start % of the abstract to achieve this.
+Many IEEE journals frown on math % in the abstract anyway. In particular, the
+Computer Society does % not want either math or citations to appear in the
+abstract.
 
 % Note that keywords are not normally used for peer review papers.
-\begin{keywords}
-Software Development, Startups, Grounded Theory.
+\begin{keywords} Software Development, Startups, Grounded Theory.
 \end{keywords}}
 
-% make the title area
-\maketitle
-
-% To allow for easy dual compilation without having to reenter the
-% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
-% not be used in maketitle, but will appear (i.e., to be "transported")
-% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
-% is not selected <OR> if conference mode is selected - because compsoc
-% conference papers position the abstract like regular (non-compsoc)
-% papers do!
-\IEEEdisplaynotcompsoctitleabstractindextext
-% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
-% compsoc under a non-conference mode.
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
-%------------------------------------------------------------------------------%
-\section{Introduction}
-\label{intro}
-% Computer Society journal papers do something a tad strange with the very
-% first section heading (almost always called "Introduction"). They place it
-% ABOVE the main text! IEEEtran.cls currently does not do this for you.
-% However, You can achieve this effect by making LaTeX jump through some
-% hoops via something like:
-%
-%\ifCLASSOPTIONcompsoc
-%  \noindent\raisebox{2\baselineskip}[0pt][0pt]%
-%  {\parbox{\columnwidth}{\section{Introduction}\label{sec:introduction}%
-%  \global\everypar=\everypar}}%
-%  \vspace{-1\baselineskip}\vspace{-\parskip}\par
-%\else
-%  \section{Introduction}\label{sec:introduction}\par
-%\fi
-%
-% Admittedly, this is a hack and may well be fragile, but seems to do the
-% trick for me. Note the need to keep any \label that may be used right
-% after \section in the above as the hack puts \section within a raised box.
-
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps (small caps for compsoc).
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-\IEEEPARstart{A}{n} impressive number of new software startups are launched 
-worldwide every day as a result of an increase of new markets, accessible 
-technologies, and venture capital \cite{8491286}.  With the term 
-\textit{software startups} we refer to those temporary organizations focused on 
-the creation of high-tech and innovative products, with little or no operating 
-history, aiming to grow by  aggressively scaling their business in highly 
-scalable markets.
+% make the title area \maketitle
+
+% To allow for easy dual compilation without having to reenter the %
+abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will % not
+be used in maketitle, but will appear (i.e., to be "transported") % here as
+\IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode % is not selected
+<OR> if conference mode is selected - because compsoc % conference papers
+position the abstract like regular (non-compsoc) % papers do!
+\IEEEdisplaynotcompsoctitleabstractindextext %
+\IEEEdisplaynotcompsoctitleabstractindextext has no effect when using % compsoc
+under a non-conference mode.
+
+% For peer review papers, you can put extra information on the cover % page as
+needed: % \ifCLASSOPTIONpeerreview % \begin{center} \bfseries EDICS Category:
+3-BBND \end{center} % \fi % % For peerreview papers, this IEEEtran command
+inserts a page break and % creates the second title. It will be ignored for
+other modes. \IEEEpeerreviewmaketitle
+
+%------------------------------------------------------------------------------%
+\section{Introduction} \label{intro} % Computer Society journal papers do
+something a tad strange with the very % first section heading (almost always
+called "Introduction"). They place it % ABOVE the main text! IEEEtran.cls
+currently does not do this for you. % However, You can achieve this effect by
+making LaTeX jump through some % hoops via something like: %
+%\ifCLASSOPTIONcompsoc %  \noindent\raisebox{2\baselineskip}[0pt][0pt]% %
+{\parbox{\columnwidth}{\section{Introduction}\label{sec:introduction}% %
+\global\everypar=\everypar}}% %  \vspace{-1\baselineskip}\vspace{-\parskip}\par
+%\else %  \section{Introduction}\label{sec:introduction}\par %\fi % %
+Admittedly, this is a hack and may well be fragile, but seems to do the % trick
+for me. Note the need to keep any \label that may be used right % after \section
+in the above as the hack puts \section within a raised box.
+
+% The very first letter is a 2 line initial drop letter followed % by the rest
+of the first word in caps (small caps for compsoc). %  % form to use if the
+first word consists of a single letter: % \IEEEPARstart{A}{demo} file is .... %
+% form to use if you need the single drop letter followed by % normal text
+(unknown if ever used by IEEE): % \IEEEPARstart{A}{}demo file is .... %  % Some
+journals put the first two words in caps: % \IEEEPARstart{T}{his demo} file is
+.... %  % Here we have the typical use of a "T" for an initial drop letter % and
+"HIS" in caps to complete the first word. \IEEEPARstart{A}{n} impressive number
+of new software startups are launched  worldwide every day as a result of an
+increase of new markets, accessible  technologies, and venture capital
+\cite{8491286}.  With the term  \textit{software startups} we refer to those
+temporary organizations focused on  the creation of high-tech and innovative
+products, with little or no operating  history, aiming to grow by  aggressively
+scaling their business in highly  scalable markets.
 
@@
 decisions and avoiding choices that could easily lead the whole business to
-failure \cite{Kajko-Mattsson2008}.
-However, despite the impressive size of the startup ecosystem 
-\cite{ISI:000243253000007}, the state-of-the-art presents a gap. Through a 
-\textit{Systematic Mapping Study} (SMS) of the literature, conducted by 
-Paternoster et al. \cite{SMS}, only few publications were found related to 
-engineering activities in startups. Moreover the majority of these studies do 
-not possess the relevance and rigour required to form a consistent body of 
-knowledge in the state-of-the-art and in the state-of-practice.
+failure \cite{Kajko-Mattsson2008}. However, despite the impressive size of the
+startup ecosystem  \cite{ISI:000243253000007}, the state-of-the-art presents a
+gap. Through a  \textit{Systematic Mapping Study} (SMS) of the literature,
+conducted by  Paternoster et al. \cite{SMS}, only few publications were found
+related to  engineering activities in startups. Moreover the majority of these
+studies do  not possess the relevance and rigour required to form a consistent
+body of  knowledge in the state-of-the-art and in the state-of-practice.
 
@@
 
-With a cross-sectional study we focused the research context in a limited 
-time-frame to emphasize the nature of uncertainty in the development activities 
-and time pressure from the market, as discussed in \cite{Maccormack2001, 
+With a cross-sectional study we focused the research context in a limited  time-
+frame to emphasize the nature of uncertainty in the development activities  and
+time pressure from the market, as discussed in \cite{Maccormack2001,
 Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and
@@
 the research focused at studying technical debt \cite{Nugroho2011,Izurieta2012}.
-On the other hand, researchers can take advantage from the GSM to 
-understand how the technical debt influences the future growth of startup 
-companies.
+On the other hand, researchers can take advantage from the GSM to  understand
+how the technical debt influences the future growth of startup  companies.
 
@@
 research process. Results are discussed in section \ref{res:gsm}, which presents
-the GSM. Section \ref{res:val}  compares results of this research to 
-literature and frameworks presented in this field. Section \ref{valt} discusses 
-the validity threats considered during this study. Section 
-\ref{sect:theory:impl} discusses the most relevant implications of the GSM. 
-Finally, this paper concludes in section \ref{conc}.
+the GSM. Section \ref{res:val}  compares results of this research to  literature
+and frameworks presented in this field. Section \ref{valt} discusses  the
+validity threats considered during this study. Section  \ref{sect:theory:impl}
+discusses the most relevant implications of the GSM.  Finally, this paper
+concludes in section \ref{conc}.
 
@@
 %reviewers: The references you use in the introduction are not very recent.
-%Look at the list of papers [1-15], most of them are published before 2005.
-%I understand that you use some references to strengthen your general arguments 
+%Look at the list of papers [1-15], most of them are published before 2005. I
+%understand that you use some references to strengthen your general arguments
 %(which can refer to older publications), but your introduction should also
-%show the relevance of your research right now. This is best shown by 
-%referring to recent (max 5 years old) literature. Think about if you can add 
-%more recent (2010-now), well published, papers in the introduction.
-
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
+%show the relevance of your research right now. This is best shown by  referring
+%to recent (max 5 years old) literature. Think about if you can add  more recent
+%(2010-now), well published, papers in the introduction.
+
+% You must have at least 2 lines in the paragraph with the drop letter (should
+% never be an issue)
 
@@
 %------------------------------------------------------------------------------%
-\section{Background}
-\label{backg}
+\section{Background} \label{backg}
 
 Modern entrepreneurship has been boosted by the advent of the consumer internet
-markets in the middle of the 1990s and culminated with the notorious 
-\textit{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126, 
+markets in the middle of the 1990s and culminated with the notorious  \textit
+{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126,
 storey1982entrepreneurship}. Several years later, with the massification of the
@@
 %------------------------------------------------------------------------------%
-\subsection{General lack of research in startups}
-Despite some studies tried to address the above mentioned issues, a SMS 
-\cite{SMS} revealed only a few SE works in this specific area, as confirmed by 
-other studies \cite{Coleman2008, Coleman2008a, Coleman2007, Sutton2000}. 
-Moreover, the studies, identified in \cite{SMS}, appear to be highly fragmented 
-and spread across different areas rather than constituting a consistent 
-\textit{body of knowledge}.
+\subsection{General lack of research in startups} Despite some studies tried to
+address the above mentioned issues, a SMS  \cite{SMS} revealed only a few SE
+works in this specific area, as confirmed by  other studies \cite{Coleman2008,
+Coleman2008a, Coleman2007, Sutton2000}.  Moreover, the studies, identified in
+\cite{SMS}, appear to be highly fragmented  and spread across different areas
+rather than constituting a consistent  \textit{body of knowledge}.
 
@@
 1994 in an article written by Carmel \cite{Camel1994a} where he studied the
-time-to-completion in a young package firm. %MUN: If this is a quote 
-%from Carmel, put it in quotation marks, like you did in the previous section 
-%with Suttons quote. Generally, check that you are consistent how you use 
-%quotations throughout the paper. 
-%CG: fixed, btw  it was just to emphasize his the focus of his study. 
-%MUN1: Ok. As you may have noticed, I have removed a lot of cases were you use 
-%italics. If you emphasize *too* much, it it rather distracting than helpful 
-%for the reader. Furthermore, it conflicts with you convention you use to 
-%highlight high level categories of GSM.
-He noticed how these companies were particularly innovative and successful, 
-advocating for the need of more research investigating software development 
-practices so as to replicate success and try to transfer it to other technology 
-sectors.
+time-to-completion in a young package firm. %MUN: If this is a quote  %from
+Carmel, put it in quotation marks, like you did in the previous section  %with
+Suttons quote. Generally, check that you are consistent how you use  %quotations
+throughout the paper.  %CG: fixed, btw  it was just to emphasize his the focus
+of his study.  %MUN1: Ok. As you may have noticed, I have removed a lot of cases
+were you use  %italics. If you emphasize *too* much, it it rather distracting
+than helpful  %for the reader. Furthermore, it conflicts with you convention you
+use to  %highlight high level categories of GSM. CG: ok. He noticed how these
+companies were particularly innovative and successful,  advocating for the need
+of more research investigating software development  practices so as to
+replicate success and try to transfer it to other technology  sectors.
 
@@
 ground running''} \cite{Yoffie1999}. Empowering the team and focus on
-methodological attributes of the processes oriented in prototyping, 
-proof-of-concepts, mock-ups and demos, to test basic functionalities, have been 
-the primary priority in startups \cite{Camel1994a}. Only when they grow, formal 
+methodological attributes of the processes oriented in prototyping,  proof-of-
+concepts, mock-ups and demos, to test basic functionalities, have been  the
+primary priority in startups \cite{Camel1994a}. Only when they grow, formal
 methodologies arise, followed by performing quality assurance and long-term
@@
 Additionally, the role of SPI has been neglected because seen as an obstacle to
-the development team's creativity and flexibility, and to 
-the need of a quick delivering product process environment \cite{Coleman2008a}.
-%MUN: the second part of the sentence is just a statement without reference. If 
-%it refers to a statement by Coleman, put his reference at the end OR find the 
-%correct reference OR delete the statement.
-%CG fixed
-Product quality is left aside in favor of minimal and suitable functionalities 
-to shorten the time-to-market. As reported in two studies of Mater and Mirel 
-\cite{Mater2000, Mirel2000}, quality aspects, mostly taken in consideration in 
-internet startups, are oriented to usability and scalability, even though the 
-market and application type heavily influences the quality-demand 
-\cite{Coleman2008, Kim2005}.
+the development team's creativity and flexibility, and to  the need of a quick
+delivering product process environment \cite{Coleman2008a}. %MUN: the second
+part of the sentence is just a statement without reference. If  %it refers to a
+statement by Coleman, put his reference at the end OR find the  %correct
+reference OR delete the statement. %CG fixed Product quality is left aside in
+favor of minimal and suitable functionalities  to shorten the time-to-market. As
+reported in two studies of Mater and Mirel  \cite{Mater2000, Mirel2000}, quality
+aspects, mostly taken in consideration in  internet startups, are oriented to
+usability and scalability, even though the  market and application type heavily
+influences the quality-demand  \cite{Coleman2008, Kim2005}.
 
@@
 \cite{Hanna2010}, software reuse \cite{Jansen2008} and open-source strategies ,
-\cite{Wall2001, Bean2005}. 
-%MUN: Can you split up the references? I.e. write:
+\cite{Wall2001, Bean2005}.  %MUN: Can you split up the references? I.e. write:
 %"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source
-%strategies \cite{Bla}.
-%CG: fixed
-%------------------------------------------------------------------------------%
-\subsection{Technical debt}
-Notwithstanding, a new area of the SE research, trying to tackle the problem of 
-the \textit{technical debt},  brings and encompasses different implications in 
-studying development in software startups \cite{Tom2013}. The metaphoric 
-neologism was originally introduced by Cunningham in 1992 
-\cite{TechnicalDebtCunn} and has recently attracted the attention of SE 
-researchers\footnote{To attest the fact that \textit{technical debt} is gaining 
-traction among researchers, we mention two important contributions which 
-characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} 
-and a dedicated workshop \cite{workshopDebt} organized by the Software 
-Engineering Institute  and ICSE.}. The concept of \textit{technical debt} is 
-well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea 
-is that developers sometimes accept compromises in a system in one dimension 
-(e.g., modularity) to meet an urgent demand in some other dimension (e.g., a 
-deadline), and that such compromises incur a ``debt'' on which ``interest'' has 
-to be paid and which the ``principal'' should be repaid at some point for the 
-long-term health of the project''}.  The compromise between high-speed and 
-high-quality engineering is faced daily by software startups, not only in terms 
-of architecture design but in multifaceted aspects (weak project management, 
-testing, process control).
+%strategies \cite{Bla}. %CG: fixed
+%------------------------------------------------------------------------------%
+\subsection{Technical debt} Notwithstanding, a new area of the SE research,
+trying to tackle the problem of  the \textit{technical debt},  brings and
+encompasses different implications in  studying development in software startups
+\cite{Tom2013}. The metaphoric  neologism was originally introduced by
+Cunningham in 1992  \cite{TechnicalDebtCunn} and has recently attracted the
+attention of SE  researchers\footnote{To attest the fact that \textit{technical
+debt} is gaining  traction among researchers, we mention two important
+contributions which  characterize the \textit{``debt landscape''} :
+\cite{Nugroho2011,Izurieta2012}  and a dedicated workshop \cite{workshopDebt}
+organized by the Software  Engineering Institute  and ICSE.}. The concept of
+\textit{technical debt} is  well illustrated in
+\cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea  is that developers
+sometimes accept compromises in a system in one dimension  (e.g., modularity) to
+meet an urgent demand in some other dimension (e.g., a  deadline), and that such
+compromises incur a ``debt'' on which ``interest'' has  to be paid and which the
+``principal'' should be repaid at some point for the  long-term health of the
+project''}.  The compromise between high-speed and  high-quality engineering is
+faced daily by software startups, not only in terms  of architecture design but
+in multifaceted aspects (weak project management,  testing, process control).
 
@@
 
-\subsection{Terminology}
-In order to set a common ground and to prevent ambiguity, we use the following 
-terminology throughout the paper:
-
-\begin{compactitem}
-\item Software development strategy: the overall approach adopted by the 
-company to carry out the product development.
-\item Engineering activities: the activities needed to bring a product from 
-idea to market. Traditional engineering activities are, among others, 
-requirement engineering, design, architecture, implementation, testing.
-\item Engineering elements: any practice, tool and artifacts contributing and 
-supporting the engineering activities.
-\item Quality attributes: those overall factors that affect run-time behavior, 
-system design, and user experience. They represent areas of concern that have 
-the potential for applications to impact across layers and tiers. Some of these 
-attributes are related to the overall system design, while others are specific 
-to run time, design time, or user centric issues \cite{Microsoft2009}.
-\item Growth: an increase in the company size respect to the initial conditions 
-in terms of either employees or users/customers, and product complexity in terms 
-of handling an increasing number of feature requests.
-\item Software product: any software product and software service.
+\subsection{Terminology} In order to set a common ground and to prevent
+ambiguity, we use the following  terminology throughout the paper:
+
+\begin{compactitem} \item Software development strategy: the overall approach
+adopted by the  company to carry out the product development. \item Engineering
+activities: the activities needed to bring a product from  idea to market.
+Traditional engineering activities are, among others,  requirement engineering,
+design, architecture, implementation, testing. \item Engineering elements: any
+practice, tool and artifacts contributing and  supporting the engineering
+activities. \item Quality attributes: those overall factors that affect run-time
+behavior,  system design, and user experience. They represent areas of concern
+that have  the potential for applications to impact across layers and tiers.
+Some of these  attributes are related to the overall system design, while others
+are specific  to run time, design time, or user centric issues
+\cite{Microsoft2009}. \item Growth: an increase in the company size respect to
+the initial conditions  in terms of either employees or users/customers, and
+product complexity in terms  of handling an increasing number of feature
+requests. \item Software product: any software product and software service.
 \end{compactitem}
 %------------------------------------------------------------------------------%
-\section{Research methodology}
-\label{resmet}
-
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=5in]{figures/completemethodology2}
-\caption{Grounded theory study process overview}
-\label{fig:gt:completemethodology}
-\end{figure*}
+\section{Research methodology} \label{resmet}
+
+\begin{figure*}[!t] \centering
+\includegraphics[width=5in]{figures/completemethodology2} \caption{Grounded
+theory study process overview} \label{fig:gt:completemethodology} \end{figure*}
 %MUN: I think you can increase the font size in the figure a bit. Try thereby
@@
 %you name it     in the text. Also, add a section (3.5) that describes shortly
-%what you did in "Theory  Validation" 
-%CG: fixed... fontsize cannot be increased uniformly among all the boxes, then 
-%I skipped that part...
-%MUN1: OK, but in the figure, the correct section number for theory validation 
-%is 3.5, not 3.4
+%what you did in "Theory  Validation"  %CG: fixed... fontsize cannot be
+increased uniformly among all the boxes, then  %I skipped that part... %MUN1:
+OK, but in the figure, the correct section number for theory validation  %is
+3.5, not 3.4. CG:fixed
 
@@
 
-Initially the boundaries of the research domain were set by means of 
-non-systematic literature survey, which provided the initial keywords used 
-further in the research process. Moreover it helped us in defining the research 
+Initially the boundaries of the research domain were set by means of  non-
+systematic literature survey, which provided the initial keywords used  further
+in the research process. Moreover it helped us in defining the research
 questions (RQs):
@@
 \item RQ-1 : What is the current state-of-practice related to software
-development strategy in early-stage startups?
-\begin{compactitem}
+development strategy in early-stage startups? \begin{compactitem}
 
 \item RQ-1.1 : How do startups structure and execute the main engineering
-activities?
-\item RQ-1.2 : How are product quality attributes considered by startups?
-\end{compactitem}
+activities? \item RQ-1.2 : How are product quality attributes considered by
+startups? \end{compactitem}
 
@@
 by the first two authors, handling conflicts by reviewing the rationale of
-decision with the third and fourth authors. When necessary we performed an 
-in-depth review of the design of the research and data collected during the 
+decision with the third and fourth authors. When necessary we performed an  in-
+depth review of the design of the research and data collected during the
 execution process.
@@
 %------------------------------------------------------------------------------%
-\subsection{Design and Execution}
-\label{desex}
+\subsection{Design and Execution} \label{desex}
 
@@
 
-\begin{table*}[!t]
-\renewcommand{\arraystretch}{1.3}
-\caption{Companies overview} 
-\label{t_interviews-stats}
-\centering
-\begin{tabular}{|c||c||c||c||c||c|}
-
-\hline
-    ID  &  Company age & Founding & Initial & Current & First product  \\
+\begin{table*}[!t] renewcommand{arraystretch}{1.3} caption{Companies overview}
+\label{t_interviews-stats} centering begin{tabular}{|c||c||c||c||c||c|}
+
+\hline     ID  &  Company age & Founding & Initial & Current & First product  \\
 &   (months)   & team   & developers  & employees  &  building time \\
-                    &      & &   &  & (months)   \\
-   \hline
-C1    & 11    & 4     & 2     & 11    & 6   \\
-C2    & 5     & 2     & 2     & 6     & 3      \\
-C3    & 18    & 4     & 4     & 4     & 6      \\
-C4    & 17    & 3     & 2     & 11    & 6     \\
-C5    & 20    & 2     & 1     & 4     & 12   \\
-C6    & 30    & 3     & 2     & 4     & 1      \\
-C7    & 12    & 2     & 1     & 7     & 4      \\
-C8    & 24    & 4     & 3     & 16    & 4    \\
-C9    & 5     & 5     & 4     & 5     & 1     \\
-C10   & 43    & 6     & 4     & 9     & 4      \\
-C11   & 36    & 3     & 3     & 6     & 2     \\
-C12   & 12    & 3     & 3     & 3     & 3      \\
-C13   & 24    & 2     & 2     & 20    & 3     \\
-
-\hline
-\end{tabular}
-\end{table*}
+&      & &   &  & (months)   \\    \hline C1    & 11    & 4     & 2     & 11
+& 6   \\ C2    & 5     & 2     & 2     & 6     & 3      \\ C3    & 18    & 4
+& 4     & 4     & 6      \\ C4    & 17    & 3     & 2     & 11    & 6     \\ C5
+& 20    & 2     & 1     & 4     & 12   \\ C6    & 30    & 3     & 2     & 4
+& 1      \\ C7    & 12    & 2     & 1     & 7     & 4      \\ C8    & 24    & 4
+& 3     & 16    & 4    \\ C9    & 5     & 5     & 4     & 5     & 1     \\ C10
+& 43    & 6     & 4     & 9     & 4      \\ C11   & 36    & 3     & 3     & 6
+& 2     \\ C12   & 12    & 3     & 3     & 3     & 3      \\ C13   & 24    & 2
+& 2     & 20    & 3     \\
+
+\hline end{tabular} end{table*}
 
@@
 %MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct?
-%CG: right
-by an average of 3 starting members, which were in majority developers. 
-Moreover, the number of current employees shows how, to different degrees, 
-companies expanded the initial teams. All companies, except C5, brought their 
-first product to market in less than 6 months from the idea conception. The 
-products consist of web-applications, launched in different nations and market 
-sectors. The growing team size and the publicly available data suggest an 
-healthy status of the businesses.
+%CG: right by an average of 3 starting members, which were in majority
+developers.  Moreover, the number of current employees shows how, to different
+degrees,  companies expanded the initial teams. All companies, except C5,
+brought their  first product to market in less than 6 months from the idea
+conception. The  products consist of web-applications, launched in different
+nations and market  sectors. The growing team size and the publicly available
+data suggest an  healthy status of the businesses.
 
@@
 questionnaires have been collected within one day after the conclusion of the
-last interview.   
-%MUN: I don't understand this: you
-%collected the data from questionnaires after ALL interviews have been 
-%finished, but on the other hand you said the overall process was iterative 
-%(data analysis and triangulation let themes emerge which in turn were used
-%in subsequent interviews). 
-%CG: fixed...
-
-%------------------------------------------------------------------------------%
-\subsection{Data collection}
-The data collection was designed to allow triangulation, which integrates 
-multiple data sources converging on the same phenomenon. After transcribing the 
-interview, we extracted conceptual relations, integrating them with the 
-questionnaire's results. A well structured case-study database allowed us to 
-easily retrieve and seek for information, assembling the evidence from the case 
-study reports, as described also in \cite{Yin1994}. The database has been stored 
-and constructed using the qualitative data analysis software package 
-AtlasTI\footnote{Available online at \url{http://www.atlasti.com/}.} - one of 
-the most suitable tools for GT \cite{Coleman2007}. 
-We overlapped interviews with questionnaire results to adjust the data 
-collection and take advantage of emergent themes and reveal possible 
-inconsistencies. The data was analyzed simultaneously and with flexibility in 
-mind in such a way that adjustments were made according to the emerging 
-findings.
-%------------------------------------------------------------------------------%
-\subsection{Data analysis}
-Before starting the analysis, a data ordering procedure was necessary since 
-interviews were spread across a multitude of topics. Therefore transcripts have 
-been structured in thematic areas accordingly to different topic cards used 
-during the interviews. We proceeded horizontally between same thematic areas of 
-transcripts to be able to identify a better number of similar concepts, rather 
-than going through an entire transcript at one time.
+last interview.    %MUN: I don't understand this: you %collected the data from
+questionnaires after ALL interviews have been  %finished, but on the other hand
+you said the overall process was iterative  %(data analysis and triangulation
+let themes emerge which in turn were used %in subsequent interviews).  %CG:
+fixed...
+
+%------------------------------------------------------------------------------%
+\subsection{Data collection} The data collection was designed to allow
+triangulation, which integrates  multiple data sources converging on the same
+phenomenon. After transcribing the  interview, we extracted conceptual
+relations, integrating them with the  questionnaire's results. A well structured
+case-study database allowed us to  easily retrieve and seek for information,
+assembling the evidence from the case  study reports, as described also in
+\cite{Yin1994}. The database has been stored  and constructed using the
+qualitative data analysis software package  AtlasTI\footnote{Available online at
+\url{http://www.atlasti.com/}.} - one of  the most suitable tools for GT
+\cite{Coleman2007}.  We overlapped interviews with questionnaire results to
+adjust the data  collection and take advantage of emergent themes and reveal
+possible  inconsistencies. The data was analyzed simultaneously and with
+flexibility in  mind in such a way that adjustments were made according to the
+emerging  findings.
+%------------------------------------------------------------------------------%
+\subsection{Data analysis} Before starting the analysis, a data ordering
+procedure was necessary since  interviews were spread across a multitude of
+topics. Therefore transcripts have  been structured in thematic areas
+accordingly to different topic cards used  during the interviews. We proceeded
+horizontally between same thematic areas of  transcripts to be able to identify
+a better number of similar concepts, rather  than going through an entire
+transcript at one time.
 
@@
 \item Labels were assigned to raw data, and a first low-level conceptualization
-was carried out using both in-vivo and open coding \cite{ColinRobson2009}.
-\item Concepts were grouped together into theoretical categories and 
-subcategories. By means of axial coding we first described the different 
-relations between subcategories, and then relations between subcategories and 
-categories.
-\item Categories were refined several times in the attempt to create different 
-levels of abstraction and adjusting concepts, aided by a simple knowledge 
-management tool.
-\item Consistency among categories was validated by exploring and analyzing 
-links among subcategories by means of selective coding.
-\item The core category - the one with the greatest explanatory power - was 
-identified by analyzing the causal relations between high-level categories.
-\end{compactitem}
+was carried out using both in-vivo and open coding \cite{ColinRobson2009}. \item
+Concepts were grouped together into theoretical categories and  subcategories.
+By means of axial coding we first described the different  relations between
+subcategories, and then relations between subcategories and  categories. \item
+Categories were refined several times in the attempt to create different  levels
+of abstraction and adjusting concepts, aided by a simple knowledge  management
+tool. \item Consistency among categories was validated by exploring and
+analyzing  links among subcategories by means of selective coding. \item The
+core category - the one with the greatest explanatory power - was  identified by
+analyzing the causal relations between high-level categories. \end{compactitem}
 
@@
 
-\begin{compactitem}
-\item \textit{Causal conditions}: the events which lead to the occurrence of 
-the phenomenon, that is our core category
-\item \textit{Context}: set of conditions in which the phenomenon can be 
-extrapolated
+\begin{compactitem} \item \textit{Causal conditions}: the events which lead to
+the occurrence of  the phenomenon, that is our core category \item
+\textit{Context}: set of conditions in which the phenomenon can be  extrapolated
 \item \textit{Intervening conditions}: the broader set of conditions from that
-the phenomenon can be generalized
-\item \textit{Action/interaction strategies}: the actions and responses that 
-occur as the result of the phenomenon
-\item \textit{Consequences}: specification of the outcomes, both intended and 
-unintended of the actions and interaction strategies
-\end{compactitem}
+the phenomenon can be generalized \item \textit{Action/interaction strategies}:
+the actions and responses that  occur as the result of the phenomenon \item
+\textit{Consequences}: specification of the outcomes, both intended and
+unintended of the actions and interaction strategies \end{compactitem}
 
@@
 %------------------------------------------------------------------------------%
-\subsection{Theory Validation}
-\label{rm:val}
-%MUN: put here what I wrote in the comment in Section "Theory validation" (i.e.
-%the section following the GSM)
-%CG: fixed
-%MUN1: This description is way too long and verbose. It should be MAX 1 column, 
-%and focus on what you present in Section 5. Theory Validation. Until the 
-%second-last paragraph you don't talk about that at all! Remove the "theory" on 
-%GT validation and what you did *during* the GT process. Keep what you need 
-%to motivate
-% - 5.1 Comparison of frameworks
-% - 5.2 Theoretical categories and existing literature
-% - 5.3 Confounding factors
-% - 5.4 Static assessment of the GSM
-%You need to explain why you did this and how it contributes to the validation 
-%of the theory.
+\subsection{Theory Validation} \label{rm:val} %MUN: put here what I wrote in the
+comment in Section "Theory validation" (i.e. %the section following the GSM)
+%CG: fixed %MUN1: This description is way too long and verbose. It should be MAX
+1 column,  %and focus on what you present in Section 5. Theory Validation. Until
+the  %second-last paragraph you don't talk about that at all! Remove the
+"theory" on  %GT validation and what you did *during* the GT process. Keep what
+you need  %to motivate % - 5.1 Comparison of frameworks % - 5.2 Theoretical
+categories and existing literature % - 5.3 Confounding factors % - 5.4 Static
+assessment of the GSM %You need to explain why you did this and how it
+contributes to the validation  %of the theory. CG: fixed, removing the initial
+verbose part.
 
@@
 and determine if the GT is sufficiently grounded, we used a systematic technique
-to assess the theory, following evaluation lists and factors presented in  
-\cite{Strauss1998}.
-
-We consider the following factors:
-
-\par\vspace {0.4cm}
-\begin{compactitem}
-\item \textit{Fit}, that describes the consistency of the theory with the real 
-data.
-\item \textit{Understanding}, that describes the degree to which participants 
-agree with the theory.
-\item \textit{Generality}, that describes the level of abstraction of the 
-theory, and how sufficiently it can be adopted by practitioners without losing 
-its relevance.
-\item \textit{Control}, that describes how well the theory enables 
-practitioners to fully understand the described situation.
-\end{compactitem}
-\par\vspace {0.4cm}
-
-In order to achieve a high-level of fit between the theory and the raw 
-empirical data, we started GT without any assumption on how the emergent 
-patterns would influence the research direction. Ideas and perceptions evolved 
-during the mapping study and interviews. In addition, constant comparative 
-methods, overturning of some early categories as new data came to light, 
-generation and testing of interim hypotheses, and constant re-evaluation of the 
-interview transcripts, ensured that researchers' bias was minimized and 
-theoretical fit maintained.
-
-To represent understandable emergent theories, concepts and categories were 
-carefully developed using appropriate and self-explanatory codes. The developing 
-theory was not immediately presented to the participants to prevent potential 
-bias in the responses. But rather, the theoretical framework has been presented 
-and described to startups practitioners not involved in the data collection, 
-whose reaction was positively expressed.
-
-From generality perspective, we developed a theory from abstract categories, 
-which allowed us to create a general guide to constantly changing situations. 
-Nevertheless, we took attention on maintaining a reasonable level of 
-abstraction, keeping focus on their sensitizing characteristics and providing 
-low-level subgroup of categories.
-
-Finally, we provided a comprehensive set of categories with detailed 
-interrelations to explain how the \textit{core-category} is achieved and the 
-reasons of its consequences, in order to enable practitioners to independently 
-understand and analyze situations, predict changes and their consequences, and 
-to be capable of revising actions or the theory itself. Using both 
-methodological tools and features provided by \textit{AtlasTI}, we explored and 
-tested:
-
-\par\vspace {0.4cm}
-\begin{compactitem}
-\item Each category and the strength of relations between them.
-\item Hypotheses, derived from and related to the emergent theory.
-\item Deviant cases to ensure robustness and general applicability.
-\end{compactitem}
-\par\vspace {0.4cm}
-
-Additionally, Strauss and Corbin provided a list of questions to assist in 
-determining how well the findings are grounded \cite{Strauss1998}. They 
-questions are as follows:
-
-\par\vspace {0.4cm}
-\begin{compactenum}
-\item Are concepts generated, and are the concepts systematically related?
-\item Are there many conceptual linkages and are the categories well developed?
-\item Is variation\footnote{Variation refers to the variety of contexts to 
-which the theory can be applied.} built into the theory and are the conditions 
-under which variation can be found built into the study and explained?
-\item Are the conditions under which variation can be found built into the 
-study and explained?
-\item Has the process been taken into account, and does the theory stand the 
-test of time?
-\item Do the theoretical findings seem significant, and to what extent?
-\end{compactenum}
+to assess the theory.
+
+Strauss and Corbin provided a list of questions to assist in  determining how
+well the findings are grounded \cite{Strauss1998}. They  questions are as
+follows:
+
+\par\vspace {0.4cm} \begin{compactenum} \item Are concepts generated, and are
+the concepts systematically related? \item Are there many conceptual linkages
+and are the categories well developed? \item Is variation\footnote{Variation
+refers to the variety of contexts to  which the theory can be applied.} built
+into the theory and are the conditions  under which variation can be found built
+into the study and explained? \item Are the conditions under which variation can
+be found built into the  study and explained? \item Has the process been taken
+into account, and does the theory stand the  test of time? \item Do the
+theoretical findings seem significant, and to what extent? \end{compactenum}
 \par\vspace {0.4cm}
@@
 diagram (see Figure \ref{fig:gt:completemethodology}), by which we established
-the relations between the concepts. At each iteration of the 
-grounded theory process, we considered and examined a concept within different 
-conditions and dimensions, trying to incorporate data from a broader range of 
-practitioners.
+the relations between the concepts. At each iteration of the  grounded theory
+process, we considered and examined a concept within different  conditions and
+dimensions, trying to incorporate data from a broader range of  practitioners.
 
@@
 \textit{comparison with the state-of-the-art} has been applied in order to
-validate the theory and enhance odds of its applicability within a wider 
-time-frame. By this comparison we  highlight the areas which have been neglected 
-by existing studies, providing possible directions for future studies (see 
+validate the theory and enhance odds of its applicability within a wider  time-
+frame. By this comparison we  highlight the areas which have been neglected  by
+existing studies, providing possible directions for future studies (see
 subsection \ref{sect:theory:validation:others}). Moreover we study the
@@
 
-%------------------------------------------------------------------------------%
-
-\section{Results: Green-field startup model}
-\label{res:gsm}
+\section{Results: Green-field startup model} label{res:gsm}
 
@@
 
-The GSM captures the underlying phenomenon of software development in 
-early-stage startups. The framework is formed by 128 sub-categories, clustered 
-in 35 groups, and finally in 7 categories at the highest level of abstraction. 
-By the means of the GSM we provide explanations of the development strategies 
-and engineering activities undertaken by startups.
-%------------------------------------------------------------------------------%
-\subsection{Framework overview}
-\label{res:gsm:frmov}
-The main concepts representing the underlying phenomenon have been grouped 
-together to form high-level categories. Figure \ref{fig:gsm} shows the network 
-of causal relationships (represented by arrows) between categories (represented 
-by blocks). Each node is linked to subsequent nodes which are called successors, 
+The GSM captures the underlying phenomenon of software development in  early-
+stage startups. The framework is formed by 128 sub-categories, clustered  in 35
+groups, and finally in 7 categories at the highest level of abstraction.  By the
+means of the GSM we provide explanations of the development strategies  and
+engineering activities undertaken by startups.
+%------------------------------------------------------------------------------%
+\subsection{Framework overview} \label{res:gsm:frmov} The main concepts
+representing the underlying phenomenon have been grouped  together to form high-
+level categories. Figure \ref{fig:gsm} shows the network  of causal
+relationships (represented by arrows) between categories (represented  by
+blocks). Each node is linked to subsequent nodes which are called successors,
 and preceding nodes, called predecessors. The relations between them are denoted
@@
 
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=4in]{figures/high-level}
-\caption{Green-field startup model}\label{fig:gsm}
-\end{figure*}
+\begin{figure*}[!t] centering includegraphics[width=4in]{figures/high-level}
+\caption{Green-field startup model}label{fig:gsm} end{figure*}
 
@@
 in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e.
-CATx) to facilitate the reading of this section.
-The network is centered around the core category, \textit{speed-up 
-development}, which is the most interconnected node in the framework reflecting 
-the fact that \textit{``it is the one (category) with the greatest explanatory 
-power''}  \cite{Strauss1998}. The meaning of the labels on the three arrows that 
-reach the core category (efficiency, effectiveness and performance) is discussed 
-in detail in section \ref{res:val}. %MUN: this refers to the validity threats 
-%section. Do you mean rather the "Theory Validation" section?
-%CG:fixed
-%MUN1: this should be \ref{sect:theory:validation:stats}, right? Section 5 is 
-%quite long and you should point to the correct subsection
+CATx) to facilitate the reading of this section. The network is centered around
+the core category, \textit{speed-up  development}, which is the most
+interconnected node in the framework reflecting  the fact that \textit{``it is
+the one (category) with the greatest explanatory  power''}  \cite{Strauss1998}.
+The meaning of the labels on the three arrows that  reach the core category
+(efficiency, effectiveness and performance) is discussed  in detail in section
+\ref{sect:theory:validation:stats}. %MUN: this refers to the validity threats
+%section. Do you mean rather the "Theory Validation" section? %CG:fixed %MUN1:
+this should be \ref{sect:theory:validation:stats}, right? Section 5 is  %quite
+long and you should point to the correct subsection. CG: fixed.
 
@@
 urgent needs\footnote{There are some exceptions where the quality aspects
-actually matter and such cases will be discussed in section \ref{res:val}%MUN: 
-%this refers to the validity threats section. Do you mean rather the "Theory 
-%Validation" section?
-%CG fixed
-%MUN1: Same here, point to the correct subsection.
-.}. At the same time, to be able to deal with such constraints, startups must 
-depend on a small group of capable and motivated individuals.
+actually matter and such cases will be discussed in section \ref{sect:an:comp
+:cat-vs-literature:conf}%MUN:  %this refers to the validity threats section. Do
+you mean rather the "Theory  %Validation" section? %CG fixed %MUN1: Same here,
+point to the correct subsection. CG:fixed .}. At the same time, to be able to
+deal with such constraints, startups must  depend on a small group of capable
+and motivated individuals.
 
@@
 performance in software development. To support a fast-paced production
-environment, engineers are required to be highly committed, co-located, 
-multi-role, and self-organized. In other words \textit{team is the catalyst of 
-development}.
-With an essential and flexible workflow, which relies on tacit knowledge 
-instead of formal documentation, startups are able to achieve very short 
-time-to-market cycles. However, each line of code, written without following 
-structures and processes, contributes to grow the \textit{accumulated technical 
-debt}, which is further increased by having almost non-existing specifications, 
-a minimal project management and a lack of automated tests.
+environment, engineers are required to be highly committed, co-located,  multi-
+role, and self-organized. In other words \textit{team is the catalyst of
+development}. With an essential and flexible workflow, which relies on tacit
+knowledge  instead of formal documentation, startups are able to achieve very
+short  time-to-market cycles. However, each line of code, written without
+following  structures and processes, contributes to grow the \textit{accumulated
+technical  debt}, which is further increased by having almost non-existing
+specifications,  a minimal project management and a lack of automated tests.
 
@@
 %------------------------------------------------------------------------------%
-\subsection{Severe lack of resources}
-\label{res:gsm:cat7}
-The concept of \textit{severe lack of resources} characterizes the uncertainty 
-of the development strategies in startups and it is composed by three 
-subcategories: \textit{time-shortage}, \textit{limited human resources} and 
-\textit{limited access to expertise}.
+\subsection{Severe lack of resources} \label{res:gsm:cat7} The concept of
+\textit{severe lack of resources} characterizes the uncertainty  of the
+development strategies in startups and it is composed by three  subcategories:
+\textit{time-shortage}, \textit{limited human resources} and  \textit{limited
+access to expertise}.
 
@@
 %------------------------------------------------------------------------------%
-\subsection{Team is the catalyst of development}
-\label{res:gsm:cat4}
-Among the different aspects fostering the speed of the development process, the 
-startups' main focus is on the characteristics of the initial team.
+\subsection{Team is the catalyst of development} \label{res:gsm:cat4} Among the
+different aspects fostering the speed of the development process, the  startups'
+main focus is on the characteristics of the initial team.
 
@@
 
-Early engineers in the founding team of a startups are typically 
-\textit{multi-role and full-stack engineers}: they are full-stack for being able 
-to tackle different problems at different levels of the technology stack 
+Early engineers in the founding team of a startups are typically  \textit{multi-
+role and full-stack engineers}: they are full-stack for being able  to tackle
+different problems at different levels of the technology stack
 (\textit{generalists developers instead of specialists}) and multi-role since
@@
 under constant pressure}, \textit{``up to 16 hours per day''}, %MUN: add Cx to
-%this statement since it is a citation? 
-%CG: not fixed. It represents a subcategory, expressed by the 70% of the 
-%interviewed startups
-%MUN1: and how should the reader know that? See my problem with italics: you 
-%use italics for both interview statements and (sub-)categories. 
-mainly constrained by a tight \textit{time shortage}.
+%this statement since it is a citation?  %CG: not fixed. It represents a
+subcategory, expressed by the 70% of the  %interviewed startups %MUN1: and how
+should the reader know that? See my problem with italics: you  %use italics for
+both interview statements and (sub-)categories. CG: skip as acked mainly
+constrained by a tight \textit{time shortage}.
 
@@
 %------------------------------------------------------------------------------%
-\subsection{Evolutionary approach}
-\label{res:gsm:cat2}
-Among different approaches to development, startups prefer to build an initial 
-prototype and iteratively refine it over time, similarly to the concept of 
+\subsection{Evolutionary approach} \label{res:gsm:cat2} Among different
+approaches to development, startups prefer to build an initial  prototype and
+iteratively refine it over time, similarly to the concept of
 \textit{``evolutionary prototyping''} \cite{EvProt}. The reason of using this
@@
 
-The objective of this evolutionary approach is to avoid wasting time in 
-``over-engineering the system'' and building complex functionalities that have 
-not been tested on real users. By releasing a very small number of good-enough 
+The objective of this evolutionary approach is to avoid wasting time in  ``over-
+engineering the system'' and building complex functionalities that have  not
+been tested on real users. By releasing a very small number of good-enough
 functionalities (see CAT3) the startup can verify the suitability of the
@@
 %------------------------------------------------------------------------------%
-\subsection{Product quality has low priority}
-\label{res:gsm:cat3}
-The interests of software startups, related to the product, are mainly 
-concentrated on building a  \textit{limited number of suitable functionalities} 
-rather than fulfilling restrictive non-functional requirements. This strategy 
-allows them to quickly release simple products with less need of preliminary 
-architectural studies.
+\subsection{Product quality has low priority} \label{res:gsm:cat3} The interests
+of software startups, related to the product, are mainly  concentrated on
+building a  \textit{limited number of suitable functionalities}  rather than
+fulfilling restrictive non-functional requirements. This strategy  allows them
+to quickly release simple products with less need of preliminary  architectural
+studies.
 
@@
 achieved. Product metrics are typically web-based \textit{statistical hypothesis
-testing}, such as A/B testing \cite{AB}. 
-%MUN: add reference or explain shortly what A/B testing is
-%CG: fixed
+testing}, such as A/B testing \cite{AB}.  %MUN: add reference or explain shortly
+what A/B testing is %CG: fixed
 
@@
 effectiveness of the minimal set of functionalities, obtained according to CAT2
-and to the concept of MVP.
-\item The \textit{product should be reasonably ready-to-scale} to be able to 
-accommodate a potential growth of the user-base\footnote{Note that startups 
-tackle fast growing markets which are particularly subject to sudden user 
-growths}. Thanks to modern cloud services, startups can \textit{externalize 
-complexity to third party solutions} achieving a good level of scalability with 
-a reasonable effort.
-\item Realizing high reliability is not an urgent priority since \textit{users 
-are fault-tolerant in innovative beta products}. In these cases, users have 
-typically a positive attitude towards the product, even though it presents 
-unreliable behaviours. In this regard, the main focus of beta testing is 
-reducing frictions between the product and the users, often incorporating 
-usability testing. In fact, the beta release is typically the first time that 
-the software is available outside of the organization that developed it. 
-\footnote{Detailed discussion of the impact of innovative products on the user 
-satisfaction is presented in subsection \ref{sect:an:comp:cat-vs-literature:conf}}.
-%MUN: do you use the Kano model in Theory Validation? Change text to refer to 
-%the correct section. [After I've read the respective section, I'd suggest to 
-%remove the Kano discussion altogether to shorten the paper. Keep the footnote, 
-%reference the correct section, but remove "according to the Kano model]
-%CG: fixed
-\end{compactitem}
+and to the concept of MVP. \item The \textit{product should be reasonably ready-
+to-scale} to be able to  accommodate a potential growth of the user-
+base\footnote{Note that startups  tackle fast growing markets which are
+particularly subject to sudden user  growths}. Thanks to modern cloud services,
+startups can \textit{externalize  complexity to third party solutions} achieving
+a good level of scalability with  a reasonable effort. \item Realizing high
+reliability is not an urgent priority since \textit{users  are fault-tolerant in
+innovative beta products}. In these cases, users have  typically a positive
+attitude towards the product, even though it presents  unreliable behaviours. In
+this regard, the main focus of beta testing is  reducing frictions between the
+product and the users, often incorporating  usability testing. In fact, the beta
+release is typically the first time that  the software is available outside of
+the organization that developed it.  \footnote{Detailed discussion of the impact
+of innovative products on the user  satisfaction is presented in subsection
+\ref{sect:an:comp:cat-vs-literature:conf}}. %MUN: do you use the Kano model in
+Theory Validation? Change text to refer to  %the correct section. [After I've
+read the respective section, I'd suggest to  %remove the Kano discussion
+altogether to shorten the paper. Keep the footnote,  %reference the correct
+section, but remove "according to the Kano model] %CG: fixed \end{compactitem}
 
@@
 %------------------------------------------------------------------------------%
-\subsection{Speed-up development}
-\label{res:gsm:cat1}
-As remarked initially, \textit{speed-up development} represents the core 
-category of the GSM. Firmly grounded as the primary objective of startups, it 
-shows the most important characteristic of developing software in the early 
-stage.
+\subsection{Speed-up development} \label{res:gsm:cat1} As remarked initially,
+\textit{speed-up development} represents the core  category of the GSM. Firmly
+grounded as the primary objective of startups, it  shows the most important
+characteristic of developing software in the early  stage.
 
@@
 %------------------------------------------------------------------------------%
-\subsection{Accumulated technical debt}
-\label{res:gsm:cat5}
-Startups achieve a very high development speed by radically ignoring aspects 
-related to documentation, structures and processes. However, although these 
-aspects are not considered important in the very first stages, they will become 
+\subsection{Accumulated technical debt} \label{res:gsm:cat5} Startups achieve a
+very high development speed by radically ignoring aspects  related to
+documentation, structures and processes. However, although these  aspects are
+not considered important in the very first stages, they will become
 increasingly more important for the productivity in the long-term, as we
@@
 modularization and frameworks from day 1} and \textit{describing critical
-interactions with third-party components only}. In particular, the use of 
-well-known standards, frameworks and conventions limits the need of formal UML 
+interactions with third-party components only}. In particular, the use of  well-
+known standards, frameworks and conventions limits the need of formal UML
 \cite{UML} diagrams and documentation, and provide a minimum level of
@@
 develop faster''}). Startups' \textit{minimal project management}  is supported
-by keeping: \textit{internal milestones short and informal}, 
-\textit{low-precision task assignment mechanism}  and a very low attention for 
-project metrics (paraphrasing C13, \textit{``the only track of progress was made 
-by looking at closed tickets''}). In this context \textit{only a final release 
+by keeping: \textit{internal milestones short and informal},  \textit{low-
+precision task assignment mechanism}  and a very low attention for  project
+metrics (paraphrasing C13, \textit{``the only track of progress was made  by
+looking at closed tickets''}). In this context \textit{only a final release
 milestone is viable}, which helps practitioners to remain focused on short term
@@
 the {use of standard/known technologies} with a \textit{simple workflow} are the
-main reasons for 
-not establishing a heavy project management process since the project and 
-technologies can be autonomously managed by the development team (with 
-characteristics of: co-location, self-organization and very small team size). 
-%MUN: do you mean small team size?
-%CG: fixed
+main reasons for  not establishing a heavy project management process since the
+project and  technologies can be autonomously managed by the development team
+(with  characteristics of: co-location, self-organization and very small team
+size).  %MUN: do you mean small team size? %CG: fixed
 
@@
 There is no time for it. Instead, there is a huge pay off in having a code that
-is understandable itself.''}
-This approach is supported by the fact that startups make use of  
-\textit{simple and informal workflow}, \textit{standard/known technologies}, 
-\textit{very small and co-located development team} with \textit{limited need of 
-formalities}.
-%------------------------------------------------------------------------------%
-\subsection{Initial growth hinders performance}
-\label{res:gsm:cat6}
-The lack of attention given in the first phases to engineering activities 
-allows startups to ship code extremely quickly (see CAT5). However, if 
-successful, the initial product becomes more and more complex over time, the 
-number of users increases and the company size starts to grow. Under these 
-circumstances the need of controlling the initial chaos forces the development 
-team to return the \textit{accumulated technical debt} instead of focusing on 
-new users' requests. Hence, \textit{initial growth hinders performance} in terms 
-of number of new user stories implemented in a certain amount of time, which 
-brings new suitable functionalities to the users.
+is understandable itself.''} This approach is supported by the fact that
+startups make use of   \textit{simple and informal workflow},
+\textit{standard/known technologies},  \textit{very small and co-located
+development team} with \textit{limited need of  formalities}.
+%------------------------------------------------------------------------------%
+\subsection{Initial growth hinders performance} \label{res:gsm:cat6} The lack of
+attention given in the first phases to engineering activities  allows startups
+to ship code extremely quickly (see CAT5). However, if  successful, the initial
+product becomes more and more complex over time, the  number of users increases
+and the company size starts to grow. Under these  circumstances the need of
+controlling the initial chaos forces the development  team to return the
+\textit{accumulated technical debt} instead of focusing on  new users' requests.
+Hence, \textit{initial growth hinders performance} in terms  of number of new
+user stories implemented in a certain amount of time, which  brings new suitable
+functionalities to the users.
 
@@
 %------------------------------------------------------------------------------%
-\subsection{Theory}
-\label{res:gsm:th}
-In order to explain and understand the development strategies in early-stage 
-software startups we construct the theory generated and supported by the above 
-presented GSM:
-
-\begin{theory}
-Focusing on limited number of suitable functionalities, and adopting partial 
-and rapid evolutionary development approaches, early-stage software startups 
-operate at high development speed, aided by skilled and highly co-located 
-developers. Through these development strategies, early-stage software startups 
-aim to find early product/market fit within extremely uncertain conditions and 
-severe lack of resources. Nevertheless, by speeding-up the development process, 
-they accumulate technical debt, causing an initial and temporary drop-down in 
-performance before setting off for further growth.
+\subsection{Theory} \label{res:gsm:th} In order to explain and understand the
+development strategies in early-stage  software startups we construct the theory
+generated and supported by the above  presented GSM:
+
+\begin{theory} Focusing on limited number of suitable functionalities, and
+adopting partial  and rapid evolutionary development approaches, early-stage
+software startups  operate at high development speed, aided by skilled and
+highly co-located  developers. Through these development strategies, early-stage
+software startups  aim to find early product/market fit within extremely
+uncertain conditions and  severe lack of resources. Nevertheless, by speeding-up
+the development process,  they accumulate technical debt, causing an initial and
+temporary drop-down in  performance before setting off for further growth.
 \end{theory}
@@
 As discussed in section \ref{res:val} %MUN: this refers to the validity threats
-%section. Do you mean rather the "Theory Validation" section?
-%CG:fixed
-, we formed the theory by applying the paradigm model, introduced by Corbin 
+%section. Do you mean rather the "Theory Validation" section? %CG:fixed , we
+formed the theory by applying the paradigm model, introduced by Corbin
 \cite{Corbin1990}. The theory is composed by different elements reflecting the
@@
 categories: \textit{product quality has low priority}, \textit{evolutionary
-approach} and \textit{team is the catalyst of development}.
-\item \textit{Phenomenon} is represented by the core category \textit{speed-up 
-development}.
-\item \textit{Context} is limited to web early-stage software startups 
-operating in conditions of severe lack of resources aiming to early find 
-product/market fit.
-\item \textit{Intervening conditions} are  summarized by the extremely 
-uncertain development environment.
-\item \textit{Action and interaction strategies} are represented by the 
-accumulation of technical debt.
+approach} and \textit{team is the catalyst of development}. \item
+\textit{Phenomenon} is represented by the core category \textit{speed-up
+development}. \item \textit{Context} is limited to web early-stage software
+startups  operating in conditions of severe lack of resources aiming to early
+find  product/market fit. \item \textit{Intervening conditions} are  summarized
+by the extremely  uncertain development environment. \item \textit{Action and
+interaction strategies} are represented by the  accumulation of technical debt.
 \item \textit{Consequences} lead to a temporary performance drop-down.
@@
 %------------------------------------------------------------------------------%
-\section{Theory Validation}
-\label{res:val}
-
-In this section we discuss the validity of the GSM by means of 
-cross-methodological observations, as discussed in section \ref{rm:val}.
-%MUN: expand this discussion a bit (motivation why you do this like you do it) 
-%and move to Section 3.5. Then write here just that you perform theory 
-%validation as discussed in Section 3.5 (and remove the bullet list below).
-%CG:fixed
-%MUN1: Ok, but fix section 3.5... you see that you don't motivate 5.1-5.4 at 
-%all there and that the current content of 3.5 is not related to what you 
-%actually do here as validation?
+\section{Theory Validation} \label{res:val}
+
+In this section we discuss the validity of the GSM by means of  cross-
+methodological observations, as discussed in section \ref{rm:val}. %MUN: expand
+this discussion a bit (motivation why you do this like you do it)  %and move to
+Section 3.5. Then write here just that you perform theory  %validation as
+discussed in Section 3.5 (and remove the bullet list below). %CG:fixed %MUN1:
+Ok, but fix section 3.5... you see that you don't motivate 5.1-5.4 at  %all
+there and that the current content of 3.5 is not related to what you  %actually
+do here as validation? CG: skip as acked
 
@@
 later stage, representing also companies in expansion phase with more than 100
-employees. %MUN: so he did not do research 
-%in a startup context? You need to clarify what you mean by "more mature 
-%companies"... I assume you still mean startups, but startups at a later stage, 
-%i.e. after first product release.
+employees. %MUN: so he did not do research  %in a startup context? You need to
+clarify what you mean by "more mature  %companies"... I assume you still mean
+startups, but startups at a later stage,  %i.e. after first product release.
 %CG: fixed
@@
 wanted it to work for us, for our own size and scale''} \cite{Coleman2008}.
-%MUN: add ref to Coleman's paper where he states this
-%CG fixed
+%MUN: add ref to Coleman's paper where he states this %CG fixed
 
@@
 is not a matter of poor knowledge and training, but rather it is the necessity
-of operating with solutions that let the company move faster. 
-\textit{``One-size-fits-all''} solutions have always found difficulty in 
-penetrating small software organizations \cite{Staples2007}. When startups began 
-establishing any SPI process, they experienced \textit{process erosions}  
-\cite{Coleman2008} %MUN: 
-%add ref (Coleman?). You should add, for each category you use from Coleman, 
-%add the reference to the paper. This makes it clear that the category comes 
-%from him (and is not one of our categories)
-%CG: fixed
-, which resolved to a barely sufficient workflow to satisfy the organizational 
-business needs.
+of operating with solutions that let the company move faster.  \textit{``One-
+size-fits-all''} solutions have always found difficulty in  penetrating small
+software organizations \cite{Staples2007}. When startups began  establishing any
+SPI process, they experienced \textit{process erosions}   \cite{Coleman2008}
+%MUN:  %add ref (Coleman?). You should add, for each category you use from
+Coleman,  %add the reference to the paper. This makes it clear that the category
+comes  %from him (and is not one of our categories) %CG: fixed , which resolved
+to a barely sufficient workflow to satisfy the organizational  business needs.
 
@@
 %do not produce heavy bureaucracy, but they allow fast communication
-%strategies... I may be wrong, but then you need to explain it better
-%CG:fixed, you are right :) ...  
-, even though the accumulated tacit knowledge is hard to manage and to transfer 
-to new hires as also discussed in category \textit{CAT6}.
+%strategies... I may be wrong, but then you need to explain it better %CG:fixed,
+you are right :) ...   , even though the accumulated tacit knowledge is hard to
+manage and to transfer  to new hires as also discussed in category
+\textit{CAT6}.
 
@@
 found that despite Coleman's study contributes to define which factors hinder
-the introduction of SPI in startups 
-%MUN: add: "of SPI in startups"? CG: fixed
+the introduction of SPI in startups  %MUN: add: "of SPI in startups"? CG: fixed
 his results are coherent with the theory generated within our research, focusing
 at explaining the overall strategies within software development in startups.
-%MUN: add "our research, focused at explaining software development in
-%general in startups." ? CG:fixed
-In this regard, a wide variability of processes is a key factor for startups.  
-In contrast with some SPI processes, such as \textit{Six Sigma} which objective 
-is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, 
-startups seem to follow the \textit{Fischer's Fundamental Theorem of natural  
-selection} \cite{Fisher}.
+%MUN: add "our research, focused at explaining software development in %general
+in startups." ? CG:fixed In this regard, a wide variability of processes is a
+key factor for startups.   In contrast with some SPI processes, such as
+\textit{Six Sigma} which objective  is minimizing unpredictability in the
+definition of workflows \cite{Sixsigma},  startups seem to follow the
+\textit{Fischer's Fundamental Theorem of natural   selection} \cite{Fisher}.
 Moving towards flexible and variable processes (easy  to adapt to the uncertain
@@
 of resources, %MUN: under which  %circumstances? This statement sounds too
-%general, i.e. don't do SPI ever... %CG: fixed 
-%And the following part,I did not fully understand. Try to clarify. CG:fixed  
-not following SPI does not predict
-major adaptiveness of startups development towards the customer's needs.
+%general, i.e. don't do SPI ever... %CG: fixed  %And the following part,I did
+not fully understand. Try to clarify. CG:fixed   not following SPI does not
+predict major adaptiveness of startups development towards the customer's needs.
 Nevertheless, to imply a major evolutionary product/market fit, certain
@@
 
-\item A desperate rush-to-market.
-\item A new and unique software market environment.
-\item A lack of experience developing software under the conditions this 
-environment imposed.
+\item A desperate rush-to-market. \item A new and unique software market
+environment. \item A lack of experience developing software under the conditions
+this  environment imposed.
 
@@
 %software development - MUN: put this sentence where you introduce
-%Baskerville's work. %CG: fixed
-, Baskerville revealed similar causal factors as the GSM.
+%Baskerville's work. %CG: fixed , Baskerville revealed similar causal factors as
+the GSM.
 
@@
 
-\begin{compactitem}
-\item Speed-up development by releasing more often the software and 
-``implanting'' customers in the development environment (similar to the 
-\textit{find the product/market fit} concept presented in CAT2). 
-\item Make heavy use of simple tools and existing components (similar to the 
-concepts expressed in CAT1).
-\item Invest time in facilitating development of scalable systems by the use of 
-simple but stable architectural solutions (see subcategory \textit{using 
-modularization and frameworks from day 1} discussed in CAT5). 
-\item Tailor the development process daily according to the intense demands for 
-speed. In this regard, he argues that the trend is to skip phases or tasks that 
-might impede the ability to deliver software quickly even though producing lower 
-quality software.
+\begin{compactitem} \item Speed-up development by releasing more often the
+software and  ``implanting'' customers in the development environment (similar
+to the  \textit{find the product/market fit} concept presented in CAT2).  \item
+Make heavy use of simple tools and existing components (similar to the  concepts
+expressed in CAT1). \item Invest time in facilitating development of scalable
+systems by the use of  simple but stable architectural solutions (see
+subcategory \textit{using  modularization and frameworks from day 1} discussed
+in CAT5).  \item Tailor the development process daily according to the intense
+demands for  speed. In this regard, he argues that the trend is to skip phases
+or tasks that  might impede the ability to deliver software quickly even though
+producing lower  quality software.
 
@@
 already implemented. It is the main strategy, which enables startups to
-externalize complexity to third party solutions explained within CAT1,  
-\textit{speed-up development}.
-\item Requirements refinement and rapid prototyping- avoid deciding precisely 
-what to build but rather iteratively extract and refine the product requirements 
-with customers and users. It represents the evolutionary development approach to 
-maintain a fast release-cycles during development (CAT2).
-\item Incremental development- starting from simple solutions allows 
-organizations to early prototype and control complexity overtime. It represents 
-the main purpose of focusing on \textit{limited number of suitable 
-functionalities} examined in \textit{product quality has low priority}  category 
-(CAT3).
-\item Great team- people are the center of a software project and it is 
-profoundly important to empower and liberate their creative mind. It is the main 
-objective of empowering developers' capabilities described in \textit{team is 
-the catalyst of development} (CAT4).
+externalize complexity to third party solutions explained within CAT1,   \textit
+{speed-up development}. \item Requirements refinement and rapid prototyping-
+avoid deciding precisely  what to build but rather iteratively extract and
+refine the product requirements  with customers and users. It represents the
+evolutionary development approach to  maintain a fast release-cycles during
+development (CAT2). \item Incremental development- starting from simple
+solutions allows  organizations to early prototype and control complexity
+overtime. It represents  the main purpose of focusing on \textit{limited number
+of suitable  functionalities} examined in \textit{product quality has low
+priority}  category  (CAT3). \item Great team- people are the center of a
+software project and it is  profoundly important to empower and liberate their
+creative mind. It is the main  objective of empowering developers' capabilities
+described in \textit{team is  the catalyst of development} (CAT4).
 \end{compactitem}
@@
 
-\begin{table*}[!t]
-\renewcommand{\arraystretch}{1.3}
-\caption{SMS \cite{SMS} and GSM} 
-\label{tab:an:literature_comp}
-\centering
+\begin{table*}[!t] renewcommand{arraystretch}{1.3} caption{SMS cite{SMS} and
+\GSM}  label{tab:an:literature_comp} centering
 \begin{tabular}{|l||c||c||c||c||c||c||c||c||c|}
 
-\hline
-Author (year) & CAT1 & CAT2 & CAT3 & CAT4 & CAT5 & CAT6 & CAT7 & Count & Ref. \\
-\hline
-Sutton (2000) & X & X & X & X & X & X & X & 7 & \cite{Sutton2000} \\
-Kajko-Mattson (2008) & X & X & X & X & X & X & X & 7 & 
-\cite{Kajko-Mattsson2008} \\
-Crowne (2002) & X & X & X & X & X & X & X & 7 & \cite{Crowne2002} \\
-Coleman (2008) & X & X & X & X & X & X & X & 7 & \cite{Coleman2008a} \\
-Coleman (2008) & X & X & X & X & X & X & X & 7 & \cite{Coleman2008} \\
-Coleman (2007) & X & X & X & X & X & X & X & 7 & \cite{Coleman2007} \\
-Carmel (1994) & X & X & X & X & X & X & X & 7 & \cite{Camel1994a} \\
-Yoffie (1999) & X & X & X & X & X & X & & 6 & \cite{Yoffie1999} \\
-Zettel (2001) & X & X & X & X & & & X & 5 & \cite{Zettel2001} \\
-Jansen (2008) & X & & X & & X & X & X & 5 & \cite{Jansen2008} \\
-Heitlager (2007) &  &  & X & X & X & X & X & 5 & \cite{Heitlager2007} \\
-Deias (2002) & X & X & & X & X & X & & 5 & \cite{Deias} \\
-Ambler (2002) & X & X &  &  & X & X & X & 5 & \cite{Ambler2002} \\
-Wood (2005) & X &  & X  & X &  &  & X & 4 & \cite{Wood2005} \\
-Tingling (2007) &  & X &  & X & X & X &  & 4 & \cite{Tingling2007} \\
-Taipale (2010) & X & X &  & X & X &  &  & 4 & \cite{Taipale2010} \\
-da Silva (2005) & X & X &  & X & X &  &  & 4 & \cite{Silva2005} \\
-Mirel (2000) & X &  & X & X & X &  &  & 4 & \cite{Mirel2000} \\
-Midler (2008) & X & X &  & X &  &  & X & 4 & \cite{Midler2008} \\
-Tanabian (2005) & X &  &  & X &  &  & X  & 3 & \cite{Tanabian2005} \\
-Stanfill (2007) & X &  &  &  & X &  & X & 3 & \cite{Stanfill2007} \\
-Mater (2000) & X & X &  &  &  & X &  & 3 & \cite{Mater2000} \\
-Kuvinka (2011) & X & X &  & X &  &  &  & 3 & \cite{Kuvinka2011} \\
-Deakins (2005) & X & X &  & X &  &  &  & 3 & \cite{Deakins2005} \\
-Yogendra (2002) & X & X &  &  &  &  &  & 2 & \cite{Yogendra2002} \\
-Wall (2001) & X &  &  &  &  &  & X & 2 & \cite{Wall2001} \\
-Su-Chuang (2007) & X &  &  & X &  &  &  & 2 & \cite{Su-Chan2007} \\
-Steenhuis (2008) &  &  &  & X &  &  & X & 2 & \cite{Steenhuis2008} \\
-Sau-ling Lai (2010) &  & X &  &  &  &  & X & 2 & \cite{Lai2010} \\
-Kakati (2003) &  & X &  & X &  &  &  & 2 & \cite{Kakati2003} \\
-Himola (2003) & X &  &  &  &  &  & X & 2 & \cite{Hilmola2003} \\
-H\"{a}sel (2010) & X &  &  & X &  &  &  & 2 & \cite{Hasel2010} \\
-Hanna (2010) & X &  &  &  &  &  & X  & 2 & \cite{Hanna2010} \\
-Bean (2005) & X &  &  &  & X &  &  & 2 & \cite{Bean2005} \\
-Kim (2005) &  & X &  &  &  &  &  & 1 & \cite{Kim2005} \\
-Fayad (1997) &  &  &  & X &  &  &  & 1 & \cite{Fayad1997} \\
-Chorev (2006) &  &  &  & X &  &  &  & 1 & \cite{Chorev2006} \\
-\hline
-Count & 29 & 22 & 13 & 26 & 18 & 14 & 20 & &\\
-%MUN: add a row which shows the sum of each column. This shows how much 
-%categories are covered overall %CG: fixed
-\hline
-\end{tabular}
-\end{table*}
+\hline Author (year) & CAT1 & CAT2 & CAT3 & CAT4 & CAT5 & CAT6 & CAT7 & Count &
+Ref. \\ \hline Sutton (2000) & X & X & X & X & X & X & X & 7 & \cite{Sutton2000}
+\\ Kajko-Mattson (2008) & X & X & X & X & X & X & X & 7 &  \cite{Kajko-
+Mattsson2008} \\ Crowne (2002) & X & X & X & X & X & X & X & 7 &
+\cite{Crowne2002} \\ Coleman (2008) & X & X & X & X & X & X & X & 7 &
+\cite{Coleman2008a} \\ Coleman (2008) & X & X & X & X & X & X & X & 7 &
+\cite{Coleman2008} \\ Coleman (2007) & X & X & X & X & X & X & X & 7 &
+\cite{Coleman2007} \\ Carmel (1994) & X & X & X & X & X & X & X & 7 &
+\cite{Camel1994a} \\ Yoffie (1999) & X & X & X & X & X & X & & 6 &
+\cite{Yoffie1999} \\ Zettel (2001) & X & X & X & X & & & X & 5 &
+\cite{Zettel2001} \\ Jansen (2008) & X & & X & & X & X & X & 5 &
+\cite{Jansen2008} \\ Heitlager (2007) &  &  & X & X & X & X & X & 5 &
+\cite{Heitlager2007} \\ Deias (2002) & X & X & & X & X & X & & 5 & \cite{Deias}
+\\ Ambler (2002) & X & X &  &  & X & X & X & 5 & \cite{Ambler2002} \\ Wood
+(2005) & X &  & X  & X &  &  & X & 4 & \cite{Wood2005} \\ Tingling (2007) &  & X
+&  & X & X & X &  & 4 & \cite{Tingling2007} \\ Taipale (2010) & X & X &  & X & X
+&  &  & 4 & \cite{Taipale2010} \\ da Silva (2005) & X & X &  & X & X &  &  & 4 &
+\cite{Silva2005} \\ Mirel (2000) & X &  & X & X & X &  &  & 4 & \cite{Mirel2000}
+\\ Midler (2008) & X & X &  & X &  &  & X & 4 & \cite{Midler2008} \\ Tanabian
+(2005) & X &  &  & X &  &  & X  & 3 & \cite{Tanabian2005} \\ Stanfill (2007) & X
+&  &  &  & X &  & X & 3 & \cite{Stanfill2007} \\ Mater (2000) & X & X &  &  &  &
+X &  & 3 & \cite{Mater2000} \\ Kuvinka (2011) & X & X &  & X &  &  &  & 3 &
+\cite{Kuvinka2011} \\ Deakins (2005) & X & X &  & X &  &  &  & 3 &
+\cite{Deakins2005} \\ Yogendra (2002) & X & X &  &  &  &  &  & 2 &
+\cite{Yogendra2002} \\ Wall (2001) & X &  &  &  &  &  & X & 2 & \cite{Wall2001}
+\\ Su-Chuang (2007) & X &  &  & X &  &  &  & 2 & \cite{Su-Chan2007} \\ Steenhuis
+(2008) &  &  &  & X &  &  & X & 2 & \cite{Steenhuis2008} \\ Sau-ling Lai (2010)
+&  & X &  &  &  &  & X & 2 & \cite{Lai2010} \\ Kakati (2003) &  & X &  & X &  &
+&  & 2 & \cite{Kakati2003} \\ Himola (2003) & X &  &  &  &  &  & X & 2 &
+\cite{Hilmola2003} \\ H\"{a}sel (2010) & X &  &  & X &  &  &  & 2 &
+\cite{Hasel2010} \\ Hanna (2010) & X &  &  &  &  &  & X  & 2 & \cite{Hanna2010}
+\\ Bean (2005) & X &  &  &  & X &  &  & 2 & \cite{Bean2005} \\ Kim (2005) &  & X
+&  &  &  &  &  & 1 & \cite{Kim2005} \\ Fayad (1997) &  &  &  & X &  &  &  & 1 &
+\cite{Fayad1997} \\ Chorev (2006) &  &  &  & X &  &  &  & 1 & \cite{Chorev2006}
+\\ \hline Count & 29 & 22 & 13 & 26 & 18 & 14 & 20 & &\\ %MUN: add a row which
+shows the sum of each column. This shows how much  %categories are covered
+overall %CG: fixed \hline \end{tabular} \end{table*}
 
@@
 papers' main contribution%MUN: which results? The papers' main contribution
-%or...? %CG: fixed
-, we were able to extrapolate information which confirmed the GSM categories.
+%or...? %CG: fixed , we were able to extrapolate information which confirmed the
+GSM categories.
 
@@
 need of empowering people, which is the element with the greatest impact on
-software development. Humans, being non-linear variables \cite{Human}
-%MUN: add ref %CG:fixed
-, are unlikely to follow repeatable prescriptive methodologies. As stated by 
-Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and every 
-methodology is to expect people to behave consistently over time, when is 
+software development. Humans, being non-linear variables \cite{Human} %MUN: add
+ref %CG:fixed , are unlikely to follow repeatable prescriptive methodologies. As
+stated by  Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and
+every  methodology is to expect people to behave consistently over time, when is
 clearly not like that''}. Especially in early-stage startups, where the whole
@@
 %MUN: Think of structuring this long subsection. Have a subsubsection for each
-%CAT, i.e. 5.2.1 Severe lack of resources (CAT7)
-%I've added to each category you mention the corresponding CATx since that  
-%identifier you use in table 2. I did not check if you discuss each and every 
-%category. Do you? If yes, ok. If not, why not? 
-%CG: partially fixed, the categories CAT3 and CAT6 are not well covered as 
-%discussed above.
+%CAT, i.e. 5.2.1 Severe lack of resources (CAT7) I've added to each category you
+%mention the corresponding CATx since that   identifier you use in table 2. I
+%did not check if you discuss each and every  category. Do you? If yes, ok. If
+%not, why not?  CG: partially fixed, the categories CAT3 and CAT6 are not well
+%covered as  discussed above.
+
+%------------------------------------------------------------------------------%
+\subsubsection{Limited resources}
+
 Starting from the category \textit{severe lack of resources} (CAT7), many
@@
 start to grow the accumulated technical debt which requires fast integration of
-more structured and standard processes overtime.
-Issues related to low-precision engineering practices during companies' growth 
-are discussed in an experience paper by Ambler \cite{Ambler2002}. He reports on 
-two case studies of growing Internet startups approaching an initial public  
-offering (IPO). %MUN:explain %acronym %CG: fixed
-Despite the two companies were facing business and team growth, attitudes 
-against the introduction of processes to structure and model the system were 
-evident. The need to re-engineer the codebase lead the researchers to suggest a 
-tailored version of the Rational Unified Process (RUP) 
-%MUN: add reference
-%CG: not fixed, not sure if needed since Ambler has the detailed 
-%explaination of his process, similar to RUP... anyway a dedicated resource 
-%would be an IBM website
-development framework. Nevertheless, even with the aid of CASE 
-tools\footnote{CASE tools are a class of software that automates many of the 
-activities involved in various engineering activities. See 
-\url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, 
-startups failed in practicing engineering activities in view of excessive lack of 
-flexibility. Developers fast moved to simpler tools such as using a white-board, 
-where they were able to provide input and insights, question assumptions, share 
-past experiences, and even talk about potential implementation strategies. 
-%MUN: Candidate for removal since details can be found in Ambler anyway: "In 
-%general the development remained highly iterative, where project teams started 
-%to have a little modeling and testing only when necessary. They didn't take 
-%the serial approach of completely developing and accepting the requirements 
-%model, analysis model and so on. This let them react swiftly and adapt 
-%rapidly according to changes into their highly competitive markets. The 
-%researchers report \textit{``modeling was streamlined''}, that is modeling 
-%what was needed and go straightforward to coding, avoiding analysis 
-%paralysis." %CG: ok
-
-The use of whiteboards in working environments supports \textit{use of 
-well-integrated and simple tools}. In both the startups studied by Ambler, 
+more structured and standard processes overtime. Issues related to low-precision
+engineering practices during companies' growth  are discussed in an experience
+paper by Ambler \cite{Ambler2002}. He reports on  two case studies of growing
+Internet startups approaching an initial public   offering (IPO). %MUN:explain
+%acronym %CG: fixed Despite the two companies were facing business and team
+growth, attitudes  against the introduction of processes to structure and model
+the system were  evident. The need to re-engineer the codebase lead the
+researchers to suggest a  tailored version of the Rational Unified Process (RUP)
+%MUN: add reference %CG: not fixed, not sure if needed since Ambler has the
+detailed  %explaination of his process, similar to RUP... anyway a dedicated
+resource  %would be an IBM website development framework. Nevertheless, even
+with the aid of CASE  tools\footnote{CASE tools are a class of software that
+automates many of the  activities involved in various engineering activities.
+See  \url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .},
+startups failed in practicing engineering activities in view of excessive lack
+of  flexibility. Developers fast moved to simpler tools such as using a white-
+board,  where they were able to provide input and insights, question
+assumptions, share  past experiences, and even talk about potential
+implementation strategies.  %MUN: Candidate for removal since details can be
+found in Ambler anyway: "In  %general the development remained highly iterative,
+where project teams started  %to have a little modeling and testing only when
+necessary. They didn't take  %the serial approach of completely developing and
+accepting the requirements  %model, analysis model and so on. This let them
+react swiftly and adapt  %rapidly according to changes into their highly
+competitive markets. The  %researchers report \textit{``modeling was
+streamlined''}, that is modeling  %what was needed and go straightforward to
+coding, avoiding analysis  %paralysis." %CG: ok
+
+%------------------------------------------------------------------------------%
+\subsubsection{Minimal project management}
+
+The use of whiteboards in working environments supports \textit{use of  well-
+integrated and simple tools}. In both the startups studied by Ambler,
 documentation was minimal, finding the ``sweet spot'' to have just barely enough
@@
 %MUN: the whole paragraph needs to be summarized to its essence. Details can be
-%read in the referred paper. 
-%CG: partially fixed, removed the detailed process description
-In the evaluation phase, the status of the development workflow within the 
-company  is analyzed. The researchers documented a state of chaos where any 
-initiatives of quantitative methodology assessment would have been a waste of 
-time. Therefore, they applied a qualitative methodology to define a model of the 
-release management process. 
-%They moved towards the plan phase where the 
-%researchers decided to develop a model to manage releases, introducing a process 
-%composed by: release scope preparation, release planning, release development, 
-%system testing, acceptance testing and release deployment. 
-Despite the effort spent by the researchers to establish a development workflow, 
-practices were not fully conducted even though some marginal benefits have been 
-reported: planned major release from four to six months; planned minor releases 
-from two to four weeks; encompassing urgent corrective releases from one to five 
-days. 
-Regarding the previously mentioned problems, only 
-partial improvements have been achieved since the major obstacle was to 
-motivate employees to change their habits. With perfect hindsight the 
+%read in the referred paper.  %CG: partially fixed, removed the detailed process
+description In the evaluation phase, the status of the development workflow
+within the  company  is analyzed. The researchers documented a state of chaos
+where any  initiatives of quantitative methodology assessment would have been a
+waste of  time. Therefore, they applied a qualitative methodology to define a
+model of the  release management process.  %They moved towards the plan phase
+where the  %researchers decided to develop a model to manage releases,
+introducing a process  %composed by: release scope preparation, release
+planning, release development,  %system testing, acceptance testing and release
+deployment.  Despite the effort spent by the researchers to establish a
+development workflow,  practices were not fully conducted even though some
+marginal benefits have been  reported: planned major release from four to six
+months; planned minor releases  from two to four weeks; encompassing urgent
+corrective releases from one to five  days.  Regarding the previously mentioned
+problems, only  partial improvements have been achieved since the major obstacle
+was to  motivate employees to change their habits. With perfect hindsight the
 researchers claim that understanding and adapting solutions by solving one
@@
 
-Prescribed development models, oriented towards SPI, are basically ignored by 
-startups, as discussed in \cite{Zettel2001}. 
-%MUN: <-- rephrase, not clear to me %CG:fixed
-To overcome the need of structure and to accelerate the development process, 
-Zettel developed a \textit{``LIghtweight Process for E-business software 
-development''} (LIPE). This lightweight development method integrates the XP 
-approach with ideas from areas of software measurement for project control and 
-process improvement as a compromise between an ad-hoc and a more rigorous 
-approach. Zettel argues that LIPE provides scalability of the development 
-process with a certain degree of flexibility, omitting parts of prescribed 
-practices. Nevertheless, we couldn't identify any empirical evaluation of LIPE 
-that supports its suitability in a startup context.
+%MUN: To which category does the following paragraph belong?  %CG: they
+represent subcategories of speed-up development  Eventually to externalize
+complexity of the developing product %MUN: solutions on what? CG: fixed , two
+other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the  wheel,
+instead of reinventing it''}, taking advantage of open source software  whenever
+it is possible, building systems quickly and effectively as a result.  Even
+though the principles in leading-edge development processes are the same
+principles of yesteryear, the way in which the fundamentals are applied has
+changed from prescriptive to agile solutions, with less concern on quality
+aspects, also confirmed by Mater \cite{Mater2000}.
+
+%------------------------------------------------------------------------------%
+\subsubsection{Simple workflows} Prescribed development models, oriented towards
+SPI, are basically ignored by  startups, as discussed in \cite{Zettel2001}.
+%MUN: <-- rephrase, not clear to me %CG:fixed To overcome the need of structure
+and to accelerate the development process,  Zettel developed a
+\textit{``LIghtweight Process for E-business software  development''} (LIPE).
+This lightweight development method integrates the XP  approach with ideas from
+areas of software measurement for project control and  process improvement as a
+compromise between an ad-hoc and a more rigorous  approach. Zettel argues that
+LIPE provides scalability of the development  process with a certain degree of
+flexibility, omitting parts of prescribed  practices. Nevertheless, we couldn't
+identify any empirical evaluation of LIPE  that supports its suitability in a
+startup context.
 
@@
 innovative product in volatile e-commerce environment. This methodology
-represents a fast evolutionary approach with flexible processes. 
-It aims to obtain customer's feedback as soon as possible for continuous 
-product improvement.%MUN: %<-- rephrase, sentence not parsable CG:fixed
-Even in this case, the spiral methodology%MUN: which model? CG:fixed
-is compliant with the nature of a random, ad-hoc and  \textit{evolutionary 
-development approach} (CAT2). Evolutionary development approaches are 
-typically more suitable for web applications, as confirmed in 
-\cite{Deshpande2001}.
-Moreover, they provide practices oriented towards rapid and iterative design 
-with complete freedom to the developer team, enabled to operate anywhere within 
-the organization.
-%MUN: <-- the whole paragraph is not that clear to me. Try to start 
-%with the category (team is catalyst), then write about Deakins spiral model
-%CG: partially fixed, the focus here is on the evolutionary development approach
+represents a fast evolutionary approach with flexible processes.  It aims to
+obtain customer's feedback as soon as possible for continuous  product
+improvement.%MUN: %<-- rephrase, sentence not parsable CG:fixed Even in this
+case, the spiral methodology%MUN: which model? CG:fixed is compliant with the
+nature of a random, ad-hoc and  \textit{evolutionary  development approach}
+(CAT2). Evolutionary development approaches are  typically more suitable for web
+applications, as confirmed in  \cite{Deshpande2001}. Moreover, they provide
+practices oriented towards rapid and iterative design  with complete freedom to
+the developer team, enabled to operate anywhere within  the organization. %MUN:
+<-- the whole paragraph is not that clear to me. Try to start  %with the
+category (team is catalyst), then write about Deakins spiral model %CG:
+partially fixed, the focus here is on the evolutionary development approach
+
+%------------------------------------------------------------------------------%
+\subsubsection{Full-stack development team}
 
 Carmel \cite{Camel1994a} reports on the need of flexible and rapid development
-solutions to shorten time-to-market, but even more important, the need of 
-high-skilled team developers, which has an impact on \textit{speed-up 
-development} (CAT1). He suggests that entrepreneurs need to look for a well 
-formed, skilled core development team and not just a set of product ideas and 
-features. Moreover the team must be empowered with full-stack and 
-self-organization settings, paired with the flexibility of minimal bureaucracy. 
-A studied company stated: \textit{``we didn't need weeks or months of detailed 
-modeling and documentation, but rather modeling the architecture a little and 
-then either exploring strategies by providing users with code or simply starting 
-work on the actual software itself''}.
-
-%MUN: To which category does the following paragraph belong? 
-%CG: they represent subcategories of speed-up development 
-Eventually to externalize complexity of the developing product
-%MUN: solutions on what? CG: fixed
-, two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the 
-wheel, instead of reinventing it''}, taking advantage of open source software 
-whenever it is possible, building systems quickly and effectively as a result. 
-Even though the principles in leading-edge development processes are the same 
-principles of yesteryear, the way in which the fundamentals are applied has 
-changed from prescriptive to agile solutions, with less concern on quality 
-aspects, also confirmed by Mater \cite{Mater2000}.
+solutions to shorten time-to-market, but even more important, the need of  high-
+skilled team developers, which has an impact on \textit{speed-up  development}
+(CAT1). He suggests that entrepreneurs need to look for a well  formed, skilled
+core development team and not just a set of product ideas and  features.
+Moreover the team must be empowered with full-stack and  self-organization
+settings, paired with the flexibility of minimal bureaucracy.  A studied company
+stated: \textit{``we didn't need weeks or months of detailed  modeling and
+documentation, but rather modeling the architecture a little and  then either
+exploring strategies by providing users with code or simply starting  work on
+the actual software itself''}.
 
@@
 %explanatory (what does "compliant with the behavioural model of the
-%phenomenon" mean?). 
-%CG: not fixed, when I wrote it I tried to have an overview of the main 
-%contributions and make it fluently by papers description. Creating subsections 
-%would mean to divide papers from different focuses. This is a trade-off where 
-%we took a decision in the past. Take it in consideration, but we can discuss 
-%it during the next meeting.
-%MUN1: Why would creating subsubsections like 5.2.1 Severe lack of resources 
-%(CAT7) remove the flow? This subdivision would give the reader clues what the 
-%subsubsection is about, he can relate it to Table 2, and provides some 
-%navigational support. Now there are nearly 2p of flowing text with now clue in 
-%what order/priority the literature is covered. I think you don't even need to 
-%rewrite much, just add the subsubsections and a proper summary.
-
-Thus, by summarizing the comparison, we have found that most of the 
-contributions of the studies are compliant with the behavioral model of the 
-phenomenon we provided.
-
-%------------------------------------------------------------------------------%
-\subsection{Confounding factors}
-\label{sect:an:comp:cat-vs-literature:conf}
+%phenomenon" mean?).  CG: not fixed, when I wrote it I tried to have an overview
+%of the main  contributions and make it fluently by papers description. Creating
+%subsections  would mean to divide papers from different focuses. This is a
+%trade-off where  we took a decision in the past. Take it in consideration, but
+%we can discuss  it during the next meeting. MUN1: Why would creating
+%subsubsections like 5.2.1 Severe lack of resources  (CAT7) remove the flow?
+%This subdivision would give the reader clues what the  subsubsection is about,
+%he can relate it to Table 2, and provides some  navigational support. Now there
+%are nearly 2p of flowing text with now clue in  what order/priority the
+%literature is covered. I think you don't even need to  rewrite much, just add
+%the subsubsections and a proper summary. CG: partially fixed
+
+%------------------------------------------------------------------------------%
+\subsubsection{Summary} Summarizing the comparison, we have found that most of
+the  contributions of the studies are compliant with the GSM model. Startups
+focusing on simple workflows and full-stack engineers face the uncertainty and
+immaturity of the company. Nevertheless few considerations are presented towards
+the consequences of a minimal project management and the use of low engineering
+activities as presented in the GSM model. However, the capability of the model
+to accommodate the results identified by previous researches give us more
+confidence in the generalizability of the GSM model.
+
+%------------------------------------------------------------------------------%
+\subsection{Confounding factors} \label{sect:an:comp:cat-vs-literature:conf}
 
@@
 identified by the SMS \cite{SMS} %MUN: do you mean the SMS? Then refer to it.
-%CG: fixed 
-%I remember you mentioned CF in the results (GSM), maybe also refer to the 
-%respective sections CG: not parsed.
-%MUN1: me neither... forget what I wrote :-)
-, related to: creativity and innovation, market requirements and application 
-type, and developer experience.
+%CG: fixed  %I remember you mentioned CF in the results (GSM), maybe also refer
+to the  %respective sections CG: not parsed. %MUN1: me neither... forget what I
+wrote :-). CG: ok , related to: creativity and innovation, market requirements
+and application  type, and developer experience.
 
 %MUN: Again, structure this into subsubsections according to the above factors.
-%I did it, but check if you agree...
-\subsubsection{Creativity and innovation} 
+%I did it, but check if you agree... \subsubsection{Creativity and innovation}
 Heitlager et al. \cite{Heitlager2007} contributed to the investigation of
-creativity and innovation in startups. 
-%MUN: the rest of this paragraph is too detailed. I'd remove this "The study 
-%reports how development approaches oriented to the product in the early-stage 
-%life-cycle, called \textit{fluid} have main impact to innovative solutions 
-%initially. After a while the need of structuring and controlling the growing 
-%company size (along with \textit{transition} and \textit{specific} phases) 
-%arises the need of project management activities and software processes for 
-%long-term scalability issues."  CG: ok.
-%MUN: Please reformulate the first sentence, I cannot parse it. CG: fixed
-The proposed model identifies the three lifecycle stages of a startup: 
-\textit{fluid}, \textit{transition} and \textit{specific}. At each stage, the 
-model presents different development approaches (product-oriented and 
-process-oriented), according to the level of innovation achieved. The study 
+creativity and innovation in startups.  %MUN: the rest of this paragraph is too
+detailed. I'd remove this "The study  %reports how development approaches
+oriented to the product in the early-stage  %life-cycle, called \textit{fluid}
+have main impact to innovative solutions  %initially. After a while the need of
+structuring and controlling the growing  %company size (along with
+\textit{transition} and \textit{specific} phases)  %arises the need of project
+management activities and software processes for  %long-term scalability
+issues."  CG: ok. %MUN: Please reformulate the first sentence, I cannot parse
+it. CG: fixed The proposed model identifies the three lifecycle stages of a
+startup:  \textit{fluid}, \textit{transition} and \textit{specific}. At each
+stage, the  model presents different development approaches (product-oriented
+and  process-oriented), according to the level of innovation achieved. The study
 reports how product-oriented development, in contrast with forced and mature
@@
 process in the long run (see how the \textit{initial growth hinders performance}
-(CAT6)). %MUN: not clear to me what you mean by "evolution 
-%of this model" and what "compliant with our assumption" means. Also, don't use 
-%only CATx in this context, but cite the category name AND CATx. CG: fixed
-From the \textit{fluid} phase to the \textit{specific} phase, the innovation 
-confounding factor seems to behave initially the same as the \textit{speed-up 
-development}, both highly focused towards a product-oriented development. 
-%MUN: <-- please provide some motivation... not obvious. CG: fixed
-Nevertheless, we were unable to ground these evidences in the GT data even 
-though those are factors that might be positively correlated 
-%MUN: "present strict correlation" %What does that mean, I don't understand. 
-%CG: fixed.
-to the choice of an early product-development orientation and the need of 
-low-precision engineering elements.
-
-\subsubsection{Market requirements and application type} 
-Market requirements \cite{Coleman2007} and application type \cite{Sutton2000, 
-Coleman2008, Coleman2008a} are two further potential confounding factors to the 
-GSM. Their main impact is related to the adoption of flexible and reactive 
-solutions for the development process. In particular, they refer to the 
-necessity of fulfillment of quality concerns that goes beyond scalability and UX 
-defined by the GSM. Especially when requirements are rigidly imposed or the 
-application domain is well-known, providing low-quality products to final users 
-might determine the failure of a startup.
+(CAT6)). %MUN: not clear to me what you mean by "evolution  %of this model" and
+what "compliant with our assumption" means. Also, don't use  %only CATx in this
+context, but cite the category name AND CATx. CG: fixed From the \textit{fluid}
+phase to the \textit{specific} phase, the innovation  confounding factor seems
+to behave initially the same as the \textit{speed-up  development}, both highly
+focused towards a product-oriented development.  %MUN: <-- please provide some
+motivation... not obvious. CG: fixed Nevertheless, we were unable to ground
+these evidences in the GT data even  though those are factors that might be
+positively correlated  %MUN: "present strict correlation" %What does that mean,
+I don't understand.  %CG: fixed. to the choice of an early product-development
+orientation and the need of  low-precision engineering elements.
+
+\subsubsection{Market requirements and application type}  Market requirements
+\cite{Coleman2007} and application type \cite{Sutton2000,  Coleman2008,
+Coleman2008a} are two further potential confounding factors to the  GSM. Their
+main impact is related to the adoption of flexible and reactive  solutions for
+the development process. In particular, they refer to the  necessity of
+fulfillment of quality concerns that goes beyond scalability and UX  defined by
+the GSM. Especially when requirements are rigidly imposed or the  application
+domain is well-known, providing low-quality products to final users  might
+determine the failure of a startup.
 
 For example, safety-critical systems possess quality concerns that must be
-considered. Even in e-commerce systems factors related to quality have 
-high-impact effects on obtaining the first customer reference, 
-%MUN: <-- cannot parse it. Remove the first part and write: "Even e-commerce 
-%systems...". Then rephrase the second part of the sentence to make it clearer. 
-%CG: fixed
-as discussed in \cite{Deakins2005,Kim2005,Silva2005}. Quality attributes  
-especially %MUN: what is "this"? CG: fixed
-characterize those applications, whose part of functionalities are already 
-successfully implemented in other software systems or whose failure might 
-inflict serious financial or social damages \cite{Bass2003}. Yet, limiting our 
-context to early-stage software startups, which typically develop first 
-cutting-edge products in beta versions, partially mitigates the high quality 
-demands of users.
+considered. Even in e-commerce systems factors related to quality have  high-
+impact effects on obtaining the first customer reference,  %MUN: <-- cannot
+parse it. Remove the first part and write: "Even e-commerce  %systems...". Then
+rephrase the second part of the sentence to make it clearer.  %CG: fixed as
+discussed in \cite{Deakins2005,Kim2005,Silva2005}. Quality attributes
+especially %MUN: what is "this"? CG: fixed characterize those applications,
+whose part of functionalities are already  successfully implemented in other
+software systems or whose failure might  inflict serious financial or social
+damages \cite{Bass2003}. Yet, limiting our  context to early-stage software
+startups, which typically develop first  cutting-edge products in beta versions,
+partially mitigates the high quality  demands of users.
 
@@
 %startups w.r.t. customer expectations and quality concerns. It's ok, but it
-%adds little value to the validation of GSM. I'd remove it.
-%To understand how to obtain processes in accordance to a set of customer 
-%needs, correlating the proper levels of customers' satisfaction, we 
-%introduce the Kano model, described in \cite{Ullah2011} and extensively 
-%studied in product development.  As shown in the Kano model, users 
-%preferences can be classified into four categories: \textit{basic}, 
-%\textit{excitement}, \textit{performance} % and \textit{indifference} areas.
-% 
-%The \textit{basic} area represents the space where customers are dissatisfied 
-%if those \textit{must-be} functionalities are not met. This is the case of 
-%similar functionalities already implemented in the current market with 
-%well-known solutions. The \textit{performance} area represents the space where 
-%customers' satisfaction is proportional to how many of functionalities are 
-%implemented. In the \textit{indifference} area the customer does not really 
-%care about the functionalities: the satisfaction remains invariant. This 
-%represents the case where startups have not (yet) achieved the right 
-%product/market fit. Finally, the \textit{excitement} area represents the space 
-%where customers will be more satisfied if functionalities are implemented, but 
-%not disappointed if they are not implemented. This is the case of extremely 
-%innovative product development.
-% 
-%Well-known domains such as e-commerce applications or other \textit{``quality 
-%critical products''} belong to the \textit{basic} area, where if a quality 
-%concern such as security in bank transfer services is missing satisfaction of 
-%customers will be profoundly affected. But as discussed, in early-stage 
-%startups that typically develop first beta products, developers usually deal 
-%with cutting-edge technologies. Those innovative solutions, if in the right 
-%market, are likely to belong to  the \textit{excitement} area, where 
-%satisfaction can hardly be negatively affected by faulty products. In fact, in 
-%the GSM model, minimal functionalities and less concern about quality aspects 
-%have low-impact on customers' satisfaction, supporting startup's ability to 
-%mainly focus on \textit{speed-up development}. Moreover operating in the 
-%\textit{excitement area} is fully compliant with an evolutionary approach and 
-%minimal functionalities since from the first product released to the market, 
-%it is unlikely to lose users because 
-% of high-quality demand.
-% 
-%Nevertheless, this aspect profoundly undermines the generalizability of the 
-%theoretical model according to the theory, in particular \textit{CAT3}. 
-%therefore, future studies within wider startups' context are advisable to 
-%extrapolate insights from those situations where quality concerns cannot be 
-%postponed or externalized. CG: ok
-
-\subsubsection{Developer experience} 
-%MUN: Ups, a confounding factor you did not mention initially. Add it... 
-%CG: fixed
-Finally, also the experience of team-members is considered as an important 
-confounding factor. Since only little evidence has been gathered during the 
-interviews, ``lack of experience'' was not considered as a sub-category in 
-\textit{severe lack of resources}. Nevertheless, startups rely at the beginning 
-on clever, but inexperienced developers \cite{Crowne2002}. Having team-members 
-with deep experience is a ``double-edge sword'' \cite{Yoffie1999}. On one hand, 
-experience might quickly provide structure and maturity to the development 
-process, on the other it might cause challenges in managing self-confident 
-overachievers that almost inevitably clash. Consequently, team management 
-requires higher control and coordination activities that inevitably hinder the 
-flexibility of the development environment, that is essential in early-stage 
-startups. Accordingly, also Coleman \cite{Coleman2007,
-Coleman2008a,Coleman2008} states that the operating strategies towards 
-\textit{``minimum processes''} in startups is not a matter of poor knowledge and 
-training, but rather it is the necessity of operating with solutions that let 
-the company move faster.
-
-%MUN: removed below paragraph since it does not add much
-%Nevertheless, since the amount of the results did not allow ``the lack of 
-%experience'' to be in the GSM model (in view of the fact that not enough data 
-%were identified \footnote{A category is formed if the codes are representative 
-%of more than the 70\% of the analyzed transcripts.}), this confounding factor 
-%requires further investigation. CG: ok
-
-%------------------------------------------------------------------------------%
-\subsection{Static assessment of the GSM}
-\label{sect:theory:validation:stats}
+%adds little value to the validation of GSM. I'd remove it. To understand how to
+%obtain processes in accordance to a set of customer  needs, correlating the
+%proper levels of customers' satisfaction, we  introduce the Kano model,
+%described in \cite{Ullah2011} and extensively  studied in product development.
+%As shown in the Kano model, users  preferences can be classified into four
+%categories: \textit{basic},  \textit{excitement}, \textit{performance}  and
+%\textit{indifference} areas.   The \textit{basic} area represents the space
+%where customers are dissatisfied  if those \textit{must-be} functionalities are
+%not met. This is the case of  similar functionalities already implemented in
+%the current market with  well-known solutions. The \textit{performance} area
+%represents the space where  customers' satisfaction is proportional to how many
+%of functionalities are  implemented. In the \textit{indifference} area the
+%customer does not really  care about the functionalities: the satisfaction
+%remains invariant. This  represents the case where startups have not (yet)
+%achieved the right  product/market fit. Finally, the \textit{excitement} area
+%represents the space  where customers will be more satisfied if functionalities
+%are implemented, but  not disappointed if they are not implemented. This is the
+%case of extremely  innovative product development.   Well-known domains such as
+%e-commerce applications or other \textit{``quality  critical products''} belong
+%to the \textit{basic} area, where if a quality  concern such as security in
+%bank transfer services is missing satisfaction of  customers will be profoundly
+%affected. But as discussed, in early-stage  startups that typically develop
+%first beta products, developers usually deal  with cutting-edge technologies.
+%Those innovative solutions, if in the right  market, are likely to belong to
+%the \textit{excitement} area, where  satisfaction can hardly be negatively
+%affected by faulty products. In fact, in  the GSM model, minimal
+%functionalities and less concern about quality aspects  have low-impact on
+%customers' satisfaction, supporting startup's ability to  mainly focus on
+%\textit{speed-up development}. Moreover operating in the  \textit{excitement
+%area} is fully compliant with an evolutionary approach and  minimal
+%functionalities since from the first product released to the market,  it is
+%unlikely to lose users because   of high-quality demand.   Nevertheless, this
+%aspect profoundly undermines the generalizability of the  theoretical model
+%according to the theory, in particular \textit{CAT3}.  therefore, future
+%studies within wider startups' context are advisable to  extrapolate insights
+%from those situations where quality concerns cannot be  postponed or
+%externalized. CG: ok
+
+\subsubsection{Developer experience}  %MUN: Ups, a confounding factor you did
+not mention initially. Add it...  %CG: fixed Finally, also the experience of
+team-members is considered as an important  confounding factor. Since only
+little evidence has been gathered during the  interviews, ``lack of experience''
+was not considered as a sub-category in  \textit{severe lack of resources}.
+Nevertheless, startups rely at the beginning  on clever, but inexperienced
+developers \cite{Crowne2002}. Having team-members  with deep experience is a
+``double-edge sword'' \cite{Yoffie1999}. On one hand,  experience might quickly
+provide structure and maturity to the development  process, on the other it
+might cause challenges in managing self-confident  overachievers that almost
+inevitably clash. Consequently, team management  requires higher control and
+coordination activities that inevitably hinder the  flexibility of the
+development environment, that is essential in early-stage  startups.
+Accordingly, also Coleman \cite{Coleman2007, Coleman2008a,Coleman2008} states
+that the operating strategies towards  \textit{``minimum processes''} in
+startups is not a matter of poor knowledge and  training, but rather it is the
+necessity of operating with solutions that let  the company move faster.
+
+%MUN: removed below paragraph since it does not add much Nevertheless, since the
+%amount of the results did not allow ``the lack of  experience'' to be in the
+%GSM model (in view of the fact that not enough data  were identified
+%\footnote{A category is formed if the codes are representative  of more than
+%the 70\ of the analyzed transcripts.}), this confounding factor  requires
+%further investigation. CG: ok
+
+%------------------------------------------------------------------------------%
+\subsection{Static assessment of the GSM} \label{sect:theory:validation:stats}
 
@@
 %methods you used on what data, then refer to that section from here (no need
-%to repeat again here). CG: fixed
-In this subsection we validate the most important causal relations presented in 
-our model by attesting their correctness  with a quantification of empirical 
-trends observed in the interview data. 
-%MUN1: The following is part of the method and should be in 3.5: "We identify 
-%trends and recurrent patterns by thoroughly analyzing and synthesizing the 
-%complete set of data obtained from the case study on which the GSM is built 
-%upon. We illustrate our observations aided by simple statistical methods, 
-%providing researchers an additional tool, supported by further companies' 
-%data, to validate our model."
+%to repeat again here). CG: fixed In this subsection we validate the most
+important causal relations presented in  our model by attesting their
+correctness  with a quantification of empirical  trends observed in the
+interview data.  %MUN1: The following is part of the method and should be in
+3.5: "We identify  %trends and recurrent patterns by thoroughly analyzing and
+synthesizing the  %complete set of data obtained from the case study on which
+the GSM is built  %upon. We illustrate our observations aided by simple
+statistical methods,  %providing researchers an additional tool, supported by
+further companies'  %data, to validate our model." CG: not important, I would
+not add this.
 
@@
 %MUN: for the paragraphs below, you need to explain what and why you do this.
-%CG: fixed, explained in methodology 
-%What do you mean by "we grounded in the interview transcripts the fact that 
-%those relations were characterized in terms of performances since 
-%time-to-market was the main concern." Not clear to me. 
-%CG: fixed, performance for us is efficiency AND effectiveness
-Before starting the evaluation of the impact of the three categories on 
-\textit{speed-up development}, we grounded in the interview transcripts the fact 
-that those relations were characterized in terms of "effectiveness" and 
-"efficiency" since time-to-market was the main concern (as confirmed in 
-\cite{SMS}):
-
-\begin{compactitem}
-\item The team enhances the overall \textit{performance} of software 
-development. Recalling that startups deal with severe lack of resources, the 
-team members are the company's main assets. They develop software with no need 
-of formalities and structures, left to the team's own capabilities to 
-efficiently develop the product. Moreover, the effectiveness is granted by 
-continuous advice given by mentors and by a management style oriented towards 
-self-organization, full-stack, and multi-role settings. The importance of a 
-skilled team for development performance is expressed in many studies such as 
-\cite{surviving-os-cockburn, Highsmith2000}, described in CAT4 and confirmed by 
-Sutton \cite{Sutton2000}.
-\item The evolutionary approach enhances the \textit{effectiveness} in 
-implementing the correct functionalities. This approach improves the company's 
-capabilities to adjust the trajectory of product development (in agreement to 
-the concept of pivoting discussed by Ries \cite{Ries2011}).
-\item Since product quality has low priority, the development is focused only on 
-implementing a limited number of suitable functionalities, enhancing its 
-\textit{efficiency}. An essential prototype (MVP) does not require to comply 
-with heavy quality constraints, enabling the team to quickly implement 
-functionalities, ready to be validated by final users.
-\end{compactitem}
+%CG: fixed, explained in methodology  %What do you mean by "we grounded in the
+interview transcripts the fact that  %those relations were characterized in
+terms of performances since  %time-to-market was the main concern." Not clear to
+me.  %CG: fixed, performance for us is efficiency AND effectiveness Before
+starting the evaluation of the impact of the three categories on  \textit{speed-
+up development}, we grounded in the interview transcripts the fact  that those
+relations were characterized in terms of "effectiveness" and  "efficiency" since
+time-to-market was the main concern (as confirmed in  \cite{SMS}):
+
+\begin{compactitem} \item The team enhances the overall \textit{performance} of
+software  development. Recalling that startups deal with severe lack of
+resources, the  team members are the company's main assets. They develop
+software with no need  of formalities and structures, left to the team's own
+capabilities to  efficiently develop the product. Moreover, the effectiveness is
+granted by  continuous advice given by mentors and by a management style
+oriented towards  self-organization, full-stack, and multi-role settings. The
+importance of a  skilled team for development performance is expressed in many
+studies such as  \cite{surviving-os-cockburn, Highsmith2000}, described in CAT4
+and confirmed by  Sutton \cite{Sutton2000}. \item The evolutionary approach
+enhances the \textit{effectiveness} in  implementing the correct
+functionalities. This approach improves the company's  capabilities to adjust
+the trajectory of product development (in agreement to  the concept of pivoting
+discussed by Ries \cite{Ries2011}). \item Since product quality has low
+priority, the development is focused only on  implementing a limited number of
+suitable functionalities, enhancing its  \textit{efficiency}. An essential
+prototype (MVP) does not require to comply  with heavy quality constraints,
+enabling the team to quickly implement  functionalities, ready to be validated
+by final users. \end{compactitem}
 
@@
 Accordingly, by analyzing the data of the 13 companies, we formulated two
-hypotheses which further describe the above illustrated relations:
-%MUN: You need to be more precise with terminology. Above you state that the 
-%two items below are "trends which further empower relations", but then, 
-%further below, you write "to assess these hypotheses". Observed trends != 
-%Hypotheses. This is also why you need to be extremely explicit and precise 
-%when explaining what the a) aim, b) method, and c) outcome/result of this 
-%section is. I'm aware that it is difficult to boil down what you wrote in the 
-%thesis to this limited space, but I think it is doable (and relevant for 
-%supporting validity of GSM). CG: fixed
-
-\begin{compactenum}
-\item The more a capable \footnote{The team capability is evaluated according 
-to the characteristics presented in CAT4.} team, with few quality constraints, 
-uses an evolutionary approach, the faster the company releases the product.
-\item The faster the company releases the product, the larger the degree of 
-\textit{accumulated technical debt}\footnote{The debt here is considered 
-regardless of context-specific debt mitigation factors and tactics.}.
-\end{compactenum}
+hypotheses which further describe the above illustrated relations: %MUN: You
+need to be more precise with terminology. Above you state that the  %two items
+below are "trends which further empower relations", but then,  %further below,
+you write "to assess these hypotheses". Observed trends !=  %Hypotheses. This is
+also why you need to be extremely explicit and precise  %when explaining what
+the a) aim, b) method, and c) outcome/result of this  %section is. I'm aware
+that it is difficult to boil down what you wrote in the  %thesis to this limited
+space, but I think it is doable (and relevant for  %supporting validity of GSM).
+CG: fixed
+
+\begin{compactenum} \item The more a capable \footnote{The team capability is
+evaluated according  to the characteristics presented in CAT4.} team, with few
+quality constraints,  uses an evolutionary approach, the faster the company
+releases the product. \item The faster the company releases the product, the
+larger the degree of  \textit{accumulated technical debt}\footnote{The debt here
+is considered  regardless of context-specific debt mitigation factors and
+tactics.}. \end{compactenum}
 
@@
 dynamic environment during the development process, given by the three
-categories that mostly affect \textit{speed-up development}.
-\end{compactitem}
+categories that mostly affect \textit{speed-up development}. \end{compactitem}
 
@@
 questionnaire results. The complete statistical procedure and provided rubrics
-to quantify the measures are illustrated and discussed in Appendix.
-%MUN: make sure it is clear that the PROCEDURE material is available in the 
-%ONLINE appendix and refer to it. Still, you need to report the actual RESULTS 
-%of you statistical analysis here... 
-%CG: partially fixed. we always referred to appendix without references until 
-%now.
-
-%MUN1: I don't really understand why you add the second clause in H01 
-%"AND with less quality constraints" --> just makes the hypothesis more 
-%difficult to understand. 
-We conducted statistical tests  using the analysis of variance to assess the 
-existence of relations between the dimensions. First we defined two null 
-hypotheses:
-
-\begin{itemize}
-\item[$H0_1$:] \textit{Startups do not release the product faster 
-when a capable team adopts a more evolutionary approach AND with less quality 
-constraints}
-\item[$H0_2$:] \textit{The execution speed does not increase the amount 
-of accumulated technical debt}. 
-\end{itemize}
-
-Then we tested $H0_1$ and $H0_2$  with an 
-one-tailed test using Pearson's product moment correlation 
-coefficient\footnote{To perform the Pearson's correlation we verified two 
-assumptions: a) data is normally distributed; b) data is on interval scale.}, 
-with positive association analysis, fixing the level of confidence to 95\%.
-We reject both $H0_1$ and $H0_2$, and conclude that in our sample:
-\begin{compactenum}
+to quantify the measures are illustrated and discussed in Appendix. %MUN: make
+sure it is clear that the PROCEDURE material is available in the  %ONLINE
+appendix and refer to it. Still, you need to report the actual RESULTS  %of you
+statistical analysis here...  %CG: partially fixed. we always referred to
+appendix without references until  %now.
+
+%MUN1: I don't really understand why you add the second clause in H01  %"AND
+with less quality constraints" --> just makes the hypothesis more  %difficult to
+understand. CG: fixed. We conducted statistical tests  using the analysis of
+variance to assess the  existence of relations between the dimensions. First we
+defined two null  hypotheses:
+
+\begin{itemize} \item[$H0_1$:] \textit{Startups do not release the product
+faster  when a capable team adopts a more evolutionary approach AND with less
+quality  constraints} \item[$H0_2$:] \textit{The execution speed does not
+increase the amount  of accumulated technical debt}.  \end{itemize}
+
+The formulation of the two hypotheses reflects the representation of the
+categories related to \textit{speed-up development} presented in the GSM model
+(see figure \ref{fig:gsm}). We tested $H0_1$ and $H0_2$  with an  one-tailed
+test using Pearson's product moment correlation  coefficient\footnote{To perform
+the Pearson's correlation we verified two  assumptions: a) data is normally
+distributed; b) data is on interval scale.},  with positive association
+analysis, fixing the level of confidence to 95\%. We reject both $H0_1$ and
+$H0_2$, and conclude that in our sample: \begin{compactenum} \item Higher values
+of \textit{Execution speed} are strongly associated  (p-value: 0.002) with
+higher values for \textit{Technical debt}.
+
 \item Higher values of \textit{Execution speed} are  strongly associated
-(p-value: 0.002) with higher values for \textit{Technical debt}.
-
-\item Higher values of \textit{Execution speed} are 
-strongly associated (p-value: 0.005) with higher values for  
-\textit{Potential capability}.
+(p-value: 0.005) with higher values for   \textit{Potential capability}.
 \end{compactenum}
@@
 %------------------------------------------------------------------------------%
-\section{Discussion} %MUN:Implications of the GSM
+\section{Implications of the GSM} %MUN:Implications of the GSM. CG:fixed
 \label{sect:theory:impl}
@@
 %next steps for research in this area. You can be creative/visionary here as
-%long as your arguments follow from the previous analysis.
-%CG: not fixed, I have shortened the content of one page. But still we cannot 
-%create the subsections you proposed since the implications are all derived 
-%from the model. Then, we can still shorten it if you want. Let's discuss it 
-%during the next skype call.
+%long as your arguments follow from the previous analysis. CG: not fixed, I have
+%shortened the content of one page. But still we cannot  create the subsections
+%you proposed since the implications are all derived  from the model. Then, we
+%can still shorten it if you want. Let's discuss it  during the next skype call.
 
@@
 %with the subsection names. Still, I would rename the section: "Discussion" -->
-%"Implications of the GSM"
-
-In this section we present relevant implications which 
-emerge from the behavior of early-stage startups, formally expressed in the GSM, 
-answering to our research question (\textit{What is the current 
-state-of-practice related to software development strategy in early-stage 
-startups?}) through its sub-questions. Although the startups we inquired in the 
-case study were spread across different nations and market sectors, we have 
-found several patterns that emerged very clearly.
+%"Implications of the GSM". CG: fixed
+
+In this section we present relevant implications which  emerge from the behavior
+of early-stage startups, formally expressed in the GSM,  answering to our
+research question (\textit{What is the current  state-of-practice related to
+software development strategy in early-stage  startups?}) through its sub-
+questions. Although the startups we inquired in the  case study were spread
+across different nations and market sectors, we have  found several patterns
+that emerged very clearly.
 
@@
 startups to quickly iterate on the product by exploring and experimenting new
-functionalities progressively rolled-out to early-adopters.
-%MUN1: Ok, I'm now "inventing" possible implications in the sections below... 
-%this is your job and you are much more qualified to do that than me :-)
+functionalities progressively rolled-out to early-adopters. %MUN1: Ok, I'm now
+"inventing" possible implications in the sections below...  %this is your job
+and you are much more qualified to do that than me :-)
 
 %MUN1:Do we (research) know enough about how to shorten time-to-market without
-%accumulating technical debt? What are possible venues for research?
-
-%------------------------------------------------------------------------------%
-\subsection{Light-weight methodology}
-Startups do not apply any standard development methodology. When focused 
-on building the first version of the product, startups do not observe any 
-specific and standard development methodologies or processes.  From an accurate 
-and in-depth investigation into daily activities, it emerged that the number of  
-the observed development \textit{practices} is extremely small and never 
-systematic (only little \textit{pair programming} sessions and rarely 
-\textit{test driven development}). Despite the wide number of publications, 
-which discuss the adoption of lightweight methodologies, we have found that 
-especially in this first stage not only methodologies but even individual 
-practices are basically neglected.
-
-The closest development approach undertaken by early-stage startups is 
-the \textit{Lean startup methodology}.  A highly evolutionary development 
-approach, centered around the quick production of a functioning prototype and 
-guided by customer feedback is a central aspect of the \textit{Lean startup 
-methodology} pioneered by Ries \cite{Ries2011}. However, startups do not 
-formally follow the cycle \textit{build-measure-learn} as it is proposed by the 
-methodology, where they should set-up an experiment to \textit{test the riskiest 
-element of the business idea}, nor make explicit use of the suggested good 
-practices.
-%MUN1: Why don't follow startups the build-measure-learn cycle proposed by 
-%Ries? Well, naively I'd say that since he published in 2011, his ideas do not 
-%have yet transferred to industry... Any other reason you can think of? Is it 
-%still too heavyweight and in conflict with the highest priority of shorting 
-%time-to-market? Is Lean still to Heavy for startups (in their early phases)?
-For example, when it comes to user feedback, startups do 
-not explicitly follow the step-by-step process of \textit{customer development}
-formally defined by Blank \cite{Blank2005}. Instead, startups absorb and 
-implement the high-level principles that outlines the methodology%MUN1:<-- 
-%which methodology?
-, reflected in the GSM by the theoretical category \textit{find the 
-product/market fit quickly}.
-
-%MUN1: paragraph below is candidate for removal
-%The greatest part of engineering activities of startups are focused on 
-%the implementation while only little attention is given to more conventional 
-%activities (project management, requirement specifications, analysis, 
-%architecture design, automatic testing, \ldots). Especially when bringing the 
-%first product to the market, the attitude of startups towards traditional 
-%engineering activities is strongly characterized by a general feeling of 
-%unnecessary. Every document, artifact or process, which does not produce a 
-%short-term tangible benefit to the software production is perceived as a waste 
-%of time.
-
-%------------------------------------------------------------------------------%
-\subsection{Focus on minimal set of functionalities}
-The first release of the product includes only a limited set of well 
-suitable functionalities focused on user experience. The beta version is usually 
-focused on a minimal and essential set of functionalities providing a reasonable 
-user experience (UX). 
-%MUN1: This rises the question, how do startups prioritize requirements. 
-%There's abundant amount of literature on how to select/prioritize reqs., but 
-%do they work in startups? Are they also too heavyweight or could they be 
-%adopted by startups in a modified form in order to increase the probabilities 
-%of success in the market?
-As confirmed by practitioners, the user must be able to 
-accomplish a task on the product without major interruptions in the flow, which 
-would cause frustration before being able to capture his attention. To realize 
-UX in a short time frame, developers focus on a limited number of suitable 
-functionalities, especially in the first version(s) of the product. Other 
-product qualities are regarded as less important: for instance, obtaining a 
-high-reliability is not a priority since early adopters of innovative product 
-are keen to accept minor bugs in exchange of new functionalities\footnote{These 
-results further highlight differences between software development in startups 
-and traditional web engineering, where quality aspects are rated as very 
-important, even more than time-to-market \cite{Offutt2002}.}. From a business 
-perspective, one might argue that established companies might take advantage of 
-the new idea and develop a better version of the product internally.
-Despite this might represent a risk, established companies usually cannot 
-afford to operate in a market of \textit{innovators} and \textit{early adopters} 
-group of users - described by the Roger's bell curve in \cite{Rogers} - because 
-of the little return of investment they would gain, as discussed by Christensen 
-in \cite{Christensen2003}. Instead, established companies have more interests 
-in acquiring a startup and all the knowledge it has gained during the 
+%accumulating technical debt? What are possible venues for research? CG: fixed
+%see summary.
+
+%------------------------------------------------------------------------------%
+\subsection{Light-weight methodology} Startups do not apply any standard
+development methodology. When focused  on building the first version of the
+product, startups do not observe any  specific and standard development
+methodologies or processes.  From an accurate  and in-depth investigation into
+daily activities, it emerged that the number of   the observed development
+\textit{practices} is extremely small and never  systematic (only little
+\textit{pair programming} sessions and rarely  \textit{test driven
+development}). Despite the wide number of publications,  which discuss the
+adoption of lightweight methodologies, we have found that  especially in this
+first stage not only methodologies but even individual  practices are basically
+neglected.
+
+The closest development approach undertaken by early-stage startups is  the
+\textit{Lean startup methodology}.  A highly evolutionary development  approach,
+centered around the quick production of a functioning prototype and  guided by
+customer feedback is a central aspect of the \textit{Lean startup  methodology}
+pioneered by Ries \cite{Ries2011}. However, startups do not  formally follow the
+cycle \textit{build-measure-learn} as it is proposed by the  methodology, where
+they should set-up an experiment to \textit{test the riskiest  element of the
+business idea}, nor make explicit use of the suggested good  practices. Instead,
+startups in the early stage prefer also to apply fast cycles of "build and fix"
+when necessary, hacking the first functionalities to present to potential
+customers and investors in order to obtain quick feedbacks. %MUN1: Why don't
+follow startups the build-measure-learn cycle proposed by  %Ries? Well, naively
+I'd say that since he published in 2011, his ideas do not  %have yet transferred
+to industry... Any other reason you can think of? Is it  %still too heavyweight
+and in conflict with the highest priority of shorting  %time-to-market? Is Lean
+still to Heavy for startups (in their early phases)? CG:fixed. Moreover, when it
+comes to user feedback, startups do  not explicitly follow the step-by-step
+process of \textit{customer development} formally defined by Blank
+\cite{Blank2005}. Instead, startups absorb and  implement the high-level
+principles that outlines the customer development methodology%MUN1:<--  %which
+methodology? CG: fixed , reflected in the GSM by the theoretical category
+\textit{find the  product/market fit quickly}.
+
+%MUN1: paragraph below is candidate for removal The greatest part of engineering
+%activities of startups are focused on  the implementation while only little
+%attention is given to more conventional  activities (project management,
+%requirement specifications, analysis,  architecture design, automatic testing,
+%\ldots). Especially when bringing the  first product to the market, the
+%attitude of startups towards traditional  engineering activities is strongly
+%characterized by a general feeling of  unnecessary. Every document, artifact or
+%process, which does not produce a  short-term tangible benefit to the software
+%production is perceived as a waste  of time. CG: ok
+
+%------------------------------------------------------------------------------%
+\subsection{Focus on minimal set of functionalities} The first release of the
+product includes only a limited set of well  suitable functionalities focused on
+user experience. The beta version is usually  focused on a minimal and essential
+set of functionalities providing a reasonable  user experience (UX). The first
+functionalities are usually set by the founders and their experience in a
+particular business domain. As soon as possible they provide demos to
+demonstrate the idea to potential customers. %MUN1: This rises the question, how
+do startups prioritize requirements.  %There's abundant amount of literature on
+how to select/prioritize reqs., but  %do they work in startups? Are they also
+too heavyweight or could they be  %adopted by startups in a modified form in
+order to increase the probabilities  %of success in the market?. CG: fixed. As
+confirmed by practitioners, the user must be able to  accomplish a task on the
+product without major interruptions in the flow, which  would cause frustration
+before being able to capture his attention. To realize  UX in a short time
+frame, developers focus on a limited number of suitable  functionalities,
+especially in the first version(s) of the product. Other  product qualities are
+regarded as less important: for instance, obtaining a  high-reliability is not a
+priority since early adopters of innovative product  are keen to accept minor
+bugs in exchange of new functionalities\footnote{These  results further
+highlight differences between software development in startups  and traditional
+web engineering, where quality aspects are rated as very  important, even more
+than time-to-market \cite{Offutt2002}.}. From a business  perspective, one might
+argue that established companies might take advantage of  the new idea and
+develop a better version of the product internally. Despite this might represent
+a risk, established companies usually cannot  afford to operate in a market of
+\textit{innovators} and \textit{early adopters}  group of users - described by
+the Roger's bell curve in \cite{Rogers} - because  of the little return of
+investment they would gain, as discussed by Christensen  in
+\cite{Christensen2003}. Instead, established companies have more interests  in
+acquiring a startup and all the knowledge it has gained during the
 \textit{customer development} process while focusing effort on acquiring larger
-shares of a ``mass market'' sector. 
-%MUN: candidate for shorting
-%Furthermore, thanks to cloud computing and 
-%more in general high availability of third party components, aspects such as 
-%performance, security and scalability can be achieved with a reasonable 
-%effort. 
-%Thus, given the general lack of quality concerns, and supported by the fact 
-%that the optimal strategy to improve the UX is by having the product used by 
-%real customers, startups can effectively focus on achieving short 
-%time-to-market.
-
-%------------------------------------------------------------------------------%
-\subsection{Empowering the team members}
-The founding team is the most important determinant of speed. More than 
-techniques and methodologies, the features of the team are the essential factors 
-to achieve high-development speed. Discussed in the theoretical category the 
-\textit{team is the catalyst of development}, this result has been confirmed in 
-multiple occasions both in the interviews and in the questionnaires. What is 
-truly important for the startup, among other factors, is the extremely 
-high-colocation which enables continuous informal and colloquial coordination 
-and discussions. 
-%$MUN1: candidate for removal
-%The initial (small) development team usually spends most of the 
-%hours of the week working closely to implement the product, often extending 
-%the conventional office hours.
-The large amount of communication, spontaneously held in informal places, has 
-been proven to be an extremely effective element of 
+shares of a ``mass market'' sector.  %MUN1: candidate for shorting %Furthermore,
+thanks to cloud computing and  %more in general high availability of third party
+components, aspects such as  %performance, security and scalability can be
+achieved with a reasonable  %effort.  %Thus, given the general lack of quality
+concerns, and supported by the fact  %that the optimal strategy to improve the
+UX is by having the product used by  %real customers, startups can effectively
+focus on achieving short  %time-to-market. CG: fixed
+
+%------------------------------------------------------------------------------%
+\subsection{Empowering the team members} The founding team is the most important
+determinant of speed. More than  techniques and methodologies, the features of
+the team are the essential factors  to achieve high-development speed. Discussed
+in the theoretical category the  \textit{team is the catalyst of development},
+this result has been confirmed in  multiple occasions both in the interviews and
+in the questionnaires. What is  truly important for the startup, among other
+factors, is the extremely  high-colocation which enables continuous informal and
+colloquial coordination  and discussions.  %$MUN1: candidate for removal %The
+initial (small) development team usually spends most of the  %hours of the week
+working closely to implement the product, often extending  %the conventional
+office hours. CG: ok. The large amount of communication, spontaneously held in
+informal places, has  been proven to be an extremely effective element of
 several successful software projects \cite{Highsmith2000}. Hence, the vastly
 \textit{informal} development environment of startups acts as a catalyst of
-speed.
-%MUN1: candidate for removal
-%Other important elements, which contribute in achieving such an initial 
-%high-speed, are: minimal set of functionalities with low concerns on product 
-%quality; evolutionary prototyping approach; smart use of advanced tools and 
-%frameworks; low precision artifacts; and the close colocation and capability 
-%of the team.
-
-%------------------------------------------------------------------------------%
-\subsection{Paying back the technical debt}
-The initial lack of structures and processes negatively affects the 
-performance when the company starts to grow. One of the primary objectives of 
-startups is to expand their business into fast growing markets. If 
-successful, the company will face a growth in terms of number of customers, 
-employees and product capabilities. Then, the increased complexity arises the 
-necessity of controlling the initially chaotic software development environment. 
-Under delicate circumstances, where an increased market demand meets a poorly 
-engineered product and process, the development team need to start returning 
-the accumulated technical debt. As reported by practitioners, introducing more 
-structures, documentation and standard procedures, cause an initial drop-down in 
-performance which severity depends on the speed of growth, the amount of 
-technical debt and the capacity of the team to absorb it. 
-%MUN1: what does literature say about technical debt? Are there strategies that 
-%maybe applicable to startups and if not why? Is there more research needed? 
-%Obviously, shortening time-to-market and preventing technical debt are two 
-%conflicting goals, but is there a chance to find a sweet spot where 
-%development speed can be balanced to achieve market success and still reducing 
-%the performance bump that hampers growth?
-
-\subsection{Summary}
-All of the above observations are extracted from the analysis of the model, 
-which is systematically derived from empirical data of the case study. Indeed, 
-with slightly different levels of adherence, the implications we presented are 
-reflected in the behavior of the great majority of companies in the sample. The 
-results of this analysis, which has been focused on the software development 
-strategy from the idea conception to the first release, indicate that startups 
-are far from adopting standard or ad-hoc development methodologies, especially 
-in this phase. The typical tendency is to focus on the team capability to 
-implement and quickly iterate on a prototype, which is released very fast. Thus, 
-in a context where even for the most \textit{lightweight} agile methodologies is 
-hard to penetrate, we believe that it is premature to discuss issues related to 
-SPI. Researchers should rather focus on the trade-off between development speed 
-and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which 
-appears to be the most important determinant for the future of 
-software development in the company.
-
-%------------------------------------------------------------------------------%
-%\subsection{Future work}
-%label{futwork}
-%MUN: Don't have future work as a subsection in threats to validity. Either 
-%merge it in "Implications of the GSM" OR into "Conclusions". CG: fixed
+speed. %MUN1: candidate for removal %Other important elements, which contribute
+in achieving such an initial  %high-speed, are: minimal set of functionalities
+with low concerns on product  %quality; evolutionary prototyping approach; smart
+use of advanced tools and  %frameworks; low precision artifacts; and the close
+colocation and capability  %of the team. CG: ok.
+
+%------------------------------------------------------------------------------%
+\subsection{Paying back the technical debt} The initial lack of structures and
+processes negatively affects the  performance when the company starts to grow.
+One of the primary objectives of  startups is to expand their business into fast
+growing markets. If  successful, the company will face a growth in terms of
+number of customers,  employees and product capabilities. Then, the increased
+complexity arises the  necessity of controlling the initially chaotic software
+development environment.  Under delicate circumstances, where an increased
+market demand meets a poorly  engineered product and process, the development
+team need to start returning  the accumulated technical debt. As reported by
+practitioners, introducing more  structures, documentation and standard
+procedures, cause an initial drop-down in  performance which severity depends on
+the speed of growth, the amount of  technical debt and the capacity of the team
+to absorb it.  %MUN1: what does literature say about technical debt? Are there
+strategies that  %maybe applicable to startups and if not why? Is there more
+research needed?  %Obviously, shortening time-to-market and preventing technical
+debt are two  %conflicting goals, but is there a chance to find a sweet spot
+where  %development speed can be balanced to achieve market success and still
+reducing  %the performance bump that hampers growth? CG: partially fixed below
+in summary.
+
+\subsection{Summary} All of the above observations are extracted from the
+analysis of the model,  which is systematically derived from empirical data of
+the case study. Indeed,  with slightly different levels of adherence, the
+implications we presented are  reflected in the behavior of the great majority
+of companies in the sample. The  results of this analysis, which has been
+focused on the software development  strategy from the idea conception to the
+first release, indicate that startups  are far from adopting standard or ad-hoc
+development methodologies, especially  in this phase. The typical tendency is to
+focus on the team capability to  implement and quickly iterate on a prototype,
+which is released very fast. Thus,  in a context where even for the most
+\textit{lightweight} agile methodologies is  hard to penetrate, we believe that
+it is premature to discuss issues related to  SPI. Researchers should rather
+focus on the trade-off between development speed  and accumulated technical debt
+\cite{Brown:2010:MTD:1882362.1882373}, which  appears to be the most important
+determinant for the future of  software development in the company.
+
+Future research can investigate the performance drop down caused by the
+necessity of returning the accumulated technical debt while expanding the
+company's operation and structuring mitigation strategies with four software
+development objectives: integrating scalable solutions with fast iterations and
+minimal set of functionalities (this allows startups to maintain effective
+planning and realistic expectations); empowering team members enabling them to
+operate horizontally in all the activities of the development environment
+simultaneously; improve desirable workflow patterns through the initiation of a
+minimal project management over time, as a natural result of emerging activities
+of tracing project progress and task assignment mechanisms; and finally, only
+when the chaos has been initially managed, planning long-term performance by
+adoption of Agile and Lean development practices.
+
+%------------------------------------------------------------------------------%
+%\subsection{Future work} %label{futwork} %MUN: Don't have future work as a
+subsection in threats to validity. Either  %merge it in "Implications of the
+GSM" OR into "Conclusions". CG: fixed
 
@@
 %influence the growth of a company after the release of the first product to the
-%market.
-
-%------------------------------------------------------------------------------%
-\section{Threats to Validity}
-\label{valt}
-%MUN: Also here, introduce some structure: first, explain shortly what types of 
-%validity threats exist for the research methodology you applied. Then, discuss 
-%each type individually. Think about which threat type is the most prominent 
-%one for your research methodology and start with that one. I did not review 
-%this section in detail now since it will change considerably after your 
-%restructuring.
-%CG: fixed
-
-%MUN1: After reading the discussion on validity threats, I think you have
-%chosen the wrong framework (Wohlin) to discuss threats. His framework is 
-%targeted at experiments and even though it has been applied in other contexts, 
-%other taxonomies are a better fit (see Yin and also Runeson). You tried to 
-%squeeze your threats and strategies into Wohlin's taxonomy which seems in some 
-%cases quite artificial and forced. I think it is better if you adopt Yin's 
-%taxonomy of validity threats. If you read Yin (p 33-38) you'll see that his 
-%taxonomy is a much better fit and that you can easier place and motivate your 
-%mitigation strategies.
-%------------------------------------------------------------------------------%
-In this section we discuss the validity of the overall results. %MUN1: threats 
-%are discussed w.r.t. the research methodology, not the results... 
-We structure the discussion according to Wohlin's taxonomy \cite{Wohlin2000}.
+%market. CG: ok
+
+%------------------------------------------------------------------------------%
+\section{Threats to Validity} \label{valt} %MUN: Also here, introduce some
+structure: first, explain shortly what types of  %validity threats exist for the
+research methodology you applied. Then, discuss  %each type individually. Think
+about which threat type is the most prominent  %one for your research
+methodology and start with that one. I did not review  %this section in detail
+now since it will change considerably after your  %restructuring. %CG: fixed
+
+%MUN1: After reading the discussion on validity threats, I think you have chosen
+%the wrong framework (Wohlin) to discuss threats. His framework is  targeted at
+%experiments and even though it has been applied in other contexts,  other
+%taxonomies are a better fit (see Yin and also Runeson). You tried to  squeeze
+%your threats and strategies into Wohlin's taxonomy which seems in some  cases
+%quite artificial and forced. I think it is better if you adopt Yin's  taxonomy
+%of validity threats. If you read Yin (p 33-38) you'll see that his  taxonomy is
+%a much better fit and that you can easier place and motivate your  mitigation
+%strategies. CG: to discuss since the ref of Yin and Runeson present very
+%similar validity threats.
+%------------------------------------------------------------------------------
+In this section we discuss the validity of the overall methodology. %MUN1:
+threats  %are discussed w.r.t. the research methodology, not the results... CG:
+fixed. We structure the discussion according to Wohlin's taxonomy
+\cite{Wohlin2000}.
 
@@
 %to explain (shortly) their relevance w.r.t. GT. I guess my comment above was
-%misleading...
-%\begin{compactitem}
+%misleading... \begin{compactitem}
 
@@
 %direct causal connection, but rather due to either a coincidence or the
-%presence of confounding factors.}.
-%\item Construct validity - which is enforced by clearly specifing operational 
-%procedures, and providing an adequate description of 
-%constructs\footnote{Constructs are complex ideas that we as humans form in 
-%order to summarize observations about things that we cannot see directly. For 
-%example a construct is \textit{``job satisfaction''}, which to be defined, 
-%requires a higher level of abstraction than a concrete object(e.g. a book).} 
-%that are used to generate the theory.
-%\item Conclusion validity - which represent the degree to which conclusions 
-%are reasonable, according to the defined relationships in the data. It 
-%involves ensuring adequate sampling procedures, appropriate analytical methods 
-%and reliable measurement procedures. Reliability requires to demonstrate that 
-%the operations of a study - such as data collection procedures - can be 
-%repeated by other researchers, obtaining the same results.
-%\end{compactitem}
+%presence of confounding factors.}. %\item Construct validity - which is
+enforced by clearly specifing operational  %procedures, and providing an
+adequate description of  %constructs\footnote{Constructs are complex ideas that
+we as humans form in  %order to summarize observations about things that we
+cannot see directly. For  %example a construct is \textit{``job satisfaction''},
+which to be defined,  %requires a higher level of abstraction than a concrete
+object(e.g. a book).}  %that are used to generate the theory. %\item Conclusion
+validity - which represent the degree to which conclusions  %are reasonable,
+according to the defined relationships in the data. It  %involves ensuring
+adequate sampling procedures, appropriate analytical methods  %and reliable
+measurement procedures. Reliability requires to demonstrate that  %the
+operations of a study - such as data collection procedures - can be  %repeated
+by other researchers, obtaining the same results. %\end{compactitem}
 
@@
 \subsection{External validity}
-%External
-One threat to external validity is the selection of subjects interviewed 
-for the study. This threat profoundly affects GT, since it is a qualitative 
+
+One threat to external validity is the selection of subjects interviewed  for
+the study. This threat profoundly affects GT, since it is a qualitative
 research method, which makes the use of semi-structured interviews, and centers
-the research on respondent's opinions. 
-In order to mitigate this threat we selected interviewees which 
-covered positions of CTOs' and CEOs'. Their perspectives on what is taking 
-place within the organization, were the only meaningful data taken in 
-consideration in the study.  In many cases there was no supporting evidence to 
-verify the opinion expressed. The questionnaire dedicated to ``data 
-triangulation'' was not of primary support because of the limited time 
-practitioners were able to grant to it. However, researchers have to accept the 
-veracity of what respondents reported during the interviews \cite{Coleman2008a}. 
-%MUN1: candidate for removal
-%Notwithstanding, the issues surrounding semi-structured interviews are vital. 
-%Within our study, although the reality might be different from what has been 
-%described, the reported data are CTOs' and CEOs' perception, which is used to 
-%base their decisions.
+the research on respondent's opinions.  In order to mitigate this threat we
+selected interviewees which  covered positions of CTOs' and CEOs'. Their
+perspectives on what is taking  place within the organization, were the only
+meaningful data taken in  consideration in the study.  In many cases there was
+no supporting evidence to  verify the opinion expressed. The questionnaire
+dedicated to ``data  triangulation'' was not of primary support because of the
+limited time  practitioners were able to grant to it. However, researchers have
+to accept the  veracity of what respondents reported during the interviews
+\cite{Coleman2008a}.  %MUN1: candidate for removal %Notwithstanding, the issues
+surrounding semi-structured interviews are vital.  %Within our study, although
+the reality might be different from what has been  %described, the reported data
+are CTOs' and CEOs' perception, which is used to  %base their decisions. CG: ok
 
@@
 domain to which the study's findings can be generalized. In particular the
-previous framework developed by Coleman \cite{Coleman2008} has allowed 
-us to find similarities, differences and broader reasoning related to 
-factors that hinder maturing processes in startups. Moreover, we analyzed 
-Brook's framework describing the \textit{basic attacks} to the challenges of 
-developing software systems and other literature covered by the SMS on startups.
+previous framework developed by Coleman \cite{Coleman2008} has allowed  us to
+find similarities, differences and broader reasoning related to  factors that
+hinder maturing processes in startups. Moreover, we analyzed  Brook's framework
+describing the \textit{basic attacks} to the challenges of  developing software
+systems and other literature covered by the SMS on startups.
 \cite{SMS}\footnote{Note that the generalizability of results is applied only to
-those startups that operate in software-intensive activities.}. 
-%MUN1: candidate for removal: detail on RM / seems out of place
-%In conclusion, regarding the validation of 
-%the subcategories, a continuous re-evaluation of the interview transcripts has 
-%been performed.
-
-%------------------------------------------------------------------------------%
-\subsection{Internal validity}
-%Internal
-To enhance the internal validity, we created a three-dimensional research 
-framework. By means of a grounded theory approach, formed by a 
-\textit{systematic mapping study}, \textit{interviews} and \textit{follow-up 
-questionnaires}, we searched for convergence among different sources of 
-information to confirm the generated theory.
+those startups that operate in software-intensive activities.}.  %MUN1:
+candidate for removal: detail on RM / seems out of place %In conclusion,
+regarding the validation of  %the subcategories, a continuous re-evaluation of
+the interview transcripts has  %been performed. CG: ok
+
+%------------------------------------------------------------------------------%
+\subsection{Internal validity} %Internal To enhance the internal validity, we
+created a three-dimensional research  framework. By means of a grounded theory
+approach, formed by a  \textit{systematic mapping study}, \textit{interviews}
+and \textit{follow-up  questionnaires}, we searched for convergence among
+different sources of  information to confirm the generated theory.
 
@@
 stated: ``Tying the emergent theory to existing literature enhances the internal
-validity, generalizability, and theoretical level of the theory building from 
-a case study research [\ldots] because the findings often rest on a very limited 
+validity, generalizability, and theoretical level of the theory building from  a
+case study research [\ldots] because the findings often rest on a very limited
 number of cases.''
 
-%MUN1: candidate for shortening
-%Little ambiguities about direction of causal influences have been reduced by 
-%introducing arrows to indicate the direction of the causal relation in the GSM 
-%model. Furthermore, by means of comparison between the GSM model and extant 
-%studies in the broader fields, we revealed that the theory resembles in wider 
-%startup development context. 
+%MUN1: candidate for shortening %Little ambiguities about direction of causal
+influences have been reduced by  %introducing arrows to indicate the direction
+of the causal relation in the GSM  %model. Furthermore, by means of comparison
+between the GSM model and extant  %studies in the broader fields, we revealed
+that the theory resembles in wider  %startup development context. CG: ok
 Nevertheless, we identified important confounding factors, related to
-innovation, market requirements and developer experience, which might 
-impact the internal validity of the GSM.
-
-%------------------------------------------------------------------------------%
-\subsection{Construct validity}
-%Construct
-One threat to this study is a possible inadequate description of constructs. In 
-order to diminish this risk, the entire study constructs have been adapted to 
-the terminology utilized by practitioners and defined at an adequate level for 
-each theoretical conceptualization. For instance, we defined \textit{Time 
-shortage} in terms of \textit{Investor pressure, CEO/business 
-pressure, Demo presentations at events and internal final deadline} as used by 
-most of the interviewees in the study. 
+innovation, market requirements and developer experience, which might  impact
+the internal validity of the GSM.
+
+%------------------------------------------------------------------------------%
+\subsection{Construct validity} %Construct One threat to this study is a
+possible inadequate description of constructs. In  order to diminish this risk,
+the entire study constructs have been adapted to  the terminology utilized by
+practitioners and defined at an adequate level for  each theoretical
+conceptualization. For instance, we defined \textit{Time  shortage} in terms of
+\textit{Investor pressure, CEO/business  pressure, Demo presentations at events
+and internal final deadline} as used by  most of the interviewees in the study.
 Moreover, during the coding of interview transcripts, we adopted explanatory
@@
 %MUN1: I get the threat, but I don't understand its mitigation: let the
-%interviewees drive the study?! Try to reformulate...
-The second important threat is caused by the fact that interviewees might 
-already be aware of the possible emergent theories analyzed by researchers. To 
-reduce the risk that some hypothesis could have been involved in the design of 
-the research, we let mainly participants drive the direction of the case study.
-
-Another threat associated with construct validity is evaluation 
-apprehension, which describes people's rejection attitude towards the 
-assessment of their behavior. 
-To address this risk we didn't attained to first answers from practitioners, but 
-rather went towards details of engineering elements and activities. %MUN1: <-- 
-%I don't understand this...
-Moreover, we developed a rigorous data collection protocol to create case study 
-databases and employ integration of multiple data collection methods.%MUN1: <-- 
-%Since you write "Moreover, ..." the reader gets the impression that this is a 
+%interviewees drive the study?! Try to reformulate... CG: fixed. The second
+important threat is caused by the fact that interviewees might  already be aware
+of the possible emergent theories analyzed by researchers. To  reduce the risk
+that some hypothesis could have been involved in the design of  the research, we
+didn't disclose any goal and emergent results of our research to the
+interviewees.
+
+Another threat associated with construct validity is evaluation  apprehension,
+which describes people's rejection attitude towards the  assessment of their
+behavior.  To address this risk we didn't attained to first answers from
+practitioners, but  rather we asked for details of artifacts and engineering
+strategies. %MUN1: <--  %I don't understand this... CG: fixed %Moreover, we
+developed a rigorous data collection protocol to create case study  %databases
+and employ integration of multiple data collection methods.%MUN1: <--  %Since
+you write "Moreover, ..." the reader gets the impression that this is a
 %further mitigation strategy to evaluation apprehension. I don't see however
-%how this mitigates this risk. Fix...
-Reporting bias was mitigated by packaging all the needed 
-material for conducting new researches in other contexts in order to criticize 
-our findings (interview package with instructions has been made available 
-online\footnote{Available at 
-\url{https://github.com/adv0r/BTH-Interview-Package}.}. %MUN1: --> What exactly 
-%is reporting bias? Let's assume that it means that you only report results, 
-%i.e. interviewee statements, that are favorable to your theory. How does 
-%publishing the auxiliary study material help to mitigate this risk? Publishing 
-%the raw data (what you not do) would only allow researchers to check the 
-%conclusions you made. The published auxiliary material helps to replicate the 
-%study, i.e. makes is possible to test your theory in a different context. 
-Moreover, two researchers not involved in the execution of the study conducted 
-a peer-review analysis of the model's constructs.
+%how this mitigates this risk. Fix... CG: right, removed because misleading.
+Reporting bias was mitigated by packaging all the needed  material for
+conducting new researches in other contexts in order to criticize  our findings
+(interview package with instructions has been made available
+online\footnote{Available at  \url{https://github.com/adv0r/BTH-Interview-
+Package}.}. %MUN1: --> What exactly  %is reporting bias? Let's assume that it
+means that you only report results,  %i.e. interviewee statements, that are
+favorable to your theory. How does  %publishing the auxiliary study material
+help to mitigate this risk? Publishing  %the raw data (what you not do) would
+only allow researchers to check the  %conclusions you made. The published
+auxiliary material helps to replicate the  %study, i.e. makes is possible to
+test your theory in a different context. CG: by this, we meant providing all the
+material to replicate the same study with other startups and compare the
+results. Moreover, two researchers not involved in the execution of the study
+conducted  a peer-review analysis of the model's constructs.
 
@@
 
-\begin{compactitem}
-\item Each category and the strength of relations between them.
-\item Hypotheses, derived from and related to the emergent theory.
-\item Deviant cases to ensure robustness and general applicability.
-\end{compactitem}
+\begin{compactitem} item Each category and the strength of relations between
+\them. item Hypotheses, derived from and related to the emergent theory. item
+\Deviant cases to ensure robustness and general applicability. end{compactitem}
 
@@
 
-According to the theoretical sampling concept, we adjusted our 
-research design and the emergent theory until only marginal results were 
-generated. Moreover to enhance reliability of the outcome conceptualizations and 
-relations, the process of coding interviews has been systematically conducted 
-following a detailed process. Then, we asked to other practitioners a 
-confirmation to further validate the generated framework and enhance the 
-reliability of the relations among the identified categories.
+According to the theoretical sampling concept, we adjusted our  research design
+and the emergent theory until only marginal results were  generated. Moreover to
+enhance reliability of the outcome conceptualizations and  relations, the
+process of coding interviews has been systematically conducted  following a
+detailed process. Then, we asked to other practitioners a  confirmation to
+further validate the generated framework and enhance the  reliability of the
+relations among the identified categories.
 
@@
 
-The  approach of generating theory allowed us to check emerging categories 
-and their properties by gathering new evidence within an iterative and 
-evolutionary approach, in view of the wide context of the research area that was 
-under investigation.
-
-%------------------------------------------------------------------------------%
-%\subsection{Subsection Heading Here}
-%Subsection text here.
-
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
-
-%\subsubsection{Subsubsection Heading Here}
-%Subsubsection text here.
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
+The  approach of generating theory allowed us to check emerging categories  and
+their properties by gathering new evidence within an iterative and  evolutionary
+approach, in view of the wide context of the research area that was  under
+investigation.
+
+%------------------------------------------------------------------------------%
+%\subsection{Subsection Heading Here} %Subsection text here.
+
+% needed in second column of first page if using \IEEEpubid %\IEEEpubidadjcol
+
+%\subsubsection{Subsubsection Heading Here} %Subsubsection text here.
+
+% An example of a floating figure using the graphicx package. % Note that \label
+must occur AFTER (or within) \caption. % For figures, \caption should occur
+after the \includegraphics. % Note that IEEEtran v1.7 and later has special
+internal code that % is designed to preserve the operation of \label within
+\caption % even when the captionsoff option is in effect. However, because % of
+issues like this, it may be the safest practice to put all your % \label just
+after \caption rather than within \caption{}. % % Reminder: the "draftcls" or
+"draftclsnofoot", not "draft", class % option should be used if it is desired
+that the figures are to be % displayed while in draft mode. %
 
 % Note that IEEE CS typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-% However, the Computer Society has been known to put floats at the bottom.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
+% results in a large percentage of a column being occupied by floats. However,
+% the Computer Society has been known to put floats at the bottom.
+
+% An example of a double column floating figure using two subfigures. % (The
+subfig.sty package must be loaded for this to work.) % The subfigure \label
+commands are set within each subfloat command, the % \label for the overall
+figure must come after \caption. % \hfil must be used as a separator to get
+equal spacing. % The subfigure.sty package works much the same way, except
+\subfigure is % used instead of \subfloat. % %\begin{figure*}[!t]
 %\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE CS papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
+%\label{fig_first_case}} %\hfil %\subfloat[Case
+II]{\includegraphics[width=2.5in]{subfigcase2}% %\label{fig_second_case}}}
+%\caption{Simulation results} %\label{fig_sim} %\end{figure*} % % Note that
+often IEEE CS papers with subfigures do not employ subfigure % captions (using
+the optional argument to \subfloat), but instead will % reference/describe all
+of them (a), (b), etc., within the main caption.
+
+% An example of a floating table. Note that, for IEEE style tables, the  %
+\caption command should come BEFORE the table. Table text will default to %
+\footnotesize as IEEE normally uses this smaller font for tables. % The \label
+must come after \caption as always. % %\begin{table}[!t] %% increase table row
+spacing, adjust to taste %\renewcommand{\arraystretch}{1.3} % if using
+array.sty, it might be a good idea to tweak the value of % \extrarowheight as
+needed to properly center the text within the cells %\caption{An Example of a
+Table} %\label{table_example} %\centering %% Some packages, such as MDW tools,
+offer better commands for making tables %% than the plain LaTeX2e tabular which
+is used here. %\begin{tabular}{|c||c|} %\hline %One & Two\\ %\hline %Three &
+Four\\ %\hline %\end{tabular} %\end{table}
 
@@
 % positioning is not used. Most IEEE journals use top floats exclusively.
-% However, Computer Society journals sometimes do use bottom floats - bear
-% this in mind when choosing appropriate optional arguments for the
-% figure/table environments.
-% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
-% floats. This can be corrected via the \fnbelowfloat command of the
+% However, Computer Society journals sometimes do use bottom floats - bear this
+% in mind when choosing appropriate optional arguments for the figure/table
+% environments. Note that, LaTeX2e, unlike IEEE journals, places footnotes above
+% bottom floats. This can be corrected via the \fnbelowfloat command of the
 % stfloats package.
 
-\section{Conclusion}
-\label{conc}
-Startups are able to produce cutting-edge software products with a wide impact 
-on the market, significantly contributing to the global economy. Software 
-development, especially in the early-stage, is at the core of the company's 
-daily activities. Despite their severely high failure-rate, the SMS \cite{SMS} 
-found that the quick proliferation of startups is not supported by an adequate 
-and scientific body of knowledge. %MUN: <-- last sentence: isn't that a 
-%conclusion from the SMS? CG: fixed
-To be able to intervene on the software development strategy with scientific and 
-engineering approaches, the first step is to understand startups' behavior. 
-Hence, in this research work, we provided an initial explanation of the 
-underlying phenomenon by means of a grounded theory case study, focusing on 
-early engineering activities, from the idea conception to the first open beta 
-release of the product.
+\section{Conclusion} \label{conc} Startups are able to produce cutting-edge
+software products with a wide impact  on the market, significantly contributing
+to the global economy. Software  development, especially in the early-stage, is
+at the core of the company's  daily activities. Despite their severely high
+failure-rate, the SMS \cite{SMS}  found that the quick proliferation of startups
+is not supported by an adequate  and scientific body of knowledge. %MUN: <--
+last sentence: isn't that a  %conclusion from the SMS? CG: fixed To be able to
+intervene on the software development strategy with scientific and  engineering
+approaches, the first step is to understand startups' behavior.  Hence, in this
+research work, we provided an initial explanation of the  underlying phenomenon
+by means of a grounded theory case study, focusing on  early engineering
+activities, from the idea conception to the first open beta  release of the
+product.
 
@@
 have been validated by means of systematic comparisons with the state-of-the-art
-results and empirical data. The systematic mapping study \cite{SMS}, 
-with 37 studies extracted, revealed a 
-multi-faceted state-of-the-art inadequate to support software development 
-activities in startup companies. %MUN: <-- you mix again SMS and GT here... fix 
-%it CG: not fixed, I would keep it since important part of the initial GT 
-%methodology
-%MUN1: Ok, I understand that. But this paper is about the GT, ONLY the GT. The 
-%SMS is published elsewhere. Maybe it is ok how you use the SMS now... let's 
-%see what Tony says...
-On the other hand, the case study conducted in 13 early-stage startups, 
-provided a broad set of empirical evidences obtained by combining different 
-research methodologies in a grounded theory approach.
-
-The overall results of our research confirm that startups possess %MUN: 
-%why unique? CG: fixed.
-characteristics of uncertainty, lack of resources and time-pressure. These 
-factors influence the software development to an extent that transforms every 
-decision related to the development strategies into a difficult trade-off for 
-the company. Moreover, although startups share different characteristics with 
-other similar SE domains (e.g. market-driven development, small companies and 
-web engineering), the unique combination of coexisting factors poses a whole new 
-set of challenges which need to be addressed by primary studies. Especially when 
-bringing the first product to market, startups' most urgent priority is 
-releasing the product as quickly as possible to verify the product/market fit 
-and to adjust the business and product trajectory according to early feedback 
-and collected metrics. In this stage, startups often discard any formal project 
-management, documentation, analysis, planning, testing and other traditional 
-process activities. Practitioners take advantage of an evolutionary prototyping 
-approach, using well-integrated tools and externalizing complexity to third 
-party solutions.
+results and empirical data. The systematic mapping study \cite{SMS},  with 37
+studies extracted, revealed a  multi-faceted state-of-the-art inadequate to
+support software development  activities in startup companies. %MUN: <-- you mix
+again SMS and GT here... fix  %it CG: not fixed, I would keep it since important
+part of the initial GT  %methodology %MUN1: Ok, I understand that. But this
+paper is about the GT, ONLY the GT. The  %SMS is published elsewhere. Maybe it
+is ok how you use the SMS now... let's  %see what Tony says... CG: ok On the
+other hand, the case study conducted in 13 early-stage startups,  provided a
+broad set of empirical evidences obtained by combining different  research
+methodologies in a grounded theory approach.
+
+The overall results of our research confirm that startups possess %MUN:  %why
+unique? CG: fixed. characteristics of uncertainty, lack of resources and time-
+pressure. These  factors influence the software development to an extent that
+transforms every  decision related to the development strategies into a
+difficult trade-off for  the company. Moreover, although startups share
+different characteristics with  other similar SE domains (e.g. market-driven
+development, small companies and  web engineering), the unique combination of
+coexisting factors poses a whole new  set of challenges which need to be
+addressed by primary studies. Especially when  bringing the first product to
+market, startups' most urgent priority is  releasing the product as quickly as
+possible to verify the product/market fit  and to adjust the business and
+product trajectory according to early feedback  and collected metrics. In this
+stage, startups often discard any formal project  management, documentation,
+analysis, planning, testing and other traditional  process activities.
+Practitioners take advantage of an evolutionary prototyping  approach, using
+well-integrated tools and externalizing complexity to third  party solutions.
 
@@
 %of chaotic/complex domain in the paper, I think not. Hence you need to
-%rephrase that (but you can not introduce new concepts here in the conclusion). %CG: fixed. 
-Therefore, from a software development perspective, working with low-precision 
-engineering activities represents the only viable strategy that startups can 
-follow in the early stage of their lifecycle.
+%rephrase that (but you can not introduce new concepts here in the conclusion).
+%CG: fixed.  Therefore, from a software development perspective, working with
+low-precision  engineering activities represents the only viable strategy that
+startups can  follow in the early stage of their lifecycle.
 
@@
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
-% use section* for acknowledgement
-\ifCLASSOPTIONcompsoc
-  % The Computer Society usually uses the plural form
-  \section*{Acknowledgments}
-\else
-  % regular IEEE prefers the singular form
-  \section*{Acknowledgment}
+% if have a single appendix: %\appendix[Proof of the Zonklar Equations] % or
+%\appendix  % for no appendix heading % do not use \section anymore after
+\appendix, only \section* % is possibly needed
+
+% use appendices with more than one appendix % then use \section to start each
+appendix % you must declare a \section before using any % \subsection or using
+\label (\appendices by itself % starts a section numbered zero.) %
+
+% use section* for acknowledgement \ifCLASSOPTIONcompsoc   % The Computer
+Society usually uses the plural form   \section*{Acknowledgments} \else   %
+regular IEEE prefers the singular form   \section*{Acknowledgment} \fi
+
+The authors would like to thank...
+
+% Can use something like this to put references on a page % by themselves when
+using endfloat and the captionsoff option. \ifCLASSOPTIONcaptionsoff   \newpage
 \fi
 
-The authors would like to thank...
-
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
+% trigger a \newpage just before the given reference % number - used to balance
+the columns on the last page % adjust value as needed - may need to be
+readjusted if % the document is modified later %\IEEEtriggeratref{8} % The
+"triggered" command can be changed if desired:
 %\IEEEtriggercmd{\enlargethispage{-5in}}
@@
 
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-\bibliography{tse}
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{John Doe}
-Biography text here.
+% can use a bibliography generated by BibTeX as a .bbl file % BibTeX
+documentation can be easily obtained at: % http://www.ctan.org/tex-
+archive/biblio/bibtex/contrib/doc/ % The IEEEtran BibTeX style support page is
+at: % http://www.michaelshell.org/tex/ieeetran/bibtex/
+\bibliographystyle{IEEEtran} % argument is your BibTeX string definitions and
+bibliography database(s) \bibliography{tse}
+
+% if you will not have a photo at all: \begin{IEEEbiographynophoto}{John Doe}
+Biography text here. \end{IEEEbiographynophoto}
+
+% insert where needed to balance the two columns on the last page with %
+biographies %\newpage
+
+\begin{IEEEbiographynophoto}{Jane Doe} Biography text here.
 \end{IEEEbiographynophoto}
 
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-\begin{IEEEbiographynophoto}{Jane Doe}
-Biography text here.
-\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
+% You can push biographies down or up by placing a \vfill before or after them.
+% The appropriate use of \vfill depends on what kind of text is on the last page
+% and whether or not the columns are being equalized.
 
@@
 
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
-\end{document}
+% Can be used to pull up biographies so that the bottom of the last one % is
+flush with the other column. %\enlargethispage{-5in}
+
+% that's all folks \end{document}

### 2013-03-04 article/main.tex
Compiles again...

--- 2013-02-26 article/main.tex
@@
 the research focused at studying technical debt \cite{Nugroho2011,Izurieta2012}.
-On the other hand, researchers can take advantage from the GSM to 
-understand how the technical debt influences the future growth of startup 
-companies.
+On the other hand, researchers can take advantage from the GSM to  understand
+how the technical debt influences the future growth of startup companies.
 
@@
 %for the reader. Furthermore, it conflicts with you convention you use to  
-%highlight high level categories of GSM.
+%highlight high level categories of GSM. CG: ok. 
 He noticed how these companies were particularly innovative and successful,  
 advocating for the need of more research investigating software development  
-practices so as to replicate success and try to transfer it to other technology 
-sectors.
+practices so as to replicate success and try to transfer it to other 
+technology sectors.
 
@@
 Additionally, the role of SPI has been neglected because seen as an obstacle to
-the development team's creativity and flexibility, and to 
-the need of a quick delivering product process environment \cite{Coleman2008a}.
+the development team's creativity and flexibility, and to  the need of a quick
+delivering product process environment \cite{Coleman2008a}. 
 %MUN: the second part of the sentence is just a statement without reference. If 
 %it refers to a statement by Coleman, put his reference at the end OR find the  
-%correct reference OR delete the statement.
-%CG fixed
+%correct reference OR delete the statement. %CG fixed 
 Product quality is left aside in favor of minimal and suitable functionalities  
@@
 \cite{Hanna2010}, software reuse \cite{Jansen2008} and open-source strategies ,
-\cite{Wall2001, Bean2005}. 
-%MUN: Can you split up the references? I.e. write:
+\cite{Wall2001, Bean2005}.  %MUN: Can you split up the references? I.e. write:
 %"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source
-%strategies \cite{Bla}.
-%CG: fixed
-%------------------------------------------------------------------------------%
-\subsection{Technical debt}
-Notwithstanding, a new area of the SE research, trying to tackle the problem of 
-the \textit{technical debt},  brings and encompasses different implications in 
-studying development in software startups \cite{Tom2013}. The metaphoric 
-neologism was originally introduced by Cunningham in 1992 
-\cite{TechnicalDebtCunn} and has recently attracted the attention of SE 
-researchers\footnote{To attest the fact that \textit{technical debt} is gaining 
-traction among researchers, we mention two important contributions which 
-characterize the \textit{``debt landscape''} : \cite{Nugroho2011,Izurieta2012} 
-and a dedicated workshop \cite{workshopDebt} organized by the Software 
-Engineering Institute  and ICSE.}. The concept of \textit{technical debt} is 
-well illustrated in \cite{Brown:2010:MTD:1882362.1882373} : \textit{``The idea 
-is that developers sometimes accept compromises in a system in one dimension 
-(e.g., modularity) to meet an urgent demand in some other dimension (e.g., a 
-deadline), and that such compromises incur a ``debt'' on which ``interest'' has 
-to be paid and which the ``principal'' should be repaid at some point for the 
-long-term health of the project''}.  The compromise between high-speed and 
-high-quality engineering is faced daily by software startups, not only in terms 
-of architecture design but in multifaceted aspects (weak project management, 
-testing, process control).
+%strategies \cite{Bla}. %CG: fixed
+
+%------------------------------------------------------------------------------%
+\subsection{Technical debt} Notwithstanding, a new area of SE research,
+trying to tackle the problem of the \textit{technical debt}, brings and
+encompasses different implications in studying development in software startups
+\cite{Tom2013}. The metaphoric neologism was originally introduced by
+Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the
+attention of SE researchers\footnote{To attest the fact that \textit{technical
+debt} is gaining  traction among researchers, we mention two important
+contributions which characterize the \textit{``debt landscape''}:
+\cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt}
+organized by the Software  Engineering Institute  and ICSE.}. The concept of
+\textit{technical debt} is  well illustrated in
+\cite{Brown:2010:MTD:1882362.1882373}: \textit{``The idea  is that developers
+sometimes accept compromises in a system in one dimension  (e.g., modularity) to
+meet an urgent demand in some other dimension (e.g., a  deadline), and that such
+compromises incur a ``debt'' on which ``interest'' has  to be paid and which the
+``principal'' should be repaid at some point for the long-term health of the
+project''}.  The compromise between high-speed and high-quality engineering is
+faced daily by software startups, not only in terms  of architecture design but
+in multifaceted aspects (weak project management, testing, process control).
 
@@
 
-\subsection{Terminology}
-In order to set a common ground and to prevent ambiguity, we use the following 
-terminology throughout the paper:
+\subsection{Terminology} In order to set a common ground and to prevent
+ambiguity, we use the following  terminology throughout the paper:
 
@@
 supporting the engineering activities.
-\item Quality attributes: those overall factors that affect run-time behavior, 
-system design, and user experience. They represent areas of concern that have 
-the potential for applications to impact across layers and tiers. Some of these 
-attributes are related to the overall system design, while others are specific 
-to run time, design time, or user centric issues \cite{Microsoft2009}.
-\item Growth: an increase in the company size respect to the initial conditions 
-in terms of either employees or users/customers, and product complexity in terms 
-of handling an increasing number of feature requests.
+\item Quality attributes: those overall factors that affect run-time
+behavior, system design, and user experience. They represent areas of concern
+that have  the potential for applications to impact across layers and tiers.
+Some of these  attributes are related to the overall system design, while others
+are specific  to run time, design time, or user centric issues
+\cite{Microsoft2009}. 
+\item Growth: an increase in the company size respect to the initial 
+conditions in terms of either employees or users/customers, and
+product complexity in terms  of handling an increasing number of feature
+requests. 
 \item Software product: any software product and software service.
 \end{compactitem}
-%------------------------------------------------------------------------------%
-\section{Research methodology}
-\label{resmet}
-
-\begin{figure*}[!t]
-\centering
-\includegraphics[width=5in]{figures/completemethodology2}
-\caption{Grounded theory study process overview}
-\label{fig:gt:completemethodology}
-\end{figure*}
+
+%------------------------------------------------------------------------------%
+\section{Research methodology} \label{resmet}
+
+\begin{figure*}[!t] \centering
+\includegraphics[width=5in]{figures/completemethodology2} \caption{Grounded
+theory study process overview} \label{fig:gt:completemethodology} \end{figure*}
 %MUN: I think you can increase the font size in the figure a bit. Try thereby
@@
 %you name it     in the text. Also, add a section (3.5) that describes shortly
-%what you did in "Theory  Validation" 
-%CG: fixed... fontsize cannot be increased uniformly among all the boxes, then 
-%I skipped that part...
+%what you did in "Theory  Validation"  %CG: fixed... fontsize cannot be
+%increased uniformly among all the boxes, then  I skipped that part... 
 %MUN1: OK, but in the figure, the correct section number for theory validation  
-%is 3.5, not 3.4
+%is 3.5, not 3.4. CG:fixed
 
@@
 development strategy in early-stage startups?
+
   \begin{compactitem}
-
   \item RQ-1.1 : How do startups structure and execute the main engineering
     activities? 
-\item RQ-1.2 : How are product quality attributes considered by startups?
+  \item RQ-1.2 : How are product quality attributes considered by
+    startups? 
   \end{compactitem}
@@
 last interview.    
-%MUN: I don't understand this: you
-%collected the data from questionnaires after ALL interviews have been 
-%finished, but on the other hand you said the overall process was iterative 
-%(data analysis and triangulation let themes emerge which in turn were used
-%in subsequent interviews). 
+%MUN: I don't understand this: you %collected the data from questionnaires 
+%after ALL interviews have been finished, but on the other hand
+%you said the overall process was iterative  (data analysis and triangulation
+%let themes emerge which in turn were used %in subsequent interviews).  
 %CG: fixed...
@@
 %------------------------------------------------------------------------------%
-\subsection{Data collection}
-The data collection was designed to allow triangulation, which integrates 
-multiple data sources converging on the same phenomenon. After transcribing the 
-interview, we extracted conceptual relations, integrating them with the 
-questionnaire's results. A well structured case-study database allowed us to 
-easily retrieve and seek for information, assembling the evidence from the case 
-study reports, as described also in \cite{Yin1994}. The database has been stored 
-and constructed using the qualitative data analysis software package 
-AtlasTI\footnote{Available online at \url{http://www.atlasti.com/}.} - one of 
-the most suitable tools for GT \cite{Coleman2007}. 
-We overlapped interviews with questionnaire results to adjust the data 
-collection and take advantage of emergent themes and reveal possible 
-inconsistencies. The data was analyzed simultaneously and with flexibility in 
-mind in such a way that adjustments were made according to the emerging 
-findings.
-%------------------------------------------------------------------------------%
-\subsection{Data analysis}
-Before starting the analysis, a data ordering procedure was necessary since 
-interviews were spread across a multitude of topics. Therefore transcripts have 
-been structured in thematic areas accordingly to different topic cards used 
-during the interviews. We proceeded horizontally between same thematic areas of 
-transcripts to be able to identify a better number of similar concepts, rather 
-than going through an entire transcript at one time.
+\subsection{Data collection} The data collection was designed to allow
+triangulation, which integrates  multiple data sources converging on the same
+phenomenon. After transcribing the  interview, we extracted conceptual
+relations, integrating them with the  questionnaire's results. A well structured
+case-study database allowed us to  easily retrieve and seek for information,
+assembling the evidence from the case  study reports, as described also in
+\cite{Yin1994}. The database has been stored  and constructed using the
+qualitative data analysis software package  AtlasTI\footnote{Available online at
+\url{http://www.atlasti.com/}.} - one of  the most suitable tools for GT
+\cite{Coleman2007}.  We overlapped interviews with questionnaire results to
+adjust the data  collection and take advantage of emergent themes and reveal
+possible  inconsistencies. The data was analyzed simultaneously and with
+flexibility in  mind in such a way that adjustments were made according to the
+emerging  findings.
+%------------------------------------------------------------------------------%
+\subsection{Data analysis} Before starting the analysis, a data ordering
+procedure was necessary since  interviews were spread across a multitude of
+topics. Therefore transcripts have  been structured in thematic areas
+accordingly to different topic cards used  during the interviews. We proceeded
+horizontally between same thematic areas of  transcripts to be able to identify
+a better number of similar concepts, rather  than going through an entire
+transcript at one time.
 
@@
 management tool. 
-\item Consistency among categories was validated by exploring and analyzing 
-links among subcategories by means of selective coding.
+\item Consistency among categories was validated by exploring and
+analyzing  links among subcategories by means of selective coding. 
 \item The core category - the one with the greatest explanatory power - was  
@@
 %------------------------------------------------------------------------------%
-\subsection{Theory Validation}
-\label{rm:val}
+\subsection{Theory Validation} \label{rm:val} 
 %MUN: put here what I wrote in the comment in Section "Theory validation" (i.e. 
@@
 %You need to explain why you did this and how it contributes to the validation  
-%of the theory.
+%of the theory. CG: fixed, removing the initial verbose part.
 
@@
 and determine if the GT is sufficiently grounded, we used a systematic technique
-to assess the theory, following evaluation lists and factors presented in  
-\cite{Strauss1998}.
-
-We consider the following factors:
-
-\par\vspace {0.4cm}
-\begin{compactitem}
-\item \textit{Fit}, that describes the consistency of the theory with the real 
-data.
-\item \textit{Understanding}, that describes the degree to which participants 
-agree with the theory.
-\item \textit{Generality}, that describes the level of abstraction of the 
-theory, and how sufficiently it can be adopted by practitioners without losing 
-its relevance.
-\item \textit{Control}, that describes how well the theory enables 
-practitioners to fully understand the described situation.
-\end{compactitem}
-\par\vspace {0.4cm}
-
-In order to achieve a high-level of fit between the theory and the raw 
-empirical data, we started GT without any assumption on how the emergent 
-patterns would influence the research direction. Ideas and perceptions evolved 
-during the mapping study and interviews. In addition, constant comparative 
-methods, overturning of some early categories as new data came to light, 
-generation and testing of interim hypotheses, and constant re-evaluation of the 
-interview transcripts, ensured that researchers' bias was minimized and 
-theoretical fit maintained.
-
-To represent understandable emergent theories, concepts and categories were 
-carefully developed using appropriate and self-explanatory codes. The developing 
-theory was not immediately presented to the participants to prevent potential 
-bias in the responses. But rather, the theoretical framework has been presented 
-and described to startups practitioners not involved in the data collection, 
-whose reaction was positively expressed.
-
-From generality perspective, we developed a theory from abstract categories, 
-which allowed us to create a general guide to constantly changing situations. 
-Nevertheless, we took attention on maintaining a reasonable level of 
-abstraction, keeping focus on their sensitizing characteristics and providing 
-low-level subgroup of categories.
-
-Finally, we provided a comprehensive set of categories with detailed 
-interrelations to explain how the \textit{core-category} is achieved and the 
-reasons of its consequences, in order to enable practitioners to independently 
-understand and analyze situations, predict changes and their consequences, and 
-to be capable of revising actions or the theory itself. Using both 
-methodological tools and features provided by \textit{AtlasTI}, we explored and 
-tested:
-
-\par\vspace {0.4cm}
-\begin{compactitem}
-\item Each category and the strength of relations between them.
-\item Hypotheses, derived from and related to the emergent theory.
-\item Deviant cases to ensure robustness and general applicability.
-\end{compactitem}
-\par\vspace {0.4cm}
-
-Additionally, Strauss and Corbin provided a list of questions to assist in 
-determining how well the findings are grounded \cite{Strauss1998}. They 
-questions are as follows:
+to assess the theory.
+
+Strauss and Corbin provided a list of questions to assist in  determining how
+well the findings are grounded \cite{Strauss1998}. They  questions are as
+follows:
 
@@
 
-%------------------------------------------------------------------------------%
-
-\section{Results: Green-field startup model}
-\label{res:gsm}
+\section{Results: Green-field startup model}\label{res:gsm}
 
@@
 in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e.
-CATx) to facilitate the reading of this section.
-The network is centered around the core category, \textit{speed-up 
-development}, which is the most interconnected node in the framework reflecting 
-the fact that \textit{``it is the one (category) with the greatest explanatory 
-power''}  \cite{Strauss1998}. The meaning of the labels on the three arrows that 
-reach the core category (efficiency, effectiveness and performance) is discussed 
-in detail in section \ref{res:val}. %MUN: this refers to the validity threats 
-%section. Do you mean rather the "Theory Validation" section?
-%CG:fixed
+CATx) to facilitate the reading of this section. The network is centered around
+the core category, \textit{speed-up  development}, which is the most
+interconnected node in the framework reflecting  the fact that \textit{``it is
+the one (category) with the greatest explanatory  power''}  \cite{Strauss1998}.
+The meaning of the labels on the three arrows that  reach the core category
+(efficiency, effectiveness and performance) is discussed  in detail in section
+\ref{sect:theory:validation:stats}. %MUN: this refers to the validity threats
+%section. Do you mean rather the "Theory Validation" section? %CG:fixed 
 %MUN1: this should be \ref{sect:theory:validation:stats}, right? Section 5 is  
-%quite long and you should point to the correct subsection
+%quite long and you should point to the correct subsection. CG: fixed.
 
@@
 urgent needs\footnote{There are some exceptions where the quality aspects
-actually matter and such cases will be discussed in section \ref{res:val}%MUN: 
-%this refers to the validity threats section. Do you mean rather the "Theory 
-%Validation" section?
-%CG fixed
-%MUN1: Same here, point to the correct subsection.
+actually matter and such cases will be discussed in section \ref{sect:an:comp
+:cat-vs-literature:conf}
+%MUN: this refers to the validity threats section. Do you mean rather the 
+%"Theory Validation" section? %CG fixed 
+%MUN1: Same here, point to the correct subsection. CG:fixed 
 .}. At the same time, to be able to deal with such constraints, startups must  
@@
 %------------------------------------------------------------------------------%
-\subsection{Severe lack of resources}
-\label{res:gsm:cat7}
+\subsection{Severe lack of resources}\label{res:gsm:cat7} 
 The concept of \textit{severe lack of resources} characterizes the uncertainty  
-of the development strategies in startups and it is composed by three 
-subcategories: \textit{time-shortage}, \textit{limited human resources} and 
-\textit{limited access to expertise}.
+of the development strategies in startups and it is composed by 
+three subcategories: \textit{time-shortage}, \textit{limited human resources} 
+and \textit{limited access to expertise}.
 
@@
 %------------------------------------------------------------------------------%
-\subsection{Team is the catalyst of development}
-\label{res:gsm:cat4}
-Among the different aspects fostering the speed of the development process, the 
-startups' main focus is on the characteristics of the initial team.
+\subsection{Team is the catalyst of development} \label{res:gsm:cat4} 
+Among the different aspects fostering the speed of the development process, 
+the startups' main focus is on the characteristics of the initial team.
 
@@
 %use italics for both interview statements and (sub-)categories. 
+%CG: skip as acked 
 mainly constrained by a tight \textit{time shortage}.
@@
 %------------------------------------------------------------------------------%
-\subsection{Evolutionary approach}
-\label{res:gsm:cat2}
+\subsection{Evolutionary approach}\label{res:gsm:cat2} 
 Among different approaches to development, startups prefer to build an initial  
@@
 %------------------------------------------------------------------------------%
-\subsection{Product quality has low priority}
-\label{res:gsm:cat3}
+\subsection{Product quality has low priority}\label{res:gsm:cat3}
 The interests of software startups, related to the product, are mainly  
@@
 testing}, such as A/B testing \cite{AB}.
-%MUN: add reference or explain shortly what A/B testing is
-%CG: fixed
+%MUN: add reference or explain shortly what A/B testing is %CG: fixed
 
@@
 \footnote{Detailed discussion of the impact of innovative products on the user  
-satisfaction is presented in subsection \ref{sect:an:comp:cat-vs-literature:conf}}.
+satisfaction is presented in subsection 
+\ref{sect:an:comp:cat-vs-literature:conf}}. 
 %MUN: do you use the Kano model in Theory Validation? Change text to refer to  
@@
 %------------------------------------------------------------------------------%
-\subsection{Speed-up development}
-\label{res:gsm:cat1}
+\subsection{Speed-up development}\label{res:gsm:cat1} 
 As remarked initially, \textit{speed-up development} represents the core  
@@
 %------------------------------------------------------------------------------%
-\subsection{Accumulated technical debt}
-\label{res:gsm:cat5}
+\subsection{Accumulated technical debt}\label{res:gsm:cat5} 
 Startups achieve a very high development speed by radically ignoring aspects  
@@
 There is no time for it. Instead, there is a huge pay off in having a code that
-is understandable itself.''}
-This approach is supported by the fact that startups make use of  
-\textit{simple and informal workflow}, \textit{standard/known technologies}, 
-\textit{very small and co-located development team} with \textit{limited need of 
-formalities}.
-%------------------------------------------------------------------------------%
-\subsection{Initial growth hinders performance}
-\label{res:gsm:cat6}
+is understandable itself.''} This approach is supported by the fact that
+startups make use of   \textit{simple and informal workflow},
+\textit{standard/known technologies},  \textit{very small and co-located
+development team} with \textit{limited need of  formalities}.
+%------------------------------------------------------------------------------%
+\subsection{Initial growth hinders performance}\label{res:gsm:cat6} 
 The lack of attention given in the first phases to engineering activities  
@@
 %------------------------------------------------------------------------------%
-\subsection{Theory}
-\label{res:gsm:th}
+\subsection{Theory}\label{res:gsm:th}
 In order to explain and understand the development strategies in early-stage  
@@
 
-\begin{theory}
-Focusing on limited number of suitable functionalities, and adopting partial 
-and rapid evolutionary development approaches, early-stage software startups 
-operate at high development speed, aided by skilled and highly co-located 
-developers. Through these development strategies, early-stage software startups 
-aim to find early product/market fit within extremely uncertain conditions and 
-severe lack of resources. Nevertheless, by speeding-up the development process, 
-they accumulate technical debt, causing an initial and temporary drop-down in 
-performance before setting off for further growth.
+\begin{theory} Focusing on limited number of suitable functionalities, and
+adopting partial  and rapid evolutionary development approaches, early-stage
+software startups  operate at high development speed, aided by skilled and
+highly co-located  developers. Through these development strategies, early-stage
+software startups  aim to find early product/market fit within extremely
+uncertain conditions and  severe lack of resources. Nevertheless, by speeding-up
+the development process,  they accumulate technical debt, causing an initial and
+temporary drop-down in  performance before setting off for further growth.
 \end{theory}
@@
 As discussed in section \ref{res:val} %MUN: this refers to the validity threats
-%section. Do you mean rather the "Theory Validation" section?
-%CG:fixed
+%section. Do you mean rather the "Theory Validation" section? %CG:fixed
 , we formed the theory by applying the paradigm model, introduced by Corbin
@@
 \item \textit{Context} is limited to web early-stage software startups  
-operating in conditions of severe lack of resources aiming to early find 
-product/market fit.
+operating in conditions of severe lack of resources aiming to early
+find  product/market fit. 
 \item \textit{Intervening conditions} are  summarized by the extremely  
@@
 %------------------------------------------------------------------------------%
-\section{Theory Validation}
-\label{res:val}
+\section{Theory Validation} \label{res:val}
 
@@
 %all there and that the current content of 3.5 is not related to what you  
-%actually do here as validation?
+%actually do here as validation? CG: skip as acked
 
@@
 later stage, representing also companies in expansion phase with more than 100
-employees. %MUN: so he did not do research 
-%in a startup context? You need to clarify what you mean by "more mature 
-%companies"... I assume you still mean startups, but startups at a later stage, 
-%i.e. after first product release.
+employees. %MUN: so he did not do research  in a startup context? You need to
+%clarify what you mean by "more mature companies"... I assume you still mean
+%startups, but startups at a later stage, i.e. after first product release.
 %CG: fixed
@@
 is not a matter of poor knowledge and training, but rather it is the necessity
-of operating with solutions that let the company move faster. 
-\textit{``One-size-fits-all''} solutions have always found difficulty in 
-penetrating small software organizations \cite{Staples2007}. When startups began 
-establishing any SPI process, they experienced \textit{process erosions}  
-\cite{Coleman2008} %MUN: 
-%add ref (Coleman?). You should add, for each category you use from Coleman, 
-%add the reference to the paper. This makes it clear that the category comes 
-%from him (and is not one of our categories)
-%CG: fixed
+of operating with solutions that let the company move faster.  \textit{``One-
+size-fits-all''} solutions have always found difficulty in  penetrating small
+software organizations \cite{Staples2007}. When startups began  establishing any
+SPI process, they experienced \textit{process erosions}   \cite{Coleman2008}
+%MUN:  %add ref (Coleman?). You should add, for each category you use from
+%Coleman, add the reference to the paper. This makes it clear that the category
+%comes from him (and is not one of our categories) %CG: fixed 
 , which resolved to a barely sufficient workflow to satisfy the organizational  
@@
 %do not produce heavy bureaucracy, but they allow fast communication
-%strategies... I may be wrong, but then you need to explain it better
-%CG:fixed, you are right :) ...  
+%strategies... I may be wrong, but then you need to explain it better %CG:fixed,
+%you are right :) ...   
 , even though the accumulated tacit knowledge is hard to manage and to transfer 
@@
 found that despite Coleman's study contributes to define which factors hinder
-the introduction of SPI in startups 
-%MUN: add: "of SPI in startups"? CG: fixed
+the introduction of SPI in startups  %MUN: add: "of SPI in startups"? CG: fixed
 his results are coherent with the theory generated within our research, focusing
@@
 practices is advisable, under circumstances of high uncertainty and severe lack
-of resources, %MUN: under which  %circumstances? This statement sounds too
+of resources, %MUN: under which circumstances? This statement sounds too
 %general, i.e. don't do SPI ever... %CG: fixed  
 %And the following part,I did not fully understand. Try to clarify. CG:fixed   
-not following SPI does not predict
-major adaptiveness of startups development towards the customer's needs.
-Nevertheless, to imply a major evolutionary product/market fit, certain
-conditions explained in GSM need to be considered.
+not following SPI does not predict major adaptiveness of startups development 
+towards the customer's needs. Nevertheless, to imply a major evolutionary 
+product/market fit, certain conditions explained in GSM need to be considered.
 
@@
 \item Speed-up development by releasing more often the software and  
-``implanting'' customers in the development environment (similar to the 
-\textit{find the product/market fit} concept presented in CAT2). 
+``implanting'' customers in the development environment (similar
+to the  \textit{find the product/market fit} concept presented in CAT2).
 \item Make heavy use of simple tools and existing components (similar to the  
@@
 simple but stable architectural solutions (see subcategory \textit{using  
-modularization and frameworks from day 1} discussed in CAT5). 
+modularization and frameworks from day 1} discussed
+in CAT5).
 \item Tailor the development process daily according to the intense demands for 
@@
 quality software.
-
 \end{compactitem}
@@
 already implemented. It is the main strategy, which enables startups to
-externalize complexity to third party solutions explained within CAT1,  
-\textit{speed-up development}.
+externalize complexity to third party solutions explained within CAT1,   \textit
+{speed-up development}. 
 \item Requirements refinement and rapid prototyping- avoid deciding precisely  
@@
 %MUN: add ref %CG:fixed
-, are unlikely to follow repeatable prescriptive methodologies. As stated by 
-Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and every 
-methodology is to expect people to behave consistently over time, when is 
+, are unlikely to follow repeatable prescriptive methodologies. As
+stated by  Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and
+every  methodology is to expect people to behave consistently over time, when is
 clearly not like that''}. Especially in early-stage startups, where the whole
@@
 start to grow the accumulated technical debt which requires fast integration of
-more structured and standard processes overtime.
-Issues related to low-precision engineering practices during companies' growth 
-are discussed in an experience paper by Ambler \cite{Ambler2002}. He reports on 
-two case studies of growing Internet startups approaching an initial public  
-offering (IPO). %MUN:explain %acronym %CG: fixed
+more structured and standard processes overtime. Issues related to low-precision
+engineering practices during companies' growth  are discussed in an experience
+paper by Ambler \cite{Ambler2002}. He reports on  two case studies of growing
+Internet startups approaching an initial public offering (IPO). %MUN:explain
+%acronym %CG: fixed 
 Despite the two companies were facing business and team growth, attitudes  
@@
 tailored version of the Rational Unified Process (RUP)
-%MUN: add reference
-%CG: not fixed, not sure if needed since Ambler has the detailed 
-%explaination of his process, similar to RUP... anyway a dedicated resource 
-%would be an IBM website
+%MUN: add reference %CG: not fixed, not sure if needed since Ambler has the
+%detailed explaination of his process, similar to RUP... anyway a dedicated
+%resource would be an IBM website
 development framework. Nevertheless, even with the aid of CASE  
 tools\footnote{CASE tools are a class of software that automates many of the  
-activities involved in various engineering activities. See 
-\url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .}, 
-startups failed in practicing engineering activities in view of excessive lack of 
-flexibility. Developers fast moved to simpler tools such as using a white-board, 
-where they were able to provide input and insights, question assumptions, share 
-past experiences, and even talk about potential implementation strategies. 
+activities involved in various engineering activities.
+See  \url{http://www.unl.csi.cuny.edu/faqs/software-enginering/tools.html} .},
+startups failed in practicing engineering activities in view of excessive lack
+of  flexibility. Developers fast moved to simpler tools such as using a white-
+board,  where they were able to provide input and insights, question
+assumptions, share  past experiences, and even talk about potential
+implementation strategies.  
 %MUN: Candidate for removal since details can be found in Ambler anyway: "In  
@@
 
-The use of whiteboards in working environments supports \textit{use of 
-well-integrated and simple tools}. In both the startups studied by Ambler, 
+%------------------------------------------------------------------------------%
+\subsubsection{Minimal project management}
+
+The use of whiteboards in working environments supports \textit{use of  well-
+integrated and simple tools}. In both the startups studied by Ambler,
 documentation was minimal, finding the ``sweet spot'' to have just barely enough
@@
 %CG: partially fixed, removed the detailed process description 
-In the evaluation phase, the status of the development workflow within the 
-company  is analyzed. The researchers documented a state of chaos where any 
-initiatives of quantitative methodology assessment would have been a waste of 
-time. Therefore, they applied a qualitative methodology to define a model of the 
-release management process. 
-%They moved towards the plan phase where the 
-%researchers decided to develop a model to manage releases, introducing a process 
-%composed by: release scope preparation, release planning, release development, 
-%system testing, acceptance testing and release deployment. 
+In the evaluation phase, the status of the development workflow
+within the  company  is analyzed. The researchers documented a state of chaos
+where any  initiatives of quantitative methodology assessment would have been a
+waste of  time. Therefore, they applied a qualitative methodology to define a
+model of the  release management process.  
+%They moved towards the plan phase where the  %researchers decided to develop a 
+%model to manage releases, introducing a process composed by: release scope 
+%preparation, release planning, release development, system testing, 
+%acceptance testing and release deployment.  
 Despite the effort spent by the researchers to establish a development workflow, 
@@
 days.  
-Regarding the previously mentioned problems, only 
-partial improvements have been achieved since the major obstacle was to 
-motivate employees to change their habits. With perfect hindsight the 
-researchers claim that understanding and adapting solutions by solving one 
-problem at a time would have increased the benefits. Then, despite the reported 
-results, starting the establishment of a simple and informal workflow aided the 
-organization to move from the stage of knowing nothing to knowing at least 
-something about their development process.
+Regarding the previously mentioned problems, only  partial improvements have 
+been achieved since the major obstacle was to  motivate employees to change 
+their habits. With perfect hindsight the researchers claim that understanding 
+and adapting solutions by solving one problem at a time would have increased 
+the benefits. Then, despite the reported results, starting the establishment of 
+a simple and informal workflow aided the organization to move from the stage of 
+knowing nothing to knowing at least something about their development process.
 
@@
 tailored solutions.
-
-Prescribed development models, oriented towards SPI, are basically ignored by 
-startups, as discussed in \cite{Zettel2001}. 
-%MUN: <-- rephrase, not clear to me %CG:fixed
-To overcome the need of structure and to accelerate the development process, 
-Zettel developed a \textit{``LIghtweight Process for E-business software 
-development''} (LIPE). This lightweight development method integrates the XP 
-approach with ideas from areas of software measurement for project control and 
-process improvement as a compromise between an ad-hoc and a more rigorous 
-approach. Zettel argues that LIPE provides scalability of the development 
-process with a certain degree of flexibility, omitting parts of prescribed 
-practices. Nevertheless, we couldn't identify any empirical evaluation of LIPE 
-that supports its suitability in a startup context.
-
-Deakins \cite{Deakins2005} designed a new spiral methodology, constructed  
-during a fieldwork with a developer team that had limited resources managing an 
-innovative product in volatile e-commerce environment. This methodology 
-represents a fast evolutionary approach with flexible processes. 
-It aims to obtain customer's feedback as soon as possible for continuous 
-product improvement.%MUN: %<-- rephrase, sentence not parsable CG:fixed
-Even in this case, the spiral methodology%MUN: which model? CG:fixed
-is compliant with the nature of a random, ad-hoc and  \textit{evolutionary 
-development approach} (CAT2). Evolutionary development approaches are 
-typically more suitable for web applications, as confirmed in 
-\cite{Deshpande2001}.
-Moreover, they provide practices oriented towards rapid and iterative design 
-with complete freedom to the developer team, enabled to operate anywhere within 
-the organization.
-%MUN: <-- the whole paragraph is not that clear to me. Try to start 
-%with the category (team is catalyst), then write about Deakins spiral model
-%CG: partially fixed, the focus here is on the evolutionary development approach
-
-Carmel \cite{Camel1994a} reports on the need of flexible and rapid development 
-solutions to shorten time-to-market, but even more important, the need of 
-high-skilled team developers, which has an impact on \textit{speed-up 
-development} (CAT1). He suggests that entrepreneurs need to look for a well 
-formed, skilled core development team and not just a set of product ideas and 
-features. Moreover the team must be empowered with full-stack and 
-self-organization settings, paired with the flexibility of minimal bureaucracy. 
-A studied company stated: \textit{``we didn't need weeks or months of detailed 
-modeling and documentation, but rather modeling the architecture a little and 
-then either exploring strategies by providing users with code or simply starting 
-work on the actual software itself''}.
 
@@
 
+%------------------------------------------------------------------------------%
+\subsubsection{Simple workflows} Prescribed development models, oriented towards
+SPI, are basically ignored by  startups, as discussed in \cite{Zettel2001}.
+%MUN: <-- rephrase, not clear to me %CG:fixed 
+To overcome the need of structure and to accelerate the development process,  
+Zettel developed a \textit{``LIghtweight Process for E-business software  
+development''} (LIPE).
+This lightweight development method integrates the XP  approach with ideas from
+areas of software measurement for project control and  process improvement as a
+compromise between an ad-hoc and a more rigorous  approach. Zettel argues that
+LIPE provides scalability of the development  process with a certain degree of
+flexibility, omitting parts of prescribed  practices. Nevertheless, we couldn't
+identify any empirical evaluation of LIPE  that supports its suitability in a
+startup context.
+
+Deakins \cite{Deakins2005} designed a new spiral methodology, constructed
+during a fieldwork with a developer team that had limited resources managing an
+innovative product in volatile e-commerce environment. This methodology
+represents a fast evolutionary approach with flexible processes.  It aims to
+obtain customer's feedback as soon as possible for continuous  product
+improvement.%MUN: %<-- rephrase, sentence not parsable CG:fixed 
+Even in this case, the spiral methodology%MUN: which model? CG:fixed 
+is compliant with the nature of a random, ad-hoc and  \textit{evolutionary  
+development approach} (CAT2). Evolutionary development approaches are  
+typically more suitable for web applications, as confirmed in  
+\cite{Deshpande2001}. Moreover, they provide practices oriented towards rapid 
+and iterative design with complete freedom to the developer team, enabled to 
+operate anywhere within the organization. 
+%MUN: <-- the whole paragraph is not that clear to me. Try to start  
+%with the category (team is catalyst), then write about Deakins spiral model 
+%CG: partially fixed, the focus here is on the evolutionary development approach
+
+%------------------------------------------------------------------------------%
+\subsubsection{Full-stack development team}
+
+Carmel \cite{Camel1994a} reports on the need of flexible and rapid development
+solutions to shorten time-to-market, but even more important, the need of  high-
+skilled team developers, which has an impact on \textit{speed-up  development}
+(CAT1). He suggests that entrepreneurs need to look for a well  formed, skilled
+core development team and not just a set of product ideas and  features.
+Moreover the team must be empowered with full-stack and  self-organization
+settings, paired with the flexibility of minimal bureaucracy.  A studied company
+stated: \textit{``we didn't need weeks or months of detailed  modeling and
+documentation, but rather modeling the architecture a little and  then either
+exploring strategies by providing users with code or simply starting  work on
+the actual software itself''}.
+
 %MUN: When you subdivide this sections, as proposed above, finish with a
@@
 %respective sections CG: not parsed. 
-%MUN1: me neither... forget what I wrote :-)
+%MUN1: me neither... forget what I wrote :-). CG: ok 
 , related to: creativity and innovation, market requirements and application  
@@
 %MUN: Please reformulate the first sentence, I cannot parse it. CG: fixed 
-The proposed model identifies the three lifecycle stages of a startup: 
-\textit{fluid}, \textit{transition} and \textit{specific}. At each stage, the 
-model presents different development approaches (product-oriented and 
-process-oriented), according to the level of innovation achieved. The study 
+The proposed model identifies the three lifecycle stages of a
+startup:  \textit{fluid}, \textit{transition} and \textit{specific}. At each
+stage, the  model presents different development approaches (product-oriented
+and  process-oriented), according to the level of innovation achieved. The study
 reports how product-oriented development, in contrast with forced and mature
@@
 Market requirements \cite{Coleman2007} and application type \cite{Sutton2000,  
-Coleman2008, Coleman2008a} are two further potential confounding factors to the 
+Coleman2008,Coleman2008a} are two further potential confounding factors to the  
 GSM. Their main impact is related to the adoption of flexible and reactive  
@@
 For example, safety-critical systems possess quality concerns that must be
-considered. Even in e-commerce systems factors related to quality have 
-high-impact effects on obtaining the first customer reference, 
+considered. Even in e-commerce systems factors related to quality have  high-
+impact effects on obtaining the first customer reference,  
 %MUN: <-- cannot parse it. Remove the first part and write: "Even e-commerce  
@@
 %startups w.r.t. customer expectations and quality concerns. It's ok, but it
-%adds little value to the validation of GSM. I'd remove it.
-%To understand how to obtain processes in accordance to a set of customer 
-%needs, correlating the proper levels of customers' satisfaction, we 
-%introduce the Kano model, described in \cite{Ullah2011} and extensively 
-%studied in product development.  As shown in the Kano model, users 
-%preferences can be classified into four categories: \textit{basic}, 
-%\textit{excitement}, \textit{performance} % and \textit{indifference} areas.
-% 
-%The \textit{basic} area represents the space where customers are dissatisfied 
-%if those \textit{must-be} functionalities are not met. This is the case of 
-%similar functionalities already implemented in the current market with 
-%well-known solutions. The \textit{performance} area represents the space where 
-%customers' satisfaction is proportional to how many of functionalities are 
-%implemented. In the \textit{indifference} area the customer does not really 
-%care about the functionalities: the satisfaction remains invariant. This 
-%represents the case where startups have not (yet) achieved the right 
-%product/market fit. Finally, the \textit{excitement} area represents the space 
-%where customers will be more satisfied if functionalities are implemented, but 
-%not disappointed if they are not implemented. This is the case of extremely 
-%innovative product development.
-% 
-%Well-known domains such as e-commerce applications or other \textit{``quality 
-%critical products''} belong to the \textit{basic} area, where if a quality 
-%concern such as security in bank transfer services is missing satisfaction of 
-%customers will be profoundly affected. But as discussed, in early-stage 
-%startups that typically develop first beta products, developers usually deal 
-%with cutting-edge technologies. Those innovative solutions, if in the right 
-%market, are likely to belong to  the \textit{excitement} area, where 
-%satisfaction can hardly be negatively affected by faulty products. In fact, in 
-%the GSM model, minimal functionalities and less concern about quality aspects 
-%have low-impact on customers' satisfaction, supporting startup's ability to 
-%mainly focus on \textit{speed-up development}. Moreover operating in the 
-%\textit{excitement area} is fully compliant with an evolutionary approach and 
-%minimal functionalities since from the first product released to the market, 
-%it is unlikely to lose users because 
-% of high-quality demand.
-% 
-%Nevertheless, this aspect profoundly undermines the generalizability of the 
-%theoretical model according to the theory, in particular \textit{CAT3}. 
-%therefore, future studies within wider startups' context are advisable to 
-%extrapolate insights from those situations where quality concerns cannot be 
-%postponed or externalized. CG: ok
+%adds little value to the validation of GSM. I'd remove it. To understand how to
+%obtain processes in accordance to a set of customer  needs, correlating the
+%proper levels of customers' satisfaction, we  introduce the Kano model,
+%described in \cite{Ullah2011} and extensively  studied in product development.
+%As shown in the Kano model, users  preferences can be classified into four
+%categories: \textit{basic},  \textit{excitement}, \textit{performance}  and
+%\textit{indifference} areas.   The \textit{basic} area represents the space
+%where customers are dissatisfied  if those \textit{must-be} functionalities are
+%not met. This is the case of  similar functionalities already implemented in
+%the current market with  well-known solutions. The \textit{performance} area
+%represents the space where  customers' satisfaction is proportional to how many
+%of functionalities are  implemented. In the \textit{indifference} area the
+%customer does not really  care about the functionalities: the satisfaction
+%remains invariant. This  represents the case where startups have not (yet)
+%achieved the right  product/market fit. Finally, the \textit{excitement} area
+%represents the space  where customers will be more satisfied if functionalities
+%are implemented, but  not disappointed if they are not implemented. This is the
+%case of extremely  innovative product development.   Well-known domains such as
+%e-commerce applications or other \textit{``quality  critical products''} belong
+%to the \textit{basic} area, where if a quality  concern such as security in
+%bank transfer services is missing satisfaction of  customers will be profoundly
+%affected. But as discussed, in early-stage  startups that typically develop
+%first beta products, developers usually deal  with cutting-edge technologies.
+%Those innovative solutions, if in the right  market, are likely to belong to
+%the \textit{excitement} area, where  satisfaction can hardly be negatively
+%affected by faulty products. In fact, in  the GSM model, minimal
+%functionalities and less concern about quality aspects  have low-impact on
+%customers' satisfaction, supporting startup's ability to  mainly focus on
+%\textit{speed-up development}. Moreover operating in the  \textit{excitement
+%area} is fully compliant with an evolutionary approach and  minimal
+%functionalities since from the first product released to the market,  it is
+%unlikely to lose users because   of high-quality demand.   Nevertheless, this
+%aspect profoundly undermines the generalizability of the  theoretical model
+%according to the theory, in particular \textit{CAT3}.  therefore, future
+%studies within wider startups' context are advisable to  extrapolate insights
+%from those situations where quality concerns cannot be  postponed or
+%externalized. CG: ok
 
@@
 interviews, ``lack of experience'' was not considered as a sub-category in  
-\textit{severe lack of resources}. Nevertheless, startups rely at the beginning 
-on clever, but inexperienced developers \cite{Crowne2002}. Having team-members 
-with deep experience is a ``double-edge sword'' \cite{Yoffie1999}. On one hand, 
-experience might quickly provide structure and maturity to the development 
-process, on the other it might cause challenges in managing self-confident 
-overachievers that almost inevitably clash. Consequently, team management 
-requires higher control and coordination activities that inevitably hinder the 
-flexibility of the development environment, that is essential in early-stage 
-startups. Accordingly, also Coleman \cite{Coleman2007,
-Coleman2008a,Coleman2008} states that the operating strategies towards 
-\textit{``minimum processes''} in startups is not a matter of poor knowledge and 
-training, but rather it is the necessity of operating with solutions that let 
-the company move faster.
-
-%MUN: removed below paragraph since it does not add much
-%Nevertheless, since the amount of the results did not allow ``the lack of 
-%experience'' to be in the GSM model (in view of the fact that not enough data 
-%were identified \footnote{A category is formed if the codes are representative 
-%of more than the 70\% of the analyzed transcripts.}), this confounding factor 
-%requires further investigation. CG: ok
-
-%------------------------------------------------------------------------------%
-\subsection{Static assessment of the GSM}
-\label{sect:theory:validation:stats}
+\textit{severe lack of resources}.
+Nevertheless, startups rely at the beginning  on clever, but inexperienced
+developers \cite{Crowne2002}. Having team-members  with deep experience is a
+``double-edge sword'' \cite{Yoffie1999}. On one hand,  experience might quickly
+provide structure and maturity to the development  process, on the other it
+might cause challenges in managing self-confident  overachievers that almost
+inevitably clash. Consequently, team management  requires higher control and
+coordination activities that inevitably hinder the  flexibility of the
+development environment, that is essential in early-stage  startups.
+Accordingly, also Coleman \cite{Coleman2007, Coleman2008a,Coleman2008} states
+that the operating strategies towards  \textit{``minimum processes''} in
+startups is not a matter of poor knowledge and  training, but rather it is the
+necessity of operating with solutions that let  the company move faster.
+
+%MUN: removed below paragraph since it does not add much Nevertheless, since the
+%amount of the results did not allow ``the lack of  experience'' to be in the
+%GSM model (in view of the fact that not enough data  were identified
+%\footnote{A category is formed if the codes are representative  of more than
+%the 70\ of the analyzed transcripts.}), this confounding factor  requires
+%further investigation. CG: ok
+
+%------------------------------------------------------------------------------%
+\subsection{Static assessment of the GSM} \label{sect:theory:validation:stats}
 
@@
 %providing researchers an additional tool, supported by further companies'  
-%data, to validate our model."
+%data, to validate our model." CG: not important, I would not add this.
 
@@
 %CG: fixed, explained in methodology  
-%What do you mean by "we grounded in the interview transcripts the fact that 
+%MUN:What do you mean by "we grounded in the interview transcripts the fact that 
 %those relations were characterized in terms of performances since  
@@
 
-\begin{compactitem}
-\item The team enhances the overall \textit{performance} of software 
-development. Recalling that startups deal with severe lack of resources, the 
-team members are the company's main assets. They develop software with no need 
-of formalities and structures, left to the team's own capabilities to 
-efficiently develop the product. Moreover, the effectiveness is granted by 
-continuous advice given by mentors and by a management style oriented towards 
-self-organization, full-stack, and multi-role settings. The importance of a 
-skilled team for development performance is expressed in many studies such as 
-\cite{surviving-os-cockburn, Highsmith2000}, described in CAT4 and confirmed by 
-Sutton \cite{Sutton2000}.
-\item The evolutionary approach enhances the \textit{effectiveness} in 
-implementing the correct functionalities. This approach improves the company's 
-capabilities to adjust the trajectory of product development (in agreement to 
-the concept of pivoting discussed by Ries \cite{Ries2011}).
-\item Since product quality has low priority, the development is focused only on 
-implementing a limited number of suitable functionalities, enhancing its 
-\textit{efficiency}. An essential prototype (MVP) does not require to comply 
-with heavy quality constraints, enabling the team to quickly implement 
-functionalities, ready to be validated by final users.
-\end{compactitem}
+\begin{compactitem} \item The team enhances the overall \textit{performance} of
+software  development. Recalling that startups deal with severe lack of
+resources, the  team members are the company's main assets. They develop
+software with no need  of formalities and structures, left to the team's own
+capabilities to  efficiently develop the product. Moreover, the effectiveness is
+granted by  continuous advice given by mentors and by a management style
+oriented towards  self-organization, full-stack, and multi-role settings. The
+importance of a  skilled team for development performance is expressed in many
+studies such as  \cite{surviving-os-cockburn, Highsmith2000}, described in CAT4
+and confirmed by  Sutton \cite{Sutton2000}. \item The evolutionary approach
+enhances the \textit{effectiveness} in  implementing the correct
+functionalities. This approach improves the company's  capabilities to adjust
+the trajectory of product development (in agreement to  the concept of pivoting
+discussed by Ries \cite{Ries2011}). \item Since product quality has low
+priority, the development is focused only on  implementing a limited number of
+suitable functionalities, enhancing its  \textit{efficiency}. An essential
+prototype (MVP) does not require to comply  with heavy quality constraints,
+enabling the team to quickly implement  functionalities, ready to be validated
+by final users. \end{compactitem}
 
@@
 %thesis to this limited space, but I think it is doable (and relevant for  
-%supporting validity of GSM). CG: fixed
+%supporting validity of GSM).
+%CG: fixed
 
 \begin{compactenum} 
-\item The more a capable \footnote{The team capability is evaluated according 
-to the characteristics presented in CAT4.} team, with few quality constraints, 
-uses an evolutionary approach, the faster the company releases the product.
+\item The more a capable \footnote{The team capability is
+evaluated according  to the characteristics presented in CAT4.} team, with few
+quality constraints,  uses an evolutionary approach, the faster the company
+releases the product. 
 \item The faster the company releases the product, the larger the degree of  
@@
 \begin{compactitem}
-
 \item \textit{Execution speed}: a metric that represents the development speed
@@
 %"AND with less quality constraints" --> just makes the hypothesis more  
-%difficult to understand. 
-We conducted statistical tests  using the analysis of variance to assess the 
-existence of relations between the dimensions. First we defined two null 
-hypotheses:
+%difficult to understand. CG: fixed. 
+We conducted statistical tests  using the analysis of
+variance to assess the  existence of relations between the dimensions. First we
+defined two null  hypotheses:
 
 \begin{itemize} 
-\item[$H0_1$:] \textit{Startups do not release the product faster 
-when a capable team adopts a more evolutionary approach AND with less quality 
-constraints}
-\item[$H0_2$:] \textit{The execution speed does not increase the amount 
-of accumulated technical debt}. 
+\item[$H0_1$:] \textit{Startups do not release the product
+faster  when a capable team adopts a more evolutionary approach AND with less
+quality  constraints} 
+\item[$H0_2$:] \textit{The execution speed does not
+increase the amount  of accumulated technical debt}.
 \end{itemize}
 
-Then we tested $H0_1$ and $H0_2$  with an 
-one-tailed test using Pearson's product moment correlation 
-coefficient\footnote{To perform the Pearson's correlation we verified two 
-assumptions: a) data is normally distributed; b) data is on interval scale.}, 
-with positive association analysis, fixing the level of confidence to 95\%.
-We reject both $H0_1$ and $H0_2$, and conclude that in our sample:
+The formulation of the two hypotheses reflects the representation of the
+categories related to \textit{speed-up development} presented in the GSM model
+(see figure \ref{fig:gsm}). We tested $H0_1$ and $H0_2$  with an  one-tailed
+test using Pearson's product moment correlation  coefficient\footnote{To perform
+the Pearson's correlation we verified two  assumptions: a) data is normally
+distributed; b) data is on interval scale.},  with positive association
+analysis, fixing the level of confidence to 95\%. We reject both $H0_1$ and
+$H0_2$, and conclude that in our sample: 
 \begin{compactenum}
@@
 
-\item Higher values of \textit{Execution speed} are 
-strongly associated (p-value: 0.005) with higher values for  
-\textit{Potential capability}.
+\item Higher values of \textit{Execution speed} are  strongly associated
+(p-value: 0.005) with higher values for   \textit{Potential capability}.
 \end{compactenum}
@@
 %MUN1:Do we (research) know enough about how to shorten time-to-market without
-%accumulating technical debt? What are possible venues for research?
+%accumulating technical debt? What are possible venues for research? CG: fixed
+%see summary.
 
@@
 \subsection{Light-weight methodology} 
-Startups do not apply any standard development methodology. When focused 
-on building the first version of the product, startups do not observe any 
+Startups do not apply any standard development methodology. When focused  on 
+building the first version of the product, startups do not observe any  
 specific and standard development methodologies or processes.  From an accurate 
@@
 
-The closest development approach undertaken by early-stage startups is 
-the \textit{Lean startup methodology}.  A highly evolutionary development 
-approach, centered around the quick production of a functioning prototype and 
-guided by customer feedback is a central aspect of the \textit{Lean startup 
-methodology} pioneered by Ries \cite{Ries2011}. However, startups do not 
-formally follow the cycle \textit{build-measure-learn} as it is proposed by the 
-methodology, where they should set-up an experiment to \textit{test the riskiest 
-element of the business idea}, nor make explicit use of the suggested good 
-practices.
+The closest development approach undertaken by early-stage startups is  the
+\textit{Lean startup methodology}.  A highly evolutionary development  approach,
+centered around the quick production of a functioning prototype and  guided by
+customer feedback is a central aspect of the \textit{Lean startup  methodology}
+pioneered by Ries \cite{Ries2011}. However, startups do not  formally follow the
+cycle \textit{build-measure-learn} as it is proposed by the  methodology, where
+they should set-up an experiment to \textit{test the riskiest  element of the
+business idea}, nor make explicit use of the suggested good  practices. Instead,
+startups in the early stage prefer also to apply fast cycles of "build and fix"
+when necessary, hacking the first functionalities to present to potential
+customers and investors in order to obtain quick feedbacks. 
 %MUN1: Why don't follow startups the build-measure-learn cycle proposed by  
@@
 %time-to-market? Is Lean still to Heavy for startups (in their early phases)? 
-For example, when it comes to user feedback, startups do 
-not explicitly follow the step-by-step process of \textit{customer development}
-formally defined by Blank \cite{Blank2005}. Instead, startups absorb and 
-implement the high-level principles that outlines the methodology%MUN1:<-- 
-%which methodology?
+% CG:fixed. 
+Moreover, when it comes to user feedback, startups do  not explicitly follow 
+the step-by-step process of \textit{customer development} formally defined by 
+Blank \cite{Blank2005}. Instead, startups absorb and  implement the high-level
+principles that outlines the customer development methodology
+%MUN1:<--  %which methodology? CG: fixed 
 , reflected in the GSM by the theoretical category \textit{find the  
@@
 
-%MUN1: paragraph below is candidate for removal
-%The greatest part of engineering activities of startups are focused on 
-%the implementation while only little attention is given to more conventional 
-%activities (project management, requirement specifications, analysis, 
-%architecture design, automatic testing, \ldots). Especially when bringing the 
-%first product to the market, the attitude of startups towards traditional 
-%engineering activities is strongly characterized by a general feeling of 
-%unnecessary. Every document, artifact or process, which does not produce a 
-%short-term tangible benefit to the software production is perceived as a waste 
-%of time.
+%MUN1: paragraph below is candidate for removal The greatest part of engineering
+%activities of startups are focused on  the implementation while only little
+%attention is given to more conventional  activities (project management,
+%requirement specifications, analysis,  architecture design, automatic testing,
+%\ldots). Especially when bringing the  first product to the market, the
+%attitude of startups towards traditional  engineering activities is strongly
+%characterized by a general feeling of  unnecessary. Every document, artifact or
+%process, which does not produce a  short-term tangible benefit to the software
+%production is perceived as a waste  of time. CG: ok
 
@@
 \subsection{Focus on minimal set of functionalities} 
-The first release of the product includes only a limited set of well 
-suitable functionalities focused on user experience. The beta version is usually 
+The first release of the product includes only a limited set of well  suitable 
+functionalities focused on user experience. The beta version is usually  
 focused on a minimal and essential set of functionalities providing a reasonable 
-user experience (UX). 
+user experience (UX). The first functionalities are usually set by the 
+founders and their experience in a particular business domain. As soon as 
+possible they provide demos to demonstrate the idea to potential customers. 
 %MUN1: This rises the question, how do startups prioritize requirements.  
@@
 %adopted by startups in a modified form in order to increase the probabilities  
-%of success in the market?
-As confirmed by practitioners, the user must be able to 
-accomplish a task on the product without major interruptions in the flow, which 
-would cause frustration before being able to capture his attention. To realize 
-UX in a short time frame, developers focus on a limited number of suitable 
+%of success in the market?. CG: fixed. 
+As confirmed by practitioners, the user must be able to  accomplish a task on 
+the product without major interruptions in the flow, which  would cause 
+frustration before being able to capture his attention. To realize  UX in a 
+short time frame, developers focus on a limited number of suitable  
 functionalities, especially in the first version(s) of the product. Other  
@@
 perspective, one might argue that established companies might take advantage of 
-the new idea and develop a better version of the product internally.
-Despite this might represent a risk, established companies usually cannot 
-afford to operate in a market of \textit{innovators} and \textit{early adopters} 
-group of users - described by the Roger's bell curve in \cite{Rogers} - because 
-of the little return of investment they would gain, as discussed by Christensen 
-in \cite{Christensen2003}. Instead, established companies have more interests 
-in acquiring a startup and all the knowledge it has gained during the 
+the new idea and develop a better version of the product internally. Despite 
+this might represent a risk, established companies usually cannot  afford to 
+operate in a market of \textit{innovators} and \textit{early adopters}  group 
+of users - described by the Roger's bell curve in \cite{Rogers} - because  of 
+the little return of investment they would gain, as discussed by Christensen  in
+\cite{Christensen2003}. Instead, established companies have more interests  in
+acquiring a startup and all the knowledge it has gained during the
 \textit{customer development} process while focusing effort on acquiring larger
 shares of a ``mass market'' sector.  
-%MUN: candidate for shorting
-%Furthermore, thanks to cloud computing and 
-%more in general high availability of third party components, aspects such as 
-%performance, security and scalability can be achieved with a reasonable 
-%effort. 
-%Thus, given the general lack of quality concerns, and supported by the fact 
-%that the optimal strategy to improve the UX is by having the product used by 
-%real customers, startups can effectively focus on achieving short 
-%time-to-market.
+%MUN1: candidate for shorting 
+%Furthermore, thanks to cloud computing and more in general high availability 
+%of third party components, aspects such as performance, security and 
+%scalability can be achieved with a reasonable effort. Thus, given the 
+%general lack of quality concerns, and supported by the fact that the optimal 
+%strategy to improve the UX is by having the product used by real customers, 
+%startups can effectively focus on achieving short time-to-market. CG: fixed
 
@@
 %$MUN1: candidate for removal 
-%The initial (small) development team usually spends most of the 
-%hours of the week working closely to implement the product, often extending 
-%the conventional office hours.
+%The initial (small) development team usually spends most of the hours of the 
+%week working closely to implement the product, often extending  
+%the conventional office hours. CG: ok. 
 The large amount of communication, spontaneously held in informal places, has  
-been proven to be an extremely effective element of 
-several successful software projects \cite{Highsmith2000}. Hence, the vastly 
-\textit{informal} development environment of startups acts as a catalyst of 
-speed.
+been proven to be an extremely effective element of several successful software 
+projects \cite{Highsmith2000}. Hence, the vastly \textit{informal} development 
+environment of startups acts as a catalyst of speed. 
 %MUN1: candidate for removal 
@@
 %frameworks; low precision artifacts; and the close colocation and capability  
-%of the team.
+%of the team. CG: ok.
 
@@
 The initial lack of structures and processes negatively affects the  
-performance when the company starts to grow. One of the primary objectives of 
-startups is to expand their business into fast growing markets. If 
-successful, the company will face a growth in terms of number of customers, 
-employees and product capabilities. Then, the increased complexity arises the 
-necessity of controlling the initially chaotic software development environment. 
-Under delicate circumstances, where an increased market demand meets a poorly 
-engineered product and process, the development team need to start returning 
-the accumulated technical debt. As reported by practitioners, introducing more 
-structures, documentation and standard procedures, cause an initial drop-down in 
-performance which severity depends on the speed of growth, the amount of 
-technical debt and the capacity of the team to absorb it. 
+performance when the company starts to grow.
+One of the primary objectives of  startups is to expand their business into fast
+growing markets. If  successful, the company will face a growth in terms of
+number of customers,  employees and product capabilities. Then, the increased
+complexity arises the  necessity of controlling the initially chaotic software
+development environment.  Under delicate circumstances, where an increased
+market demand meets a poorly  engineered product and process, the development
+team need to start returning  the accumulated technical debt. As reported by
+practitioners, introducing more  structures, documentation and standard
+procedures, cause an initial drop-down in  performance which severity depends on
+the speed of growth, the amount of  technical debt and the capacity of the team
+to absorb it.  
 %MUN1: what does literature say about technical debt? Are there strategies that 
@@
 %the performance bump that hampers growth? 
-
-\subsection{Summary}
-All of the above observations are extracted from the analysis of the model, 
-which is systematically derived from empirical data of the case study. Indeed, 
-with slightly different levels of adherence, the implications we presented are 
-reflected in the behavior of the great majority of companies in the sample. The 
-results of this analysis, which has been focused on the software development 
-strategy from the idea conception to the first release, indicate that startups 
-are far from adopting standard or ad-hoc development methodologies, especially 
-in this phase. The typical tendency is to focus on the team capability to 
-implement and quickly iterate on a prototype, which is released very fast. Thus, 
-in a context where even for the most \textit{lightweight} agile methodologies is 
-hard to penetrate, we believe that it is premature to discuss issues related to 
-SPI. Researchers should rather focus on the trade-off between development speed 
-and accumulated technical debt \cite{Brown:2010:MTD:1882362.1882373}, which 
-appears to be the most important determinant for the future of 
-software development in the company.
-
-%------------------------------------------------------------------------------%
-%\subsection{Future work}
-%label{futwork}
+%CG: partially fixed below in summary.
+
+\subsection{Summary} All of the above observations are extracted from the
+analysis of the model,  which is systematically derived from empirical data of
+the case study. Indeed,  with slightly different levels of adherence, the
+implications we presented are  reflected in the behavior of the great majority
+of companies in the sample. The  results of this analysis, which has been
+focused on the software development  strategy from the idea conception to the
+first release, indicate that startups  are far from adopting standard or ad-hoc
+development methodologies, especially  in this phase. The typical tendency is to
+focus on the team capability to  implement and quickly iterate on a prototype,
+which is released very fast. Thus,  in a context where even for the most
+\textit{lightweight} agile methodologies is  hard to penetrate, we believe that
+it is premature to discuss issues related to  SPI. Researchers should rather
+focus on the trade-off between development speed  and accumulated technical debt
+\cite{Brown:2010:MTD:1882362.1882373}, which  appears to be the most important
+determinant for the future of  software development in the company.
+
+Future research can investigate the performance drop down caused by the
+necessity of returning the accumulated technical debt while expanding the
+company's operation and structuring mitigation strategies with four software
+development objectives: integrating scalable solutions with fast iterations and
+minimal set of functionalities (this allows startups to maintain effective
+planning and realistic expectations); empowering team members enabling them to
+operate horizontally in all the activities of the development environment
+simultaneously; improve desirable workflow patterns through the initiation of a
+minimal project management over time, as a natural result of emerging activities
+of tracing project progress and task assignment mechanisms; and finally, only
+when the chaos has been initially managed, planning long-term performance by
+adoption of Agile and Lean development practices.
+
+%------------------------------------------------------------------------------%
+%\subsection{Future work} %label{futwork} 
 %MUN: Don't have future work as a subsection in threats to validity. Either  
@@
 %influence the growth of a company after the release of the first product to the
-%market.
-
-%------------------------------------------------------------------------------%
-\section{Threats to Validity}
-\label{valt}
+%market. CG: ok
+
+%------------------------------------------------------------------------------%
+\section{Threats to Validity} \label{valt} 
 %MUN: Also here, introduce some structure: first, explain shortly what types of 
@@
 %this section in detail now since it will change considerably after your  
-%restructuring.
-%CG: fixed
-
-%MUN1: After reading the discussion on validity threats, I think you have
-%chosen the wrong framework (Wohlin) to discuss threats. His framework is 
-%targeted at experiments and even though it has been applied in other contexts, 
-%other taxonomies are a better fit (see Yin and also Runeson). You tried to 
-%squeeze your threats and strategies into Wohlin's taxonomy which seems in some 
-%cases quite artificial and forced. I think it is better if you adopt Yin's 
-%taxonomy of validity threats. If you read Yin (p 33-38) you'll see that his 
-%taxonomy is a much better fit and that you can easier place and motivate your 
-%mitigation strategies.
-%------------------------------------------------------------------------------%
-In this section we discuss the validity of the overall results. %MUN1: threats 
-%are discussed w.r.t. the research methodology, not the results... 
+%restructuring. %CG: fixed
+
+%MUN1: After reading the discussion on validity threats, I think you have chosen
+%the wrong framework (Wohlin) to discuss threats. His framework is  targeted at
+%experiments and even though it has been applied in other contexts,  other
+%taxonomies are a better fit (see Yin and also Runeson). You tried to  squeeze
+%your threats and strategies into Wohlin's taxonomy which seems in some  cases
+%quite artificial and forced. I think it is better if you adopt Yin's  taxonomy
+%of validity threats. If you read Yin (p 33-38) you'll see that his  taxonomy is
+%a much better fit and that you can easier place and motivate your  mitigation
+%strategies. CG: to discuss since the ref of Yin and Runeson present very
+%similar validity threats.
+%------------------------------------------------------------------------------
+In this section we discuss the validity of the overall methodology. 
+%MUN1: threats  %are discussed w.r.t. the research methodology, not the 
+%results... CG: fixed. 
 We structure the discussion according to Wohlin's taxonomy \cite{Wohlin2000}.
@@
 \subsection{External validity}
-%External
-One threat to external validity is the selection of subjects interviewed 
-for the study. This threat profoundly affects GT, since it is a qualitative 
+
+One threat to external validity is the selection of subjects interviewed  for
+the study. This threat profoundly affects GT, since it is a qualitative
 research method, which makes the use of semi-structured interviews, and centers
-the research on respondent's opinions. 
-In order to mitigate this threat we selected interviewees which 
-covered positions of CTOs' and CEOs'. Their perspectives on what is taking 
-place within the organization, were the only meaningful data taken in 
-consideration in the study.  In many cases there was no supporting evidence to 
-verify the opinion expressed. The questionnaire dedicated to ``data 
-triangulation'' was not of primary support because of the limited time 
-practitioners were able to grant to it. However, researchers have to accept the 
-veracity of what respondents reported during the interviews \cite{Coleman2008a}. 
-%MUN1: candidate for removal
+the research on respondent's opinions.  In order to mitigate this threat we
+selected interviewees which  covered positions of CTOs' and CEOs'. Their
+perspectives on what is taking  place within the organization, were the only
+meaningful data taken in  consideration in the study.  In many cases there was
+no supporting evidence to  verify the opinion expressed. The questionnaire
+dedicated to ``data  triangulation'' was not of primary support because of the
+limited time  practitioners were able to grant to it. However, researchers have
+to accept the  veracity of what respondents reported during the interviews
+\cite{Coleman2008a}.  %MUN1: candidate for removal 
 %Notwithstanding, the issues surrounding semi-structured interviews are vital.  
@@
 %described, the reported data are CTOs' and CEOs' perception, which is used to  
-%base their decisions.
+%base their decisions. CG: ok
 
@@
 domain to which the study's findings can be generalized. In particular the
-previous framework developed by Coleman \cite{Coleman2008} has allowed 
-us to find similarities, differences and broader reasoning related to 
-factors that hinder maturing processes in startups. Moreover, we analyzed 
-Brook's framework describing the \textit{basic attacks} to the challenges of 
-developing software systems and other literature covered by the SMS on startups.
+previous framework developed by Coleman \cite{Coleman2008} has allowed  us to
+find similarities, differences and broader reasoning related to  factors that
+hinder maturing processes in startups. Moreover, we analyzed  Brook's framework
+describing the \textit{basic attacks} to the challenges of  developing software
+systems and other literature covered by the SMS on startups.
 \cite{SMS}\footnote{Note that the generalizability of results is applied only to
@@
 %the subcategories, a continuous re-evaluation of the interview transcripts has 
-%been performed.
-
-%------------------------------------------------------------------------------%
-\subsection{Internal validity}
-%Internal
+%been performed. CG: ok
+
+%------------------------------------------------------------------------------%
+\subsection{Internal validity} %Internal 
 To enhance the internal validity, we created a three-dimensional research  
 framework. By means of a grounded theory approach, formed by a  
-\textit{systematic mapping study}, \textit{interviews} and \textit{follow-up 
-questionnaires}, we searched for convergence among different sources of 
-information to confirm the generated theory.
+\textit{systematic mapping study}, \textit{interviews}
+and \textit{follow-up  questionnaires}, we searched for convergence among
+different sources of  information to confirm the generated theory.
 
@@
 stated: ``Tying the emergent theory to existing literature enhances the internal
-validity, generalizability, and theoretical level of the theory building from 
-a case study research [\ldots] because the findings often rest on a very limited 
+validity, generalizability, and theoretical level of the theory building from  a
+case study research [\ldots] because the findings often rest on a very limited
 number of cases.''
@@
 %studies in the broader fields, we revealed that the theory resembles in wider  
-%startup development context. 
+%startup development context. CG: ok
 Nevertheless, we identified important confounding factors, related to
-innovation, market requirements and developer experience, which might 
-impact the internal validity of the GSM.
-
-%------------------------------------------------------------------------------%
-\subsection{Construct validity}
-%Construct
+innovation, market requirements and developer experience, which might impact
+the internal validity of the GSM.
+
+%------------------------------------------------------------------------------%
+\subsection{Construct validity} %Construct 
 One threat to this study is a possible inadequate description of constructs. In 
@@
 each theoretical conceptualization. For instance, we defined \textit{Time  
-shortage} in terms of \textit{Investor pressure, CEO/business 
-pressure, Demo presentations at events and internal final deadline} as used by 
-most of the interviewees in the study. 
+shortage} in terms of \textit{Investor pressure, CEO/business  pressure, Demo 
+presentations at events and internal final deadline} as used by  most of the 
+interviewees in the study.
 Moreover, during the coding of interview transcripts, we adopted explanatory
@@
 %MUN1: I get the threat, but I don't understand its mitigation: let the
-%interviewees drive the study?! Try to reformulate...
+%interviewees drive the study?! Try to reformulate... CG: fixed. 
 The second important threat is caused by the fact that interviewees might  
@@
 reduce the risk that some hypothesis could have been involved in the design of  
-the research, we let mainly participants drive the direction of the case study.
-
-Another threat associated with construct validity is evaluation 
-apprehension, which describes people's rejection attitude towards the 
-assessment of their behavior. 
-To address this risk we didn't attained to first answers from practitioners, but 
-rather went towards details of engineering elements and activities. %MUN1: <-- 
-%I don't understand this...
+the research, we didn't disclose any goal and emergent results of our research 
+to the interviewees.
+
+Another threat associated with construct validity is evaluation  apprehension,
+which describes people's rejection attitude towards the  assessment of their
+behavior.  To address this risk we didn't attained to first answers from
+practitioners, but  rather we asked for details of artifacts and engineering
+strategies. %MUN1: <--  %I don't understand this... CG: fixed 
 Moreover, we developed a rigorous data collection protocol to create case study 
@@
 %further mitigation strategy to evaluation apprehension. I don't see however
-%how this mitigates this risk. Fix...
-Reporting bias was mitigated by packaging all the needed 
-material for conducting new researches in other contexts in order to criticize 
-our findings (interview package with instructions has been made available 
-online\footnote{Available at 
-\url{https://github.com/adv0r/BTH-Interview-Package}.}. %MUN1: --> What exactly 
+%how this mitigates this risk. Fix... CG: right, removed because misleading.
+Reporting bias was mitigated by packaging all the needed  material for
+conducting new researches in other contexts in order to criticize  our findings
+(interview package with instructions has been made available
+online\footnote{Available at  \url{https://github.com/adv0r/BTH-Interview-
+Package}.}. %MUN1: --> What exactly  
 %is reporting bias? Let's assume that it means that you only report results,  
@@
 %study, i.e. makes is possible to test your theory in a different context. 
-Moreover, two researchers not involved in the execution of the study conducted 
-a peer-review analysis of the model's constructs.
+%CG: by this, we meant providing all the material to replicate the same study 
+%with other startups and compare the results. 
+Moreover, two researchers not involved in the execution of the study
+conducted  a peer-review analysis of the model's constructs.
 
@@
 
-The  approach of generating theory allowed us to check emerging categories 
-and their properties by gathering new evidence within an iterative and 
-evolutionary approach, in view of the wide context of the research area that was 
-under investigation.
-
-%------------------------------------------------------------------------------%
-%\subsection{Subsection Heading Here}
-%Subsection text here.
-
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
-
-%\subsubsection{Subsubsection Heading Here}
-%Subsubsection text here.
-
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-
-% Note that IEEE CS typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-% However, the Computer Society has been known to put floats at the bottom.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE CS papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals use top floats exclusively.
-% However, Computer Society journals sometimes do use bottom floats - bear
-% this in mind when choosing appropriate optional arguments for the
-% figure/table environments.
-% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
-% floats. This can be corrected via the \fnbelowfloat command of the
-% stfloats package.
-
-\section{Conclusion}
-\label{conc}
+The  approach of generating theory allowed us to check emerging categories  and
+their properties by gathering new evidence within an iterative and  evolutionary
+approach, in view of the wide context of the research area that was  under
+investigation.
+
+%------------------------------------------------------------------------------%
+
+\section{Conclusion} \label{conc} 
 Startups are able to produce cutting-edge software products with a wide impact  
@@
 have been validated by means of systematic comparisons with the state-of-the-art
-results and empirical data. The systematic mapping study \cite{SMS}, 
-with 37 studies extracted, revealed a 
-multi-faceted state-of-the-art inadequate to support software development 
-activities in startup companies. %MUN: <-- you mix again SMS and GT here... fix 
-%it CG: not fixed, I would keep it since important part of the initial GT 
-%methodology
+results and empirical data. The systematic mapping study \cite{SMS},  with 37
+studies extracted, revealed a  multi-faceted state-of-the-art inadequate to
+support software development  activities in startup companies. 
+%MUN: <-- you mix again SMS and GT here... fix it CG: not fixed, I would keep 
+%it since important part of the initial GT methodology 
 %MUN1: Ok, I understand that. But this paper is about the GT, ONLY the GT. The  
 %SMS is published elsewhere. Maybe it is ok how you use the SMS now... let's  
-%see what Tony says...
+%see what Tony says... CG: ok 
 On the other hand, the case study conducted in 13 early-stage startups,  
@@
 
-The overall results of our research confirm that startups possess %MUN: 
-%why unique? CG: fixed.
+The overall results of our research confirm that startups possess 
+%MUN:  %why unique? CG: fixed. 
 characteristics of uncertainty, lack of resources and time-pressure. These  
-factors influence the software development to an extent that transforms every 
-decision related to the development strategies into a difficult trade-off for 
-the company. Moreover, although startups share different characteristics with 
-other similar SE domains (e.g. market-driven development, small companies and 
-web engineering), the unique combination of coexisting factors poses a whole new 
-set of challenges which need to be addressed by primary studies. Especially when 
-bringing the first product to market, startups' most urgent priority is 
-releasing the product as quickly as possible to verify the product/market fit 
-and to adjust the business and product trajectory according to early feedback 
-and collected metrics. In this stage, startups often discard any formal project 
-management, documentation, analysis, planning, testing and other traditional 
-process activities. Practitioners take advantage of an evolutionary prototyping 
-approach, using well-integrated tools and externalizing complexity to third 
-party solutions.
+factors influence the software development to an extent that
+transforms every  decision related to the development strategies into a
+difficult trade-off for  the company. Moreover, although startups share
+different characteristics with  other similar SE domains (e.g. market-driven
+development, small companies and  web engineering), the unique combination of
+coexisting factors poses a whole new  set of challenges which need to be
+addressed by primary studies. Especially when  bringing the first product to
+market, startups' most urgent priority is  releasing the product as quickly as
+possible to verify the product/market fit  and to adjust the business and
+product trajectory according to early feedback  and collected metrics. In this
+stage, startups often discard any formal project  management, documentation,
+analysis, planning, testing and other traditional  process activities.
+Practitioners take advantage of an evolutionary prototyping  approach, using
+well-integrated tools and externalizing complexity to third  party solutions.
 
@@
 %of chaotic/complex domain in the paper, I think not. Hence you need to
-%rephrase that (but you can not introduce new concepts here in the conclusion). %CG: fixed. 
+%rephrase that (but you can not introduce new concepts here in the conclusion).
+%CG: fixed.  
 Therefore, from a software development perspective, working with low-precision  

### 2013-03-04 article/main.tex
Fixed potential problems with references starting on a new line

@@
 worldwide every day as a result of an increase of new markets, accessible
-technologies, and venture capital \cite{8491286}. With the term
+technologies, and venture capital~\cite{8491286}. With the term
 \textit{software startups} we refer to those temporary organizations focused on
@@
 fail within two years from their creation, primarily due to self-destruction
-rather than competition \cite{Crowne2002}. Operating in a chaotic, rapidly
+rather than competition~\cite{Crowne2002}. Operating in a chaotic, rapidly
 evolving and uncertain domain, software startups face intense time-pressure from
-the market and are exposed to tough competition \cite{Maccormack2001,
+the market and are exposed to tough competition~\cite{Maccormack2001,
 Eisenhardt1998}. In order to succeed in this environment, it is crucial to
 choose the right features to build and be able to quickly adapt the product to
-new requests constrained by very limited amount of resources \cite{Sutton2000}.
+new requests constrained by very limited amount of resources~\cite{Sutton2000}.
 
@@
 since they develop software in a context where processes can hardly follow a
-prescriptive methodology \cite{Sutton2000, Coleman2005}. Despite startups share
+prescriptive methodology~\cite{Sutton2000, Coleman2005}. Despite startups share
 some characteristics with similar domains (e.g. small and web companies), the
 combination of different factors makes the specific development context quite
-unique \cite{Blank2005, Sutton2000}. Therefore, research is needed to support
-their engineering activities \cite{Coleman2005}, guiding practitioners in taking
+unique~\cite{Blank2005, Sutton2000}. Therefore, research is needed to support
+their engineering activities~\cite{Coleman2005}, guiding practitioners in taking
 decisions and avoiding choices that could easily lead the whole business to
-failure \cite{Kajko-Mattsson2008}. However, despite the impressive size of the
-startup ecosystem  \cite{ISI:000243253000007}, the state-of-the-art presents a
+failure~\cite{Kajko-Mattsson2008}. However, despite the impressive size of the
+startup ecosystem~\cite{ISI:000243253000007}, the state-of-the-art presents a
 gap. Through a \textit{Systematic Mapping Study} (SMS) of the literature,
-conducted by  Paternoster et al. \cite{SMS}, only few publications were found
+conducted by Paternoster et al.~\cite{SMS}, only few publications were found
 related to engineering activities in startups. Moreover the majority of these
@@
 
-With a cross-sectional study we focused the research context in a limited  time-
-frame to emphasize the nature of uncertainty in the development activities  and
-time pressure from the market, as discussed in \cite{Maccormack2001,
+With a cross-sectional study we focused the research context in a limited 
+time-frame to emphasize the nature of uncertainty in the development activities 
+and time pressure from the market, as discussed in~\cite{Maccormack2001,
 Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and
@@
 commonalities between the issues related to software development in startups and
-the research focused at studying technical debt \cite{Nugroho2011,Izurieta2012}.
+the research focused at studying technical debt~\cite{Nugroho2011,Izurieta2012}.
 On the other hand, researchers can take advantage from the GSM to understand
@@
 The remainder of this paper is structured as follows: we provide relevant
-information about startups in section \ref{backg}. Section \ref{resmet}
+information about startups in section~\ref{backg}. Section~\ref{resmet}
 introduces the research questions and discusses the design and execution of the
-research process. Results are discussed in section \ref{res:gsm}, which presents
-the GSM. Section \ref{res:val}  compares results of this research to  literature
-and frameworks presented in this field. Section \ref{valt} discusses  the
-validity threats considered during this study. Section  \ref{sect:theory:impl}
+research process. Results are discussed in section~\ref{res:gsm}, which presents
+the GSM. Section~\ref{res:val} compares results of this research to literature
+and frameworks presented in this field. Section~\ref{valt} discusses the
+validity threats considered during this study. Section~\ref{sect:theory:impl}
 discusses the most relevant implications of the GSM. Finally, this paper
-concludes in section \ref{conc}.
+concludes in section~\ref{conc}.
 
@@
 markets in the middle of the 1990s and culminated with the notorious \textit
-{dot-com bubble burst} of 2000 \cite{Perkins:1999:IBI:555126,
+{dot-com bubble burst} of 2000~\cite{Perkins:1999:IBI:555126,
 storey1982entrepreneurship}. Several years later, with the massification of the
@@
 The easy access to vast potential markets and the low cost of services
-distribution are appealing condition for modern entrepreneurs \cite{Marmer2011}.
+distribution are appealing condition for modern entrepreneurs~\cite{Marmer2011}.
 Inspired by success stories, a large number of software businesses are created
 every day. However, the great majority of these companies fail within two years
-from their creation \cite{Crowne2002}.
+from their creation~\cite{Crowne2002}.
 
 Looking at the number of new business incubators which appeared in the last
-decade one can estimate the importance of startups \cite{Grimaldi2005}. The wave
+decade one can estimate the importance of startups~\cite{Grimaldi2005}. The wave
 of disruption in new technologies has led companies to be more and more
 competitive, forcing themselves to radical organizational and innovational
-renewals,  which bring many companies to the attempt of behaving like startups
-\cite{Christensen1997}.
+renewals, which bring many companies to the attempt of behaving like 
+startups~\cite{Christensen1997}.
 
 The implementation of methodologies to structure and control the development
-activities in startups is a major challenge for engineers \cite{Coleman2008}.
+activities in startups is a major challenge for engineers~\cite{Coleman2008}.
 Generally, the management of software development is achieved through the
@@
 organizations should take at each stage of production and provide assistance in
-making estimates, developing plans and measuring quality \cite{Chroust1996}. In
+making estimates, developing plans and measuring quality~\cite{Chroust1996}. In
 the last decades, several models have been introduced to control software
 development activities. However, their application in startup companies doesn't
-report significant benefits \cite{Coleman2008a, Coleman2008, Sutton2000}.
+report significant benefits~\cite{Coleman2008a, Coleman2008, Sutton2000}.
 
@@
 startups as creative and flexible in nature and reluctant to introduce process
-or bureaucratic measures which may result ineffective \cite{Sutton2000}. Also
+or bureaucratic measures which may result ineffective~\cite{Sutton2000}. Also
 Bach refers to startups as \textit{``a bunch of energetic and committed people
-without defined development processes''} \cite{Bach1998}. In fact, startups have
+without defined development processes''}~\cite{Bach1998}. In fact, startups have
 very limited resources and typically wish to use them to support product
-development instead of establishing processes \cite{Coleman2008, Heitlager2007}.
+development instead of establishing processes~\cite{Coleman2008, Heitlager2007}.
 Some attempts to taylor lightweight processes to startups reported basic failure
 of their application: \textit{``Everyone is busy, and software engineering
-practices are often one of the first places developers cut corners''}
-\cite{Martin2007}. Rejecting the notion of repeatable and controlled processes,
-startups prominently take advantage of unpredictable, reactive and low-precision
-\cite{surviving-os-cockburn} engineering practices \cite{Sutton2000,
-Tanabian2005, Chorev2006, Kakati2003}.
+practices are often one of the first places developers cut 
+corners''}~\cite{Martin2007}. Rejecting the notion of repeatable and controlled 
+processes, startups prominently take advantage of unpredictable, reactive and 
+low-precision~\cite{surviving-os-cockburn} engineering 
+practices~\cite{Sutton2000,Tanabian2005, Chorev2006, Kakati2003}.
 
 Moreover, most startups develop packaged applications rather than software for
-a specific client \cite{genome2012}. Issues related to this domain are addressed
+a specific client~\cite{genome2012}. Issues related to this domain are addressed
 in literature by the area known as \textit{market-driven software development}
 (sometimes called \textit{packaged software development} or \textit{COTS
-software development} \cite{Alves2006}). Among other results, researchers
-emphasize the importance of time-to-market as a key strategic objective
-\cite{dagMDR,sawyer99} for companies operating in this domain.  Other peculiar
-aspects which influence the software development are related to requirements,
-which are reported to be often ``invented by the software company''
-\cite{512553}, ``rarely documented'' \cite{Karlsson02challengesin}, and can be
-validated only after the product is released to market
-\cite{dahl2003,Keil:1995}. Under these circumstances, failure of product
-launches are largely due to ``products not meeting customer needs''
-\cite{Alves2006}.
+software development}~\cite{Alves2006}). Among other results, researchers
+emphasize the importance of time-to-market as a key strategic 
+objective~\cite{dagMDR,sawyer99} for companies operating in this domain. Other 
+peculiar aspects which influence the software development are related to 
+requirements, which are reported to be often ``invented by the software 
+company''~\cite{512553}, ``rarely documented''~\cite{Karlsson02challengesin}, 
+and can be validated only after the product is released to 
+market~\cite{dahl2003,Keil:1995}. Under these circumstances, failure of product
+launches are largely due to ``products not meeting customer 
+needs''~\cite{Alves2006}.
 
@@
 with workflows that leave them the ability to quickly change the direction
-according to the targeted market  \cite{Heitlager2007,Sutton2000}. In this
+according to the targeted market~\cite{Heitlager2007,Sutton2000}. In this
 regard, many startups focus on team productivity, asserting more control to the
-employees instead of providing them rigid guidelines \cite{Tanabian2005,
+employees instead of providing them rigid guidelines~\cite{Tanabian2005,
 Chorev2006, Kakati2003}.
@@
 \subsection{General lack of research in startups} Despite some studies tried to
-address the above mentioned issues, a SMS  \cite{SMS} revealed only a few SE
-works in this specific area, as confirmed by  other studies \cite{Coleman2008,
-Coleman2008a, Coleman2007, Sutton2000}.  Moreover, the studies, identified in
-\cite{SMS}, appear to be highly fragmented  and spread across different areas
+address the above mentioned issues, a SMS~\cite{SMS} revealed only a few SE
+works in this specific area, as confirmed by other studies~\cite{Coleman2008,
+Coleman2008a, Coleman2007, Sutton2000}. Moreover, the studies, identified 
+in~\cite{SMS}, appear to be highly fragmented and spread across different areas
 rather than constituting a consistent \textit{body of knowledge}.
 
-A research published in 2000 by Sutton \cite{Sutton2000} confirmed a general
+A research published in 2000 by Sutton~\cite{Sutton2000} confirmed a general
 lack of studies in this area, claiming that \textit{``software startups
 represent a segment that has been mostly neglected in process studies''} and it
-has been further confirmed with the empirical studies of Coleman et al.
-\cite{Coleman2008,Coleman2008a,Coleman2007} eight years later.
+has been further confirmed with the empirical studies of Coleman et 
+al.~\cite{Coleman2008,Coleman2008a,Coleman2007} eight years later.
 
@@
 The word \textit{startup} appeared in the SE literature for the first time in
-1994 in an article written by Carmel \cite{Camel1994a} where he studied the
+1994 in an article written by Carmel~\cite{Camel1994a} where he studied the
 time-to-completion in a young package firm. %MUN: If this is a quote  
@@
 First insights reveal how software startups are product-oriented in the first
-period of their development phase \cite{Heitlager2007}. Despite good
+period of their development phase~\cite{Heitlager2007}. Despite good
 achievements at the beginning, software development and organizational
-management increase in complexity \cite{1456074,Banker1998} causing
+management increase in complexity~\cite{1456074,Banker1998} causing
 deterioration of performance over time. Briefly, the necessity of establishing
@@
 
-A study of Kajko-Mattsson \cite{Kajko-Mattsson2008}, which investigated a
+A study of Kajko-Mattsson~\cite{Kajko-Mattsson2008}, which investigated a
 Swedish software startup, reported a heavy lack of requirements gathering
@@
 approaching to an upcoming IPO started to require processes to focus on scalable
-solutions, in view of the growing company's size in terms of users and employees
-\cite{Ambler2002}. In this regard, Crowne \cite{Crowne2002} specifies different
-stages through which software startups evolve. Starting without any established
-workflows, startups grow over time, creating and stabilizing processes to
-eventually improving them only when sufficiently mature.
+solutions, in view of the growing company's size in terms of users and 
+employees~\cite{Ambler2002}. In this regard, Crowne~\cite{Crowne2002} specifies 
+different stages through which software startups evolve. Starting without any 
+established workflows, startups grow over time, creating and stabilizing 
+processes to eventually improving them only when sufficiently mature.
 
 But when startups have no time for training and orienting activities, as
-discussed in \cite{Sutton2000}, their main focus remains on team capabilities
+discussed in~\cite{Sutton2000}, their main focus remains on team capabilities
 instead of prescriptive processes, hiring people who can \textit{``hit the
-ground running''} \cite{Yoffie1999}. Empowering the team and focus on
-methodological attributes of the processes oriented in prototyping,  proof-of-
-concepts, mock-ups and demos, to test basic functionalities, have been  the
-primary priority in startups \cite{Camel1994a}. Only when they grow, formal
+ground running''}~\cite{Yoffie1999}. Empowering the team and focus on
+methodological attributes of the processes oriented in prototyping, 
+proof-of-concepts, mock-ups and demos, to test basic functionalities, have been  
+the primary priority in startups~\cite{Camel1994a}. Only when they grow, formal
 methodologies arise, followed by performing quality assurance and long-term
-planning processes \cite{Yoffie1999}.
+planning processes~\cite{Yoffie1999}.
 
 The maturity of a company affects the extent to which processes are adopted.
-Tingling \cite{Tingling2007} reports how introducing Extreme Programming (XP)
-principles \cite{Beck:2004:EPE:1076267} in the development process was
+Tingling~\cite{Tingling2007} reports how introducing Extreme Programming (XP)
+principles~\cite{Beck:2004:EPE:1076267} in the development process was
 challenging because of the need of trained team-members for fully implementing
-the methodology. Da Silva and Kon \cite{Silva2005} were able to start with all
+the methodology. Da Silva and Kon~\cite{Silva2005} were able to start with all
 the XP practices in place only after six months of coaching the team, trying to
@@
 practices were inevitably implemented to adapt the processes to the undertaken
-startups' context \cite{Deias}.
+startups' context~\cite{Deias}.
 
 As partially discussed above, contributions to flexibility and reactiveness of
-the development process has been conducted prominently by means of \textit{Lean}
-\cite{Gautam2008} and \textit{Agile} \cite{Abrahamsson2002} methodologies (also
-reported in \cite{Taipale2010, Kuvinka2011}), where the extreme uncertain
-conditions lead startups to learn fast from trials and errors with a strong
-customer relationship in order to avoid wasting time in building wrong
-functionalities and prevent rapidly exhaustion of resources \cite{Midler2008,
-Hilmola2003, Sutton2000}. Customer involvement in software development has also
-been discussed by Yogendra \cite{Yogendra2002} as an important factor to
-encourage an early alignment of business concerns to the technology strategies,
-because both are salient considerations to be successful \cite{Tingling2007}.
+the development process has been conducted prominently by means of 
+\textit{Lean}~\cite{Gautam2008} and \textit{Agile}~\cite{Abrahamsson2002} 
+methodologies (also reported in~\cite{Taipale2010, Kuvinka2011}), where the 
+extreme uncertain conditions lead startups to learn fast from trials and errors 
+with a strong customer relationship in order to avoid wasting time in building 
+wrong functionalities and prevent rapidly exhaustion of 
+resources~\cite{Midler2008,Hilmola2003, Sutton2000}. Customer involvement in 
+software development has also been discussed by Yogendra~\cite{Yogendra2002} as 
+an important factor to encourage an early alignment of business concerns to the 
+technology strategies, because both are salient considerations to be 
+successful~\cite{Tingling2007}.
 
@@
 workflow that helps them quickly to change direction according to the target
-market \cite{Sutton2000}. In this regard, many startups have been focused on
+market~\cite{Sutton2000}. In this regard, many startups have been focused on
 team productivity, providing more control to the employees instead of providing
-them rigid guidelines \cite{Tanabian2005, Chorev2006, Kakati2003}.
+them rigid guidelines~\cite{Tanabian2005, Chorev2006, Kakati2003}.
 
@@
 light-weight processes such as a design of development process based on XP,
-proposed by Zettel \cite{Zettel2001}. The process consists of a set of
+proposed by Zettel~\cite{Zettel2001}. The process consists of a set of
 activities and artifacts (in addition to some important roles) defined in order
@@
 introducing a \textit{Helical Model for managing e-commerce development
-environment} \cite{Deakins2005}. Also in this case, the author prescribed broad
+environment}~\cite{Deakins2005}. Also in this case, the author prescribed broad
 guidelines for a rapid high-quality development process, which underwent limited
@@
 The problem of \textit{one-size-fits-all}, related to the SPI representations
-for startups, is described by Fayad \cite{Fayad1997}. The author discusses the
+for startups, is described by Fayad~\cite{Fayad1997}. The author discusses the
 problem in actuating the same \textit{best-practices} criteria for established
-companies in 10-person software startups. Thoroughly remarked in
-\cite{Sutton2000}, Sutton states that problems of SPI in software startups arise
-because of: the dynamism of the development process, which precludes
+companies in 10-person software startups. Thoroughly remarked 
+in~\cite{Sutton2000}, Sutton states that problems of SPI in software startups 
+arise because of: the dynamism of the development process, which precludes
 repeatability; organizational maturity, that cannot be maintained from startups
@@
 the development team's creativity and flexibility, and to the need of a quick
-delivering product process environment \cite{Coleman2008a}. 
+delivering product process environment~\cite{Coleman2008a}. 
 %MUN: the second part of the sentence is just a statement without reference. If 
@@
 Product quality is left aside in favor of minimal and suitable functionalities 
-to shorten the time-to-market. As reported in two studies of Mater and Mirel  
-\cite{Mater2000, Mirel2000}, quality aspects, mostly taken in consideration in  
-internet startups, are oriented to usability and scalability, even though the  
-market and application type heavily influences the quality-demand  
-\cite{Coleman2008, Kim2005}.
+to shorten the time-to-market. As reported in two studies of Mater and 
+Mirel~\cite{Mater2000, Mirel2000}, quality aspects, mostly taken in 
+consideration in internet startups, are oriented to usability and scalability, 
+even though the market and application type heavily influences the 
+quality-demand~\cite{Coleman2008, Kim2005}.
 
@@
 functionality, studies suggest to externalize the complexity of parts of the
-project to third party solutions by means of outsourcing activities
-\cite{Hanna2010}, software reuse \cite{Jansen2008} and open-source strategies ,
-\cite{Wall2001, Bean2005}.  %MUN: Can you split up the references? I.e. write:
-%"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and open-source
-%strategies \cite{Bla}. %CG: fixed
+project to third party solutions by means of outsourcing 
+activities~\cite{Hanna2010}, software reuse~\cite{Jansen2008} and open-source 
+strategies~\cite{Wall2001, Bean2005}. %MUN: Can you split up the references? 
+%I.e. write:"outsourcing activities \cite{Bla}, software reuse \cite{Bla}, and 
+%open-source strategies \cite{Bla}. %CG: fixed
 
@@
 trying to tackle the problem of the \textit{technical debt}, brings and
-encompasses different implications in studying development in software startups
-\cite{Tom2013}. The metaphoric neologism was originally introduced by
-Cunningham in 1992 \cite{TechnicalDebtCunn} and has recently attracted the
+encompasses different implications in studying development in software 
+startups~\cite{Tom2013}. The metaphoric neologism was originally introduced by
+Cunningham in 1992~\cite{TechnicalDebtCunn} and has recently attracted the
 attention of SE researchers\footnote{To attest the fact that \textit{technical
@@
 contributions which characterize the \textit{``debt landscape''}:
-\cite{Nugroho2011,Izurieta2012} and a dedicated workshop \cite{workshopDebt}
+\cite{Nugroho2011,Izurieta2012} and a dedicated workshop~\cite{workshopDebt}
 organized by the Software Engineering Institute and ICSE.}. The concept of
-\textit{technical debt} is  well illustrated in
-\cite{Brown:2010:MTD:1882362.1882373}: \textit{``The idea  is that developers
+\textit{technical debt} is well illustrated 
+in~\cite{Brown:2010:MTD:1882362.1882373}: \textit{``The idea is that developers
 sometimes accept compromises in a system in one dimension (e.g., modularity) to
@@
 Since \textit{``all decisions related to product development are trade-off
-situations''} \cite{Hilmola2003}, generally startups optimize workflows to the
+situations''}~\cite{Hilmola2003}, generally startups optimize workflows to the
 dynamic context they are involved into. They typically adopt any development
 style that might work to support their first needs in what is called the
-\textit{``Just do it''} school of software startups \cite{Ries2011}.
+\textit{``Just do it''} school of software startups~\cite{Ries2011}.
 Additionally, as remarked by Coleman, \textit{``many managers just decide to
 apply what they know, as their experience tells them it is merely common
-sense''} \cite{Coleman2008}.
+sense''}~\cite{Coleman2008}.
 
@@
 Some of these attributes are related to the overall system design, while others
-are specific  to run time, design time, or user centric issues
-\cite{Microsoft2009}. 
+are specific to run time, design time, or user centric 
+issues~\cite{Microsoft2009}. 
 \item Growth: an increase in the company size respect to the initial 
@@
 time that goes from idea conception to the first open beta release of the
-software product} \footnote{The above mentioned goal is structured according to
-the GQM paradigm \cite{Basili1992}.}.
-
-Initially the boundaries of the research domain were set by means of  non-
-systematic literature survey, which provided the initial keywords used  further
-in the research process. Moreover it helped us in defining the research
+software product}\footnote{The above mentioned goal is structured according to
+the GQM paradigm~\cite{Basili1992}.}.
+
+Initially the boundaries of the research domain were set by means of 
+non-systematic literature survey, which provided the initial keywords used  
+further in the research process. Moreover it helped us in defining the research
 questions (RQs):
@@
 We investigated the software development approach undertaken by practitioners
-of startups. Following a \textit{Grounded Theory} (GT) methodology
-\cite{Glaser1978}, we executed 13 semi-structured interviews integrated with
-follow-up questionnaires. From this, we elaborated and extracted the Green-field
-startup model (GSM) explaining the underlying phenomenon of software development
-in startups.
+of startups. Following a \textit{Grounded Theory} (GT) 
+methodology~\cite{Glaser1978}, we executed 13 semi-structured interviews 
+integrated with follow-up questionnaires. From this, we elaborated and 
+extracted the Green-field startup model (GSM) explaining the underlying 
+phenomenon of software development in startups.
 
@@
 best approach to answer the question - What is going on here?''} and defined by
-its creators \cite{ColinRobson2009} as \textit{``a set of well-developed
+its creators~\cite{ColinRobson2009} as \textit{``a set of well-developed
 categories ( e.g. themes, concepts ) that are systematically interrelated
@@
 Despite GT has its roots in the social sciences, it has been increasingly used
-by Information System researchers and has been almost ignored by the SE research
-\cite{Coleman2007}. In 1997 Bertelsen advocated for a need of more qualitative
-research in SE \cite{Bertelsen1997}, and in the last decade we assisted to an
-increasing number of publications that used a GT-like approach in SE
-\cite{Coleman2008a,Coleman2007, Sulayman2012}.
+by Information System researchers and has been almost ignored by the SE 
+research~\cite{Coleman2007}. In 1997 Bertelsen advocated for a need of more 
+qualitative research in SE~\cite{Bertelsen1997}, and in the last decade we 
+assisted to an increasing number of publications that used a GT-like approach 
+in SE~\cite{Coleman2008a,Coleman2007, Sulayman2012}.
 
@@
 in the last years, researchers should indicate which \textit{``implementation''}
-of the theory is being used \cite{Coleman2007}. After the methodology was
+of the theory is being used~\cite{Coleman2007}. After the methodology was
 initially introduced, its original authors had divergences regarding how the
@@
 methodology, which leaves to the researchers an higher degree of freedom. The
-latter approach empowers researchers' \textit{``theoretical sensitivity''}
-\cite{Corbin1990}, and encourages them to outline the research problem
-beforehand. Since the knowledge obtained from parallel execution of the SMS
-\cite{SMS} and our direct experience with startups companies provided a good
+latter approach empowers researchers' \textit{``theoretical 
+sensitivity''}~\cite{Corbin1990}, and encourages them to outline the research 
+problem beforehand. Since the knowledge obtained from parallel execution of the 
+SMS~\cite{SMS} and our direct experience with startups companies provided a good
 initial level of knowledge, in this study we use the Corbin and Strauss
-approach \cite{Strauss1998}.
+approach~\cite{Strauss1998}.
 
@@
 release of the first product in highly scalable markets. Moreover, since most
-startups are web-oriented companies \cite{tc-webstartups-link, Allen2003}, in
+startups are web-oriented companies~\cite{tc-webstartups-link, Allen2003}, in
 our case study, we inquired web companies with little or no operating history.
@@
 A complete overview of the overall case study methodology and execution is
-shown in Figure \ref{fig:gt:completemethodology}, which presents how we tailored
+shown in Figure~\ref{fig:gt:completemethodology}, which presents how we tailored
 the general GT methodology to our specific needs.
@@
 The initial design of the case study is supported by the results of the
-\textit{Systematic Mapping Study} (SMS) \cite{SMS} which contributed to define
+\textit{Systematic Mapping Study} (SMS)~\cite{SMS} which contributed to define
 the initial questions for both questionnaires and semi-structured interviews.
-The process depicted in Figure \ref{fig:gt:completemethodology} is evolutionary
+The process depicted in Figure~\ref{fig:gt:completemethodology} is evolutionary
 and affects the design at each new iteration. In \textit{Data Collection} the
@@
 The process details are described in the following subsections, structured
-according to the five macro phases depicted in bold in Figure
-\ref{fig:gt:completemethodology}.
+according to the five macro phases depicted in bold in 
+Figure~\ref{fig:gt:completemethodology}.
 
@@
 This case study started based on a SMS of the literature regarding software
-development in startups \cite{SMS}. By means of this SMS we refined our research
+development in startups~\cite{SMS}. By means of this SMS we refined our research
 questions. The research questions have been narrowed down to outline the domain
 of our study, even though still broad enough to allow emerging new theories and
-to adjust the scope of the research \cite{Corbin1990}.
+to adjust the scope of the research~\cite{Corbin1990}.
 
 The sampling of candidate companies took place in two distinct phases. At first
-we executed an initial convenience sampling \cite{Dawson2009}, which led to the
+we executed an initial convenience sampling~\cite{Dawson2009}, which led to the
 identification of eight companies. Then we included five additional startups
@@
 the sample according to the emerging theory. The characteristics of the sampled
-companies are reported in Table \ref{t_interviews-stats}.
+companies are reported in Table~\ref{t_interviews-stats}.
 
@@
 artifacts\footnote{The interview package and follow-up questionnaire has been
-released under MIT License \cite{MITLicense} and it is available for download on
-Github \cite{GitHubInterviewPackage}. We encourage anyone interested in pushing
+released under MIT License~\cite{MITLicense} and it is available for download on
+Github~\cite{GitHubInterviewPackage}. We encourage anyone interested in pushing
 this work forward, to fork the repository and contribute to it. If used in a
@@
 questionnaires have been tailored to each startup, partially taking advantage of
-the repertory grid principles \cite{Edwards2009}. To obtain better chances of
+the repertory grid principles~\cite{Edwards2009}. To obtain better chances of
 quickly obtaining responses, questionnaires have been sent to respondents
@@
 case-study database allowed us to easily retrieve and seek for information,
-assembling the evidence from the case  study reports, as described also in
-\cite{Yin1994}. The database has been stored  and constructed using the
+assembling the evidence from the case study reports, as described also 
+in~\cite{Yin1994}. The database has been stored and constructed using the
 qualitative data analysis software package AtlasTI\footnote{Available online at
-\url{http://www.atlasti.com/}.} - one of  the most suitable tools for GT
-\cite{Coleman2007}.  We overlapped interviews with questionnaire results to
+\url{http://www.atlasti.com/}.} - one of the most suitable tools for 
+GT~\cite{Coleman2007}. We overlapped interviews with questionnaire results to
 adjust the data collection and take advantage of emergent themes and reveal
@@
 \item Labels were assigned to raw data, and a first low-level conceptualization
-was carried out using both in-vivo and open coding \cite{ColinRobson2009}. 
+was carried out using both in-vivo and open coding~\cite{ColinRobson2009}. 
 \item Concepts were grouped together into theoretical categories and 
@@
 \textit{``direct quoting from the transcript gives more expressive power to the
-data''} \cite{ColinRobson2009} combined with the more descriptive procedure of
+data''}~\cite{ColinRobson2009} combined with the more descriptive procedure of
 open coding. Following the example of other grounded theories, developed in
-related areas such as Information Systems \cite{Orlikowski1993} and Software
-Process Improvement \cite{Coleman2006}, we performed the high-level
+related areas such as Information Systems~\cite{Orlikowski1993} and Software
+Process Improvement~\cite{Coleman2006}, we performed the high-level
 conceptualization during creation of categories, in the process of refining
@@
 Given the large number of codes, categories, properties, propositions and
-related questions that evolved from the analytical process \cite{Corbin1990},
+related questions that evolved from the analytical process~\cite{Corbin1990},
 an important activity that helped the coding process was memoing. It constituted
@@
 theoretical framework is presented in the form of a network of categories and
-subcategories that are linked together according to cause-effect relationships
-\cite{Corbin1990}. During the formation of the theoretical framework the
-researchers operated by means of a  bottom-up approach. From the empirical data
-and coding process, the framework developed into two different levels: a
-detailed level representing the network of subcategories (identified mainly by
-the axial coding process), and a high-level representation of the main
+subcategories that are linked together according to cause-effect 
+relationships~\cite{Corbin1990}. During the formation of the theoretical 
+framework the researchers operated by means of a bottom-up approach. From the 
+empirical data and coding process, the framework developed into two different 
+levels: a detailed level representing the network of subcategories (identified 
+mainly by the axial coding process), and a high-level representation of the main
 categories network (identified mainly by the selective coding process).
@@
 
-As mentioned in Section \ref{desex}, emergent theories have been tested by
+As mentioned in Section~\ref{desex}, emergent theories have been tested by
 integrating additional companies into the sample, selected following the
-principle of theoretical sampling \cite{Yin1994}.
+principle of theoretical sampling~\cite{Yin1994}.
 
@@
 interview execution, where we systematically analyzed the theoretical framework
-by means of the paradigm model introduced by Corbin \cite{Corbin1990}. The
+by means of the paradigm model introduced by Corbin~\cite{Corbin1990}. The
 paradigm model is composed by:
@@
 Strauss and Corbin provided a list of questions to assist in determining how
-well the findings are grounded \cite{Strauss1998}. They  questions are as
+well the findings are grounded~\cite{Strauss1998}. They questions are as
 follows:
@@
 coding process. They were systematically related through the use of a network
-diagram (see Figure \ref{fig:gt:completemethodology}), by which we established
+diagram (see Figure~\ref{fig:gt:completemethodology}), by which we established
 the relations between the concepts. At each iteration of the grounded theory
@@
 existing studies, providing possible directions for future studies (see
-subsection \ref{sect:theory:validation:others}). Moreover we study the
+subsection~\ref{sect:theory:validation:others}). Moreover we study the
 confounding factors which could interfere with the application of the GSM model
-(see subsection \ref{sect:an:comp:cat-vs-literature:conf}).
-
-Then, we performed an \textit{evaluation with empirical data} (see subsection
-\ref{sect:theory:validation:stats}), quantifying scores for each company
-involved in our empirical study by defining metrics based on a set of rubrics
-and evaluating startups accordingly, through an analysis of interview
+(see subsection~\ref{sect:an:comp:cat-vs-literature:conf}).
+
+Then, we performed an \textit{evaluation with empirical data} (see 
+subsection~\ref{sect:theory:validation:stats}), quantifying scores for each 
+company involved in our empirical study by defining metrics based on a set of 
+rubrics and evaluating startups accordingly, through an analysis of interview
 transcripts and codes. The complete statistical procedure and provided rubrics
@@
 representing the underlying phenomenon have been grouped together to form high-
-level categories. Figure \ref{fig:gsm} shows the network  of causal
+level categories. Figure~\ref{fig:gsm} shows the network of causal
 relationships (represented by arrows) between categories (represented by
@@
 categories using the \textit{italic} font style. For the main categories shown
-in Figure \ref{fig:gsm}, when necessary we also make use of identifiers (i.e.
+in Figure~\ref{fig:gsm}, when necessary we also make use of identifiers (i.e.
 CATx) to facilitate the reading of this section. The network is centered around
@@
 interconnected node in the framework reflecting the fact that \textit{``it is
-the one (category) with the greatest explanatory  power''}  \cite{Strauss1998}.
+the one (category) with the greatest explanatory power''}~\cite{Strauss1998}.
 The meaning of the labels on the three arrows that reach the core category
-(efficiency, effectiveness and performance) is discussed  in detail in section
-\ref{sect:theory:validation:stats}. %MUN: this refers to the validity threats
+(efficiency, effectiveness and performance) is discussed in detail in 
+section~\ref{sect:theory:validation:stats}. 
+%MUN: this refers to the validity threats
 %section. Do you mean rather the "Theory Validation" section? %CG:fixed 
-%MUN1: this should be \ref{sect:theory:validation:stats}, right? Section 5 is  
+%MUN1: this should be~\ref{sect:theory:validation:stats}, right? Section 5 is  
 %quite long and you should point to the correct subsection. CG: fixed.
@@
 urgent needs\footnote{There are some exceptions where the quality aspects
-actually matter and such cases will be discussed in section \ref{sect:an:comp
+actually matter and such cases will be discussed in section~\ref{sect:an:comp
 :cat-vs-literature:conf}
@@
 performance in software development. To support a fast-paced production
-environment, engineers are required to be highly committed, co-located,  multi-
-role, and self-organized. In other words \textit{team is the catalyst of
+environment, engineers are required to be highly committed, co-located, 
+multi-role, and self-organized. In other words \textit{team is the catalyst of
 development}. With an essential and flexible workflow, which relies on tacit
@@
 In the following subsections we explain in detail all the categories presented
-in Figure \ref{fig:gsm}, and conclude in subsection \ref{res:gsm:th} with the
+in Figure~\ref{fig:gsm}, and conclude in subsection~\ref{res:gsm:th} with the
 final theory. In the explanations we use identifiers of the companies presented
-in Table \ref{t_interviews-stats} (i.e. Cx) to highlight statements recorded
+in Table~\ref{t_interviews-stats} (i.e. Cx) to highlight statements recorded
 from the interviewees.
@@
 decisions related to software development are fundamentally trade-off
-situations (confirming Himola's results \cite{Hilmola2003}).
+situations (confirming Himola's results~\cite{Hilmola2003}).
 %------------------------------------------------------------------------------%
@@
 prototype and iteratively refine it over time, similarly to the concept of
-\textit{``evolutionary prototyping''} \cite{EvProt}. The reason of using this
+\textit{``evolutionary prototyping''}~\cite{EvProt}. The reason of using this
 approach is that startups want to validate the product against the market as
@@
 survival in the short term (recalling Ries' definition of \textit{Minimum
-Viable Product} (MVP) \cite{Ries2011}).
+Viable Product} (MVP)~\cite{Ries2011}).
 
@@
 startups attempt to \textit{find what is valuable for customers} with activities
-in agreement with Blank's \textit{Customer validation} stage \cite{Blank2005}.
+in agreement with Blank's \textit{Customer validation} stage~\cite{Blank2005}.
 
@@
 achieved. Product metrics are typically web-based \textit{statistical hypothesis
-testing}, such as A/B testing \cite{AB}.
+testing}, such as A/B testing~\cite{AB}.
 %MUN: add reference or explain shortly what A/B testing is %CG: fixed
@@
 usability testing. In fact, the beta release is typically the first time that 
-the software is available outside of the organization that developed it.  
-\footnote{Detailed discussion of the impact of innovative products on the user  
-satisfaction is presented in subsection 
-\ref{sect:an:comp:cat-vs-literature:conf}}. 
+the software is available outside of the organization that developed 
+it\footnote{Detailed discussion of the impact of innovative products on the user 
+satisfaction is presented in 
+subsection~\ref{sect:an:comp:cat-vs-literature:conf}.}. 
 %MUN: do you use the Kano model in Theory Validation? Change text to refer to  
@@
 tools to manage low-precision lists of features to implement, written in form of
-self-explanatory user stories \cite{AgilePlan}. Practitioners intensively use
+self-explanatory user stories~\cite{AgilePlan}. Practitioners intensively use
 physical tools such as post-it notes and whiteboards, which help in making
@@
 interactions with third-party components only}. In particular, the use of well-
-known standards, frameworks and conventions limits the need of formal UML
-\cite{UML} diagrams and documentation, and provide a minimum level of
+known standards, frameworks and conventions limits the need of formal 
+UML~\cite{UML} diagrams and documentation, and provide a minimum level of
 maintainability costs. Additionally, since quality aspects are not a main
@@
 
-As discussed in section \ref{res:val} %MUN: this refers to the validity threats
+As discussed in section~\ref{res:val} %MUN: this refers to the validity threats
 %section. Do you mean rather the "Theory Validation" section? %CG:fixed
-, we formed the theory by applying the paradigm model, introduced by Corbin
-\cite{Corbin1990}. The theory is composed by different elements reflecting the
-paradigm model:
+, we formed the theory by applying the paradigm model, introduced by 
+Corbin~\cite{Corbin1990}. The theory is composed by different elements 
+reflecting the paradigm model:
 
@@
 In this section we discuss the validity of the GSM by means of 
-cross-methodological observations, as discussed in section \ref{rm:val}. 
+cross-methodological observations, as discussed in section~\ref{rm:val}. 
 %MUN: expand this discussion a bit (motivation why you do this like you do it)  
@@
 defined from the GSM that are in support of previous models developed by
-Coleman \cite{Coleman2007,Coleman2008a, Coleman2008}, Baskerville
-\cite{Internet} and finally Brooks \cite{BrooksJr1987}.
+Coleman~\cite{Coleman2007,Coleman2008a, Coleman2008}, 
+Baskerville~\cite{Internet} and finally Brooks~\cite{BrooksJr1987}.
 
@@
 letter of them and they are always adapted, so we adapted ours to the way we
-wanted it to work for us, for our own size and scale''} \cite{Coleman2008}.
+wanted it to work for us, for our own size and scale''}~\cite{Coleman2008}.
 %MUN: add ref to Coleman's paper where he states this %CG fixed
@@
 depends on the context in which the study is conducted and on the background
-knowledge of the researchers \cite{Pearl2011}.}  in our study since we were not
+knowledge of the researchers~\cite{Pearl2011}.} in our study since we were not
 able to ground this concept in the data.
@@
 being responsive in our position. That works well and I don't want to lose
-it.''} \cite{Coleman2007}.
+it.''}~\cite{Coleman2007}.
 
@@
 size-fits-all''} solutions have always found difficulty in penetrating small
-software organizations \cite{Staples2007}. When startups began  establishing any
-SPI process, they experienced \textit{process erosions}   \cite{Coleman2008}
+software organizations~\cite{Staples2007}. When startups began establishing any
+SPI process, they experienced \textit{process erosions}~\cite{Coleman2008}
 %MUN:  %add ref (Coleman?). You should add, for each category you use from
@@
 staff to carry out tasks with less direct supervision, greater delegation of
-responsibilities and a more generally consensual environment \cite{Coleman2008}.
+responsibilities and a more generally consensual environment~\cite{Coleman2008}.
 Nevertheless, software development managers and founders have still impact on
@@
 In contrast with some SPI processes, such as \textit{Six Sigma} which objective 
-is minimizing unpredictability in the definition of workflows \cite{Sixsigma}, 
+is minimizing unpredictability in the definition of workflows~\cite{Sixsigma}, 
 startups seem to follow the \textit{Fischer's Fundamental Theorem of natural  
-selection} \cite{Fisher}.
+selection}~\cite{Fisher}.
 Moving towards flexible and variable processes (easy to adapt to the uncertain
@@
 
-Also Baskerville \cite{Internet} refers to SPI as a development approach
+Also Baskerville~\cite{Internet} refers to SPI as a development approach
 typically effective only in large-scale, long-term development efforts with
@@
 
-With a wider research focus, Brooks \cite{BrooksJr1987} discussed what
+With a wider research focus, Brooks~\cite{BrooksJr1987} discussed what
 challenges are involved in constructing software products. Brooks divides
@@
 In this subsection we report the evaluation of our theory in contrast with the
-main contributions of the studies retrieved and discussed within the SMS
-\cite{SMS}. For each retrieved article we present the main results, mapping them
-to categories in accordance to the GSM model. Table \ref{tab:an:literature_comp}
-shows the main contributions of the 37 identified studies, where an 'X' is
-placed in correspondence to the theoretical category they considered. The table
-is sorted by the number of GSM categories covered by the studies.
+main contributions of the studies retrieved and discussed within the 
+SMS~\cite{SMS}. For each retrieved article we present the main results, mapping 
+them to categories in accordance to the GSM model. 
+Table~\ref{tab:an:literature_comp} shows the main contributions of the 37 
+identified studies, where an 'X' is placed in correspondence to the theoretical 
+category they considered. The table is sorted by the number of GSM categories 
+covered by the studies.
 
@@
 \renewcommand{\arraystretch}{1.3}
-\caption{SMS \cite{SMS} and GSM}
+\caption{SMS~\cite{SMS} and GSM}
 \label{tab:an:literature_comp}
@@
 development} (CAT4). The importance of people has been widely discussed in other
-studies in SE (Cooper \cite{Cooper1986}, DeMarco \cite{peopleware-demarco1999},
-Coleman \cite{Coleman2004}, Valtanen \cite{Valtanen2008},  Adolph
-\cite{Adolph2011} and Cockburn \cite{people-first-order}), advocating for the
-need of empowering people, which is the element with the greatest impact on
-software development. Humans, being non-linear variables \cite{Human} 
+studies in SE (Cooper~\cite{Cooper1986}, DeMarco~\cite{peopleware-demarco1999},
+Coleman~\cite{Coleman2004}, Valtanen~\cite{Valtanen2008}, 
+Adolph~\cite{Adolph2011} and Cockburn~\cite{people-first-order}), advocating for 
+the need of empowering people, which is the element with the greatest impact on
+software development. Humans, being non-linear variables~\cite{Human} 
 %MUN: add ref %CG:fixed
 , are unlikely to follow repeatable prescriptive methodologies. As
-stated by  Highsmith \cite{Highsmith2000}, \textit{``A drawback of each and
+stated by Highsmith~\cite{Highsmith2000}, \textit{``A drawback of each and
 every methodology is to expect people to behave consistently over time, when is
@@
 authors describe startups facing with inescapable constraints defined by
-engineering/business concerns and constant time-pressure \cite{Camel1994a,
+engineering/business concerns and constant time-pressure~\cite{Camel1994a,
 Cugola98softwareprocesses:}. Because of their young and inexperienced working
 conditions, startups present limited resources in terms of management and
-strategic alliances \cite{Sutton2000}.
-
-Crowne \cite{Crowne2002} identifies four main stages of a startup's lifecycle:
+strategic alliances~\cite{Sutton2000}.
+
+Crowne~\cite{Crowne2002} identifies four main stages of a startup's lifecycle:
 startup, stabilization, growth and maturity. Inexperienced practitioners
@@
 engineering practices during companies' growth are discussed in an experience
-paper by Ambler \cite{Ambler2002}. He reports on  two case studies of growing
+paper by Ambler~\cite{Ambler2002}. He reports on two case studies of growing
 Internet startups approaching an initial public offering (IPO). %MUN:explain
@@
 
-The use of whiteboards in working environments supports \textit{use of  well-
-integrated and simple tools}. In both the startups studied by Ambler,
+The use of whiteboards in working environments supports \textit{use of 
+well-integrated and simple tools}. In both the startups studied by Ambler,
 documentation was minimal, finding the ``sweet spot'' to have just barely enough
@@
 
-Kajko-Mattsson et al. \cite{Kajko-Mattsson2008} investigated a Swedish software
+Kajko-Mattsson et al.~\cite{Kajko-Mattsson2008} investigated a Swedish software
 startup company involved in mobile applications, reporting the following issues:
@@
 model of the release management process. 
-%They moved towards the plan phase where the  %researchers decided to develop a 
+%They moved towards the plan phase where the researchers decided to develop a 
 %model to manage releases, introducing a process composed by: release scope 
@@
 %MUN: solutions on what? CG: fixed 
-, two other studies \cite{Wall2001, Bean2005} suggest to \textit{``reuse the  
+, two other studies~\cite{Wall2001, Bean2005} suggest to \textit{``reuse the 
 wheel, instead of reinventing it''}, taking advantage of open source software 
@@
 changed from prescriptive to agile solutions, with less concern on quality
-aspects, also confirmed by Mater \cite{Mater2000}.
+aspects, also confirmed by Mater~\cite{Mater2000}.
 
@@
 \subsubsection{Simple workflows} Prescribed development models, oriented towards
-SPI, are basically ignored by  startups, as discussed in \cite{Zettel2001}.
+SPI, are basically ignored by startups, as discussed in~\cite{Zettel2001}.
 %MUN: <-- rephrase, not clear to me %CG:fixed 
@@
 
-Deakins \cite{Deakins2005} designed a new spiral methodology, constructed
+Deakins~\cite{Deakins2005} designed a new spiral methodology, constructed
 during a fieldwork with a developer team that had limited resources managing an
@@
 development approach} (CAT2). Evolutionary development approaches are 
-typically more suitable for web applications, as confirmed in  
-\cite{Deshpande2001}. Moreover, they provide practices oriented towards rapid 
+typically more suitable for web applications, as confirmed 
+in~\cite{Deshpande2001}. Moreover, they provide practices oriented towards rapid 
 and iterative design with complete freedom to the developer team, enabled to 
@@
 
-Carmel \cite{Camel1994a} reports on the need of flexible and rapid development
+Carmel~\cite{Camel1994a} reports on the need of flexible and rapid development
 solutions to shorten time-to-market, but even more important, the need of high-
@@
 those variables which and have not been considered in the model and might
-interfere with the theoretical framework positively or negatively
-\cite{ColinRobson2009}. We report those factors that have been explicitly
-identified by the SMS \cite{SMS} %MUN: do you mean the SMS? Then refer to it.
+interfere with the theoretical framework positively or 
+negatively~\cite{ColinRobson2009}. We report those factors that have been 
+explicitly identified by the SMS~\cite{SMS} 
+%MUN: do you mean the SMS? Then refer to it.
 %CG: fixed  
@@
 \subsubsection{Creativity and innovation}
-Heitlager et al. \cite{Heitlager2007} contributed to the investigation of
+Heitlager et al.~\cite{Heitlager2007} contributed to the investigation of
 creativity and innovation in startups. 
@@
 \subsubsection{Market requirements and application type} 
-Market requirements \cite{Coleman2007} and application type \cite{Sutton2000,  
+Market requirements~\cite{Coleman2007} and application type~\cite{Sutton2000, 
 Coleman2008,Coleman2008a} are two further potential confounding factors to the 
@@
 For example, safety-critical systems possess quality concerns that must be
-considered. Even in e-commerce systems factors related to quality have  high-
-impact effects on obtaining the first customer reference,  
+considered. Even in e-commerce systems factors related to quality have 
+high-impact effects on obtaining the first customer reference, 
 %MUN: <-- cannot parse it. Remove the first part and write: "Even e-commerce  
@@
 %CG: fixed 
-as discussed in \cite{Deakins2005,Kim2005,Silva2005}. Quality attributes
+as discussed in~\cite{Deakins2005,Kim2005,Silva2005}. Quality attributes
 especially %MUN: what is "this"? CG: fixed 
@@
 successfully implemented in other software systems or whose failure might 
-inflict serious financial or social damages \cite{Bass2003}. Yet, limiting our  
+inflict serious financial or social damages~\cite{Bass2003}. Yet, limiting our 
 context to early-stage software startups, which typically develop first 
@@
 %proper levels of customers' satisfaction, we  introduce the Kano model,
-%described in \cite{Ullah2011} and extensively  studied in product development.
+%described in~\cite{Ullah2011} and extensively  studied in product development.
 %As shown in the Kano model, users  preferences can be classified into four
@@
 Nevertheless, startups rely at the beginning on clever, but inexperienced
-developers \cite{Crowne2002}. Having team-members  with deep experience is a
-``double-edge sword'' \cite{Yoffie1999}. On one hand,  experience might quickly
+developers~\cite{Crowne2002}. Having team-members with deep experience is a
+``double-edge sword''~\cite{Yoffie1999}. On one hand, experience might quickly
 provide structure and maturity to the development process, on the other it
@@
 development environment, that is essential in early-stage startups.
-Accordingly, also Coleman \cite{Coleman2007, Coleman2008a,Coleman2008} states
+Accordingly, also Coleman~\cite{Coleman2007, Coleman2008a,Coleman2008} states
 that the operating strategies towards \textit{``minimum processes''} in
@@
 
-We argued in section \ref{res:gsm:frmov}, that the most essential feature
-startups aim to achieve is high speed. As illustrated in Figure \ref{fig:gsm},
+We argued in section~\ref{res:gsm:frmov}, that the most essential feature
+startups aim to achieve is high speed. As illustrated in Figure~\ref{fig:gsm},
 the core category (\textit{speed-up development}) is influenced by: the degree
@@
 that those relations were characterized in terms of "effectiveness" and 
-"efficiency" since time-to-market was the main concern (as confirmed in  
-\cite{SMS}):
-
-\begin{compactitem} \item The team enhances the overall \textit{performance} of
+"efficiency" since time-to-market was the main concern (as confirmed 
+in~\cite{SMS}):
+
+\begin{compactitem} 
+\item The team enhances the overall \textit{performance} of
 software development. Recalling that startups deal with severe lack of
@@
 importance of a skilled team for development performance is expressed in many
-studies such as  \cite{surviving-os-cockburn, Highsmith2000}, described in CAT4
-and confirmed by  Sutton \cite{Sutton2000}. \item The evolutionary approach
-enhances the \textit{effectiveness} in  implementing the correct
-functionalities. This approach improves the company's  capabilities to adjust
-the trajectory of product development (in agreement to  the concept of pivoting
-discussed by Ries \cite{Ries2011}). \item Since product quality has low
-priority, the development is focused only on  implementing a limited number of
-suitable functionalities, enhancing its  \textit{efficiency}. An essential
-prototype (MVP) does not require to comply  with heavy quality constraints,
-enabling the team to quickly implement  functionalities, ready to be validated
-by final users. \end{compactitem}
+studies such as~\cite{surviving-os-cockburn, Highsmith2000}, described in CAT4
+and confirmed by Sutton~\cite{Sutton2000}. 
+\item The evolutionary approach enhances the \textit{effectiveness} in 
+implementing the correct functionalities. This approach improves the company's 
+capabilities to adjust the trajectory of product development (in agreement to 
+the concept of pivoting discussed by Ries~\cite{Ries2011}). 
+\item Since product quality has low priority, the development is focused only on 
+implementing a limited number of suitable functionalities, enhancing its 
+\textit{efficiency}. An essential prototype (MVP) does not require to comply 
+with heavy quality constraints, enabling the team to quickly implement 
+functionalities, ready to be validated by final users. 
+\end{compactitem}
 
@@
 \begin{compactenum} 
-\item The more a capable \footnote{The team capability is
+\item The more a capable\footnote{The team capability is
 evaluated according to the characteristics presented in CAT4.} team, with few
@@
 categories related to \textit{speed-up development} presented in the GSM model
-(see figure \ref{fig:gsm}). We tested $H0_1$ and $H0_2$  with an  one-tailed
+(see figure~\ref{fig:gsm}). We tested $H0_1$ and $H0_2$ with an one-tailed
 test using Pearson's product moment correlation coefficient\footnote{To perform
@@
 the model. According to the paradigm model those are the causal condition which
-are linked to the core-category (see Figure \ref{fig:gsm}). There are four
+are linked to the core-category (see Figure~\ref{fig:gsm}). There are four
 high-level relationships which have not been validated by quantitative
@@
 customer feedback is a central aspect of the \textit{Lean startup methodology}
-pioneered by Ries \cite{Ries2011}. However, startups do not  formally follow the
+pioneered by Ries~\cite{Ries2011}. However, startups do not formally follow the
 cycle \textit{build-measure-learn} as it is proposed by the methodology, where
@@
 the step-by-step process of \textit{customer development} formally defined by 
-Blank \cite{Blank2005}. Instead, startups absorb and  implement the high-level
+Blank~\cite{Blank2005}. Instead, startups absorb and implement the high-level
 principles that outlines the customer development methodology
@@
 and traditional web engineering, where quality aspects are rated as very 
-important, even more than time-to-market \cite{Offutt2002}.}. From a business  
+important, even more than time-to-market~\cite{Offutt2002}.}. From a business 
 perspective, one might argue that established companies might take advantage of 
@@
 operate in a market of \textit{innovators} and \textit{early adopters} group 
-of users - described by the Roger's bell curve in \cite{Rogers} - because  of 
-the little return of investment they would gain, as discussed by Christensen  in
-\cite{Christensen2003}. Instead, established companies have more interests  in
-acquiring a startup and all the knowledge it has gained during the
+of users - described by the Roger's bell curve in~\cite{Rogers} - because of 
+the little return of investment they would gain, as discussed by Christensen 
+in~\cite{Christensen2003}. Instead, established companies have more interests 
+in acquiring a startup and all the knowledge it has gained during the 
 \textit{customer development} process while focusing effort on acquiring larger
@@
 been proven to be an extremely effective element of several successful software 
-projects \cite{Highsmith2000}. Hence, the vastly \textit{informal} development 
+projects~\cite{Highsmith2000}. Hence, the vastly \textit{informal} development 
 environment of startups acts as a catalyst of speed. 
@@
 it is premature to discuss issues related to SPI. Researchers should rather
-focus on the trade-off between development speed  and accumulated technical debt
-\cite{Brown:2010:MTD:1882362.1882373}, which  appears to be the most important
-determinant for the future of  software development in the company.
+focus on the trade-off between development speed and accumulated technical 
+debt~\cite{Brown:2010:MTD:1882362.1882373}, which appears to be the most 
+important determinant for the future of software development in the company.
 
@@
 %results... CG: fixed. 
-We structure the discussion according to Wohlin's taxonomy \cite{Wohlin2000}.
+We structure the discussion according to Wohlin's taxonomy~\cite{Wohlin2000}.
 
@@
 limited time practitioners were able to grant to it. However, researchers have
-to accept the  veracity of what respondents reported during the interviews
-\cite{Coleman2008a}.  %MUN1: candidate for removal 
+to accept the veracity of what respondents reported during the 
+interviews~\cite{Coleman2008a}.  %MUN1: candidate for removal 
 %Notwithstanding, the issues surrounding semi-structured interviews are vital.  
@@
 domain to which the study's findings can be generalized. In particular the
-previous framework developed by Coleman \cite{Coleman2008} has allowed  us to
+previous framework developed by Coleman~\cite{Coleman2008} has allowed us to
 find similarities, differences and broader reasoning related to factors that
@@
 describing the \textit{basic attacks} to the challenges of developing software
-systems and other literature covered by the SMS on startups.
-\cite{SMS}\footnote{Note that the generalizability of results is applied only to
-those startups that operate in software-intensive activities.}.  
+systems and other literature covered by the SMS on 
+startups~\cite{SMS}\footnote{Note that the generalizability of results is 
+applied only to those startups that operate in software-intensive activities.}.
 %MUN1: candidate for removal: detail on RM / seems out of place 
@@
 theory validation we highlighted and examined similarities, contrasts and
-explanations as discussed in \cite{Eisenhardt2007}. In this regard Eisenhardt
+explanations as discussed in~\cite{Eisenhardt2007}. In this regard Eisenhardt
 stated: ``Tying the emergent theory to existing literature enhances the internal
@@
 contexts to attest relationships among conceptualizations of an examined
-phenomenon (see \cite{Coleman2007, Basri, Coleman2008a}). Those relationships
+phenomenon (see~\cite{Coleman2007, Basri, Coleman2008a}). Those relationships
 between sampling and analysis should be verified in such a way that emerging
@@
 statement: \textit{``By the time three or four sets of data have been analysed,
-the majority of useful concepts will have been discovered.''}
-\cite{giardinoEtAl}. To obtain major certainty of the obtained results, what we
-applied was the final validation.
+the majority of useful concepts will have been 
+discovered.''}~\cite{giardinoEtAl}. To obtain major certainty of the obtained 
+results, what we applied was the final validation.
 
 Despite we couldn't apply any power analysis procedures, in grounded theory the
-sample size depended on when data saturation occurred. Ramer in \cite{Ramer},
+sample size depended on when data saturation occurred. Ramer in~\cite{Ramer},
 comparing quantitative to qualitative researches, states: \textit{``reaching
@@
 development, especially in the early-stage, is at the core of the company's 
-daily activities. Despite their severely high failure-rate, the SMS \cite{SMS}  
+daily activities. Despite their severely high failure-rate, the SMS~\cite{SMS} 
 found that the quick proliferation of startups is not supported by an adequate 
@@
 have been validated by means of systematic comparisons with the state-of-the-art
-results and empirical data. The systematic mapping study \cite{SMS},  with 37
+results and empirical data. The systematic mapping study~\cite{SMS}, with 37
 studies extracted, revealed a multi-faceted state-of-the-art inadequate to
@@
 increasing startup phenomenon. By making a comparison with Berry's definition of
-SE \cite{Berry1992}, we would like to see the rise of a new discipline -
+SE~\cite{Berry1992}, we would like to see the rise of a new discipline -
 \textit{startup engineering} - which can be defined as \textit{the use of

### 2013-03-04 article/main.tex
Revision 0.4. Sent to Tony for review

@@
 the software development strategies conducted by practitioners, in the period of
-time that goes from idea conception to the first open beta release. A
-\textit{grounded theory} approach guided the execution of a case study to
-explore the state-of-practice. As result, the Greenfield startup model (GSM)
-reveals the urgent priority of startups of releasing the product as quickly as
-possible to verify the product/market fit and to adjust the business and product
-trajectory according to early collected user feedback. Nevertheless, the need of
+time that goes from idea conception to the first open beta release. A grounded 
+theory approach guided the execution of a case study to explore the 
+state-of-practice. As result, the Greenfield startup model (GSM) reveals the 
+urgent priority of startups of releasing the product as quickly as possible to 
+verify the product/market fit and to adjust the business and product trajectory 
+according to early collected user feedback. Nevertheless, the need of 
 shortening time-to-market, by speeding-up the development through low-precision
@@
 
-From a software engineering perspective development in startups is challenging
+From a software engineering perspective, development in startups is challenging
 since they develop software in a context where processes can hardly follow a
@@
 combination of different factors makes the specific development context quite
-unique~\cite{Blank2005, Sutton2000}. Therefore, research is needed to support
+unique~\cite{Blank2005,Sutton2000}. Therefore, research is needed to support
 their engineering activities~\cite{Coleman2005}, guiding practitioners in taking
@@
 This research aims to understand how software development strategies are
-engineered by practitioners in startup companies in terms of level of:
-structure, planning and control of software projects, in the period of time
-that goes from idea conception to the first open beta release of the software
-product.
-
-With a cross-sectional study we focused the research context in a limited 
+engineered by practitioners in startup companies in terms of: structure, 
+planning and control of software projects, in the period of time that goes from 
+idea conception to the first open beta release of the software product.
+
+With a cross-sectional study we focused the research context on a limited 
 time-frame to emphasize the nature of uncertainty in the development activities 
 and time pressure from the market, as discussed in~\cite{Maccormack2001,
-Eisenhardt1998}. We performed semi-structured in-depth interviews with CEOs and
-CTOs of startups, covering a wide spectrum of thematics and iteratively
-adjusting the direction of the research according to the emerging evidences.
+Eisenhardt1998}. We performed semi-structured, in-depth interviews with CEOs and
+CTOs of startups, covering a wide spectrum of themes and iteratively
+adjusting the direction of the research according to the emerging evidence.
 
@@
 the GSM. Section~\ref{res:val} compares results of this research to literature
-and frameworks presented in this field. Section~\ref{valt} discusses the
-validity threats considered during this study. Section~\ref{sect:theory:impl}
-discusses the most relevant implications of the GSM. Finally, this paper
+and frameworks presented in this field. Section~\ref{sect:theory:impl}
+discusses the most relevant implications of the GSM. Section~\ref{valt} 
+discusses the validity threats considered during this study. The paper
 concludes in section~\ref{conc}.
@@
 Modern entrepreneurship has been boosted by the advent of the consumer internet
-markets in the middle of the 1990s and culminated with the notorious \textit
-{dot-com bubble burst} of 2000~\cite{Perkins:1999:IBI:555126,
+markets in the middle of the 1990s and culminated with the notorious 
+\textit{dot-com bubble burst} of 2000~\cite{Perkins:1999:IBI:555126,
 storey1982entrepreneurship}. Several years later, with the massification of the
 internet and mobile devices, we are now assisting to an impressive proliferation
-of software ventures - metaphorically referred as the \textit{startup bubble}.
-The easy access to vast potential markets and the low cost of services
-distribution are appealing condition for modern entrepreneurs~\cite{Marmer2011}.
-Inspired by success stories, a large number of software businesses are created
-every day. However, the great majority of these companies fail within two years
-from their creation~\cite{Crowne2002}.
+of software ventures - metaphorically referred to as the \textit{startup 
+bubble}. The easy access to vast potential markets and the low cost of services
+distribution are appealing conditions for modern
+entrepreneurs~\cite{Marmer2011}. Inspired by success stories, a large number of 
+software businesses are created every day. However, the great majority of these 
+companies fail within two years from their creation~\cite{Crowne2002}.
 
@@
 a specific client~\cite{genome2012}. Issues related to this domain are addressed
-in literature by the area known as \textit{market-driven software development}
+in literature by \textit{market-driven software development}
 (sometimes called \textit{packaged software development} or \textit{COTS
@@
 works in this specific area, as confirmed by other studies~\cite{Coleman2008,
-Coleman2008a, Coleman2007, Sutton2000}. Moreover, the studies, identified 
-in~\cite{SMS}, appear to be highly fragmented and spread across different areas
-rather than constituting a consistent \textit{body of knowledge}.
-
-A research published in 2000 by Sutton~\cite{Sutton2000} confirmed a general
-lack of studies in this area, claiming that \textit{``software startups
-represent a segment that has been mostly neglected in process studies''} and it
-has been further confirmed with the empirical studies of Coleman et 
+Coleman2008a, Coleman2007, Sutton2000}. Moreover, the studies appear to be 
+highly fragmented and spread across different areas rather than constituting a 
+consistent \textit{body of knowledge}.
+
+Sutton~\cite{Sutton2000} confirmed in 2000 a general lack of studies in this 
+area, claiming that \textit{``software startups represent a segment that has 
+been mostly neglected in process studies''} and it has been further confirmed 
+with the empirical studies by Coleman et 
 al.~\cite{Coleman2008,Coleman2008a,Coleman2007} eight years later.
@@
 
-A study of Kajko-Mattsson~\cite{Kajko-Mattsson2008}, which investigated a
+A study by Kajko-Mattsson~\cite{Kajko-Mattsson2008}, investigating a
 Swedish software startup, reported a heavy lack of requirements gathering
@@
 absence of documentation to track the status and progress of the process, and
-defective releases. Accordingly, Ambler et al. report how two startups
-approaching to an upcoming IPO started to require processes to focus on scalable
-solutions, in view of the growing company's size in terms of users and 
-employees~\cite{Ambler2002}. In this regard, Crowne~\cite{Crowne2002} specifies 
+defective releases. Accordingly, Ambler et al.~\cite{Ambler2002} reported how 
+two startups approaching to an upcoming IPO started to require processes to 
+focus on scalable solutions, in view of the growing company's size in terms of 
+users and employees. In this regard, Crowne~\cite{Crowne2002} specifies 
 different stages through which software startups evolve. Starting without any 
@@
 But when startups have no time for training and orienting activities, as
-discussed in~\cite{Sutton2000}, their main focus remains on team capabilities
-instead of prescriptive processes, hiring people who can \textit{``hit the
-ground running''}~\cite{Yoffie1999}. Empowering the team and focus on
-methodological attributes of the processes oriented in prototyping, 
-proof-of-concepts, mock-ups and demos, to test basic functionalities, have been  
-the primary priority in startups~\cite{Camel1994a}. Only when they grow, formal
+discussed by Sutton~\cite{Sutton2000}, their main focus remains on team 
+capabilities instead of prescriptive processes, hiring people who can 
+\textit{``hit the ground running''}~\cite{Yoffie1999}. Empowering the team and 
+focus on methodological attributes of the processes oriented towards 
+prototyping, proof-of-concepts, mock-ups and demos, to test basic 
+functionalities, have been  the primary priority in startups~\cite{Camel1994a}. 
+Only when they grow, formal
 methodologies arise, followed by performing quality assurance and long-term
@@
 
-As partially discussed above, contributions to flexibility and reactiveness of
-the development process has been conducted prominently by means of 
-\textit{Lean}~\cite{Gautam2008} and \textit{Agile}~\cite{Abrahamsson2002} 
-methodologies (also reported in~\cite{Taipale2010, Kuvinka2011}), where the 
-extreme uncertain conditions lead startups to learn fast from trials and errors 
-with a strong customer relationship in order to avoid wasting time in building 
-wrong functionalities and prevent rapidly exhaustion of 
-resources~\cite{Midler2008,Hilmola2003, Sutton2000}. Customer involvement in 
-software development has also been discussed by Yogendra~\cite{Yogendra2002} as 
-an important factor to encourage an early alignment of business concerns to the 
-technology strategies, because both are salient considerations to be 
-successful~\cite{Tingling2007}.
+Contributions to flexibility and reactiveness of the development process has 
+been conducted prominently by means of \textit{Lean}~\cite{Gautam2008} and 
+\textit{Agile}~\cite{Abrahamsson2002} methodologies (also reported 
+in~\cite{Taipale2010, Kuvinka2011}), where the extreme uncertain conditions lead 
+startups to learn fast from trial and error with a strong customer 
+relationship in order to avoid wasting time in building wrong functionalities 
+and prevent rapidly exhaustion of resources~\cite{Midler2008,Hilmola2003, 
+Sutton2000}. Customer involvement in software development has also been 
+discussed by Yogendra~\cite{Yogendra2002} as an important factor to encourage an 
+early alignment of business concerns to the technology strategies, because both 
+are salient considerations to be successful~\cite{Tingling2007}.
 
@@
 proposed by Zettel~\cite{Zettel2001}. The process consists of a set of
-activities and artifacts (in addition to some important roles) defined in order
-to identify responsibilities and tools to utilize. But, despite the promising
-benefits reported by Zettel, we were not able to identify any future evaluation
-in real-world settings.
-
-Another attempt of SPI in startups has been conducted by Deakins et al.
-introducing a \textit{Helical Model for managing e-commerce development
-environment}~\cite{Deakins2005}. Also in this case, the author prescribed broad
-guidelines for a rapid high-quality development process, which underwent limited
-testing only in academic settings.
+activities and artifacts to identify responsibilities and tools to utilize. But, 
+despite the promising benefits reported by Zettel, we were not able to identify 
+any evaluation in real-world settings. Another attempt of SPI in startups has 
+been conducted by Deakins et al.~\cite{Deakins2005}, introducing broad
+guidelines for a rapid high-quality development process, which underwent 
+however only limited testing in academic settings.
 
@@
 for startups, is described by Fayad~\cite{Fayad1997}. The author discusses the
-problem in actuating the same \textit{best-practices} criteria for established
-companies in 10-person software startups. Thoroughly remarked 
-in~\cite{Sutton2000}, Sutton states that problems of SPI in software startups 
-arise because of: the dynamism of the development process, which precludes
-repeatability; organizational maturity, that cannot be maintained from startups
-in view of lack of corporate direction; severe lack of resources, both human and
-technological for process definition, implementation, management, and training.
-In conclusion, the primary benefits of SPI do not address startups, which
-instead of promoting product quality, aim to minimize time-to-market.
-
-Additionally, the role of SPI has been neglected because seen as an obstacle to
-the development team's creativity and flexibility, and to the need of a quick
-delivering product process environment~\cite{Coleman2008a}. 
+problem in actuating the same best-practices criteria for established
+companies in 10-person software startups. Sutton~\cite{Sutton2000} remarks that 
+problems of SPI in software startups arise due to: the dynamism of the 
+development process, which precludes repeatability; organizational maturity, 
+that cannot be maintained from startups in view of lack of corporate direction; 
+severe lack of resources, both human and technological for process definition, 
+implementation, management, and training. In conclusion, the primary benefits 
+of SPI do not address startups, which instead of promoting product quality, aim 
+to minimize time-to-market.
+
+Additionally, the role of SPI has been neglected because it is seen as an 
+obstacle to the development team's creativity and flexibility, and to the need 
+of a quick delivering product process environment~\cite{Coleman2008a}. 
 %MUN: the second part of the sentence is just a statement without reference. If 
@@
 Product quality is left aside in favor of minimal and suitable functionalities 
-to shorten the time-to-market. As reported in two studies of Mater and 
-Mirel~\cite{Mater2000, Mirel2000}, quality aspects, mostly taken in 
-consideration in internet startups, are oriented to usability and scalability, 
-even though the market and application type heavily influences the 
+to shorten the time-to-market. As reported in two studies by Mater and 
+Mirel~\cite{Mater2000, Mirel2000}, quality aspects mostly taken in 
+consideration in internet startups, are oriented towards usability and 
+scalability, even though the market and application type heavily influences the 
 quality-demand~\cite{Coleman2008, Kim2005}.
@@
 %------------------------------------------------------------------------------%
-\subsection{Technical debt} Notwithstanding, a new area of SE research,
-trying to tackle the problem of the \textit{technical debt}, brings and
-encompasses different implications in studying development in software 
-startups~\cite{Tom2013}. The metaphoric neologism was originally introduced by
-Cunningham in 1992~\cite{TechnicalDebtCunn} and has recently attracted the
-attention of SE researchers\footnote{To attest the fact that \textit{technical
+\subsection{Technical debt} 
+Notwithstanding, a new area of SE research, trying to tackle the problem of 
+\textit{technical debt}, brings and encompasses different implications in 
+studying development in software startups~\cite{Tom2013}. The metaphoric 
+neologism was originally introduced by Cunningham in 
+1992~\cite{TechnicalDebtCunn} and has recently attracted the
+attention of SE researchers\footnote{To attest that \textit{technical
 debt} is gaining traction among researchers, we mention two important
@@
 organized by the Software Engineering Institute and ICSE.}. The concept of
-\textit{technical debt} is well illustrated 
-in~\cite{Brown:2010:MTD:1882362.1882373}: \textit{``The idea is that developers
-sometimes accept compromises in a system in one dimension (e.g., modularity) to
-meet an urgent demand in some other dimension (e.g., a deadline), and that such
-compromises incur a ``debt'' on which ``interest'' has to be paid and which the
-``principal'' should be repaid at some point for the long-term health of the
-project''}. The compromise between high-speed and high-quality engineering is
-faced daily by software startups, not only in terms of architecture design but
-in multifaceted aspects (weak project management, testing, process control).
+technical debt is well illustrated in~\cite{Brown:2010:MTD:1882362.1882373}: 
+\textit{``The idea is that developers sometimes accept compromises in a system 
+in one dimension (e.g., modularity) to meet an urgent demand in some other 
+dimension (e.g., a deadline), and that such compromises incur a ``debt'' on 
+which ``interest'' has to be paid and which the ``principal'' should be repaid 
+at some point for the long-term health of the project''}. The compromise 
+between high-speed and high-quality engineering is faced daily by software 
+startups, not only in terms of architecture design but in multifaceted aspects 
+(weak project management, testing, process control).
 
@@
 issues~\cite{Microsoft2009}. 
-\item Growth: an increase in the company size respect to the initial 
+\item Growth: an increase in the company size with respect to the initial 
 conditions in terms of either employees or users/customers, and
@@
 requests. 
-\item Software product: any software product and software service.
+\item Software product: any software product and/or software service.
 \end{compactitem}
@@
 The aim of this research is to \textit{understand how software development
-strategies are engineered by practitioners in startup companies in terms of
-level of: structure, planning and control of software projects, in the period of
+strategies are engineered by practitioners in startup companies in terms of: 
+structure, planning and control of software projects, in the period of
 time that goes from idea conception to the first open beta release of the
@@
 
-Initially the boundaries of the research domain were set by means of 
+Initially, the boundaries of the research domain were set by means of a
 non-systematic literature survey, which provided the initial keywords used  
@@
 its creators~\cite{ColinRobson2009} as \textit{``a set of well-developed
-categories ( e.g. themes, concepts ) that are systematically interrelated
+categories (e.g. themes, concepts) that are systematically interrelated
 through statements of relations to form a theoretical framework that explains
@@
 
-Following the GT principles we attempted to capture the most relevant aspects
-of software development from startup practitioners, letting a theory emerge
-from the interviews and adjusting the research hypotheses and questions as we
-proceeded through. During these interviews we collected data related to
+Following the GT principles, we captured the most relevant aspects of software 
+development from startup practitioners, letting a theory emerge from the 
+interviews and adjusting the research hypotheses and questions as we
+proceeded. During these interviews we collected data related to
 engineering activities undertaken by startups in the time frame between the idea
@@
 
-As suggested by Coleman, in view of the different versions of GT which emerged
-in the last years, researchers should indicate which \textit{``implementation''}
-of the theory is being used~\cite{Coleman2007}. After the methodology was
-initially introduced, its original authors had divergences regarding how the
-theory should be formed by analyzing the data (Strauss and Corbin on one side,
-Glaser on the other). Glaser advocated for a more \textit{puristic} approach,
-where the theoretical categories should \textit{``naturally emerge from the
-data''}, whilst Strauss and Corbin formulated an updated version of the
-methodology, which leaves to the researchers an higher degree of freedom. The
-latter approach empowers researchers' \textit{``theoretical 
-sensitivity''}~\cite{Corbin1990}, and encourages them to outline the research 
-problem beforehand. Since the knowledge obtained from parallel execution of the 
-SMS~\cite{SMS} and our direct experience with startups companies provided a good
+As suggested by Coleman, in view of the different versions of GT, researchers 
+should indicate which \textit{``implementation''} of the theory is being 
+used~\cite{Coleman2007}. After the GT methodology was initially introduced, its 
+original authors had divergences regarding how the theory should be formed by 
+analyzing the data (Strauss and Corbin on one side, Glaser on the other). 
+Glaser advocated for a more \textit{puristic} approach, where the theoretical 
+categories should \textit{``naturally emerge from the data''}, whilst Strauss 
+and Corbin formulated an updated version of the methodology, which leaves to 
+the researchers an higher degree of freedom. The latter approach empowers 
+researchers' \textit{``theoretical sensitivity''}~\cite{Corbin1990}, and 
+encourages them to outline the research problem beforehand. Since the knowledge 
+obtained from parallel execution of the SMS~\cite{SMS} and our direct experience 
+with startups companies provided a good
 initial level of knowledge, in this study we use the Corbin and Strauss
@@
 by the first two authors, handling conflicts by reviewing the rationale of
-decision with the third and fourth authors. When necessary we performed an in-
-depth review of the design of the research and data collected during the
+decision with the third and fourth authors. When necessary we performed an 
+in-depth review of the design of the research and data collected during the
 execution process.
@@
 This case study started based on a SMS of the literature regarding software
-development in startups~\cite{SMS}. By means of this SMS we refined our research
-questions. The research questions have been narrowed down to outline the domain
-of our study, even though still broad enough to allow emerging new theories and
+development in startups~\cite{SMS}, refining our research questions. The 
+research questions have been narrowed down to outline the domain
+of our study, even though still broad enough to allow new theories to emerge and
 to adjust the scope of the research~\cite{Corbin1990}.
@@
 %MUN: I've shortened the text and added absolute numbers. Is 2009-2012 correct?
-%CG: right by an average of 3 starting members, which were in majority
-developers. Moreover, the number of current employees shows how, to different
+%CG: right,
+by an average of 3 starting members, which were in majority developers. 
+Moreover, the number of current employees shows how, to different
 degrees, companies expanded the initial teams. All companies, except C5,
-brought their first product to market in less than 6 months from the idea
+brought their first product to market within less than 6 months from the idea
 conception. The products consist of web-applications, launched in different
@@
 %------------------------------------------------------------------------------%
-\subsection{Data collection} The data collection was designed to allow
-triangulation, which integrates multiple data sources converging on the same
-phenomenon. After transcribing the interview, we extracted conceptual
-relations, integrating them with the questionnaire's results. A well structured
-case-study database allowed us to easily retrieve and seek for information,
-assembling the evidence from the case study reports, as described also 
-in~\cite{Yin1994}. The database has been stored and constructed using the
-qualitative data analysis software package AtlasTI\footnote{Available online at
-\url{http://www.atlasti.com/}.} - one of the most suitable tools for 
-GT~\cite{Coleman2007}. We overlapped interviews with questionnaire results to
-adjust the data collection and take advantage of emergent themes and reveal
-possible inconsistencies. The data was analyzed simultaneously and with
-flexibility in mind in such a way that adjustments were made according to the
-emerging findings.
-%------------------------------------------------------------------------------%
-\subsection{Data analysis} Before starting the analysis, a data ordering
-procedure was necessary since interviews were spread across a multitude of
-topics. Therefore transcripts have been structured in thematic areas
-accordingly to different topic cards used during the interviews. We proceeded
-horizontally between same thematic areas of transcripts to be able to identify
-a better number of similar concepts, rather than going through an entire
-transcript at one time.
+\subsection{Data collection} 
+The data collection was designed to allow triangulation, which integrates 
+multiple data sources converging on the same phenomenon. After transcribing the 
+interview, we extracted conceptual relations, integrating them with the 
+questionnaire's results. A well structured case-study database allowed us to 
+easily retrieve and seek for information, assembling the evidence from the case 
+study reports, as described also by Yin~\cite{Yin1994}. The database has been 
+stored and constructed using the qualitative data analysis software package 
+AtlasTI\footnote{Available online at \url{http://www.atlasti.com/}.} - one of 
+the most suitable tools for GT~\cite{Coleman2007}. We overlapped interviews with 
+questionnaire results to adjust the data collection and take advantage of 
+emergent themes and reveal possible inconsistencies. The data was analyzed 
+simultaneously and with flexibility in mind in such a way that adjustments were 
+made according to the emerging findings.
+
+%------------------------------------------------------------------------------%
+\subsection{Data analysis} 
+Before starting the analysis, a data ordering procedure was necessary since 
+interviews were spread across a multitude of topics. Therefore transcripts have 
+been structured in thematic areas accordingly to different topic cards used 
+during the interviews. We proceeded horizontally between same thematic areas of 
+transcripts to be able to identify a better number of similar concepts, rather 
+than going through an entire transcript at one time.
 
@@
 memos to adjust the new emerging theory. To improve the speed of the coding
-process, we analyzed all the transcripts one part at-the-time: transcripts
+process, we analyzed all the transcripts one part at-a-time: transcripts
 followed the semi-structured interview format, and this approach enhanced the
@@
 Strauss and Corbin provided a list of questions to assist in determining how
-well the findings are grounded~\cite{Strauss1998}. They questions are as
-follows:
-
-\par\vspace {0.4cm} 
+well the findings are grounded~\cite{Strauss1998}:
+
 \begin{compactenum} 
@@
 \end{compactenum}
-\par\vspace {0.4cm}
 
@@
 coding process. They were systematically related through the use of a network
-diagram (see Figure~\ref{fig:gt:completemethodology}), by which we established
-the relations between the concepts. At each iteration of the grounded theory
-process, we considered and examined a concept within different conditions and
-dimensions, trying to incorporate data from a broader range of practitioners.
+diagram by which we established the relations between the concepts. At each 
+iteration of the grounded theory process, we considered and examined a concept 
+within different conditions and dimensions, trying to incorporate data from a 
+broader range of practitioners.
 
@@
 The process has been designed by the researches in different steps, explaining
-and making clear the purpose of each of them. Thus, the same processes enable
-other research to conduct the same study within similar contexts. Moreover
-\textit{comparison with the state-of-the-art} has been applied in order to
-validate the theory and enhance odds of its applicability within a wider time-
-frame. By this comparison we highlight the areas which have been neglected by
-existing studies, providing possible directions for future studies (see
-subsection~\ref{sect:theory:validation:others}). Moreover we study the
-confounding factors which could interfere with the application of the GSM model
-(see subsection~\ref{sect:an:comp:cat-vs-literature:conf}).
-
-Then, we performed an \textit{evaluation with empirical data} (see 
-subsection~\ref{sect:theory:validation:stats}), quantifying scores for each 
+the purpose of each of them. Thus, the same processes enable other researches 
+to replicate the study within similar contexts. Moreover, a comparison with the 
+state-of-the-art has been performed in order to validate the theory and to 
+strengthen its applicability within a wider time-frame. By this 
+comparison we highlight the areas which have been neglected by
+existing studies, providing possible directions for future studies
+(subsections~\ref{sect:theory:validation:others} 
+and~\ref{sect:theory:validation:sms}). Furthermore, we studied the
+confounding factors which could interfere with the application of the GSM
+(subsection~\ref{sect:an:comp:cat-vs-literature:conf}).
+
+At last, we performed an evaluation with empirical data 
+(subsection~\ref{sect:theory:validation:stats}), quantifying scores for each 
 company involved in our empirical study by defining metrics based on a set of 
 rubrics and evaluating startups accordingly, through an analysis of interview
-transcripts and codes. The complete statistical procedure and provided rubrics
-to quantify the measures are illustrated in Appendix.
+transcripts and codes. The complete statistical procedure and used rubrics
+to quantify the measures are provided in the Appendix [ADD URL].
 
@@
 
-The GSM captures the underlying phenomenon of software development in early-
-stage startups. The framework is formed by 128 sub-categories, clustered in 35
-groups, and finally in 7 categories at the highest level of abstraction. By the
-means of the GSM we provide explanations of the development strategies and
-engineering activities undertaken by startups.
-%------------------------------------------------------------------------------%
-\subsection{Framework overview} \label{res:gsm:frmov} The main concepts
-representing the underlying phenomenon have been grouped together to form high-
-level categories. Figure~\ref{fig:gsm} shows the network of causal
-relationships (represented by arrows) between categories (represented by
-blocks). Each node is linked to subsequent nodes which are called successors,
+The GSM captures the underlying phenomenon of software development in 
+early-stage startups. The framework is formed by 128 sub-categories, clustered 
+in 35 groups, and finally in 7 categories at the highest level of abstraction. 
+By the means of the GSM we provide explanations of the development strategies 
+and engineering activities undertaken by startups.
+
+%------------------------------------------------------------------------------%
+\subsection{Model overview} \label{res:gsm:frmov} 
+The main concepts representing the underlying phenomenon have been grouped 
+together to form high-level categories. Figure~\ref{fig:gsm} shows the network 
+of causal relationships (represented by arrows) between categories (represented 
+by blocks). Each node is linked to subsequent nodes which are called successors,
 and preceding nodes, called predecessors. The relations between them are denoted
@@
 
-In forthcoming the explanation of the GSM we emphasize labels of theoretical
+In the forthcoming explanation of the GSM we emphasize labels of theoretical
 categories using the \textit{italic} font style. For the main categories shown
@@
 implementing an essential set of functionalities and it is one of the main
-reasons why the \textit{product quality has low priority} respect to other more
-urgent needs\footnote{There are some exceptions where the quality aspects
-actually matter and such cases will be discussed in section~\ref{sect:an:comp
-:cat-vs-literature:conf}
+reasons why the \textit{product quality has low priority} with respect to other 
+more urgent needs\footnote{There are some exceptions where the quality aspects
+actually matter and such cases will be discussed in 
+section~\ref{sect:an:comp:cat-vs-literature:conf}
 %MUN: this refers to the validity threats section. Do you mean rather the 
@@
 
-In the following subsections we explain in detail all the categories presented
-in Figure~\ref{fig:gsm}, and conclude in subsection~\ref{res:gsm:th} with the
+In the following subsections we explain the categories presented in 
+Figure~\ref{fig:gsm}, and conclude in subsection~\ref{res:gsm:th} with the
 final theory. In the explanations we use identifiers of the companies presented
@@
 
-Early engineers in the founding team of a startups are typically \textit{multi-
+Early engineers in the founding team of startups are typically \textit{multi-
 role and full-stack engineers}: they are full-stack for being able to tackle
@@
 comments: \textit{``We had a hacker culture/environment, people hacking stuff
-without formally analyzing it, but breaking it down and finding a way around.
-Ticket on trello and go.''}
+without formally analyzing it, but breaking it down and finding a way around.''}
 
@@
 prototype and iteratively refine it over time, similarly to the concept of
-\textit{``evolutionary prototyping''}~\cite{EvProt}. The reason of using this
-approach is that startups want to validate the product against the market as
-soon as possible, finding the proper product/market fit. Indeed, they can focus
-on developing only parts of the system they want to validate instead of working
-on developing a whole new system. Then, as the prototype is released, users
-detect opportunities for new functionalities and improvements, providing their
-feedback to developers.
+\textit{``evolutionary prototyping''}~\cite{EvProt}. Startups want to validate 
+the product against the market as soon as possible, finding the proper 
+product/market fit. Indeed, they can focus on developing only parts of the 
+system they want to validate instead of working on developing a whole new 
+system. Then, as the prototype is released, users detect opportunities for new 
+functionalities and improvements, providing their feedback to developers.
 
@@
 suitable class of software development approaches are clearly highly
-evolutionary in nature. In fact, as \textit{uncertain conditions make long-term
+evolutionary in nature. As \textit{uncertain conditions make long-term
 planning not viable}, startups cannot base their work on assumptions without
@@
 and the project knowledge is not documented, even a little change in their
-composition (e.g. a developer falls ill) can have a very high impact on the
+composition (e.g. a developer falls ill) can have a high impact on the
 overall product development. Furthermore, startups operate in a continuous
@@
 
-The objective of this evolutionary approach is to avoid wasting time in ``over-
-engineering the system'' and building complex functionalities that have not
-been tested on real users. By releasing a very small number of good-enough
+The objective of this evolutionary approach is to avoid wasting time on 
+``over-engineering the system'' and building complex functionalities that have 
+not been tested on real users. By releasing a very small number of good-enough
 functionalities (see CAT3) the startup can verify the suitability of the
@@
 the software is available outside of the organization that developed 
-it\footnote{Detailed discussion of the impact of innovative products on the user 
+it\footnote{A discussion of the impact of innovative products on the user 
 satisfaction is presented in 
@@
 
-As mentioned before, in order to \textit{speed-up development}, startups adopt
-evolutionary approaches supported by a solid team focusing on implementing a
+In order to \textit{speed-up development}, startups adopt evolutionary 
+approaches supported by a solid team focusing on implementing a
 minimal set of suitable functionalities. Startups \textit{keep simple and
@@
 adoption of informal workflows is facilitated by the fact that teams are
-typically self-organized and \textit{developers have big responsibilities} .
+typically self-organized and \textit{developers have big responsibilities}.
 
@@
 category of tools is represented by advanced version control systems, which are
-not only used to manage the code-base, but for many other purposes such as:
-assigning tasks; trace responsibilities; configuration management; issue
-management; automatic deployment; discussions and code reviews.
+not only used to manage the code-base, but for also other purposes such as
+task assignment, responsibility tracing, configuration and issue 
+management, automatic deployment, and code reviews.
 
@@
 Other important factors that positively impact the speed of development are the
-team's desire to: \textit{create disruptive technologies}; to
-\textit{demonstrate personal abilities}; and to \textit{have the product used in
+team's desire to \textit{create disruptive technologies}, to
+\textit{demonstrate personal abilities}, and to \textit{have the product used in
 the market}. As reported by practitioners, these factors are essential to
@@
 practitioners, such way of working can be an effective strategy only in the
-short term since it can lead to poorly maintainable code and developer burnouts
+short term since it can lead to poorly maintainable code and developer burnout
 in the long run.
+
 %------------------------------------------------------------------------------%
@@
 Instead of traditional requirement engineering activities, startups make use of
-\textit{informal specification of functionalities} by means of ticked-based
+\textit{informal specification of functionalities} by means of ticket-based
 tools to manage low-precision lists of features to implement, written in form of
@@
 severe \textit{lack of automated testing}, which is mainly replaced by manual
-smoke tests executed by: trying the product internally; seeking for critical
-faults; and letting early adopters report non-critical bugs. Paraphrasing C3,
-\textit{``Trying the product internally allows us to get rid of 50\% of bugs of
-important functionalities. Meanwhile users report bugs of secondary
-functionalities, eventually allowing us to mitigate the lack of testing. Indeed,
-staying one week in production enables us to identify 90\% of bugs''.} The lack
-of complete automated tests is partially also motivated by the fact that
-\textit{users are fault-tolerant in innovative beta products} and by the
-limited number of functionalities, which allow the team to easily control
+smoke tests. Paraphrasing C3, \textit{``Trying the product internally allows us 
+to get rid of 50\% of bugs of important functionalities. Meanwhile users report 
+bugs of secondary functionalities, eventually allowing us to mitigate the lack 
+of testing. Indeed, staying one week in production enables us to identify 90\% 
+of bugs''.} The lack of complete automated tests is partially also motivated by 
+the fact that \textit{users are fault-tolerant in innovative beta products} and 
+by the limited number of functionalities, which allow the team to easily control
 critical bugs. However, in certain cases where components of the system might
@@
 project and technologies can be autonomously managed by the development team
-(with characteristics of: co-location, self-organization and very small team
+(characterized by co-location, self-organization and very small team
 size).  %MUN: do you mean small team size? %CG: fixed
@@
 development team} with \textit{limited need of formalities}.
+
 %------------------------------------------------------------------------------%
@@
 becomes of major relevance. C11 states: \textit{``Yet, it is still better to
-have a reasonable drop-down in performance when team grows than lose time in the
-beginning''}.
+have a reasonable drop-down in performance when the team grows than lose time 
+in the beginning''}.
+
 %------------------------------------------------------------------------------%
@@
 Coleman~\cite{Coleman2007,Coleman2008a, Coleman2008}, 
-Baskerville~\cite{Internet} and finally Brooks~\cite{BrooksJr1987}.
-
-Starting from the Coleman's work, we refer to him since only he has, to the
-best of our knowledge, conducted a similar research in startups' context, even
-though with a different focus. Coleman investigated factors in software
-development that hinder initiatives of software process improvement (SPI) in a
-later stage, representing also companies in expansion phase with more than 100
+Baskerville~\cite{Internet} and Brooks~\cite{BrooksJr1987}.
+
+We refer to Coleman's work since only he has, to the best of our knowledge, 
+conducted similar research in startups' context, even though with a different 
+focus. Coleman investigated factors in software development that hinder 
+initiatives of software process improvement (SPI) in a later stage, 
+representing also companies in the expansion phase with more than 100
 employees. %MUN: so he did not do research  in a startup context? You need to
@@
 is not a matter of poor knowledge and training, but rather it is the necessity
-of operating with solutions that let the company move faster. \textit{``One-
-size-fits-all''} solutions have always found difficulty in penetrating small
-software organizations~\cite{Staples2007}. When startups began establishing any
-SPI process, they experienced \textit{process erosions}~\cite{Coleman2008}
+of operating with solutions that let the company move faster. 
+\textit{``One-size-fits-all''} solutions have always found difficulty in 
+penetrating small software organizations~\cite{Staples2007}. When startups 
+began establishing any SPI process, they experienced \textit{process 
+erosions}~\cite{Coleman2008}
 %MUN:  %add ref (Coleman?). You should add, for each category you use from
@@
 , even though the accumulated tacit knowledge is hard to manage and to transfer 
-to new hires as also discussed in category \textit{CAT6}.
+to new hires as also discussed in CAT6.
 
@@
 
-Here we present the basic attacks (or mitigation strategies) presented in 1986
-by Brooks on the conceptual essence ascertained within the GSM model
-description:
+Here we present the basic mitigation strategies presented by Brooks in 1986 on 
+the conceptual essence ascertained within the GSM model description:
 
@@
 
-\item Buy versus build- the most radical possible solution for constructing
+\item Buy versus build: the most radical possible solution for constructing
 software is not to construct it at all, taking advantage of what others have
@@
 {speed-up development}. 
-\item Requirements refinement and rapid prototyping- avoid deciding precisely 
+\item Requirements refinement and rapid prototyping: avoid deciding precisely 
 what to build but rather iteratively extract and refine the product requirements 
@@
 maintain a fast release-cycles during development (CAT2). 
-\item Incremental development- starting from simple solutions allows 
+\item Incremental development: starting from simple solutions allows 
 organizations to early prototype and control complexity overtime. It represents 
@@
 (CAT3). 
-\item Great team- people are the center of a software project and it is 
+\item Great team: people are the center of a software project and it is 
 profoundly important to empower and liberate their creative mind. It is the main 
@@
 
-Brooks' forecast about software development strategies is revealed to be
+Brooks' forecast about software development strategies revealed to be
 extremely accurate, according to the state-of-practice in modern startups.
+
 %------------------------------------------------------------------------------%
@@
 
-In this subsection we report the evaluation of our theory in contrast with the
+In this subsection we report the evaluation of our theory with respect to the
 main contributions of the studies retrieved and discussed within the 
 SMS~\cite{SMS}. For each retrieved article we present the main results, mapping 
-them to categories in accordance to the GSM model. 
+them to categories in accordance to the GSM. 
 Table~\ref{tab:an:literature_comp} shows the main contributions of the 37 
@@
 startups failed in practicing engineering activities in view of excessive lack
-of flexibility. Developers fast moved to simpler tools such as using a white-
-board, where they were able to provide input and insights, question
+of flexibility. Developers fast moved to simpler tools such as using a 
+white-board, where they were able to provide input and insights, question
 assumptions, share past experiences, and even talk about potential
@@
 obtain customer's feedback as soon as possible for continuous product
-improvement.%MUN: %<-- rephrase, sentence not parsable CG:fixed 
-Even in this case, the spiral methodology%MUN: which model? CG:fixed 
+improvement. %MUN: %<-- rephrase, sentence not parsable CG:fixed 
+Even in this case, the spiral methodology %MUN: which model? CG:fixed 
 is compliant with the nature of a random, ad-hoc and \textit{evolutionary 
@@
 Carmel~\cite{Camel1994a} reports on the need of flexible and rapid development
-solutions to shorten time-to-market, but even more important, the need of high-
-skilled team developers, which has an impact on \textit{speed-up development}
-(CAT1). He suggests that entrepreneurs need to look for a well formed, skilled
+solutions to shorten time-to-market, but even more important, the need of 
+high-skilled team developers, which has an impact on \textit{speed-up 
+development} (CAT1). 
+He suggests that entrepreneurs need to look for a well formed, skilled
 core development team and not just a set of product ideas and features.
@@
 \subsubsection{Summary} Summarizing the comparison, we have found that most of
-the contributions of the studies are compliant with the GSM model. Startups
+the contributions of the studies are compliant with the GSM. Startups
 focusing on simple workflows and full-stack engineers face the uncertainty and
@@
 the consequences of a minimal project management and the use of low engineering
-activities as presented in the GSM model. However, the capability of the model
-to accommodate the results identified by previous researches give us more
-confidence in the generalizability of the GSM model.
+activities as presented in the GSM. However, the capability of the model
+to accommodate the results identified by previous research provides
+confidence in the generalizability of the GSM.
 
@@
 %MUN: Please reformulate the first sentence, I cannot parse it. CG: fixed 
-The proposed model identifies the three lifecycle stages of a
-startup: \textit{fluid}, \textit{transition} and \textit{specific}. At each
+The proposed model identifies three lifecycle stages of a startup: 
+\textit{fluid}, \textit{transition} and \textit{specific}. At each
 stage, the model presents different development approaches (product-oriented
@@
 inevitably clash. Consequently, team management requires higher control and
-coordination activities that inevitably hinder the flexibility of the
-development environment, that is essential in early-stage startups.
+coordination activities that hinder the flexibility of the development 
+environment, that is essential in early-stage startups.
 Accordingly, also Coleman~\cite{Coleman2007, Coleman2008a,Coleman2008} states
@@
 \item The faster the company releases the product, the larger the degree of 
-\textit{accumulated technical debt}\footnote{The debt here is considered 
+accumulated technical debt\footnote{The debt here is considered 
 regardless of context-specific debt mitigation factors and tactics.}. 
@@
 questionnaire results. The complete statistical procedure and provided rubrics
-to quantify the measures are illustrated and discussed in Appendix. 
+to quantify the measures are illustrated and discussed in Appendix [ADD URL]. 
 %MUN: make sure it is clear that the PROCEDURE material is available in the  
@@
 %difficult to understand. CG: fixed. 
-We conducted statistical tests using the analysis of
-variance to assess the existence of relations between the dimensions. First we
-defined two null hypotheses:
-
-\begin{itemize} 
-\item[$H0_1$:] \textit{Startups do not release the product
-faster when a capable team adopts a more evolutionary approach AND with less
-quality constraints} 
-\item[$H0_2$:] \textit{The execution speed does not
-increase the amount of accumulated technical debt}.
-\end{itemize}
+We conducted statistical tests using analysis of variance to assess the 
+existence of relations between the dimensions. First we defined two null 
+hypotheses:
+
+\begin{compactitem} 
+\item[$H0_1$:] \textit{Startups do not release the product faster when a 
+capable team adopts a more evolutionary approach AND with less quality 
+constraints.}
+\item[$H0_2$:] \textit{The execution speed does not increase the amount of 
+accumulated technical debt.}
+\end{compactitem}
 
@@
 analysis, fixing the level of confidence to 95\%. We reject both $H0_1$ and
-$H0_2$, and conclude that in our sample: 
+$H0_2$, and conclude that: 
 \begin{compactenum}
@@
 research question (\textit{What is the current state-of-practice related to
-software development strategy in early-stage startups?}) through its sub-
-questions. Although the startups we inquired in the case study were spread
+software development strategy in early-stage startups?}) through its 
+sub-questions. Although the startups we inquired in the case study were spread
 across different nations and market sectors, we have found several patterns
@@
 Developing software without releasing it for a long period of time, coming up
-with very complex solutions often is not a viable option. Business demands,
+with very complex solutions is often not a viable option. Business demands,
 investors interests, market pressures and the general uncertain conditions force
@@
 business idea}, nor make explicit use of the suggested good practices. Instead,
-startups in the early stage prefer also to apply fast cycles of "build and fix"
+startups in the early stage prefer to apply fast cycles of "build and fix"
 when necessary, hacking the first functionalities to present to potential
-customers and investors in order to obtain quick feedbacks. 
+customers and investors in order to obtain quick feedback. 
 %MUN1: Why don't follow startups the build-measure-learn cycle proposed by  
@@
 Blank~\cite{Blank2005}. Instead, startups absorb and implement the high-level
-principles that outlines the customer development methodology
-%MUN1:<--  %which methodology? CG: fixed 
-, reflected in the GSM by the theoretical category \textit{find the 
-product/market fit quickly}.
+principles that outlines the customer development methodology, reflected in the 
+GSM by the theoretical category \textit{find the product/market fit quickly}.
 
@@
 The first release of the product includes only a limited set of well suitable 
-functionalities focused on user experience. The beta version is usually 
-focused on a minimal and essential set of functionalities providing a reasonable 
-user experience (UX). The first functionalities are usually set by the 
-founders and their experience in a particular business domain. As soon as 
-possible they provide demos to demonstrate the idea to potential customers. 
+functionalities focused on user experience. These functionalities are 
+usually set by the founders and their experience in a particular business 
+domain. As soon as possible they provide demos to demonstrate the idea to 
+potential customers. 
 %MUN1: This rises the question, how do startups prioritize requirements.  
@@
 
-Future research can investigate the performance drop down caused by the
+Future research can investigate the performance drop-down caused by the
 necessity of returning the accumulated technical debt while expanding the
@@
 %------------------------------------------------------------------------------
-In this section we discuss the validity of the overall methodology. 
+In this section we discuss the validity of the overall research methodology. 
 %MUN1: threats  %are discussed w.r.t. the research methodology, not the 
@@
 which describes people's rejection attitude towards the assessment of their
-behavior. To address this risk we didn't attained to first answers from
-practitioners, but rather we asked for details of artifacts and engineering
+behavior. To address this risk we didn't attain to the first answers from
+practitioners, but rather asked for details of artifacts and engineering
 strategies. %MUN1: <--  %I don't understand this... CG: fixed 
 Moreover, we developed a rigorous data collection protocol to create case study 
-databases and employ integration of multiple data collection methods.%MUN1: <-- 
+databases and employ integration of multiple data collection methods. 
+%MUN1: <-- 
 %Since you write "Moreover, ..." the reader gets the impression that this is a
@@
 Reporting bias was mitigated by packaging all the needed material for
-conducting new researches in other contexts in order to criticize our findings
+conducting new research, in other contexts, in order to criticize our findings
 (interview package with instructions has been made available
@@
 the majority of useful concepts will have been 
-discovered.''}~\cite{giardinoEtAl}. To obtain major certainty of the obtained 
-results, what we applied was the final validation.
+discovered.''}~\cite{giardinoEtAl}.
 
@@
 development, especially in the early-stage, is at the core of the company's 
-daily activities. Despite their severely high failure-rate, the SMS~\cite{SMS} 
-found that the quick proliferation of startups is not supported by an adequate 
-and scientific body of knowledge. %MUN: <-- last sentence: isn't that a  
-%conclusion from the SMS? CG: fixed 
+daily activities. Despite their severely high failure-rate, the SMS 
+by Paternoster et al.~\cite{SMS} found that the quick proliferation of startups 
+is not supported by an adequate and scientific body of knowledge. 
+%MUN: <-- last sentence: isn't that a conclusion from the SMS? CG: fixed 
 To be able to intervene on the software development strategy with scientific and 
@@
 among practitioners with the aim of explaining how development strategies are
-engineered startups. The explanatory capability and correctness of the model
-have been validated by means of systematic comparisons with the state-of-the-art
-results and empirical data. The systematic mapping study~\cite{SMS}, with 37
+engineered in startups. The explanatory capability and correctness of the model
+has been validated by means of systematic comparisons with the state-of-the-art 
+and empirical data. The systematic mapping study~\cite{SMS}, with 37
 studies extracted, revealed a multi-faceted state-of-the-art inadequate to

### 2013-05-13 article/main.tex
ready for tony review

@@
 sub-questions. Although the startups we inquired in the case study were spread
-across different nations and market sectors, we have found several patterns
+across different nations and market sectors, we have found some patterns
 that emerged very clearly.
@@
 functionalities progressively rolled-out to early-adopters.
+
+Introducing a software development process in this early phases of high pressure can be challenging for startups.
+However, some development approaches, mostly investigated in mature companies, could help in shortening time-to-market.
+Among several methodologies that can be found in literature, the rapid application development (RAD) \cite{mcconnell2004professional} facilitates prototyping and promotes a continuous interactive process that allows early-adopters to understand, modify and approve a working product that meets their needs.
+A study of Gerbel et.
+al.
+states that the use of RAD, which promotes minimal planning and rapid prototyping, allows software to be released faster and makes easier to change requirements \cite{Gerber_practicalimplications}.
 %MUN1: Ok, I'm now "inventing" possible implications in the sections below...  
@@
 systematic (only little \textit{pair programming} sessions and rarely 
-\textit{test driven development}). Despite the wide number of publications, 
-which discuss the adoption of lightweight methodologies, we have found that 
-especially in this first stage not only methodologies but even individual 
-practices are basically neglected.
+\textit{test driven development}).
 
@@
 
+While the absence of a basic process might enable startups to focus more on the product, some studies claim that companies can take advantage from the use of some engineering activities even in the early stages \cite{989006}.
+For instance an experience report conducted by Taipale \cite{Taipale2010} shows how their startup took advantage from tailoring some simple XP practices to their needs.
 %MUN1: paragraph below is candidate for removal The greatest part of engineering
@@
 potential customers.
+
+To ship a product with the right features built in, startups need to prioritize and filter.
+From an engineering point of view, most startups do not explicitly apply traditional Requirement Engineering (RE) activities to collect and manage requirements.
+However, some studies have found that by integrating simple techniques, such as Persona and Scenario, companies can improve the effectiveness of the requirements elicitation even whether the final users is mostly unknown \cite{1531030}.
+Another study suggest that using a lightweight project-initiation framework such as the Agile Inception Deck can help in preventing premature failure of the software project due to wrong understanding of the project requirements \cite{1667597}.
 %MUN1: This rises the question, how do startups prioritize requirements.  
@@
 growing markets. If successful, the company will face a growth in terms of
-number of customers, employees and product capabilities. Then, the increased
-complexity arises the necessity of controlling the initially chaotic software
-development environment. Under delicate circumstances, where an increased
+number of customers, employees and product capabilities.
+Then, the increased complexity requires the team to start controlling the initial chaotic software development environment.
+Under delicate circumstances, where an increased
 market demand meets a poorly engineered product and process, the development
@@
 to absorb it.
+
+As stated by many interviewee, modern coding platforms (e.g.
+github\footnote{http://www.github.com}) and the use of well-known frameworks can be leveraged to mitigate the technical debt.
+Coding platforms allow developers to integrate in one place several engineering activities such as requirements lists, issue tracking, code repository, documentation, continuous integration, release and configuration management.
+Frameworks include support programs, compilers, code libraries and tool sets to enable the initial development of a project with limited overhead.
+
 %MUN1: what does literature say about technical debt? Are there strategies that 
